### 2020-08-23 paper/appendix.Rmd
init

---
title: "Appendix M"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 4
    fig_height: 4
    fig_caption: true
    highlight: tango
---
<!-- Run this to compile:
R -e "rmarkdown::render('appendix.Rmd', rmarkdown::pdf_document(), output_file='appendix.pdf', clean=TRUE)"
-->

# Basic / Classical RMT with Empirical Data

## Requirements and Overview
All you *really* need for basic RMT analysis is either:

1. a large (1000x1000) symmetric square matrix
2. a large sample of smaller symmetric square matrices
   - e.g. 10 or more 100x100 symmetric matrices

Provided your data / system has $N$ entities, and each entity has a
series of scalar values associated with them (e.g. you have $S_1, S_2, \dots, S_N$
where $S_i = \{t_{i_1}, ...\;, t_{i_{M_i}}\}$, $t_{i_j} \in \mathbb{R}$),
you can always generate a symmetric matrix with any function
that symmetrically compares the two series. I.e. if we have
$F$ such that $F(x, y) = F(y, x)$ for real series $x$ and $y$,
then the matrix $[F(S_i, S_j)]$ will be symmetric. Sensible choices of
$F$ might be Pearson's correlation coefficient, mutual information,
cross-correlation (convolution), or other distance / similarity metrics.

However, if you want to be able to actually find anything
interpretable (or interesting), you probably want a system that can be
roughly modeled as a number of somewhat distinct but interacting (or
similar, or connected) entities. E.g. particles bouncing off each
other, energy levels of a Uranium atom, distances between cars on a
freeway or in a parking lot, stock prices in a stock market, etc. It
is not clear how one would make sense of an RMT analysis of the
correlations of a static system (e.g. correlations between voxels or
ROIs of a structural MRI image).

The standard RMT procedure is then to compute some curves (metrics),
and compare (usually by eye) those curves to those predicted by RMT.
Typically, empirical results are compared to only the Gaussian
Orthogonal Ensemble (GOE) case, and/or the Poisson case, since each of
these represent two interpretational extremes.

The GOE is essentially a matrix where each entry is sampled from
the standard normal distribution. More precisely, it is the set of random
matrices:

$$A = [a_{ij}] \quad\mathrm{where}\quad A = A^T, \quad a_{ij} ~
\begin{cases}
  N(0, 1) &\mathrm{if}\; i \ne j\\
  \sqrt{2}N(0, 1) &\mathrm{if \; i = j}\\
\end{cases}
$$

If your matrix or matrices look like they are coming from a GOE
ensemble, you conclude the interactions in your system are highly
chaotic/random/noisy, i.e. that there is no real information (in the
mathematical and statistical sense) to be extracted there. This is a
pretty big deal, so if you see GOE behaviour, that's important.

The Poisson case is less clear to me currently, but it represents a
different kind of physical / mathematical extreme, and seems to represent
a more *regular* system. E.g. harmonic oscillators (where a particle's
position depends only on the field and its current position, and
not the positions of any other particles) produce some RMT metrics
similar to Poisson systems (Guhr & Mehta, 1998).

## Computational Approach
1. Compute a symmetric correlation, covariance, similarity, distance, mutual
   information, convolution, etc. matrix for your system
2. Compute the eigenvalues of that matrix
3. Sort the eigenvalues in ascending order
4. Compute the eigenvalue step / counting function
5. Unfold the eigenvalues
   - i.e. fit a smooth curve from the eigenvalues to the step function
   - use the predicted eigenvalues from this fit as the new unfolded values
6. Compute the spectral observables (curves) of the unfolded eigenvalues
7. Compare the computed curves to those predicted by theory

# Equations
## Fundamentals
### Dirac Delta Function

The function $\delta(x-t)$, *Dirac's delta* is informally defined, for constant $t$ as:

$$\delta(x-t) =
\begin{cases}
  \infty & \quad \text{if  } x = t \\
  0 & \quad \text{if  } x \ne t
\end{cases}$$

and where $\int_{-\infty}^{\infty}\delta(x - t) \;\mathrm{d}x = 1$.
Mostly it is a clever tool to place an infinitely thin but smooth
"spike" at a value to allow discrete-valued functions or sums to be
written continuously. It can be justified formally (as an entity used
only in integrals, and as the limit of a series of increasingly narrowing
bump functions with fixed area under the curve).

## Eigenvalue Step Function
Suppose we have our ordered eigenvalues $\{\lambda_1, ...,
\lambda_N\}$. Then $\eta(\lambda)$, the number of eigenvalues on
$[-\infty, \;\lambda]$, the eigenvalue **step** or **staircase
function** is:

$$\eta(\lambda) = \int_{-\infty}^{\lambda}
\sum_{n=1}^{N}\delta(\lambda - \lambda_n) \;\mathrm{d}\lambda \;\; =
\;\; \text{the numbers eigenvalues } \lambda_i \text{ less than
}\lambda$$

![Step Function for N=10000 GOE Matrix](figures/step_function_zoomed_out.png)
**Figure 1.** Example step function from a N=10 000 GOE matrix:

![Step Function for N=10000 GOE Matrix](figures/step_function_zoomed_in.png)
**Figure 2.** Step function fine structure.

In code:
```python
def stepFunction(eigs: np.array, x: np.array) -> np.array:
    if x == eigs:
        return np.arange(0, len(eigs) + 1)  # shortcut, return [0, 1, ... len(eigs)]
    steps = np.empty(len(eigs))
    for i, val in x:
        steps[i] = len(eigs[eigs <= val])  # count eigs <= each x
    return steps
```

## Unfolding
In order to compare the eigenvalues observed from a system to those
predicted from theory, we have to "unfold" the eigenvalues so that
they exist on the same scale / as those of the classical RMT ensembles.
This is a crucial and necessary step, and can be thought of as roughly
analagous to denoising (paraphrasing Guhr & Mehta, it is "removing the
system-specific mean levels to allow comparison of the unversal
components").

To unfold, we decompose the step function--in practice, by fitting a
polynomial or spline to $\{\lambda, \eta(\lambda)\}$--into a smooth
part $\tilde\eta(\lambda)$ and fluctuating part
$\eta_{\mathrm{fl}}(\lambda)$ such that

$$\eta(\lambda) = \tilde\eta(\lambda) + \eta_{\mathrm{fl}}(\lambda)$$

In certain special cases / systems where the physical model is fully
specified and understood, the functional forms of these smooth and
fluctuating parts can be calculated precisely. But for novel systems
being investigated with RMT, especially when there is no clear or
explicit underlying model, this unfolding procedure is essentially
guesswork.

In practice, it is most common to fit a smooth spline, or polynomial
of degree 5-11 $f$ to $\{\lambda, \eta(\lambda)\}$, and hope that $f$
is a good approximation to $\tilde\eta$. We then define

$$\{u_i = f(\lambda_i) \;\; | \;\; i = 1, 2, ..., N\}$$

to be the unfolded eigenvalues.

### Unfolding Algorithm
The essence of the unfolding process is a simple fitting procedure.

```python
def polynomial_unfold(eigs, degree):
    steps = np.empty([len(eigs)])  # allocate array for step function values
    for i, eig in enumerate(eigs):
        steps[i] = len(eigs[eigs < eig])  # count eigenvalues less than eig
    # steps = np.arange(0, len(eigs) + 1)  # alternate faster computation
    fit = polyfit(eigs, steps, degree)
    unfolded = polyval(eigs, fit)
    return unfolded
```
However, in practice the large eigenvalues of the system may act as anchor
points during fitting, heavily destroying the ability of the unfolding
process to extract information from the system. These large eigenvalues
must be dealt with in some manner (trimmed, fit separately). Additionally,
more complex unfolding procedures are also possible (using detrending, signal
processing approaches, and more local fitting).

### Unfolding Issues
* Long-range spectral observables are wildly sensitive to the unfolding procedure.
* The unfolding procedure is extremely arbitrary
  * choice of fitting procedure
    * e.g. polynomial, spline, EMD, other signal processing methods, detrending or not
  * fitting parameters
    * e.g. degree of polynomial / number of knots or type of spline, other fitting parameters
    * proportion of eigenvalues to fit (e.g. trimming eigenvalues of "large" absolute value)
    * fit accuracy (e.g. in theory we fit a continuous function $\eta(\lambda)$, but in practice
      we only evaluate $\eta(\lambda)$ on some grid of our choice
  * no way to test validity for unknown ensembles / systems
    * i.e. bad unfolding can make GOE systems look not GOE, or non-GOE systems look GOE (!!)

# Spectral Observables (Empirical Tests of RMT)
The spectral observables are the statistics of RMT. They are
statistics in the formal / general sense, i.e. they summarize random
variables. However, there are very, very few statistical tests in RMT.
One compares empirical results with RMT theory in the usual way
physicists do, e.g. via "binocular inspection".

## Nearest-Neighbour Spacing Distribution (Short Range)
Given observed eigenvalues $\{\lambda_1 \le ... \le \lambda_N\}$, we define

$$S_i = \lambda_{i+1} - \lambda_{i}, \qquad i = 1, 2, \dots N-1$$

the distances between the eigenvalues, and

$$D = \langle S_i \rangle$$

the mean distance between them. Then the (relative) level spacings are

$$s_i = S_i / D.$$

From theory, we know $p(s)$, is the probability of the spacings
between nearest neighboring eigenvalues. For certain cases (matrix
ensembles, or uncorrelated / Poisson case) the specific form of the
distribution is known. By computing the histogram (or kernel density
estimate) based on the empirically observed $s_i$ calculated from the
unfolded eigenvalues, we can compare our observations to the distribution
expected from theory.

For the Poisson case:

$$p(s) = \mathrm{exp}(-s)$$

For the GOE, $\beta = 1$,

$$p_1(s) = \frac{\pi}{2}s\;\mathrm{exp}\left(-\frac{\pi}{4}s^2\right)$$

For the GUE, $\beta = 2$,

$$p_2(s) = \frac{32}{\pi^2}s\;\mathrm{exp}\left(-\frac{4}{\pi}s^2\right)$$

For the GSE, $\beta = 4$,

$$p_4(s) = \frac{262144}{729\pi^3}s\;\mathrm{exp}\left(-\frac{64}{9\pi}s^2\right)$$

Note the domain of $p(s)$ is always $[0, \infty]$.

![NNSD](figures/nnsd_curves.png)
**Figure.** Predicted and actual distributions from a 1000x1000 GOE matrix.

## Level Number Variance $\Sigma^2(L)$

Let $N(L, c)$ count the number of eigenvalues on the interval $[c,\;c+L]$ "on the unfolded scale", i.e. $u_1 = \tilde\eta(\lambda_1)$. I.e. if we had the eigenvalues $E = \{1, 2, 3, 4.5, 4.7, 5, 6, 7, 8\}$, then:

$$
\begin{aligned}
  N(3, 3) \;&=\; |\{\lambda \in E : \lambda \in [3, 6]| &=& \quad|\{3, 4.5, 4.7, 5, 6\}| &= 5 \\
  N(3, 1.6) \;&=\; |\{\lambda \in E : \lambda \in [3, 4.6]| &=& \quad|\{3, 4.5\}| &= 2 \\
  N(100, 7) \;&=\; |\{\lambda \in E : \lambda \in [7, 107]| &=& \quad|\{7, 8\}| &= 2 \\
  N(100, 7) \;&=\; |\{\lambda \in E : \lambda \in [7, 107]| &=& \quad|\{7, 8\}| &= 2 \\
\end{aligned}
$$

Then we define:

$$\Sigma^2(L) = \left\langle N(L, c)^2 \right\rangle_c -  \left\langle N(L,c) \right\rangle_c^2,$$

the level number variance.

### $\Sigma^2(L)$ Algorithm
Algorithmically, for a particular value of $L$, we might compute $\Sigma^2$ like:
```python
def sigma_iter(unfolded: np.array, L: float, c_iters: int = 100):
    # allocate arrays to save counts of levels and square of those counts
    levels = np.empty((c_iters), dtype=np.float64)
    levels_sq = np.empty((c_iters), dtype=np.float64)

    # randomly choose c within the unfolded scale, and repeat c_iters times
    for i in range(c_iters):
        c_start = np.random.uniform(np.min(unfolded), np.max(unfolded))
        start, end = c_start - L/2, c_start + L/2
        # count number of eigenvalues in [start, end]
        n_within = len(unfolded[(unfolded >= start) & (unfolded <= end)])
        levels[i], levels_sq[i] = n_within, n_within**2

    ave = np.mean(levels)
    av_of_levels_sq = ave * ave
    av_of_sq_levels = np.mean(levels_sq)
    return av_of_sq_levels - av_of_levels_sq
```

For the classical ensembles and large $L$,
$\Sigma^2$, $\gamma = 0.5772\dots$ Euler's constant, $\Sigma^2$ has
approximations (valid to order $1/L$):

$$
\begin{aligned}
 \Sigma^2_\mathrm{Poisson}(L) =&\;L \\
 \Sigma^2_\mathrm{GOE}(L) =&\; \frac{2}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma + 1 - \frac{\pi^2}{8} \right) \\
 \Sigma^2_\mathrm{GUE}(L) =&\; \frac{1}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma + 1 \right) \\
 \Sigma^2_\mathrm{GSE}(L) =&\; \frac{1}{2\pi^2}\left( \mathrm{ln}(4\pi L) + \gamma + 1 + \frac{\pi^2}{8} \right) \\
\end{aligned}
$$

![Level Variance](figures/levelvar_curves.png)
**Figure.** Predicted and actual curves from a 1000x1000 GOE matrix.

There are also precise forms that are more accurate for small $L$.

## Spectral Rigidity $\Delta_3(L)$
If $\langle f \rangle_{t}$ denotes the average of $f$ over all $t$, and
$\hat{\eta}(\xi)$ represents the the number of levels on the unfolded scale, then

\begin{equation}
  \label{eq:delta3}
  \Delta_3(L) = {\left\langle \min_{A, B} \frac{1}{L} \int_{c}^{c+L}(\hat{\eta}(\lambda) -A\lambda - B)^2 \;\mathrm{d}\lambda\right\rangle}_{c}
\end{equation}

![Spectral Rigidity](figures/rigidity_calculation.png)
**Figure.** Average deviation of a section of the staircase function from a straight line.

is the spectral rigidity. Note that $A$ and $B$ may be different for each choice of $c$, $L$. For the motivation, note that if $(\hat{\eta}(\lambda) -A\lambda - B)^2 = 0$ then $\hat\eta$ is linear, i.e. the staircase function is extremely rigid. Also note that if $y = A\lambda + B$ is the least squares linear fit to $\hat\eta(\lambda)$ on $[c, \;c+L]$, then $A$ and $B$ minimize the integral.

### $\Delta_3(L)$ Algorithm
Algorithmically then, for a particular $L$ and $c$ value, we might calculate $\Delta_3(L)$ as:

```python
def spectralIter(eigs: np.array, unfolded: np.array, L, c_start, interval_gridsize) -> float:
    # c_start is in space of eigs
    grid = np.linspace(c_start - L/2, c_start + L/2, interval_gridsize)
    # compute step function values for each value in grid
    steps = stepFunction(eigs, grid)
    A, B = computeLinearFit(grid, steps)  # get slope A and intercept B from fit
    fit = A*grid + B                      # fit the line to the step function
    y = (steps - fit)**2                  # compute squared deviation of fit
    delta3 = np.trapz(y, grid)            # numerically integrate the values
    return delta3 / L
```
choosing $c$ values randomly and as many times as desired for stability / precision, for each $L$.

For the classical ensembles, the shape of $\Delta_3$ is known, and can be approximated to order $1/L$ by:

$$
\begin{aligned}
 \Delta_{3,\mathrm{Pois}son}(L) =&\;L/15 \\
 \Delta_{3,\mathrm{GOE}}(L) =&\; \frac{1}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma - \frac{5}{4} - \frac{\pi^2}{8} \right) \\
 \Delta_{3,\mathrm{GUE}}(L) =&\; \frac{1}{2\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma - \frac{5}{4} \right) \\
 \Delta_{3,\mathrm{GSE}}(L) =&\; \frac{1}{4\pi^2}\left( \mathrm{ln}(4\pi L) + \gamma - \frac{5}{4} + \frac{\pi^2}{8} \right) \\
\end{aligned}
$$

![Spectral Rigidity](figures/rigidity_curves.png)
**Figure.** Predicted Spectral Rigidity curves, with points from actual GOE data.

# Theory
There are two ways to look at RMT and justify its applications in
fMRI: mathematical and physical. RMT was was ultimately inspired by
and developed to solve complex, many-body problems in statistical
quantum mechanics (Guhr & Mehta, 1998). It was only recently with the
explosion of big data and more powerful computational capacities that
RMT has started to become relevant to other complex systems (Guhr &
Mehta, 1998).

## Interpretations
"The successful application of GRMT to
data shows that these data fluctuate stochastically in a manner
consistent with GRMT. Since GRMT can be derived from a maximum entropy
principle [18], this statement is tantamount to saying that the data
under study carry no information content" (Guhr & Mehta, 1998, p.202)

### RMT-Deviating Eigenvalues
"For complex quantum systems, RMT predictions represent an average
over all possible interactions [8–10]. Deviations from the universal
predictions of RMT identify system specific, non- random properties of
the system under consideration, provid- ing clues about the underlying
interactions [11–13]. Recent studies [14,15] applying RMT methods to
analyze the properties of C show that ~98% of the eigenvalues of C
agree with RMT predictions, suggesting a considerable de- gree of
randomness in the measured cross correlations. It is also found that
there are deviations from RMT predictions for ~2% of the largest
eigenvalues. These results prompt the following questions:" (Plerou et
al., 2002)

This seems to be a promising way to distinguish task from rest.

## Limitations
"It is clear that *RMT cannot ever reproduce a given data set in its
full detail*. It can only yield the distribution function of, and the
correlations between, the data points. An example is the nuclear level
spectrum at neutron threshold, the first case ever studied in a
statistically significant fashion. The actual sequence of levels
observed experimentally remains inaccessible to RMT predictions, while
the distribution of spacings and the correlations between levels can
be predicted. The same holds for the stochastic fluctuations of
nuclear cross sections, for universal conductance fluctu- ations of
mesoscopic probes, and for all other applications of RMT." (Guhr &
Mehta, 1998, p.202)

### Practical / Computational Issues

### Issues Related to fMRI

### 2020-08-23 code/extract.py

@@ -1,0 +1,58 @@
+import argparse
+import nibabel as nib
+import numpy as np
+import os
+import sys
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from pathlib import Path
+
+
+def mkdirp(path: Path) -> None:
+    try:
+        os.makedirs(path, exist_ok=True)
+    except Exception as e:
+        print(
+            f"Error making directory {path}. Another program may have modified the file "
+            "while this script was running.",
+            file=sys.stderr,
+        )
+        print("Original error:", file=sys.stderr)
+        raise e
+
+
+def res(string: str) -> str:
+    return str(Path(string).resolve())
+
+
+parser = argparse.ArgumentParser(description="Handle eigenvalue extraction")
+parser.add_argument("bold", metavar="<bold.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("mask", metavar="<mask.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("outfile", metavar="out.npy", type=res, nargs=1, action="store")
+
+args = parser.parse_args()
+
+img = nib.load(args.bold[0]).get_fdata()
+mask = np.array(nib.load(args.mask[0]).get_fdata(), dtype=bool)
+
+N, t = (np.prod(img.shape[:-1]), img.shape[-1])
+img = img.reshape([N, t])
+mask = mask.reshape(-1)
+
+# remove dead or constant voxels
+for i, signal in enumerate(img):
+    if np.sum(signal) == 0 or np.sum(signal * signal) == 0 or np.std(signal, ddof=1) == 0:
+        mask[i] = False
+
+brain = img[mask, :]
+
+eigs = Eigenvalues.from_time_series(brain, covariance=False, trim_zeros=False)
+vals = eigs.vals
+
+outfile = Path(args.outfile[0])
+shape_outfile = outfile.parent / outfile.name.replace("eigs", "shapes")
+parent = outfile.resolve().parent.resolve()
+mkdirp(parent)
+np.save(outfile, vals, allow_pickle=False)
+np.save(shape_outfile, np.array(brain.shape, dtype=int), allow_pickle=False)
+print(f"Saved eigenvalues to {args.outfile[0]}")

### 2020-08-23 code/learning.py

@@ -1,0 +1,460 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import pandas as pd
+import re
+import seaborn as sbn
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from empyricalRMT.ensemble import GOE, Poisson
+from glob import glob
+from matplotlib.pyplot import Axes
+from numpy import ndarray
+from pathlib import Path
+from tqdm import tqdm
+
+from typing import Any, Dict, List, Union
+from typing_extensions import Literal
+
+from rmt._filenames import precomputed_subgroup_paths_from_args
+from rmt._precompute import (
+    precompute_brody,
+    precompute_largest,
+    precompute_levelvar,
+    precompute_marchenko,
+    precompute_rigidity,
+)
+from rmt._types import Observable
+from rmt._utilities import _kde, _percentile_boot
+
+SubjField = Literal["group", "runs"]
+Subject = Dict[SubjField, Union[str, ndarray]]
+Subjects = Dict[str, Subject]
+
+repro = Path(__file__).resolve().parent
+DATA_ROOT = repro / "data"
+
+LEARNING_DATA = DATA_ROOT / "Task_Learning" / "rmt"
+os.makedirs(LEARNING_DATA, exist_ok=True)
+
+OUTDIRS = {"SLEPT": "", "NOSLEEP": ""}
+
+PLOT_OUTDIR = LEARNING_DATA / "plots"
+SUMMARY_OUTDIR = LEARNING_DATA / "summary"
+
+# fmt: off
+TRIM_ARGS     = r"(1,-1)"  # must be indices of trims, tuple, no spaces
+UNFOLD_ARGS   = {"smoother": "poly", "degree": 7, "detrend": False}
+LEVELVAR_ARGS = {"L": np.arange(0.5, 10, 0.1), "tol": 0.001, "max_L_iters": 50000}
+RIGIDITY_ARGS = {"L": np.arange(2, 20, 0.5)}
+BRODY_ARGS    = {"method": "mle"}
+# fmt: on
+
+
+def get_subjects_dict() -> Dict[str, Subject]:
+    eig_paths: List[Path] = np.sort(
+        [Path(p).resolve() for p in glob(str(LEARNING_DATA.parent) + "/**/*eigs*", recursive=True)]
+    )
+    subj_ids = np.unique(list(map(lambda p: p.stem.replace("eigs-", "")[:2], eig_paths)))
[…]

### 2020-08-23 code/print_extrema.py

@@ -1,0 +1,40 @@
+import numpy as np
+
+from glob import glob
+from pathlib import Path
+
+
+def print_extrema() -> None:
+    DATA_ROOT = Path(__file__).resolve().parent.parent / "data"
+
+    np.set_printoptions(precision=6, linewidth=150)
+
+    # shape_paths = np.sort([Path(file) for file in glob(f"{DATA_ROOT}/**/*shapes*.npy", recursive=True)])
+    shapes = [
+        np.load(file, allow_pickle=True)
+        for file in glob(f"{DATA_ROOT}/**/*shapes*.npy", recursive=True)
+    ]
+    paths = np.sort(
+        [Path(file).resolve() for file in glob(f"{DATA_ROOT}/**/*eigs*.npy", recursive=True)]
+    )
+    eigs_all = [
+        np.load(path, allow_pickle=True)
+        for path in glob(f"{DATA_ROOT}/**/*eigs*.npy", recursive=True)
+    ]
+    # parents = [path.parent.relative_to(DATA_ROOT) for path in paths]
+    parents = [path.parent for path in paths]
+    mins, maxs = (
+        np.array([eigs[:20] for eigs in eigs_all]),
+        np.array([eigs[-20:] for eigs in eigs_all]),
+    )
+    for eigs, shape, mn, mx, parent in zip(eigs_all, shapes, mins, maxs, parents):
+        print(str(parent))
+        print(f"shape: {shape}")
+        # print(f"mins:  {np.round(mn, 1)}")
+        print(f"maxs:  {np.round(mx, 1)}")
+        print(f"mean:  {np.round(np.mean(eigs), 1)}")
+        print("=" * 80)
+
+
+if __name__ == "__main__":
+    print_extrema()

### 2020-08-23 code/print_shapes.py

@@ -1,0 +1,26 @@
+import numpy as np
+
+from rmt._data_constants import DATASETS
+
+
+def print_shapes() -> None:
+    for dataset_name, dataset in DATASETS.items():
+        print(f"{'='*80}\n{dataset_name.upper()}")
+        for subgroup, eigpaths in dataset.items():
+            shape_paths = [str(p.resolve()).replace("eigs-", "shapes-") for p in eigpaths]
+            shapes = [np.array(np.load(shape), dtype=int) for shape in shape_paths]
+            ns, ts = (
+                np.array([shape[0] for shape in shapes]),
+                np.array([shape[1] for shape in shapes]),
+            )
+            n_mean, t_mean = int(np.round(np.mean(ns))), int(np.round(np.mean(ts)))
+            print(f"{'-'*80}\n{subgroup}")
+            print(f"N subjects: {len(eigpaths)}")
+            print("        {:>7}   {:>7}".format("n", "t"))
+            print("   min: {:7d}   {:7d}".format(int(ns.min()), int(ts.min())))
+            print("   max: {:7d}   {:7d}".format(int(ns.max()), int(ts.max())))
+            print("  mean: {:7d}   {:7d}".format(n_mean, t_mean))
+
+
+if __name__ == "__main__":
+    print_shapes()

### 2020-08-23 code/print_shapes_fullpre.py

@@ -1,0 +1,17 @@
+import numpy as np
+
+from rmt._data_constants import DATASETS_FULLPRE
+
+for dataset_name, dataset in DATASETS_FULLPRE.items():
+    print(f"{'='*80}\n{dataset_name.upper()}")
+    for subgroup, eigpaths in dataset.items():
+        shape_paths = [str(p.resolve()).replace("eigs-", "shapes-") for p in eigpaths]
+        shapes = [np.array(np.load(shape), dtype=int) for shape in shape_paths]
+        ns, ts = np.array([shape[0] for shape in shapes]), np.array([shape[1] for shape in shapes])
+        n_mean, t_mean = int(np.round(np.mean(ns))), int(np.round(np.mean(ts)))
+        print(f"{'-'*80}\n{subgroup}")
+        print(f"N subjects: {len(eigpaths)}")
+        print("        {:>7}   {:>7}".format("n", "t"))
+        print("   min: {:7d}   {:7d}".format(int(ns.min()), int(ts.min())))
+        print("   max: {:7d}   {:7d}".format(int(ns.max()), int(ts.max())))
+        print("  mean: {:7d}   {:7d}".format(n_mean, t_mean))

### 2020-08-23 code/rmt/_data_constants.py

@@ -1,0 +1,348 @@
+"""Holds constants related to filepaths."""
+import pandas as pd
+import sys
+
+from copy import deepcopy
+from glob import glob
+from pathlib import Path
+from pprint import pprint
+from typing import Any, Callable, Dict, List, Tuple
+
+from rmt._types import Dataset
+
+
+DATA_ROOT = Path(__file__).resolve().parent.parent.parent / "data"
+FULLPRE_DIRNAME = "rmt_fullpre"
+
+# fmt: off
+LEARNING_DATA                  = DATA_ROOT / "Rest_v_LearningRecall" / "rmt"
+OSTEO_DATA                     = DATA_ROOT / "Rest_w_Healthy_v_OsteoPain" / "rmt"
+PSYCHOLOGICAL_DATA             = DATA_ROOT / "Rest_w_VigilanceAttention" / "rmt"
+PARKINSONS_DATA                = DATA_ROOT / "Park_v_Control" / "rmt"
+REFLECTIVE_SUMMED              = DATA_ROOT / "Rest_v_Various_MultiEcho" / "rmt" / "summed"
+REFLECTIVE_INTERLEAVED         = DATA_ROOT / "Rest_v_Various_MultiEcho" / "rmt" / "interleaved"
+
+LEARNING_DATA_FULLPRE          = DATA_ROOT / "Rest_v_LearningRecall" / FULLPRE_DIRNAME
+OSTEO_DATA_FULLPRE             = DATA_ROOT / "Rest_w_Healthy_v_OsteoPain" / FULLPRE_DIRNAME
+PSYCHOLOGICAL_DATA_FULLPRE     = DATA_ROOT / "Rest_w_VigilanceAttention" / FULLPRE_DIRNAME
+PARKINSONS_DATA_FULLPRE        = DATA_ROOT / "Park_v_Control" / FULLPRE_DIRNAME
+REFLECTIVE_SUMMED_FULLPRE      = DATA_ROOT / "Rest_v_Various_MultiEcho" / FULLPRE_DIRNAME / "summed"
+REFLECTIVE_INTERLEAVED_FULLPRE = DATA_ROOT / "Rest_v_Various_MultiEcho" / FULLPRE_DIRNAME / "interleaved"
+
+VIGILANCE_SES1                 = PSYCHOLOGICAL_DATA / "selfreported_vigilance_ses-1.csv"
[…]

### 2020-08-23 code/rmt/_defaults.py

@@ -1,0 +1,9 @@
+import numpy as np
+
+# fmt: off
+TRIM_ARG_DEFAULTS     = "(1,:)"
+UNFOLD_ARG_DEFAULTS   = {"smoother": "poly", "degree": 7, "detrend": False}
+LEVELVAR_ARG_DEFAULTS = {"L": np.arange(0.5, 10, 0.1), "tol": 0.01, "max_L_iters": 50000}
+RIGIDITY_ARG_DEFAULTS = {"L": np.arange(2, 20, 0.5)}
+BRODY_ARG_DEFAULTS    = {"method": "mle"}
+# fmt: on

### 2020-08-23 code/rmt/_filenames.py

@@ -1,0 +1,171 @@
+"""Utilities for working with paths and filenames"""
+from pathlib import Path
+from typing import Any, Dict, List, Tuple
+
+import numpy as np
+
+from rmt._data_constants import DATA_ROOT, DATASETS, DATASETS_FULLPRE, PRECOMPUTE_OUTDIRS, PRECOMPUTE_OUTDIRS_FULLPRE
+from rmt._defaults import BRODY_ARG_DEFAULTS, LEVELVAR_ARG_DEFAULTS, RIGIDITY_ARG_DEFAULTS, UNFOLD_ARG_DEFAULTS
+from rmt._types import Observable
+
+
+def _prefix(trim_args: str, unfold_args: dict) -> str:
+    unfold_args = {**UNFOLD_ARG_DEFAULTS, **unfold_args}
+    prefix = (
+        f"{'' if (trim_args == '(1,:)') else f'tr{trim_args}_'}"
+        f"{unfold_args['smoother']}_deg{unfold_args['degree']}"
+        f"{'_detrended' if unfold_args['detrend'] else ''}"
+    )
+    return prefix
+
+
+def relpath(p: Path) -> str:
+    return str(p.relative_to(DATA_ROOT))
+
+
+def argstrings_from_args(args: Any) -> Tuple[str, Dict[Observable, str]]:
+    """Generate unique "argument strings" (useful for naming files) from `args`
+
+    Parameters
+    ----------
+    args: Args
+        Contains the unfolding, trimming, normalization, etc options defined in
+        run.py
+
+    Returns
+    -------
+    unfold_str: str
+        A string that uniquely and readably identifies the unfolding arguments
+
+    prefix_dict: Dict[str, str]
+        A dict with keys "prefix", "rigidity", "levelvar", "brody", "marchenko", "nnsd",
+        and "largest", and values that are strings that uniquely correspond to
+        the arguments
+    """
+    # defaults
+    unfold_args = {**UNFOLD_ARG_DEFAULTS, **args.unfold}  # splat (yes, that's the name for this syntax)
+    levelvar_args = {**LEVELVAR_ARG_DEFAULTS, **args.levelvar}
+    rigidity_args = {**RIGIDITY_ARG_DEFAULTS, **args.rigidity}
+    brody_args = {**BRODY_ARG_DEFAULTS, **args.brody}
+
+    prefix = _prefix(args.trim, unfold_args)
[…]

### 2020-08-23 code/rmt/_precompute.py

@@ -1,0 +1,391 @@
+import numpy as np
+import os
+import pandas as pd
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from pathlib import Path
+from tqdm import tqdm
+from typing import Any, List
+
+from rmt._data_constants import DATASETS, DATASETS_FULLPRE
+from rmt._filenames import precomputed_subgroup_paths_from_args, relpath
+from rmt._types import DataSummaryPaths
+
+
+def precompute_largest(eigpaths: List[Path], out: Path, force: bool = False, silent: bool = False) -> Path:
+    """Take the eigenvalues saved in `eigpaths`, compute the largest, and save that in a DataFrame
+    in `out`
+
+    Parameters
+    ----------
+    eigpaths: List[Path]
+        The values of either DATASETS or DATASETS_FULLPRE
+
+    out: Path
+        See usage below.
+
+    force: bool
+        If False (default), don't recompute the values if they already exist.
+
+    silent: bool
+        If False (default) display a tqdm progress bar while calculating.
+
+    Returns
+    -------
+    pickle: Path
+        Path to the pickle file saving the precomputed values.
+    """
+    if not force and out.exists():
+        return out
+    largest = pd.DataFrame(index=["largest"], columns=[path.stem for path in eigpaths])
+    desc = "{} - Largest"
+    pbar = tqdm(total=len(eigpaths), desc=desc.format("eigs-XX"), disable=silent)
+    for path in eigpaths:
+        eigname = path.stem
+        vals = np.load(path)
+        maximum = vals[-1]
+        largest.loc["largest", eigname] = maximum
+        pbar.set_description(desc=desc.format(path.stem))
+        pbar.update()
+    pbar.close()
+    # print(marchenko_df)
+    largest.to_pickle(out)
+    return out
+
+
+def precompute_marchenko(eigpaths: List[Path], out: Path, force: bool = False, silent: bool = False) -> Path:
+    """Take the eigenvalues saved in `eigpaths`, compute the Marchenko-Pastur
+    endpoints (both shifted and unshifted), and save that in a DataFrame at
+    `out`. DataFrame will also contain information rated to proportion of
+    eigenvalues within those bounds "noise_ratio" and "noise_ratio" shifted.
+
+    Parameters
+    ----------
+    eigpaths: List[Path]
+        The values of either DATASETS or DATASETS_FULLPRE
+
+    out: Path
+        See usage below.
+
+    force: bool
[…]

### 2020-08-23 code/rmt/_types.py

@@ -1,0 +1,8 @@
+from pathlib import Path
+from typing import Dict, List, Tuple
+from typing_extensions import Literal
+
+Dataset = Dict[str, List[Path]]
+Pair = Tuple[Path, Path]
+Observable = Literal["eigs", "rigidity", "levelvar", "brody", "nnsd", "largest", "marchenko"]
+DataSummaryPaths = Dict[str, Dict[Observable, Path]]

### 2020-08-23 code/rmt/_utilities.py

@@ -1,0 +1,114 @@
+import numpy as np
+import pandas as pd
+import seaborn as sbn
+
+
+from numpy import ndarray
+from statsmodels.nonparametric.kde import KDEUnivariate as KDE
+from typing import Any, Union
+
+
+def _percentile_boot(df: pd.DataFrame, B: int = 2000) -> pd.DataFrame:
+    """Compute the percentile bootstrapped mean and lower and upper 95% CI bounds
+
+    Parameters
+    ----------
+    df: DataFrame
+        DataFrame of observable values, with each column being a subject.
+        Assumes index has been correctly set (e.g. is "L" or "beta")
+
+    B: int
+        Number of bootstrap resamples.
+
+    Returns
+    -------
+    bootstraps: DataFrame
+        DataFrame with columns "mean", "low, "high", and with index same as df.
+    """
+    m, n = df.shape
+    out = pd.DataFrame(index=df.index, columns=["mean", "low", "high"], dtype=float)
+    for i in range(m):  # get bootstrap estimates for each row
+        boot_resamples = np.random.choice(df.iloc[i, :], size=(B, n), replace=True)
+        boot_means = np.sort(boot_resamples.mean(axis=1))
+        mean = np.mean(boot_means)
+        low, high = np.percentile(boot_means, [0.5, 99.5])
+        out.iloc[i, :] = [mean, low, high]
+    return out
+
+
+def _configure_sbn_style(prepend_black: bool = True) -> None:
+    """
+    NOTE
+    ----
+    This must be called every time *before* creating new figs and axes. In order
+    to fully ensure all plots get the correct style.
+    """
+    palette = sbn.color_palette("dark").copy()
+    if prepend_black:
+        palette.insert(0, (0.0, 0.0, 0.0))
+    sbn.set()
+    sbn.set_palette(palette)
+
+
+def _kde(values: ndarray, grid: ndarray, bw: Union[float, str] = "scott") -> ndarray:
+    """Calculate KDE for observed spacings.
+
+    Parameters
+    ----------
+    values: ndarray
+        the values used to compute (fit) the kernel density estimate
+
+    grid: ndarray
+        the grid of values over which to evaluate the computed KDE curve
+
+    axes: pyplot.Axes
+        the current axes object to be modified
+
+    bw: bandwidh
+        The `bw` argument for statsmodels KDEUnivariate .fit
+
[…]

### 2020-08-23 code/rmt/args.py

@@ -1,0 +1,199 @@
+import argparse
+import numpy as np
+
+from pathlib import Path
+from typing import Dict, Any
+
+
+def res(string: str) -> str:
+    return str(Path(string).resolve())
+
+
+def trim_parse(string: str) -> str:
+    s = string.replace(" ", "")
+    if s[0] != "(" or s[-1] != ")" or s.find(".") != -1:
+        raise ValueError("`trim_indices` must by tuple of ints of length 2.")
+    return str(s)
+
+
+def smoother_parse(string: str) -> str:
+    if string != "poly":
+        raise ValueError("`smoother` argument currently only implemented for 'poly'.")
+    return "poly"
+
+
+def degree_parse(string: str) -> int:
+    try:
+        deg = int(string)
+    except BaseException as e:
+        raise ValueError("`degree` must be in int.") from e
+    return deg
+
+
+def bool_parse(string: str) -> bool:
+    lower = string.lower()
+    if lower == "true" or lower == "t":
+        return True
+    if lower == "false" or lower == "f":
+        return False
+    return bool(string)
+
+
+def brody_parse(string: str) -> str:
+    if string in ["mle", "spacing"]:
+        return string
+    raise ValueError("Brody fit method must be either 'mle' or 'spacing'.")
+
+
+parser = argparse.ArgumentParser(description="Run with RMT args")
+parser.add_argument(
+    "-t", "--trim", metavar="<trim indices>", type=trim_parse, nargs=1, action="store", default="(1,-1)"
+)
+parser.add_argument(
+    "-s", "--smoother", metavar="<smoother>", type=smoother_parse, nargs=1, action="store", default="poly"
+)
+parser.add_argument(
+    "-d", "--degree", metavar="<smoother degree>", type=degree_parse, nargs=1, action="store", default=7
+)
+parser.add_argument("--detrend", metavar="<detrend>", type=bool_parse, nargs="?", action="store", default=False)
+parser.add_argument("-b", "--brody", metavar="<brody>", type=brody_parse, nargs=1, action="store", default="mle")
+parser.add_argument("--normalize", metavar="<normalize>", type=bool_parse, nargs="?", action="store", default=False)
+parser.add_argument("--fullpre", action="store_true")
+
[…]

### 2020-08-23 code/rmt/comparisons.py

@@ -1,0 +1,792 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import pandas as pd
+import seaborn as sbn
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from empyricalRMT.ensemble import GOE, Poisson
+from numpy import ndarray
+from pathlib import Path
+from pandas import DataFrame
+from scipy.stats import mannwhitneyu, ttest_ind, ttest_rel, wilcoxon
+from sklearn.discriminant_analysis import LinearDiscriminantAnalysis as LDA
+from sklearn.linear_model import LogisticRegression
+from sklearn.model_selection import cross_val_score, LeaveOneOut
+from sklearn.neighbors import KNeighborsClassifier as KNN
+from sklearn.preprocessing import MinMaxScaler
+from sklearn.svm import SVC
+from tqdm import tqdm
+
+from typing import Any, Dict, List, Optional, Tuple
+
+from rmt._data_constants import (
+    DATASETS,
+    DATASETS_FULLPRE,
+    DUDS,
+    PAIRED_DATA,
+    STAT_OUTDIRS,
+    STAT_OUTDIRS_FULLPRE,
+)
+from rmt._types import Observable
+from rmt._filenames import _prefix, precomputed_subgroup_paths_from_args, relpath, stats_fnames
+from rmt._utilities import _percentile_boot, _configure_sbn_style, _cohen_d
+
+
+def _trimmed_from_args(vals: ndarray, trim_args: str) -> ndarray:
+    """Parse `trim_args` and return the appropriate values from `vals`"""
+    if trim_args in ["(1,:)", "", "(0,:)"]:
+        vals = vals[1:]  # smallest eigenvalue is always spurious here
+    else:
+        low, high = eval(trim_args)
+        vals = vals[low:high]
+    return vals
+
+
+def group_differences_summary(g1: ndarray, g2: ndarray, rowname: str, dataset_name: str, comparison: str) -> DataFrame:
+    """Given ndarrays of scalar RMT features `g1` and `g2`, compute various
+    descriptive statistics and return them as a DataFrame.
+
+    Parameters
+    ----------
+    rowname: str
+        The str to use for the data row label (index).
+
+    dataset_name: str
+        The Dataset key name (just used here as a label).
+
+    comparison: str
+        The label for the comparison.
+
+    Returns
+    -------
+    df: DataFrame
+        The summary DataFrame.
+    """
[…]

### 2020-08-23 code/rmt/plot_datasets.py

@@ -1,0 +1,471 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import pandas as pd
+import seaborn as sbn
+
+
+from typing import Any, List
+
+from rmt._data_constants import DATA_ROOT, DATASETS, DATASETS_FULLPRE
+from rmt._filenames import argstrings_from_args, relpath
+from rmt._precompute import precompute_dataset
+from rmt._utilities import _percentile_boot
+from rmt.args import TRIM_IDX, NORMALIZE, UNFOLD_ARGS, FULLPRE, ARGS
+from rmt.comparisons import Pairings
+from rmt.summarize import supplement_stat_dfs, compute_all_preds_df
+
+# def plot_marchenko(datapaths: DataSummaryPaths, title: str = None, outdir: Path = None) -> None:
+def plot_marchenko(args: Any, show: bool = False) -> None:
+    """Create violin plots of the distributions of marchenko noise percents
+
+    Parameters
+    ----------
+    args: Args
+        Contains the unfolding, trimming, normalization, etc options defined in
+        run.py
+
+    show: bool
+        If False (default) just save the plot. Otherwise, call plt.show() and do
+        NOT save
+    """
+    # label is always g1_v_g2, we want "attention" to be orange, "nonattend"
+    # to be black
+    if not show:
+        outdir = DATA_ROOT / "plots"
+        os.makedirs(outdir, exist_ok=True)
+
+    datasets = DATASETS_FULLPRE if args.fullpre else DATASETS
+
+    total = 0
+    fig, axes = plt.subplots(nrows=3, nocols=8)
+    for dataset_name in datasets:
+        for groupname, observables in precompute_dataset(dataset_name, args, silent=True).items():
+            total += 1  # count how much plots we need
+
+    fig, axes = plt.subplots(nrows=3, nocols=8)
+    for dataset_name in datasets:
+        dfs, dfs_noise = [], []
+        for groupname, observables in precompute_dataset(dataset_name, args, silent=True).items():
+            df_full = pd.read_pickle(observables["marchenko"])
+            df = pd.DataFrame(df_full.loc["noise_ratio", :])
+            df_noise = pd.DataFrame(df_full.loc["noise_ratio_shifted", :])
+            n = len(df)
+            df["subgroup"] = [groupname for _ in range(n)]
+            df_noise["subgroup"] = [groupname for _ in range(n)]
+            dfs.append(df)
+            dfs_noise.append(df_noise)
+        df = pd.concat(dfs)
[…]

### 2020-08-23 code/rmt/summarize.py

@@ -1,0 +1,96 @@
+import pandas as pd
+
+from pathlib import Path
+from typing import Any, Tuple
+
+from rmt._data_constants import DATA_ROOT, DATASETS, DATASETS_FULLPRE
+from rmt._filenames import relpath, stats_fnames
+from rmt._precompute import precompute_dataset
+from rmt.comparisons import Pairings
+
+
+def compute_all_diffs_dfs(args: Any, silent: bool = False) -> Path:
+    _, diffs_out = stats_fnames(args, args.normalize, extension="csv")
+    all_diffs_out = DATA_ROOT / diffs_out
+    DUDS = ["control_pre_v_parkinsons", "park_pre_v_parkinsons", "control_v_park_pre", "control_v_control_pre"]
+
+    all_diffs = []
+    datasets = DATASETS_FULLPRE if args.fullpre else DATASETS
+    for dataset_name, dataset in datasets.items():
+        if dataset_name == "SINGLE_SUBJECT":
+            continue
+        precompute_dataset(dataset_name, args=args, force_all=False, silent=silent)
+        pairings = Pairings(args, dataset_name)
+        for pair in pairings.pairs:
+            if pair.label in DUDS:
+                continue
+            diffs = pair.paired_differences(args)
+            all_diffs.append(diffs)
+
+    diffs_df = pd.concat(all_diffs)
+    diffs_df.to_csv(all_diffs_out)
+    if not silent:
+        print(diffs_df)
+    print(f"Saved all differences to {all_diffs_out}")
+    return all_diffs_out
+
+
+def compute_all_preds_df(args: Any, silent: bool = False, force: bool = False) -> Path:
+    preds_out, _ = stats_fnames(args, args.normalize, extension="csv")
+    all_preds_out = DATA_ROOT / preds_out
+    if not force and all_preds_out.exists():
+        return all_preds_out
+    DUDS = ["control_pre_v_parkinsons", "park_pre_v_parkinsons", "control_v_park_pre", "control_v_control_pre"]
+    datasets = DATASETS_FULLPRE if args.fullpre else DATASETS
+
+    all_preds = []
+    for dataset_name, dataset in datasets.items():
+        if dataset_name == "SINGLE_SUBJECT":
+            continue
+        precompute_dataset(dataset_name, args=args, force_all=False, silent=silent)
+        pairings = Pairings(args, dataset_name)
+        for pair in pairings.pairs:
[…]

### 2020-08-23 code/run.py

@@ -1,0 +1,393 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import seaborn as sbn
+import sys
+
+from typing import Any
+from warnings import filterwarnings
+
+from rmt._data_constants import DATA_ROOT, DATASETS, DATASETS_FULLPRE
+from rmt._filenames import preview_precompute_outpaths
+from rmt._precompute import precompute_dataset
+from rmt.args import ARGS
+from rmt.summarize import compute_all_diffs_dfs, compute_all_preds_df, supplement_stat_dfs
+
+
+def plot_learning_fails(args: Any):
+    pass
+
+
+def all_pred_means_by_algo(
+    dataset_name: str = None,
+    comparison: str = None,
+    subtract_guess: bool = True,
+    fullpre: bool = True,
+    normalize: bool = False,
+    silent: bool = False,
+    force: bool = False,
+) -> None:
+    FULLPRE = ARGS.fullpre
+
+    UNFOLD = [5, 7, 9, 11, 13]
+
+    dfs = []
+    for trim in ["(1,-1)", "(1,-20)"]:
+        ARGS.trim = trim
+        for norm in [normalize]:
+            ARGS.normalize = norm
+            for degree in UNFOLD:
+                ARGS.unfold["degree"] = degree
+                supplemented = supplement_stat_dfs(diffs=None, preds=compute_all_preds_df(ARGS, silent=silent))[1]
+                df = pd.read_csv(supplemented)
+                df["Degree"] = ARGS.unfold["degree"]
+                df["Trim"] = ARGS.trim
+                dfs.append(df)
+
+    compare: pd.DataFrame = pd.concat(dfs)
+    features = ["Raw Eigs", "Largest20", "Largest", "Noise", "Noise (shift)", "Rigidity", "Levelvar"]
+    compare = compare[
+        [
+            "Dataset",
+            "Algorithm",
+            "Comparison",
+            "Degree",
+            "Trim",
+            "Raw Eigs",
+            "Largest20",
+            "Largest",
+            "Noise",
+            "Noise (shift)",
+            "Rigidity",
+            "Levelvar",
+            "Guess",
+        ]
+    ]
+    if dataset_name is not None:
+        compare = compare[compare["Dataset"] == dataset_name]
+    if comparison is not None:
+        compare = compare[compare["Comparison"] == comparison]
+
[…]

### 2020-08-23 code/save_shapes.py

@@ -1,0 +1,55 @@
+import argparse
+import nibabel as nib
+import numpy as np
+import os
+import sys
+
+from pathlib import Path
+
+
+def mkdirp(path: Path) -> None:
+    try:
+        os.makedirs(path, exist_ok=True)
+    except Exception as e:
+        print(
+            f"Error making directory {path}. Another program may have modified the file "
+            "while this script was running.",
+            file=sys.stderr,
+        )
+        print("Original error:", file=sys.stderr)
+        raise e
+
+
+def res(string: str) -> str:
+    return str(Path(string).resolve())
+
+
+parser = argparse.ArgumentParser(description="Handle eigenvalue extraction")
+parser.add_argument("bold", metavar="<bold.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("mask", metavar="<mask.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("outfile", metavar="out.npy", type=res, nargs=1, action="store")
+
+args = parser.parse_args()
+
+img = nib.load(args.bold[0]).get_fdata()
+mask = np.array(nib.load(args.mask[0]).get_fdata(), dtype=bool)
+
+N, t = (np.prod(img.shape[:-1]), img.shape[-1])
+img = img.reshape([N, t])
+mask = mask.reshape(-1)
+
+# remove dead or constant voxels
+for i, signal in enumerate(img):
+    if (
+        np.sum(signal) == 0
+        or np.sum(signal * signal) == 0
+        or np.std(signal, ddof=1) == 0
+    ):
+        mask[i] = False
+
+brain = img[mask, :]
+
+parent = Path(args.outfile[0]).resolve().parent.resolve()
+mkdirp(parent)
+np.save(args.outfile[0], np.array(brain.shape, dtype=int), allow_pickle=False)
+print(f"Saved shape to {args.outfile[0]}")

### 2020-08-23 code/view_results.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+TEMPFILE="/tmp/all_found_images.txt"
+fd --hidden --no-ignore --regex ".*$1.*.png" | sort > $TEMPFILE;
+num_pics=$(< $TEMPFILE wc -l);
+thumb_height=500;
+thumb_width=960;
+montage_width=3840;
+per_row=$(($montage_width / $thumb_width));
+nrows=$(($num_pics / $per_row));
+is_dangling=$(($num_pics % $per_row > 0));
+if [ "$is_dangling" -eq 1 ]; then
+    nrows=$(($nrows + 1))
+fi
+montage_height=$(($nrows * thumb_height + 500));
+echo "Displaying plots for regex \".*$1.*.png\":"
+feh -m \
+    --borderless \
+    --stretch \
+    --thumb-height $thumb_height \
+    --thumb-width $thumb_width \
+    -H $montage_height \
+    -W $montage_width \
+    --image-bg white \
+    -f $TEMPFILE \
+    --title "regex == $1" \
+    --thumb-title "regex == $1" \
+    ;
+rm -rf $TEMPFILE;
+

### 2020-08-23 data/Park_v_Control/afni_extract.py

@@ -1,0 +1,57 @@
+import argparse
+import nibabel as nib
+import numpy as np
+import os
+import platform
+import sys
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from glob import glob
+from pathlib import Path
+
+
+def mkdirp(path: Path) -> None:
+    try:
+        os.makedirs(path, exist_ok=True)
+    except Exception as e:
+        print(
+            f"Error making directory {path}. Another program may have modified the file "
+            "while this script was running.",
+            file=sys.stderr,
+        )
+        print("Original error:", file=sys.stderr)
+        raise e
+
+
+def res(string: str) -> str:
+    return str(Path(string).resolve())
+
+
+parser = argparse.ArgumentParser(description="Handle eigenvalue extraction")
+parser.add_argument("bold", metavar="<bold.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("outfile", metavar="out.npy", type=res, nargs=1, action="store")
+
+args = parser.parse_args()
+
+img = nib.load(args.bold[0]).get_fdata()
+N, t = (np.prod(img.shape[:-1]), img.shape[-1])
+img = img.reshape([N, t])
+mask = np.ones([N], dtype=bool)
+
+# remove dead or constant voxels
+for i, signal in enumerate(img):
+    if np.sum(signal) == 0 or np.sum(signal * signal) == 0 or np.std(signal, ddof=1) == 0:
+        mask[i] = False
+
+brain = img[mask, :]
+
+eigs = Eigenvalues.from_time_series(brain, covariance=False, trim_zeros=False)
+vals = eigs.vals
+
+outfile = Path(args.outfile[0])
+shape_outfile = outfile.parent / outfile.name.replace("eigs", "shapes")
+parent = outfile.resolve().parent.resolve()
+mkdirp(parent)
+np.save(outfile, vals, allow_pickle=False)
+np.save(shape_outfile, np.array(brain.shape, dtype=int), allow_pickle=False)
+print(f"Saved eigenvalues to {args.outfile[0]}")

### 2020-08-23 data/Park_v_Control/afni_extract.sh

@@ -1,0 +1,277 @@
+mkdir -p rmt/afni_preproc/{controls,parkinsons} &&
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-1_bold.nii.gz rmt/afni_preproc/controls/eigs-01_1.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-2_bold.nii.gz rmt/afni_preproc/controls/eigs-01_2.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-3_bold.nii.gz rmt/afni_preproc/controls/eigs-01_3.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-4_bold.nii.gz rmt/afni_preproc/controls/eigs-01_4.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-5_bold.nii.gz rmt/afni_preproc/controls/eigs-01_5.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4101_ses-1_task-ANT_run-6_bold.nii.gz rmt/afni_preproc/controls/eigs-01_6.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4103_ses-1_task-ANT_run-1_bold.nii.gz rmt/afni_preproc/controls/eigs-03_1.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4103_ses-1_task-ANT_run-2_bold.nii.gz rmt/afni_preproc/controls/eigs-03_2.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4103_ses-1_task-ANT_run-3_bold.nii.gz rmt/afni_preproc/controls/eigs-03_3.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4103_ses-1_task-ANT_run-4_bold.nii.gz rmt/afni_preproc/controls/eigs-03_4.npy;
+python3 afni_extract.py derivatives/afni_preproc/controls/sub-RC4103_ses-1_task-ANT_run-5_bold.nii.gz rmt/afni_preproc/controls/eigs-03_5.npy;
[…]

### 2020-08-23 data/Park_v_Control/extract.sh

@@ -1,0 +1,204 @@
+mkdir -p rmt/controls rmt/parkinsons &&
+echo "Starting parkinsons";
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-6_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-6_stripped_mask.nii.gz rmt/parkinsons/eigs-27_6.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-5_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-5_stripped_mask.nii.gz rmt/parkinsons/eigs-27_5.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-4_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-4_stripped_mask.nii.gz rmt/parkinsons/eigs-27_4.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-3_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-3_stripped_mask.nii.gz rmt/parkinsons/eigs-27_3.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-2_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-2_stripped_mask.nii.gz rmt/parkinsons/eigs-27_2.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-1_stripped.nii.gz derivatives/parkinsons/sub-RC4227_ses-1_task-ANT_run-1_stripped_mask.nii.gz rmt/parkinsons/eigs-27_1.npy;
+python3 ../../extract.py derivatives/parkinsons/sub-RC4226_ses-1_task-ANT_run-6_stripped.nii.gz derivatives/parkinsons/sub-RC4226_ses-1_task-ANT_run-6_stripped_mask.nii.gz rmt/parkinsons/eigs-26_6.npy;
[…]

### 2020-08-23 data/Park_v_Control/extract_fullpre.sh

@@ -1,0 +1,201 @@
+mkdir -p rmt_fullpre/raw/controls rmt_fullpre/raw/parkinsons &&
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_6.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-5_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-5_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_5.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-4_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-4_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_4.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-3_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-3_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_3.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-2_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-2_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_2.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-1_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-1_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_1.npy ;
+python3 ../../extract.py data/control/sub-RC4130_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/control/sub-RC4130_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-30_6.npy ;
[…]

### 2020-08-23 data/Park_v_Control/extract_fullpre_control.sh

@@ -1,0 +1,101 @@
+mkdir -p rmt_fullpre/raw/controls rmt_fullpre/raw/parkinsons &&
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_6.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-5_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-5_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_5.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-4_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-4_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_4.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-3_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-3_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_3.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-2_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-2_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_2.npy ;
+python3 ../../extract.py data/control/sub-RC4131_ses-1_task-ANT_run-1_fullpre_stripped.nii.gz data/control/sub-RC4131_ses-1_task-ANT_run-1_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-31_1.npy ;
+python3 ../../extract.py data/control/sub-RC4130_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/control/sub-RC4130_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/controls/eigs-30_6.npy ;
[…]

### 2020-08-23 data/Park_v_Control/extract_fullpre_park.sh

@@ -1,0 +1,100 @@
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_6.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-5_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-5_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_5.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-4_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-4_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_4.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-3_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-3_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_3.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-2_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-2_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_2.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4227_ses-1_task-ANT_run-1_fullpre_stripped.nii.gz data/parkinsons/sub-RC4227_ses-1_task-ANT_run-1_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-27_1.npy ;
+python3 ../../extract.py data/parkinsons/sub-RC4226_ses-1_task-ANT_run-6_fullpre_stripped.nii.gz data/parkinsons/sub-RC4226_ses-1_task-ANT_run-6_fullpre_stripped_mask.nii.gz rmt_fullpre/raw/parkinsons/eigs-26_6.npy ;
[…]

### 2020-08-23 data/Park_v_Control/full_preprocess.py

@@ -1,0 +1,175 @@
+"""
+The main challenge here is that we have to delete intermediates in order to
+prevent completely filling up the HD with junk. However, if some intermediate
+step fails, we want to keep the last relevant pre-processing step.
+"""
+import argparse
+import nipype
+import os
+import pickle
+
+from glob import glob
+from nipype.interfaces.fsl import MCFLIRT, SliceTimer, BET
+from pathlib import Path
+from time import ctime
+from typing import List, Tuple
+from warnings import filterwarnings
+
+
+DATA = Path(__file__).resolve().parent / "data"
+LOGFILE = (
+    Path(__file__).resolve().parent
+    / f"{ctime().replace(' ', '_').replace(':', '.')}_pid{os.getpid()}__preprocess.log"
+)
+
+MCFLIRT_SUFFIX = "mcflirted.nii.gz"
+SLICETIME_SUFFIX = "stationary.nii.gz"
+BET_SUFFIX = "fullpre_stripped.nii.gz"
+
+
+class Log:
+    def __init__(self, logfile: Path = LOGFILE) -> None:
+        self.log: List[str] = []
+        if logfile.exists():
+            os.remove(logfile)
+        self.out = logfile
+
+    def append(self, s: str) -> None:
+        self.log.append(s)
+        with open(self.out, "a+") as log:
+            log.write(f"{s}\n")
+
+    def print(self) -> None:
+        print("\n".join(self.log))
+
+
+def mcflirt(infile: Path) -> Path:
+    # motion-correction FIRST supported by e.g.
+    # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6736626/
+    # given we are dealing with FSL
+    outfile = Path(str(infile).replace("bold.nii.gz", MCFLIRT_SUFFIX))
+    if outfile.exists():
+        return outfile
+    cmd = MCFLIRT(
+        in_file=str(infile),
+        out_file=str(outfile),
+        output_type="NIFTI_GZ",
+        save_mats=False,
+        save_rms=False,
+        stages=3,
+        stats_imgs=False,
+        terminal_output="stream",
+        mean_vol=False,  # speed up things
+    )
+
+    results = cmd.run()
+    return Path(results.outputs.out_file)
+
+
+def slicetime_correct(infile: Path, timings: Path, TR: float, slice_direction: str = "z") -> Path:
[…]

### 2020-08-23 data/Park_v_Control/fully_preprocess.sh

@@ -1,0 +1,201 @@
+mkdir -p rmt_fullpre &&
+python3 full_preprocess.py -i data/control/sub-RC4113_ses-1_task-ANT_run-3_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4113_ses-1_task-ANT_run-4_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4113_ses-1_task-ANT_run-5_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4113_ses-1_task-ANT_run-6_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-1_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-2_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-3_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-4_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-5_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4114_ses-1_task-ANT_run-6_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4115_ses-1_task-ANT_run-1_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i data/control/sub-RC4115_ses-1_task-ANT_run-2_bold.nii.gz --TR 2.4 --timings slicetimes.txt --direction z ;
[…]

### 2020-08-23 data/Park_v_Control/preprocess.sh

@@ -1,0 +1,203 @@
+mkdir -p derivatives/controls derivatives/parkinsons &&
+bet data/control/sub-RC4113_ses-1_task-ANT_run-3_bold.nii.gz data/control/sub-RC4113_ses-1_task-ANT_run-3_stripped.nii.gz -F &
+bet data/control/sub-RC4113_ses-1_task-ANT_run-4_bold.nii.gz data/control/sub-RC4113_ses-1_task-ANT_run-4_stripped.nii.gz -F &
+bet data/control/sub-RC4113_ses-1_task-ANT_run-5_bold.nii.gz data/control/sub-RC4113_ses-1_task-ANT_run-5_stripped.nii.gz -F &
+bet data/control/sub-RC4113_ses-1_task-ANT_run-6_bold.nii.gz data/control/sub-RC4113_ses-1_task-ANT_run-6_stripped.nii.gz -F &
+bet data/control/sub-RC4114_ses-1_task-ANT_run-1_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-1_stripped.nii.gz -F &
+bet data/control/sub-RC4114_ses-1_task-ANT_run-2_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-2_stripped.nii.gz -F &
+bet data/control/sub-RC4114_ses-1_task-ANT_run-3_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-3_stripped.nii.gz -F &
+bet data/control/sub-RC4114_ses-1_task-ANT_run-4_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-4_stripped.nii.gz -F ;
+bet data/control/sub-RC4114_ses-1_task-ANT_run-5_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-5_stripped.nii.gz -F &
+bet data/control/sub-RC4114_ses-1_task-ANT_run-6_bold.nii.gz data/control/sub-RC4114_ses-1_task-ANT_run-6_stripped.nii.gz -F &
+bet data/control/sub-RC4115_ses-1_task-ANT_run-1_bold.nii.gz data/control/sub-RC4115_ses-1_task-ANT_run-1_stripped.nii.gz -F &
[…]

### 2020-08-23 data/Rest_v_LearningRecall/extract.sh

@@ -1,0 +1,432 @@
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-08_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-08_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_08.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-07_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-07_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_07.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-06_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-06_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_06.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-05_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-05_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_05.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-04_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-04_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_04.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-03_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-03_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_03.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-02_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-02_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_02.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-satellite_run-01_stripped.nii.gz all_fmri/sub-24_ses-2_task-satellite_run-01_stripped_mask.nii.gz rmt/task/ses-2_eigs-24_01.npy ;
[…]

### 2020-08-23 data/Rest_v_LearningRecall/extract_fullpre_rest.sh

@@ -1,0 +1,48 @@
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-02_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-02_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-03_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-03_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-04_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-04_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-05_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-05_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-06_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-06_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
[…]

### 2020-08-23 data/Rest_v_LearningRecall/extract_fullpre_task.sh

@@ -1,0 +1,384 @@
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-02_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-03_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-04_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-05_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-06_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-07_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-satellite_run-08_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-satellite_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-satellite_run-02_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-satellite_run-03_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-satellite_run-04_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction z ;
[…]

### 2020-08-23 data/Rest_v_LearningRecall/extract_fully_preprocessed.sh

@@ -1,0 +1,433 @@
+mkdir -p rmt_fullpre/rest rmt_fullpre/task &&
+python3 ../../extract.py all_fmri/sub-24_ses-2_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-24_ses-2_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-2_eigs-24_1.npy ;
+python3 ../../extract.py all_fmri/sub-24_ses-1_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-24_ses-1_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-1_eigs-24_1.npy ;
+python3 ../../extract.py all_fmri/sub-23_ses-2_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-23_ses-2_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-2_eigs-23_1.npy ;
+python3 ../../extract.py all_fmri/sub-23_ses-1_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-23_ses-1_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-1_eigs-23_1.npy ;
+python3 ../../extract.py all_fmri/sub-22_ses-2_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-22_ses-2_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-2_eigs-22_1.npy ;
+python3 ../../extract.py all_fmri/sub-22_ses-1_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-22_ses-1_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-1_eigs-22_1.npy ;
+python3 ../../extract.py all_fmri/sub-21_ses-2_task-rest_run-01_fullpre_stripped.nii.gz all_fmri/sub-21_ses-2_task-rest_run-01_fullpre_stripped_mask.nii.gz rmt_fullpre/rest/ses-2_eigs-21_1.npy ;
[…]

### 2020-08-23 data/Rest_v_LearningRecall/full_preprocess.py

@@ -1,0 +1,175 @@
+"""
+The main challenge here is that we have to delete intermediates in order to
+prevent completely filling up the HD with junk. However, if some intermediate
+step fails, we want to keep the last relevant pre-processing step.
+"""
+import argparse
+import nipype
+import os
+import pickle
+
+from glob import glob
+from nipype.interfaces.fsl import MCFLIRT, SliceTimer, BET
+from pathlib import Path
+from time import ctime
+from typing import List, Tuple
+from warnings import filterwarnings
+
+
+DATA = Path(__file__).resolve().parent / "all_fmri"
+LOGFILE = (
+    Path(__file__).resolve().parent
+    / f"{ctime().replace(' ', '_').replace(':', '.')}_pid{os.getpid()}__preprocess.log"
+)
+
+MCFLIRT_SUFFIX = "mcflirted.nii.gz"
+SLICETIME_SUFFIX = "stationary.nii.gz"
+BET_SUFFIX = "fullpre_stripped.nii.gz"
+
+
+class Log:
+    def __init__(self, logfile: Path = LOGFILE) -> None:
+        self.log: List[str] = []
+        if logfile.exists():
+            os.remove(logfile)
+        self.out = logfile
+
+    def append(self, s: str) -> None:
+        self.log.append(s)
+        with open(self.out, "a+") as log:
+            log.write(f"{s}\n")
+
+    def print(self) -> None:
+        print("\n".join(self.log))
+
+
+def mcflirt(infile: Path) -> Path:
+    # motion-correction FIRST supported by e.g.
+    # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6736626/
+    # given we are dealing with FSL
+    outfile = Path(str(infile).replace("bold.nii.gz", MCFLIRT_SUFFIX))
+    if outfile.exists():
+        return outfile
+    cmd = MCFLIRT(
+        in_file=str(infile),
+        out_file=str(outfile),
+        output_type="NIFTI_GZ",
+        save_mats=False,
+        save_rms=False,
+        stages=3,
+        stats_imgs=False,
+        terminal_output="stream",
+        mean_vol=False,  # speed up things
+    )
+
+    results = cmd.run()
+    return Path(results.outputs.out_file)
+
+
+def slicetime_correct(infile: Path, timings: Path, TR: float, slice_direction: str = "z") -> Path:
[…]

### 2020-08-23 data/Rest_v_LearningRecall/fully_preprocess.sh

@@ -1,0 +1,432 @@
+python3 full_preprocess.py -i all_fmri/sub-01_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-01_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-02_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-02_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-03_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-03_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-04_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-04_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-05_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-05_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-06_ses-1_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
+python3 full_preprocess.py -i all_fmri/sub-06_ses-2_task-rest_run-01_bold.nii.gz --TR 2.0 --timings slicetimes_rest.txt --direction z ;
[…]

### 2020-08-23 data/Rest_v_LearningRecall/preprocess.sh

@@ -1,0 +1,433 @@
+bet all_fmri/sub-01_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-01_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-01_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-01_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-02_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-02_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-02_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-02_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-03_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-03_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-03_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-03_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-04_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-04_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-04_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-04_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-05_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-05_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-05_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-05_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-06_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-06_ses-1_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-06_ses-2_task-rest_run-01_bold.nii.gz all_fmri/sub-06_ses-2_task-rest_run-01_stripped.nii.gz -F ;
+bet all_fmri/sub-07_ses-1_task-rest_run-01_bold.nii.gz all_fmri/sub-07_ses-1_task-rest_run-01_stripped.nii.gz -F ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/echo_interleave.py

@@ -1,0 +1,56 @@
+import nibabel as nib
+import numpy as np
+
+from glob import glob
+from numba import jit, prange
+from numpy import ndarray
+from pathlib import Path
+from tqdm import tqdm
+
+from typing import Any, Optional
+
+GLOB = str(Path(__file__).resolve().parent / "all_fmri" / "*bold.nii.gz")
+
+
+def something(v: Optional[Any]) -> bool:
+    return v is not None
+
+
+def make_cheaty_nii(orig: nib.Nifti1Image, array: np.array) -> nib.Nifti1Image:
+    """clone the header and extraneous info from `orig` and data in `array`
+    into a new Nifti1Image object, for plotting
+    """
+    affine = orig.affine
+    header = orig.header
+    # return new_img_like(orig, array, copy_header=True)
+    return nib.Nifti1Image(dataobj=array, affine=affine, header=header)
+
+
+@jit(nopython=True, cache=True, parallel=True)
+def interweave(m1: ndarray, m2: ndarray, m3: ndarray) -> ndarray:
+    x, y, z, T = m1.shape
+    m = np.empty((x, y, z, 3 * T), dtype=np.float64)
+    for t in prange(T):
+        m[:, :, :, 3 * t] = m1[:, :, :, t]
+        m[:, :, :, 3 * t + 1] = m2[:, :, :, t]
+        m[:, :, :, 3 * t + 2] = m3[:, :, :, t]
+    return m
+
+
+niis = [Path(p).resolve() for p in glob(GLOB)]
+echo1 = sorted(list(filter(something, [p if str(p).find("echo-1") > 0 else None for p in niis])))
+echo2 = sorted(list(filter(something, [p if str(p).find("echo-2") > 0 else None for p in niis])))
+echo3 = sorted(list(filter(something, [p if str(p).find("echo-3") > 0 else None for p in niis])))
+
+e1: Path
+e2: Path
+e3: Path
+for e1, e2, e3 in tqdm(zip(echo1, echo2, echo3), desc="Interweaving", total=len(echo1)):
+    orig = nib.load(str(e1))
+    mat1 = nib.load(str(e1)).get_fdata()
+    mat2 = nib.load(str(e2)).get_fdata()
+    mat3 = nib.load(str(e3)).get_fdata()
+    interweaved = interweave(mat1, mat2, mat3)
+    outfile = str(e1).replace("echo-1", "interweaved")
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/echo_interleave_fullpre.py

@@ -1,0 +1,83 @@
+import gc
+import multiprocessing as mp
+import nibabel as nib
+import numpy as np
+
+from glob import glob
+from multiprocessing import Pool
+from numba import jit
+from numpy import ndarray
+from pathlib import Path
+from tqdm import tqdm
+
+from typing import Any, Optional
+
+NIIS_GLOB = str(Path(__file__).resolve().parent / "all_fmri" / "*fullpre_stripped.nii.gz")
+MASKS_GLOB = str(Path(__file__).resolve().parent / "all_fmri" / "*fullpre_stripped_mask.nii.gz")
+
+
+def something(v: Optional[Any]) -> bool:
+    return v is not None
+
+
+def make_cheaty_nii(orig: nib.Nifti1Image, array: np.array) -> nib.Nifti1Image:
+    """clone the header and extraneous info from `orig` and data in `array`
+    into a new Nifti1Image object, for plotting
+    """
+    affine = orig.affine
+    header = orig.header
+    # return new_img_like(orig, array, copy_header=True)
+    return nib.Nifti1Image(dataobj=array, affine=affine, header=header)
+
+
+@jit(nopython=True, cache=True)
+def interweave(m1: ndarray, m2: ndarray, m3: ndarray) -> ndarray:
+    x, y, z, T = m1.shape
+    m = np.empty((x, y, z, 3 * T), dtype=np.float64)
+    for t in range(T):
+        m[:, :, :, 3 * t] = m1[:, :, :, t]
+        m[:, :, :, 3 * t + 1] = m2[:, :, :, t]
+        m[:, :, :, 3 * t + 2] = m3[:, :, :, t]
+    return m
+
+
+def interleave_echoes(e1, e2, e3):
+    orig = nib.load(str(e1))
+    mat1 = nib.load(str(e1)).get_fdata()
+    mat2 = nib.load(str(e2)).get_fdata()
+    mat3 = nib.load(str(e3)).get_fdata()
+    interweaved = interweave(mat1, mat2, mat3)
+    outfile = str(e1).replace("echo-1", "interweaved")
+    nib.save(make_cheaty_nii(orig, interweaved), outfile)
+
+
+def join_masks(m1, m2, m3):
+    orig = nib.load(str(m1))
+    only_brain = np.copy(np.array(orig.get_fdata(), dtype=bool))
+    only_brain &= np.array(nib.load(str(m2)).get_fdata(), dtype=bool)
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/echo_sum.py

@@ -1,0 +1,40 @@
+import nibabel as nib
+import numpy as np
+
+from glob import glob
+from pathlib import Path
+from pprint import pprint
+
+from typing import Any, Optional
+
+
+def something(v: Optional[Any]) -> bool:
+    return v is not None
+
+
+def make_cheaty_nii(orig: nib.Nifti1Image, array: np.array) -> nib.Nifti1Image:
+    """clone the header and extraneous info from `orig` and data in `array`
+    into a new Nifti1Image object, for plotting
+    """
+    affine = orig.affine
+    header = orig.header
+    # return new_img_like(orig, array, copy_header=True)
+    return nib.Nifti1Image(dataobj=array, affine=affine, header=header)
+
+
+niis = [Path(p).resolve() for p in glob("all_fmri/*bold.nii.gz")]
+echo1 = sorted(list(filter(something, [p if str(p).find("echo-1") > 0 else None for p in niis])))
+echo2 = sorted(list(filter(something, [p if str(p).find("echo-2") > 0 else None for p in niis])))
+echo3 = sorted(list(filter(something, [p if str(p).find("echo-3") > 0 else None for p in niis])))
+
+e1: Path
+e2: Path
+e3: Path
+for e1, e2, e3 in zip(echo1, echo2, echo3):
+    orig = nib.load(str(e1))
+    nii = orig.get_fdata()
+    nii = nii + nib.load(str(e2)).get_fdata()
+    nii = nii + nib.load(str(e3)).get_fdata()
+    outfile = str(e1).replace("echo-1", "summed")
+    nib.save(make_cheaty_nii(orig, nii), outfile)
+    print(f"Saved {e1.name.replace('run-01', 'run-0X')} summed echo image to {outfile}")

### 2020-08-23 data/Rest_v_Various_MultiEcho/echo_sum_fullpre.py

@@ -1,0 +1,70 @@
+import gc
+import multiprocessing as mp
+import nibabel as nib
+import numpy as np
+import sys
+
+from glob import glob
+from multiprocessing import Pool
+from pathlib import Path
+from pprint import pprint
+from tqdm import tqdm
+from typing import Any, Optional
+
+
+def is_some(v: Optional[Any]) -> bool:
+    return v is not None
+
+
+def make_cheaty_nii(orig: nib.Nifti1Image, array: np.array) -> nib.Nifti1Image:
+    """clone the header and extraneous info from `orig` and data in `array`
+    into a new Nifti1Image object, for plotting
+    """
+    affine = orig.affine
+    header = orig.header
+    # return new_img_like(orig, array, copy_header=True)
+    return nib.Nifti1Image(dataobj=array, affine=affine, header=header)
+
+
+niis = [Path(p).resolve() for p in glob("all_fmri/*fullpre_stripped.nii.gz")]
+echo1 = sorted(list(filter(is_some, [p if str(p).find("echo-1") > 0 else None for p in niis])))
+echo2 = sorted(list(filter(is_some, [p if str(p).find("echo-2") > 0 else None for p in niis])))
+echo3 = sorted(list(filter(is_some, [p if str(p).find("echo-3") > 0 else None for p in niis])))
+
+masks = [Path(p).resolve() for p in glob("all_fmri/*fullpre_stripped_mask.nii.gz")]
+mask1 = sorted(list(filter(is_some, [p if str(p).find("echo-1") > 0 else None for p in masks])))
+mask2 = sorted(list(filter(is_some, [p if str(p).find("echo-2") > 0 else None for p in masks])))
+mask3 = sorted(list(filter(is_some, [p if str(p).find("echo-3") > 0 else None for p in masks])))
+
+
+def sum_images(e1: Path, e2: Path, e3: Path):
+    orig = nib.load(str(e1))
+    nii = orig.get_fdata()
+    nii = nii + nib.load(str(e2)).get_fdata()
+    nii = nii + nib.load(str(e3)).get_fdata()
+    outfile = str(e1).replace("echo-1", "summed")
+    nib.save(make_cheaty_nii(orig, nii), outfile)
+
+
+def join_masks(m1: Path, m2: Path, m3: Path):
+    orig = nib.load(str(m1))
+    only_brain = np.copy(np.array(orig.get_fdata(), dtype=bool))
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract.sh

@@ -1,0 +1,497 @@
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_summed_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_summed_stripped_mask.nii.gz rmt/summed_eigs-31_04.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_echo-3_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_echo-3_stripped_mask.nii.gz rmt/echo-3_eigs-31_04.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_echo-2_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_echo-2_stripped_mask.nii.gz rmt/echo-2_eigs-31_04.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_echo-1_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_echo-1_stripped_mask.nii.gz rmt/echo-1_eigs-31_04.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_summed_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_summed_stripped_mask.nii.gz rmt/summed_eigs-31_03.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_echo-3_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_echo-3_stripped_mask.nii.gz rmt/echo-3_eigs-31_03.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_echo-2_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_echo-2_stripped_mask.nii.gz rmt/echo-2_eigs-31_03.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_echo-1_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_echo-1_stripped_mask.nii.gz rmt/echo-1_eigs-31_03.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_fullpre.sh

@@ -1,0 +1,249 @@
+mkdir -p rmt_fullpre/interleaved/task rmt_fullpre/summed/task &&
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_interweaved_fullpre_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-31_4.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_interweaved_fullpre_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-31_3.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-02_interweaved_fullpre_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-02_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-31_2.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-01_interweaved_fullpre_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-01_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-31_1.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-04_interweaved_fullpre_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-04_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-30_4.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-03_interweaved_fullpre_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-03_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-30_3.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-02_interweaved_fullpre_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-02_interweaved_fullpre_stripped_mask.nii.gz rmt_fullpre/interleaved/task/eigs-30_2.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_interleaved.sh

@@ -1,0 +1,156 @@
+mkdir -p rmt/interleaved/task rmt/interleaved/rest &&
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-04_interweaved_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-04_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-31_4.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-03_interweaved_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-03_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-31_3.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-02_interweaved_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-02_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-31_2.npy ;
+python3 ../../extract.py all_fmri/sub-31_task-cuedSGT_run-01_interweaved_stripped.nii.gz all_fmri/sub-31_task-cuedSGT_run-01_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-31_1.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-04_interweaved_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-04_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-30_4.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-03_interweaved_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-03_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-30_3.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-02_interweaved_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-02_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-30_2.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-cuedSGT_run-01_interweaved_stripped.nii.gz all_fmri/sub-30_task-cuedSGT_run-01_interweaved_stripped_mask.nii.gz rmt/interleaved/task/eigs-30_1.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_reshaped.sh

@@ -1,0 +1,63 @@
+mkdir -p rmt/interleaved/rest_reshaped rmt/summed/rest_reshaped
+python3 extract_reshaped_resting.py all_fmri/sub-01_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-01.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-02_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-02.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-03_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-03.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-04_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-04.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-05_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-05.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-06_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-06.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-07_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-07.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-08_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-08.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-09_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-09.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-10_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-10.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-11_task-rest_run-01_interweaved_stripped.nii.gz rmt/interleaved/rest_reshaped/eigs-11.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_reshaped_fullpre.sh

@@ -1,0 +1,63 @@
+mkdir -p rmt_fullpre/interleaved/rest_reshaped rmt_fullpre/summed/rest_reshaped
+python3 extract_reshaped_resting.py all_fmri/sub-01_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-01.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-02_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-02.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-03_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-03.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-04_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-04.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-05_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-05.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-06_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-06.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-07_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-07.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-08_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-08.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-09_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-09.npy ;
+python3 extract_reshaped_resting.py all_fmri/sub-10_task-rest_run-01_interweaved_fullpre_stripped.nii.gz rmt_fullpre/interleaved/rest_reshaped/eigs-10.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_reshaped_resting.py

@@ -1,0 +1,72 @@
+import argparse
+import nibabel as nib
+import numpy as np
+import os
+
+from empyricalRMT.eigenvalues import Eigenvalues
+from numba import jit
+from numpy import ndarray
+from pathlib import Path
+from skimage.transform import rescale
+
+DATA_ROOT = Path(__file__).resolve().parent
+FMRI_ROOT = DATA_ROOT / "all_fmri"
+TARGET_SHAPE = (64, 64, 33, 260)
+
+
+def make_cheaty_nii(orig: nib.Nifti1Image, array: np.array) -> nib.Nifti1Image:
+    """clone the header and extraneous info from `orig` and data in `array`
+    into a new Nifti1Image object, for plotting
+    """
+    affine = orig.affine
+    header = orig.header
+    # return new_img_like(orig, array, copy_header=True)
+    return nib.Nifti1Image(dataobj=array, affine=affine, header=header)
+
+
+def res(string: str) -> str:
+    return str(Path(string).resolve())
+
+
+@jit(nopython=True, cache=True, fastmath=True)
+def remask(img: ndarray, mask: ndarray):
+    x, y, z, T = img.shape
+    for i in range(x):
+        for j in range(y):
+            for k in range(z):
+                signal = img[i, j, k, :]
+                if np.sum(signal) == 0 or np.sum(signal * signal) == 0 or np.std(signal) == 0:
+                    mask[i, j, k, 0] = False
+
+
+parser = argparse.ArgumentParser(description="Handle reshaping")
+parser.add_argument("bold", metavar="<bold.nii.gz>", type=res, nargs=1, action="store")
+parser.add_argument("outfile", metavar="out.npy", type=res, nargs=1, action="store")
+
+args = parser.parse_args()
+
+img = nib.load(args.bold[0]).get_fdata()
+
+factors = [t / m for m, t in zip(img.shape, TARGET_SHAPE)]  # scaling factors per dimension
+factors[-1] = 1.0
+rescaled = rescale(img, factors, clip=False, preserve_range=True)
+
+mask = np.ones(rescaled.shape[:-1] + (1,), dtype=bool)
+remask(rescaled, mask)
+
+N, t = (np.prod(rescaled.shape[:-1]), rescaled.shape[-1])
+rescaled = rescaled.reshape([N, t])
+mask = mask.reshape(-1)
+
+brain = rescaled[mask, :]
+
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/extract_summed.sh

@@ -1,0 +1,31 @@
+python3 ../../extract.py all_fmri/sub-31_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-31_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-31_1.npy ;
+python3 ../../extract.py all_fmri/sub-30_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-30_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-30_1.npy ;
+python3 ../../extract.py all_fmri/sub-29_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-29_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-29_1.npy ;
+python3 ../../extract.py all_fmri/sub-28_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-28_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-28_1.npy ;
+python3 ../../extract.py all_fmri/sub-27_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-27_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-27_1.npy ;
+python3 ../../extract.py all_fmri/sub-26_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-26_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-26_1.npy ;
+python3 ../../extract.py all_fmri/sub-25_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-25_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-25_1.npy ;
+python3 ../../extract.py all_fmri/sub-24_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-24_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-24_1.npy ;
+python3 ../../extract.py all_fmri/sub-23_task-rest_run-01_summed_stripped.nii.gz all_fmri/sub-23_task-rest_run-01_summed_stripped_mask.nii.gz rmt/summed/rest/eigs-23_1.npy ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/fully_preprocess.sh

@@ -1,0 +1,465 @@
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/fully_preprocess_rest.sh

@@ -1,0 +1,93 @@
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-02_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-03_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-1_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-2_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-04_task-rest_run-01_echo-3_bold.nii.gz --TR 3.0 --timings slicetimes_rest.txt --direction k ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/fully_preprocess_task.sh

@@ -1,0 +1,372 @@
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-01_echo-1_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-01_echo-2_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-01_echo-3_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-02_echo-1_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-02_echo-2_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-02_echo-3_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-03_echo-1_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-03_echo-2_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-03_echo-3_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-04_echo-1_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-04_echo-2_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
+python3 full_preprocess.py -i all_fmri/sub-01_task-cuedSGT_run-04_echo-3_bold.nii.gz --TR 2.0 --timings slicetimes_task.txt --direction k ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/preprocess.sh

@@ -1,0 +1,621 @@
+bet all_fmri/sub-01_task-cuedSGT_run-01_echo-1_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-01_echo-1_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-01_echo-2_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-01_echo-2_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-01_echo-3_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-01_echo-3_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-01_summed_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-01_summed_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-02_echo-1_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-02_echo-1_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-02_echo-2_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-02_echo-2_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-02_echo-3_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-02_echo-3_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-02_summed_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-02_summed_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-03_echo-1_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-03_echo-1_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-03_echo-2_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-03_echo-2_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-03_echo-3_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-03_echo-3_stripped.nii.gz -F ;
[…]

### 2020-08-23 data/Rest_v_Various_MultiEcho/preprocess_interleaved.sh

@@ -1,0 +1,155 @@
+bet all_fmri/sub-01_task-cuedSGT_run-01_interweaved_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-01_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-02_interweaved_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-02_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-03_interweaved_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-03_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-cuedSGT_run-04_interweaved_bold.nii.gz all_fmri/sub-01_task-cuedSGT_run-04_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-01_task-rest_run-01_interweaved_bold.nii.gz all_fmri/sub-01_task-rest_run-01_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-02_task-cuedSGT_run-01_interweaved_bold.nii.gz all_fmri/sub-02_task-cuedSGT_run-01_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-02_task-cuedSGT_run-02_interweaved_bold.nii.gz all_fmri/sub-02_task-cuedSGT_run-02_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-02_task-cuedSGT_run-03_interweaved_bold.nii.gz all_fmri/sub-02_task-cuedSGT_run-03_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-02_task-cuedSGT_run-04_interweaved_bold.nii.gz all_fmri/sub-02_task-cuedSGT_run-04_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-02_task-rest_run-01_interweaved_bold.nii.gz all_fmri/sub-02_task-rest_run-01_interweaved_stripped.nii.gz -F ;
+bet all_fmri/sub-03_task-cuedSGT_run-01_interweaved_bold.nii.gz all_fmri/sub-03_task-cuedSGT_run-01_interweaved_stripped.nii.gz -F ;
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/extract.sh

@@ -1,0 +1,76 @@
+python3 ../../extract.py derivatives/pain/placebo/sub-75_task-rest_bold.nii.gz derivatives/pain/placebo/sub-75_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-75.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-74_task-rest_bold.nii.gz derivatives/pain/placebo/sub-74_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-74.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-73_task-rest_bold.nii.gz derivatives/pain/placebo/sub-73_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-73.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-72_task-rest_bold.nii.gz derivatives/pain/placebo/sub-72_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-72.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-68_task-rest_bold.nii.gz derivatives/pain/placebo/sub-68_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-68.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-67_task-rest_bold.nii.gz derivatives/pain/placebo/sub-67_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-67.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-64_task-rest_bold.nii.gz derivatives/pain/placebo/sub-64_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-64.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-63_task-rest_bold.nii.gz derivatives/pain/placebo/sub-63_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-63.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-62_task-rest_bold.nii.gz derivatives/pain/placebo/sub-62_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-62.npy &
+python3 ../../extract.py derivatives/pain/placebo/sub-60_task-rest_bold.nii.gz derivatives/pain/placebo/sub-60_task-rest_bold_mask.nii.gz rmt/pain/placebo/eigs-60.npy &
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/extract_fullpre.sh

@@ -1,0 +1,77 @@
+mkdir -p rmt_fullpre/pain/placebo rmt_fullpre/pain/duloxetine rmt_fullpre/controls &&
+python3 ../../extract.py pain/data/placebo/sub-75/func/sub-75_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-75/func/sub-75_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-75.npy
+python3 ../../extract.py pain/data/placebo/sub-74/func/sub-74_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-74/func/sub-74_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-74.npy
+python3 ../../extract.py pain/data/placebo/sub-73/func/sub-73_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-73/func/sub-73_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-73.npy
+python3 ../../extract.py pain/data/placebo/sub-72/func/sub-72_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-72/func/sub-72_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-72.npy
+python3 ../../extract.py pain/data/placebo/sub-68/func/sub-68_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-68/func/sub-68_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-68.npy
+python3 ../../extract.py pain/data/placebo/sub-67/func/sub-67_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-67/func/sub-67_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-67.npy
+python3 ../../extract.py pain/data/placebo/sub-64/func/sub-64_task-rest_fullpre_stripped.nii.gz pain/data/placebo/sub-64/func/sub-64_task-rest_fullpre_stripped_mask.nii.gz rmt_fullpre/pain/placebo/eigs-64.npy
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/full_preprocess.py

@@ -1,0 +1,172 @@
+"""
+The main challenge here is that we have to delete intermediates in order to
+prevent completely filling up the HD with junk. However, if some intermediate
+step fails, we want to keep the last relevant pre-processing step.
+"""
+import argparse
+import nipype
+import os
+import pickle
+
+from glob import glob
+from nipype.interfaces.fsl import MCFLIRT, SliceTimer, BET
+from pathlib import Path
+from time import ctime
+from typing import List, Tuple
+from warnings import filterwarnings
+
+
+DATA = Path(__file__).resolve().parent
+LOGFILE = (
+    Path(__file__).resolve().parent
+    / f"{ctime().replace(' ', '_').replace(':', '.')}_pid{os.getpid()}__preprocess.log"
+)
+
+MCFLIRT_SUFFIX = "mcflirted.nii.gz"
+SLICETIME_SUFFIX = "stationary.nii.gz"
+BET_SUFFIX = "fullpre_stripped.nii.gz"
+
+
+class Log:
+    def __init__(self, logfile: Path = LOGFILE) -> None:
+        self.log: List[str] = []
+        if logfile.exists():
+            os.remove(logfile)
+        self.out = logfile
+
+    def append(self, s: str) -> None:
+        self.log.append(s)
+        with open(self.out, "a+") as log:
+            log.write(f"{s}\n")
+
+    def print(self) -> None:
+        print("\n".join(self.log))
+
+
+def mcflirt(infile: Path) -> Path:
+    # motion-correction FIRST supported by e.g.
+    # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6736626/
+    # given we are dealing with FSL
+    outfile = Path(str(infile).replace("bold.nii.gz", MCFLIRT_SUFFIX))
+    if outfile.exists():
+        return outfile
+    cmd = MCFLIRT(
+        in_file=str(infile),
+        out_file=str(outfile),
+        output_type="NIFTI_GZ",
+        save_mats=False,
+        save_rms=False,
+        stages=3,
+        stats_imgs=False,
+        terminal_output="stream",
+        mean_vol=False,  # speed up things
+    )
+
+    results = cmd.run()
+    return Path(results.outputs.out_file)
+
+
+def slicetime_correct(infile: Path, timings: Path, TR: float, slice_direction: str = "z") -> Path:
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/fully_preprocess.sh

@@ -1,0 +1,76 @@
+python3 full_preprocess.py -i controls/data/sub-01/func/sub-01_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-02/func/sub-02_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-03/func/sub-03_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-04/func/sub-04_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-05/func/sub-05_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-06/func/sub-06_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-07/func/sub-07_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-08/func/sub-08_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-09/func/sub-09_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-10/func/sub-10_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-11/func/sub-11_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-12/func/sub-12_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
+python3 full_preprocess.py -i controls/data/sub-13/func/sub-13_task-rest_bold.nii.gz --TR 2.5 --timings slicetimes.txt --direction z ;
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/preprocess.sh

@@ -1,0 +1,77 @@
+mkdir -p derivatives/controls derivatives/pain/duloxetine derivatives/pain/placebo &&
+bet controls/data/sub-01/func/sub-01_task-rest_bold.nii.gz derivatives/controls/sub-01_task-rest_bold.nii.gz -F &
+bet controls/data/sub-02/func/sub-02_task-rest_bold.nii.gz derivatives/controls/sub-02_task-rest_bold.nii.gz -F &
+bet controls/data/sub-03/func/sub-03_task-rest_bold.nii.gz derivatives/controls/sub-03_task-rest_bold.nii.gz -F &
+bet controls/data/sub-04/func/sub-04_task-rest_bold.nii.gz derivatives/controls/sub-04_task-rest_bold.nii.gz -F &
+bet controls/data/sub-05/func/sub-05_task-rest_bold.nii.gz derivatives/controls/sub-05_task-rest_bold.nii.gz -F &
+bet controls/data/sub-06/func/sub-06_task-rest_bold.nii.gz derivatives/controls/sub-06_task-rest_bold.nii.gz -F &
+bet controls/data/sub-07/func/sub-07_task-rest_bold.nii.gz derivatives/controls/sub-07_task-rest_bold.nii.gz -F &
+bet controls/data/sub-08/func/sub-08_task-rest_bold.nii.gz derivatives/controls/sub-08_task-rest_bold.nii.gz -F &
+bet controls/data/sub-09/func/sub-09_task-rest_bold.nii.gz derivatives/controls/sub-09_task-rest_bold.nii.gz -F &
+bet controls/data/sub-10/func/sub-10_task-rest_bold.nii.gz derivatives/controls/sub-10_task-rest_bold.nii.gz -F &
+bet controls/data/sub-11/func/sub-11_task-rest_bold.nii.gz derivatives/controls/sub-11_task-rest_bold.nii.gz -F &
+bet controls/data/sub-12/func/sub-12_task-rest_bold.nii.gz derivatives/controls/sub-12_task-rest_bold.nii.gz -F &
+bet controls/data/sub-13/func/sub-13_task-rest_bold.nii.gz derivatives/controls/sub-13_task-rest_bold.nii.gz -F &
[…]

### 2020-08-23 data/Rest_w_Healthy_v_OsteoPain/saveshapes.sh

@@ -1,0 +1,76 @@
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-75_task-rest_bold.nii.gz derivatives/pain/placebo/sub-75_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-75.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-74_task-rest_bold.nii.gz derivatives/pain/placebo/sub-74_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-74.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-73_task-rest_bold.nii.gz derivatives/pain/placebo/sub-73_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-73.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-72_task-rest_bold.nii.gz derivatives/pain/placebo/sub-72_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-72.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-68_task-rest_bold.nii.gz derivatives/pain/placebo/sub-68_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-68.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-67_task-rest_bold.nii.gz derivatives/pain/placebo/sub-67_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-67.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-64_task-rest_bold.nii.gz derivatives/pain/placebo/sub-64_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-64.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-63_task-rest_bold.nii.gz derivatives/pain/placebo/sub-63_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-63.npy &
+python3 ../../save_shapes.py derivatives/pain/placebo/sub-62_task-rest_bold.nii.gz derivatives/pain/placebo/sub-62_task-rest_bold_mask.nii.gz rmt/pain/placebo/shapes-62.npy &
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/extract.sh

@@ -1,0 +1,22 @@
+python3 ../../extract.py derivatives/sub-22_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-22_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-22.npy ; 
+python3 ../../extract.py derivatives/sub-21_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-21_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-21.npy ;
+python3 ../../extract.py derivatives/sub-20_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-20_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-20.npy ;
+python3 ../../extract.py derivatives/sub-19_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-19_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-19.npy ;
+python3 ../../extract.py derivatives/sub-18_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-18_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-18.npy ;
+python3 ../../extract.py derivatives/sub-17_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-17_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-17.npy ;
+python3 ../../extract.py derivatives/sub-16_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-16_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-16.npy ;
+python3 ../../extract.py derivatives/sub-15_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-15_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/eigs-15.npy ;
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/extract_fullpre.sh

@@ -1,0 +1,89 @@
+mkdir -p rmt/full_pre &&
+python3 ../../extract.py data/all_bold/sub-22_ses-2_task-rest_acq-fullbrain_run-2_fullpre_stripped.nii.gz data/all_bold/sub-22_ses-2_task-rest_acq-fullbrain_run-2_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-2_eigs-22_2.npy ;
+python3 ../../extract.py data/all_bold/sub-22_ses-2_task-rest_acq-fullbrain_run-1_fullpre_stripped.nii.gz data/all_bold/sub-22_ses-2_task-rest_acq-fullbrain_run-1_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-2_eigs-22_1.npy ;
+python3 ../../extract.py data/all_bold/sub-22_ses-1_task-rest_acq-fullbrain_run-2_fullpre_stripped.nii.gz data/all_bold/sub-22_ses-1_task-rest_acq-fullbrain_run-2_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-1_eigs-22_2.npy ;
+python3 ../../extract.py data/all_bold/sub-22_ses-1_task-rest_acq-fullbrain_run-1_fullpre_stripped.nii.gz data/all_bold/sub-22_ses-1_task-rest_acq-fullbrain_run-1_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-1_eigs-22_1.npy ;
+python3 ../../extract.py data/all_bold/sub-21_ses-2_task-rest_acq-fullbrain_run-2_fullpre_stripped.nii.gz data/all_bold/sub-21_ses-2_task-rest_acq-fullbrain_run-2_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-2_eigs-21_2.npy ;
+python3 ../../extract.py data/all_bold/sub-21_ses-2_task-rest_acq-fullbrain_run-1_fullpre_stripped.nii.gz data/all_bold/sub-21_ses-2_task-rest_acq-fullbrain_run-1_fullpre_stripped_mask.nii.gz rmt/full_pre/ses-2_eigs-21_1.npy ;
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/extract_remain.sh

@@ -1,0 +1,66 @@
+python3 ../../extract.py derivatives/sub-01_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz derivatives/sub-01_ses-1_task-rest_acq-fullbrain_run-2_stripped_mask.nii.gz rmt/ses-1_eigs-01_2.npy ;
+python3 ../../extract.py derivatives/sub-01_ses-2_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-01_ses-2_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/ses-2_eigs-01_1.npy ;
+python3 ../../extract.py derivatives/sub-01_ses-2_task-rest_acq-fullbrain_run-2_stripped.nii.gz derivatives/sub-01_ses-2_task-rest_acq-fullbrain_run-2_stripped_mask.nii.gz rmt/ses-2_eigs-01_2.npy ;
+python3 ../../extract.py derivatives/sub-02_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz derivatives/sub-02_ses-1_task-rest_acq-fullbrain_run-2_stripped_mask.nii.gz rmt/ses-1_eigs-02_2.npy ;
+python3 ../../extract.py derivatives/sub-02_ses-2_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-02_ses-2_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/ses-2_eigs-02_1.npy ;
+python3 ../../extract.py derivatives/sub-02_ses-2_task-rest_acq-fullbrain_run-2_stripped.nii.gz derivatives/sub-02_ses-2_task-rest_acq-fullbrain_run-2_stripped_mask.nii.gz rmt/ses-2_eigs-02_2.npy ;
+python3 ../../extract.py derivatives/sub-03_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz derivatives/sub-03_ses-1_task-rest_acq-fullbrain_run-2_stripped_mask.nii.gz rmt/ses-1_eigs-03_2.npy ;
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/full_preprocess.py

@@ -1,0 +1,174 @@
+"""
+The main challenge here is that we have to delete intermediates in order to
+prevent completely filling up the HD with junk. However, if some intermediate
+step fails, we want to keep the last relevant pre-processing step.
+"""
+import argparse
+import nipype
+import os
+import pickle
+
+from glob import glob
+from nipype.interfaces.fsl import MCFLIRT, SliceTimer, BET
+from pathlib import Path
+from time import ctime
+from typing import List, Tuple
+from warnings import filterwarnings
+
+
+DATA = Path(__file__).resolve().parent / "data/all_bold"
+LOGFILE = (
+    Path(__file__).resolve().parent
+    / f"{ctime().replace(' ', '_').replace(':', '.')}_pid{os.getpid()}__preprocess.log"
+)
+
+MCFLIRT_SUFFIX = "mcflirted.nii.gz"
+SLICETIME_SUFFIX = "stationary.nii.gz"
+BET_SUFFIX = "fullpre_stripped.nii.gz"
+
+class Log:
+    def __init__(self, logfile: Path = LOGFILE) -> None:
+        self.log: List[str] = []
+        if logfile.exists():
+            os.remove(logfile)
+        self.out = logfile
+
+    def append(self, s: str) -> None:
+        self.log.append(s)
+        with open(self.out, "a+") as log:
+            log.write(f"{s}\n")
+
+    def print(self) -> None:
+        print("\n".join(self.log))
+
+
+def mcflirt(infile: Path) -> Path:
+    # motion-correction FIRST supported by e.g.
+    # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6736626/
+    # given we are dealing with FSL
+    outfile = Path(str(infile).replace("bold.nii.gz", MCFLIRT_SUFFIX))
+    if outfile.exists():
+        return outfile
+    cmd = MCFLIRT(
+        in_file=str(infile),
+        out_file=str(outfile),
+        output_type="NIFTI_GZ",
+        save_mats=False,
+        save_rms=False,
+        stages=3,
+        stats_imgs=False,
+        terminal_output="stream",
+        mean_vol=False,  # speed up things
+    )
+
+    results = cmd.run()
+    return Path(results.outputs.out_file)
+
+
+def slicetime_correct(infile: Path, timings: Path, TR: float, slice_direction: str = "z") -> Path:
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/fully_preprocess.sh

@@ -1,0 +1,88 @@
+python3 full_preprocess.py -i data/all_bold/sub-01_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-01_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-2_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-02_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-02_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-2_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-03_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-03_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
+python3 full_preprocess.py -i data/all_bold/sub-03_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz -t slicetimes.txt --TR 3.0 --direction k ;
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/preprocess.sh

@@ -1,0 +1,23 @@
+mkdir -p derivatives &&
+bet data/sub-01/ses-1/func/sub-01_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-01_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-02/ses-1/func/sub-02_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-02_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-03/ses-1/func/sub-03_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-03_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-04/ses-1/func/sub-04_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-04_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-05/ses-1/func/sub-05_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-05_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-06/ses-1/func/sub-06_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-06_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-07/ses-1/func/sub-07_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-07_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-08/ses-1/func/sub-08_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-08_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/sub-09/ses-1/func/sub-09_ses-1_task-rest_acq-fullbrain_run-1_bold.nii.gz derivatives/sub-09_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/preprocess_remain.sh

@@ -1,0 +1,66 @@
+bet data/all_bold/sub-01_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-01_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-01_ses-2_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-02_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-02_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F &
+bet data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-02_ses-2_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-03_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-03_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-03_ses-2_task-rest_acq-fullbrain_run-1_bold.nii.gz data/all_bold/sub-03_ses-2_task-rest_acq-fullbrain_run-1_stripped.nii.gz -F ;
+bet data/all_bold/sub-03_ses-2_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-03_ses-2_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
+bet data/all_bold/sub-04_ses-1_task-rest_acq-fullbrain_run-2_bold.nii.gz data/all_bold/sub-04_ses-1_task-rest_acq-fullbrain_run-2_stripped.nii.gz -F &
[…]

### 2020-08-23 data/Rest_w_VigilanceAttention/saveshapes.sh

@@ -1,0 +1,22 @@
+python3 ../../save_shapes.py derivatives/sub-22_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-22_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-22.npy ; 
+python3 ../../save_shapes.py derivatives/sub-21_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-21_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-21.npy ;
+python3 ../../save_shapes.py derivatives/sub-20_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-20_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-20.npy ;
+python3 ../../save_shapes.py derivatives/sub-19_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-19_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-19.npy ;
+python3 ../../save_shapes.py derivatives/sub-18_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-18_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-18.npy ;
+python3 ../../save_shapes.py derivatives/sub-17_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-17_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-17.npy ;
+python3 ../../save_shapes.py derivatives/sub-16_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-16_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-16.npy ;
+python3 ../../save_shapes.py derivatives/sub-15_ses-1_task-rest_acq-fullbrain_run-1_stripped.nii.gz derivatives/sub-15_ses-1_task-rest_acq-fullbrain_run-1_stripped_mask.nii.gz rmt/shapes-15.npy ;
[…]

### 2020-08-23 data/Task_Learning/extract_all.sh

@@ -1,0 +1,192 @@
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-01_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-01_stripped_mask.nii.gz nosleep/eigs-02_1.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-02_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-02_stripped_mask.nii.gz nosleep/eigs-02_2.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-03_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-03_stripped_mask.nii.gz nosleep/eigs-02_3.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-04_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-04_stripped_mask.nii.gz nosleep/eigs-02_4.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-05_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-05_stripped_mask.nii.gz nosleep/eigs-02_5.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-06_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-06_stripped_mask.nii.gz nosleep/eigs-02_6.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-07_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-07_stripped_mask.nii.gz nosleep/eigs-02_7.npy;
+python3 ../../extract.py nosleep/sub-02_ses-1_task-satellite_run-08_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-08_stripped_mask.nii.gz nosleep/eigs-02_8.npy;
+python3 ../../extract.py nosleep/sub-03_ses-1_task-satellite_run-01_stripped.nii.gz nosleep/sub-03_ses-1_task-satellite_run-01_stripped_mask.nii.gz nosleep/eigs-03_1.npy;
[…]

### 2020-08-23 data/Task_Learning/saveshapes.sh

@@ -1,0 +1,192 @@
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-01_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-01_stripped_mask.nii.gz nosleep/shapes-02_1.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-02_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-02_stripped_mask.nii.gz nosleep/shapes-02_2.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-03_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-03_stripped_mask.nii.gz nosleep/shapes-02_3.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-04_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-04_stripped_mask.nii.gz nosleep/shapes-02_4.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-05_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-05_stripped_mask.nii.gz nosleep/shapes-02_5.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-06_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-06_stripped_mask.nii.gz nosleep/shapes-02_6.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-07_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-07_stripped_mask.nii.gz nosleep/shapes-02_7.npy;
+python3 ../../save_shapes.py nosleep/sub-02_ses-1_task-satellite_run-08_stripped.nii.gz nosleep/sub-02_ses-1_task-satellite_run-08_stripped_mask.nii.gz nosleep/shapes-02_8.npy;
+python3 ../../save_shapes.py nosleep/sub-03_ses-1_task-satellite_run-01_stripped.nii.gz nosleep/sub-03_ses-1_task-satellite_run-01_stripped_mask.nii.gz nosleep/shapes-03_1.npy;
[…]

### 2020-08-23 data/bundle.sh

@@ -1,0 +1,67 @@
+BASE=/home/derek/Desktop/RandomMatrixFMRI/Writing/repro/data;
+
+mkdir -p $BASE/plots/Rest_v_LearningRecall;
+mkdir -p $BASE/plots/Rest_v_Various_SingleSubject;
+mkdir -p $BASE/plots/Rest_w_Healthy_v_OsteoPain;
+mkdir -p $BASE/plots/Rest_w_VigilanceAttention/task-attend;
+mkdir -p $BASE/plots/Rest_w_VigilanceAttention/vigilance;
+mkdir -p $BASE/plots/Rest_w_VigilanceAttention/weekly-attend;
+
+cp $BASE/Rest_v_LearningRecall/rmt/plots/poly_deg7_rig-maxL-20_rest_v_task.png $BASE/plots/Rest_v_LearningRecall
+cp $BASE/Rest_v_LearningRecall/rmt/plots/poly_deg7_rest_v_task_nnsd.png $BASE/plots/Rest_v_LearningRecall
+cp $BASE/Rest_v_LearningRecall/rmt/plots/poly_deg7_brody-mle.png $BASE/plots/Rest_v_LearningRecall
+cp $BASE/Rest_v_LearningRecall/rmt/plots/marchenko_shifted.png $BASE/plots/Rest_v_LearningRecall
+cp $BASE/Rest_v_LearningRecall/rmt/plots/marchenko.png $BASE/plots/Rest_v_LearningRecall
+cp $BASE/Rest_v_LearningRecall/rmt/plots/poly_deg7_var-maxL-10_rest_v_task.png $BASE/plots/Rest_v_LearningRecall
+
+cp $BASE/Rest_v_Various_SingleSubject/rmt/plots/poly_deg7_rig-maxL-20_rest_v_task.png $BASE/plots/Rest_v_Various_SingleSubject
+cp $BASE/Rest_v_Various_SingleSubject/rmt/plots/poly_deg7_rest_v_task_nnsd.png $BASE/plots/Rest_v_Various_SingleSubject
+cp $BASE/Rest_v_Various_SingleSubject/rmt/plots/poly_deg7_brody-mle.png $BASE/plots/Rest_v_Various_SingleSubject
+cp $BASE/Rest_v_Various_SingleSubject/rmt/plots/marchenko.png $BASE/plots/Rest_v_Various_SingleSubject
[…]

### 2020-08-23 data/count_runs.sh

@@ -1,0 +1,1 @@
+find -iname "*bold.nii.gz" | grep -v afni | grep -v derivatives | grep -v prefrontal | grep -v sleep | grep -v slept | grep -v summed | grep -v interweave | grep -v -x 'echo-[23]' | sort -r | wc -l

### 2020-08-23 data/count_scans.sh

@@ -1,0 +1,1 @@
+find -iname "*bold.nii.gz" | grep -v afni | grep -v derivatives | grep -v prefrontal | grep -v sleep | grep -v slept | grep -v -x 'echo-[123]' | sort -r | wc -l

### 2020-08-23 data/get_runs.sh

@@ -1,0 +1,1 @@
+find -iname "*bold.nii.gz" | grep -v afni | grep -v derivatives | grep -v prefrontal | grep -v sleep | grep -v slept | grep -v summed | grep -v interweave | grep -v -x 'echo-[23]' | sort -r 

### 2021-09-08 paper/latex/NetworkNeuroscience/NETNSample.tex
add latex start

%% Documentclass:
%% Network Neuroscience
\documentclass[NETN,manuscript]{stjour-new}

%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Article Type:
%% Default is Research.

%% Or, choose one of these options:
%% Research, Methods, Data, Review, and Perspective

\articletype{Perspective}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% author definitions should be placed here:

%% example definition
\def\taupav{\tau_{\mathrm{Pav}}}

%% common path to your figure files (for organization purposes)
\graphicspath{{figs/}}

\begin{document}
\title[Shortened Version of Title]{Title of Article}
\subtitle{Subtitle Here}

%% If shortened title for running head is needed so that the article title can fit
%%   in the running head, use [] argument, ie,
%%
%%   \title[Shortened Title for Running Head]{Title of Article}
%%   \subtitle{Subtitle Here}

%% Since we use \affil{} in the argument of the \author command, we
%% need to supply another version of the author names, without \affil{}
%% to be used for running heads:

\author[Author Names]
{Author Names with affiliations\affil{1},
Another Name\affil{2}, Still another Name\affil{2},\\
\and Final Name\affil{1}}

\affiliation{1}{Department, Institution, City, Country}

%ie.
%\affiliation{1}{Gatsby Computational Neuroscience Unit, University
%College London, London, United Kingdom}

\affiliation{2}{Another Department, Institution, City, Country}

%ie
%\affiliation{2}{Center for Studies in
%Behavioral Neurobiology, Concordia University, Montreal, Quebec,
%Canada}

\correspondingauthor{Author Name}{Corresponding author email address}

% ie,
%\correspondingauthor{Ritwik K. Niyogi}{ritwik.niyogi@gatsby.ucl.ac.uk}

\keywords{(a series of capitalized words, separated with commas)}

%ie
%\keywords{Work, Leisure, Normative, Microscopic,  Reinforcement Learning, Economics}

\begin{abstract}
Abstract text here.
\end{abstract}

\begin{authorsummary}
Author summary here. 
\end{authorsummary}

\section{Sample Section}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\subsection{Sample Subsection}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\subsubsection{Sample Subsubsection}
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.
Text here. Text here. Text here. Text here.

\section{Sample equations}
\begin{equation}
\label{eq:rhoCHT}
\rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
\left[C_L(\taupav+\tau_L) \right]   +
\displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
\Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
\Biggr]            }      {P +
\mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
\displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
\Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
}
\end{equation}
As long as
$RI - K_LP > 
\frac{1}{\beta}$
\begin{equation}
%\def\theequation{5.1}
\left.\begin{array}{lrcl}
&\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
(P+\taupav )}    \\[12pt]
\hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
K_LP)-1}  
\label{eq:analytical_linear}
\end{array}\right\}\hbox to 1.25in{\hfill}
\end{equation} 
Text finishing first page.
Text finishing first page.
Text finishing first page.
Text finishing first page.
Text finishing first page.
Text finishing first page.
Text finishing first page.

\begin{boxedtext}{Comparative Analysis of Different Classes of Networks} 
Going beyond the examination of shared topological features across
nervous systems, the generalized mathematical language of graph theory
also offers tools for the comparison of the organization of brain
networks to other classes of network studied
by different scientific
disciplines.

Many real-world systems operate as some sort of
interaction or communication network, including, for example, social
networks, gene regulatory networks, computer networks, and
transportation networks. Similar to brain networks, many of these
real-world networks display an efficient small-world organization, a
pronounced community structure with densely connected modules, as well
as the formation of hubs and rich clubs. Going beyond the
comparison of networks within the class of nervous systems, the field
of `comparative network analysis' examines commonalities and
differences across a range of network classes.

\begin{figure}
\includegraphics[width=\hsize]{Fig1}
\caption{Here is the caption.}
\end{figure}
\end{boxedtext}

\newpage
\begin{boxedtext}{Comparative Analysis of Different Classes of Networks} 
Going beyond the examination of shared topological features across
nervous systems, the generalized mathematical language of graph theory
also offers tools for the comparison of the organization of brain
networks to other classes of network studied by different scientific
disciplines. Many real-world systems operate as some sort of
interaction or communication network, including, for example, social
networks, gene regulatory networks, computer networks, and
transportation networks. Similar to brain networks, many of these
real-world networks display an efficient small-world organization, a
pronounced community structure with densely connected modules, as well
as the formation of hubs and rich clubs. Going beyond the
comparison of networks within the class of nervous systems, the field
of `comparative network analysis' examines commonalities and
differences across a range of network classes.

\begin{table}[ht]
\caption{Here is the caption.}

\centerline{\begin{tabular}{|c|c|c|}
\hline
one&two&three\\
\hline
four&five&six\\
\hline
\end{tabular}}
\end{table}
\end{boxedtext}

\subsection{Technical Terms}
All NETN article types require Technical Terms.

Identify approximately 10 key terms that are mentioned in your article and whose usage and definition may not be familiar across the broad readership of the journal. 
Provide brief (20-word or less) definitions for each term, avoiding in these definitions the use of jargon, or highly technical or specialized language. 
When the article is typeset, the Technical Terms will appear in the margins at or near their first mention in the text.

In your manuscript, bold the first occurrence of each \textbf{Technical Term} and then provide a list of the terms and their definitions at the end of the manuscript after the references.

\subsection{Simple code sample}

\begin{code}
\begin{verbatim}
procedure bubbleSort( A : list of sortable items )
    n = length(A)
    repeat
       newn = 0
       for i = 1 to n-1 inclusive do
          if A[i-1] > A[i] then
             swap(A[i-1], A[i])
             newn = i
          end if
       end for
       n = newn
    until n = 0
end procedure
\end{verbatim}
\end{code}

\subsection{Algorithm environment}
%% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
%% See documentation for algorithmic.sty for more information on formatting algorithms.

\begin{algorithm}[h]
\caption{A sample in an algorithm environment.}
\begin{algorithmic}
\If {$i\geq maxval$}
    \State $i\gets 0$
\Else
    \If {$i+k\leq maxval$}
        \State $i\gets i+k$
    \EndIf
\EndIf
\end{algorithmic}
\end{algorithm}

\section{Itemized Lists}

\subsection{Roman list:}

\begin{enumerate}
\item[(i)] at high 
payoffs, subjects work almost continuously.
\item[(ii)] at low payoffs, they 
engage in leisure all at once, in long bouts after working.
\item[(iii)] subjects work continuously for the entire price duration, as long as
the price is not very long;
\item[(iv)] the duration of leisure bouts is variable.
\end{enumerate}

\subsection{Numbered list:}

\begin{enumerate}
\item at high 
payoffs, subjects work almost continuously, engaging in little leisure
inbetween work bouts; 
\item at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by
Y-AB, to be published separately);  
\item the duration of leisure bouts is variable.
\end{enumerate}

\subsection{Bulleted list:}

\begin{itemize}
\item at high 
payoffs, subjects work almost continuously, engaging in little leisure
inbetween work bouts; 
\item at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by
Y-AB, to be published separately);  
\item the duration of leisure bouts is variable.
\end{itemize}

\subsection{Description list:}
\begin{description}
\item[High payoffs:] at high 
payoffs, subjects work almost continuously, engaging in little leisure
inbetween work bouts; 
\item[Low payoffs:] at low payoffs, they 
engage in leisure all at once, in long bouts after working, rather
than distributing the same amount of leisure time into multiple short
leisure bouts; 
\item[Continuous work:] subjects work continuously for the entire price duration, as long as
the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); 
\item[Duration:] the duration of leisure bouts is variable.
\end{description}

\newpage
\section{Sample citations}
For general information on the correct form for citations using
the APA 6 format, see the following sites:
\href{https://owl.english.purdue.edu/owl/resource/560/02/}
{APA 6, In-text citations, The Basics} and
\href{https://owl.english.purdue.edu/owl/resource/560/03/}
{APA 6, In-text citations}

\section{Natbib citation mark up}

\subsection{Single citations}
\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90}+&\dogray Jones et al. (1990)\\
\verb+\citet[chap. 2]{jon90}+&\dogray Jones et al. (1990, chap. 2)\\
    \verb+\citep{jon90}+	    &\dogray   	(Jones et al., 1990)\\
    \verb+\citep[chap. 2]{jon90}+ 	&\dogray    	(Jones et al., 1990, chap. 2)\\
    \verb+\citep[see][]{jon90}+ 	 &\dogray    	(see Jones et al., 1990)\\
    \verb+\citep[see][chap. 2]{jon90}+ 	&\dogray    	(see Jones et al., 1990, chap. 2)\\
    \verb+\citet*{jon90}+ 	    &\dogray    	Jones, Baker, and Williams (1990)\\
    \verb+\citep*{jon90}+	    & \dogray   	(Jones, Baker, and Williams,
    1990) \\
\end{tabular}

For example, some citations from the NETNbibsamp.bib database:

citet: \citet{bullmore2009complex}, citep: \citep{gomez2009analysis}, and
citep*: \citep*{de2012cortical}

\subsection{Multiple citations}
Multiple citations may be made by including more than one citation
key in the \verb+\cite+ command argument.

\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90,jam91}+&\dogray Jones et al. (1990); James et al. (1991)\\
\verb+\citep{jon90,jam91}+&\dogray (Jones et al., 1990; James et al. 1991)\\
\verb+\citep{jon90,jon91}+&\dogray (Jones et al., 1990, 1991)\\
\verb+\citep{jon90a,jon90b}+&\dogray (Jones et al., 1990a,b)\\
\end{tabular}

For example, multiple citations from the bibsamp.bib database:
citet: \citet{nooner2012nki,hutchison2013resting},
 citep:
\citep{tagliazucchi2012dynamic, de2012cortical}

As you see, the citations are automatically hyperlinked to their
reference in the bibliography.

\newpage

\section{Sample figures}

\begin{figure}[h] 
\centerline{\includegraphics[width=\textwidth]{Fig1.pdf}}
\caption{(Colour online) \textbf{Task and key features of the
 data.} \\
 A) Cumulative handling time (CHT) task. Grey bars denote work
(depressing a lever), white gaps show leisure. The subject must
accumulate work up to a total period of time called the
\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
$2$s each time the price is attained, during which the lever is retracted so it cannot
work; not shown).}
\label{fig:task_data}
\end{figure}
%% this command ends a page but does not fill the bottom with white space:
\eject

\begin{figure}[ht] 
\widefigure{\fullpagewidth}{Fig1.pdf}
\caption{(Colour online) \textbf{Task and key features of the
 data.} \\
 A) Cumulative handling time (CHT) task. Grey bars denote work
(depressing a lever), white gaps show leisure. The subject must
accumulate work up to a total period of time called the
\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
$2$s each time the price is attained, during which the lever is retracted so it cannot
work; not shown).}
\label{fig:task_data2}
\end{figure}

\newpage
\section{Sample tables}

\begin{table}[!ht]
\caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
\label{tab:label}
\centering
\begin{tabular}{lc}
\hline
 Run  & Time (min)  \\
\hline
  $l1$  & 260   \\
  $l2$  & 300   \\
  $l3$  & 340   \\
  $h1$  & 270   \\
  $h2$  & 250   \\
  $h3$  & 380   \\
  $r1$  & 370   \\
  $r2$  & 390   \\
\hline
\multicolumn{2}{l}{$^{a}$Table note text here.}
\end{tabular}
\end{table}

\begin{table}[ht]
\widecaption{Sample table taken from [treu03]\label{tbl-1}}
\begin{widetable}
\advance\tabcolsep-1pt
\footnotesize
\begin{tabular}{ccrrccccccccc}
\hline
\bf 
POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
&\multicolumn1c{\bf Y} &\bf
RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
IAP2 &\bf star &\bf E &\bf Comment\\
\hline
0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
edge \\
\hline\\[-6pt]
\multicolumn{13}{l}{%
Table 2 is published in its entirety in the electronic
edition of the {\it Astrophysical Journal}.}\\[3pt]
\multicolumn{13}{l}{%
$^a$ Sample footnote for table 2.}\\[3pt]
\multicolumn{13}{l}{%
$^b$ Another sample footnote for table 2.}
\end{tabular}
\end{widetable}
\end{table}

\begin{table}[p]
\rotatebox{90}{\vbox{\hsize=\textheight
\caption{Here is a caption for a table that is found in landscape
mode.}
\begin{tabular}{ccrrccccccccc}
\hline
\bf 
POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
&\multicolumn1c{\bf Y} &\bf
RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
IAP2 &\bf star &\bf E &\bf Comment\\
\hline
0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
edge \\
\hline\\[-6pt]
\multicolumn{13}{l}{%
Table 2 is published in its entirety in the electronic
edition of the {\it Astrophysical Journal}.}\\[3pt]
\multicolumn{13}{l}{%
$^a$ Sample footnote for table 2.}\\[3pt]
\multicolumn{13}{l}{%
$^b$ Another sample footnote for table 2.}
\end{tabular}
}}
\end{table}
\clearpage
\begin{boxedtext}{Tools for comparison of networks} 
Going beyond the examination of shared topological features across
nervous systems, the generalized mathematical language of graph theory
also offers tools for the comparison of the organization of brain
networks to other classes of network studied by different scientific
disciplines.

From $\mathcal{W}$, we can estimate the variability in the fluctuations of the functional connection between nodes $i$ and $j$ over time as:
\begin{equation}
s_{ij}=\sqrt{\frac{1}{T-L}\sum_{t=1}^{T-L+1}(W_{ij}(t) - m_{ij})}
\end{equation}
where $m_{ij}=\frac{1}{T-L+1}\sum_{t=1}^{T-L+1}W_{ij}(t)$ is the mean
dynamic functional connectivity over time.

Many real-world systems operate as some sort of
interaction or communication network, including, for example, social
networks, gene regulatory networks, computer networks, and
transportation networks. Similar to brain networks, many of these
real-world networks display an efficient small-world organization, a
pronounced community structure with densely connected modules, as well
as the formation of hubs and rich clubs. Going beyond the
comparison of networks within the class of nervous systems, the field
of `comparative network analysis' examines commonalities and
differences across a range of network classes.
\end{boxedtext}

Example of table continuing over pages:

\begin{center}
\begin{longtable}{ccc@{}}
\caption{ApJ costs from 1991 to 2013
\label{tab:table}} \\[2pt]
\hline
\bf Year & \bf Subscription & \bf Publication \\
 & \bf cost &\bf charges\\
 & \bf(\$) & \bf (\$/page)\\
\hline
\endfirsthead

\multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
\multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
\hline
\bf Year & \bf Subscription & \bf Publication \\
 & \bf cost &\bf charges\\
 & \bf(\$) & \bf (\$/page)\\
\hline
\endhead
\\\hline
\\[-8pt]
\multicolumn{3}{r}{\it Table continued on next page}\\ 
\endfoot

\hline
\endlastfoot

1991 & 600 & 100 \\
1992 & 650 & 105 \\
1993 & 550 & 103 \\
1994 & 450 & 110 \\
1995 & 410 & 112 \\
1996 & 400 & 114 \\
1997 & 525 & 115 \\
1998 & 590 & 116 \\
1999 & 575 & 115 \\
2000 & 450 & 103 \\
2001 & 490 &  90 \\
2002 & 500 &  88 \\
2003 & 450 &  90 \\
2004 & 460 &  88 \\
2005 & 440 &  79 \\
2006 & 350 &  77 \\
2007 & 325 &  70 \\
2008 & 320 &  65 \\
2009 & 190 &  68 \\
2010 & 280 &  70 \\
2011 & 275 &  68 \\
2012 & 150 &  56 \\
2013 & 140 &  55 \\
\end{longtable}
\end{center}

\acknowledgments
The Funder and award ID information you input at submission will be introduced by the publisher under a Funding Information head during production. 
Please use this space for any additional acknowledgements and verbiage required by your funders.

%% ie.,

% The authors thank Laurence Aitchison for fruitful discussions.  RKN
% and PD received funding from the Gatsby Charitable Foundation. Y-AB,
% RBS, KC and PS received funding from Canadian Institutes of Health
% Research grant $MOP74577$,
% Fond de recherche Qu\'{e}bec - Sant\'{e} (Group grant to the Groupe 
% de recherche en neurobiologie comportementale, Shimon Amir, P.I.), and
% Concordia University Research Chair (Tier I).

\section{Supporting Information}
This is an optional section. Please use this space to provide information about any supporting information referred to in your manuscript.

\section{Competing Interests}
This is an optional section. If you declared a conflict of interest when you submitted your manuscript, please  use this space to provide details about this conflict.

\section{Making Your Bibliography for a Network Neuroscience Article}
{\it Network Neuroscience} uses the APA author-date  bibliography style,
apacite.bst. For more
information on apacite, for examples in how to make your .bib file and more, see:\\
\href{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}

\noindent
(In spite of the mention of apacite cite commands, please use only
Natbib commands for in text citations, as shown above.)

\subsection{BibTeX}
You will need to use BibTeX to form your bibliography; typing in the
references would be
a huge and unpleasant task. Look at the NETNSample.bbl file and you'll see why
typing in the bibitems would be difficult.

For a good basic introduction to using BibTeX, see\\
\href{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}

When you use BibTeX, the form of the bibliography will be correct. You
don't need to supply a bibliography style, since that is built into
the stjour.cls file when the NETN option is used
(\verb+\documentclass[NETN]{stjour}+).

\newpage
\subsection{Sample citations}

Here are some samples using \verb+\citep{}+:\\
\citep{bullmore2009complex,
gomez2009analysis,
rubinov2011weight,
power2014methods,
sporns2015modular,
scheeringa2012eeg,
fortunato2007resolution,
reichardt2006statistical,
smith2009correspondence,
sporns2011networks,
liegeois2014cerebral}

And more using \verb+\citet{}+\\
\citet{allen2012tracking,
calhoun2014chronnectome,
liu2013time,
fisher1915frequency,
gonzalez2014spatial,
damaraju2014dynamic,
hutchison2013resting,
de2012cortical,
tagliazucchi2012dynamic,
nooner2012nki}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%
%% The bibliography

\bibliography{NETNbibsamp}

\section{Technical Terms}

\textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

\textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

\textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

\textbf{Technical Term} a key term that is mentioned in an NETN article and whose usage and definition may not be familiar across the broad readership of the journal.

%% No appendices allowed in Network Neuroscience style
%\appendix

\end{document}

### 2021-09-08 paper/latex/preamble.tex
add latex start

\documentclass{article} % This command is used to set the type of document you are working on such as an article, book, or presentation

\usepackage{geometry} % This package allows the editing of the page layout
\usepackage{amsmath}  % This package allows the use of a large range of mathematical formula, commands, and symbols
\usepackage{graphicx}  % This package allows the importing of images

\newcommand{\question}[2][]{
    \begin{flushleft}
    \textbf{Question #1}: \textit{#2}
    \end{flushleft}
}
\newcommand{\sol}{\textbf{Solution}:} %Use if you want a boldface solution line
\newcommand{\maketitletwo}[2][]{
    \begin{center}
        \Large{\textbf{Assignment #1}}

        Course Title

        \vspace{5pt}
        \normalsize{Matthew Frenkel}

        \today
        \vspace{15pt}
    \end{center}
}

### 2021-09-08 paper/latex/rmt.tex
add latex start

\input{preamble.tex}
\begin{document}

\maketitletwo[5]  % Optional argument is assignment number
%Keep a blank space between maketitletwo and \question[1]

\question[1]{Here is my first question}

YOUR SOLUTION HERE

\question[2]{Here is my second question}

YOUR SOLUTION HERE

\question[3]{What is the \Large{$\int_0^2 x^2 \, dx $}\normalsize{. Show all steps}}

\begin{align*}
\int_0^2 x^2 &= \left. \frac{x^3}{3} \right|_0^2 \\
                &= \frac{2^3}{3}-\frac{0^3}{3}\\
                &= \frac{8}{3}
\end{align*}

\end{document}

### 2021-09-09 paper/latex/rmt.tex
updatre

@@
 \input{preamble.tex}
+
+\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
+\author{Derek M. Berger \and Gurpreet S. Matharoo \and Jacob Levman}
+\date{\today}
+
 \begin{document}
+\maketitle
 
-\maketitletwo[5]  % Optional argument is assignment number
-%Keep a blank space between maketitletwo and \question[1]
+\begin{abstract}
+    Previous studies have investigated the potential of using analytic techniques from Random Matrix
+    Theory (RMT) to investigate magnetic resonance imaging (MRI) data. We assess the potential
+    application of RMT-based features for the analysis of functional MRI (fMRI) across diverse
+    datasets. As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain
+    features extracted from voxel-wise functional connectivity to contain information useful for
+    classifying between various psychological processes, (2) assess these features’
+    predictive—rather than explanatory—value, (3) investigate the effect of varying RMT analysis
+    methods on the robustness of study findings, and (4) make general-purpose code publicly
+    available for users to extract these features from a wide variety of data. We find preliminary
+    evidence suggesting that RMT-inspired features may have unique potential in analyses of fMRI
+    functional connectivity.
 
-\question[1]{Here is my first question}
+    \subparagraph{Index Terms:} feature extraction, functional connectivity, functional magnetic
+    resonance imaging, random matrix theory, voxelwise connectivity
 
-YOUR SOLUTION HERE
+    \subparagraph{Authors}
+    Derek Berger is with the department of Computer Science, St. Francis Xavier University, Antigonish
+    NS B2G 2W5, Canada (e-mail: dberger@stfx.ca).
 
-\question[2]{Here is my second question}
+    Gurpreet Matharoo is with ACENET, and the Department of Physics, St. Francis Xavier University,
+    Antigonish NS B2G 2W5, Canada (e-mail: gmatharo@stfx.ca).
 
-YOUR SOLUTION HERE
+    Jacob Levman is with the Department of Computer Science, St. Francis Xavier University,
+    Antigonish NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).
+\end{abstract}
 
-\question[3]{What is the \Large{$\int_0^2 x^2 \, dx $}\normalsize{. Show all steps}}
+\section{Introduction}
 
-\begin{align*}
-\int_0^2 x^2 &= \left. \frac{x^3}{3} \right|_0^2 \\
-                &= \frac{2^3}{3}-\frac{0^3}{3}\\
-                &= \frac{8}{3}
-\end{align*}
+In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
+(BOLD) signals are related to neural activity. It is common to investigate statistical relationships
+between the BOLD signals through functional connectivity analyses, where correlations between
+collections of these signals are examined to infer connections between different voxels or regions
+of interest (ROIs) within the brain.
+
+Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
+complex functional connectivity networks are ubiquitous [1]–[4]. This complexity suggests fMRI is a
+candidate to be studied using Random Matrix Theory (RMT), a set of mathematical tools originally
+developed some 50 years ago to solve complex problems in nuclear physics [5], [6]. In various
+physical systems where nuclei can have different possible states, precisely modeling the entirety of
+the interactions between nuclei is often intractable. For such large, complex systems, the
+eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
+of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
+spectrum statistically, complex systems can be investigated by comparison to the universal
+properties predicted by RMT.
+
+\begin{sloppypar}
+These universal properties have been observed in diverse phenomena. In small-scale systems, RMT
+universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
+disordered mesoscopic systems [5]–[11], and at larger scales, RMT has been applied to atmospheric
+physics [12], stock cross-correlations [13], social networks [14], random networks [15],
+network-formation in liquids [16], [17] and amorphous clusters [18]–[21]. Within biological systems,
+RMT has also been used to successfully model aspects of  amino acid functional relationships [22],
+synchrony in epileptic seizures [23], and in protein-protein interactions both in different species
+[24] and breast cancer [25].
+\end{sloppypar}
+
+In recent years, RMT analyses have been applied to study brain functioning.  The earliest  study
+demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
+resemble those of the Gaussian Orthogonal Ensemble (GOE) [26].  Somewhat more recently, RMT was used
+to evaluate the quality of whole brain features extracted from fMRI data [27], [28]. RMT has also
+been used in diffusion MRI to aid in the selection of the number of components to employ in
+principal-component reduction analysis and denoising [29]–[31].
+
+Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
+differences between rest and task states [32], between subjects with and without ADHD [33], and
+between pain and non-pain states [34]. Across these three studies, the spectra of resting or
+low-attention states exhibited properties closest to the GOE.
+
+These findings suggest that certain aspects of psychological processes might be characterized, in
+part, by features computed from the eigenvalues of fMRI correlation matrices, and that these
+features might vary in an interpretable manner across psychological processes. If this is the case,
+RMT could have unique potential in characterizing the functioning of the human brain.
+
+Here, we use a novel, voxel-based approach, and expand the applications of RMT to analyze fMRI scans
+from  diverse datasets1. We extract RMT-inspired whole-brain features from voxelwise functional
+connectivity data, and assess the predictive value of these features. We also examine in detail the
+robustness of the RMT methodology to various common analytic choices.
+
+\section{Materials and Methods}
+
+\subsection{Datasets}
+
+As our primary intention is to examine the potential utility of RMT-derived features across a wide
+variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
+psychological processes, we examine a wide selection of fMRI datasets available on the OpenNeuro
+platform [36].  This allows for a broad assessment of the potential of voxelwise RMT features.
+
+However, since previous studies employing RMT for feature extraction [33], [34] have partially
+interpreted their findings with respect to attentional processes, we have elected to use datasets
+that involve attention in some way. This includes measures of attention or vigilance, or comparisons
+between a condition involving more attention and focus (e.g. a task state) or conditions which
+should involve considerably less attention or focus (e.g. resting-state).
+
+\begin{center}
+\texttt{\textbf{INSERT TABLE 1 HERE}}
+\end{center}
+
+Dataset scan parameters and acquisition details are summarized in Table I, and sample and subgroup
+sizes are summarized in Table II. For all datasets, we use the raw (unpreprocessed) scans, and
+perform our own preprocessing (detailed later). Given the complexity of these datasets, we highlight
+below only those aspects relevant to the current paper.
+
+\subsubsection{Dataset 1}
+Dataset 1 (referred to in Fig. and tables with the label "PSYCH") has detailed and extensive
+information available in  [37]. Subjects completed a battery of psychological measurements, and
+various resting-state fMRI (rs-fMRI) images were obtained. Only the whole-brain 7-Tesla rs-fMRI
+scans were used for the current paper, to make the results comparable to the other included
+datasets. For the purpose of keeping with the "attention" theme of the analyses in this paper, three
+pairings of subgroups were constructed via z-score splits of the various sum-scores of scale items
+involving concepts closely related to attention.
+
+\paragraph{Trait Attention}
+The PANAS-X [38] is a 9-point Likert-type self-report measure that queries the extent to which
+subjects agree particular affective terms have applied to them over a recent time period (e.g. a few
+weeks). An ad hoc (i.e. face-valid) sum-score was constructed from the scale items: "attentive",
+"alert", "interested", "concentrating", and reverse-scored items "sluggish", "tired", "sleepy",
+"drowsy". Subjects with a z-score greater / less than zero on this sum-score were placed in the high
+/ low "trait attention" category
+
+\paragraph{Task Attention}
+The Conjunctive Continuous Performance Task (CCPT [39]) is a behavioural measure of sustained
+attention where subjects must react only if they see a red square among other different coloured
+figures. For this paper, a sum score was constructed by subtracting the number of false positive and
+false negative responses from the number of success responses, and subjects with a z-score greater /
+less than zero on this scale were placed in high / low "task attention" category.
+
+\paragraph{Vigilance}
+The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
+state, e.g. "my thoughts involved other people", "I thought about something positive" (translated
+from German) [40]. One question assesses alertness ("I was fully awake" / "war ich vollkommen wach")
+and was used to create high / low "vigilance" conditions in the same manner as the previous two splits.
+
+\subsubsection{Dataset 2}
+Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in [41]. Subjects
+included healthy controls ("nopain" condition) and individuals with knee osteoarthritis ("pain"
+condition) treated for 2 weeks with either placebo ("pain" condition) or duloxetine ("duloxetine"
+condition). An additional grouping was also created with both the placebo and duloxetine data
+("allpain" condition), allowing for six subgroup comparisons. All subjects were imaged with
+whole-brain rs-fMRI. Scan parameters are documented in Table I. The use of rs-fMRI and RMT for
+examining pain makes this dataset complementary to [34].
+
+\subsubsection{Dataset 3}
+Dataset 3  (referred to here with the label "REFLECT", due to the "reflective” nature of the
+cognitive tasks) is described in full detail by DuPre, Luh and Spreng [42]. Both multi-echo rs- and
+task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks involved viewing
+various affective images (see [43] for details) and performing tasks related to autobiographical
+memory, prospection, or theory-of-mind reasoning. Continuing with the theme of broadly comparing
+high versus low attention throughout this paper, the different tasks were simply treated generically
+as "attending", in contrast to the non-attentive resting-state scans. This resulted in four task
+state scans and one resting-state scan per subject.
+
+In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
+the "summed" condition, echoes were combined via summation, a method which has value in providing an
+image with overall increased contrast [44]. In the "interleave" condition, echoes were interleaved
+at each voxel: that is, if we denote the voxel's ith signal value from echo j as , then the
+interleaved voxel of the new image has the time-series
+\([t_{1,1}, t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
+
+This creates a physically-meaningful 4D
+image that contains the full, original multi-echo fMRI time-series data without any alterations. It
+should be noted that the purpose of these combination methods is not to create a 4D volume that
+allows optimal detection of the purely task-linked BOLD signal, but to ensure all scan information
+was used with minimal additional analytic assumptions.
+
+\subsubsection{Dataset 4}
+Dataset 4 (referred to in Fig. and tables with the label "PARK" or "PARKINSONS") is documented and
+described in detail in [45]. Scan parameters are documented in Table I. Subjects with non-demented
+Parkinson’s Disease (PD) and healthy controls performed a number of repetitions of the Attention
+Network Task (ANT) [46]. Subgroup details are available in Table I. As the ANT task demands high
+attention, it is not as straightforward to create "high" or "low" attention groups. Instead, since
+attention presumably is high and varies little in this dataset, and subgroups instead differ by
+virtue of the presence or absence of PD, any results here serve as a useful comparison point for the
+results from studies that should clearly involve attentional differences. This dataset also included
+extensively preprocessed scans, which were included in analyses to examine the effects of
+preprocessing on the extracted RMT features.
+
+\subsubsection{Dataset 5}
+Dataset 5 (referred to in Fig. and tables with the label "LEARN") is documented and described in
+detail in [47]. In this study, both rs-fMRI and task fMRI scans were available for all subjects,
+allowing for within-subject comparisons. The complex task involved multiple phases, wherein subjects
+were exposed to named, black-and-white outlined satellite images related by family-resemblance of
+the components (see Fig. 1 of [47]). After this learning task, subjects were tested on their memory
+of the various learned features. Scans corresponding to both the initial learning and subsequent
+testing phase were included in the high-attention "task" subgroup for the purposes of the analysis
+of this paper. Resting state scans were included in the low-attention "rest" subgroup.
+
+\subsection{Preprocessing}
+Some consideration should be given to whether it is appropriate to apply standard preprocessing
+steps before extracting RMT features, as preprocessing steps should be developed, tested, and
+employed always with respect to a set of inferential goals and a specific analytic framework.
+However, these preprocessing steps are increasingly recognized to have complex effects on fMRI
+analyses [48]. For example, spatial smoothing has clear benefits when employing a general linear
+model [49]–[51], but is less motivated in functional network analyses with small ROIs, where the
+effect of smoothing can be considerable [50], [52], [53], with these effects interacting complexly
+with factors like field strength and image resolution [54]. Similarly, image registration is usually
+required to allow comparisons across (or even within) subjects, but the choice of registration
+strategy can have dramatic effects on functional network analyses [55].
+
+The RMT features extracted here are based on functional correlations rather than ROI activations,
+and are completely insensitive to voxel location (spatial re-arrangement of an image’s voxels does
+not change the eigenvalues of the correlation matrix). Thus RMT features summarize all voxel-wise
+correlations in a whole-brain and location-free manner, which brings into question some of the usual
+motivations for performing spatial smoothing, or registration to identify ROIs. In addition, if
+image dimensions are identical between subjects, the RMT features (by virtue of their definition and
+the unfolding procedure) already allow for direct comparisons. Thus, the usual motivations for
+registration are also no longer clear when extracting RMT features.
+
+Given these particularities, as well as the novelty of the RMT-based methods and the diversity of
+the datasets, we elected to minimize the complexity of the preprocessing pipeline to ensure that
+data processing is kept as standardized as possible. Ultimately, different preprocessing pipelines
+for RMT fMRI analyses can be expected to have different benefits and shortcomings depending on the
+question being investigated, and possibly also the specific RMT features being relied upon, but such
+an investigation is beyond the scope of the current study.
+
+For this paper’s analyses, brain extraction is necessary to limit analyses to the appropriate
+voxels, motion correction is necessary to allow proper brain extraction, and both motion-correction
+and slice-time correction simplify the interpretation of voxel-voxel correlations across voxel
+locations. Whereas physiological noise regression and bias field correction would perhaps be
+appropriate here, there is insufficient data to perform these steps for all datasets (Table II).
+Thus, motion-correction was performed with FSL's MCFLIRT [56], slicetime correction with the default
+method in FSL, and brain extraction was performed with BET [57]. Due to differences in the size of
+the REFLECT data, the higher-resolution resting-state data was interpolated to have the same spatial
+dimensions as the task data.
+
+\section{Conclusion}
+
+\paragraph{Declarations}
+Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
+Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
+Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
+Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
+(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
+R0168020), and a St. Francis Xavier University UCR grant to JL.
 

### 2021-09-09 paper/latex/rmt.tex
paste in main content

@@
 
+\subsection{RMT-Derived Features}
+This paper examines features informed by various aspects of RMT, and which can be extracted from
+empirical correlation matrices. To keep with previous studies, we limit our analyses to features
+extracted only from the eigenvalues of the matrices of correlations between all voxel signals.
+
+Two classes of features were examined in this study: those that require unfolding, and those that do
+not. In order to compare the observed eigenvalue distribution to some theoretical predictions, the
+eigenvalues must be unfolded [5], [6]. The details and motivation behind the unfolding procedure are
+well documented in [5]. Practically, the unfolding process can be viewed as a smoothing and
+rescaling procedure where the originally observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is
+smoothly mapped to the unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
+if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing function is not known, and
+so must be approximated [5], [6]. Typically, a polynomial is chosen [58].
+
+A number of statistics (or "spectral observables" [6]) can then be computed from this unfolded
+spectrum. In the literature, the two most common points of comparison for these unfolded features
+are typically the features of the Poisson or Gaussian Diagonal Ensemble (GDE) and the GOE (see e.g.
+[12], [14], [15], [24]–[26], [32], [33]). The GOE (or GDE) consists of real, symmetric matrices
+where entries (or only diagonal entries) are normal random variables with zero mean and unit
+variance. When the empirically observed features are close to those of the GOE, this is usually
+interpreted as indicating a high degree of randomness and statistical independence in the data, and
+when the observed features are closer to those expected in the GDE or Poisson case, this is
+interpreted as indicating a high degree of coherence or dependence [12], [14], [15], [24]–[26],
+[32]–[34].
+
+The Nearest-Neighbours Spacings Distribution (NNSD) is the most basic RMT feature, and is the
+distribution of the spacings (differences) of adjacent unfolded eigenvalues. Normally, the NNSD is
+analysed visually via a histogram of the distribution of spacings, however it is also possible to
+analyse it more formally by fitting the Brody distribution [59]:
+
+\[
+p(s) = \alpha (1 + \beta) s^\alpha \exp({\beta s^{\alpha+1}}),
+\quad \alpha = \left[ \Gamma \left( \frac{\beta+2}{\beta+1} \right) \right]^{\beta+1}
+\]
+
+where \(\beta\) is free to vary within \([0, 1]\), with \(\beta = 0\)  producing the expected NNSD
+for the GDE case, and \(\beta = 1\) producing the expected NNSD for the GOE case. In this paper, we
+fit, via maximum likelihood, the empirically observed spacings with the Brody distribution, and
+report the Brody parameter, \(\beta\), as an important unfolded feature to investigate.
+
+Whereas the NNSD and Brody parameter characterize the local properties of eigenvalues close on the
+spectrum, the long-range properties relating eigenvalues distant on the spectrum can be
+characterized by both the spectral rigidity and level number variance [5], [6]. The spectral
+rigidity, \(\Delta_3(L)\), is calculated for any real value \(L\) as:
+
+\[
+\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
+\]
+
+where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
+\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
+\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
+line to \(\eta(\lambda)\) on \([c, c+L]\) [5]. The level number variance, \(\Sigma^2(L)\), is
+closely related to the spectral rigidity [6], and is calculated as:
+
+\[
+\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
+\]
+
+where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
+and \(\langle \cdot \rangle_c\) are as above [5]. Both the spectral rigidity and level number
+variance are also computed and investigated in this study.
+
+In addition to the unfolded features suggested by RMT, there are some potential "raw" features that
+do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
+information related to the system, and is also investigated as an RMT-inspired feature.
+
+The distribution of eigenvalues of large rectangular matrices with finite-variance, independent and
+identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP) distribution
+[60], [61]:
+
+\[
+\rho(\lambda) =
+\frac{1}{2\pi\sigma^2}
+\frac{\sqrt{(\lambda_+ - \lambda)(\lambda - \lambda_-)}}{r\lambda}, \;
+\lambda_{\pm} = \sigma^2\left(1 \pm \sqrt{\lambda}\right)^2
+\]
+
+where \(\sigma^2\) is the true variance of the distribution of each i.i.d. entry (equal to 1 in the
+case of correlation matrices), and where \(r = N / T\), for matrices of dimension \(N \times T\).
+While real-world data will not, in general, be i.i.d., the extent to which spectra overlap with or
+are well-described by the MP distribution has been used to guide analytical decisions regarding the
+relevance of empirically observed eigenvalues. That is, since the majority of the eigenvalues of
+random correlation matrices will fall within \([\lambda_{-}, \lambda_{+}]\), eigenvalues observed to
+fall within this region are typically considered to be the result of noise [29]–[31], [62]. We
+examine as a potential feature the proportion of eigenvalues that fall within this range, and call
+this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
+adjustments of [63] which takes into account the larger degree of correlation in empirical data.
+
+\subsection{RMT Feature Extraction}
+The entire 4-dimensional \(m \times n \times n \times p \times t\) scan array for each run,
+excluding voxels with constant signals (which lead to undefined correlations), and excluding voxels
+identified as non-brain via preprocessing, was flattened to a two-dimensional \(m \cdot n \cdot p
+\times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
+value of \(t\) for a scan can be inferred from the number of volumes in Table I. The various RMT
+features mentioned above were then extracted via the empyricalRMT library [64] developed and made
+publicly available as part of this study, as well as via specific code for this paper (which is
+documented and available online [65]).
+
+As both the spectral rigidity and level number variance can be highly sensitive to the choice of
+smoothing procedure used in the unfolding method [58], [66]–[69], all unfolded features in this
+study were calculated using polynomials of degree 5, 7, 9, 11, and 13 as the smoothing technique. As
+much of the variance in the unfolded spectral statistics can be due to the largest eigenvalues
+acting as anchor points during smoothing [66], [69], all unfolded features were also calculated
+twice (in addition to each polynomial degree) to examine the effects of trimming on spectral
+observables: once with only the largest eigenvalue omitted, and once with the largest 20 eigenvalues
+omitted.
+
+\subsection{Feature Evaluation}
+To evaluate the extracted RMT features, for each possible pairing of subgroups in each dataset, each
+feature was used to predict subject subgroup membership. That is, for each dataset, and for each
+possible pairing of subgroups, all scans belonging to a subgroup were pooled together and treated as
+independent samples of the subgroup of interest. So, for example, in the LEARN dataset, this
+resulted in a comparison of 384 (\(16 \times 4\)) task scans, to 48 (\(2 \times 24\)) rest scans. When it was not
+reasonable to compare scans across sessions, sessions were analyzed separately. This was the case
+for the PSYCH dataset, where psychological measurements, and thus subgroups, differed from session
+to session.
+
+A wide variety of standard, out-of-the-box classifiers available in the Python scikit-learn library
+[70] were then employed to gauge the predictive value of the RMT features. To allow for a wide range
+of flexibilities, classifiers employing k-nearest neighbours (KNN), with k values in \(\{1, 2, 3,
+\dots, 10, 20\}\), logistic regression (LR), linear discriminant analysis (LDA), and a
+support-vector machine (SVM) were employed. Default parameter values were used when possible (see
+[65] for precise function calls used).
+
+The accuracy of each binary classifier was then validated with leave-one-out cross-validation
+(LOOCV) on all datasets, for all combinations of trimming and unfolding choices.
+
+\section{Results}
+
+\subsection{Overview}
+Due to space limitations, we highlight only key results here. Extensive overviews of the full
+results are available in the supplementary materials. Visual summaries of the prediction accuracies
+for all features and analytic choices are presented in Fig. S1--4. Prediction accuracies for only the
+multivariate RMT features are summarized in Fig. S5--7, and the distributions of the scalar features
+are presented in Fig. S8--10. Prediction accuracies for the scalar features across analytic choices
+are summarized in Fig. S11--12.
+
+Numeric (mean) summaries across datasets, broken down by classifier and feature are available in
+Tables S11--18. Complete (unsummarized) tables of all predictions for all combinations of features,
+unfolding degree, and trimming are available as supplementary material in the spreadsheet
+\texttt{prediction\_tables\_fullpre.ods}. Unless otherwise noted, subsequent figures and analyses were
+computed using the preprocessed data: full results for the raw data are available as spreadsheets in
+the supplementary material in \texttt{prediction\_tables.ods}.
+
+With the exception of the LEARNING dataset, the RMT-inspired features all demonstrated some
+predictive utility (Fig. S1-4). For example, there were strong differences in RMT features between
+task and rest conditions for the REFLECT dataset, but the features were of variable predictive
+utility in the PSYCH dataset, sometimes outperforming the naïve classifier, and sometimes not.
+Although the RMT features were useful in the comparison involving duloxetine and controls (no pain)
+in the OSTEO dataset, these features mostly did not aid in classifying between any of the other
+possible pairings of subgroups in this dataset (Fig. S1-4).
+
+Although the performance of classifiers trained on RMT features was variable, there was a tendency
+for the raw and largest eigenvalues to be the most predictively useful features, especially in the
+REFLECT and PARK data in discriminating between Parkinson's and controls (Fig. S5-7. S13; but see
+Discussion).
+
+\subsection{Raw Features}
+As evident from Fig. S8-10, the scalar raw features (largest eigenvalue, noise ratio, and shifted
+noise ratio) varied more within than between subgroups. They were not generally useful for
+classifying conditions (Tables S15-17), with two or three notable exceptions.
+
+The noise ratio (but not shifted noise ratio) differed consistently between the duloxetine and
+control groups in the OSTEO data, allowing for prediction accuracies of up to 72\% compared to 51\%
+for the naive classifier (Fig. S11-12 top-right panel, Table S5). The largest eigenvalue was also
+useful in distinguishing between subjects self-reporting high versus low vigilance across sessions:
+accuracies up to 64\% and 65\% (compared to naïve accuracies of 55\% and 50\%) for the first and
+second sessions, respectively (Table S6). Finally, for the REFLECT data, task and rest were
+perfectly distinguishable with various features and with some classifiers.
+
+The sorted, raw eigenvalues, however, more consistently contained information that allowed some
+separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
+consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
+had the most consistent predictive value for distinguishing individuals with Parkinson's from
+controls, or distinguishing those with osteopathic pain taking duloxetine from controls ("nopain" in
+Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
+comparing those differing in self-reported vigilance and high vs. low scorers in the attention task
+(Table S7--8).
+
+In all these relationships, normalizing the eigenvalues to the interval \([0,1]\) prior to
+prediction considerably improved prediction accuracy (Tables S9-12).
+
+\paragraph{Role of Classifer}
+The raw eigenvalues generally had the most predictive utility when employing either an SVM
+classifier or KNN classifier with k > 3 (Table S11, first row). When the raw eigenvalues were first
+min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
+predicting subgroup memberships (Table S12), however without normalization, KNN with k = 8 or k = 9
+generally best predicted subgroup membership across datasets (Table S11).
+
+\subsection{Unfolded Features}
+The Brody parameter rarely resulted in predictions exceeding naïve accuracies in these comparisons:
+for most unfoldings and comparisons, the Brody parameter was either constant and equal for all
+subjects (e.g. exactly 0.999 or 0.0001) or deviated from these values only for a small number of
+subjects. This is because the observed NNSDs were usually well outside of the GOE-GDE bounds.
+
+The predictive utility of the unfolded RMT features varied considerably across datasets and
+classifiers. However, some results were robust: for example, the spectral rigidity and level number
+variance strongly differentiated controls from individuals prescribed duloxetine for osteopathic
+pain, and this differentiation persisted regardless of unfolding degree or trimming (Fig. 5-8, top-right panels).
+
+Despite the insensitivity to unfolding in this dataset, there was still nontrivial variability in
+accuracy across classifiers when classifying duloxetine vs controls, with the less flexible
+classifiers (logistic regression, SVM, and KNN with k=10, 20) generally performing best (Tables
+S1-2).
+
+For other datasets, the predictive utility of the unfolded features was highly dependent on the
+unfolding degree, and sometimes trim proportions. These differences are illustrated in detail for
+the Parkinson's data in Fig. S14. For example, with minimal trimming, subgroup overlap in spectral
+rigidity was minimized for unfolding degrees 5 and 9 (Fig. S14, first row), but with greater
+trimming, this was not the case (Fig. S14, second row). For the level number variance with minimal
+trimming, subgroup overlap appeared smaller for lower unfolding degrees (Fig. S14, third row),
+however, with a greater degree of trimming, separation was clearer for slightly larger degrees (Fig.
+S14, fourth row).
+
+Across datasets, the unfolding degree, compared to trimming, was a much more significant source of
+variance in the overall predictive utility of the unfolded features (e.g. note the minimal changes
+from trimming in Fig. S5 and S6, compared to the overall differences in prediction patterns from
+differing choices of unfolding degree in Fig. S3 and S4). However, neither the choice of unfolding
+degree nor amount of trimming had any easily summarizable effects on the predictive value of the
+various RMT metrics: the optimal trimming and unfolding degree was different for each dataset and
+comparison.
+
+\paragraph{Role of Classifier}. Across all datasets, unfoldings, and trimmings, the spectral rigidity was
+generally best able to separate between subgroups when either k was large (10 or 20) in KNN, or
+Logistic Regression or SVM was employed (Table S13). The pattern was similar for the level number
+variance, although the predictive utility of the level number variance was, overall, slightly more
+variable than the spectral rigidity (Table S4).
+
+\section{Discussion}
+In this study, certain features inspired by RMT and derived from the sorted eigenvalues of the full,
+whole-brain voxelwise correlation matrix were found to contain information with the potential to
+meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
+from previous studies [26], [33], [34], this suggests RMT may be a useful analytic and theoretical
+tool for understanding functional connectivity.
+
+Although the choice of unfolding and trimming often had nontrivial effects on the overall shape and
+values of the various spectral observables, and thus the predictive utility, in cases where there
+was clear predictive utility, there was some evidence that the unfolding and trimming procedures did
+not reverse the relations between subgroups (Fig. 1-3, S14). For example, regardless of the
+unfolding methodology, it was always the case that control subjects had spectral observables
+farthest from those of the GOE (Fig. 1-3) compared to those given duloxetine for osteopathic pain,
+whereas when comparing subject's with Parkinson's to controls, Parkinson's subjects consistently
+showed the greatest deviation from the GOE (Fig. S14).
+
+In particular, for some psychological phenomena examined in this study (specifically, the
+resting-state, voxelwise functional connectivity of those prescribed duloxetine for osteopathic pain
+versus the resting-state, voxelwise functional connectivity of controls), the spectral rigidity and
+level number variance contained information that was valuable in differentiating between subgroups.
+Across classifiers for this subgroup comparison, maximum mean LOOCV classification accuracies
+improved upon the 51\% accuracy of the naïve classifier by 11-26\% for the spectral rigidity and
+16-26\% for the level number variance (Tables S1-2).
+
+In the above case, these unfolded features allowed for significantly improved predictions over those
+using the raw (not unfolded) eigenvalues (Table S3). It is possible that some of these predictive
+improvements are because both the unfolding procedure and calculation of the spectral observables
+both smooth and transform all data to common scales. For example,  rescaling the raw eigenvalues to
+the interval [0, 1] also resulted in comparable prediction accuracies (Table S4).
+
+For comparison, for small datasets of rs- or task-fMRI data, binary classification using whole-brain
+features is challenging. For subject level classification, state of the art deep learning methods
+improve upon guessing by 17\% for autism [71] or 3-30\% for ADHD [72], 16\% for severe depression [73],
+and 23\% for obsessive compulsive disorder [74]. Manual feature engineering with more separable
+conditions (e.g. schizophrenia) can result in classification accuracies well above 90\% [75], and
+with very large data, sophisticated custom feature extraction methods can achieve near perfect
+accuracies at classifying task vs. rest [76]. However, for functional connectivity data and with
+classical machine learning algorithms (such as SVM) we in general only expect large prediction
+accuracies (e.g. greater than 80\%) when the group functional connectivities are already strongly
+separated (e.g. Cohen’s d > 1.0 [77]).
+
+Nevertheless, the features extracted in this study are somewhat unique in that they are small,
+spatially-insensitive summaries of the complete fMRI data. In particular, reducing the functional
+connectivity to just the largest 20 eigenvalues results in predictions comparable to using the
+entire set of eigenvalues (Tables S12, S18). That such a reduction can lead to accuracies comparable
+to other methods suggests that the system eigenvalues have utility beyond information reduction, and
+should perhaps be regarded themselves as signatures of functional connectivity patterns.
+
+\subsection{Confounds}
+While some of the RMT features allowed some classifiers to nearly perfectly separate conditions for
+the REFLECT dataset, this should almost certainly be regarded as an artifact of the differing scan
+parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
+correlation matrix is primarily determined by such differences in dimensions [78], as are the noise
+ratios, which include these dimensions in their definitions [29], [30]. Likewise, similar strong
+separation was observed in the PARK data when comparing the minimally pre-processed to fully
+pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps (spike
+removal, spatial smoothing, co-registration to larger T1 image [45]) which will also almost
+certainly alter the size of the largest eigenvalue.
+
+When there was predictive utility in the unfolded features,	 this utility was dependent on the
+unfolding procedure and choice of classifier. For example, the highest nontrivial LOOCV mean
+accuracy across all datasets was 0.79, corresponding to the duloxetine vs. controls subgroup
+comparison in the OSTEO dataset, whereas a poor choice of classifier in this same comparison could
+lead to prediction accuracies of 0.51, equivalent to the accuracy of a naive classifier (Table S2).
+
+There was variability in the predictive utility of the unfolding features, and this is likely
+because the unfolded features varied considerably based on the degree of the smoothing polynomial
+and extent of trimming. For example, the spread of the spectral rigidity and level number variance
+curves are shown for all subjects in the PARK data in Fig. S14. Clearly, both the curve shape,
+average magnitude, and extent of overlap is strongly dependent on the unfolding and trimming
+procedure.
+
+The unidimensional features, such as the largest eigenvalue or those based on the Marchenko-Pastur
+distribution, were most useful in predicting subgroup membership in trivial cases, such as in the
+REFLECT data, where the underlying scan dimensions differed between comparison groups, or in the
+PARK data, where the largest eigenvalue clearly distinguished the degree of preprocessing (Fig.
+S11-12). As these differences are due to methods, and not neuropsychological phenomena, they are
+perhaps of limited interest. To the extent that these features are scalar summaries of much
+higher-dimensional, 4D data, the limited utility of the scalar features should not be particularly
+surprising—it would be quite extraordinary if different neuropsychological states could be reliably
+separated based on such a reduction.
+
+However, that there are not consistently strong differences in these unidimensional features between
+subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
+noise-reduction procedures (such as [29], [30]), is unlikely to introduce bias into subsequent
+analyses. It perhaps also suggests these scalar features might be useful for quality control within
+a dataset: if there are differences between subgroups or individuals on one of the individual RMT
+features, then perhaps this is cause for more careful investigation.
+
+The unfolded features (spectral rigidity and level number variance) were more promising, however
+their sensitivity to the unfolding procedure means that great caution must be employed if they are
+to be used as features in future research. Ideally, the utility of these unfolded features (and all
+RMT features provided in this software release) need to be rigorously validated with wholly
+independent datasets addressing similar conditions. In addition, as simply using the raw, min-max
+normalized eigenvalues generally resulted in similar classification accuracies as the unfolded
+features, further study will be required to justify the use of these more complicated features.
+
+\subsection{Preprocessing}
+Across datasets, the features extracted from the minimally preprocessed images resulted in slightly
+lower prediction accuracies than the raw images for the unfolded features (rigidity, level
+variance), and slightly improved accuracies for some of the raw features (all eigenvalues, largest
+eigenvalue(s), noise ratio). However, these differences were quite small relative to the performance
+differences between datasets or between classifiers1.
+
+Similar patterns were observed in the PARK dataset. Fig. S5-7 show that more extensive preprocessing
+resulted in greater prediction accuracies for raw and largest eigenvalue(s), but decreased
+accuracies for the unfolded features. Given a particular level of preprocessing, however, it is
+clear from the spread of prediction accuracies that there was still substantial overall variance in
+prediction accuracy due to the choices of unfolding and classifier.
+
+\subsection{Comparison with Previous Studies}
+
+We generally find different results from previous studies [26], [33], [34] which have used the
+spectral rigidity or level number variance to differentiate between attention conditions. As [26]
+extracted rigidity curves from the correlations of electroencephalographic (EEG) signals, and [34]
+and [33] extracted spectral observables from the correlations of the mean signals of anatomical
+ROIs, these differences are to be expected. A single voxel signal amalgamates information from far
+fewer neurons than an EEG or mean ROI signal, and it is possible that any broad connections between
+RMT spectral observables and attention arise only at higher levels of organization, or after certain
+pre-processing steps (e.g. spatial smoothing) not performed in this study.
+
+We do find some modest evidence of greater overall correlation in high attention conditions: there
+was a slight tendency for the largest eigenvalue to be smaller when attention is low (Fig. S8).
+However, this is only suggestive, and future studies will be needed to determine how and under what
+conditions these RMT features reliably relate to neuropsychological phenomena.
+
+\subsection{Limitations}
+
+To make analyses across datasets at least minimally comparable, only limited preprocessing (motion
+and slicetime correction, skull-stripping) was performed in this study. In particular, in the
+absence of physiological noise regression, there must be caution in assuming any of the observed
+patterns are uniquely due to changes in functional connectivity. This is perhaps especially notable
+as the RMT features were most consistently capable of distinguishing between individuals taking
+duloxetine versus controls, or between controls and individuals with Parkinson's. It seems much more
+likely that physiological confounds could be involved in these comparisons in contrast to other
+comparisons (such as task-state versus resting-state, or in the differing levels of attention in the
+PSYCH dataset).
+
+The correlation matrices calculated for each run are computed from the full time-series data. No
+attempt was made to control for stimulus presentation times, under the assumption that the broad,
+whole-brain summarizing nature of the RMT features should be able to detect broad, overall
+differences between subgroups or conditions. However, differences were not consistently observed
+between conditions across all datasets. It is possible more careful selection of regions of the
+time-series in task conditions will be necessary in order for RMT features to reflect the
+differences between these conditions.
+
+While the extraction of eigenvalues from the correlations of all voxels has the advantage of not
+relying on a specific parcellation scheme, and does not impose anatomical or locality assumptions
+that ROI-based analyses do, it is also possible that this limited the utility of the RMT features.
+In particular, we found the spectral observables (NNSD, spectral rigidity, level number variance) to
+frequently fall outside the GOE/GDE bounds that have typically been observed in previous studies.
+Future studies should examine the sensitivity of these RMT features to the level of locality of
+analyses, or choice of size and location of ROIs.
+
 \section{Conclusion}

### 2021-09-09 paper/latex/preamble.tex
updatre
get biblio working

@@
-\documentclass{article} % This command is used to set the type of document you are working on such as an article, book, or presentation
+\documentclass[10pt, letterpaper]{article}
+\usepackage[utf8]{inputenc}
 
-\usepackage{geometry} % This package allows the editing of the page layout
-\usepackage{amsmath}  % This package allows the use of a large range of mathematical formula, commands, and symbols
-\usepackage{graphicx}  % This package allows the importing of images
+\usepackage{geometry}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{biblatex}
+\addbibresource{rmt.bib}
 

### 2021-09-09 paper/latex/rmt.tex
get biblio working
get one citation working

@@
 Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
-complex functional connectivity networks are ubiquitous [1]–[4]. This complexity suggests fMRI is a
-candidate to be studied using Random Matrix Theory (RMT), a set of mathematical tools originally
-developed some 50 years ago to solve complex problems in nuclear physics [5], [6]. In various
-physical systems where nuclei can have different possible states, precisely modeling the entirety of
-the interactions between nuclei is often intractable. For such large, complex systems, the
-eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
-of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
-spectrum statistically, complex systems can be investigated by comparison to the universal
-properties predicted by RMT.
+complex functional connectivity networks are ubiquitous
+(\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013})
+[1]–[4]. This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory
+(RMT), a set of mathematical tools originally developed some 50 years ago to solve complex problems
+in nuclear physics [5], [6]. In various physical systems where nuclei can have different possible
+states, precisely modeling the entirety of the interactions between nuclei is often intractable. For
+such large, complex systems, the eigenvalues (or spectra) efficiently summarize the totality of the
+interactions between components of the system, and RMT describes the expected behaviour of such
+eigenvalues. By analysing the spectrum statistically, complex systems can be investigated by
+comparison to the universal properties predicted by RMT.
 
@@
 
+\section{References}
+\printbibliography{}
+
 \end{document}

### 2021-09-12 paper/latex/rmt.tex
wip

--- 2021-09-09 paper/latex/rmt.tex
@@
 Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
-complex functional connectivity networks are ubiquitous [1]–[4]. This complexity suggests fMRI is a
-candidate to be studied using Random Matrix Theory (RMT), a set of mathematical tools originally
-developed some 50 years ago to solve complex problems in nuclear physics [5], [6]. In various
+complex functional connectivity networks are ubiquitous
+(\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}).
+This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory (RMT), a set
+of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
+physics (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). In various
 physical systems where nuclei can have different possible states, precisely modeling the entirety of
@@
 universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
-disordered mesoscopic systems [5]–[11], and at larger scales, RMT has been applied to atmospheric
-physics [12], stock cross-correlations [13], social networks [14], random networks [15],
-network-formation in liquids [16], [17] and amorphous clusters [18]–[21]. Within biological systems,
-RMT has also been used to successfully model aspects of  amino acid functional relationships [22],
-synchrony in epileptic seizures [23], and in protein-protein interactions both in different species
-[24] and breast cancer [25].
+disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has been applied to atmospheric
+physics (\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
+network-formation in liquids (\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009})  and amorphous clusters (\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}).
+Within biological systems,
+RMT has also been used to successfully model aspects of  amino acid functional relationships (\cite{bhadolaTargetingFunctionalMotifs2016}),
+synchrony in epileptic seizures (\cite{osorioPhasesynchronizationRandommatrixBased2011}), and in protein-protein interactions both in different species (\cite{agrawalQuantifyingRandomnessProtein2014}) and breast cancer (\cite{raiRandomnessPreservedPatterns2015}).
 \end{sloppypar}
@@
 demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) [26].  Somewhat more recently, RMT was used
-to evaluate the quality of whole brain features extracted from fMRI data [27], [28]. RMT has also
+resemble those of the Gaussian Orthogonal Ensemble (GOE) (\cite{sebaRandomMatrixAnalysis2003}).
+Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
+fMRI data (\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}). RMT has also
 been used in diffusion MRI to aid in the selection of the number of components to employ in
-principal-component reduction analysis and denoising [29]–[31].
+principal-component reduction analysis and denoising
+(\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}).
 
 Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states [32], between subjects with and without ADHD [33], and
-between pain and non-pain states [34]. Across these three studies, the spectra of resting or
-low-attention states exhibited properties closest to the GOE.
+differences between rest and task states (\cite{wangSpectralPropertiesTemporal2015}), between
+subjects with and without ADHD (\cite{wangRandomMatrixTheory2016}), and between pain and non-pain
+states (\cite{matharooSpontaneousBackpainAlters2020}). Across these three studies, the spectra of
+resting or low-attention states exhibited properties closest to the GOE.
 
@@
 
-However, since previous studies employing RMT for feature extraction [33], [34] have partially
-interpreted their findings with respect to attentional processes, we have elected to use datasets
-that involve attention in some way. This includes measures of attention or vigilance, or comparisons
-between a condition involving more attention and focus (e.g. a task state) or conditions which
-should involve considerably less attention or focus (e.g. resting-state).
+However, since previous studies employing RMT for feature extraction
+(\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}) have partially interpreted
+their findings with respect to attentional processes, we have elected to use datasets that involve
+attention in some way. This includes measures of attention or vigilance, or comparisons between a
+condition involving more attention and focus (e.g. a task state) or conditions which should involve
+considerably less attention or focus (e.g. resting-state).
 
@@
 Dataset 1 (referred to in Fig. and tables with the label "PSYCH") has detailed and extensive
-information available in  [37]. Subjects completed a battery of psychological measurements, and
-various resting-state fMRI (rs-fMRI) images were obtained. Only the whole-brain 7-Tesla rs-fMRI
-scans were used for the current paper, to make the results comparable to the other included
-datasets. For the purpose of keeping with the "attention" theme of the analyses in this paper, three
-pairings of subgroups were constructed via z-score splits of the various sum-scores of scale items
-involving concepts closely related to attention.
+information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
+psychological measurements, and various resting-state fMRI (rs-fMRI) images were obtained. Only the
+whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
+the other included datasets. For the purpose of keeping with the "attention" theme of the analyses
+in this paper, three pairings of subgroups were constructed via z-score splits of the various
+sum-scores of scale items involving concepts closely related to attention.
 
 \paragraph{Trait Attention}
-The PANAS-X [38] is a 9-point Likert-type self-report measure that queries the extent to which
-subjects agree particular affective terms have applied to them over a recent time period (e.g. a few
-weeks). An ad hoc (i.e. face-valid) sum-score was constructed from the scale items: "attentive",
-"alert", "interested", "concentrating", and reverse-scored items "sluggish", "tired", "sleepy",
-"drowsy". Subjects with a z-score greater / less than zero on this sum-score were placed in the high
-/ low "trait attention" category
+The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
+that queries the extent to which subjects agree particular affective terms have applied to them over
+a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
+the scale items: "attentive", "alert", "interested", "concentrating", and reverse-scored items
+"sluggish", "tired", "sleepy", "drowsy". Subjects with a z-score greater / less than zero on this
+sum-score were placed in the high / low "trait attention" category
 
 \paragraph{Task Attention}
-The Conjunctive Continuous Performance Task (CCPT [39]) is a behavioural measure of sustained
-attention where subjects must react only if they see a red square among other different coloured
-figures. For this paper, a sum score was constructed by subtracting the number of false positive and
-false negative responses from the number of success responses, and subjects with a z-score greater /
-less than zero on this scale were placed in high / low "task attention" category.
+The Conjunctive Continuous Performance Task (CCPT;
+\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
+where subjects must react only if they see a red square among other different coloured figures. For
+this paper, a sum score was constructed by subtracting the number of false positive and false
+negative responses from the number of success responses, and subjects with a z-score greater / less
+than zero on this scale were placed in high / low "task attention" category.
 
@@
 state, e.g. "my thoughts involved other people", "I thought about something positive" (translated
-from German) [40]. One question assesses alertness ("I was fully awake" / "war ich vollkommen wach")
-and was used to create high / low "vigilance" conditions in the same manner as the previous two splits.
+from German) \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
+alertness ("I was fully awake" / "war ich vollkommen wach") and was used to create high / low
+"vigilance" conditions in the same manner as the previous two splits.
 
 \subsubsection{Dataset 2}
-Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in [41]. Subjects
-included healthy controls ("nopain" condition) and individuals with knee osteoarthritis ("pain"
-condition) treated for 2 weeks with either placebo ("pain" condition) or duloxetine ("duloxetine"
-condition). An additional grouping was also created with both the placebo and duloxetine data
-("allpain" condition), allowing for six subgroup comparisons. All subjects were imaged with
-whole-brain rs-fMRI. Scan parameters are documented in Table I. The use of rs-fMRI and RMT for
-examining pain makes this dataset complementary to [34].
+Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in
+\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls ("nopain"
+condition) and individuals with knee osteoarthritis ("pain" condition) treated for 2 weeks with
+either placebo ("pain" condition) or duloxetine ("duloxetine" condition). An additional grouping was
+also created with both the placebo and duloxetine data ("allpain" condition), allowing for six
+subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
+documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
+complementary to \cite{matharooSpontaneousBackpainAlters2020}.
 
@@
 Dataset 3  (referred to here with the label "REFLECT", due to the "reflective” nature of the
-cognitive tasks) is described in full detail by DuPre, Luh and Spreng [42]. Both multi-echo rs- and
-task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks involved viewing
-various affective images (see [43] for details) and performing tasks related to autobiographical
-memory, prospection, or theory-of-mind reasoning. Continuing with the theme of broadly comparing
-high versus low attention throughout this paper, the different tasks were simply treated generically
-as "attending", in contrast to the non-attentive resting-state scans. This resulted in four task
-state scans and one resting-state scan per subject.
+cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
+multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
+involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
+and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
+Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
+different tasks were simply treated generically as "attending", in contrast to the non-attentive
+resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
 
@@
 the "summed" condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast [44]. In the "interleave" condition, echoes were interleaved
-at each voxel: that is, if we denote the voxel's ith signal value from echo j as , then the
-interleaved voxel of the new image has the time-series
-\([t_{1,1}, t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
+image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the "interleave"
+condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
+from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
+t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
 
-This creates a physically-meaningful 4D
-image that contains the full, original multi-echo fMRI time-series data without any alterations. It
-should be noted that the purpose of these combination methods is not to create a 4D volume that
-allows optimal detection of the purely task-linked BOLD signal, but to ensure all scan information
-was used with minimal additional analytic assumptions.
+This creates a physically-meaningful 4D image that contains the full, original multi-echo fMRI
+time-series data without any alterations. It should be noted that the purpose of these combination
+methods is not to create a 4D volume that allows optimal detection of the purely task-linked BOLD
+signal, but to ensure all scan information was used with minimal additional analytic assumptions.
 
@@
 Dataset 4 (referred to in Fig. and tables with the label "PARK" or "PARKINSONS") is documented and
-described in detail in [45]. Scan parameters are documented in Table I. Subjects with non-demented
-Parkinson’s Disease (PD) and healthy controls performed a number of repetitions of the Attention
-Network Task (ANT) [46]. Subgroup details are available in Table I. As the ANT task demands high
-attention, it is not as straightforward to create "high" or "low" attention groups. Instead, since
-attention presumably is high and varies little in this dataset, and subgroups instead differ by
-virtue of the presence or absence of PD, any results here serve as a useful comparison point for the
-results from studies that should clearly involve attentional differences. This dataset also included
-extensively preprocessed scans, which were included in analyses to examine the effects of
-preprocessing on the extracted RMT features.
+described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
+in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
+number of repetitions of the Attention Network Task (ANT)
+(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
+task demands high attention, it is not as straightforward to create "high" or "low" attention
+groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
+instead differ by virtue of the presence or absence of PD, any results here serve as a useful
+comparison point for the results from studies that should clearly involve attentional differences.
+This dataset also included extensively preprocessed scans, which were included in analyses to
+examine the effects of preprocessing on the extracted RMT features.
 
@@
 Dataset 5 (referred to in Fig. and tables with the label "LEARN") is documented and described in
-detail in [47]. In this study, both rs-fMRI and task fMRI scans were available for all subjects,
-allowing for within-subject comparisons. The complex task involved multiple phases, wherein subjects
-were exposed to named, black-and-white outlined satellite images related by family-resemblance of
-the components (see Fig. 1 of [47]). After this learning task, subjects were tested on their memory
-of the various learned features. Scans corresponding to both the initial learning and subsequent
-testing phase were included in the high-attention "task" subgroup for the purposes of the analysis
-of this paper. Resting state scans were included in the low-attention "rest" subgroup.
+detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
+were available for all subjects, allowing for within-subject comparisons. The complex task involved
+multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
+related by family-resemblance of the components (see Fig. 1 of
+\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
+memory of the various learned features. Scans corresponding to both the initial learning and
+subsequent testing phase were included in the high-attention "task" subgroup for the purposes of the
+analysis of this paper. Resting state scans were included in the low-attention "rest" subgroup.
 
@@
 However, these preprocessing steps are increasingly recognized to have complex effects on fMRI
-analyses [48]. For example, spatial smoothing has clear benefits when employing a general linear
-model [49]–[51], but is less motivated in functional network analyses with small ROIs, where the
-effect of smoothing can be considerable [50], [52], [53], with these effects interacting complexly
-with factors like field strength and image resolution [54]. Similarly, image registration is usually
-required to allow comparisons across (or even within) subjects, but the choice of registration
-strategy can have dramatic effects on functional network analyses [55].
+analyses (\cite{parkerBenefitSliceTiming2019}). For example, spatial smoothing has clear benefits
+when employing a general linear model
+(\cite{miklEffectsSpatialSmoothing2008,alakorkkoEffectsSpatialSmoothing2017,worsleyAnalysisFMRITimeSeries1995}),
+but is less motivated in functional network analyses with small ROIs, where the effect of smoothing
+can be considerable
+(\cite{alakorkkoEffectsSpatialSmoothing2017,goelmanMaximizingNegativeCorrelations2014,trianaEffectsSpatialSmoothing2020},
+with these effects interacting complexly with factors like field strength and image resolution
+(\cite{triantafyllouEffectSpatialSmoothing2006}). Similarly, image registration is usually required
+to allow comparisons across (or even within) subjects, but the choice of registration strategy can
+have dramatic effects on functional network analyses (\cite{magalhaesImpactNormalizationSegmentation2014}).
 

### 2021-09-13 paper/latex/rmt.tex
wip

@@
 appropriate here, there is insufficient data to perform these steps for all datasets (Table II).
-Thus, motion-correction was performed with FSL's MCFLIRT [56], slicetime correction with the default
-method in FSL, and brain extraction was performed with BET [57]. Due to differences in the size of
-the REFLECT data, the higher-resolution resting-state data was interpolated to have the same spatial
-dimensions as the task data.
+Thus, motion-correction was performed with FSL's MCFLIRT
+(\cite{jenkinsonImprovedOptimizationRobust2002}), slicetime correction with the default method in
+FSL, and brain extraction was performed with BET (\cite{smithFastRobustAutomated2002}). Due to
+differences in the size of the REFLECT data, the higher-resolution resting-state data was
+interpolated to have the same spatial dimensions as the task data.
 
@@
 not. In order to compare the observed eigenvalue distribution to some theoretical predictions, the
-eigenvalues must be unfolded [5], [6]. The details and motivation behind the unfolding procedure are
-well documented in [5]. Practically, the unfolding process can be viewed as a smoothing and
-rescaling procedure where the originally observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is
-smoothly mapped to the unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
-if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing function is not known, and
-so must be approximated [5], [6]. Typically, a polynomial is chosen [58].
+eigenvalues must be unfolded (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}).
+The details and motivation behind the unfolding procedure are well documented in
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process can be viewed as a
+smoothing and rescaling procedure where the originally observed spectrum \(\lambda_1 \le \dots \le
+\lambda_n\) is smoothly mapped to the unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n
+\approx 1\), if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing function is
+not known, and so must be approximated
+(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). Typically, a polynomial is
+chosen (\cite{abul-magdUnfoldingSpectrumChaotic2014}).
 
-A number of statistics (or "spectral observables" [6]) can then be computed from this unfolded
-spectrum. In the literature, the two most common points of comparison for these unfolded features
-are typically the features of the Poisson or Gaussian Diagonal Ensemble (GDE) and the GOE (see e.g.
-[12], [14], [15], [24]–[26], [32], [33]). The GOE (or GDE) consists of real, symmetric matrices
-where entries (or only diagonal entries) are normal random variables with zero mean and unit
-variance. When the empirically observed features are close to those of the GOE, this is usually
-interpreted as indicating a high degree of randomness and statistical independence in the data, and
-when the observed features are closer to those expected in the GDE or Poisson case, this is
-interpreted as indicating a high degree of coherence or dependence [12], [14], [15], [24]–[26],
-[32]–[34].
+A number of statistics (or "spectral observables"; \cite{mehtaRandomMatrices2004}) can then be
+computed from this unfolded spectrum. In the literature, the two most common points of comparison
+for these unfolded features are typically the features of the Poisson or Gaussian Diagonal Ensemble
+(GDE) and the GOE (see e.g.
+\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}). The
+GOE (or GDE) consists of real, symmetric matrices where entries (or only diagonal entries) are
+normal random variables with zero mean and unit variance. When the empirically observed features are
+close to those of the GOE, this is usually interpreted as indicating a high degree of randomness and
+statistical independence in the data, and when the observed features are closer to those expected in
+the GDE or Poisson case, this is interpreted as indicating a high degree of coherence or dependence
+(\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}).
 
@@
 analysed visually via a histogram of the distribution of spacings, however it is also possible to
-analyse it more formally by fitting the Brody distribution [59]:
+analyse it more formally by fitting the Brody distribution (\cite{brodyStatisticalMeasureRepulsion1973}):
 
@@
 spectrum, the long-range properties relating eigenvalues distant on the spectrum can be
-characterized by both the spectral rigidity and level number variance [5], [6]. The spectral
-rigidity, \(\Delta_3(L)\), is calculated for any real value \(L\) as:
+characterized by both the spectral rigidity and level number variance
+(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The spectral rigidity,
+\(\Delta_3(L)\), is calculated for any real value \(L\) as:
 
@@
 \(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
-line to \(\eta(\lambda)\) on \([c, c+L]\) [5]. The level number variance, \(\Sigma^2(L)\), is
-closely related to the spectral rigidity [6], and is calculated as:
+line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The level
+number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
+(\cite{mehtaRandomMatrices2004}), and is calculated as:
 
@@
 where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
-and \(\langle \cdot \rangle_c\) are as above [5]. Both the spectral rigidity and level number
-variance are also computed and investigated in this study.
+and \(\langle \cdot \rangle_c\) are as above (\cite{guhrRandommatrixTheoriesQuantum1998a}). Both the
+spectral rigidity and level number variance are also computed and investigated in this study.
 
@@
 The distribution of eigenvalues of large rectangular matrices with finite-variance, independent and
-identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP) distribution
-[60], [61]:
+identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP)
+distribution (\cite{marcenkoDISTRIBUTIONEIGENVALUESSETS1967,yaoLargeSampleCovariance2015}):
 
@@
 random correlation matrices will fall within \([\lambda_{-}, \lambda_{+}]\), eigenvalues observed to
-fall within this region are typically considered to be the result of noise [29]–[31], [62]. We
-examine as a potential feature the proportion of eigenvalues that fall within this range, and call
-this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
-adjustments of [63] which takes into account the larger degree of correlation in empirical data.
+fall within this region are typically considered to be the result of noise
+(\cite{veraartDenoisingDiffusionMRI2016,veraartDiffusionMRINoise2016,ulfarssonDimensionEstimationNoisy2008,macmahonCommunityDetectionCorrelation2015}).
+We examine as a potential feature the proportion of eigenvalues that fall within this range, and
+call this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
+adjustments of \cite{almogUncoveringFunctionalSignature2019} which takes into account the larger
+degree of correlation in empirical data.
 
@@
 publicly available as part of this study, as well as via specific code for this paper (which is
-documented and available online [65]).
+documented and available online (\cite{bergerderekm.DMBergerRandommatrixfmriV02020}).
 
 As both the spectral rigidity and level number variance can be highly sensitive to the choice of
-smoothing procedure used in the unfolding method [58], [66]–[69], all unfolded features in this
-study were calculated using polynomials of degree 5, 7, 9, 11, and 13 as the smoothing technique. As
-much of the variance in the unfolded spectral statistics can be due to the largest eigenvalues
-acting as anchor points during smoothing [66], [69], all unfolded features were also calculated
-twice (in addition to each polynomial degree) to examine the effects of trimming on spectral
-observables: once with only the largest eigenvalue omitted, and once with the largest 20 eigenvalues
-omitted.
+smoothing procedure used in the unfolding method
+(\cite{abul-magdUnfoldingSpectrumChaotic2014,moralesImprovedUnfoldingDetrending2011,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012}),
+all unfolded features in this study were calculated using polynomials of degree 5, 7, 9, 11, and 13
+as the smoothing technique. As much of the variance in the unfolded spectral statistics can be due
+to the largest eigenvalues acting as anchor points during smoothing
+(\cite{moralesImprovedUnfoldingDetrending2011,abueleninEffectUnfoldingSpectral2012}), all unfolded
+features were also calculated twice (in addition to each polynomial degree) to examine the effects
+of trimming on spectral observables: once with only the largest eigenvalue omitted, and once with
+the largest 20 eigenvalues omitted.
 
@@
 A wide variety of standard, out-of-the-box classifiers available in the Python scikit-learn library
-[70] were then employed to gauge the predictive value of the RMT features. To allow for a wide range
-of flexibilities, classifiers employing k-nearest neighbours (KNN), with k values in \(\{1, 2, 3,
-\dots, 10, 20\}\), logistic regression (LR), linear discriminant analysis (LDA), and a
-support-vector machine (SVM) were employed. Default parameter values were used when possible (see
-[65] for precise function calls used).
+(\cite{pedregosaScikitlearnMachineLearning2011}) were then employed to gauge the predictive value of
+the RMT features. To allow for a wide range of flexibilities, classifiers employing k-nearest
+neighbours (KNN), with k values in \(\{1, 2, 3, \dots, 10, 20\}\), logistic regression (LR), linear
+discriminant analysis (LDA), and a support-vector machine (SVM) were employed. Default parameter
+values were used when possible (see \cite{bergerderekm.DMBergerRandommatrixfmriV02020} for precise
+function calls used).
 

### 2021-09-13 paper/latex/rmt.tex
formatting

@@
 
-\begin{center}
-\texttt{\textbf{INSERT TABLE 1 HERE}}
-\end{center}
+\begin{table}[h!]
+\small
+\centering
+\begin{tabular}{ c c c c c c c }
+\hline
+\textbf{ID}  & \textbf{FOV (mm)}  & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{Time (s)} \\
+\hline
+PSYCH   &  192              & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0 & 300 & 15 \\
+OSTEO   &  -                & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5 & 300 & 12.5 \\
+REFLECT &  210              & 72 x 72 x 46   & 3.0 x 3.0 x 3.0 & 2.5 & 204 & 10.2 \\
+REFLECT &  240              & 64 x 64 x 33   & 3.8 x 3.8 x 3.8 & 2.0 & 260 & 8.7 \\
+PARK    &  240 x 240 x 129  & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4 & 149 & 5.96 \\
+LEARN   &  -                & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0 & 196 & 6.5 \\
+\hline
+\end{tabular}
+\caption{
+    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P slices with dimensions M x N.
+}
+\label{table:1}
+\end{table}
 
@@
 
+\begin{table}[h!]
+\small
+\centering
+\begin{tabular}{ l l c c c }
+\hline
+\textbf{ID}  & \textbf{Subgroup / Condition}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
+\hline
+PSYCH   &  high\_vigilance    & 12 / 11 & resting-state                    & 2 \\
+        &  low\_vigilance     & 10 / 11 & resting-state                    & 2 \\
+        &  trait\_attend      & 12 / 15 & resting-state                    & 2 \\
+        &  trait\_non-attend  & 10 / 7  & resting-state                    & 2 \\
+        &  task\_attend       & 13 / 15 & resting-state                    & 2 \\
+        &  task\_nonattend    &  9 / 7  & resting-state                    & 2 \\
+\hline
+LEARN   &  task              & 24      & learn image sketches & 16 \\
+        &  rest              & 24      & resting-state                    & 2 \\
+\hline
+REFLECT &  control (no pain) & 20      & resting-state                    & 1 \\
+        &  duloxetine        & 19      & resting-state                    & 1 \\
+        &  placebo           & 37      & resting-state                    & 1 \\
+        &  allpain (duloxetine + placebo) & 56         & resting-state    & 1 \\
+\hline
+PARK    &  Parkinson's       & 17      & ANT                              & 6 \\
+        &  control           & 17      & ANT                              & 6 \\
+\hline
+ECHO    &  summed/interleave & 31      & resting-state                    & 1 \\
+        &  summed/interleave & 31      & imagery/reflection               & 4 \\
+\hline
+\end{tabular}
+\caption{
+    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P slices with dimensions M x N.
+}
+\label{table:2}
+\end{table}
+
 \subsubsection{Dataset 1}
-Dataset 1 (referred to in Fig. and tables with the label "PSYCH") has detailed and extensive
+Dataset 1 (referred to in Fig. and tables with the label ``PSYCH'') has detailed and extensive
 information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
@@
 whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
-the other included datasets. For the purpose of keeping with the "attention" theme of the analyses
+the other included datasets. For the purpose of keeping with the ``attention'' theme of the analyses
 in this paper, three pairings of subgroups were constructed via z-score splits of the various
@@
 a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
-the scale items: "attentive", "alert", "interested", "concentrating", and reverse-scored items
-"sluggish", "tired", "sleepy", "drowsy". Subjects with a z-score greater / less than zero on this
-sum-score were placed in the high / low "trait attention" category
+the scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
+``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects with a z-score greater / less than zero on this
+sum-score were placed in the high / low ``trait attention'' category
 
@@
 negative responses from the number of success responses, and subjects with a z-score greater / less
-than zero on this scale were placed in high / low "task attention" category.
+than zero on this scale were placed in high / low ``task attention'' category.
 
@@
 The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
-state, e.g. "my thoughts involved other people", "I thought about something positive" (translated
-from German) \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
-alertness ("I was fully awake" / "war ich vollkommen wach") and was used to create high / low
-"vigilance" conditions in the same manner as the previous two splits.
+state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
+from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}). One question assesses
+alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
+``vigilance'' conditions in the same manner as the previous two splits.
 
 \subsubsection{Dataset 2}
-Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in
-\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls ("nopain"
-condition) and individuals with knee osteoarthritis ("pain" condition) treated for 2 weeks with
-either placebo ("pain" condition) or duloxetine ("duloxetine" condition). An additional grouping was
-also created with both the placebo and duloxetine data ("allpain" condition), allowing for six
+Dataset 2 (referred to in Fig. and tables with the label ``OSTEO'') is detailed in
+\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
+condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
+either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
+also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
 subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
@@
 \subsubsection{Dataset 3}
-Dataset 3  (referred to here with the label "REFLECT", due to the "reflective” nature of the
+Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
 cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
@@
 Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
-different tasks were simply treated generically as "attending", in contrast to the non-attentive
+different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
 resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
@@
 In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
-the "summed" condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the "interleave"
+the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
+image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the ``interleave''
 condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
@@
 \subsubsection{Dataset 4}
-Dataset 4 (referred to in Fig. and tables with the label "PARK" or "PARKINSONS") is documented and
+Dataset 4 (referred to in Fig. and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
 described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
@@
 (\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
-task demands high attention, it is not as straightforward to create "high" or "low" attention
+task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
 groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
@@
 \subsubsection{Dataset 5}
-Dataset 5 (referred to in Fig. and tables with the label "LEARN") is documented and described in
+Dataset 5 (referred to in Fig. and tables with the label ``LEARN'') is documented and described in
 detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
@@
 memory of the various learned features. Scans corresponding to both the initial learning and
-subsequent testing phase were included in the high-attention "task" subgroup for the purposes of the
-analysis of this paper. Resting state scans were included in the low-attention "rest" subgroup.
+subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of the
+analysis of this paper. Resting state scans were included in the low-attention ``rest'' subgroup.
 
@@
 
-A number of statistics (or "spectral observables"; \cite{mehtaRandomMatrices2004}) can then be
+A number of statistics (or ``spectral observables''; \cite{mehtaRandomMatrices2004}) can then be
 computed from this unfolded spectrum. In the literature, the two most common points of comparison
@@
 
-In addition to the unfolded features suggested by RMT, there are some potential "raw" features that
+In addition to the unfolded features suggested by RMT, there are some potential ``raw'' features that
 do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
@@
 had the most consistent predictive value for distinguishing individuals with Parkinson's from
-controls, or distinguishing those with osteopathic pain taking duloxetine from controls ("nopain" in
+controls, or distinguishing those with osteopathic pain taking duloxetine from controls (``nopain'' in
 Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
@@
 The raw eigenvalues generally had the most predictive utility when employing either an SVM
-classifier or KNN classifier with k > 3 (Table S11, first row). When the raw eigenvalues were first
-min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
-predicting subgroup memberships (Table S12), however without normalization, KNN with k = 8 or k = 9
-generally best predicted subgroup membership across datasets (Table S11).
+classifier or KNN classifier with \(k > 3\) (Table S11, first row). When the raw eigenvalues were
+first min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
+predicting subgroup memberships (Table S12), however without normalization, KNN with \(k = 8\) or
+\(k = 9\) generally best predicted subgroup membership across datasets (Table S11).
 
@@
 meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
-from previous studies [26], [33], [34], this suggests RMT may be a useful analytic and theoretical
-tool for understanding functional connectivity.
+from previous studies
+(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020})
+this suggests RMT may be a useful analytic and theoretical tool for understanding functional
+connectivity.
 
@@
 features is challenging. For subject level classification, state of the art deep learning methods
-improve upon guessing by 17\% for autism [71] or 3-30\% for ADHD [72], 16\% for severe depression [73],
-and 23\% for obsessive compulsive disorder [74]. Manual feature engineering with more separable
-conditions (e.g. schizophrenia) can result in classification accuracies well above 90\% [75], and
-with very large data, sophisticated custom feature extraction methods can achieve near perfect
-accuracies at classifying task vs. rest [76]. However, for functional connectivity data and with
-classical machine learning algorithms (such as SVM) we in general only expect large prediction
-accuracies (e.g. greater than 80\%) when the group functional connectivities are already strongly
-separated (e.g. Cohen’s d > 1.0 [77]).
+improve upon guessing by 17\% for autism (\cite{bengs4DSpatioTemporalDeep2020}) or 3-30\% for ADHD
+(\cite{riazDeepFMRIEndtoendDeep2020}), 16\% for severe depression
+(\cite{ramasubbuAccuracyAutomatedClassification2016}), and 23\% for obsessive compulsive disorder
+(\cite{takagiNeuralMarkerObsessiveCompulsive2017}). Manual feature engineering with more separable
+conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
+(\cite{duHighClassificationAccuracy2012}), and with very large data, sophisticated custom feature
+extraction methods can achieve near perfect accuracies at classifying task vs. rest
+(\cite{zhangCharacterizingDifferentiatingTaskbased2016}). However, for functional connectivity data
+and with classical machine learning algorithms (such as SVM) we in general only expect large
+prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
+strongly separated (e.g. Cohen’s d > 1.0; \cite{dansereauStatisticalPowerPrediction2017}).
 
@@
 parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
-correlation matrix is primarily determined by such differences in dimensions [78], as are the noise
-ratios, which include these dimensions in their definitions [29], [30]. Likewise, similar strong
-separation was observed in the PARK data when comparing the minimally pre-processed to fully
-pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps (spike
-removal, spatial smoothing, co-registration to larger T1 image [45]) which will also almost
-certainly alter the size of the largest eigenvalue.
+correlation matrix is primarily determined by such differences in dimensions
+(\cite{yinLimitLargestEigenvalue1988}), as are the noise ratios, which include these dimensions in
+their definitions (\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}). Likewise,
+similar strong separation was observed in the PARK data when comparing the minimally pre-processed
+to fully pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps
+(spike removal, spatial smoothing, co-registration to larger T1 image;
+\cite{madhyasthaDynamicConnectivityRest2015}) which will also almost certainly alter the size of the
+largest eigenvalue.
 
@@
 subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
-noise-reduction procedures (such as [29], [30]), is unlikely to introduce bias into subsequent
-analyses. It perhaps also suggests these scalar features might be useful for quality control within
-a dataset: if there are differences between subgroups or individuals on one of the individual RMT
-features, then perhaps this is cause for more careful investigation.
+noise-reduction procedures (such as
+\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}), is unlikely to introduce bias
+into subsequent analyses. It perhaps also suggests these scalar features might be useful for quality
+control within a dataset: if there are differences between subgroups or individuals on one of the
+individual RMT features, then perhaps this is cause for more careful investigation.
 
@@
 
-We generally find different results from previous studies [26], [33], [34] which have used the
-spectral rigidity or level number variance to differentiate between attention conditions. As [26]
-extracted rigidity curves from the correlations of electroencephalographic (EEG) signals, and [34]
-and [33] extracted spectral observables from the correlations of the mean signals of anatomical
-ROIs, these differences are to be expected. A single voxel signal amalgamates information from far
-fewer neurons than an EEG or mean ROI signal, and it is possible that any broad connections between
-RMT spectral observables and attention arise only at higher levels of organization, or after certain
-pre-processing steps (e.g. spatial smoothing) not performed in this study.
+We generally find different results from previous studies
+(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
+which have used the spectral rigidity or level number variance to differentiate between attention
+conditions. As \cite{sebaRandomMatrixAnalysis2003} extracted rigidity curves from the correlations
+of electroencephalographic (EEG) signals, and \cite{wangRandomMatrixTheory2016} and
+\cite{matharooSpontaneousBackpainAlters2020} extracted spectral observables from the correlations of
+the mean signals of anatomical ROIs, these differences are to be expected. A single voxel signal
+amalgamates information from far fewer neurons than an EEG or mean ROI signal, and it is possible
+that any broad connections between RMT spectral observables and attention arise only at higher
+levels of organization, or after certain pre-processing steps (e.g. spatial smoothing) not performed
+in this study.
 

### 2021-09-13 paper/latex/NetworkNeuroscience/rmt.tex
update

--- 2021-09-13 paper/latex/rmt.tex
@@
     predictive—rather than explanatory—value, (3) investigate the effect of varying RMT analysis
-    methods on the robustness of study findings, and (4) make general-purpose code publicly
-    available for users to extract these features from a wide variety of data. We find preliminary
-    evidence suggesting that RMT-inspired features may have unique potential in analyses of fMRI
-    functional connectivity.
+    methods on the robustness of study findings, and (4) make code publicly available to extract
+    these features from a wide variety of data. We find preliminary evidence suggesting that
+    RMT-inspired features may have unique potential in analyses of fMRI functional connectivity.
 
-    \subparagraph{Index Terms:} feature extraction, functional connectivity, functional magnetic
-    resonance imaging, random matrix theory, voxelwise connectivity
+    \subparagraph{Keywords:} feature extraction, functional connectivity, functional magnetic
+    resonance imaging, random matrix theory, voxelwise connectivity, predictive analysis
 

### 2021-09-15 paper/latex/NetworkNeuroscience/rmt.tex
update

@@
-\input{preamble.tex}
+\documentclass[NETN,manuscript]{stjour-new}
+\articletype{Research}
+\usepackage{graphicx}
+\graphicspath{ {figs/} }
+% \input{preamble.tex}
+% \documentclass{article}
+% \usepackage[utf8]{inputenc}
+% \usepackage{geometry}
+% \usepackage{amsmath}
+% \usepackage{graphicx}
+% \usepackage[style=apa]{biblatex}
+% \usepackage{natbib}
+% \usepackage[font=footnotesize,labelfont=bf]{caption}
+% \addbibresource{rmt.bib}
+% \bibliographystyle{plain}
+% \bibliography{rmt}
+\begin{document}
 
 \title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
-\author{Derek M. Berger \and Gurpreet S. Matharoo \and Jacob Levman}
+\author[Derek M. Berger, Gurpreet S. Matharoo, Jacob Levman]{
+    Derek M. Berger\affil{1},
+    Gurpreet S. Matharoo\affil{2}
+    \and Jacob Levman\affil{3}
+}
+\affiliation{1}{Computer Science, St. Francis Xavier University, Antigonish, Canada}
+\affiliation{2}{Physics, St. Francis Xavier University, Antigonish, Canada}
+\affiliation{3}{Computer Science, St. Francis Xavier University, Antigonish, Canada}
+
 \date{\today}
 
-\begin{document}
-\maketitle
+% \maketitle
+
+% \keywords{random matrix theory, feature extraction, functional connectivity,
+% functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
 
@@
     application of RMT-based features for the analysis of functional MRI (fMRI) across diverse
+
     datasets. As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain
@@
     RMT-inspired features may have unique potential in analyses of fMRI functional connectivity.
-
-    \subparagraph{Keywords:} feature extraction, functional connectivity, functional magnetic
-    resonance imaging, random matrix theory, voxelwise connectivity, predictive analysis
-
-    \subparagraph{Authors}
-    Derek Berger is with the department of Computer Science, St. Francis Xavier University, Antigonish
-    NS B2G 2W5, Canada (e-mail: dberger@stfx.ca).
-
-    Gurpreet Matharoo is with ACENET, and the Department of Physics, St. Francis Xavier University,
-    Antigonish NS B2G 2W5, Canada (e-mail: gmatharo@stfx.ca).
-
-    Jacob Levman is with the Department of Computer Science, St. Francis Xavier University,
-    Antigonish NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).
 \end{abstract}
@@
 complex functional connectivity networks are ubiquitous
-(\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}).
+\citep{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
 This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory (RMT), a set
 of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
-physics (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). In various
+physics \citep{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. In various
 physical systems where nuclei can have different possible states, precisely modeling the entirety of
@@
 disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has been applied to atmospheric
-physics (\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
-network-formation in liquids (\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009})  and amorphous clusters (\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}).
+physics \citep{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
+network-formation in liquids \citep{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}  and amorphous clusters \citep{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
 Within biological systems,
-RMT has also been used to successfully model aspects of  amino acid functional relationships (\cite{bhadolaTargetingFunctionalMotifs2016}),
-synchrony in epileptic seizures (\cite{osorioPhasesynchronizationRandommatrixBased2011}), and in protein-protein interactions both in different species (\cite{agrawalQuantifyingRandomnessProtein2014}) and breast cancer (\cite{raiRandomnessPreservedPatterns2015}).
+RMT has also been used to successfully model aspects of  amino acid functional relationships \citep{bhadolaTargetingFunctionalMotifs2016},
+synchrony in epileptic seizures \citep{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein interactions both in different species \citep{agrawalQuantifyingRandomnessProtein2014} and breast cancer \citep{raiRandomnessPreservedPatterns2015}.
 \end{sloppypar}
@@
 demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) (\cite{sebaRandomMatrixAnalysis2003}).
+resemble those of the Gaussian Orthogonal Ensemble (GOE) \citep{sebaRandomMatrixAnalysis2003}.
 Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
-fMRI data (\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}). RMT has also
+fMRI data \citep{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
 been used in diffusion MRI to aid in the selection of the number of components to employ in
 principal-component reduction analysis and denoising
-(\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}).
+\citep{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
 
 Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states (\cite{wangSpectralPropertiesTemporal2015}), between
-subjects with and without ADHD (\cite{wangRandomMatrixTheory2016}), and between pain and non-pain
-states (\cite{matharooSpontaneousBackpainAlters2020}). Across these three studies, the spectra of
+differences between rest and task states \citep{wangSpectralPropertiesTemporal2015}, between
+subjects with and without ADHD \citep{wangRandomMatrixTheory2016}, and between pain and non-pain
+states \citep{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
 resting or low-attention states exhibited properties closest to the GOE.
@@
 However, since previous studies employing RMT for feature extraction
-(\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}) have partially interpreted
+\citep{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020} have partially interpreted
 their findings with respect to attentional processes, we have elected to use datasets that involve
@@
 
-Dataset scan parameters and acquisition details are summarized in Table I, and sample and subgroup
-sizes are summarized in Table II. For all datasets, we use the raw (unpreprocessed) scans, and
-perform our own preprocessing (detailed later). Given the complexity of these datasets, we highlight
-below only those aspects relevant to the current paper.
+Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
+and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
+(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
+these datasets, we highlight below only those aspects relevant to the current paper.
 
@@
 \subsubsection{Dataset 1}
-Dataset 1 (referred to in Fig. and tables with the label ``PSYCH'') has detailed and extensive
+Dataset 1 (referred to in figures and tables with the label ``PSYCH'') has detailed and extensive
 information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
@@
 \paragraph{Trait Attention}
-The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
+The PANAS-X \citep{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure
 that queries the extent to which subjects agree particular affective terms have applied to them over
@@
 The Conjunctive Continuous Performance Task (CCPT;
-\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
+\cite{shalevConjunctiveContinuousPerformance2011} is a behavioural measure of sustained attention
 where subjects must react only if they see a red square among other different coloured figures. For
@@
 state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
-from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}). One question assesses
+from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
 alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
@@
 subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
-documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
+documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
 complementary to \cite{matharooSpontaneousBackpainAlters2020}.
@@
 cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
-multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
-involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
+multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
+involved viewing various affective images \citep[see][for details]{sprengPatternsBrainActivity2010}
 and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
@@
 the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the ``interleave''
+image with overall increased contrast \citep{kunduMultiechoFMRIReview2017}. In the ``interleave''
 condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
@@
 described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
-in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
-number of repetitions of the Attention Network Task (ANT)
-(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
+in Table \ref{table:1}. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
+number of repetitions of the Attention Network Task
+\citep[ANT;][]{fanActivationAttentionalNetworks2005}. Subgroup details are available in Table \ref{table:1}. As the ANT
 task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
@@
 multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
-related by family-resemblance of the components (see Fig. 1 of
-\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
+related by family-resemblance of the components \citep[see figure 1
+of][]{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
 memory of the various learned features. Scans corresponding to both the initial learning and
-subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of the
-analysis of this paper. Resting state scans were included in the low-attention ``rest'' subgroup.
+subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of
+the analysis of this paper. Resting state scans were included in the low-attention ``rest''
+subgroup.
 
@@
 However, these preprocessing steps are increasingly recognized to have complex effects on fMRI
-analyses (\cite{parkerBenefitSliceTiming2019}). For example, spatial smoothing has clear benefits
+analyses \citep{parkerBenefitSliceTiming2019}. For example, spatial smoothing has clear benefits
 when employing a general linear model
-(\cite{miklEffectsSpatialSmoothing2008,alakorkkoEffectsSpatialSmoothing2017,worsleyAnalysisFMRITimeSeries1995}),
+\citep{miklEffectsSpatialSmoothing2008,alakorkkoEffectsSpatialSmoothing2017,worsleyAnalysisFMRITimeSeries1995},
 but is less motivated in functional network analyses with small ROIs, where the effect of smoothing
 can be considerable
-(\cite{alakorkkoEffectsSpatialSmoothing2017,goelmanMaximizingNegativeCorrelations2014,trianaEffectsSpatialSmoothing2020},
+\citep{alakorkkoEffectsSpatialSmoothing2017,goelmanMaximizingNegativeCorrelations2014,trianaEffectsSpatialSmoothing2020},
 with these effects interacting complexly with factors like field strength and image resolution
-(\cite{triantafyllouEffectSpatialSmoothing2006}). Similarly, image registration is usually required
+\citep{triantafyllouEffectSpatialSmoothing2006}. Similarly, image registration is usually required
 to allow comparisons across (or even within) subjects, but the choice of registration strategy can
-have dramatic effects on functional network analyses (\cite{magalhaesImpactNormalizationSegmentation2014}).
+have dramatic effects on functional network analyses \citep{magalhaesImpactNormalizationSegmentation2014}.
 
@@
 locations. Whereas physiological noise regression and bias field correction would perhaps be
-appropriate here, there is insufficient data to perform these steps for all datasets (Table II).
+appropriate here, there is insufficient data to perform these steps for all datasets (Table \ref{table:2}).
 Thus, motion-correction was performed with FSL's MCFLIRT
-(\cite{jenkinsonImprovedOptimizationRobust2002}), slicetime correction with the default method in
-FSL, and brain extraction was performed with BET (\cite{smithFastRobustAutomated2002}). Due to
+\citep{jenkinsonImprovedOptimizationRobust2002}, slicetime correction with the default method in
+FSL, and brain extraction was performed with BET \citep{smithFastRobustAutomated2002}. Due to
 differences in the size of the REFLECT data, the higher-resolution resting-state data was
@@
 not. In order to compare the observed eigenvalue distribution to some theoretical predictions, the
-eigenvalues must be unfolded (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}).
+eigenvalues must be unfolded \citep{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}.
 The details and motivation behind the unfolding procedure are well documented in
@@
 not known, and so must be approximated
-(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). Typically, a polynomial is
-chosen (\cite{abul-magdUnfoldingSpectrumChaotic2014}).
+\citep{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically, a polynomial is
+chosen \citep{abul-magdUnfoldingSpectrumChaotic2014}.
 
-A number of statistics (or ``spectral observables''; \cite{mehtaRandomMatrices2004}) can then be
+A number of statistics or ``spectral observables'' \citep{mehtaRandomMatrices2004} can then be
 computed from this unfolded spectrum. In the literature, the two most common points of comparison
 for these unfolded features are typically the features of the Poisson or Gaussian Diagonal Ensemble
-(GDE) and the GOE (see e.g.
-\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}). The
+(GDE) and the GOE
+\cite[see e.g.][]{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}. The
 GOE (or GDE) consists of real, symmetric matrices where entries (or only diagonal entries) are
@@
 the GDE or Poisson case, this is interpreted as indicating a high degree of coherence or dependence
-(\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}).
+\citep{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}.
 
@@
 analysed visually via a histogram of the distribution of spacings, however it is also possible to
-analyse it more formally by fitting the Brody distribution (\cite{brodyStatisticalMeasureRepulsion1973}):
+analyse it more formally by fitting the Brody distribution \citep{brodyStatisticalMeasureRepulsion1973}:
 
@@
 characterized by both the spectral rigidity and level number variance
-(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The spectral rigidity,
+\citep{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The spectral rigidity,
 \(\Delta_3(L)\), is calculated for any real value \(L\) as:
@@
 \(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
-line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The level
+line to \(\eta(\lambda)\) on \([c, c+L]\) \citep{guhrRandommatrixTheoriesQuantum1998a}. The level
 number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
-(\cite{mehtaRandomMatrices2004}), and is calculated as:
+\citep{mehtaRandomMatrices2004}, and is calculated as:
 
@@
 where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
-and \(\langle \cdot \rangle_c\) are as above (\cite{guhrRandommatrixTheoriesQuantum1998a}). Both the
+and \(\langle \cdot \rangle_c\) are as above \citep{guhrRandommatrixTheoriesQuantum1998a}. Both the
 spectral rigidity and level number variance are also computed and investigated in this study.
@@
 identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP)
-distribution (\cite{marcenkoDISTRIBUTIONEIGENVALUESSETS1967,yaoLargeSampleCovariance2015}):
+distribution \citep{marcenkoDISTRIBUTIONEIGENVALUESSETS1967,yaoLargeSampleCovariance2015}:
 
@@
 fall within this region are typically considered to be the result of noise
-(\cite{veraartDenoisingDiffusionMRI2016,veraartDiffusionMRINoise2016,ulfarssonDimensionEstimationNoisy2008,macmahonCommunityDetectionCorrelation2015}).
+\citep{veraartDenoisingDiffusionMRI2016,veraartDiffusionMRINoise2016,ulfarssonDimensionEstimationNoisy2008,macmahonCommunityDetectionCorrelation2015}.
 We examine as a potential feature the proportion of eigenvalues that fall within this range, and
@@
 \times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
-value of \(t\) for a scan can be inferred from the number of volumes in Table I. The various RMT
+value of \(t\) for a scan can be inferred from the number of volumes in Table \ref{table:1}. The various RMT
 features mentioned above were then extracted via the empyricalRMT library [64] developed and made
 publicly available as part of this study, as well as via specific code for this paper (which is
-documented and available online (\cite{bergerderekm.DMBergerRandommatrixfmriV02020}).
+documented and available online \citep{bergerderekm.DMBergerRandommatrixfmriV02020}.
 
@@
 smoothing procedure used in the unfolding method
-(\cite{abul-magdUnfoldingSpectrumChaotic2014,moralesImprovedUnfoldingDetrending2011,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012}),
+\citep{abul-magdUnfoldingSpectrumChaotic2014,moralesImprovedUnfoldingDetrending2011,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012},
 all unfolded features in this study were calculated using polynomials of degree 5, 7, 9, 11, and 13
@@
 to the largest eigenvalues acting as anchor points during smoothing
-(\cite{moralesImprovedUnfoldingDetrending2011,abueleninEffectUnfoldingSpectral2012}), all unfolded
+\citep{moralesImprovedUnfoldingDetrending2011,abueleninEffectUnfoldingSpectral2012}, all unfolded
 features were also calculated twice (in addition to each polynomial degree) to examine the effects
@@
 A wide variety of standard, out-of-the-box classifiers available in the Python scikit-learn library
-(\cite{pedregosaScikitlearnMachineLearning2011}) were then employed to gauge the predictive value of
+\citep{pedregosaScikitlearnMachineLearning2011} were then employed to gauge the predictive value of
 the RMT features. To allow for a wide range of flexibilities, classifiers employing k-nearest
@@
 from previous studies
-(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020})
+\citep{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
 this suggests RMT may be a useful analytic and theoretical tool for understanding functional
@@
 showed the greatest deviation from the GOE (Fig. S14).
+
+\begin{figure}[ht]
+\centerline{\includegraphics[width=\textwidth]{figure1}} \caption{
+    (Colour online) Spectral rigidity for controls (nopain) vs. those with osteopathic pain taking
+    duloxetine, trimming largest eigenvalue only. Solid lines indicate group mean rigidity, and
+    shaded regions correspond to 99\% percentile bootstrapped intervals for each group. Vertical
+    axes vary to better depict group-wise overlap.}
+\label{fig:rigidity}
+\end{figure}
+
+\begin{figure}[ht]
+\centerline{\includegraphics[width=\textwidth]{figure2}} \caption{
+    (Colour online) Level variance for controls (nopain) vs. those with osteopathic pain taking
+    duloxetine. Solid lines indicate group mean level variance, and shaded regions correspond to
+    99\% percentile bootstrapped intervals for each group.  Vertical axes vary to better depict
+    group-wise overlap.
+}
+\label{fig:levelvar}
+\end{figure}
+
+\begin{figure}[ht]
+\centerline{\includegraphics[width=\textwidth]{figure3}} \caption{
+    (Colour online) Nearest Neighbours Spacings Distribution for controls (nopain) vs. those with
+    osteopathic pain taking duloxetine. Solid lines indicate Brody distribution maximum likelihood
+    estimates, and transparent regions depict the overlapping spacings histograms for all subjects.
+}
+\label{fig:nnsd}
+\end{figure}
 
@@
 features is challenging. For subject level classification, state of the art deep learning methods
-improve upon guessing by 17\% for autism (\cite{bengs4DSpatioTemporalDeep2020}) or 3-30\% for ADHD
-(\cite{riazDeepFMRIEndtoendDeep2020}), 16\% for severe depression
-(\cite{ramasubbuAccuracyAutomatedClassification2016}), and 23\% for obsessive compulsive disorder
-(\cite{takagiNeuralMarkerObsessiveCompulsive2017}). Manual feature engineering with more separable
+improve upon guessing by 17\% for autism \citep{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
+\citep{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
+\citep{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive compulsive disorder
+\citep{takagiNeuralMarkerObsessiveCompulsive2017}. Manual feature engineering with more separable
 conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
-(\cite{duHighClassificationAccuracy2012}), and with very large data, sophisticated custom feature
+\citep{duHighClassificationAccuracy2012}, and with very large data, sophisticated custom feature
 extraction methods can achieve near perfect accuracies at classifying task vs. rest
-(\cite{zhangCharacterizingDifferentiatingTaskbased2016}). However, for functional connectivity data
+\citep{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional connectivity data
 and with classical machine learning algorithms (such as SVM) we in general only expect large
 prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
-strongly separated (e.g. Cohen’s d > 1.0; \cite{dansereauStatisticalPowerPrediction2017}).
+strongly separated (e.g. Cohen’s \(d > 1.0\); \cite{dansereauStatisticalPowerPrediction2017}.
 
@@
 correlation matrix is primarily determined by such differences in dimensions
-(\cite{yinLimitLargestEigenvalue1988}), as are the noise ratios, which include these dimensions in
-their definitions (\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}). Likewise,
+\citep{yinLimitLargestEigenvalue1988}, as are the noise ratios, which include these dimensions in
+their definitions \citep{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}. Likewise,
 similar strong separation was observed in the PARK data when comparing the minimally pre-processed
@@
 (spike removal, spatial smoothing, co-registration to larger T1 image;
-\cite{madhyasthaDynamicConnectivityRest2015}) which will also almost certainly alter the size of the
+\cite{madhyasthaDynamicConnectivityRest2015} which will also almost certainly alter the size of the
 largest eigenvalue.
@@
 subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
-noise-reduction procedures (such as
-\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}), is unlikely to introduce bias
+noise-reduction procedures
+\citep[as in][]{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}, is unlikely to introduce bias
 into subsequent analyses. It perhaps also suggests these scalar features might be useful for quality
@@
 We generally find different results from previous studies
-(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
+\citep{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
 which have used the spectral rigidity or level number variance to differentiate between attention
@@
 
-\section{References}
-\printbibliography{}
+% \section{References}
+% \bibliography{rmt.bib}
+% \references{rmt.bib}
+
+% \bibstyle{plain}
+% \bibdata{rmt}
+\bibliography{rmt}
+
+\section{Technical Terms}
+
+\textbf{Functional Connectivity} the correlations and/or interactions between the timeseries of
+various regions of interest in fMRI scans.
+
+\textbf{Voxelwise Functional Connectivity} the correlations and/or interactions between the
+timeseries of all voxels in an fMRI scan.
+
+\textbf{Random Matrix} A random matrix is a matrix where each entry is sampled from some probability
+distribution.
+
+\textbf{Random Matrix Theory} or RMT is an area of mathematics concerned with the properties of
+large random matrices with identically and independently distributed (\emph{i.i.d.}) entries. RMT often investigates
+the limiting distributions of the eigenvalues of such matrices.
+
+\textbf{Gaussian Orthogonal Ensemble} or GOE. The set of real symmetric matrices with entries
+distributed according to the standard Gaussian.
 

### 2021-09-16 paper/latex/NetworkNeuroscience/cover_letter.tex
add cover letter

\documentclass[NETN,manuscript]{stjour-new}
\articletype{Research}
\usepackage{graphicx}
\graphicspath{ {figs/} }
\usepackage{hyperref}
% \input{preamble.tex}
% \usepackage[utf8]{inputenc}
% \usepackage{geometry}
% \usepackage{amsmath}
% \usepackage{graphicx}
% \usepackage[style=apa]{biblatex}
% \usepackage{natbib}
% \usepackage[font=footnotesize,labelfont=bf]{caption}
% \addbibresource{rmt.bib}
% \bibliographystyle{plain}
% \bibliography{rmt}
\begin{document}

\title{Cover Letter}
\date{\today}

\author[Derek Berger, Gurpreet Matharoo, Jacob Levman]{
    Derek M. Berger\affil{1},
    Gurpreet S. Matharoo\affil{2}
    \and Jacob Levman\affil{3}
}
\affiliation{1}{Computer Science, St. Francis Xavier University, Antigonish, Canada, dberger@stfx.ca}
\affiliation{2}{Physics, St. Francis Xavier University, Antigonish, Canada, gmatharo@stfx.ca}
\affiliation{3}{Computer Science, St. Francis Xavier University, Antigonish, Canada, jlevman@stfx.ca}

% \keywords{random matrix theory, feature extraction, functional connectivity,
% functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}

To the Editors of Network Neuroscience

We believe our manuscript \emph{Random Matrix Theory Tools for the Analysis of Functional Magnetic
Resonance Imaging Examinations} would be an excellent fit for Network Neuroscience as a
\emph{Research} article. Our use of random matrix theory (RMT) for the analysis of functional
connectivity in functional magnetic resonance imaging (fMRI) falls clearly within the \href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope}
of \emph{Network Neuroscience}.

To the best of our knowledge, our voxelwise analysis of fMRI using RMT features is novel\(\mathbf{^{[1]}}\): the RMT
features computed here summarize the correlations of \emph{all} fMRI voxels (rather than those from
some seed voxels), but previous works (documented in the manuscript) have been ROI focused. We also
present results in a predictive framework using cross-validation on a wide variety of datasets,
which we believe is more rigorous than an approach based solely on statistical significance for a
single dataset. Given there are very few uses of RMT for fMRI analysis, we hope you will find this
manuscript to be worth your while.

Should you consider to review our manuscript, we believe the following reviewers would be suitable:

\begin{itemize}
\item Sarika Jalan (sarika@iiti.ac.in) \\
\small
~~~ Centre for Biosciences and Biomedical Engineering, IIT Indore India
\normalsize
\item Javeria Hashmi (javeria.hashmi@dal.ca) \\
\small
~~~ Department of Anesthesia, Pain Management and Perioperative Medicine Dalhousie University, Canada
\normalsize
\item Nivedita Deo (ndeo@physics.du.ac.in) \\
\small
~~~ Department of Physics and Astrophysics University of Delhi, India
\normalsize
\item Guillermo A. Cecchi (gcecchi@us.ibm.com) \\
\small
~~~ Thomas J. Watson Research Center, Yorktown Heights, NY USA
\normalsize
\item Randy McIntosh (rmcintosh@research.baycrest.org) \\
\small
~~~ Rotman Research Institute at Baycrest Health Sciences, Toronto, Canada
\normalsize
\end{itemize}

All authors approve the manuscript in its submitted form, and have no conflicts of interest,
financial or otherwise. [Names, affiliations, email addresses and funding information]

\textbf{[1]} A poster of some preliminary results of our study was presented at the ISMRM annual
conference \citep{bergerOpenSourceRandom2021}.

Sincerely, \\
Corresponding Author \\
Institution Title \\
Institution/Affiliation Name \\
Institution Address \\
Corresponding author e-mail address \\

~\\
\textbf{Additional Contact} [should the corresponding author be unavailable] \\
Institution Title \\
Institution/Affiliation Name \\
Institution Address \\
e-mail address \\

\bibliography{rmt}

\end{document}

### 2021-09-16 paper/latex/NetworkNeuroscience/rmt.tex
add cover letter

@@
 \usepackage{graphicx}
+\usepackage{cleveref}
 \graphicspath{ {figs/} }
@@
 
-\date{\today}
+% \date{\today}
 
-% \maketitle
-
-% \keywords{random matrix theory, feature extraction, functional connectivity,
-% functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
+\keywords{random matrix theory, feature extraction, functional connectivity,
+functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
 
@@
     application of RMT-based features for the analysis of functional MRI (fMRI) across diverse
-
     datasets. As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain
@@
 (BOLD) signals are related to neural activity. It is common to investigate statistical relationships
-between the BOLD signals through functional connectivity analyses, where correlations between
+between the BOLD signals through \textbf{functional connectivity} analyses, where correlations between
 collections of these signals are examined to infer connections between different voxels or regions
@@
 \citep{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
-This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory (RMT), a set
+This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
 of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
@@
 universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
-disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has been applied to atmospheric
-physics \citep{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
-network-formation in liquids \citep{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}  and amorphous clusters \citep{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
-Within biological systems,
-RMT has also been used to successfully model aspects of  amino acid functional relationships \citep{bhadolaTargetingFunctionalMotifs2016},
-synchrony in epileptic seizures \citep{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein interactions both in different species \citep{agrawalQuantifyingRandomnessProtein2014} and breast cancer \citep{raiRandomnessPreservedPatterns2015}.
+disordered mesoscopic systems
+\citep{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
+and at larger scales, RMT has been applied to atmospheric physics
+\citep{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
+\citep{plerouRandomMatrixApproach2002} , social networks
+\citep{jalanUncoveringRandomnessSuccess2014}, random networks
+\citep{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in liquids
+\citep{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}  and
+amorphous clusters
+\citep{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
+Within biological systems, RMT has also been used to successfully model aspects of  amino acid
+functional relationships \citep{bhadolaTargetingFunctionalMotifs2016}, synchrony in epileptic
+seizures \citep{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein
+interactions both in different species \citep{agrawalQuantifyingRandomnessProtein2014} and breast
+cancer \citep{raiRandomnessPreservedPatterns2015}.
 \end{sloppypar}
@@
 states \citep{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
-resting or low-attention states exhibited properties closest to the GOE.
-
-These findings suggest that certain aspects of psychological processes might be characterized, in
-part, by features computed from the eigenvalues of fMRI correlation matrices, and that these
-features might vary in an interpretable manner across psychological processes. If this is the case,
-RMT could have unique potential in characterizing the functioning of the human brain.
+resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
+certain aspects of psychological processes might be characterized, in part, by features computed
+from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
+interpretable manner across psychological processes. If this is the case, RMT could have unique
+potential in characterizing the functioning of the human brain.
 
 Here, we use a novel, voxel-based approach, and expand the applications of RMT to analyze fMRI scans
-from  diverse datasets1. We extract RMT-inspired whole-brain features from voxelwise functional
-connectivity data, and assess the predictive value of these features. We also examine in detail the
-robustness of the RMT methodology to various common analytic choices.
+from  diverse datasets. We extract RMT-inspired whole-brain features from \textbf{voxelwise
+functional connectivity} data, and assess the predictive value of these features. We also examine in
+detail the robustness of the RMT methodology to various common analytic choices.
 
@@
 variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
-psychological processes, we examine a wide selection of fMRI datasets available on the OpenNeuro
-platform [36].  This allows for a broad assessment of the potential of voxelwise RMT features.
+psychological processes, we examine a wide selection of fMRI datasets available on the
+\href{https://openneuro.org/}{OpenNeuro} platform [36].  This allows for a broad assessment of the
+potential of voxelwise RMT features.
 
-However, since previous studies employing RMT for feature extraction
-\citep{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020} have partially interpreted
-their findings with respect to attentional processes, we have elected to use datasets that involve
-attention in some way. This includes measures of attention or vigilance, or comparisons between a
-condition involving more attention and focus (e.g. a task state) or conditions which should involve
-considerably less attention or focus (e.g. resting-state).
+However, since previous studies employing RMT for feature extraction have partially interpreted
+their findings with respect to attentional processes
+\citep{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have elected to use
+datasets that involve attention in some way. This includes measures of attention or vigilance, or
+comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
+which should involve considerably less attention or focus (e.g. resting-state).
+
+Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
+and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
+(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
+these datasets, we highlight below only those aspects relevant to the current paper.
 
@@
 \end{table}
-
-Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
-and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
-(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
-these datasets, we highlight below only those aspects relevant to the current paper.
 
@@
 
-\paragraph{Trait Attention}
+\emph{Trait Attention}.
 The PANAS-X \citep{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure
@@
 
-\paragraph{Task Attention}
+\emph{Task Attention}.
 The Conjunctive Continuous Performance Task (CCPT;
@@
 
-\paragraph{Vigilance}
+\emph{Vigilance}.
 The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
@@
 \subsubsection{Dataset 2}
-Dataset 2 (referred to in Fig. and tables with the label ``OSTEO'') is detailed in
+Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
 \cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
@@
 image with overall increased contrast \citep{kunduMultiechoFMRIReview2017}. In the ``interleave''
-condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
-from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
+condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
+from echo \(j\) as \(t_{j,i}\), then the interleaved voxel of the new image has the time-series \([t_{1,1},
 t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
@@
 \subsubsection{Dataset 4}
-Dataset 4 (referred to in Fig. and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
+Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
 described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
@@
 \subsubsection{Dataset 5}
-Dataset 5 (referred to in Fig. and tables with the label ``LEARN'') is documented and described in
+Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
 detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
@@
 image dimensions are identical between subjects, the RMT features (by virtue of their definition and
-the unfolding procedure) already allow for direct comparisons. Thus, the usual motivations for
+the \textbf{unfolding} procedure) already allow for direct comparisons. Thus, the usual motivations for
 registration are also no longer clear when extracting RMT features.
@@
 We examine as a potential feature the proportion of eigenvalues that fall within this range, and
-call this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
+call this the \textbf{noise ratio}. In addition, we calculate a broader, shifted noise ratio, based on the
 adjustments of \cite{almogUncoveringFunctionalSignature2019} which takes into account the larger
@@
 \times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
-value of \(t\) for a scan can be inferred from the number of volumes in Table \ref{table:1}. The various RMT
-features mentioned above were then extracted via the empyricalRMT library [64] developed and made
-publicly available as part of this study, as well as via specific code for this paper (which is
-documented and available online \citep{bergerderekm.DMBergerRandommatrixfmriV02020}.
+value of \(t\) for a scan can be inferred from the number of volumes in Table \ref{table:1}. The
+various RMT features mentioned above were then extracted via the \texttt{empyricalRMT} library
+\citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} developed and made publicly available as part
+of this study, as well as via code for this paper \citep{bergerderekm.DMBergerRandommatrixfmriV02020}.
 
@@
 discriminant analysis (LDA), and a support-vector machine (SVM) were employed. Default parameter
-values were used when possible (see \cite{bergerderekm.DMBergerRandommatrixfmriV02020} for precise
-function calls used).
+values were used when possible \citep[see][for precise function calls used]{bergerderekm.DMBergerRandommatrixfmriV02020}.
 
@@
 results are available in the supplementary materials. Visual summaries of the prediction accuracies
-for all features and analytic choices are presented in Fig. S1--4. Prediction accuracies for only the
-multivariate RMT features are summarized in Fig. S5--7, and the distributions of the scalar features
-are presented in Fig. S8--10. Prediction accuracies for the scalar features across analytic choices
-are summarized in Fig. S11--12.
+for all features and analytic choices are presented in figures S1--4. Prediction accuracies for only the
+multivariate RMT features are summarized in figures S5--7, and the distributions of the scalar features
+are presented in figures S8--10. Prediction accuracies for the scalar features across analytic choices
+are summarized in figures S11--12.
 
@@
 With the exception of the LEARNING dataset, the RMT-inspired features all demonstrated some
-predictive utility (Fig. S1-4). For example, there were strong differences in RMT features between
+predictive utility (figures S1-4). For example, there were strong differences in RMT features between
 task and rest conditions for the REFLECT dataset, but the features were of variable predictive
@@
 in the OSTEO dataset, these features mostly did not aid in classifying between any of the other
-possible pairings of subgroups in this dataset (Fig. S1-4).
+possible pairings of subgroups in this dataset (figures S1-4).
 
@@
 for the raw and largest eigenvalues to be the most predictively useful features, especially in the
-REFLECT and PARK data in discriminating between Parkinson's and controls (Fig. S5-7. S13; but see
+REFLECT and PARK data in discriminating between Parkinson's and controls (figures S5-7. S13; but see
 Discussion).
@@
 \subsection{Raw Features}
-As evident from Fig. S8-10, the scalar raw features (largest eigenvalue, noise ratio, and shifted
+As evident from figures S8-10, the scalar raw features (largest eigenvalue, noise ratio, and shifted
 noise ratio) varied more within than between subgroups. They were not generally useful for
@@
 control groups in the OSTEO data, allowing for prediction accuracies of up to 72\% compared to 51\%
-for the naive classifier (Fig. S11-12 top-right panel, Table S5). The largest eigenvalue was also
+for the naive classifier (figures S11-12 top-right panel, Table S5). The largest eigenvalue was also
 useful in distinguishing between subjects self-reporting high versus low vigilance across sessions:
@@
 separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
-consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
+consistently demonstrated predictive utility (Tables S7--8). As visible in figures S13, the eigenvalues
 had the most consistent predictive value for distinguishing individuals with Parkinson's from
@@
 
-\paragraph{Role of Classifer}
+\subsubsection{Role of Classifer}
 The raw eigenvalues generally had the most predictive utility when employing either an SVM
@@
 variance strongly differentiated controls from individuals prescribed duloxetine for osteopathic
-pain, and this differentiation persisted regardless of unfolding degree or trimming (Fig. 5-8, top-right panels).
+pain, and this differentiation persisted regardless of unfolding degree or trimming
+(\cref{fig:levelvar,fig:nnsd,fig:rigidity}, top-right panels).
 
@@
 unfolding degree, and sometimes trim proportions. These differences are illustrated in detail for
-the Parkinson's data in Fig. S14. For example, with minimal trimming, subgroup overlap in spectral
+the Parkinson's data in figures S14. For example, with minimal trimming, subgroup overlap in spectral
 rigidity was minimized for unfolding degrees 5 and 9 (Fig. S14, first row), but with greater
@@
 
-\paragraph{Role of Classifier}. Across all datasets, unfoldings, and trimmings, the spectral rigidity was
+\subsubsection{Role of Classifier}. Across all datasets, unfoldings, and trimmings, the spectral rigidity was
 generally best able to separate between subgroups when either k was large (10 or 20) in KNN, or
@@
 prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
-strongly separated (e.g. Cohen’s \(d > 1.0\); \cite{dansereauStatisticalPowerPrediction2017}.
+strongly separated (e.g. Cohen’s \(d > 1.0\); \citealp{dansereauStatisticalPowerPrediction2017}).
 
@@
 (spike removal, spatial smoothing, co-registration to larger T1 image;
-\cite{madhyasthaDynamicConnectivityRest2015} which will also almost certainly alter the size of the
-largest eigenvalue.
+\citealp{madhyasthaDynamicConnectivityRest2015}) which will also almost certainly alter the size of
+the largest eigenvalue.
 
@@
 
-\section{Conclusion}
+% \section{Conclusion}
 
-\paragraph{Declarations}
-Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
-Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
-Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
-Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
-(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
-R0168020), and a St. Francis Xavier University UCR grant to JL.
+% \paragraph{Declarations}
+% Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
+% Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
+% Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
+% Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
+% (R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
+% R0168020), and a St. Francis Xavier University UCR grant to JL.
 
-% \section{References}
-% \bibliography{rmt.bib}
-% \references{rmt.bib}
-
-% \bibstyle{plain}
-% \bibdata{rmt}
 \bibliography{rmt}
@@
 various regions of interest in fMRI scans.
-
-\textbf{Voxelwise Functional Connectivity} the correlations and/or interactions between the
-timeseries of all voxels in an fMRI scan.
 
@@
 
+\textbf{Voxelwise Functional Connectivity} the full matrix of correlations and/or interactions
+between the timeseries of \emph{all} voxels in an fMRI scan.
+
+\textbf{unfolding} A procedure wherein the eigenvalues of an empirically-observed matrix are
+re-mapped to a common space to allow comparison with the spectra of certain known ensembles.
+
 \textbf{Gaussian Orthogonal Ensemble} or GOE. The set of real symmetric matrices with entries
@@
 
+\textbf{The Nearest-Neighbors Spacings Distribution} or NNSD is the is the distribution of the
+spacings (differences) of the adjacent unfolded and sorted eigenvalues.
+
+\textbf{Noise Ratio} The proportion of eigenvalues that fall within the maximum and minimum expected
+limits of the Marchenko-Pastur distribution.
+
 \end{document}

### 2021-09-20 paper/latex/NetworkNeuroscience/cover_letter.tex
update letter

@@
-\documentclass[NETN,manuscript]{stjour-new}
-\articletype{Research}
+\documentclass{article}
+\usepackage[utf8]{inputenc}
+\usepackage{hyperref}
+\usepackage{url}
+% \usepackage{apalike}
+% \documentclass[NETN,manuscript]{stjour-new}
+% \articletype{Research}
 \usepackage{graphicx}
 \graphicspath{ {figs/} }
-\usepackage{hyperref}
 % \input{preamble.tex}
-% \usepackage[utf8]{inputenc}
-% \usepackage{geometry}
-% \usepackage{amsmath}
-% \usepackage{graphicx}
-% \usepackage[style=apa]{biblatex}
+\usepackage{geometry}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{natbib}
+% \usepackage[backend=biber,
+%             style=authoryear,
+%             natbib=true,
+% ]{biblatex}
+% \addbibresource{cover.bib}
+\usepackage{fancyhdr}
 % \usepackage{natbib}
@@
 % \bibliographystyle{plain}
-% \bibliography{rmt}
+% \bibstyle{plain}
+% \bibliography{cover_letter}
+
 \begin{document}
+\bibliographystyle{plain}
+\thispagestyle{fancy}
+% \addtolength{\topmargin}{-63.87003pt}
+\fancyhead[C]{\includegraphics[height=0.5in]{stfx_text.png}}
+\fancyhead[L]{\includegraphics[height=0.5in]{crest.png}}
+\fancyhead[R]{
+    \sffamily
+    \scriptsize
+    4130 University Avenue \\
+    B2G 2W5 \\
+    Antigonish, Nova Scotia \\
+    Canada
+}
+\normalsize
 
 \title{Cover Letter}
-\date{\today}
+% \begin{center}
+% \includegraphics[width=4in]{stfx_logo.png}
+% \end{center}
 
-\author[Derek Berger, Gurpreet Matharoo, Jacob Levman]{
-    Derek M. Berger\affil{1},
-    Gurpreet S. Matharoo\affil{2}
-    \and Jacob Levman\affil{3}
-}
-\affiliation{1}{Computer Science, St. Francis Xavier University, Antigonish, Canada, dberger@stfx.ca}
-\affiliation{2}{Physics, St. Francis Xavier University, Antigonish, Canada, gmatharo@stfx.ca}
-\affiliation{3}{Computer Science, St. Francis Xavier University, Antigonish, Canada, jlevman@stfx.ca}
+% \author[Derek Berger, Gurpreet Matharoo, Jacob Levman]{
+%     Derek M. Berger\affil{1},
+%     Gurpreet S. Matharoo\affil{2}
+%     \and Jacob Levman\affil{3}
+% }
+% \affiliation{1}{Computer Science, St. Francis Xavier University, Antigonish, Canada, dberger@stfx.ca}
+% \affiliation{2}{Physics, St. Francis Xavier University, Antigonish, Canada, gmatharo@stfx.ca}
+% \affiliation{3}{Computer Science, St. Francis Xavier University, Antigonish, Canada, jlevman@stfx.ca}
 
@@
 
-To the Editors of Network Neuroscience
+~\\~\\
+To the Editors of Network Neuroscience, \\
 
-We believe our manuscript \emph{Random Matrix Theory Tools for the Analysis of Functional Magnetic
-Resonance Imaging Examinations} would be an excellent fit for Network Neuroscience as a
-\emph{Research} article. Our use of random matrix theory (RMT) for the analysis of functional
-connectivity in functional magnetic resonance imaging (fMRI) falls clearly within the \href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope}
-of \emph{Network Neuroscience}.
+We hope you will consider our manuscript \emph{Random Matrix Theory Tools for the Analysis of
+Functional Magnetic Resonance Imaging Examinations} for publication in Network Neuroscience as a
+\emph{Research} article.
+
+In our manuscript, we apply some of the analytical and statistical tools of random matrix theory
+(RMT) to analyze a wide variety of fMRI data in a predictive framework, and release open-source
+software to facilitate such analyses with a wide variety of data.
+
+Our use of random matrix theory (RMT) for the analysis of functional
+connectivity in functional magnetic resonance imaging (fMRI) falls clearly within the
+\href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope} of \emph{Network
+Neuroscience}.
 
@@
 
-Should you consider to review our manuscript, we believe the following reviewers would be suitable:
+This paper is not under consideration at any other journal, although a poster of some preliminary
+results of our study was presented at the International Symposium for Magnetic Resonance in Medicine
+2021 annual conference \citep{bergerOpenSourceRandom2021}. Our lab has REB ethics approval for
+conducting retrospective analyses, although we are unaware of the need for this, this being an
+analysis of anonymized public data. All authors approve the manuscript in its submitted form, and
+have no conflicts of interest, financial or otherwise.
+
+Should you consider to review our manuscript, we believe the following reviewers would be suitable
+due to expertise in this areas of research:
 
@@
 
-All authors approve the manuscript in its submitted form, and have no conflicts of interest,
-financial or otherwise. [Names, affiliations, email addresses and funding information]
+Sincerely,\\ \textbf{Corresponding Author}\\
+Jacob Levman, PhD (jlevman@stfx.ca) \\
+Canada Research Chair in Bioinformatics \\
+Department of Computer Science \\
+St. Francis Xavier University
 
-\textbf{[1]} A poster of some preliminary results of our study was presented at the ISMRM annual
-conference \citep{bergerOpenSourceRandom2021}.
+\textbf{Additional Contacts} [should the corresponding author be unavailable]\\
+Derek Berger (dberger@stfx.ca), Department of Computer Science, St. Francis Xavier University\\
+Gurpreet S. Matharoo (gmatharo@stfx.ca), Department of Physics, St. Francis Xavier University
 
-Sincerely, \\
-Corresponding Author \\
-Institution Title \\
-Institution/Affiliation Name \\
-Institution Address \\
-Corresponding author e-mail address \\
+\today{}
 
-~\\
-\textbf{Additional Contact} [should the corresponding author be unavailable] \\
-Institution Title \\
-Institution/Affiliation Name \\
-Institution Address \\
-e-mail address \\
+[Names, affiliations, email addresses and funding information]
 
-\bibliography{rmt}
+% \printbibliography % biblatex
+\bibliography{cover.bib}  % natbib
 

### 2021-09-21 paper/latex/NetworkNeuroscience/cover_letter.tex
dramatically improve cover letter

@@
 % \addtolength{\topmargin}{-63.87003pt}
+\setlength{\headheight}{30.0pt}
 \fancyhead[C]{\includegraphics[height=0.5in]{stfx_text.png}}
@@
 
-~\\~\\
-To the Editors of Network Neuroscience, \\
+\noindent
+\today, \\
+To the Many Editors of Network Neuroscience, \\
 
@@
 Functional Magnetic Resonance Imaging Examinations} for publication in Network Neuroscience as a
-\emph{Research} article.
+\emph{Research} article. In it, we apply mathematical tools from random matrix theory (RMT) to
+analyze a wide variety of functional magnetic resonance imaging (fMRI) data, and also release
+open-source software to facilitate such analyses with a wider variety of data beyond fMRI.
 
-In our manuscript, we apply some of the analytical and statistical tools of random matrix theory
-(RMT) to analyze a wide variety of fMRI data in a predictive framework, and release open-source
-software to facilitate such analyses with a wide variety of data.
+Numerous fMRI studies examine correlations amongst regions of interest (ROI) in the brain. These
+``functional connectivity'' analyses typically use ROIs or seed regions in order to simplify
+connectivity computations \citep{huettelFunctionalMagneticResonance2004}. Instead, we use
+statistics from RMT to summarize \emph{all} voxelwise correlations with no reductions. In
+addition, rather than simply reporting findings which are statistically significant, we report all
+analyses from a cross-validated predictive framework, where multiple machine learnign models are
+compared against both each other and proper baselines (e.g. niave / dummy classifiers).
 
-Our use of random matrix theory (RMT) for the analysis of functional
-connectivity in functional magnetic resonance imaging (fMRI) falls clearly within the
-\href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope} of \emph{Network
-Neuroscience}.
-
-To the best of our knowledge, our voxelwise analysis of fMRI using RMT features is novel\(\mathbf{^{[1]}}\): the RMT
-features computed here summarize the correlations of \emph{all} fMRI voxels (rather than those from
-some seed voxels), but previous works (documented in the manuscript) have been ROI focused. We also
-present results in a predictive framework using cross-validation on a wide variety of datasets,
-which we believe is more rigorous than an approach based solely on statistical significance for a
-single dataset. Given there are very few uses of RMT for fMRI analysis, we hope you will find this
-manuscript to be worth your while.
+We show that, across a variety of data and with fairly small data, and without any model tuning,
+that these RMT reductions have real predictive utility and potential. This suggests RMT has
+potential as a powerful method for summarizing functional connectivity information. As much of the
+research in network neuroscience involves fMRI, functional connectivity, and statistical physics, as
+per the \href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope} of \emph{Network
+Neuroscience}, we believe our manuscript to be an excellent fit for submission to your journal.
 
@@
 analysis of anonymized public data. All authors approve the manuscript in its submitted form, and
-have no conflicts of interest, financial or otherwise.
+have no conflicts of interest, financial or otherwise. Should you consider to review our manuscript,
+we have suggested a number of potential reviewers on your submission website that we believe would
+be suitable due to expertise in RMT and similar interdisciplinary research:
 
-Should you consider to review our manuscript, we believe the following reviewers would be suitable
-due to expertise in this areas of research:
+% \begin{itemize}
+% \item Sarika Jalan (sarika@iiti.ac.in) \\
+% \small
+% ~~~ Centre for Biosciences and Biomedical Engineering, IIT Indore India
+% \normalsize
+% \item Javeria Hashmi (javeria.hashmi@dal.ca) \\
+% \small
+% ~~~ Department of Anesthesia, Pain Management and Perioperative Medicine Dalhousie University, Canada
+% \normalsize
+% \item Nivedita Deo (ndeo@physics.du.ac.in) \\
+% \small
+% ~~~ Department of Physics and Astrophysics University of Delhi, India
+% \normalsize
+% \item Guillermo A. Cecchi (gcecchi@us.ibm.com) \\
+% \small
+% ~~~ Thomas J. Watson Research Center, Yorktown Heights, NY USA
+% \normalsize
+% \item Randy McIntosh (rmcintosh@research.baycrest.org) \\
+% \small
+% ~~~ Rotman Research Institute at Baycrest Health Sciences, Toronto, Canada
+% \normalsize
+% \end{itemize}
 
-\begin{itemize}
-\item Sarika Jalan (sarika@iiti.ac.in) \\
-\small
-~~~ Centre for Biosciences and Biomedical Engineering, IIT Indore India
-\normalsize
-\item Javeria Hashmi (javeria.hashmi@dal.ca) \\
-\small
-~~~ Department of Anesthesia, Pain Management and Perioperative Medicine Dalhousie University, Canada
-\normalsize
-\item Nivedita Deo (ndeo@physics.du.ac.in) \\
-\small
-~~~ Department of Physics and Astrophysics University of Delhi, India
-\normalsize
-\item Guillermo A. Cecchi (gcecchi@us.ibm.com) \\
-\small
-~~~ Thomas J. Watson Research Center, Yorktown Heights, NY USA
-\normalsize
-\item Randy McIntosh (rmcintosh@research.baycrest.org) \\
-\small
-~~~ Rotman Research Institute at Baycrest Health Sciences, Toronto, Canada
-\normalsize
-\end{itemize}
+\bigskip
 
-Sincerely,\\ \textbf{Corresponding Author}\\
+\noindent
+Sincerely, \\
+
+\noindent
+\textbf{Corresponding Author}\\
 Jacob Levman, PhD (jlevman@stfx.ca) \\
@@
 Department of Computer Science \\
-St. Francis Xavier University
+St. Francis Xavier University \\
 
+\noindent
 \textbf{Additional Contacts} [should the corresponding author be unavailable]\\
@@
 
-\today{}
-
-[Names, affiliations, email addresses and funding information]
-
 % \printbibliography % biblatex
+\renewcommand{\bibsection}{}  % https://tex.stackexchange.com/questions/132646/how-to-remove-the-references-title
 \bibliography{cover.bib}  % natbib

### 2021-09-21 paper/latex/NetworkNeuroscience/cover_letter.tex
tweaks

@@
 \bibliographystyle{plain}
-\thispagestyle{fancy}
-% \addtolength{\topmargin}{-63.87003pt}
+\fancypagestyle{nofooter}{\fancyfoot{}}
+\thispagestyle{nofooter}  %  https://tex.stackexchange.com/questions/136384/remove-footer-without-removing-the-header
 \setlength{\headheight}{30.0pt}
@@
 addition, rather than simply reporting findings which are statistically significant, we report all
-analyses from a cross-validated predictive framework, where multiple machine learnign models are
-compared against both each other and proper baselines (e.g. niave / dummy classifiers).
+analyses from a cross-validated predictive framework, where multiple machine learning models are
+compared against both each other and proper baselines (e.g. naive / dummy classifiers).
 
 We show that, across a variety of data and with fairly small data, and without any model tuning,
-that these RMT reductions have real predictive utility and potential. This suggests RMT has
-potential as a powerful method for summarizing functional connectivity information. As much of the
+that these RMT reductions have real predictive utility and potential. This suggests RMT could
+be a powerful and efficient method for summarizing functional connectivity information. As much of the
 research in network neuroscience involves fMRI, functional connectivity, and statistical physics, as
@@
 
-This paper is not under consideration at any other journal, although a poster of some preliminary
-results of our study was presented at the International Symposium for Magnetic Resonance in Medicine
-2021 annual conference \citep{bergerOpenSourceRandom2021}. Our lab has REB ethics approval for
-conducting retrospective analyses, although we are unaware of the need for this, this being an
-analysis of anonymized public data. All authors approve the manuscript in its submitted form, and
-have no conflicts of interest, financial or otherwise. Should you consider to review our manuscript,
-we have suggested a number of potential reviewers on your submission website that we believe would
-be suitable due to expertise in RMT and similar interdisciplinary research:
+Should you consider to review our manuscript, we have suggested a number of potential reviewers on
+your submission website that we believe would be suitable due to expertise in RMT and similar
+interdisciplinary research. This paper is not under consideration at any other journal (although a
+poster of some preliminary results of our study was presented at the International Symposium for
+Magnetic Resonance in Medicine 2021 annual conference \citep{bergerOpenSourceRandom2021}). Our lab
+has REB ethics approval for conducting retrospective analyses (although we are unaware of the need,
+this being an analysis of anonymized public data). All authors approve the manuscript in its
+submitted form, and have no conflicts of interest, financial or otherwise.
 

### 2021-09-21 paper/latex/NetworkNeuroscience/cover_letter.tex
get complete bibliography

@@
-\documentclass{article}
+\documentclass[10pt]{article}
 \usepackage[utf8]{inputenc}
-\usepackage{hyperref}
+\usepackage{csquotes}
+\usepackage[british]{babel}
 \usepackage{url}
-% \usepackage{apalike}
-% \documentclass[NETN,manuscript]{stjour-new}
-% \articletype{Research}
 \usepackage{graphicx}
 \graphicspath{ {figs/} }
-% \input{preamble.tex}
 \usepackage{geometry}
@@
 \usepackage{graphicx}
-\usepackage{natbib}
-% \usepackage[backend=biber,
-%             style=authoryear,
-%             natbib=true,
-% ]{biblatex}
-% \addbibresource{cover.bib}
+% \usepackage[square,numbers]{natbib}
+\usepackage[
+    backend=biber,
+    style=authoryear,
+    natbib=true,
+    sortlocale=en_US,
+    url=false,
+    doi=true,
+    eprint=false
+]{biblatex}
+\addbibresource{cover.bib}
 \usepackage{fancyhdr}
-% \usepackage{natbib}
-% \usepackage[font=footnotesize,labelfont=bf]{caption}
-% \addbibresource{rmt.bib}
-% \bibliographystyle{plain}
-% \bibstyle{plain}
-% \bibliography{cover_letter}
+\usepackage{color}
+\definecolor{darkblue}{rgb}{0.,0.,0.4}
+\definecolor{darkred}{rgb}{0.5,0.,0.}
+\usepackage[pdftex,colorlinks=true,linkcolor=darkblue,citecolor=darkred,urlcolor=blue]{hyperref}
 
 \begin{document}
-\bibliographystyle{plain}
+% \bibliographystyle{apa}
 \fancypagestyle{nofooter}{\fancyfoot{}}
@@
 \emph{Research} article. In it, we apply mathematical tools from random matrix theory (RMT) to
-analyze a wide variety of functional magnetic resonance imaging (fMRI) data, and also release
-open-source software to facilitate such analyses with a wider variety of data beyond fMRI.
+analyze a wide variety of functional magnetic resonance imaging (fMRI) data, and release
+open-source software \citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} to facilitate such analyses with a wider variety of data beyond fMRI.
 
@@
 your submission website that we believe would be suitable due to expertise in RMT and similar
-interdisciplinary research. This paper is not under consideration at any other journal (although a
+interdisciplinary research. This paper is not under consideration at any other journa, although a
 poster of some preliminary results of our study was presented at the International Symposium for
-Magnetic Resonance in Medicine 2021 annual conference \citep{bergerOpenSourceRandom2021}). Our lab
+Magnetic Resonance in Medicine 2021 annual conference \citep{bergerOpenSourceRandom2021}. Our lab
 has REB ethics approval for conducting retrospective analyses (although we are unaware of the need,
@@
 
+\small
 \noindent
@@
 % \printbibliography % biblatex
-\renewcommand{\bibsection}{}  % https://tex.stackexchange.com/questions/132646/how-to-remove-the-references-title
-\bibliography{cover.bib}  % natbib
+% \renewcommand{\bibsection}{}  % https://tex.stackexchange.com/questions/132646/how-to-remove-the-references-title
+% \smallskip
+% \bibliographystyle{mystyle}
+% \bibliography{cover}  % natbib
+
+\footnotesize
+\AtNextBibliography{\footnotesize}
+\printbibliography[heading=none]
 

### 2021-11-17 paper/latex/generic/meta.tex
get generic article format working with agsm style

\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
\date{\today}
\author{ \href{https://orcid.org/0000-0003-4733-0624}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Derek M.~Berger}\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  St. Francis Xavier University\\
  Anigonish, Nova Scotia, Canada \\
  \texttt{dberger@stfx.ca} \\
  %% examples of more authors
  \AND
  \href{https://orcid.org/0000-0002-9604-3157}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jacob ~Levman} \\
  Canada Research Chair\\
  St. Francis Xavier University\\
  Anigonish, Nova Scotia, Canada \\
  \texttt{jlevman@stfx.ca} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

### 2021-11-17 paper/latex/generic/notes.tex
get generic article format working with agsm style

--- 2021-09-16 paper/latex/NetworkNeuroscience/rmt.tex
@@
-\documentclass[NETN,manuscript]{stjour-new}
-\articletype{Research}
+\documentclass[10pt]{article}
+\usepackage{arxiv}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage[english]{babel} % hyphenation
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsmath}        % https://tex.stackexchange.com/questions/41035/what-is-causing-undefined-control-sequence
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{bm}             % proper bolding
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{verbatim}
+% \usepackage{pdflscape}
 \usepackage{graphicx}
+\usepackage{doi}
 \usepackage{cleveref}
+
+\usepackage[round]{natbib}
+\setcitestyle{citesep={;}, aysep={,}}
+% \bibliographystyle{plainnat}
+% NOTE: Must be Better Bibtex for agsm style to work!
+\bibliographystyle{agsm}
+
+\newcommand{\tsub}[2]{\mathbf{#1}_{\text{#2}}}
+\DeclareMathOperator*{\argmin}{\arg\min}
 \graphicspath{ {figs/} }
-% \input{preamble.tex}
-% \documentclass{article}
-% \usepackage[utf8]{inputenc}
-% \usepackage{geometry}
-% \usepackage{amsmath}
-% \usepackage{graphicx}
-% \usepackage[style=apa]{biblatex}
-% \usepackage{natbib}
-% \usepackage[font=footnotesize,labelfont=bf]{caption}
-% \addbibresource{rmt.bib}
-% \bibliographystyle{plain}
-% \bibliography{rmt}
+
 \begin{document}
+\include{meta}
 
-\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
-\author[Derek M. Berger, Gurpreet S. Matharoo, Jacob Levman]{
-    Derek M. Berger\affil{1},
-    Gurpreet S. Matharoo\affil{2}
-    \and Jacob Levman\affil{3}
-}
-\affiliation{1}{Computer Science, St. Francis Xavier University, Antigonish, Canada}
-\affiliation{2}{Physics, St. Francis Xavier University, Antigonish, Canada}
-\affiliation{3}{Computer Science, St. Francis Xavier University, Antigonish, Canada}
+\maketitle
 
-% \date{\today}
-
-\keywords{random matrix theory, feature extraction, functional connectivity,
-functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
 \end{abstract}
+
+\keywords{random matrix theory, feature extraction, functional connectivity,
+functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
+Dataset scan parameters and acquisition details are summarized in Table \ref{tab:1}, and sample
 and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
@@
 \begin{table}[h!]
-\small
   \centering
@@
   }
-\label{table:1}
+  \label{tab:1}
 \end{table}
@@
 subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
-documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
+documented in Table \ref{tab:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
 complementary to \cite{matharooSpontaneousBackpainAlters2020}.
@@
 described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
-in Table \ref{table:1}. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
+in Table \ref{tab:1}. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
 number of repetitions of the Attention Network Task
-\citep[ANT;][]{fanActivationAttentionalNetworks2005}. Subgroup details are available in Table \ref{table:1}. As the ANT
+\citep[ANT;][]{fanActivationAttentionalNetworks2005}. Subgroup details are available in Table \ref{tab:1}. As the ANT
 task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
@@
 \times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
-value of \(t\) for a scan can be inferred from the number of volumes in Table \ref{table:1}. The
+value of \(t\) for a scan can be inferred from the number of volumes in Table \ref{tab:1}. The
 various RMT features mentioned above were then extracted via the \texttt{empyricalRMT} library
@@
 
-\bibliography{rmt}
+\begin{comment}
+\hyperref[tab:existing-attempts]{Table \ref{tab:existing-attempts}}).
+\begin{table}
+	\centering
+	\begin{tabular}{lll}
+		\toprule
+		Feature     & Name     & Shape  \\
+		\midrule
+    ROI means                                                                                  & \(\tsub{\mu}{ROI}\)                        & \((n, T)\)    \\
+    ROI standard deviations (sds)                                                              & \(\tsub{\sigma}{ROI}\)                     & \((n, T)\)    \\
+    ROI descriptive sequences                                                                  & \((\tsub{\mu}{ROI}, \tsub{\sigma}{ROI})\)  & \((2, n, T)\) or \((2n, T)\) \\
+    Correlation of ROI means                                                                   & \(\mathbf{r}_{\mu,\text{ROI}}\)            & \((n, n)\)    \\
+    Correlation of ROI sds                                                                     & \(\mathbf{r}_{\sigma,\text{ROI}}\)         & \((n, n)\)    \\
+    Correlation descriptives \([\mathbf{r}_{\mu,\text{ROI}}; \mathbf{r}_{\sigma\text{,ROI}}]\) & \(\tsub{r}{ROI}\)                          & \((2, n, n)\) or \((2n, n)\)\\
+    T-thresholded Laplacian eigenvalues of \(\mathbf{r}_{\mu,\text{ROI}}\)                     & \(\mathbf{\lambda}_{\mu,\text{ROI},T}\)    & \(n-1\)       \\
+    T-thresholded Laplacian eigenvalues of \(\mathbf{r}_{\sigma\text{,ROI}}\)                  & \(\mathbf{\lambda}_{\sigma,\text{ROI},T}\) & \(n-1\)       \\
+    Concatenated T-thresholded Laplacian eigenvalues                                           & \(\mathbf{\lambda}_{\text{ROI},T}\)        & \((2, n-1)\)  \\
+    Eigenvalues of \(\mathbf{r}_{\mu,\text{ROI}}\)                                             & \(\mathbf{\lambda}_{\mu,\text{ROI}}\)      & \(n-1\)       \\
+    Eigenvalues of \(\mathbf{r}_{\sigma\text{,ROI}}\)                                          & \(\mathbf{\lambda}_{\sigma,\text{ROI}}\)   & \(n-1\)       \\
+    Concatenated eigenvalues \([\mathbf{r}_{\sigma\text{,ROI}}; \mathbf{r}_{\sigma\text{,ROI}}]\)  & \(\tsub{\lambda}{ROI}\)                & \((2, n-1)\)  \\
+		\bottomrule
+	\end{tabular}
+	\label{tab:feature-shapes}
+	\caption{ROI-based feature shapes.
+\(n\) = number of ROIs.
+\([A;B]\) = concatenation of \(A\) and \(B\) along new initial (channel) dimension for deep learning, or along initial dimension, for classical ML.}
+\end{table}
 
-\section{Technical Terms}
+\begin{table}
+  \centering
+  \begin{tabular}{rr}
+  \toprule
+  T &   \(N\) subjects \\
+  \midrule
+   78 &   26 \\
+  116 &  134 \\
+  124 &    6 \\
+  146 &   84 \\
+  152 &   42 \\
+  176 &  227 \\
+  177 &    1 \\
+  196 &  158 \\
+  202 &    1 \\
+  206 &   34 \\
+  232 &    1 \\
+  234 &    2 \\
+  236 &  103 \\
+  246 &   63 \\
+  296 &  140 \\
+  316 &   13 \\
+  \bottomrule
+\end{tabular}\label{tab:timepoints}
+\caption{Heterogeneity in ABIDE number of timepoints. Mean number of timepoints = 193, median = 176. Standard deviation = 58.1}
+\end{table}
 
-\textbf{Functional Connectivity} the correlations and/or interactions between the timeseries of
-various regions of interest in fMRI scans.
+\begin{table}
+  \centering
+  \begin{tabular}{cr}
+  \toprule
+  TR (s) &   \(N\) subjects \\
+  \midrule
+  1.5  &      81 \\
+  1.7 &       56 \\
+  2    &     519 \\
+  2.2  &      26 \\
+  2.5  &      58 \\
+  3    &     131 \\
+  \bottomrule
+  \end{tabular}\label{tab:trs}
+\caption{Heterogeneity in ABIDE repetition times. Note many subjects have different TR values very close to 1.666 due to rounding issue, so are just reported here as 1.7 for simplicity}
+\end{table}
+\end{comment}
 
-\textbf{Random Matrix} A random matrix is a matrix where each entry is sampled from some probability
-distribution.
-
-\textbf{Random Matrix Theory} or RMT is an area of mathematics concerned with the properties of
-large random matrices with identically and independently distributed (\emph{i.i.d.}) entries. RMT often investigates
-the limiting distributions of the eigenvalues of such matrices.
-
-\textbf{Voxelwise Functional Connectivity} the full matrix of correlations and/or interactions
-between the timeseries of \emph{all} voxels in an fMRI scan.
-
-\textbf{unfolding} A procedure wherein the eigenvalues of an empirically-observed matrix are
-re-mapped to a common space to allow comparison with the spectra of certain known ensembles.
-
-\textbf{Gaussian Orthogonal Ensemble} or GOE. The set of real symmetric matrices with entries
-distributed according to the standard Gaussian.
-
-\textbf{The Nearest-Neighbors Spacings Distribution} or NNSD is the is the distribution of the
-spacings (differences) of the adjacent unfolded and sorted eigenvalues.
-
-\textbf{Noise Ratio} The proportion of eigenvalues that fall within the maximum and minimum expected
-limits of the Marchenko-Pastur distribution.
+% \bibliographystyle{unsrtnat}
+\bibliography{notes}
 

### 2021-11-17 paper/latex/PMB/cover_letter.tex
setup for PMB

--- 2021-09-21 paper/latex/NetworkNeuroscience/cover_letter.tex
@@
-\documentclass[10pt]{article}
+\documentclass[10pt,a4paper]{article}
 \usepackage[utf8]{inputenc}
@@
 \today, \\
-To the Many Editors of Network Neuroscience, \\
+To the Editors of Network Neuroscience, \\
 
@@
 analyze a wide variety of functional magnetic resonance imaging (fMRI) data, and release
-open-source software \citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} to facilitate such analyses with a wider variety of data beyond fMRI.
+open-source software \citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} to facilitate such analyses with a wide variety of data.
 
@@
 \footnotesize
+\begin{center}
+\noindent\rule{6cm}{0.4pt}
+\end{center}
+
 \AtNextBibliography{\footnotesize}

### 2021-11-17 paper/latex/PMB/title.tex
setup for PMB

--- 2021-09-13 paper/latex/NetworkNeuroscience/rmt.tex
@@
-\input{preamble.tex}
+\documentclass[12pt]{iopart}
+\usepackage{graphicx}
+\usepackage{cleveref}
+\graphicspath{ {figs/} }
+% \input{preamble.tex}
+% \documentclass{article}
+% \usepackage[utf8]{inputenc}
+% \usepackage{geometry}
+% \usepackage{amsmath}
+% \usepackage{graphicx}
+% \usepackage[style=apa]{biblatex}
+% \usepackage{natbib}
+% \usepackage[font=footnotesize,labelfont=bf]{caption}
+% \addbibresource{rmt.bib}
+% \bibliographystyle{plain}
+% \bibliography{rmt}
+% \bibstyle{harvard}
+\begin{document}
 
-\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
-\author{Derek M. Berger \and Gurpreet S. Matharoo \and Jacob Levman}
-\date{\today}
+\title[Random Matrix Theory for fMRI]{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
+\author{D M Berger$^1$, Gurpreet S. Matharoo$^2$, Jacob Levman$^3$}
+\address{$^1$ Computer Science, St. Francis Xavier University, Antigonish, Canada}
+\address{$^2$ Physics, St. Francis Xavier University, Antigonish, Canada}
+\address{$^3$ Computer Science, St. Francis Xavier University, Antigonish, Canada}
+\ead{jlevman@stfx.ca}
 
-\begin{document}
-\maketitle
+% \date{\today}
 
@@
     RMT-inspired features may have unique potential in analyses of fMRI functional connectivity.
-
-    \subparagraph{Keywords:} feature extraction, functional connectivity, functional magnetic
-    resonance imaging, random matrix theory, voxelwise connectivity, predictive analysis
-
-    \subparagraph{Authors}
-    Derek Berger is with the department of Computer Science, St. Francis Xavier University, Antigonish
-    NS B2G 2W5, Canada (e-mail: dberger@stfx.ca).
-
-    Gurpreet Matharoo is with ACENET, and the Department of Physics, St. Francis Xavier University,
-    Antigonish NS B2G 2W5, Canada (e-mail: gmatharo@stfx.ca).
-
-    Jacob Levman is with the Department of Computer Science, St. Francis Xavier University,
-    Antigonish NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).
 \end{abstract}
 
-\section{Introduction}
+\keywords{random matrix theory, feature extraction, functional connectivity,
+functional magnetic resonance imaging, voxelwise connectivity, predictive analysis}
 
-In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
-(BOLD) signals are related to neural activity. It is common to investigate statistical relationships
-between the BOLD signals through functional connectivity analyses, where correlations between
-collections of these signals are examined to infer connections between different voxels or regions
-of interest (ROIs) within the brain.
-
-Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
-complex functional connectivity networks are ubiquitous
-(\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}).
-This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory (RMT), a set
-of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
-physics (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). In various
-physical systems where nuclei can have different possible states, precisely modeling the entirety of
-the interactions between nuclei is often intractable. For such large, complex systems, the
-eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
-of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
-spectrum statistically, complex systems can be investigated by comparison to the universal
-properties predicted by RMT.
-
-\begin{sloppypar}
-These universal properties have been observed in diverse phenomena. In small-scale systems, RMT
-universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
-disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has been applied to atmospheric
-physics (\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
-network-formation in liquids (\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009})  and amorphous clusters (\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}).
-Within biological systems,
-RMT has also been used to successfully model aspects of  amino acid functional relationships (\cite{bhadolaTargetingFunctionalMotifs2016}),
-synchrony in epileptic seizures (\cite{osorioPhasesynchronizationRandommatrixBased2011}), and in protein-protein interactions both in different species (\cite{agrawalQuantifyingRandomnessProtein2014}) and breast cancer (\cite{raiRandomnessPreservedPatterns2015}).
-\end{sloppypar}
-
-In recent years, RMT analyses have been applied to study brain functioning.  The earliest  study
-demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) (\cite{sebaRandomMatrixAnalysis2003}).
-Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
-fMRI data (\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}). RMT has also
-been used in diffusion MRI to aid in the selection of the number of components to employ in
-principal-component reduction analysis and denoising
-(\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}).
-
-Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states (\cite{wangSpectralPropertiesTemporal2015}), between
-subjects with and without ADHD (\cite{wangRandomMatrixTheory2016}), and between pain and non-pain
-states (\cite{matharooSpontaneousBackpainAlters2020}). Across these three studies, the spectra of
-resting or low-attention states exhibited properties closest to the GOE.
-
-These findings suggest that certain aspects of psychological processes might be characterized, in
-part, by features computed from the eigenvalues of fMRI correlation matrices, and that these
-features might vary in an interpretable manner across psychological processes. If this is the case,
-RMT could have unique potential in characterizing the functioning of the human brain.
-
-Here, we use a novel, voxel-based approach, and expand the applications of RMT to analyze fMRI scans
-from  diverse datasets1. We extract RMT-inspired whole-brain features from voxelwise functional
-connectivity data, and assess the predictive value of these features. We also examine in detail the
-robustness of the RMT methodology to various common analytic choices.
-
-\section{Materials and Methods}
-
-\subsection{Datasets}
-
-As our primary intention is to examine the potential utility of RMT-derived features across a wide
-variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
-psychological processes, we examine a wide selection of fMRI datasets available on the OpenNeuro
-platform [36].  This allows for a broad assessment of the potential of voxelwise RMT features.
-
-However, since previous studies employing RMT for feature extraction
-(\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}) have partially interpreted
-their findings with respect to attentional processes, we have elected to use datasets that involve
-attention in some way. This includes measures of attention or vigilance, or comparisons between a
-condition involving more attention and focus (e.g. a task state) or conditions which should involve
-considerably less attention or focus (e.g. resting-state).
-
-\begin{table}[h!]
-\small
-\centering
-\begin{tabular}{ c c c c c c c }
-\hline
-\textbf{ID}  & \textbf{FOV (mm)}  & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{Time (s)} \\
-\hline
-PSYCH   &  192              & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0 & 300 & 15 \\
-OSTEO   &  -                & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5 & 300 & 12.5 \\
-REFLECT &  210              & 72 x 72 x 46   & 3.0 x 3.0 x 3.0 & 2.5 & 204 & 10.2 \\
-REFLECT &  240              & 64 x 64 x 33   & 3.8 x 3.8 x 3.8 & 2.0 & 260 & 8.7 \\
-PARK    &  240 x 240 x 129  & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4 & 149 & 5.96 \\
-LEARN   &  -                & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0 & 196 & 6.5 \\
-\hline
-\end{tabular}
-\caption{
-    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P slices with dimensions M x N.
-}
-\label{table:1}
-\end{table}
-
-Dataset scan parameters and acquisition details are summarized in Table I, and sample and subgroup
-sizes are summarized in Table II. For all datasets, we use the raw (unpreprocessed) scans, and
-perform our own preprocessing (detailed later). Given the complexity of these datasets, we highlight
-below only those aspects relevant to the current paper.
-
-\begin{table}[h!]
-\small
-\centering
-\begin{tabular}{ l l c c c }
-\hline
-\textbf{ID}  & \textbf{Subgroup / Condition}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
-\hline
-PSYCH   &  high\_vigilance    & 12 / 11 & resting-state                    & 2 \\
-        &  low\_vigilance     & 10 / 11 & resting-state                    & 2 \\
-        &  trait\_attend      & 12 / 15 & resting-state                    & 2 \\
-        &  trait\_non-attend  & 10 / 7  & resting-state                    & 2 \\
-        &  task\_attend       & 13 / 15 & resting-state                    & 2 \\
-        &  task\_nonattend    &  9 / 7  & resting-state                    & 2 \\
-\hline
-LEARN   &  task              & 24      & learn image sketches & 16 \\
-        &  rest              & 24      & resting-state                    & 2 \\
-\hline
-REFLECT &  control (no pain) & 20      & resting-state                    & 1 \\
-        &  duloxetine        & 19      & resting-state                    & 1 \\
-        &  placebo           & 37      & resting-state                    & 1 \\
-        &  allpain (duloxetine + placebo) & 56         & resting-state    & 1 \\
-\hline
-PARK    &  Parkinson's       & 17      & ANT                              & 6 \\
-        &  control           & 17      & ANT                              & 6 \\
-\hline
-ECHO    &  summed/interleave & 31      & resting-state                    & 1 \\
-        &  summed/interleave & 31      & imagery/reflection               & 4 \\
-\hline
-\end{tabular}
-\caption{
-    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P slices with dimensions M x N.
-}
-\label{table:2}
-\end{table}
-
-\subsubsection{Dataset 1}
-Dataset 1 (referred to in Fig. and tables with the label ``PSYCH'') has detailed and extensive
-information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
-psychological measurements, and various resting-state fMRI (rs-fMRI) images were obtained. Only the
-whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
-the other included datasets. For the purpose of keeping with the ``attention'' theme of the analyses
-in this paper, three pairings of subgroups were constructed via z-score splits of the various
-sum-scores of scale items involving concepts closely related to attention.
-
-\paragraph{Trait Attention}
-The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
-that queries the extent to which subjects agree particular affective terms have applied to them over
-a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
-the scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
-``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects with a z-score greater / less than zero on this
-sum-score were placed in the high / low ``trait attention'' category
-
-\paragraph{Task Attention}
-The Conjunctive Continuous Performance Task (CCPT;
-\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
-where subjects must react only if they see a red square among other different coloured figures. For
-this paper, a sum score was constructed by subtracting the number of false positive and false
-negative responses from the number of success responses, and subjects with a z-score greater / less
-than zero on this scale were placed in high / low ``task attention'' category.
-
-\paragraph{Vigilance}
-The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
-state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
-from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}). One question assesses
-alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
-``vigilance'' conditions in the same manner as the previous two splits.
-
-\subsubsection{Dataset 2}
-Dataset 2 (referred to in Fig. and tables with the label ``OSTEO'') is detailed in
-\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
-condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
-either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
-also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
-subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
-documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
-complementary to \cite{matharooSpontaneousBackpainAlters2020}.
-
-\subsubsection{Dataset 3}
-Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
-cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
-multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
-involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
-and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
-Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
-different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
-resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
-
-In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
-the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the ``interleave''
-condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
-from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
-t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
-
-This creates a physically-meaningful 4D image that contains the full, original multi-echo fMRI
-time-series data without any alterations. It should be noted that the purpose of these combination
-methods is not to create a 4D volume that allows optimal detection of the purely task-linked BOLD
-signal, but to ensure all scan information was used with minimal additional analytic assumptions.
-
-\subsubsection{Dataset 4}
-Dataset 4 (referred to in Fig. and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
-described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
-in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
-number of repetitions of the Attention Network Task (ANT)
-(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
-task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
-groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
-instead differ by virtue of the presence or absence of PD, any results here serve as a useful
-comparison point for the results from studies that should clearly involve attentional differences.
-This dataset also included extensively preprocessed scans, which were included in analyses to
-examine the effects of preprocessing on the extracted RMT features.
-
-\subsubsection{Dataset 5}
-Dataset 5 (referred to in Fig. and tables with the label ``LEARN'') is documented and described in
-detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
-were available for all subjects, allowing for within-subject comparisons. The complex task involved
-multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
-related by family-resemblance of the components (see Fig. 1 of
-\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
-memory of the various learned features. Scans corresponding to both the initial learning and
-subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of the
-analysis of this paper. Resting state scans were included in the low-attention ``rest'' subgroup.
-
-\subsection{Preprocessing}
-Some consideration should be given to whether it is appropriate to apply standard preprocessing
-steps before extracting RMT features, as preprocessing steps should be developed, tested, and
-employed always with respect to a set of inferential goals and a specific analytic framework.
-However, these preprocessing steps are increasingly recognized to have complex effects on fMRI
-analyses (\cite{parkerBenefitSliceTiming2019}). For example, spatial smoothing has clear benefits
-when employing a general linear model
-(\cite{miklEffectsSpatialSmoothing2008,alakorkkoEffectsSpatialSmoothing2017,worsleyAnalysisFMRITimeSeries1995}),
-but is less motivated in functional network analyses with small ROIs, where the effect of smoothing
-can be considerable
-(\cite{alakorkkoEffectsSpatialSmoothing2017,goelmanMaximizingNegativeCorrelations2014,trianaEffectsSpatialSmoothing2020},
-with these effects interacting complexly with factors like field strength and image resolution
-(\cite{triantafyllouEffectSpatialSmoothing2006}). Similarly, image registration is usually required
-to allow comparisons across (or even within) subjects, but the choice of registration strategy can
-have dramatic effects on functional network analyses (\cite{magalhaesImpactNormalizationSegmentation2014}).
-
-The RMT features extracted here are based on functional correlations rather than ROI activations,
-and are completely insensitive to voxel location (spatial re-arrangement of an image’s voxels does
-not change the eigenvalues of the correlation matrix). Thus RMT features summarize all voxel-wise
-correlations in a whole-brain and location-free manner, which brings into question some of the usual
-motivations for performing spatial smoothing, or registration to identify ROIs. In addition, if
-image dimensions are identical between subjects, the RMT features (by virtue of their definition and
-the unfolding procedure) already allow for direct comparisons. Thus, the usual motivations for
-registration are also no longer clear when extracting RMT features.
-
-Given these particularities, as well as the novelty of the RMT-based methods and the diversity of
-the datasets, we elected to minimize the complexity of the preprocessing pipeline to ensure that
-data processing is kept as standardized as possible. Ultimately, different preprocessing pipelines
-for RMT fMRI analyses can be expected to have different benefits and shortcomings depending on the
-question being investigated, and possibly also the specific RMT features being relied upon, but such
-an investigation is beyond the scope of the current study.
-
-For this paper’s analyses, brain extraction is necessary to limit analyses to the appropriate
-voxels, motion correction is necessary to allow proper brain extraction, and both motion-correction
-and slice-time correction simplify the interpretation of voxel-voxel correlations across voxel
-locations. Whereas physiological noise regression and bias field correction would perhaps be
-appropriate here, there is insufficient data to perform these steps for all datasets (Table II).
-Thus, motion-correction was performed with FSL's MCFLIRT
-(\cite{jenkinsonImprovedOptimizationRobust2002}), slicetime correction with the default method in
-FSL, and brain extraction was performed with BET (\cite{smithFastRobustAutomated2002}). Due to
-differences in the size of the REFLECT data, the higher-resolution resting-state data was
-interpolated to have the same spatial dimensions as the task data.
-
-\subsection{RMT-Derived Features}
-This paper examines features informed by various aspects of RMT, and which can be extracted from
-empirical correlation matrices. To keep with previous studies, we limit our analyses to features
-extracted only from the eigenvalues of the matrices of correlations between all voxel signals.
-
-Two classes of features were examined in this study: those that require unfolding, and those that do
-not. In order to compare the observed eigenvalue distribution to some theoretical predictions, the
-eigenvalues must be unfolded (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}).
-The details and motivation behind the unfolding procedure are well documented in
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process can be viewed as a
-smoothing and rescaling procedure where the originally observed spectrum \(\lambda_1 \le \dots \le
-\lambda_n\) is smoothly mapped to the unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n
-\approx 1\), if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing function is
-not known, and so must be approximated
-(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). Typically, a polynomial is
-chosen (\cite{abul-magdUnfoldingSpectrumChaotic2014}).
-
-A number of statistics (or ``spectral observables''; \cite{mehtaRandomMatrices2004}) can then be
-computed from this unfolded spectrum. In the literature, the two most common points of comparison
-for these unfolded features are typically the features of the Poisson or Gaussian Diagonal Ensemble
-(GDE) and the GOE (see e.g.
-\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}). The
-GOE (or GDE) consists of real, symmetric matrices where entries (or only diagonal entries) are
-normal random variables with zero mean and unit variance. When the empirically observed features are
-close to those of the GOE, this is usually interpreted as indicating a high degree of randomness and
-statistical independence in the data, and when the observed features are closer to those expected in
-the GDE or Poisson case, this is interpreted as indicating a high degree of coherence or dependence
-(\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}).
-
-The Nearest-Neighbours Spacings Distribution (NNSD) is the most basic RMT feature, and is the
-distribution of the spacings (differences) of adjacent unfolded eigenvalues. Normally, the NNSD is
-analysed visually via a histogram of the distribution of spacings, however it is also possible to
-analyse it more formally by fitting the Brody distribution (\cite{brodyStatisticalMeasureRepulsion1973}):
-
-\[
-p(s) = \alpha (1 + \beta) s^\alpha \exp({\beta s^{\alpha+1}}),
-\quad \alpha = \left[ \Gamma \left( \frac{\beta+2}{\beta+1} \right) \right]^{\beta+1}
-\]
-
-where \(\beta\) is free to vary within \([0, 1]\), with \(\beta = 0\)  producing the expected NNSD
-for the GDE case, and \(\beta = 1\) producing the expected NNSD for the GOE case. In this paper, we
-fit, via maximum likelihood, the empirically observed spacings with the Brody distribution, and
-report the Brody parameter, \(\beta\), as an important unfolded feature to investigate.
-
-Whereas the NNSD and Brody parameter characterize the local properties of eigenvalues close on the
-spectrum, the long-range properties relating eigenvalues distant on the spectrum can be
-characterized by both the spectral rigidity and level number variance
-(\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The spectral rigidity,
-\(\Delta_3(L)\), is calculated for any real value \(L\) as:
-
-\[
-\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\]
-
-where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
-\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
-\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
-line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The level
-number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
-(\cite{mehtaRandomMatrices2004}), and is calculated as:
-
-\[
-\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\]
-
-where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
-and \(\langle \cdot \rangle_c\) are as above (\cite{guhrRandommatrixTheoriesQuantum1998a}). Both the
-spectral rigidity and level number variance are also computed and investigated in this study.
-
-In addition to the unfolded features suggested by RMT, there are some potential ``raw'' features that
-do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
-information related to the system, and is also investigated as an RMT-inspired feature.
-
-The distribution of eigenvalues of large rectangular matrices with finite-variance, independent and
-identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP)
-distribution (\cite{marcenkoDISTRIBUTIONEIGENVALUESSETS1967,yaoLargeSampleCovariance2015}):
-
-\[
-\rho(\lambda) =
-\frac{1}{2\pi\sigma^2}
-\frac{\sqrt{(\lambda_+ - \lambda)(\lambda - \lambda_-)}}{r\lambda}, \;
-\lambda_{\pm} = \sigma^2\left(1 \pm \sqrt{\lambda}\right)^2
-\]
-
-where \(\sigma^2\) is the true variance of the distribution of each i.i.d. entry (equal to 1 in the
-case of correlation matrices), and where \(r = N / T\), for matrices of dimension \(N \times T\).
-While real-world data will not, in general, be i.i.d., the extent to which spectra overlap with or
-are well-described by the MP distribution has been used to guide analytical decisions regarding the
-relevance of empirically observed eigenvalues. That is, since the majority of the eigenvalues of
-random correlation matrices will fall within \([\lambda_{-}, \lambda_{+}]\), eigenvalues observed to
-fall within this region are typically considered to be the result of noise
-(\cite{veraartDenoisingDiffusionMRI2016,veraartDiffusionMRINoise2016,ulfarssonDimensionEstimationNoisy2008,macmahonCommunityDetectionCorrelation2015}).
-We examine as a potential feature the proportion of eigenvalues that fall within this range, and
-call this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
-adjustments of \cite{almogUncoveringFunctionalSignature2019} which takes into account the larger
-degree of correlation in empirical data.
-
-\subsection{RMT Feature Extraction}
-The entire 4-dimensional \(m \times n \times n \times p \times t\) scan array for each run,
-excluding voxels with constant signals (which lead to undefined correlations), and excluding voxels
-identified as non-brain via preprocessing, was flattened to a two-dimensional \(m \cdot n \cdot p
-\times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
-value of \(t\) for a scan can be inferred from the number of volumes in Table I. The various RMT
-features mentioned above were then extracted via the empyricalRMT library [64] developed and made
-publicly available as part of this study, as well as via specific code for this paper (which is
-documented and available online (\cite{bergerderekm.DMBergerRandommatrixfmriV02020}).
-
-As both the spectral rigidity and level number variance can be highly sensitive to the choice of
-smoothing procedure used in the unfolding method
-(\cite{abul-magdUnfoldingSpectrumChaotic2014,moralesImprovedUnfoldingDetrending2011,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012}),
-all unfolded features in this study were calculated using polynomials of degree 5, 7, 9, 11, and 13
-as the smoothing technique. As much of the variance in the unfolded spectral statistics can be due
-to the largest eigenvalues acting as anchor points during smoothing
-(\cite{moralesImprovedUnfoldingDetrending2011,abueleninEffectUnfoldingSpectral2012}), all unfolded
-features were also calculated twice (in addition to each polynomial degree) to examine the effects
-of trimming on spectral observables: once with only the largest eigenvalue omitted, and once with
-the largest 20 eigenvalues omitted.
-
-\subsection{Feature Evaluation}
-To evaluate the extracted RMT features, for each possible pairing of subgroups in each dataset, each
-feature was used to predict subject subgroup membership. That is, for each dataset, and for each
-possible pairing of subgroups, all scans belonging to a subgroup were pooled together and treated as
-independent samples of the subgroup of interest. So, for example, in the LEARN dataset, this
-resulted in a comparison of 384 (\(16 \times 4\)) task scans, to 48 (\(2 \times 24\)) rest scans. When it was not
-reasonable to compare scans across sessions, sessions were analyzed separately. This was the case
-for the PSYCH dataset, where psychological measurements, and thus subgroups, differed from session
-to session.
-
-A wide variety of standard, out-of-the-box classifiers available in the Python scikit-learn library
-(\cite{pedregosaScikitlearnMachineLearning2011}) were then employed to gauge the predictive value of
-the RMT features. To allow for a wide range of flexibilities, classifiers employing k-nearest
-neighbours (KNN), with k values in \(\{1, 2, 3, \dots, 10, 20\}\), logistic regression (LR), linear
-discriminant analysis (LDA), and a support-vector machine (SVM) were employed. Default parameter
-values were used when possible (see \cite{bergerderekm.DMBergerRandommatrixfmriV02020} for precise
-function calls used).
-
-The accuracy of each binary classifier was then validated with leave-one-out cross-validation
-(LOOCV) on all datasets, for all combinations of trimming and unfolding choices.
-
-\section{Results}
-
-\subsection{Overview}
-Due to space limitations, we highlight only key results here. Extensive overviews of the full
-results are available in the supplementary materials. Visual summaries of the prediction accuracies
-for all features and analytic choices are presented in Fig. S1--4. Prediction accuracies for only the
-multivariate RMT features are summarized in Fig. S5--7, and the distributions of the scalar features
-are presented in Fig. S8--10. Prediction accuracies for the scalar features across analytic choices
-are summarized in Fig. S11--12.
-
-Numeric (mean) summaries across datasets, broken down by classifier and feature are available in
-Tables S11--18. Complete (unsummarized) tables of all predictions for all combinations of features,
-unfolding degree, and trimming are available as supplementary material in the spreadsheet
-\texttt{prediction\_tables\_fullpre.ods}. Unless otherwise noted, subsequent figures and analyses were
-computed using the preprocessed data: full results for the raw data are available as spreadsheets in
-the supplementary material in \texttt{prediction\_tables.ods}.
-
-With the exception of the LEARNING dataset, the RMT-inspired features all demonstrated some
-predictive utility (Fig. S1-4). For example, there were strong differences in RMT features between
-task and rest conditions for the REFLECT dataset, but the features were of variable predictive
-utility in the PSYCH dataset, sometimes outperforming the naïve classifier, and sometimes not.
-Although the RMT features were useful in the comparison involving duloxetine and controls (no pain)
-in the OSTEO dataset, these features mostly did not aid in classifying between any of the other
-possible pairings of subgroups in this dataset (Fig. S1-4).
-
-Although the performance of classifiers trained on RMT features was variable, there was a tendency
-for the raw and largest eigenvalues to be the most predictively useful features, especially in the
-REFLECT and PARK data in discriminating between Parkinson's and controls (Fig. S5-7. S13; but see
-Discussion).
-
-\subsection{Raw Features}
-As evident from Fig. S8-10, the scalar raw features (largest eigenvalue, noise ratio, and shifted
-noise ratio) varied more within than between subgroups. They were not generally useful for
-classifying conditions (Tables S15-17), with two or three notable exceptions.
-
-The noise ratio (but not shifted noise ratio) differed consistently between the duloxetine and
-control groups in the OSTEO data, allowing for prediction accuracies of up to 72\% compared to 51\%
-for the naive classifier (Fig. S11-12 top-right panel, Table S5). The largest eigenvalue was also
-useful in distinguishing between subjects self-reporting high versus low vigilance across sessions:
-accuracies up to 64\% and 65\% (compared to naïve accuracies of 55\% and 50\%) for the first and
-second sessions, respectively (Table S6). Finally, for the REFLECT data, task and rest were
-perfectly distinguishable with various features and with some classifiers.
-
-The sorted, raw eigenvalues, however, more consistently contained information that allowed some
-separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
-consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
-had the most consistent predictive value for distinguishing individuals with Parkinson's from
-controls, or distinguishing those with osteopathic pain taking duloxetine from controls (``nopain'' in
-Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
-comparing those differing in self-reported vigilance and high vs. low scorers in the attention task
-(Table S7--8).
-
-In all these relationships, normalizing the eigenvalues to the interval \([0,1]\) prior to
-prediction considerably improved prediction accuracy (Tables S9-12).
-
-\paragraph{Role of Classifer}
-The raw eigenvalues generally had the most predictive utility when employing either an SVM
-classifier or KNN classifier with \(k > 3\) (Table S11, first row). When the raw eigenvalues were
-first min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
-predicting subgroup memberships (Table S12), however without normalization, KNN with \(k = 8\) or
-\(k = 9\) generally best predicted subgroup membership across datasets (Table S11).
-
-\subsection{Unfolded Features}
-The Brody parameter rarely resulted in predictions exceeding naïve accuracies in these comparisons:
-for most unfoldings and comparisons, the Brody parameter was either constant and equal for all
-subjects (e.g. exactly 0.999 or 0.0001) or deviated from these values only for a small number of
-subjects. This is because the observed NNSDs were usually well outside of the GOE-GDE bounds.
-
-The predictive utility of the unfolded RMT features varied considerably across datasets and
-classifiers. However, some results were robust: for example, the spectral rigidity and level number
-variance strongly differentiated controls from individuals prescribed duloxetine for osteopathic
-pain, and this differentiation persisted regardless of unfolding degree or trimming (Fig. 5-8, top-right panels).
-
-Despite the insensitivity to unfolding in this dataset, there was still nontrivial variability in
-accuracy across classifiers when classifying duloxetine vs controls, with the less flexible
-classifiers (logistic regression, SVM, and KNN with k=10, 20) generally performing best (Tables
-S1-2).
-
-For other datasets, the predictive utility of the unfolded features was highly dependent on the
-unfolding degree, and sometimes trim proportions. These differences are illustrated in detail for
-the Parkinson's data in Fig. S14. For example, with minimal trimming, subgroup overlap in spectral
-rigidity was minimized for unfolding degrees 5 and 9 (Fig. S14, first row), but with greater
-trimming, this was not the case (Fig. S14, second row). For the level number variance with minimal
-trimming, subgroup overlap appeared smaller for lower unfolding degrees (Fig. S14, third row),
-however, with a greater degree of trimming, separation was clearer for slightly larger degrees (Fig.
-S14, fourth row).
-
-Across datasets, the unfolding degree, compared to trimming, was a much more significant source of
-variance in the overall predictive utility of the unfolded features (e.g. note the minimal changes
-from trimming in Fig. S5 and S6, compared to the overall differences in prediction patterns from
-differing choices of unfolding degree in Fig. S3 and S4). However, neither the choice of unfolding
-degree nor amount of trimming had any easily summarizable effects on the predictive value of the
-various RMT metrics: the optimal trimming and unfolding degree was different for each dataset and
-comparison.
-
-\paragraph{Role of Classifier}. Across all datasets, unfoldings, and trimmings, the spectral rigidity was
-generally best able to separate between subgroups when either k was large (10 or 20) in KNN, or
-Logistic Regression or SVM was employed (Table S13). The pattern was similar for the level number
-variance, although the predictive utility of the level number variance was, overall, slightly more
-variable than the spectral rigidity (Table S4).
-
-\section{Discussion}
-In this study, certain features inspired by RMT and derived from the sorted eigenvalues of the full,
-whole-brain voxelwise correlation matrix were found to contain information with the potential to
-meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
-from previous studies
-(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020})
-this suggests RMT may be a useful analytic and theoretical tool for understanding functional
-connectivity.
-
-Although the choice of unfolding and trimming often had nontrivial effects on the overall shape and
-values of the various spectral observables, and thus the predictive utility, in cases where there
-was clear predictive utility, there was some evidence that the unfolding and trimming procedures did
-not reverse the relations between subgroups (Fig. 1-3, S14). For example, regardless of the
-unfolding methodology, it was always the case that control subjects had spectral observables
-farthest from those of the GOE (Fig. 1-3) compared to those given duloxetine for osteopathic pain,
-whereas when comparing subject's with Parkinson's to controls, Parkinson's subjects consistently
-showed the greatest deviation from the GOE (Fig. S14).
-
-In particular, for some psychological phenomena examined in this study (specifically, the
-resting-state, voxelwise functional connectivity of those prescribed duloxetine for osteopathic pain
-versus the resting-state, voxelwise functional connectivity of controls), the spectral rigidity and
-level number variance contained information that was valuable in differentiating between subgroups.
-Across classifiers for this subgroup comparison, maximum mean LOOCV classification accuracies
-improved upon the 51\% accuracy of the naïve classifier by 11-26\% for the spectral rigidity and
-16-26\% for the level number variance (Tables S1-2).
-
-In the above case, these unfolded features allowed for significantly improved predictions over those
-using the raw (not unfolded) eigenvalues (Table S3). It is possible that some of these predictive
-improvements are because both the unfolding procedure and calculation of the spectral observables
-both smooth and transform all data to common scales. For example,  rescaling the raw eigenvalues to
-the interval [0, 1] also resulted in comparable prediction accuracies (Table S4).
-
-For comparison, for small datasets of rs- or task-fMRI data, binary classification using whole-brain
-features is challenging. For subject level classification, state of the art deep learning methods
-improve upon guessing by 17\% for autism (\cite{bengs4DSpatioTemporalDeep2020}) or 3-30\% for ADHD
-(\cite{riazDeepFMRIEndtoendDeep2020}), 16\% for severe depression
-(\cite{ramasubbuAccuracyAutomatedClassification2016}), and 23\% for obsessive compulsive disorder
-(\cite{takagiNeuralMarkerObsessiveCompulsive2017}). Manual feature engineering with more separable
-conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
-(\cite{duHighClassificationAccuracy2012}), and with very large data, sophisticated custom feature
-extraction methods can achieve near perfect accuracies at classifying task vs. rest
-(\cite{zhangCharacterizingDifferentiatingTaskbased2016}). However, for functional connectivity data
-and with classical machine learning algorithms (such as SVM) we in general only expect large
-prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
-strongly separated (e.g. Cohen’s d > 1.0; \cite{dansereauStatisticalPowerPrediction2017}).
-
-Nevertheless, the features extracted in this study are somewhat unique in that they are small,
-spatially-insensitive summaries of the complete fMRI data. In particular, reducing the functional
-connectivity to just the largest 20 eigenvalues results in predictions comparable to using the
-entire set of eigenvalues (Tables S12, S18). That such a reduction can lead to accuracies comparable
-to other methods suggests that the system eigenvalues have utility beyond information reduction, and
-should perhaps be regarded themselves as signatures of functional connectivity patterns.
-
-\subsection{Confounds}
-While some of the RMT features allowed some classifiers to nearly perfectly separate conditions for
-the REFLECT dataset, this should almost certainly be regarded as an artifact of the differing scan
-parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
-correlation matrix is primarily determined by such differences in dimensions
-(\cite{yinLimitLargestEigenvalue1988}), as are the noise ratios, which include these dimensions in
-their definitions (\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}). Likewise,
-similar strong separation was observed in the PARK data when comparing the minimally pre-processed
-to fully pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps
-(spike removal, spatial smoothing, co-registration to larger T1 image;
-\cite{madhyasthaDynamicConnectivityRest2015}) which will also almost certainly alter the size of the
-largest eigenvalue.
-
-When there was predictive utility in the unfolded features,	 this utility was dependent on the
-unfolding procedure and choice of classifier. For example, the highest nontrivial LOOCV mean
-accuracy across all datasets was 0.79, corresponding to the duloxetine vs. controls subgroup
-comparison in the OSTEO dataset, whereas a poor choice of classifier in this same comparison could
-lead to prediction accuracies of 0.51, equivalent to the accuracy of a naive classifier (Table S2).
-
-There was variability in the predictive utility of the unfolding features, and this is likely
-because the unfolded features varied considerably based on the degree of the smoothing polynomial
-and extent of trimming. For example, the spread of the spectral rigidity and level number variance
-curves are shown for all subjects in the PARK data in Fig. S14. Clearly, both the curve shape,
-average magnitude, and extent of overlap is strongly dependent on the unfolding and trimming
-procedure.
-
-The unidimensional features, such as the largest eigenvalue or those based on the Marchenko-Pastur
-distribution, were most useful in predicting subgroup membership in trivial cases, such as in the
-REFLECT data, where the underlying scan dimensions differed between comparison groups, or in the
-PARK data, where the largest eigenvalue clearly distinguished the degree of preprocessing (Fig.
-S11-12). As these differences are due to methods, and not neuropsychological phenomena, they are
-perhaps of limited interest. To the extent that these features are scalar summaries of much
-higher-dimensional, 4D data, the limited utility of the scalar features should not be particularly
-surprising—it would be quite extraordinary if different neuropsychological states could be reliably
-separated based on such a reduction.
-
-However, that there are not consistently strong differences in these unidimensional features between
-subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
-noise-reduction procedures (such as
-\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}), is unlikely to introduce bias
-into subsequent analyses. It perhaps also suggests these scalar features might be useful for quality
-control within a dataset: if there are differences between subgroups or individuals on one of the
-individual RMT features, then perhaps this is cause for more careful investigation.
-
-The unfolded features (spectral rigidity and level number variance) were more promising, however
-their sensitivity to the unfolding procedure means that great caution must be employed if they are
-to be used as features in future research. Ideally, the utility of these unfolded features (and all
-RMT features provided in this software release) need to be rigorously validated with wholly
-independent datasets addressing similar conditions. In addition, as simply using the raw, min-max
-normalized eigenvalues generally resulted in similar classification accuracies as the unfolded
-features, further study will be required to justify the use of these more complicated features.
-
-\subsection{Preprocessing}
-Across datasets, the features extracted from the minimally preprocessed images resulted in slightly
-lower prediction accuracies than the raw images for the unfolded features (rigidity, level
-variance), and slightly improved accuracies for some of the raw features (all eigenvalues, largest
-eigenvalue(s), noise ratio). However, these differences were quite small relative to the performance
-differences between datasets or between classifiers1.
-
-Similar patterns were observed in the PARK dataset. Fig. S5-7 show that more extensive preprocessing
-resulted in greater prediction accuracies for raw and largest eigenvalue(s), but decreased
-accuracies for the unfolded features. Given a particular level of preprocessing, however, it is
-clear from the spread of prediction accuracies that there was still substantial overall variance in
-prediction accuracy due to the choices of unfolding and classifier.
-
-\subsection{Comparison with Previous Studies}
-
-We generally find different results from previous studies
-(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
-which have used the spectral rigidity or level number variance to differentiate between attention
-conditions. As \cite{sebaRandomMatrixAnalysis2003} extracted rigidity curves from the correlations
-of electroencephalographic (EEG) signals, and \cite{wangRandomMatrixTheory2016} and
-\cite{matharooSpontaneousBackpainAlters2020} extracted spectral observables from the correlations of
-the mean signals of anatomical ROIs, these differences are to be expected. A single voxel signal
-amalgamates information from far fewer neurons than an EEG or mean ROI signal, and it is possible
-that any broad connections between RMT spectral observables and attention arise only at higher
-levels of organization, or after certain pre-processing steps (e.g. spatial smoothing) not performed
-in this study.
-
-We do find some modest evidence of greater overall correlation in high attention conditions: there
-was a slight tendency for the largest eigenvalue to be smaller when attention is low (Fig. S8).
-However, this is only suggestive, and future studies will be needed to determine how and under what
-conditions these RMT features reliably relate to neuropsychological phenomena.
-
-\subsection{Limitations}
-
-To make analyses across datasets at least minimally comparable, only limited preprocessing (motion
-and slicetime correction, skull-stripping) was performed in this study. In particular, in the
-absence of physiological noise regression, there must be caution in assuming any of the observed
-patterns are uniquely due to changes in functional connectivity. This is perhaps especially notable
-as the RMT features were most consistently capable of distinguishing between individuals taking
-duloxetine versus controls, or between controls and individuals with Parkinson's. It seems much more
-likely that physiological confounds could be involved in these comparisons in contrast to other
-comparisons (such as task-state versus resting-state, or in the differing levels of attention in the
-PSYCH dataset).
-
-The correlation matrices calculated for each run are computed from the full time-series data. No
-attempt was made to control for stimulus presentation times, under the assumption that the broad,
-whole-brain summarizing nature of the RMT features should be able to detect broad, overall
-differences between subgroups or conditions. However, differences were not consistently observed
-between conditions across all datasets. It is possible more careful selection of regions of the
-time-series in task conditions will be necessary in order for RMT features to reflect the
-differences between these conditions.
-
-While the extraction of eigenvalues from the correlations of all voxels has the advantage of not
-relying on a specific parcellation scheme, and does not impose anatomical or locality assumptions
-that ROI-based analyses do, it is also possible that this limited the utility of the RMT features.
-In particular, we found the spectral observables (NNSD, spectral rigidity, level number variance) to
-frequently fall outside the GOE/GDE bounds that have typically been observed in previous studies.
-Future studies should examine the sensitivity of these RMT features to the level of locality of
-analyses, or choice of size and location of ROIs.
-
-\section{Conclusion}
-
-\paragraph{Declarations}
-Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
-Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
-Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
-Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
-(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
-R0168020), and a St. Francis Xavier University UCR grant to JL.
-
-\section{References}
-\printbibliography{}
+\submitto{\PMB}
+\maketitle
 

### 2022-10-06 paper/bookdown/rmt.Rmd
update

--- 2021-09-12 paper/latex/rmt.tex
@@
-\input{preamble.tex}
+---
+title: "Random Matrix Theory Tools for the Analysis of Functional Magnetic-Resonance Imaging Examinations"
+fontsize: 10pt
+urlcolor: blue
+linkcolor: blue
+# mainfont: "Garamond Premier Pro"
+# biblio-style: "apalike"
+bibliography: "rmt.bib"  # exported from Zotero with better BibLaTeX extension
+link-citations: true
+output:
+    # bookdown::pdf_document2:
+    bookdown::pdf_book:
+        latex_engine: xelatex
+        toc: false
+        # toc_depth: 2
+        number_sections: false
+        highlight: tango
+        # https://bookdown.org/yihui/rmarkdown-cookbook/figure-placement.html
+        # unicode-math needed for bold italic math
+        # (https://tex.stackexchange.com/questions/431013/error-extended-mathchar-used-as-mathchar-when-using-bm)
+        extra_dependencies: ["float", "unicode-math"]
+        mathjax: default
+        includes:
+            in_header: preamble.tex
+        split_bib: true
+        fig_width: 6
+        fig_height: 4
+        link-citations: true
+citation-style: ieee.csl # https://www.zotero.org/styles
+---
 
-\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
-\author{Derek M. Berger \and Gurpreet S. Matharoo \and Jacob Levman}
-\date{\today}
+# Abstract
 
-\begin{document}
-\maketitle
-
-\begin{abstract}
 Previous studies have investigated the potential of using analytic techniques from Random Matrix
 Theory (RMT) to investigate magnetic resonance imaging (MRI) data. We assess the potential
-    application of RMT-based features for the analysis of functional MRI (fMRI) across diverse
-    datasets. As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain
-    features extracted from voxel-wise functional connectivity to contain information useful for
-    classifying between various psychological processes, (2) assess these features’
-    predictive—rather than explanatory—value, (3) investigate the effect of varying RMT analysis
-    methods on the robustness of study findings, and (4) make general-purpose code publicly
-    available for users to extract these features from a wide variety of data. We find preliminary
-    evidence suggesting that RMT-inspired features may have unique potential in analyses of fMRI
-    functional connectivity.
+application of RMT-based features for the analysis of functional MRI (fMRI) across diverse datasets.
+As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain features extracted
+from voxel-wise functional connectivity to contain information useful for classifying between
+various psychological processes, (2) assess these features’ predictive—rather than
+explanatory—value, (3) investigate the effect of varying RMT analysis methods on the robustness of
+study findings, and (4) make general-purpose code publicly available for users to extract these
+features from a wide variety of data. We find preliminary evidence suggesting that RMT-inspired
+features may have unique potential in analyses of fMRI functional connectivity.
 
-    \subparagraph{Index Terms:} feature extraction, functional connectivity, functional magnetic
-    resonance imaging, random matrix theory, voxelwise connectivity
+## Index Terms / Keywords
 
-    \subparagraph{Authors}
+feature extraction, functional connectivity, functional magnetic resonance imaging, random matrix
+theory, voxelwise connectivity
+
+## Funding Info
+
+Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
+Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
+Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
+Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
+(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
+R0168020), and a St. Francis Xavier University UCR grant to JL.
+
+## Author Info
+
 Derek Berger is with the department of Computer Science, St. Francis Xavier University, Antigonish
@@
 
-    Jacob Levman is with the Department of Computer Science, St. Francis Xavier University,
-    Antigonish NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).
-\end{abstract}
+Jacob Levman is with the Department of Computer Science, St. Francis Xavier University, Antigonish
+NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).
 
-\section{Introduction}
+\newpage
+
+# Introduction
 
@@
 Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
-complex functional connectivity networks are ubiquitous
-(\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}).
-This complexity suggests fMRI is a candidate to be studied using Random Matrix Theory (RMT), a set
-of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
-physics (\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}). In various
+complex functional connectivity networks are ubiquitous [1]–[4]. This complexity suggests fMRI is a
+candidate to be studied using Random Matrix Theory (RMT), a set of mathematical tools originally
+developed some 50 years ago to solve complex problems in nuclear physics [5], [6]. In various
 physical systems where nuclei can have different possible states, precisely modeling the entirety of
@@
 
-\begin{sloppypar}
 These universal properties have been observed in diverse phenomena. In small-scale systems, RMT
 universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
-disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has been applied to atmospheric
-physics (\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations \cite{plerouRandomMatrixApproach2002} , social networks \cite{jalanUncoveringRandomnessSuccess2014}, random networks \cite{bandyopadhyayUniversalityComplexNetworks2007},
-network-formation in liquids (\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009})  and amorphous clusters (\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}).
-Within biological systems,
-RMT has also been used to successfully model aspects of  amino acid functional relationships (\cite{bhadolaTargetingFunctionalMotifs2016}),
-synchrony in epileptic seizures (\cite{osorioPhasesynchronizationRandommatrixBased2011}), and in protein-protein interactions both in different species (\cite{agrawalQuantifyingRandomnessProtein2014}) and breast cancer (\cite{raiRandomnessPreservedPatterns2015}).
-\end{sloppypar}
+disordered mesoscopic systems [5]–[11], and at larger scales, RMT has been applied to atmospheric
+physics [12], stock cross-correlations [13], social networks [14], random networks [15],
+network-formation in liquids [16], [17] and amorphous clusters [18]–[21]. Within biological systems,
+RMT has also been used to successfully model aspects of  amino acid functional relationships [22],
+synchrony in epileptic seizures [23], and in protein-protein interactions both in different species
+[24] and breast cancer [25].
 
@@
 demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) (\cite{sebaRandomMatrixAnalysis2003}).
-Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
-fMRI data (\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}). RMT has also
+resemble those of the Gaussian Orthogonal Ensemble (GOE) [26].  Somewhat more recently, RMT was used
+to evaluate the quality of whole brain features extracted from fMRI data [27], [28]. RMT has also
 been used in diffusion MRI to aid in the selection of the number of components to employ in
-principal-component reduction analysis and denoising
-(\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}).
+principal-component reduction analysis and denoising [29]–[31].
 
 Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states (\cite{wangSpectralPropertiesTemporal2015}), between
-subjects with and without ADHD (\cite{wangRandomMatrixTheory2016}), and between pain and non-pain
-states (\cite{matharooSpontaneousBackpainAlters2020}). Across these three studies, the spectra of
-resting or low-attention states exhibited properties closest to the GOE.
+differences between rest and task states [32], between subjects with and without ADHD [33], and
+between pain and non-pain states [34]. Across these three studies, the spectra of resting or
+low-attention states exhibited properties closest to the GOE.
 
@@
 
-\section{Materials and Methods}
+# Materials and Methods
 
-\subsection{Datasets}
+## Datasets
 
@@
 
-However, since previous studies employing RMT for feature extraction
-(\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}) have partially interpreted
-their findings with respect to attentional processes, we have elected to use datasets that involve
-attention in some way. This includes measures of attention or vigilance, or comparisons between a
-condition involving more attention and focus (e.g. a task state) or conditions which should involve
-considerably less attention or focus (e.g. resting-state).
+However, since previous studies employing RMT for feature extraction [33], [34] have partially
+interpreted their findings with respect to attentional processes, we have elected to use datasets
+that involve attention in some way. This includes measures of attention or vigilance, or comparisons
+between a condition involving more attention and focus (e.g. a task state) or conditions which
+should involve considerably less attention or focus (e.g. resting-state).
 
-\begin{center}
-\texttt{\textbf{INSERT TABLE 1 HERE}}
-\end{center}
+```{r , echo=FALSE, results='asis'}
+  cat(' Table: (\\#tab:scan-params) Summary of Scan Parameters.
+ID = Identifier for paper.
+FOV = Field of View.
+TR = Time of Repetition (seconds).
+Time = total duration (minutes) of each scan.
+Dimensions listed as M × N × P, indicate P slices with dimensions M × N.
 
-Dataset scan parameters and acquisition details are summarized in Table I, and sample and subgroup
-sizes are summarized in Table II. For all datasets, we use the raw (unpreprocessed) scans, and
-perform our own preprocessing (detailed later). Given the complexity of these datasets, we highlight
-below only those aspects relevant to the current paper.
+  |   ID   |     FOV (mm) | Dimensions | Voxel Size (mm) |  TR  | Volumes | Time (m) |
+  |:------:|:------------:|:-----------:|:---------------:|:----:|:-------:|:--------:|
+  | PSYCH  | 192          | 128×128×70 | 1.5×1.5×1.5     | 3.0  |     300 |  15     |
+  | OSTEO  | -            | 64×64×36   | 3.4×3.4×3.0     | 2.5  |     300 |  12.5   |
+  | REFLECT| 210          | 72×72×46   | 3.0×3.0×3.0     | 3.0  |     204 |  10.2   |
+  |        | 240          | 64×64×33   | 3.8×3.8×3.8     | 2.0  |     260 |  8.7    |
+  | PARK   | 240×240×129  | 80×80×43   | 3.0×3.0×3.0     | 2.4  |     149 |  5.96   |
+  | LEARN  |  -           | 64×64×36   | 3.0×3.0×3.0     | 2.0  |     195 |  6.5    |')
 
-\subsubsection{Dataset 1}
-Dataset 1 (referred to in Fig. and tables with the label "PSYCH") has detailed and extensive
-information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
-psychological measurements, and various resting-state fMRI (rs-fMRI) images were obtained. Only the
-whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
-the other included datasets. For the purpose of keeping with the "attention" theme of the analyses
-in this paper, three pairings of subgroups were constructed via z-score splits of the various
-sum-scores of scale items involving concepts closely related to attention.
+```
 
-\paragraph{Trait Attention}
-The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
-that queries the extent to which subjects agree particular affective terms have applied to them over
-a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
-the scale items: "attentive", "alert", "interested", "concentrating", and reverse-scored items
-"sluggish", "tired", "sleepy", "drowsy". Subjects with a z-score greater / less than zero on this
-sum-score were placed in the high / low "trait attention" category
+## A Simple Linear MLP
 
-\paragraph{Task Attention}
-The Conjunctive Continuous Performance Task (CCPT;
-\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
-where subjects must react only if they see a red square among other different coloured figures. For
-this paper, a sum score was constructed by subtracting the number of false positive and false
-negative responses from the number of success responses, and subjects with a z-score greater / less
-than zero on this scale were placed in high / low "task attention" category.
+A citation to [@abueleninSpectralUnfoldingChaotic2018a;
+@abul-magdModellingHighwaytrafficHeadway2007; @alakorkkoEffectsSpatialSmoothing2017].
 
-\paragraph{Vigilance}
-The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
-state, e.g. "my thoughts involved other people", "I thought about something positive" (translated
-from German) \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
-alertness ("I was fully awake" / "war ich vollkommen wach") and was used to create high / low
-"vigilance" conditions in the same manner as the previous two splits.
+As a concrete example, imagine we are build a simple ANN from $\mathbbm{R}$ to $\mathbbm{R}^2$ by
+way of two larger intermediate layers. We might have for example
 
-\subsubsection{Dataset 2}
-Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in
-\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls ("nopain"
-condition) and individuals with knee osteoarthritis ("pain" condition) treated for 2 weeks with
-either placebo ("pain" condition) or duloxetine ("duloxetine" condition). An additional grouping was
-also created with both the placebo and duloxetine data ("allpain" condition), allowing for six
-subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
-documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
-complementary to \cite{matharooSpontaneousBackpainAlters2020}.
+```python
+ann = Sequential([
+    Linear(1, 2),
+    Linear(2, 1),
+])
+```
 
-\subsubsection{Dataset 3}
-Dataset 3  (referred to here with the label "REFLECT", due to the "reflective” nature of the
-cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
-multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
-involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
-and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
-Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
-different tasks were simply treated generically as "attending", in contrast to the non-attentive
-resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
+which learns the weights
 
-In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
-the "summed" condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the "interleave"
-condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
-from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
-t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
+$$
+\begin{aligned}
+\mathbf{W}_1 &= \begin{bmatrix} w_1 & w_2 \end{bmatrix} & \text{layer 1: matrix of shape (1, 2), output shape (1, 2)} \\
+\mathbf{W}_2 &= \begin{bmatrix} w_3 \\ w_4 \end{bmatrix}& \text{layer 2: matrix of shape (2, 1), output shape (1, 1)} \\
+\end{aligned}
+$$
 
-This creates a physically-meaningful 4D image that contains the full, original multi-echo fMRI
-time-series data without any alterations. It should be noted that the purpose of these combination
-methods is not to create a 4D volume that allows optimal detection of the purely task-linked BOLD
-signal, but to ensure all scan information was used with minimal additional analytic assumptions.
+Assuming ReLUs and no biases, the entire equation for this network is
+$y = \mathbf{W}_2 \cdot \text{ReLU}(\mathbf{W}_1 x) + b$ (final layer should always have a bias), that is:
 
-\subsubsection{Dataset 4}
-Dataset 4 (referred to in Fig. and tables with the label "PARK" or "PARKINSONS") is documented and
-described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
-in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
-number of repetitions of the Attention Network Task (ANT)
-(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
-task demands high attention, it is not as straightforward to create "high" or "low" attention
-groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
-instead differ by virtue of the presence or absence of PD, any results here serve as a useful
-comparison point for the results from studies that should clearly involve attentional differences.
-This dataset also included extensively preprocessed scans, which were included in analyses to
-examine the effects of preprocessing on the extracted RMT features.
+$$
+\begin{aligned}
+\mathbf{W}_1 x & = \begin{bmatrix} w_1 \cdot x & \;\; w_2 \cdot x \end{bmatrix} \\
+             y & = \begin{bmatrix}  w_3 \cdot \text{ReLU}(w_1 \cdot x) + w_4 \cdot \text{ReLU}(w_2 \cdot x) + b \end{bmatrix} \\
+\end{aligned}
+$$
 
-\subsubsection{Dataset 5}
-Dataset 5 (referred to in Fig. and tables with the label "LEARN") is documented and described in
-detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
-were available for all subjects, allowing for within-subject comparisons. The complex task involved
-multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
-related by family-resemblance of the components (see Fig. 1 of
-\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
-memory of the various learned features. Scans corresponding to both the initial learning and
-subsequent testing phase were included in the high-attention "task" subgroup for the purposes of the
-analysis of this paper. Resting state scans were included in the low-attention "rest" subgroup.
+This in fact reduces to:
 
-\subsection{Preprocessing}
-Some consideration should be given to whether it is appropriate to apply standard preprocessing
-steps before extracting RMT features, as preprocessing steps should be developed, tested, and
-employed always with respect to a set of inferential goals and a specific analytic framework.
-However, these preprocessing steps are increasingly recognized to have complex effects on fMRI
-analyses (\cite{parkerBenefitSliceTiming2019}). For example, spatial smoothing has clear benefits
-when employing a general linear model
-(\cite{miklEffectsSpatialSmoothing2008,alakorkkoEffectsSpatialSmoothing2017,worsleyAnalysisFMRITimeSeries1995}),
-but is less motivated in functional network analyses with small ROIs, where the effect of smoothing
-can be considerable
-(\cite{alakorkkoEffectsSpatialSmoothing2017,goelmanMaximizingNegativeCorrelations2014,trianaEffectsSpatialSmoothing2020},
-with these effects interacting complexly with factors like field strength and image resolution
-(\cite{triantafyllouEffectSpatialSmoothing2006}). Similarly, image registration is usually required
-to allow comparisons across (or even within) subjects, but the choice of registration strategy can
-have dramatic effects on functional network analyses (\cite{magalhaesImpactNormalizationSegmentation2014}).
+$$
+y = \begin{cases}
+w_1 w_3 x + w_2 w_4 x + b  & \text{ if }  sgn(x) =   sgn(w_1) \text{ and } sgn(x)   = sgn(w_2) \\
+w_1 w_3 x + b              & \text{ if }  sgn(x) =   sgn(w_1) \text{ and } sgn(x) \ne sgn(w_2) \\
+w_2 w_4 x + b              & \text{ if }  sgn(x) \ne sgn(w_1) \text{ and } sgn(x)   = sgn(w_2) \\
+b                          & x = 0 \\
+\end{cases}
+$$
 
-The RMT features extracted here are based on functional correlations rather than ROI activations,
-and are completely insensitive to voxel location (spatial re-arrangement of an image’s voxels does
-not change the eigenvalues of the correlation matrix). Thus RMT features summarize all voxel-wise
-correlations in a whole-brain and location-free manner, which brings into question some of the usual
-motivations for performing spatial smoothing, or registration to identify ROIs. In addition, if
-image dimensions are identical between subjects, the RMT features (by virtue of their definition and
-the unfolding procedure) already allow for direct comparisons. Thus, the usual motivations for
-registration are also no longer clear when extracting RMT features.
+In particular, without loss of generality we can take $w_1 < 0$ and $w_2 > 0$ and reduce this to:
 
-Given these particularities, as well as the novelty of the RMT-based methods and the diversity of
-the datasets, we elected to minimize the complexity of the preprocessing pipeline to ensure that
-data processing is kept as standardized as possible. Ultimately, different preprocessing pipelines
-for RMT fMRI analyses can be expected to have different benefits and shortcomings depending on the
-question being investigated, and possibly also the specific RMT features being relied upon, but such
-an investigation is beyond the scope of the current study.
+$$
+y = \begin{cases}
+w_2 w_4 x + b  & \text{ if }  x \ge 0 \\
+w_1 w_3 x + b  & \text{ if }  x  < 0 \\
+\end{cases}
+$$
 
-For this paper’s analyses, brain extraction is necessary to limit analyses to the appropriate
-voxels, motion correction is necessary to allow proper brain extraction, and both motion-correction
-and slice-time correction simplify the interpretation of voxel-voxel correlations across voxel
-locations. Whereas physiological noise regression and bias field correction would perhaps be
-appropriate here, there is insufficient data to perform these steps for all datasets (Table II).
-Thus, motion-correction was performed with FSL's MCFLIRT [56], slicetime correction with the default
-method in FSL, and brain extraction was performed with BET [57]. Due to differences in the size of
-the REFLECT data, the higher-resolution resting-state data was interpolated to have the same spatial
-dimensions as the task data.
-
-\subsection{RMT-Derived Features}
-This paper examines features informed by various aspects of RMT, and which can be extracted from
-empirical correlation matrices. To keep with previous studies, we limit our analyses to features
-extracted only from the eigenvalues of the matrices of correlations between all voxel signals.
-
-Two classes of features were examined in this study: those that require unfolding, and those that do
-not. In order to compare the observed eigenvalue distribution to some theoretical predictions, the
-eigenvalues must be unfolded [5], [6]. The details and motivation behind the unfolding procedure are
-well documented in [5]. Practically, the unfolding process can be viewed as a smoothing and
-rescaling procedure where the originally observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is
-smoothly mapped to the unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
-if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing function is not known, and
-so must be approximated [5], [6]. Typically, a polynomial is chosen [58].
-
-A number of statistics (or "spectral observables" [6]) can then be computed from this unfolded
-spectrum. In the literature, the two most common points of comparison for these unfolded features
-are typically the features of the Poisson or Gaussian Diagonal Ensemble (GDE) and the GOE (see e.g.
-[12], [14], [15], [24]–[26], [32], [33]). The GOE (or GDE) consists of real, symmetric matrices
-where entries (or only diagonal entries) are normal random variables with zero mean and unit
-variance. When the empirically observed features are close to those of the GOE, this is usually
-interpreted as indicating a high degree of randomness and statistical independence in the data, and
-when the observed features are closer to those expected in the GDE or Poisson case, this is
-interpreted as indicating a high degree of coherence or dependence [12], [14], [15], [24]–[26],
-[32]–[34].
-
-The Nearest-Neighbours Spacings Distribution (NNSD) is the most basic RMT feature, and is the
-distribution of the spacings (differences) of adjacent unfolded eigenvalues. Normally, the NNSD is
-analysed visually via a histogram of the distribution of spacings, however it is also possible to
-analyse it more formally by fitting the Brody distribution [59]:
-
-\[
-p(s) = \alpha (1 + \beta) s^\alpha \exp({\beta s^{\alpha+1}}),
-\quad \alpha = \left[ \Gamma \left( \frac{\beta+2}{\beta+1} \right) \right]^{\beta+1}
-\]
-
-where \(\beta\) is free to vary within \([0, 1]\), with \(\beta = 0\)  producing the expected NNSD
-for the GDE case, and \(\beta = 1\) producing the expected NNSD for the GOE case. In this paper, we
-fit, via maximum likelihood, the empirically observed spacings with the Brody distribution, and
-report the Brody parameter, \(\beta\), as an important unfolded feature to investigate.
-
-Whereas the NNSD and Brody parameter characterize the local properties of eigenvalues close on the
-spectrum, the long-range properties relating eigenvalues distant on the spectrum can be
-characterized by both the spectral rigidity and level number variance [5], [6]. The spectral
-rigidity, \(\Delta_3(L)\), is calculated for any real value \(L\) as:
-
-\[
-\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\]
-
-where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
-\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
-\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
-line to \(\eta(\lambda)\) on \([c, c+L]\) [5]. The level number variance, \(\Sigma^2(L)\), is
-closely related to the spectral rigidity [6], and is calculated as:
-
-\[
-\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\]
-
-where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
-and \(\langle \cdot \rangle_c\) are as above [5]. Both the spectral rigidity and level number
-variance are also computed and investigated in this study.
-
-In addition to the unfolded features suggested by RMT, there are some potential "raw" features that
-do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
-information related to the system, and is also investigated as an RMT-inspired feature.
-
-The distribution of eigenvalues of large rectangular matrices with finite-variance, independent and
-identically distributed (\textit{i.i.d.}) entries also converges to the Marchenko-Pastur (MP) distribution
-[60], [61]:
-
-\[
-\rho(\lambda) =
-\frac{1}{2\pi\sigma^2}
-\frac{\sqrt{(\lambda_+ - \lambda)(\lambda - \lambda_-)}}{r\lambda}, \;
-\lambda_{\pm} = \sigma^2\left(1 \pm \sqrt{\lambda}\right)^2
-\]
-
-where \(\sigma^2\) is the true variance of the distribution of each i.i.d. entry (equal to 1 in the
-case of correlation matrices), and where \(r = N / T\), for matrices of dimension \(N \times T\).
-While real-world data will not, in general, be i.i.d., the extent to which spectra overlap with or
-are well-described by the MP distribution has been used to guide analytical decisions regarding the
-relevance of empirically observed eigenvalues. That is, since the majority of the eigenvalues of
-random correlation matrices will fall within \([\lambda_{-}, \lambda_{+}]\), eigenvalues observed to
-fall within this region are typically considered to be the result of noise [29]–[31], [62]. We
-examine as a potential feature the proportion of eigenvalues that fall within this range, and call
-this the noise ratio. In addition, we calculate a broader, shifted noise ratio, based on the
-adjustments of [63] which takes into account the larger degree of correlation in empirical data.
-
-\subsection{RMT Feature Extraction}
-The entire 4-dimensional \(m \times n \times n \times p \times t\) scan array for each run,
-excluding voxels with constant signals (which lead to undefined correlations), and excluding voxels
-identified as non-brain via preprocessing, was flattened to a two-dimensional \(m \cdot n \cdot p
-\times t\) matrix, and then the \(t - 1\) eigenvalues of that correlation matrix were computed. The
-value of \(t\) for a scan can be inferred from the number of volumes in Table I. The various RMT
-features mentioned above were then extracted via the empyricalRMT library [64] developed and made
-publicly available as part of this study, as well as via specific code for this paper (which is
-documented and available online [65]).
-
-As both the spectral rigidity and level number variance can be highly sensitive to the choice of
-smoothing procedure used in the unfolding method [58], [66]–[69], all unfolded features in this
-study were calculated using polynomials of degree 5, 7, 9, 11, and 13 as the smoothing technique. As
-much of the variance in the unfolded spectral statistics can be due to the largest eigenvalues
-acting as anchor points during smoothing [66], [69], all unfolded features were also calculated
-twice (in addition to each polynomial degree) to examine the effects of trimming on spectral
-observables: once with only the largest eigenvalue omitted, and once with the largest 20 eigenvalues
-omitted.
-
-\subsection{Feature Evaluation}
-To evaluate the extracted RMT features, for each possible pairing of subgroups in each dataset, each
-feature was used to predict subject subgroup membership. That is, for each dataset, and for each
-possible pairing of subgroups, all scans belonging to a subgroup were pooled together and treated as
-independent samples of the subgroup of interest. So, for example, in the LEARN dataset, this
-resulted in a comparison of 384 (\(16 \times 4\)) task scans, to 48 (\(2 \times 24\)) rest scans. When it was not
-reasonable to compare scans across sessions, sessions were analyzed separately. This was the case
-for the PSYCH dataset, where psychological measurements, and thus subgroups, differed from session
-to session.
-
-A wide variety of standard, out-of-the-box classifiers available in the Python scikit-learn library
-[70] were then employed to gauge the predictive value of the RMT features. To allow for a wide range
-of flexibilities, classifiers employing k-nearest neighbours (KNN), with k values in \(\{1, 2, 3,
-\dots, 10, 20\}\), logistic regression (LR), linear discriminant analysis (LDA), and a
-support-vector machine (SVM) were employed. Default parameter values were used when possible (see
-[65] for precise function calls used).
-
-The accuracy of each binary classifier was then validated with leave-one-out cross-validation
-(LOOCV) on all datasets, for all combinations of trimming and unfolding choices.
-
-\section{Results}
-
-\subsection{Overview}
-Due to space limitations, we highlight only key results here. Extensive overviews of the full
-results are available in the supplementary materials. Visual summaries of the prediction accuracies
-for all features and analytic choices are presented in Fig. S1--4. Prediction accuracies for only the
-multivariate RMT features are summarized in Fig. S5--7, and the distributions of the scalar features
-are presented in Fig. S8--10. Prediction accuracies for the scalar features across analytic choices
-are summarized in Fig. S11--12.
-
-Numeric (mean) summaries across datasets, broken down by classifier and feature are available in
-Tables S11--18. Complete (unsummarized) tables of all predictions for all combinations of features,
-unfolding degree, and trimming are available as supplementary material in the spreadsheet
-\texttt{prediction\_tables\_fullpre.ods}. Unless otherwise noted, subsequent figures and analyses were
-computed using the preprocessed data: full results for the raw data are available as spreadsheets in
-the supplementary material in \texttt{prediction\_tables.ods}.
-
-With the exception of the LEARNING dataset, the RMT-inspired features all demonstrated some
-predictive utility (Fig. S1-4). For example, there were strong differences in RMT features between
-task and rest conditions for the REFLECT dataset, but the features were of variable predictive
-utility in the PSYCH dataset, sometimes outperforming the naïve classifier, and sometimes not.
-Although the RMT features were useful in the comparison involving duloxetine and controls (no pain)
-in the OSTEO dataset, these features mostly did not aid in classifying between any of the other
-possible pairings of subgroups in this dataset (Fig. S1-4).
-
-Although the performance of classifiers trained on RMT features was variable, there was a tendency
-for the raw and largest eigenvalues to be the most predictively useful features, especially in the
-REFLECT and PARK data in discriminating between Parkinson's and controls (Fig. S5-7. S13; but see
-Discussion).
-
-\subsection{Raw Features}
-As evident from Fig. S8-10, the scalar raw features (largest eigenvalue, noise ratio, and shifted
-noise ratio) varied more within than between subgroups. They were not generally useful for
-classifying conditions (Tables S15-17), with two or three notable exceptions.
-
-The noise ratio (but not shifted noise ratio) differed consistently between the duloxetine and
-control groups in the OSTEO data, allowing for prediction accuracies of up to 72\% compared to 51\%
-for the naive classifier (Fig. S11-12 top-right panel, Table S5). The largest eigenvalue was also
-useful in distinguishing between subjects self-reporting high versus low vigilance across sessions:
-accuracies up to 64\% and 65\% (compared to naïve accuracies of 55\% and 50\%) for the first and
-second sessions, respectively (Table S6). Finally, for the REFLECT data, task and rest were
-perfectly distinguishable with various features and with some classifiers.
-
-The sorted, raw eigenvalues, however, more consistently contained information that allowed some
-separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
-consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
-had the most consistent predictive value for distinguishing individuals with Parkinson's from
-controls, or distinguishing those with osteopathic pain taking duloxetine from controls ("nopain" in
-Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
-comparing those differing in self-reported vigilance and high vs. low scorers in the attention task
-(Table S7--8).
-
-In all these relationships, normalizing the eigenvalues to the interval \([0,1]\) prior to
-prediction considerably improved prediction accuracy (Tables S9-12).
-
-\paragraph{Role of Classifer}
-The raw eigenvalues generally had the most predictive utility when employing either an SVM
-classifier or KNN classifier with k > 3 (Table S11, first row). When the raw eigenvalues were first
-min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
-predicting subgroup memberships (Table S12), however without normalization, KNN with k = 8 or k = 9
-generally best predicted subgroup membership across datasets (Table S11).
-
-\subsection{Unfolded Features}
-The Brody parameter rarely resulted in predictions exceeding naïve accuracies in these comparisons:
-for most unfoldings and comparisons, the Brody parameter was either constant and equal for all
-subjects (e.g. exactly 0.999 or 0.0001) or deviated from these values only for a small number of
-subjects. This is because the observed NNSDs were usually well outside of the GOE-GDE bounds.
-
-The predictive utility of the unfolded RMT features varied considerably across datasets and
-classifiers. However, some results were robust: for example, the spectral rigidity and level number
-variance strongly differentiated controls from individuals prescribed duloxetine for osteopathic
-pain, and this differentiation persisted regardless of unfolding degree or trimming (Fig. 5-8, top-right panels).
-
-Despite the insensitivity to unfolding in this dataset, there was still nontrivial variability in
-accuracy across classifiers when classifying duloxetine vs controls, with the less flexible
-classifiers (logistic regression, SVM, and KNN with k=10, 20) generally performing best (Tables
-S1-2).
-
-For other datasets, the predictive utility of the unfolded features was highly dependent on the
-unfolding degree, and sometimes trim proportions. These differences are illustrated in detail for
-the Parkinson's data in Fig. S14. For example, with minimal trimming, subgroup overlap in spectral
-rigidity was minimized for unfolding degrees 5 and 9 (Fig. S14, first row), but with greater
-trimming, this was not the case (Fig. S14, second row). For the level number variance with minimal
-trimming, subgroup overlap appeared smaller for lower unfolding degrees (Fig. S14, third row),
-however, with a greater degree of trimming, separation was clearer for slightly larger degrees (Fig.
-S14, fourth row).
-
-Across datasets, the unfolding degree, compared to trimming, was a much more significant source of
-variance in the overall predictive utility of the unfolded features (e.g. note the minimal changes
-from trimming in Fig. S5 and S6, compared to the overall differences in prediction patterns from
-differing choices of unfolding degree in Fig. S3 and S4). However, neither the choice of unfolding
-degree nor amount of trimming had any easily summarizable effects on the predictive value of the
-various RMT metrics: the optimal trimming and unfolding degree was different for each dataset and
-comparison.
-
-\paragraph{Role of Classifier}. Across all datasets, unfoldings, and trimmings, the spectral rigidity was
-generally best able to separate between subgroups when either k was large (10 or 20) in KNN, or
-Logistic Regression or SVM was employed (Table S13). The pattern was similar for the level number
-variance, although the predictive utility of the level number variance was, overall, slightly more
-variable than the spectral rigidity (Table S4).
-
-\section{Discussion}
-In this study, certain features inspired by RMT and derived from the sorted eigenvalues of the full,
-whole-brain voxelwise correlation matrix were found to contain information with the potential to
-meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
-from previous studies [26], [33], [34], this suggests RMT may be a useful analytic and theoretical
-tool for understanding functional connectivity.
-
-Although the choice of unfolding and trimming often had nontrivial effects on the overall shape and
-values of the various spectral observables, and thus the predictive utility, in cases where there
-was clear predictive utility, there was some evidence that the unfolding and trimming procedures did
-not reverse the relations between subgroups (Fig. 1-3, S14). For example, regardless of the
-unfolding methodology, it was always the case that control subjects had spectral observables
-farthest from those of the GOE (Fig. 1-3) compared to those given duloxetine for osteopathic pain,
-whereas when comparing subject's with Parkinson's to controls, Parkinson's subjects consistently
-showed the greatest deviation from the GOE (Fig. S14).
-
-In particular, for some psychological phenomena examined in this study (specifically, the
-resting-state, voxelwise functional connectivity of those prescribed duloxetine for osteopathic pain
-versus the resting-state, voxelwise functional connectivity of controls), the spectral rigidity and
-level number variance contained information that was valuable in differentiating between subgroups.
-Across classifiers for this subgroup comparison, maximum mean LOOCV classification accuracies
-improved upon the 51\% accuracy of the naïve classifier by 11-26\% for the spectral rigidity and
-16-26\% for the level number variance (Tables S1-2).
-
-In the above case, these unfolded features allowed for significantly improved predictions over those
-using the raw (not unfolded) eigenvalues (Table S3). It is possible that some of these predictive
-improvements are because both the unfolding procedure and calculation of the spectral observables
-both smooth and transform all data to common scales. For example,  rescaling the raw eigenvalues to
-the interval [0, 1] also resulted in comparable prediction accuracies (Table S4).
-
-For comparison, for small datasets of rs- or task-fMRI data, binary classification using whole-brain
-features is challenging. For subject level classification, state of the art deep learning methods
-improve upon guessing by 17\% for autism [71] or 3-30\% for ADHD [72], 16\% for severe depression [73],
-and 23\% for obsessive compulsive disorder [74]. Manual feature engineering with more separable
-conditions (e.g. schizophrenia) can result in classification accuracies well above 90\% [75], and
-with very large data, sophisticated custom feature extraction methods can achieve near perfect
-accuracies at classifying task vs. rest [76]. However, for functional connectivity data and with
-classical machine learning algorithms (such as SVM) we in general only expect large prediction
-accuracies (e.g. greater than 80\%) when the group functional connectivities are already strongly
-separated (e.g. Cohen’s d > 1.0 [77]).
-
-Nevertheless, the features extracted in this study are somewhat unique in that they are small,
-spatially-insensitive summaries of the complete fMRI data. In particular, reducing the functional
-connectivity to just the largest 20 eigenvalues results in predictions comparable to using the
-entire set of eigenvalues (Tables S12, S18). That such a reduction can lead to accuracies comparable
-to other methods suggests that the system eigenvalues have utility beyond information reduction, and
-should perhaps be regarded themselves as signatures of functional connectivity patterns.
-
-\subsection{Confounds}
-While some of the RMT features allowed some classifiers to nearly perfectly separate conditions for
-the REFLECT dataset, this should almost certainly be regarded as an artifact of the differing scan
-parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
-correlation matrix is primarily determined by such differences in dimensions [78], as are the noise
-ratios, which include these dimensions in their definitions [29], [30]. Likewise, similar strong
-separation was observed in the PARK data when comparing the minimally pre-processed to fully
-pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps (spike
-removal, spatial smoothing, co-registration to larger T1 image [45]) which will also almost
-certainly alter the size of the largest eigenvalue.
-
-When there was predictive utility in the unfolded features,	 this utility was dependent on the
-unfolding procedure and choice of classifier. For example, the highest nontrivial LOOCV mean
-accuracy across all datasets was 0.79, corresponding to the duloxetine vs. controls subgroup
-comparison in the OSTEO dataset, whereas a poor choice of classifier in this same comparison could
-lead to prediction accuracies of 0.51, equivalent to the accuracy of a naive classifier (Table S2).
-
-There was variability in the predictive utility of the unfolding features, and this is likely
-because the unfolded features varied considerably based on the degree of the smoothing polynomial
-and extent of trimming. For example, the spread of the spectral rigidity and level number variance
-curves are shown for all subjects in the PARK data in Fig. S14. Clearly, both the curve shape,
-average magnitude, and extent of overlap is strongly dependent on the unfolding and trimming
-procedure.
-
-The unidimensional features, such as the largest eigenvalue or those based on the Marchenko-Pastur
-distribution, were most useful in predicting subgroup membership in trivial cases, such as in the
-REFLECT data, where the underlying scan dimensions differed between comparison groups, or in the
-PARK data, where the largest eigenvalue clearly distinguished the degree of preprocessing (Fig.
-S11-12). As these differences are due to methods, and not neuropsychological phenomena, they are
-perhaps of limited interest. To the extent that these features are scalar summaries of much
-higher-dimensional, 4D data, the limited utility of the scalar features should not be particularly
-surprising—it would be quite extraordinary if different neuropsychological states could be reliably
-separated based on such a reduction.
-
-However, that there are not consistently strong differences in these unidimensional features between
-subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
-noise-reduction procedures (such as [29], [30]), is unlikely to introduce bias into subsequent
-analyses. It perhaps also suggests these scalar features might be useful for quality control within
-a dataset: if there are differences between subgroups or individuals on one of the individual RMT
-features, then perhaps this is cause for more careful investigation.
-
-The unfolded features (spectral rigidity and level number variance) were more promising, however
-their sensitivity to the unfolding procedure means that great caution must be employed if they are
-to be used as features in future research. Ideally, the utility of these unfolded features (and all
-RMT features provided in this software release) need to be rigorously validated with wholly
-independent datasets addressing similar conditions. In addition, as simply using the raw, min-max
-normalized eigenvalues generally resulted in similar classification accuracies as the unfolded
-features, further study will be required to justify the use of these more complicated features.
-
-\subsection{Preprocessing}
-Across datasets, the features extracted from the minimally preprocessed images resulted in slightly
-lower prediction accuracies than the raw images for the unfolded features (rigidity, level
-variance), and slightly improved accuracies for some of the raw features (all eigenvalues, largest
-eigenvalue(s), noise ratio). However, these differences were quite small relative to the performance
-differences between datasets or between classifiers1.
-
-Similar patterns were observed in the PARK dataset. Fig. S5-7 show that more extensive preprocessing
-resulted in greater prediction accuracies for raw and largest eigenvalue(s), but decreased
-accuracies for the unfolded features. Given a particular level of preprocessing, however, it is
-clear from the spread of prediction accuracies that there was still substantial overall variance in
-prediction accuracy due to the choices of unfolding and classifier.
-
-\subsection{Comparison with Previous Studies}
-
-We generally find different results from previous studies [26], [33], [34] which have used the
-spectral rigidity or level number variance to differentiate between attention conditions. As [26]
-extracted rigidity curves from the correlations of electroencephalographic (EEG) signals, and [34]
-and [33] extracted spectral observables from the correlations of the mean signals of anatomical
-ROIs, these differences are to be expected. A single voxel signal amalgamates information from far
-fewer neurons than an EEG or mean ROI signal, and it is possible that any broad connections between
-RMT spectral observables and attention arise only at higher levels of organization, or after certain
-pre-processing steps (e.g. spatial smoothing) not performed in this study.
-
-We do find some modest evidence of greater overall correlation in high attention conditions: there
-was a slight tendency for the largest eigenvalue to be smaller when attention is low (Fig. S8).
-However, this is only suggestive, and future studies will be needed to determine how and under what
-conditions these RMT features reliably relate to neuropsychological phenomena.
-
-\subsection{Limitations}
-
-To make analyses across datasets at least minimally comparable, only limited preprocessing (motion
-and slicetime correction, skull-stripping) was performed in this study. In particular, in the
-absence of physiological noise regression, there must be caution in assuming any of the observed
-patterns are uniquely due to changes in functional connectivity. This is perhaps especially notable
-as the RMT features were most consistently capable of distinguishing between individuals taking
-duloxetine versus controls, or between controls and individuals with Parkinson's. It seems much more
-likely that physiological confounds could be involved in these comparisons in contrast to other
-comparisons (such as task-state versus resting-state, or in the differing levels of attention in the
-PSYCH dataset).
-
-The correlation matrices calculated for each run are computed from the full time-series data. No
-attempt was made to control for stimulus presentation times, under the assumption that the broad,
-whole-brain summarizing nature of the RMT features should be able to detect broad, overall
-differences between subgroups or conditions. However, differences were not consistently observed
-between conditions across all datasets. It is possible more careful selection of regions of the
-time-series in task conditions will be necessary in order for RMT features to reflect the
-differences between these conditions.
-
-While the extraction of eigenvalues from the correlations of all voxels has the advantage of not
-relying on a specific parcellation scheme, and does not impose anatomical or locality assumptions
-that ROI-based analyses do, it is also possible that this limited the utility of the RMT features.
-In particular, we found the spectral observables (NNSD, spectral rigidity, level number variance) to
-frequently fall outside the GOE/GDE bounds that have typically been observed in previous studies.
-Future studies should examine the sensitivity of these RMT features to the level of locality of
-analyses, or choice of size and location of ROIs.
-
-\section{Conclusion}
-
-\paragraph{Declarations}
-Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
-Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
-Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
-Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
-(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
-R0168020), and a St. Francis Xavier University UCR grant to JL.
-
-\section{References}
-\printbibliography{}
-
-\end{document}
+# References

### 2022-10-06 code/ismrm_plots.py

@@ -1,0 +1,533 @@
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import pandas as pd
+import seaborn as sbn
+
+
+from warnings import filterwarnings
+from pathlib import Path
+from typing import Any, List, Union
+from typing_extensions import Literal
+
+from rmt.args import ARGS
+from rmt.comparisons import Pairings
+
+# from rmt.plot_datasets import plot_largest, plot_pred_levelvar
+from rmt.summarize import supplement_stat_dfs, compute_all_preds_df
+from rmt._utilities import _percentile_boot
+
+from empyricalRMT.brody import brody_dist, fit_brody_mle
+from empyricalRMT.eigenvalues import Eigenvalues
+from empyricalRMT.ensemble import GOE, GDE
+
+Fmt = Literal["svg", "png"]
+
+ISMRM_FOLDER = Path.home() / "Documents/ISMRM"
+
+# fmt: off
+FEATURES = [
+    "Raw Eigs",
+    # "Largest",
+    # "Largest20",
+    # "Noise",
+    # "Noise (shift)",
+    # "Brody",
+    "Rigidity",
+    "Levelvar",
+]
+HCOLORS = [
+    "#000000",  # Raw Eigs
+    # "#c10000",  # Largest
+    "#cccccc",  # Largest
+    "#0066b3",  # Largest
+    "#a80000",  # Largest20
+    "#06B93A",  # Noise
+    "#058C2C",  # Noise (shift)
+    "#EA00FF",  # brody
+    "#FD8208",  # rigidity
+    "#D97007",  # levelvar
+]
+COLORDICT = dict(zip(FEATURES, HCOLORS))
+RAWEIGS_COLOR = ["#777777"]
+# fmt: on
+
+UNFOLDS = [5, 7, 9, 11, 13]
+N_SUBPLOTS = 17
+
+# see bottom of https://matplotlib.org/tutorials/introductory/customizing.html for all options
+FONT_RC = {
+    "family": "sans-serif",
+    "style": "normal",
+    "variant": "normal",  # "small-caps" is other
+    # "weight": "normal",
+    "weight": "medium",
+    # "size": 20.0,
+    "size": 16.0,
+    "serif": "Lato",
+    "sans-serif": "CMU Serif",
+}
[…]

### 2022-10-20 paper/paper/SPIE/article.tex
start SPIE paper

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tocloft}

\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}

\author[a]{Derek Berger}
\author[a]{Jacob Levman}
\author[b]{Gurpreet M. Matharoo}
\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Canada, B2G 2W5}

% \author[a]{First Author}
% \author[a]{Second Author}
% \author[b]{Third Author}
% \author[a,b,*]{Fourth Author}
% \affil[a]{University Name, Faculty Group, Department, Street Address, City, Country, Postal Code}
% \affil[b]{Company Name, Street Address, City, Country, Postal Code}

\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table}
\begin{document}
\maketitle

% \begin{abstract}
% This document shows the required format and appearance of a manuscript prepared for SPIE journals.
% It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
% following journals require the use of structured abstracts in manuscript submissions:
% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
% Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
% structured or single-paragraph, the abstract should be a summary of the paper and not an
% introduction. Because the abstract may be used in abstracting and indexing databases, it should be
% self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
% objectives, methodology used, results obtained, and their significance. A list of up to six
% keywords should immediately follow.
% \end{abstract}

% See https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
% for info on the structured abstract
%
% Purpose: This section presents the significance and aims, stating the broad impact and the
% rationale or motivation of the work, including potentially some background.
%
% Approach: This section briefly describes the materials and methods, including the statistical
% analyses, used in the research.
%
% Results: This section provides a core summary of the findings from the work, including study
% numbers, quantitative analyses, or discoveries.
%
% Conclusions: This section summarizes and interprets the approach and findings from the work,
% stating also the importance or impact of the findings.

% Purpose:
% Data-intensive modeling could provide insight on the broad variability in outcomes in
% spine surgery. Previous studies were limited to analysis of demographic and clinical
% characteristics. We report an analytic framework called “SpineCloud” that incorporates
% quantitative features extracted from perioperative images to predict spine surgery outcome.

% Approach:
% A retrospective study was conducted in which patient demographics, imaging, and outcome
% data were collected. Image features were automatically computed from perioperative CT.
% Postoperative 3- and 12-month functional and pain outcomes were analyzed in terms of improvement
% relative to the preoperative state. A boosted decision tree classifier was trained to predict
% outcome using demographic and image features as predictor variables. Predictions were computed
% based on SpineCloud and conventional demographic models, and features associated with poor outcome
% were identified from weighting terms evident in the boosted tree.

% Results
% Neither approach was predictive of 3- or 12-month outcomes based on preoperative data
% alone in the current, preliminary study. However, SpineCloud predictions incorporating image
% features obtained during and immediately following surgery (i.e., intraoperative and immediate
% postoperative images) exhibited significant improvement in area under the receiver operating
% characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3 months and AUC=0.69 (CI95=0.55
% to 0.82) at 12 months.

% Conclusions
% Predictive modeling of lumbar spine surgery outcomes was improved by incorporation of
% image-based features compared to analysis based on conventional demographic data. The SpineCloud
% framework could improve understanding of factors underlying outcome variability and warrants
% further investigation and validation in a larger patient cohort.
\begin{abstract}
\section*{Purpose}
Random matrix theory (RMT) is increasingly becoming a useful tool for understanding complex systems.
Prior studies have examined functional Magnetic Resonance Imaging (fMRI) images using tools from
RMT, with some success. However, RMT computations are complex, and sensitive to numerous analytic
choices. We systematically investigate the usefulness of RMT in a variety of fMRI datasets.

\section*{Approach}
We develop open-source software to efficiently compute RMT features from fMRI images, and examine
the cross-validated predictive potential of eigenvalue and RMT-based features ("eigenfeatures") used
with typical classificatory machine-learning algorithms. We systematically vary pre-processing extent,
normalization procedure, RMT unfolding procedure, and classifiers to examine the utility of these
eigenfeatures across various datasets.

\section*{Results}
We find

\section*{Conclusions}
This document shows the required format and appearance of a manuscript prepared for SPIE journals.
It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
following journals require the use of structured abstracts in manuscript submissions:
\textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
structured or single-paragraph, the abstract should be a summary of the paper and not an
introduction. Because the abstract may be used in abstracting and indexing databases, it should be
self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
objectives, methodology used, results obtained, and their significance. A list of up to six keywords
should immediately follow.
\end{abstract}

% Include a list of up to six keywords after the abstract
\keywords{optics, photonics, light, lasers, journal manuscripts, LaTeX template}

% Include email contact information for corresponding author
{\noindent \footnotesize\textbf{*}Fourth author name,  \linkable{myemail@university.edu} }

\begin{spacing}{2}   % use double spacing for rest of manuscript

\section{Introduction}
\label{sect:intro}  % \label{} allows reference to this section
This document shows the format and appearance of a manuscript prepared for submission to an SPIE journal. Note that this template is only intended to be used as a guideline for author convenience. It is designed for optimum clarity and ease of reading for editors and reviewers, but the template does not reflect the final page layout of a published journal paper. Accepted papers are professionally typeset in XML according to the layout and design of the journal.

\subsection{Use of This Document}

This document is prepared using LaTeX2e\cite{Lamport94,Goossens97} with the class file {\ttfamily spieman.cls}.  The LaTeX source file used to create this document is {\ttfamily article.tex}, which contains important formatting information embedded in it. Authors may use it as a template to create their own manuscript. While LaTeX properly handles most formatting issues, the author may occasionally need to intervene to obtain a satisfactorily  formatted manuscript.

\subsection{English}

Authors are strongly encouraged to follow the principles of sound technical writing, as found in Refs.~\citenum{Alred03} and \citenum{Perelman97}, for example. In addition, good English usage is essential. Authors whose native language is not English may wish to collaborate with a colleague whose English skills are more advanced. Alternatively, you may wish to have your manuscript professionally edited prior to submission by Editage, our recommended independent editorial service: \linkable{https://www.editage.com/spie/}. SPIE authors will receive a 15\% discount off their services. A spell checker can be helpful to discover misspelled words, but authors should also proofread their papers carefully prior to submission. Manuscripts that do not meet acceptable English standards or lack clarity may be rejected.

\subsection{Page Setup and Fonts}

All text and figures, including footnotes, must fit inside a text area 6.5 in.\ wide by 9 in.\ high (16.51 by 22.86 cm). Manuscripts must be formatted for US letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1 in.\ on the bottom, and 1 in.\ on the left and right.

The Times New Roman font is used throughout the manuscript, in the sizes and styles shown in Table~\ref{tab:fonts}. If this font is not available, use a similar serif font. The manuscript should not contain headers or footers. Pages should be numbered.

\begin{table}[ht]
\caption{Fonts sizes and styles.}
\label{tab:fonts}
\begin{center}
\begin{tabular}{|l|l|} %% this creates two columns
%% |l|l| to left justify each column entry
%% |c|c| to center each column entry
%% use of \rule[]{}{} below opens up each row
\hline
\rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
\hline\hline
\rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
\hline
\rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
\hline
\end{tabular}
\end{center}
\end{table}

\section{Parts of Manuscript}

This section describes the normal structure of a manuscript and how each part should be handled. The appropriate vertical spacing between various parts of this document is achieved in LaTeX through the proper use of defined constructs, such as \verb|\section{}|.

\subsection{Title and Author Information}
\label{sect:title}
The article title appears left justified at the top of the first page. The title font is 16 pt., bold. The rules for capitalizing the title are the same as for sentences; only the first word, proper nouns, and acronyms should be capitalized. Do not begin titles with articles (for example, a, an, the) or prepositions (for example, on, by, etc.). The word ``novel'' should not appear in the title, as publication will imply novelty. Avoid the use of acronyms in the title, unless they are widely understood. Appendix A contains more about acronyms.

The list of authors immediately follows the title, 18 points below. The font is 12 pt., bold and the author names are left justified. The author affiliations and addresses follow the names, in 10-pt., normal font and left justified. For multiple affiliations, each affiliation should appear on a separate line. Superscript letters (a, b, c, etc.) should be used to associate multiple authors with their respective affiliations. The corresponding author should be identified with an asterisk, and that person's email address should be provided below the keywords.

\subsection{Abstract}
The abstract should be a summary of the paper and not an introduction. Because the abstract may be used in abstracting journals, it should be self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the objectives, methodology used, results obtained, and their significance. Please note that the following journals require the use of structured abstracts in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on the website of the journal.

\subsection{Subject terms/Keywords}
Keywords are required. Please provide 3-6 keywords related to your paper.

\subsection{Body of Paper}
The body of the paper consists of numbered sections that present the main findings. These sections should be organized to best present the material.

To provide transition elements in your paper, it is important to refer back (or forward) to specific sections. Such references are made by indicating the section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1 contained a description...'' If the word Section, Reference, Equation, or Figure starts a sentence, it is spelled out. When occurring in the middle of a sentence, these words are abbreviated Sec., Ref., Eq., and Fig.

At the first occurrence of an acronym, spell it out followed by the acronym in parentheses, for example, charge-coupled diode (CCD).

\subsection{Footnotes}
Textual footnotes should be used rarely to present important documentary or explanatory material whose inclusion in the text would be distracting.\footnote{Example of a footnote.} Due to problems with HTML display, use of footnotes should generally be avoided. If absolutely necessary, the footnote mark must come at the end of a sentence. To insert a footnote, use the {\verb|\footnote{}|} command.

\subsection{Appendices}
Brief appendices may be included when necessary, such as derivations of equations, proofs of theorems, and details of algorithms. Equations and figures appearing in appendices should continue sequential numbering from earlier in the paper.

\subsection{Disclosures}
Conflicts of interest should be declared under a separate header, above Acknowledgments. If the authors have no competing interests to declare, then a statement should be included declaring no conflicts of interest. For assistance generating a disclosure statement, see the form available from  the International Committee of Medical Journal Editors website: \linkable{http://www.icmje.org/conflicts-of-interest/}

\subsection{Acknowledgments}
Acknowledgments and funding information should be added after the conclusion, and before references. Include grant numbers and the full name of the funding body. The acknowledgments section does not have a section number.

\subsection{Data, Materials, and Code Availability}
Availability of data, materials, and/or code used in the research results reported in the manuscript may be declared under the heading "Data, Materials, and Code Availability," following the Acknowledgments section. As relevant, provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).

\subsection{References}
The References section lists books, articles, and reports that are cited in the paper. This section does not have a section number. The references are numbered in the order in which they are cited. Examples of the format to be followed are given at the end of this document.

The reference list at the end of this document is created using BibTeX, which looks through the file {\ttfamily report.bib} for the entries cited in the LaTeX source file.  The format of the reference list is determined by the bibliography style file {\ttfamily spiejour.bst}, as specified in the \\ \verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may be directly formatted in the LaTeX source file.

For books\cite{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.  Journal-article references \cite{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.  A reference to a proceedings paper or a chapter in an edited book\cite{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.  References to an article in the SPIE Proceedings may include the conference name, as shown in Ref.~\citenum{Hanson93c}.

The references are numbered in the order of their first citation. Citations to the references are made using superscripts, as demonstrated in the preceding paragraph. One may also directly refer to a reference within the text, for example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more references should be separated by a comma with no space between them. Multiple sequential references should be displayed with a dash between the first and last numbers \cite{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.

\subsubsection{Reference linking and DOIs}
A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a digital object, such as a journal article or a book chapter, that provides a persistent link to its location on the internet. The use of DOIs allows readers to easily access cited articles. Authors should include the DOI at the end of each reference in brackets if a DOI is available. See examples at the end of this manuscript. A free DOI lookup service is available from CrossRef at \\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will facilitate reference linking and is highly recommended.

In the present LaTeX template, the author needs to add the DOI reference by including it in a ``note'' in the bibliography file, as shown in the file {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}. The DOI may be used by the reader to locate that document with the link: {\verb+http://dx.doi.org10.1117/12.154577+}.

\subsection{Biographies}
A brief professional biography of approximately 75 words may be provided for each author, if available. Biographies should be placed at the end of the paper, after the references. Personal information such as hobbies or birthplace/birthdate should not be included. Author photographs are not published.

\section{Section Formatting}
\label{sect:sections}
In LaTeX, a new section is created with the \verb|\section{}| command, which automatically numbers the sections. Sections will be numbered sequentially, starting with the first section after the abstract, except for the acknowledgments and references. (Note that numbering of section headings is not required, but the numbering must be consistent if used.) All section headings should be left justified.

Main section headings are in 12-pt. bold font, left-justified and in title case, where important words are capitalized.

Paragraphs that immediately follow a section heading are leading paragraphs and should not be indented, according to standard publishing style. The same goes for leading paragraphs of subsections and sub-subsections. Subsequent paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no additional space between paragraphs. In LaTeX, paragraphs are separated by blank lines in the source file. Indentation of the first line of a paragraph may be avoided by starting it with \verb|\noindent|.

\subsection{Subsection Headings}
All important words in a subsection (level 1) header are capitalized. Subsection numbers consist of the section number, followed by a period, and the subsection number within that section, without a period at the end. The heading is left justified and its font is 12 pt. italic.

\subsubsection{Sub-subsection headings}
The first word of a sub-subsection is capitalized. The rest of the text is not capitalized, except for proper names and acronyms (the latter should only be used if well known). The heading is left justified and its font is 11 pt. italic.

\section{Figures and Tables}

\subsection{Figures}

Figures are numbered in the order in which they are called out in the text. They should appear in the document in numerical order and as close as possible to their first reference in the text. It may be necessary to move figures or tables around to enhance readability. LaTeX will attempt to place figures at the top or bottom of a page in which they are first referenced.

Figures, along with their captions, should be separated from the main text by  0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or graph. Figure captions start with the abbreviation ``Fig'' in front of the figure number, followed by a period, and the text in 10-pt. font. See Fig.~\ref{fig:example} for an example.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[height=5.5cm]{mcr3b.eps}
\end{tabular}
\end{center}
\caption
{ \label{fig:example}
Example of a figure caption. }
\end{figure}

Authors may wish to create figures consisting of two or more images, in which case, they should be neatly arranged in a rectangular array.  In no case, should the article's text be wrapped around a figure. Figure~\ref{fig:example2} shows two side-by-side images. When a figure contains more than one image, the author must submit them as a single image file. Further details about figure formatting can be found in the author guidelines for each specific SPIE journal: \\
\linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
\\
(a) \hspace{5.1cm} (b)
\end{tabular}
\end{center}
\caption
{ \label{fig:example2}
Example of a figure containing multiple images: (a) sun and (b) blob. Figures containing multiple images must be submitted to SPIE as a single image file.}
\end{figure}

\subsection{Tables}
Tables are numbered in the order in which they are referenced. They should appear in the document in numerical order and as close as possible to their first reference in the text. It is preferable to have tables appear at the top or bottom of the page, if possible. Table captions are handled identically to those for figures, except that they appear above the table. See Table~\ref{tab:fonts} for an example.

\subsection{Multimedia}
Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player. The recommended maximum size for each multimedia file is 10-12 MB. Authors must insert a representative still image from the video file in the manuscript as a figure. The caption label will be linked by the publisher to the actual video file. The video may also be mentioned in an existing figure caption. Multimedia files are treated in the same manner as figures and they will be numbered sequentially with normal figures.  The video number, file type, and file size should be included in parentheses at the end of the figure caption. See Figure \ref{vid:satellite} for an example.

\begin{video}
\begin{center}
{\includegraphics[height=5cm]{satellite.eps}}
\\
\end{center}
\caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
\end{video}

\appendix    % this command starts appendixes

\section{Miscellaneous Formatting Details}
\label{sect:misc}
At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.

\subsection{Formatting Equations}
Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.  Important equations appear on their own line.  Such equations are centered.  For example, ``The expression for the field of view is
\begin{equation}
\label{eq:fov}
2 a = \frac{(b + 1)}{3c} \, ,
\end{equation}
where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.

Equations are considered to be part of a sentence and should be punctuated accordingly. In the above example, a comma appears after the equation because the next line is a subordinate clause. If the equation ends the sentence, a period should follow the equation. The line following an equation should not be indented unless it is meant to start a new paragraph. Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.

References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise. Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..

\subsection{Formatting Theorems}

To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.  The text of the theorem continues on the same line in normal, 10-pt. font, achieved in LaTeX using \verb|\footnotesize|.  For example,

\vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}

% \disclosures
\subsection*{Disclosures}
Conflicts of interest should be declared under a separate header. If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.

\subsection* {Acknowledgments}
This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

\subsection* {Data, Materials, and Code Availability}
As relevant, the availability of data, materials, and/or software code used in the research results reported in the manuscript may be declared in this section. (Note: this section is required for the \textit{Journal of Biomedical Optics} and \textit{Neurophotonics}.) Provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).

%%%%% References %%%%%

\bibliography{report}   % bibliography data in report.bib
\bibliographystyle{spiejour}   % makes bibtex use spiejour.bst

%%%%% Biographies of authors %%%%%

\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering. He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.  He is the author of more than 50 journal papers and has written three book chapters. His current research interests include optical interconnects, holography, and optoelectronic systems. He is a member of SPIE.

\vspace{1ex}
\noindent Biographies and photographs of the other authors are not available.

\listoffigures
\listoftables

\end{spacing}
\end{document}

### 2022-10-20 paper/paper/SPIE/article.tex
wip

@@
 \author[a]{Derek Berger}
-\author[a]{Jacob Levman}
+\author[a,*]{Jacob Levman}
 \author[b]{Gurpreet M. Matharoo}
@@
 
-% Conclusions
-% Predictive modeling of lumbar spine surgery outcomes was improved by incorporation of
+% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved by incorporation of
 % image-based features compared to analysis based on conventional demographic data. The SpineCloud
@@
 \begin{abstract}
+
 \section*{Purpose}
-Random matrix theory (RMT) is increasingly becoming a useful tool for understanding complex systems.
+Random matrix theory (RMT) is an increasingly useful tool for understanding large, complex systems.
 Prior studies have examined functional Magnetic Resonance Imaging (fMRI) images using tools from
 RMT, with some success. However, RMT computations are complex, and sensitive to numerous analytic
-choices. We systematically investigate the usefulness of RMT in a variety of fMRI datasets.
+choices, and previous investigations have been mostly explanatory / descriptive. We systematically
+investigate the usefulness of RMT in a variety of fMRI datasets in a predictive framework.
 
@@
 \section*{Results}
-We find
+We find eiegenfeatures to be broadly predictive in a wide variety of contexts, and that RMT-specific
+features like the spectral rigidity and level number variance have considerable potential in
+feature-engineering for fMRI. On median, we find ...
 
 \section*{Conclusions}
-This document shows the required format and appearance of a manuscript prepared for SPIE journals.
-It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
-following journals require the use of structured abstracts in manuscript submissions:
-\textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
-Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
-Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
-structured or single-paragraph, the abstract should be a summary of the paper and not an
-introduction. Because the abstract may be used in abstracting and indexing databases, it should be
-self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
-objectives, methodology used, results obtained, and their significance. A list of up to six keywords
-should immediately follow.
+Use RMT yo.
+
+% This document shows the required format and appearance of a manuscript prepared for SPIE journals.
+% It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
+% following journals require the use of structured abstracts in manuscript submissions:
+% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
+% Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
+% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
+% structured or single-paragraph, the abstract should be a summary of the paper and not an
+% introduction. Because the abstract may be used in abstracting and indexing databases, it should be
+% self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
+% objectives, methodology used, results obtained, and their significance. A list of up to six keywords
+% should immediately follow.
 \end{abstract}
@@
 % Include a list of up to six keywords after the abstract
-\keywords{optics, photonics, light, lasers, journal manuscripts, LaTeX template}
+\keywords{random matrix, spectral rigidity, level number variance, fMRI, classification, machine-learning}
 
 % Include email contact information for corresponding author
-{\noindent \footnotesize\textbf{*}Fourth author name,  \linkable{myemail@university.edu} }
+{\noindent \footnotesize\textbf{*}Jacob Levman,  \linkable{jlevman@stfx.ca} }
 

### 2022-10-31 paper/paper/SPIE/article.tex
[SHARE]

@@
 Random matrix theory (RMT) is an increasingly useful tool for understanding large, complex systems.
-Prior studies have examined functional Magnetic Resonance Imaging (fMRI) images using tools from
+Prior studies have examined functional Magnetic Resonance Imaging (fMRI) scans using tools from
 RMT, with some success. However, RMT computations are complex, and sensitive to numerous analytic
@@
 features like the spectral rigidity and level number variance have considerable potential in
-feature-engineering for fMRI. On median, we find ...
+feature-engineering for fMRI. On median, we find ... However, we also find the utility of the
+extracted eigenfeatures to depend strongly and complexly on various analytic decisions.
 
 \section*{Conclusions}
-Use RMT yo.
+Consider using RMT.
 
@@
 \label{sect:intro}  % \label{} allows reference to this section
-This document shows the format and appearance of a manuscript prepared for submission to an SPIE journal.
-Note that this template is only intended to be used as a guideline for author convenience.
-It is designed for optimum clarity and ease of reading for editors and reviewers, but the template does not reflect the final page layout of a published journal paper.
-Accepted papers are professionally typeset in XML according to the layout and design of the journal.
+% This document shows the format and appearance of a manuscript prepared for submission to an SPIE
+% journal. Note that this template is only intended to be used as a guideline for author convenience.
+% It is designed for optimum clarity and ease of reading for editors and reviewers, but the template
+% does not reflect the final page layout of a published journal paper. Accepted papers are
+% professionally typeset in XML according to the layout and design of the journal.
 
-\subsection{Use of This Document}
+In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
+(BOLD) signals are related to neural activity. It is common to investigate statistical relationships
+between the BOLD signals through \textbf{functional connectivity} analyses, where correlations between
+collections of these signals are examined to infer connections between different voxels or regions
+of interest (ROIs) within the brain.
 
-This document is prepared using LaTeX2e\cite{Lamport94,Goossens97} with the class file {\ttfamily spieman.cls}.
-The LaTeX source file used to create this document is {\ttfamily article.tex}, which contains important formatting information embedded in it.
-Authors may use it as a template to create their own manuscript.
-While LaTeX properly handles most formatting issues, the author may occasionally need to intervene to obtain a satisfactorily  formatted manuscript.
+Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
+complex functional connectivity networks are ubiquitous
+\citenum{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
+This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
+of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
+physics \citenum{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. In various
+physical systems where nuclei can have different possible states, precisely modeling the entirety of
+the interactions between nuclei is often intractable. For such large, complex systems, the
+eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
+of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
+spectrum statistically, complex systems can be investigated by comparison to the universal
+properties predicted by RMT.
 
-\subsection{English}
+In recent years, RMT analyses have been applied to study brain functioning.  The earliest  study
+demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
+resemble those of the Gaussian Orthogonal Ensemble (GOE) \citenum{sebaRandomMatrixAnalysis2003}.
+Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
+fMRI data \citenum{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
+been used in diffusion MRI to aid in the selection of the number of components to employ in
+principal-component reduction analysis and denoising
+\citenum{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
 
-Authors are strongly encouraged to follow the principles of sound technical writing, as found in Refs.~\citenum{Alred03} and \citenum{Perelman97}, for example.
-In addition, good English usage is essential.
-Authors whose native language is not English may wish to collaborate with a colleague whose English skills are more advanced.
-Alternatively, you may wish to have your manuscript professionally edited prior to submission by Editage, our recommended independent editorial service: \linkable{https://www.editage.com/spie/}.
-SPIE authors will receive a 15\% discount off their services.
-A spell checker can be helpful to discover misspelled words, but authors should also proofread their papers carefully prior to submission.
-Manuscripts that do not meet acceptable English standards or lack clarity may be rejected.
+Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
+differences between rest and task states \citenum{wangSpectralPropertiesTemporal2015}, between
+subjects with and without ADHD \citenum{wangRandomMatrixTheory2016}, and between pain and non-pain
+states \citenum{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
+resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
+certain aspects of psychological processes might be characterized, in part, by features computed
+from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
+interpretable manner across psychological processes. If this is the case, RMT could have unique
+potential in characterizing the functioning of the human brain.
+
+\section{Datasets}
+
+As our primary intention is to examine the potential utility of RMT-derived features across a wide
+variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
+psychological processes, we examine a wide selection of fMRI datasets available on the
+\href{https://openneuro.org/}{OpenNeuro} platform [36].  This allows for a broad assessment of the
+potential of voxelwise RMT features.
+
+However, since previous studies employing RMT for feature extraction have partially interpreted
+their findings with respect to attentional processes
+\citenum{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have elected to use
+datasets that involve attention in some way. This includes measures of attention or vigilance, or
+comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
+which should involve considerably less attention or focus (e.g. resting-state).
+
+Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
+and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
+(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
+these datasets, we highlight below only those aspects relevant to the current paper.
+
+\begin{table}[h!]
+\small
+\centering
+\begin{tabular}{ c c c c c c c }
+\hline
+\textbf{ID}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} \\
+\hline
+PSYCH     & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0  & 300  \\
+PARK      & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4  & 149  \\
+OSTEO     & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5  & 300  \\
+BILING    & 100 x 100 x 72 & 1.8 x 1.8 x 1.8 & 0.88 & 823  \\
+LEARN     & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0  & 195  \\
+DEPRESS   & 112 x 112 x 25 & 2.0 x 2.0 x 5.0 & 2.5 & 100   \\
+AGING     & 74 x 74 x 32   & 3.0 x 3.0 x 4.0 & 2.0 & 300   \\
+\hline
+\end{tabular}
+\caption{
+    ID = Identifier for paper. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P slices with dimensions M x N.
+}
+\label{table:1}
+\end{table}
+
+\begin{table}[h!]
+\small
+\centering
+\begin{tabular}{ l l c c c }
+\hline
+\textbf{ID}  & \textbf{Subgroup / Condition}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
+\hline
+PSYCH   &  high\_vigilance    & 12 / 11 & resting-state                    & 2 \\
+        &  low\_vigilance     & 10 / 11 & resting-state                    & 2 \\
+        &  trait\_attend      & 12 / 15 & resting-state                    & 2 \\
+        &  trait\_non-attend  & 10 / 7  & resting-state                    & 2 \\
+        &  task\_attend       & 13 / 15 & resting-state                    & 2 \\
+        &  task\_nonattend    &  9 / 7  & resting-state                    & 2 \\
+\hline
+LEARN   &  task              & 24      & learn image sketches & 16 \\
+        &  rest              & 24      & resting-state                    & 2 \\
+\hline
+REFLECT &  control (no pain) & 20      & resting-state                    & 1 \\
+        &  duloxetine        & 19      & resting-state                    & 1 \\
+        &  placebo           & 37      & resting-state                    & 1 \\
+        &  allpain (duloxetine + placebo) & 56         & resting-state    & 1 \\
+\hline
+PARK    &  Parkinson's       & 17      & ANT                              & 6 \\
+        &  control           & 17      & ANT                              & 6 \\
+\hline
+ECHO    &  summed/interleave & 31      & resting-state                    & 1 \\
+        &  summed/interleave & 31      & imagery/reflection               & 4 \\
+\hline
+\end{tabular}
+\caption{
+    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P slices with dimensions M x N.
+}
+\label{table:2}
+\end{table}
+
+\section{Methods}
 

### 2022-10-31 code/rmt/updated_dataset.py

@@ -561,6 +561,7 @@ def levelvars(
     print(f"Saved levelvars to {outfile}")
     return df
 
+
 def find_bad_labelings() -> None:
     """There was a bug where the cached files had erroneous labels. It is now fixed, but
     this was the test of that.
@@ -575,12 +576,19 @@ def find_bad_labelings() -> None:
                 print(data.info)
 
             for degree in [3, 5, 7, 9]:
-                lvars = levelvars(dataset=data, degree=degree, parallel=True)
-                rigs = rigidities(dataset=data, degree=degree, parallel=True)
-                if (len(lvars.y.unique()) == 1):
-                    print(f"Problem for {source.name} {preproc.name} degree={degree} levelvars")
-                if (len(rigs.y.unique()) == 1):
-                    print(f"Problem for {source.name} {preproc.name} degree={degree} rigidities")
+                try:
+                    lvars = levelvars(dataset=data, degree=degree, parallel=True)
+                    rigs = rigidities(dataset=data, degree=degree, parallel=True)
+                    if len(lvars.y.unique()) == 1:
+                        print(
+                            f"Problem for {source.name} {preproc.name} degree={degree} levelvars"
+                        )
+                    if len(rigs.y.unique()) == 1:
+                        print(
+                            f"Problem for {source.name} {preproc.name} degree={degree} rigidities"
+                        )
+                except IndexError:
+                    print(f"Problem for {source.name} {preproc.name} degree={degree}")
 
 
 if __name__ == "__main__":

### 2022-10-31 code/rmt/updated_dataset.py
[SHARE]

@@ -568,7 +568,11 @@ def find_bad_labelings() -> None:
     """
     for source in UpdatedDataset:
         for preproc in PreprocLevel:
-            data = UpdatedProcessedDataset(source=source, preproc_level=preproc)
+            try:
+                data = UpdatedProcessedDataset(source=source, preproc_level=preproc)
+            except IndexError:
+                print(f"Problem for {source.name} {preproc.name}")
+                continue
 
             if len(np.unique(data.labels()).ravel()) == 1:
                 print(f"Defective: {source} w preproc={preproc}")

### 2022-10-31 paper/paper/SPIE/article.tex
wip - paper

@@
 \hline
-\textbf{ID}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} \\
+\textbf{ID}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
 \hline
-PSYCH     & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0  & 300  \\
-PARK      & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4  & 149  \\
-OSTEO     & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5  & 300  \\
-BILING    & 100 x 100 x 72 & 1.8 x 1.8 x 1.8 & 0.88 & 823  \\
-LEARN     & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0  & 195  \\
-DEPRESS   & 112 x 112 x 25 & 2.0 x 2.0 x 5.0 & 2.5 & 100   \\
-AGING     & 74 x 74 x 32   & 3.0 x 3.0 x 4.0 & 2.0 & 300   \\
+AGING     & 74 x 74 x 32   & 3.0 x 3.0 x 4.0 & 2.0  & 300 & 62  \\
+BILING    & 100 x 100 x 72 & 1.8 x 1.8 x 1.8 & 0.88 & 823 & 90  \\
+DEPRESS   & 112 x 112 x 25 & 2.0 x 2.0 x 5.0 & 2.5  & 100 & 72  \\
+LEARN     & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0  & 195 & 432 \\
+OSTEO     & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5  & 300 & 74  \\
+PARK      & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4  & 149 & 552 \\
+PSYCH     & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0  & 300 & 90  \\
 \hline
@@
     Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P slices with dimensions M x N.
+    indicate P horizontal / axial slices each with dimensions M x N.
 }
@@
 \hline
-\textbf{ID}  & \textbf{Subgroup / Condition}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
+\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
 \hline
-PSYCH   &  high\_vigilance    & 12 / 11 & resting-state                    & 2 \\
-        &  low\_vigilance     & 10 / 11 & resting-state                    & 2 \\
-        &  trait\_attend      & 12 / 15 & resting-state                    & 2 \\
-        &  trait\_non-attend  & 10 / 7  & resting-state                    & 2 \\
-        &  task\_attend       & 13 / 15 & resting-state                    & 2 \\
-        &  task\_nonattend    &  9 / 7  & resting-state                    & 2 \\
+AGING   &  older             & 28      & resting-state         & 1 \\
+        &  younger           & 34      & resting-state         & 1 \\
+\hline
+BILING  &  bilingual          & 59      & resting-state        & 1 \\
+        &  monolingual        & 33      & resting-state        & 1 \\
+\hline
+DEPRESS &  depression         & 51      & resting-state        & 1 \\
+        &  control            & 21      & resting-state        & 1 \\
 \hline
@@
 \hline
-REFLECT &  control (no pain) & 20      & resting-state                    & 1 \\
-        &  duloxetine        & 19      & resting-state                    & 1 \\
-        &  placebo           & 37      & resting-state                    & 1 \\
-        &  allpain (duloxetine + placebo) & 56         & resting-state    & 1 \\
+OSTEO   &  duloxetine        & 19      & taking duloxetine     & 1 \\
+        &  pain              & 37      & placebo               & 1 \\
+        &  nopain            & 20      & control               & 1 \\
 \hline
-PARK    &  Parkinson's       & 17      & ANT                              & 6 \\
-        &  control           & 17      & ANT                              & 6 \\
+PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
+        &  control           & 21      & ANT                   & 12 \\
 \hline
-ECHO    &  summed/interleave & 31      & resting-state                    & 1 \\
-        &  summed/interleave & 31      & imagery/reflection               & 4 \\
+PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
+        &  nonigilant         & 11     & resting-state        & 3 \\
+        &  trait-attentive     & 11    & resting-state        & 3 \\
+        &  trait-non-attentive & 11    & resting-state        & 3 \\
+        &  task-attentive     & 11     & resting-state        & 3 \\
+        &  task-nonattentive  & 11     & resting-state        & 3 \\
 \hline

### 2022-10-31 code/rmt/updated_dataset.py

@@ -595,5 +595,17 @@ def find_bad_labelings() -> None:
                     print(f"Problem for {source.name} {preproc.name} degree={degree}")
 
 
+def print_subgroup_counts() -> None:
+    for source in sorted(UpdatedDataset, key=lambda d: d.name):
+        data = UpdatedProcessedDataset(
+            source=source, preproc_level=PreprocLevel.BrainExtract
+        )
+        print(data.source.name)
+        print(data.info.groupby(["label"]).count()["sid"])
+        print(data.info.drop(columns=["session", "run"]).groupby(["sid"]).count())
+        print("="*80)
+
+
 if __name__ == "__main__":
-    find_bad_labelings()
+    # find_bad_labelings()
+    print_subgroup_counts()

### 2022-10-31 paper/paper/SPIE/article.tex
WATCH OUT FOR CITEP

@@
 We develop open-source software to efficiently compute RMT features from fMRI images, and examine
-the cross-validated predictive potential of eigenvalue and RMT-based features ("eigenfeatures") used
+the cross-validated predictive potential of eigenvalue and RMT-based features (``eigenfeatures'') used
 with typical classificatory machine-learning algorithms. We systematically vary pre-processing extent,
@@
 psychological processes, we examine a wide selection of fMRI datasets available on the
-\href{https://openneuro.org/}{OpenNeuro} platform [36].  This allows for a broad assessment of the
+\href{https://openneuro.org/}{OpenNeuro} platform (36).  This allows for a broad assessment of the
 potential of voxelwise RMT features.
@@
 their findings with respect to attentional processes
-\citenum{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have elected to use
+\citenum{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
 datasets that involve attention in some way. This includes measures of attention or vigilance, or
 comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
-which should involve considerably less attention or focus (e.g. resting-state).
+which should involve less attention or focus (e.g. resting-state, pain medication vs. none).
+
+\subsection{Dataset: PSYCH}
+The PSYCH Dataset  is detailed in \citenum{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
+battery of psychological measurements, and various resting-state fMRI (rs-fMRI) whole-brain 7-Tesla
+scans were obtained. For the purpose of keeping with the ``attention'' theme of the analyses in this
+paper, three pairings of subgroups were constructed via z-score splits of the various sum-scores of
+scale items involving concepts closely related to attention.
+
+\emph{Trait Attention}.
+The PANAS-X \citenum{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
+queries the extent to which subjects agree particular affective terms have applied to them over a
+recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) scale was constructed from the
+scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
+``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects were then placed in the high / low ``trait
+attention'' category based on a median split of the average z-score across items.
+
+\emph{Task Attention}.
+The Conjunctive Continuous Performance Task (CCPT;
+\citenum{shalevConjunctiveContinuousPerformance2011} is a behavioural measure of sustained attention
+where subjects must react only if they see a red square among other different coloured figures. For
+this paper, a sum score was constructed by subtracting the number of false positive and false
+negative responses from the number of success responses, and subjects with a z-score greater / less
+than zero on this scale were placed in high / low ``task attention'' category.
+
+\emph{Vigilance}.
+The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
+state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
+from German; \citenum{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
+alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
+``vigilance'' conditions in the same manner as the previous two splits.
+
+\subsubsection{Dataset 2}
+Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
+\citenum{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
+condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
+either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
+also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
+subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
+documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
+complementary to \citenum{matharooSpontaneousBackpainAlters2020}.
+
+\subsubsection{Dataset 3}
+Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
+cognitive tasks) is described in full detail by \citenum{dupreMultiechoFMRIReplication2016}. Both
+multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
+involved viewing various affective images (see \citenum{sprengPatternsBrainActivity2010} for details)
+and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
+Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
+different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
+resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
+
+In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
+the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
+image with overall increased contrast \citenum{kunduMultiechoFMRIReview2017}. In the ``interleave''
+condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
+from echo \(j\) as \(t_{j,i}\), then the interleaved voxel of the new image has the time-series \([t_{1,1},
+t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
+
+This creates a physically-meaningful 4D image that contains the full, original multi-echo fMRI
+time-series data without any alterations. It should be noted that the purpose of these combination
+methods is not to create a 4D volume that allows optimal detection of the purely task-linked BOLD
+signal, but to ensure all scan information was used with minimal additional analytic assumptions.
+
+\subsubsection{Dataset 4}
+Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
+described in detail in \citenum{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
+in Table \ref{table:1}. Subjects with non-demented Parkinson's Disease (PD) and healthy controls performed a
+number of repetitions of the Attention Network Task
+(ANT; \citenum{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
+task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
+groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
+instead differ by virtue of the presence or absence of PD, any results here serve as a useful
+comparison point for the results from studies that should clearly involve attentional differences.
+This dataset also included extensively preprocessed scans, which were included in analyses to
+examine the effects of preprocessing on the extracted RMT features.
+
+\subsubsection{Dataset 5}
+Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
+detail in \citenum{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
+were available for all subjects, allowing for within-subject comparisons. The complex task involved
+multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
+related by family-resemblance of the components see figure 1 of \citenum{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
+memory of the various learned features. Scans corresponding to both the initial learning and
+subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of
+the analysis of this paper. Resting state scans were included in the low-attention ``rest''
+subgroup.
 
@@
 \begin{table}[h!]
+\caption{
+    ID = Identifier for paper. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P horizontal / axial slices each with dimensions M x N.
+}
+\label{table:1}
 \small
@@
 \end{tabular}
-\caption{
-    ID = Identifier for paper. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P horizontal / axial slices each with dimensions M x N.
-}
-\label{table:1}
+
 \end{table}
@@
 \begin{table}[h!]
+\caption{
+    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
+    Task = fMRI task. ANT = Attention Network Task \citenum{fanActivationAttentionalNetworks2005}
+}
+\label{table:2}
 \small
@@
 \end{tabular}
-\caption{
-    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P slices with dimensions M x N.
-}
-\label{table:2}
 \end{table}
@@
 \subsection{Data, Materials, and Code Availability}
-Availability of data, materials, and/or code used in the research results reported in the manuscript may be declared under the heading "Data, Materials, and Code Availability," following the Acknowledgments section.
+Availability of data, materials, and/or code used in the research results reported in the manuscript may be declared under the heading ``Data, Materials, and Code Availability,'' following the Acknowledgments section.
 As relevant, provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).
@@
 
-For books\cite{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.
-Journal-article references \cite{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.
-A reference to a proceedings paper or a chapter in an edited book\cite{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.
+For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.
+Journal-article references \citenum{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.
+A reference to a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.
 References to an article in the SPIE Proceedings may include the conference name, as shown in Ref.~\citenum{Hanson93c}.
@@
 One may also directly refer to a reference within the text, for example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more references should be separated by a comma with no space between them.
-Multiple sequential references should be displayed with a dash between the first and last numbers \cite{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
+Multiple sequential references should be displayed with a dash between the first and last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
 

### 2022-10-31 paper/paper/SPIE/article.tex
WIP on paper

@@
 
-Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
-complex functional connectivity networks are ubiquitous
-\citenum{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
+Whether in the presence of experimental stimuli, or the relative absence, as in resting state fMRI (rs-fMRI),
+complex patterns in functional connectivity networks are ubiquitous
+\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
 This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
 of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
-physics \citenum{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. In various
+physics \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}
+
+In various
 physical systems where nuclei can have different possible states, precisely modeling the entirety of
@@
 
+\subsection{RMT and Neurobiological Signals}
+
 In recent years, RMT analyses have been applied to study brain functioning. The earliest study
 demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) \citenum{sebaRandomMatrixAnalysis2003}.
+resemble those of the Gaussian Orthogonal Ensemble (GOE) \cite{sebaRandomMatrixAnalysis2003}.
 Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
-fMRI data \citenum{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
+fMRI data \cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
 been used in diffusion MRI to aid in the selection of the number of components to employ in
 principal-component reduction analysis and denoising
-\citenum{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
+\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
 
 Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states \citenum{wangSpectralPropertiesTemporal2015}, between
-subjects with and without ADHD \citenum{wangRandomMatrixTheory2016}, and between pain and non-pain
-states \citenum{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
+differences between rest and task states \cite{wangSpectralPropertiesTemporal2015}, between
+subjects with and without ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain
+states \cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
 resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
@@
 from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
-interpretable manner across psychological processes. If this is the case, RMT could have unique
-potential in characterizing the functioning of the human brain.
+interpretable manner across psychological processes. If this is the case, RMT could aid
+in understanding the functioning of the human brain.
+
+\subsection{Eigenfeatures}
+
+The basic insight of RMT is thus that the \textit{eigenvalues alone} may provide interesting
+information about highly complex Neurobiological systems. However, in practice, the eigenvalues computed from
+empirically observed data are noisy. RMT is statistical in nature, describing only the
+\textit{expected} behaviour of the system spectra. To this end, a number of summary statistics or
+``spectral observables'' \cite{mehtaRandomMatrices2004} can be computed from empirically observed
+spectra, with these spectral observables sometimes being better suited for further analysis or
+comparison to theory. Two such summary statistics that have been somewhat popular are the
+\textit{spectral rigidity} and \textit{level number variance} (``rigidity'' and ``level variance''
+for short; details in Section \ref{sect:methods}).
+
+However, from a predictive standpoint, these summary statistics may eliminate predictively-useful
+information. If the basic RMT insight is that the eigenvalues are useful features to understand a
+system, then those eigenvalues alone (or other simple transformations of them) ought also to be
+predictively useful. We examine a number of such \textit{eigenfeatures} in this study.
+
+\subsection{Limitations of Previous Work}
+
+However, these previous studies took largely descriptive / explanatory approaches, noting only
+differences in RMT across subgroups and/or conditions, in many cases, with no formal statistical
+model or testing of the significance and robustness of these differences. No code was provided to
+allow the reproduction of the results, and the extraction of the spectral rigidity and level
+variance is non-trivial, and computationally demanding. Computation of these spectral observables
+requires an \textit{unfolding}
+procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}, which is,
+unfortunately, a poorly-documented fitting procedure that requires making a number of
+difficult-to-justify decisions, and these decisions can dramatically impact RMT conclusions
+\cite{abul-magdUnfoldingSpectrumChaotic2014,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012,moralesImprovedUnfoldingDetrending2011}.
+On top of the flexibility in the implementation and application of RMT to empirical data, there
+is also the flexibility introduced by the choices in the complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
+
+What is missing is a rigorous, systematic, reproducible investigation of the *predictice* value of
+RMT metrics across a wide variety of data, RMT analytic choices, choice of eigenfeature, and choice
+of preprocessing. The current study aims to rectify this.
 
@@
 
-As our primary intention is to examine the potential utility of RMT-derived features across a wide
+To examine the potential utility of RMT-derived features across a wide
 variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
@@
 their findings with respect to attentional processes
-\citenum{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
+\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
 datasets that involve attention in some way. This includes measures of attention or vigilance, or
@@
 \subsection{Dataset: PSYCH}
-The PSYCH Dataset  is detailed in \citenum{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
+The PSYCH Dataset  is detailed in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
 battery of psychological measurements, and various resting-state fMRI (rs-fMRI) whole-brain 7-Tesla
@@
 \emph{Trait Attention}.
-The PANAS-X \citenum{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
+The PANAS-X \cite{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
 queries the extent to which subjects agree particular affective terms have applied to them over a
@@
 \emph{Task Attention}.
-The Conjunctive Continuous Performance Task (CCPT;
-\citenum{shalevConjunctiveContinuousPerformance2011} is a behavioural measure of sustained attention
+The Conjunctive Continuous Performance Task \cite{shalevConjunctiveContinuousPerformance2011} (CCPT)
+ is a behavioural measure of sustained attention
 where subjects must react only if they see a red square among other different coloured figures. For
@@
 state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
-from German; \citenum{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
+from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
 alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
@@
 Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
-\citenum{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
+\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
 condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
@@
 documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
-complementary to \citenum{matharooSpontaneousBackpainAlters2020}.
+complementary to \cite{matharooSpontaneousBackpainAlters2020}.
 
@@
 Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
-cognitive tasks) is described in full detail by \citenum{dupreMultiechoFMRIReplication2016}. Both
+cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
 multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
-involved viewing various affective images (see \citenum{sprengPatternsBrainActivity2010} for details)
+involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
 and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
@@
 the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast \citenum{kunduMultiechoFMRIReview2017}. In the ``interleave''
+image with overall increased contrast \cite{kunduMultiechoFMRIReview2017}. In the ``interleave''
 condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
@@
 Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
-described in detail in \citenum{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
+described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
 in Table \ref{table:1}. Subjects with non-demented Parkinson's Disease (PD) and healthy controls performed a
 number of repetitions of the Attention Network Task
-(ANT; \citenum{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
+(ANT; \cite{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
 task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
@@
 Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
-detail in \citenum{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
+detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
 were available for all subjects, allowing for within-subject comparisons. The complex task involved
 multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
-related by family-resemblance of the components see figure 1 of \citenum{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
+related by family-resemblance of the components see figure 1 of \cite{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
 memory of the various learned features. Scans corresponding to both the initial learning and
@@
     ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
-    Task = fMRI task. ANT = Attention Network Task \citenum{fanActivationAttentionalNetworks2005}
+    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
 }
@@
 \section{Methods}
+\label{sect:methods}
+
+Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity, \(\Delta_3(L)\), is
+calculated for any positive real value \(L < \max(\Lambda)\) as:
+\[
+\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
+\]
+where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
+\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
+\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
+line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The spectral
+rigidty for a value \(L\) is thus the \textit{average nonlinearity}
+
+The level
+number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
+\cite{mehtaRandomMatrices2004}, and is calculated as:
+
+\[
+\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
+\]
+
+where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
+and \(\langle \cdot \rangle_c\) are as above \cite{guhrRandommatrixTheoriesQuantum1998a}. Both the
+spectral rigidity and level number variance are also computed and investigated in this study.
 

### 2022-10-31 code/scripts/summarize_everything_updated.py
[WIP]

@@ -314,7 +314,7 @@ def summarize_performance_by_aggregation(
     df = df.loc[~df.data.str.contains("Ses")]
 
     # cleaner view
-    """
+    # """
     sbn.displot(
         data=df,
         x="auroc",
@@ -339,10 +339,10 @@ def summarize_performance_by_aggregation(
             label="guess" if i == 0 else None,
         )
     plt.show()
-    """
+    # """
 
     # violin plot
-    """
+    # """
     sbn.catplot(
         data=df,
         y="auroc",
@@ -367,28 +367,29 @@ def summarize_performance_by_aggregation(
         )
     resize_fig()
     plt.show()
-    """
+    # """
 
     # # Very useful: shows best RMT performance due to outlier performances
-    # sbn.catplot(
-    #     data=df,
-    #     y="auroc",
-    #     x="subgroup",
-    #     hue="feature_group",
-    #     palette=FEATURE_GROUP_PALETTE,
-    #     hue_order=list(FEATURE_GROUP_PALETTE.keys()),
-    #     kind="box",
-    #     col="preproc",
-    #     linewidth=0.5,
-    #     fliersize=0.75,  # outlier markers
-    #     legend_out=False,
-    # )
-    # fig = plt.gcf()
-    # # sbn.move_legend(fig, loc="upper right")
-    # for ax in fig.axes:
-    #     ax.set_xticklabels(ax.get_xticklabels(), rotation=45)
-    # resize_fig()
-    # plt.show()
+    sbn.catplot(
+        data=df,
+        y="auroc",
+        x="subgroup",
+        hue="feature_group",
+        palette=FEATURE_GROUP_PALETTE,
+        hue_order=list(FEATURE_GROUP_PALETTE.keys()),
+        kind="box",
+        col="preproc",
+        col_order=PREPROC_ORDER,
+        linewidth=0.5,
+        fliersize=0.75,  # outlier markers
+        legend_out=False,
+    )
+    fig = plt.gcf()
+    # sbn.move_legend(fig, loc="upper right")
+    for ax in fig.axes:
+        ax.set_xticklabels(ax.get_xticklabels(), rotation=45)
+    resize_fig()
+    plt.show()
 
     # Compare by classifier: AUROC < 0.5 = overfitting
     sbn.catplot(
@@ -401,6 +402,7 @@ def summarize_performance_by_aggregation(
         hue_order=list(FEATURE_GROUP_PALETTE.keys()),
         kind="box",
         row="preproc",
[…]

### 2022-11-01 paper/paper/SPIE/article.tex
WIP on paper

@@
 
+With the increase and exapansion of digitial and electronic
+Extracting useful features from complex data is a perennial challenge.
+
 In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
@@
 
-In recent years, RMT analyses have been applied to study brain functioning. The earliest study
+Recently, RMT has been applied to study brain functioning. The earliest study
 demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
@@
 differences in RMT across subgroups and/or conditions, in many cases, with no formal statistical
-model or testing of the significance and robustness of these differences. No code was provided to
+model nor formal testing of the significance and robustness of these differences. No code was provided to
 allow the reproduction of the results, and the extraction of the spectral rigidity and level

### 2022-11-01 code/rmt/enumerables.py
implement timeseries extraction

@@ -1,6 +1,9 @@
 from enum import Enum
 from pathlib import Path
-from typing import List, Optional
+from typing import Callable, List, Optional
+
+import numpy as np
+from numpy import ndarray
 
 ROOT = Path(__file__).resolve().parent.parent.parent
 DATA = ROOT / "data/updated"
@@ -217,6 +220,55 @@ class TrimMethod(Enum):
 
 class NormMethod(Enum):
     pass
+
+
+class SeriesKind(Enum):
+    # measures of location
+    Mean = "mean"
+    Max = "max"
+    Min = "min"
+    Median = "median"
+    Percentile95 = "p95"
+    Percentile05 = "p05"
+    # measures of scale
+    StdDev = "sd"
+    IQR = "iqr"
+    Range = "range"
+    RobustRange = "r90"
+
+    def suffix(self) -> str:
+        return {
+            SeriesKind.Mean: "tseries.mean.npy",
+            SeriesKind.Max: "tseries.max.npy",
+            SeriesKind.Min: "tseries.min.npy",
+            SeriesKind.Median: "tseries.median.npy",
+            SeriesKind.Percentile95: "tseries.p95.npy",
+            SeriesKind.Percentile05: "tseries.p05.npy",
+            SeriesKind.StdDev: "tseries.sd.npy",
+            SeriesKind.IQR: "tseries.iqr.npy",
+            SeriesKind.Range: "tseries.range.npy",
+            SeriesKind.RobustRange: "tseries.r90.npy",
+        }[self]
+
+    def reducer(self) -> Callable[[ndarray], ndarray]:
+        return {
+            # measures of location
+            SeriesKind.Mean: lambda x: np.mean(x, axis=0),
+            SeriesKind.Max: lambda x: np.max(x, axis=0),
+            SeriesKind.Min: lambda x: np.min(x, axis=0),
+            SeriesKind.Median: lambda x: np.median(x, axis=0),
+            SeriesKind.Percentile95: lambda x: np.percentile(x, 95, axis=0),
+            SeriesKind.Percentile05: lambda x: np.percentile(x, 5, axis=0),
+            # measures of scale
+            SeriesKind.StdDev: lambda x: np.std(x, ddof=1, axis=0),
+            SeriesKind.IQR: lambda x: np.abs(
+                np.diff(np.percentile(x, q=[25, 75], axis=0))
+            ),
[…]

### 2022-11-01 code/rmt/preprocess/extract.py

@@ -26,24 +26,21 @@ from numpy import ndarray
 from tqdm.contrib.concurrent import process_map
 
 from rmt.constants import DATA_ROOT as DATA
-from rmt.constants import (
-    EIGS_SUFFIX,
-    EXTRACTED_SUFFIX,
-    MASK_SUFFIX,
-    MNI_REGISTERED_SUFFIX,
-    MOTION_CORRECTED_SUFFIX,
-    NII_SUFFIX,
-    RMT_DIR,
-    SLICETIME_SUFFIX,
-    UPDATED,
-)
+from rmt.constants import EIGS_SUFFIX, NII_SUFFIX, UPDATED
+from rmt.enumerables import SeriesKind
 
 if os.environ.get("CC_CLUSTER") == "niagara":
     ROOT = Path("/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri").resolve()
     DATA = Path("/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri/data").resolve()
-    UPDATED = Path("/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri/data/updated").resolve()
-    RMT_DIR = Path("/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri/data/updated/rmt").resolve()
-    os.environ["MPLCONFIGDIR"] = str(Path("/gpfs/fs0/scratch/j/jlevman/dberger/.mplconfig"))
+    UPDATED = Path(
+        "/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri/data/updated"
+    ).resolve()
+    RMT_DIR = Path(
+        "/gpfs/fs0/scratch/j/jlevman/dberger/random-matrix-fmri/data/updated/rmt"
+    ).resolve()
+    os.environ["MPLCONFIGDIR"] = str(
+        Path("/gpfs/fs0/scratch/j/jlevman/dberger/.mplconfig")
+    )
 
 
 class Loadable(ABC):
@@ -58,6 +55,39 @@ class Loadable(ABC):
 class RMTComputatable(Loadable):
     def __init__(self, source: Path) -> None:
         super().__init__(source)
+
+    def compute_timeseries(self, force: bool = False) -> None:
+        for kind in SeriesKind:
+            long_outfile = (
+                str(self.source)
+                .replace(NII_SUFFIX, kind.suffix())
+                .replace("data/updated", "data/updated/rmt")
+                .replace("func/", "")
+            )
[…]

### 2022-11-01 code/rmt/preprocess/extract.py
[WIP]

@@ -148,8 +148,22 @@ def compute_eigs(path: Path) -> None:
         print(f"Got error: {e}")
 
 
+def compute_all_tseries(path: Path) -> None:
+    try:
+        rmt = RMTComputatable(path)
+        rmt.compute_timeseries()
+    except Exception as e:
+        traceback.print_exc()
+        print(f"Got error: {e}")
+
+
 if __name__ == "__main__":
     paths = get_paths()
     process_map(
-        compute_eigs, paths, desc="Computing eigenvalues", chunksize=1, max_workers=12
+        # compute_eigs, paths, desc="Computing eigenvalues", chunksize=1, max_workers=12
+        compute_all_tseries,
+        paths,
+        desc="Computing timeseries",
+        chunksize=1,
+        max_workers=12,
     )

### 2022-11-01 code/rmt/preprocess/extract.py
[SHARE]

@@ -75,7 +75,7 @@ class RMTComputatable(Loadable):
             mask = np.var(arr2d, axis=1) > 0
             arr2d = arr2d[mask]
 
-            reduced: ndarray = kind.reduce(arr2d).ravel()
+            reduced: ndarray = kind.reducer(arr2d).ravel()
             if len(reduced) != arr2d.shape[1]:
                 raise RuntimeError(
                     f"Shape mismatch for series type: {kind}. "

### 2022-11-01 code/rmt/preprocess/extract.py
[SHARE]

@@ -75,7 +75,8 @@ class RMTComputatable(Loadable):
             mask = np.var(arr2d, axis=1) > 0
             arr2d = arr2d[mask]
 
-            reduced: ndarray = kind.reducer(arr2d).ravel()
+            reducer = kind.reducer()
+            reduced: ndarray = reducer(arr2d).ravel()
             if len(reduced) != arr2d.shape[1]:
                 raise RuntimeError(
                     f"Shape mismatch for series type: {kind}. "

### 2022-11-01 paper/paper/SPIE/article.tex
WIP - timeseries prediction

@@
 
+In addition these issues stemming from ``research degrees of freedom'' (cite Gelman), there is also
+a basic baseline comparison issue: RMT is mathematically non-trivial, the extraction of RMT
+spectral statistics is computationally demanding, and eigenvalues, while easy to understand
+mathematically, are difficult to interpret relative to simpler potential reductions of fMRI data.
+
 What is missing is a rigorous, systematic, reproducible investigation of the *predictice* value of
 RMT metrics across a wide variety of data, RMT analytic choices, choice of eigenfeature, and choice
-of preprocessing. The current study aims to rectify this.
+of preprocessing, and for these RMT eigenfeatures to be compared to simpler alternative baselines.
+The current study aims to do just this.
 

### 2022-11-01 code/rmt/enumerables.py

@@ -20,6 +20,55 @@ Rest_w_VigilanceAttention/
 """
 
 
+class SeriesKind(Enum):
+    # measures of location
+    Mean = "mean"
+    Max = "max"
+    Min = "min"
+    Median = "median"
+    Percentile95 = "p95"
+    Percentile05 = "p05"
+    # measures of scale
+    StdDev = "sd"
+    IQR = "iqr"
+    Range = "range"
+    RobustRange = "r90"
+
+    def suffix(self) -> str:
+        return {
+            SeriesKind.Mean: ".tseries.mean.npy",
+            SeriesKind.Max: ".tseries.max.npy",
+            SeriesKind.Min: ".tseries.min.npy",
+            SeriesKind.Median: ".tseries.median.npy",
+            SeriesKind.Percentile95: ".tseries.p95.npy",
+            SeriesKind.Percentile05: ".tseries.p05.npy",
+            SeriesKind.StdDev: ".tseries.sd.npy",
+            SeriesKind.IQR: ".tseries.iqr.npy",
+            SeriesKind.Range: ".tseries.range.npy",
+            SeriesKind.RobustRange: ".tseries.r90.npy",
+        }[self]
+
+    def reducer(self) -> Callable[[ndarray], ndarray]:
+        return {
+            # measures of location
+            SeriesKind.Mean: lambda x: np.mean(x, axis=0),
+            SeriesKind.Max: lambda x: np.max(x, axis=0),
+            SeriesKind.Min: lambda x: np.min(x, axis=0),
+            SeriesKind.Median: lambda x: np.median(x, axis=0),
+            SeriesKind.Percentile95: lambda x: np.percentile(x, 95, axis=0),
+            SeriesKind.Percentile05: lambda x: np.percentile(x, 5, axis=0),
+            # measures of scale
+            SeriesKind.StdDev: lambda x: np.std(x, ddof=1, axis=0),
+            SeriesKind.IQR: lambda x: np.abs(
+                np.diff(np.percentile(x, q=[25, 75], axis=0), axis=0)
+            ),
+            SeriesKind.Range: lambda x: np.max(x, axis=0) - np.min(x, axis=0),
+            SeriesKind.RobustRange: lambda x: np.abs(
+                np.diff(np.percentile(x, q=[5, 95], axis=0), axis=0)
+            ),
+        }[self]
+
+
 class PreprocLevel(Enum):
     BrainExtract = 0
     SliceTimeAlign = 1
[…]

### 2022-11-01 code/rmt/tseries_predict.py

@@ -1,0 +1,476 @@
+# fmt: off
+import sys  # isort: skip
+from pathlib import Path  # isort: skip
+ROOT = Path(__file__).resolve().parent.parent
+sys.path.append(str(ROOT))
+# fmt: on
+
+import traceback
+from abc import ABC, abstractproperty
+from argparse import ArgumentParser, Namespace
+from dataclasses import dataclass
+from enum import Enum
+from itertools import combinations
+from math import ceil
+from pathlib import Path
+from pprint import pprint
+from typing import (
+    Any,
+    Callable,
+    Dict,
+    List,
+    Mapping,
+    Optional,
+    Sequence,
+    Tuple,
+    Type,
+    Union,
+    cast,
+    no_type_check,
+)
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import pytest
+import seaborn as sbn
+from matplotlib.axes import Axes
+from numpy import ndarray
+from pandas import DataFrame, Series
+from sklearn.ensemble import GradientBoostingClassifier as GBC
+from sklearn.ensemble import RandomForestClassifier as RF
+from sklearn.model_selection import (
+    ParameterGrid,
+    StratifiedKFold,
+    cross_val_score,
+    cross_validate,
+)
+from sklearn.neighbors import KNeighborsClassifier as KNN
+from sklearn.preprocessing import LabelEncoder, MinMaxScaler, OneHotEncoder, minmax_scale
+from sklearn.svm import SVC
+from tqdm import tqdm
+from tqdm.contrib.concurrent import process_map
+from typing_extensions import Literal
+
+from rmt.enumerables import PreprocLevel, SeriesKind, UpdatedDataset
+from rmt.updated_dataset import UpdatedProcessedDataset
+from rmt.updated_features import UpdatedFeature
+
+PROJECT = ROOT.parent
+RESULTS = PROJECT / "results"
+
+# from source code
+KNN_DEFAULTS = dict(
+    weights="uniform",
+    algorithm="auto",
+    leaf_size=30,
+    p=2,
+    metric="minkowski",
+    metric_params=None,
+    n_jobs=None,
+)
+
+
+class KNN3(KNN):
+    def __init__(self) -> None:
+        super().__init__(n_neighbors=3, **KNN_DEFAULTS)
+
+
+class KNN5(KNN):
+    def __init__(self) -> None:
[…]

### 2022-11-01 code/rmt/updated_dataset.py

@@ -28,7 +28,7 @@ from tqdm import tqdm
 from tqdm.contrib.concurrent import process_map
 from typing_extensions import Literal
 
-from rmt.enumerables import PreprocLevel, TrimMethod, UpdatedDataset
+from rmt.enumerables import PreprocLevel, SeriesKind, TrimMethod, UpdatedDataset
 from rmt.preprocess.unify_attention_data import get_comparison_df, median_split_labels
 
 CACHE_DIR = ROOT.parent / "__OBSERVABLES_CACHE__"
@@ -72,9 +72,12 @@ class UpdatedProcessedDataset:
         # self.path_info = data[self.source]
         self.id = f"{self.source.name}__preproc={self.preproc_level.name}"
 
-    def get_information_frame(self) -> DataFrame:
+    def get_information_frame(self, tseries: SeriesKind | None = None) -> DataFrame:
         """Get a DataFrame linking each path to all phenotypic info"""
-        files = self.source.eig_files(self.preproc_level)
+        if tseries is None:
+            files = self.source.eig_files(self.preproc_level)
+        else:
+            files = self.source.tseries_files(self.preproc_level, tseries)
         dfs = []
 
         if self.source is UpdatedDataset.Older:
@@ -603,7 +606,7 @@ def print_subgroup_counts() -> None:
         print(data.source.name)
         print(data.info.groupby(["label"]).count()["sid"])
         print(data.info.drop(columns=["session", "run"]).groupby(["sid"]).count())
-        print("="*80)
+        print("=" * 80)
 
 
 if __name__ == "__main__":

### 2022-11-01 paper/paper/SPIE/article.tex
[WIP]

@@
 
-With the increase and exapansion of digitial and electronic
-Extracting useful features from complex data is a perennial challenge.
+\textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go back and add
+the citations and ajust the tone.
+
+Random Matrix Theory (RMT) has obvious potential. It has helped to understand the generalization
+behaviour of neural networks [add citation], can be used to guide any principal components analysis
+reduction [cite: Google search: "random matrix" "principal components"], has been used to help
+understand phenomena as diverse as [include the usual citations to buses, traffic, physics, etc],
+and can provide insights wherever there is a system of components interacting with some
+independence [cite Cambridge RMT handbook, others].
+
+At its most rudimentary, RMT simply describes the statistically-expected behaviour of the
+eigenvalues of certain classes—or \textit{ensembles}—of random matrices under certain assumptions or
+conditions. A random matrix is, generally, a matrix in which the entries can be considered to be
+independently and identically distributed (iid), and where the iid distribution is a classic
+mathemtical distribution. For example, one of the simplest ensembles [cite handbook] is the Gaussian
+Unitary Ensemble (GUE), wherein [desc of ensemble properties] and which appears to be useful in
+describing [real-world examples].
+
+This means that if we examine a system and find that the distribution of eigenvalues of that system
+resemble some RMT predicted spectrum, we potentially gain insight into the nature of that system.
+If, in addition, we find a large deviance from the spectrum predicted by RMT, we find contrapositive
+evidence that the system is \textit{not} like those described by RMT. Roughly, systems
+well-described by RMT are highyl random and independent, whereas systems that are not so
+well-described are either highly dependent, non-random, or both.
+
+Despite the mathematical elegance and simplicity, RMT has yielded insights in variosu complex domains.
 

### 2022-11-01 code/scripts/summarize_everything_updated.py

@@ -1129,6 +1129,9 @@ if __name__ == "__main__":
     simplefilter(action="ignore", category=PerformanceWarning)
     pd.options.display.max_rows = 1000
     pd.options.display.max_info_rows = 1000
+    df = load_all_renamed()
+    df.to_json(PROJECT / "EVERYTHING.json")
+    print(f"Saved all combined data to {PROJECT / 'EVERYTHING.json'}")
     # pd.options.display.large_repr = "info"
 
     # print_correlations(by=["data", "classifier", "preproc", "feature"])
@@ -1138,7 +1141,7 @@ if __name__ == "__main__":
     # get_overfit_scores()
     # summarize_performance_by_aggregation(metric="auroc", summarizer="median")
 
-    generate_all_topk_plots()
+    # generate_all_topk_plots()
     sys.exit()
 
     # df = load_all_renamed()

### 2022-11-01 paper/paper/SPIE/article.tex
[WIP]

@@
 If, in addition, we find a large deviance from the spectrum predicted by RMT, we find contrapositive
-evidence that the system is \textit{not} like those described by RMT. Roughly, systems
-well-described by RMT are highyl random and independent, whereas systems that are not so
+evidence that the system is \textit{not} like those described by RMT. Roughly, complex systems
+well-described by RMT are highly random and independent, whereas systems that are not so
 well-described are either highly dependent, non-random, or both.
+
+In this sense, RMT has considerable \textit{explanatory} and \textit{statistical} value: the more we
+see that a system resembles an RMT system, the less we can distinguish it from one of random,
+non-interacting components, and the less the system resembles an RMT system, the more
+fundamentally complex and interactive the interactions.
+
+For a system as complex as the human brain, RMT is thus tantalizing. It would be quite something
+if we could extract some matrix reduction of brain activity, and if the RMT description of this
+matrix either aligned nicely with RMT theory, or otherwise was broadly predictively useful.
 

### 2022-11-01 paper/paper/SPIE/article.tex
[WIP]

@@
 
-In this sense, RMT has considerable \textit{explanatory} and \textit{statistical} value: the more we
+In this sense, RMT might have real \textit{explanatory} and \textit{statistical} value: the more we
 see that a system resembles an RMT system, the less we can distinguish it from one of random,
@@
 
-Despite the mathematical elegance and simplicity, RMT has yielded insights in variosu complex domains.
+[Transition to fMRI now...]
 

### 2022-11-04 paper/paper/SPIE/article.tex
[WIP]

@@
 \usepackage{tocloft}
+\usepackage{doi}
 
@@
 
+Also
+
 \section*{Conclusions}
@@
 \textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go back and add
-the citations and ajust the tone.
+the citations and adjust the tone.
 
@@
 At its most rudimentary, RMT simply describes the statistically-expected behaviour of the
-eigenvalues of certain classes—or \textit{ensembles}—of random matrices under certain assumptions or
+eigenvalues of some classes—or \textit{ensembles}—of random matrices under certain assumptions or
 conditions. A random matrix is, generally, a matrix in which the entries can be considered to be
-independently and identically distributed (iid), and where the iid distribution is a classic
-mathemtical distribution. For example, one of the simplest ensembles [cite handbook] is the Gaussian
+independently and identically distributed (\textit{iid}). For example, one of the simplest ensembles [cite handbook] is the Gaussian
 Unitary Ensemble (GUE), wherein [desc of ensemble properties] and which appears to be useful in
@@
 
-\subsection{Eigenfeatures}
+\subsection{Motivating Eigenfeatures}
 
@@
 predictively useful. We examine a number of such \textit{eigenfeatures} in this study.
+
+In addition, the functional connectivity - the matrix of correlations of various collections of
+voxels of an fMRI scan - is a priori a useful representation of the fMRI data. However, the full
+correlation matrix between all voxels is itself often computationally infeasible to work with, and
+thus is reduced in various ways prior to being used in analyses.
+
+Typical reductions include the use of a ``seed'' voxel or collection of voxels (i.e. ROI): the mean
+signal over that ROI is correlated with all other N ROIs of interest to generate a reduced functional
+connectivity matrix (or image) with one correlation value at each non-seed ROI. This sort of reduction reduces the functional connectivity to N values. This approach strongly
+biases the analysis to the seed ROI.
+
+Likewise, one can work with ROI mean signals, and work with the \(N \times N\) matrix of
+these ROI correlations. This analysis [maybe cite Poldrack book for this?] is, \textit{a priori},
+sensitive to the choice of ROIs, and, in the case of anatomical atlases / parcellations, may also
+lack theoretical justification.
+
+A major disadvantage of both the above approaches, besides requiring parcellation into ROIs, is
+that most such parcellation methods are atlas-based, and thus require image registration, and,
+general, an entire preprocessing pipeline to ensure that registration is successful.
+
+RMT suggests a potentially useful reduction in the form of the eigenvalues of the voxelwise
+functional connectivity. An \(N \times t\) matrix \(\bf{A}\) of \(N\) time series of length \(t\),
+and where \(t \ll N\) will have a symmetric correlation (or covariance) matrix with \(t - 1\)
+non-zero, positive real-valued eigenvalues \(\Lambda\). These eigenvalues can be computed highly
+efficiently via transposition of the voxelwise correlation or covariance matrix.
 
@@
 
-To examine the potential utility of RMT-derived features across a wide
-variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
-psychological processes, we examine a wide selection of fMRI datasets available on the
-\href{https://openneuro.org/}{OpenNeuro} platform (36).  This allows for a broad assessment of the
-potential of voxelwise RMT features.
+\subsection{Overview}
 
-However, since previous studies employing RMT for feature extraction have partially interpreted
-their findings with respect to attentional processes
-\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
-datasets that involve attention in some way. This includes measures of attention or vigilance, or
-comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
-which should involve less attention or focus (e.g. resting-state, pain medication vs. none).
+In order not to inundate readers with dataset details, and because this is an exploratory
+investigation of RMT which is not committed to any specific theory, we refer the reader
+to the original publications and/or data releases when greater detail is needed, and
+highlight only the most basic aspects of each here.
 
-\subsection{Dataset: PSYCH}
+We selected fMRI data publicly available on the \href{https://openneuro.org/}{OpenNeuro} platform
+(36). Selection criteria was somewhat subjective, but we required (1), that all scans have the same
+spatial and temporal resolutions, (2), the the dataset comprise 10+ human subjects, and (3), that it
+was reasonable to classify at the level of an entire scan, i.e. neither task nor event timings were
+needed to determine class membership. This yielded seven datasets comprising XX total subjects, and
+over 5000 total fMRI volumes.
+
+Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
+and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
+(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
+these datasets, we highlight below only those aspects relevant to the current paper.
+
+\subsubsection{Aging}
+This dataset\cite{wahlheimIntrinsicFunctionalConnectivity2021} included 34 subjects with a mean age
+of 22 years (range: 18-32 years), and 28 subjects with a mean age of 70 years (range: 61-80 years).
+
+% The final sample included 34 younger adults (1832 years old, Mage = 22.21, SD = 3.65; 20 female)
+% and 28 older adults (61-80 years old, Mage = 69.82, SD = 5.64; 20 female). Table 1 shows that
+% younger and older adults had comparable education and working memory capacity, the latter measured
+% by forward and backward digit span (48). Older adults had higher vocabulary scores (49) and slower
+% processing speed (50) than younger adults.
+% Resting-state data comprised 300 measurements collected over one 10-minute run. This scan duration
+% produces reliable functional connectivity estimates within and across participants (57).
+% Participants were instructed to remain still and awake with their eyes open. No stimuli appeared
+% during this scan, and the display was black. Functional scans were collected using an echo-planar
+% image sequence sensitive to blood-oxygen-level-dependent (BOLD) contrast (T2*; 32 slices with 4.0
+% mm thickness and no skip, TE = 30 ms, TR = 2000 ms, flip angle = 70, FOV = 220 mm, matrix size =
+% 74 × 74 × 32 voxels, A/P phase encoding direction). Slices were collected in a descending order
+% that covered the entire cortex and partial cerebellum. At the beginning of the resting-state scan,
+% the scanner acquired and discarded two dummy scans.
+
+\subsubsection{Bilinguality}
+
+This dataset\cite{ds001747:1.0.0, goldcarrieelizabethExploringRestingState2018} examined English and
+Spanish-speaking monolinguals and multilinguals during a prolonged resting state. The original study
+grouped participants into three classes: early vs late bilinguals vs monolingual controls, and found
+connectivity differences only between the early and late
+bilinguals\cite{goldcarrieelizabethExploringRestingState2018}. We compared for our two classes just
+"bilingual" vs "monolingual". The data is most notable for the short TR (875ms) combined with long
+acquisition time (12 minutes).
+
+% Resting state fMRI scans were performed using a T2*-weighted EPI sequence with the following
+% parameters: TR=875ms; TE=43.6ms; slice thickness=1.8mm; acquisition matrix=100 × 100; flip
+% angle=55°; number of slices=72; field of view=180 × 180mm; voxel size=1.8 × 1.8 × 1.8mm;
+% multi-band factor=8; volumes per run=823 (total scan time=12min). EPI scans were oriented
+% parallel to the long axis of the hippocampus.
+
+\subsubsection{Learning}
+
+This dataset\cite{schapiroHumanHippocampalReplay2018, schapiroHumanHippocampalReplay2020}. In this study, both rs-fMRI and task fMRI scans
+were available for all subjects, allowing for within-subject comparisons. The task is complex and difficult to summarize here adequately, but involved
+multiple phases where subjects were presented with various related abstract images, and then later tested for their memory of certain aspects of the learned images.
+For our study, we simply note that these task scans ought to be broadly more attentionally  and cognitively demanding than the resting state, and instead simply try to classify task vs.
+test scans using eigenfeatures.
+
+\subsubsection{Dataset: PSYCH}
 The PSYCH Dataset  is detailed in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
@@
 
-\subsubsection{Dataset 3}
-Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
-cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
-multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
-involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
-and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
-Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
-different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
-resting-state scans. This resulted in four task state scans and one resting-state scan per subject.
-
-In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
-the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
-image with overall increased contrast \cite{kunduMultiechoFMRIReview2017}. In the ``interleave''
-condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
-from echo \(j\) as \(t_{j,i}\), then the interleaved voxel of the new image has the time-series \([t_{1,1},
-t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).
-
-This creates a physically-meaningful 4D image that contains the full, original multi-echo fMRI
-time-series data without any alterations. It should be noted that the purpose of these combination
-methods is not to create a 4D volume that allows optimal detection of the purely task-linked BOLD
-signal, but to ensure all scan information was used with minimal additional analytic assumptions.
-
 \subsubsection{Dataset 4}
@@
 \subsubsection{Dataset 5}
-Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
-detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
-were available for all subjects, allowing for within-subject comparisons. The complex task involved
-multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
-related by family-resemblance of the components see figure 1 of \cite{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
-memory of the various learned features. Scans corresponding to both the initial learning and
-subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of
-the analysis of this paper. Resting state scans were included in the low-attention ``rest''
-subgroup.
-
-Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
-and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
-(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
-these datasets, we highlight below only those aspects relevant to the current paper.
 
@@
 spectral rigidity and level number variance are also computed and investigated in this study.
+
+\section{Results}
+\label{sect:results}
+
+First not non-predictable data, and then no longer discuss it in subsequent features.
+
+\subsection{Normalization}
+NO meaningful effect of normalization on AUROC distributions at level of gross feature, or feature
+group, so can safely ignore this.
+NO meaningful effect of normalization for any feature group by classifier either, so
+normalization doesn't seem to matter here.
+Maybe add table of correlations by groupings here
+
+\subsection{Classifier}
+When the data is predictable, timeseries features almost all have the mode / max <= 0.5 for all
+classifiers and subgroups, whereas eigenfeatures (RMT, eigenvalue only, eigenvalue + RMT) more often
+than not have mode greater than 0.5, and bulk of the distribution > 0.5.
+
+KNN3 consistently has least favourable AUROC distributions across subgroups and features
 

### 2022-11-04 code/scripts/summarize_everything_updated.py

@@ -5,6 +5,7 @@ ROOT = Path(__file__).resolve().parent.parent
 sys.path.append(str(ROOT))
 # fmt: on
 
+import re
 from typing import Literal
 from warnings import simplefilter
 
@@ -19,7 +20,9 @@ from matplotlib.lines import Line2D
 from matplotlib.patches import Patch
 from pandas import DataFrame
 from pandas.errors import PerformanceWarning
+from scipy.optimize import NonlinearConstraint, minimize, minimize_scalar
 from seaborn import FacetGrid
+from statsmodels.distributions.empirical_distribution import ECDF
 from tqdm import tqdm
 
 from rmt.updated_features import FEATURE_OUTFILES as PATHS
@@ -135,6 +138,20 @@ SUBGROUP_ORDER = [
     # "WeeklyAttentionSes2 - trait_attend v trait_nonattend",
 ]
 
+"""Defined as comparisons where bulk (>50%) of AUROC distribution is right of
+0.5 under at least one preproc regime
+"""
+OVERALL_PREDICTIVE_GROUP_ORDER = [
+    "Learning - rest v task",
+    "Aging - younger v older",
+    "Osteo - nopain v duloxetine",
+    "Osteo - nopain v pain",
+    "Osteo - pain v duloxetine",
+    "Parkinsons - ctrl v park",  # questionable, only tseries seem predictive
+    "TaskAttention - task_attend v task_nonattend",
+    "Vigilance - vigilant v nonvigilant",
+]
+
 
 CLASSIFIER_ORDER = [
     # "RandomForestClassifier",
@@ -151,6 +168,10 @@ PREPROC_ORDER = [
     "SliceTimeAlign",
     "MotionCorrect",
     "MNIRegister",
+]
+NORM_ORDER = [
+    False,
+    True,
 ]
 
 
@@ -236,6 +257,12 @@ def load_tseries() -> DataFrame:
     df["slice"] = "all"
     df["trim"] = "none"
     df.loc[:, "data"] = df["data"].str.replace("Older", "Aging")
+    df.loc[:, "classifier"] = (
+        df["classifier"].str.replace("GradientBoostingClassifier", "GBDT").copy()
+    )
+    df.loc[:, "classifier"] = (
+        df["classifier"].str.replace("RandomForestClassifier", "RF").copy()
+    )
     return df
 
 
@@ -1243,17 +1270,20 @@ def print_correlations(by: list[str]) -> None:
 
 
 def clean_titles(
[…]

### 2022-11-04 paper/paper/SPIE/article.tex
[WIP]

@@
 \section*{Purpose}
-Random matrix theory (RMT) is an increasingly useful tool for understanding large, complex systems.
-Prior studies have examined functional Magnetic Resonance Imaging (fMRI) scans using tools from
-RMT, with some success. However, RMT computations are complex, and sensitive to numerous analytic
-choices, and previous investigations have been mostly explanatory / descriptive. We systematically
-investigate the usefulness of RMT in a variety of fMRI datasets in a predictive framework.
+Random matrix theory (RMT) is an increasingly useful tool for understanding
+large, complex systems. Prior studies have examined functional Magnetic
+Resonance Imaging (fMRI) scans using tools from RMT, with some success.
+However, RMT computations are highly sensitive to a number of analytic choices,
+and the robustness of findings involving RMT remain in question. We
+systematically investigate the usefulness of RMT on a wide variety of fMRI
+datasets using a rigorous predictive framework.
 
 \section*{Approach}
-We develop open-source software to efficiently compute RMT features from fMRI images, and examine
-the cross-validated predictive potential of eigenvalue and RMT-based features (``eigenfeatures'') used
-with typical classificatory machine-learning algorithms. We systematically vary pre-processing extent,
-normalization procedure, RMT unfolding procedure, and classifiers to examine the utility of these
-eigenfeatures across various datasets.
+We develop open-source software to efficiently compute RMT features from fMRI
+images, and examine the cross-validated predictive potential of eigenvalue and
+RMT-based features (``eigenfeatures'') with classic machine-learning
+classifiers. We systematically vary pre-processing extent, normalization
+procedures, RMT unfolding procedures, and feature selection, and compare the
+impact of these analytic choices on the distributions of cross-validated
+prediction performance for each combination of dataset binary classification
+task, classifier, and feature. To deal with class imbalance, we use the
+area under the receiver operating characteristic (AUROC) as the main
+performance metric.
 
 \section*{Results}
-We find eiegenfeatures to be broadly predictive in a wide variety of contexts, and that RMT-specific
-features like the spectral rigidity and level number variance have considerable potential in
-feature-engineering for fMRI. On median, we find ... However, we also find the utility of the
-extracted eigenfeatures to depend strongly and complexly on various analytic decisions.
-
-Also
+% df.groupby(["subgroup", "gross_feature"])["auroc"].median()
+Across all classification tasks and analytic choices, we find eiegenfeatures to
+more often have predictive utility than not (82.4\% of median AUROCs > 0.5;
+median AUROC range across classification tasks 0.47 - 0.64). Simple baseline
+reductions on source timeseries, by contrast, were less useful (58.8\% of
+median AUROCs > 0.5, median AUROC range across classification tasks 0.42 -
+0.62). Additionally, eigenfeature AUROC distributions were overall much more
+right-tailed than baseline features suggesting greater predictive potential.
+However, performance distributions were wide, and often significantly affected by
+analytic choices.
 
 \section*{Conclusions}
-Consider using RMT.
+Eigenfeatures clearly have potential for understanding and reducing fMRI
+functional connectivity. However, the utility of these features is strongly
+dependent on analytic decisions, suggesting considerable caution when
+interprting past and future studies applying RMT.
 

### 2022-11-04 code/scripts/summarize_everything_updated.py

@@ -237,6 +237,10 @@ def load_all_renamed() -> DataFrame:
         df.comparison == "trait_nonattend v trait_attend"
     )
     df = df.loc[~dupes]
+    df["subgroup"] = df["data"] + " - " + df["comparison"]
+    df["feature_group"] = df["feature"].apply(feature_grouping)
+    df["slice_group"] = df["slice"].apply(slice_grouping)
+    df["gross_feature"] = df["feature"].apply(gross_feature_grouping)
     return df
 
 
@@ -263,13 +267,28 @@ def load_tseries() -> DataFrame:
     df.loc[:, "classifier"] = (
         df["classifier"].str.replace("RandomForestClassifier", "RF").copy()
     )
+    dupes = (df.data == "TaskAttentionSes2") & (
+        df.comparison == "task_nonattend v task_attend"
+    )
+    df = df.loc[~dupes]
+    dupes = (df.data == "WeeklyAttentionSes2") & (
+        df.comparison == "trait_nonattend v trait_attend"
+    )
+    df = df.loc[~dupes]
+    df["subgroup"] = df["data"] + " - " + df["comparison"]
+    df["feature_group"] = df["feature"].apply(feature_grouping)
+    df["slice_group"] = df["slice"].apply(slice_grouping)
+    df["gross_feature"] = df["feature"].apply(gross_feature_grouping)
     return df
 
 
-def load_combined() -> DataFrame:
+def load_combined(drop_ses: bool = True) -> DataFrame:
     df = load_all_renamed()
     ts = load_tseries()
-    return pd.concat([df, ts], axis=0)
+    df = pd.concat([df, ts], axis=0)
+    if drop_ses:
+        df = df.loc[~df.data.str.contains("Ses")]
+    return df
 
 
 @MEMORY.cache
@@ -1405,13 +1424,8 @@ def make_kde_plots() -> None:
     sbn.set_style("ticks")
 
     print("Loading data...", end="", flush=True)
-    print(" done")
     df = load_combined()
-    df["subgroup"] = df["data"] + " - " + df["comparison"]
-    df["feature_group"] = df["feature"].apply(feature_grouping)
-    df["slice_group"] = df["slice"].apply(slice_grouping)
-    df["gross_feature"] = df["feature"].apply(gross_feature_grouping)
-    df = df.loc[~df.data.str.contains("Ses")]
+    print(" done")
 
[…]

### 2022-11-04 paper/paper/SPIE/article.tex
[WIP]

@@
 \usepackage{doi}
+
+\newcommand{\code}[1]{\small \texttt{#1} \normalsize}
 
@@
 
-% \begin{abstract}
-% This document shows the required format and appearance of a manuscript prepared for SPIE journals.
-% It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
-% following journals require the use of structured abstracts in manuscript submissions:
-% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
-% Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
-% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
-% structured or single-paragraph, the abstract should be a summary of the paper and not an
-% introduction. Because the abstract may be used in abstracting and indexing databases, it should be
-% self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
-% objectives, methodology used, results obtained, and their significance. A list of up to six
-% keywords should immediately follow.
-% \end{abstract}
+% \begin{abstract} This document shows the required format and appearance of a
+% manuscript prepared for SPIE journals. It is prepared using LaTeX2e with the
+% class file \texttt{spieman.cls}. Please note that the following journals
+% require the use of structured abstracts in manuscript submissions:
+% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the
+% \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for
+% the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Guidelines
+% are available on the journal website. Whether structured or single-paragraph,
+% the abstract should be a summary of the paper and not an introduction.
+% Because the abstract may be used in abstracting and indexing databases, it
+% should be self-contained (i.e., no numerical references) and substantive in
+% nature, presenting concisely the objectives, methodology used, results
+% obtained, and their significance. A list of up to six keywords should
+% immediately follow. \end{abstract}
 
-% See https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
+% See
+% https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
 % for info on the structured abstract
 %
-% Purpose: This section presents the significance and aims, stating the broad impact and the
-% rationale or motivation of the work, including potentially some background.
+% Purpose: This section presents the significance and aims, stating the broad
+% impact and the rationale or motivation of the work, including potentially
+% some background.
 %
-% Approach: This section briefly describes the materials and methods, including the statistical
-% analyses, used in the research.
+% Approach: This section briefly describes the materials and methods, including
+% the statistical analyses, used in the research.
 %
-% Results: This section provides a core summary of the findings from the work, including study
-% numbers, quantitative analyses, or discoveries.
+% Results: This section provides a core summary of the findings from the work,
+% including study numbers, quantitative analyses, or discoveries.
 %
-% Conclusions: This section summarizes and interprets the approach and findings from the work,
-% stating also the importance or impact of the findings.
+% Conclusions: This section summarizes and interprets the approach and findings
+% from the work, stating also the importance or impact of the findings.
 
-% Purpose:
-% Data-intensive modeling could provide insight on the broad variability in outcomes in
-% spine surgery. Previous studies were limited to analysis of demographic and clinical
-% characteristics. We report an analytic framework called “SpineCloud” that incorporates
-% quantitative features extracted from perioperative images to predict spine surgery outcome.
+% Purpose: Data-intensive modeling could provide insight on the broad
+% variability in outcomes in spine surgery. Previous studies were limited to
+% analysis of demographic and clinical characteristics. We report an analytic
+% framework called “SpineCloud” that incorporates quantitative features
+% extracted from perioperative images to predict spine surgery outcome.
 
-% Approach:
-% A retrospective study was conducted in which patient demographics, imaging, and outcome
-% data were collected. Image features were automatically computed from perioperative CT.
-% Postoperative 3- and 12-month functional and pain outcomes were analyzed in terms of improvement
-% relative to the preoperative state. A boosted decision tree classifier was trained to predict
-% outcome using demographic and image features as predictor variables. Predictions were computed
-% based on SpineCloud and conventional demographic models, and features associated with poor outcome
-% were identified from weighting terms evident in the boosted tree.
+% Approach: A retrospective study was conducted in which patient demographics,
+% imaging, and outcome data were collected. Image features were automatically
+% computed from perioperative CT. Postoperative 3- and 12-month functional and
+% pain outcomes were analyzed in terms of improvement relative to the
+% preoperative state. A boosted decision tree classifier was trained to predict
+% outcome using demographic and image features as predictor variables.
+% Predictions were computed based on SpineCloud and conventional demographic
+% models, and features associated with poor outcome were identified from
+% weighting terms evident in the boosted tree.
 
-% Results
-% Neither approach was predictive of 3- or 12-month outcomes based on preoperative data
-% alone in the current, preliminary study. However, SpineCloud predictions incorporating image
-% features obtained during and immediately following surgery (i.e., intraoperative and immediate
-% postoperative images) exhibited significant improvement in area under the receiver operating
-% characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3 months and AUC=0.69 (CI95=0.55
-% to 0.82) at 12 months.
+% Results Neither approach was predictive of 3- or 12-month outcomes based on
+% preoperative data alone in the current, preliminary study. However,
+% SpineCloud predictions incorporating image features obtained during and
+% immediately following surgery (i.e., intraoperative and immediate
+% postoperative images) exhibited significant improvement in area under the
+% receiver operating characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3
+% months and AUC=0.69 (CI95=0.55 to 0.82) at 12 months.
 
-% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved by incorporation of
-% image-based features compared to analysis based on conventional demographic data. The SpineCloud
-% framework could improve understanding of factors underlying outcome variability and warrants
-% further investigation and validation in a larger patient cohort.
+% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved
+% by incorporation of image-based features compared to analysis based on
+% conventional demographic data. The SpineCloud framework could improve
+% understanding of factors underlying outcome variability and warrants further
+% investigation and validation in a larger patient cohort.
 \begin{abstract}
@@
 Across all classification tasks and analytic choices, we find eiegenfeatures to
-more often have predictive utility than not (82.4\% of median AUROCs > 0.5;
+more often have predictive utility than not (82.4\% of median AUROCs \(>\) 0.5;
 median AUROC range across classification tasks 0.47 - 0.64). Simple baseline
 reductions on source timeseries, by contrast, were less useful (58.8\% of
-median AUROCs > 0.5, median AUROC range across classification tasks 0.42 -
+median AUROCs \(>\) 0.5, median AUROC range across classification tasks 0.42 -
 0.62). Additionally, eigenfeature AUROC distributions were overall much more
@@
 
-% This document shows the required format and appearance of a manuscript prepared for SPIE journals.
-% It is prepared using LaTeX2e with the class file \texttt{spieman.cls}. Please note that the
-% following journals require the use of structured abstracts in manuscript submissions:
-% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the \textit{Journal of
-% Medical Imaging}. Structured abstracts are encouraged for the \textit{Journal of
-% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the journal website. Whether
-% structured or single-paragraph, the abstract should be a summary of the paper and not an
-% introduction. Because the abstract may be used in abstracting and indexing databases, it should be
-% self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the
-% objectives, methodology used, results obtained, and their significance. A list of up to six keywords
-% should immediately follow.
+% This document shows the required format and appearance of a manuscript
+% prepared for SPIE journals. It is prepared using LaTeX2e with the class file
+% \texttt{spieman.cls}. Please note that the following journals require the use
+% of structured abstracts in manuscript submissions: \textit{Neurophotonics},
+% the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical
+% Imaging}. Structured abstracts are encouraged for the \textit{Journal of
+% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the
+% journal website. Whether structured or single-paragraph, the abstract should
+% be a summary of the paper and not an introduction. Because the abstract may
+% be used in abstracting and indexing databases, it should be self-contained
+% (i.e., no numerical references) and substantive in nature, presenting
+% concisely the objectives, methodology used, results obtained, and their
+% significance. A list of up to six keywords should immediately follow.
 \end{abstract}
@@
 \label{sect:intro}  % \label{} allows reference to this section
-% This document shows the format and appearance of a manuscript prepared for submission to an SPIE
-% journal. Note that this template is only intended to be used as a guideline for author convenience.
-% It is designed for optimum clarity and ease of reading for editors and reviewers, but the template
-% does not reflect the final page layout of a published journal paper. Accepted papers are
-% professionally typeset in XML according to the layout and design of the journal.
+% This document shows the format and appearance of a manuscript prepared for
+% submission to an SPIE journal. Note that this template is only intended to be
+% used as a guideline for author convenience. It is designed for optimum
+% clarity and ease of reading for editors and reviewers, but the template does
+% not reflect the final page layout of a published journal paper. Accepted
+% papers are professionally typeset in XML according to the layout and design
+% of the journal.
 
-\textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go back and add
-the citations and adjust the tone.
+\textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go
+back and add the citations and adjust the tone.
 
-Random Matrix Theory (RMT) has obvious potential. It has helped to understand the generalization
-behaviour of neural networks [add citation], can be used to guide any principal components analysis
-reduction [cite: Google search: "random matrix" "principal components"], has been used to help
-understand phenomena as diverse as [include the usual citations to buses, traffic, physics, etc],
-and can provide insights wherever there is a system of components interacting with some
-independence [cite Cambridge RMT handbook, others].
+Random Matrix Theory (RMT) has obvious potential. It has helped to understand
+the generalization behaviour of neural networks [add citation], can be used to
+guide any principal components analysis reduction [cite: Google search: "random
+matrix" "principal components"], has been used to help understand phenomena as
+diverse as [include the usual citations to buses, traffic, physics, etc], and
+can provide insights wherever there is a system of components interacting with
+some independence [cite Cambridge RMT handbook, others].
 
-At its most rudimentary, RMT simply describes the statistically-expected behaviour of the
-eigenvalues of some classes—or \textit{ensembles}—of random matrices under certain assumptions or
-conditions. A random matrix is, generally, a matrix in which the entries can be considered to be
-independently and identically distributed (\textit{iid}). For example, one of the simplest ensembles [cite handbook] is the Gaussian
-Unitary Ensemble (GUE), wherein [desc of ensemble properties] and which appears to be useful in
-describing [real-world examples].
+At its most rudimentary, RMT simply describes the statistically-expected
+behaviour of the eigenvalues of some classes—or \textit{ensembles}—of random
+matrices under certain assumptions or conditions. A random matrix is,
+generally, a matrix in which the entries can be considered to be independently
+and identically distributed (\textit{iid}). For example, one of the simplest
+ensembles [cite handbook] is the Gaussian Unitary Ensemble (GUE), wherein [desc
+of ensemble properties] and which appears to be useful in describing
+[real-world examples].
 
-This means that if we examine a system and find that the distribution of eigenvalues of that system
-resemble some RMT predicted spectrum, we potentially gain insight into the nature of that system.
-If, in addition, we find a large deviance from the spectrum predicted by RMT, we find contrapositive
-evidence that the system is \textit{not} like those described by RMT. Roughly, complex systems
-well-described by RMT are highly random and independent, whereas systems that are not so
-well-described are either highly dependent, non-random, or both.
+This means that if we examine a system and find that the distribution of
+eigenvalues of that system resemble some RMT predicted spectrum, we potentially
+gain insight into the nature of that system. If, in addition, we find a large
+deviance from the spectrum predicted by RMT, we find contrapositive evidence
+that the system is \textit{not} like those described by RMT. Roughly, complex
+systems well-described by RMT are highly random and independent, whereas
+systems that are not so well-described are either highly dependent, non-random,
+or both.
 
-In this sense, RMT might have real \textit{explanatory} and \textit{statistical} value: the more we
-see that a system resembles an RMT system, the less we can distinguish it from one of random,
-non-interacting components, and the less the system resembles an RMT system, the more
+In this sense, RMT might have real \textit{explanatory} and
+\textit{statistical} value: the more we see that a system resembles an RMT
+system, the less we can distinguish it from one of random, non-interacting
+components, and the less the system resembles an RMT system, the more
 fundamentally complex and interactive the interactions.
 
-For a system as complex as the human brain, RMT is thus tantalizing. It would be quite something
-if we could extract some matrix reduction of brain activity, and if the RMT description of this
-matrix either aligned nicely with RMT theory, or otherwise was broadly predictively useful.
+For a system as complex as the human brain, RMT is thus tantalizing. It would
+be quite something if we could extract some matrix reduction of brain activity,
+and if the RMT description of this matrix either aligned nicely with RMT
+theory, or otherwise was broadly predictively useful.
 
@@
 
-In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
-(BOLD) signals are related to neural activity. It is common to investigate statistical relationships
-between the BOLD signals through \textbf{functional connectivity} analyses, where correlations between
-collections of these signals are examined to infer connections between different voxels or regions
-of interest (ROIs) within the brain.
+In functional magnetic resonance imaging (fMRI), changes in the
+blood-oxygenation-level-dependent (BOLD) signals are related to neural
+activity. It is common to investigate statistical relationships between the
+BOLD signals through \textbf{functional connectivity} analyses, where
+correlations between collections of these signals are examined to infer
+connections between different voxels or regions of interest (ROIs) within the
+brain.
 
-Whether in the presence of experimental stimuli, or the relative absence, as in resting state fMRI (rs-fMRI),
-complex patterns in functional connectivity networks are ubiquitous
+Whether in the presence of experimental stimuli, or the relative absence, as in
+resting state fMRI (rs-fMRI), complex patterns in functional connectivity
+networks are ubiquitous
 \cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
-This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
-of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
-physics \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}
+This complexity suggests fMRI is a candidate to be studied using \textbf{Random
+Matrix Theory} (RMT), a set of mathematical tools originally developed some 50
+years ago to solve complex problems in nuclear physics
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}
 
-In various
-physical systems where nuclei can have different possible states, precisely modeling the entirety of
-the interactions between nuclei is often intractable. For such large, complex systems, the
-eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
-of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
-spectrum statistically, complex systems can be investigated by comparison to the universal
-properties predicted by RMT.
+In various physical systems where nuclei can have different possible states,
+precisely modeling the entirety of the interactions between nuclei is often
+intractable. For such large, complex systems, the eigenvalues (or spectra)
+efficiently summarize the totality of the interactions between components of
+the system, and RMT describes the expected behaviour of such eigenvalues. By
+analysing the spectrum statistically, complex systems can be investigated by
+comparison to the universal properties predicted by RMT.
 
@@
 Recently, RMT has been applied to study brain functioning. The earliest study
-demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
-resemble those of the Gaussian Orthogonal Ensemble (GOE) \cite{sebaRandomMatrixAnalysis2003}.
-Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
-fMRI data \cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
-been used in diffusion MRI to aid in the selection of the number of components to employ in
-principal-component reduction analysis and denoising
+demonstrated that spectra of the correlations between electroencephalographic
+(EEG) signals closely resemble those of the Gaussian Orthogonal Ensemble (GOE)
+\cite{sebaRandomMatrixAnalysis2003}. Somewhat more recently, RMT was used to
+evaluate the quality of whole brain features extracted from fMRI data
+\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has
+also been used in diffusion MRI to aid in the selection of the number of
+components to employ in principal-component reduction analysis and denoising
 \cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
 
-Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
-differences between rest and task states \cite{wangSpectralPropertiesTemporal2015}, between
-subjects with and without ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain
-states \cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
-resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
-certain aspects of psychological processes might be characterized, in part, by features computed
-from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
-interpretable manner across psychological processes. If this is the case, RMT could aid
-in understanding the functioning of the human brain.
+Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to
+investigate differences between rest and task states
+\cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
+ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain states
+\cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the
+spectra of resting or low-attention states exhibited properties closest to the
+GOE. These findings suggest that certain aspects of psychological processes
+might be characterized, in part, by features computed from the eigenvalues of
+fMRI correlation matrices, and that these features might vary in an
+interpretable manner across psychological processes. If this is the case, RMT
+could aid in understanding the functioning of the human brain.
 
@@
 
-The basic insight of RMT is thus that the \textit{eigenvalues alone} may provide interesting
-information about highly complex Neurobiological systems. However, in practice, the eigenvalues computed from
-empirically observed data are noisy. RMT is statistical in nature, describing only the
-\textit{expected} behaviour of the system spectra. To this end, a number of summary statistics or
-``spectral observables'' \cite{mehtaRandomMatrices2004} can be computed from empirically observed
-spectra, with these spectral observables sometimes being better suited for further analysis or
-comparison to theory. Two such summary statistics that have been somewhat popular are the
-\textit{spectral rigidity} and \textit{level number variance} (``rigidity'' and ``level variance''
-for short; details in Section \ref{sect:methods}).
+The basic insight of RMT is thus that the \textit{eigenvalues alone} may
+provide interesting information about highly complex Neurobiological systems.
+However, in practice, the eigenvalues computed from empirically observed data
+are noisy. RMT is statistical in nature, describing only the \textit{expected}
+behaviour of the system spectra. To this end, a number of summary statistics or
+``spectral observables'' \cite{mehtaRandomMatrices2004} can be computed from
+empirically observed spectra, with these spectral observables sometimes being
+better suited for further analysis or comparison to theory. Two such summary
+statistics that have been somewhat popular are the \textit{spectral rigidity}
+and \textit{level number variance} (``rigidity'' and ``level variance'' for
+short; details in Section \ref{sect:methods}).
 
-However, from a predictive standpoint, these summary statistics may eliminate predictively-useful
-information. If the basic RMT insight is that the eigenvalues are useful features to understand a
-system, then those eigenvalues alone (or other simple transformations of them) ought also to be
-predictively useful. We examine a number of such \textit{eigenfeatures} in this study.
+However, from a predictive standpoint, these summary statistics may eliminate
+predictively-useful information. If the basic RMT insight is that the
+eigenvalues are useful features to understand a system, then those eigenvalues
+alone (or other simple transformations of them) ought also to be predictively
+useful. We examine a number of such \textit{eigenfeatures} in this study.
 
-In addition, the functional connectivity - the matrix of correlations of various collections of
-voxels of an fMRI scan - is a priori a useful representation of the fMRI data. However, the full
-correlation matrix between all voxels is itself often computationally infeasible to work with, and
-thus is reduced in various ways prior to being used in analyses.
+In addition, the functional connectivity - the matrix of correlations of
+various collections of voxels of an fMRI scan - is a priori a useful
+representation of the fMRI data. However, the full correlation matrix between
+all voxels is itself often computationally infeasible to work with, and thus is
+reduced in various ways prior to being used in analyses.
 
-Typical reductions include the use of a ``seed'' voxel or collection of voxels (i.e. ROI): the mean
-signal over that ROI is correlated with all other N ROIs of interest to generate a reduced functional
-connectivity matrix (or image) with one correlation value at each non-seed ROI. This sort of reduction reduces the functional connectivity to N values. This approach strongly
-biases the analysis to the seed ROI.
+Typical reductions include the use of a ``seed'' voxel or collection of voxels
+(i.e. ROI): the mean signal over that ROI is correlated with all other N ROIs
+of interest to generate a reduced functional connectivity matrix (or image)
+with one correlation value at each non-seed ROI. This sort of reduction reduces
+the functional connectivity to N values. This approach strongly biases the
+analysis to the seed ROI.
 
-Likewise, one can work with ROI mean signals, and work with the \(N \times N\) matrix of
-these ROI correlations. This analysis [maybe cite Poldrack book for this?] is, \textit{a priori},
-sensitive to the choice of ROIs, and, in the case of anatomical atlases / parcellations, may also
-lack theoretical justification.
+Likewise, one can work with ROI mean signals, and work with the \(N \times N\)
+matrix of these ROI correlations. This analysis [maybe cite Poldrack book for
+this?] is, \textit{a priori}, sensitive to the choice of ROIs, and, in the case
+of anatomical atlases / parcellations, may also lack theoretical justification.
 
-A major disadvantage of both the above approaches, besides requiring parcellation into ROIs, is
-that most such parcellation methods are atlas-based, and thus require image registration, and,
-general, an entire preprocessing pipeline to ensure that registration is successful.
+A major disadvantage of both the above approaches, besides requiring
+parcellation into ROIs, is that most such parcellation methods are atlas-based,
+and thus require image registration, and, general, an entire preprocessing
+pipeline to ensure that registration is successful.
 
-RMT suggests a potentially useful reduction in the form of the eigenvalues of the voxelwise
-functional connectivity. An \(N \times t\) matrix \(\bf{A}\) of \(N\) time series of length \(t\),
-and where \(t \ll N\) will have a symmetric correlation (or covariance) matrix with \(t - 1\)
-non-zero, positive real-valued eigenvalues \(\Lambda\). These eigenvalues can be computed highly
-efficiently via transposition of the voxelwise correlation or covariance matrix.
+RMT suggests a potentially useful reduction in the form of the eigenvalues of
+the voxelwise functional connectivity. An \(N \times t\) matrix \(\bf{A}\) of
+\(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
+correlation (or covariance) matrix with \(t - 1\) non-zero, positive
+real-valued eigenvalues \(\Lambda\). These eigenvalues can be computed highly
+efficiently via transposition of the voxelwise correlation or covariance
+matrix.
 
@@
 
-However, these previous studies took largely descriptive / explanatory approaches, noting only
-differences in RMT across subgroups and/or conditions, in many cases, with no formal statistical
-model nor formal testing of the significance and robustness of these differences. No code was provided to
-allow the reproduction of the results, and the extraction of the spectral rigidity and level
-variance is non-trivial, and computationally demanding. Computation of these spectral observables
-requires an \textit{unfolding}
-procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}, which is,
-unfortunately, a poorly-documented fitting procedure that requires making a number of
-difficult-to-justify decisions, and these decisions can dramatically impact RMT conclusions
+However, these previous studies took largely descriptive / explanatory
+approaches, noting only differences in RMT across subgroups and/or conditions,
+in many cases, with no formal statistical model nor formal testing of the
+significance and robustness of these differences. No code was provided to allow
+the reproduction of the results, and the extraction of the spectral rigidity
+and level variance is non-trivial, and computationally demanding. Computation
+of these spectral observables requires an \textit{unfolding}
+procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004},
+which is, unfortunately, a poorly-documented fitting procedure that requires
+making a number of difficult-to-justify decisions, and these decisions can
+dramatically impact RMT conclusions
 \cite{abul-magdUnfoldingSpectrumChaotic2014,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012,moralesImprovedUnfoldingDetrending2011}.
-On top of the flexibility in the implementation and application of RMT to empirical data, there
-is also the flexibility introduced by the choices in the complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
+On top of the flexibility in the implementation and application of RMT to
+empirical data, there is also the flexibility introduced by the choices in the
+complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
 
-In addition these issues stemming from ``research degrees of freedom'' (cite Gelman), there is also
-a basic baseline comparison issue: RMT is mathematically non-trivial, the extraction of RMT
-spectral statistics is computationally demanding, and eigenvalues, while easy to understand
-mathematically, are difficult to interpret relative to simpler potential reductions of fMRI data.
+In addition these issues stemming from ``research degrees of freedom'' (cite
+Gelman), there is also a basic baseline comparison issue: RMT is mathematically
+non-trivial, the extraction of RMT spectral statistics is computationally
+demanding, and eigenvalues, while easy to understand mathematically, are
+difficult to interpret relative to simpler potential reductions of fMRI data.
 
-What is missing is a rigorous, systematic, reproducible investigation of the *predictice* value of
-RMT metrics across a wide variety of data, RMT analytic choices, choice of eigenfeature, and choice
-of preprocessing, and for these RMT eigenfeatures to be compared to simpler alternative baselines.
-The current study aims to do just this.
+What is missing is a rigorous, systematic, reproducible investigation of the
+*predictice* value of RMT metrics across a wide variety of data, RMT analytic
+choices, choice of eigenfeature, and choice of preprocessing, and for these RMT
+eigenfeatures to be compared to simpler alternative baselines. The current
+study aims to do just this.
 
 \section{Datasets}
+\label{sec:datasets}
 
@@
 
-In order not to inundate readers with dataset details, and because this is an exploratory
-investigation of RMT which is not committed to any specific theory, we refer the reader
-to the original publications and/or data releases when greater detail is needed, and
-highlight only the most basic aspects of each here.
+We selected fMRI data publicly available on the
+\href{https://openneuro.org/}{OpenNeuro} platform (36). Selection criteria was
+somewhat subjective, but we required (1), that all scans have the same spatial
+and temporal resolutions, (2), the the dataset comprise 10+ human subjects, and
+(3), that it was reasonable to classify at the level of an entire scan, i.e.
+neither task nor event timings were needed to determine class membership. This
+yielded seven datasets comprising XX total subjects, and over 5000 total fMRI
+volumes.
 
-We selected fMRI data publicly available on the \href{https://openneuro.org/}{OpenNeuro} platform
-(36). Selection criteria was somewhat subjective, but we required (1), that all scans have the same
-spatial and temporal resolutions, (2), the the dataset comprise 10+ human subjects, and (3), that it
-was reasonable to classify at the level of an entire scan, i.e. neither task nor event timings were
-needed to determine class membership. This yielded seven datasets comprising XX total subjects, and
-over 5000 total fMRI volumes.
+In order not to inundate readers with dataset details, and because this is an
+exploratory investigation of RMT which is not committed to any specific
+theory\footnote{In fact, datasets and, later, the classification tasks, were
+chosen without investigation into any findings or publications associated with
+the data, both in order to reduce bias in our decisions, and because the
+original authors analytic / statistical approaches and intentions have very
+limited relevance to our whole-brain, voxelwise, multiverse predictive
+approach.}, we refer the reader to the original publications and/or data
+releases when greater detail is needed, and highlight only the most basic
+aspects of each dataset here.
 
-Dataset scan parameters and acquisition details are summarized in Table \ref{table:1}, and sample
-and subgroup sizes are summarized in Table \ref{table:2}. For all datasets, we use the raw
-(unpreprocessed) scans, and perform our own preprocessing (detailed later). Given the complexity of
-these datasets, we highlight below only those aspects relevant to the current paper.
+Dataset scan parameters and acquisition details are summarized in Table
+\ref{table:1}, and sample and subgroup sizes are summarized in Table
+\ref{table:2}.
 
-\subsubsection{Aging}
-This dataset\cite{wahlheimIntrinsicFunctionalConnectivity2021} included 34 subjects with a mean age
-of 22 years (range: 18-32 years), and 28 subjects with a mean age of 70 years (range: 61-80 years).
+\subsection{Datasets and Classification Tasks}
 
-% The final sample included 34 younger adults (1832 years old, Mage = 22.21, SD = 3.65; 20 female)
-% and 28 older adults (61-80 years old, Mage = 69.82, SD = 5.64; 20 female). Table 1 shows that
-% younger and older adults had comparable education and working memory capacity, the latter measured
-% by forward and backward digit span (48). Older adults had higher vocabulary scores (49) and slower
-% processing speed (50) than younger adults.
-% Resting-state data comprised 300 measurements collected over one 10-minute run. This scan duration
-% produces reliable functional connectivity estimates within and across participants (57).
-% Participants were instructed to remain still and awake with their eyes open. No stimuli appeared
-% during this scan, and the display was black. Functional scans were collected using an echo-planar
-% image sequence sensitive to blood-oxygen-level-dependent (BOLD) contrast (T2*; 32 slices with 4.0
-% mm thickness and no skip, TE = 30 ms, TR = 2000 ms, flip angle = 70, FOV = 220 mm, matrix size =
-% 74 × 74 × 32 voxels, A/P phase encoding direction). Slices were collected in a descending order
-% that covered the entire cortex and partial cerebellum. At the beginning of the resting-state scan,
-% the scanner acquired and discarded two dummy scans.
+The \textit{Aging} dataset\cite{wahlheimIntrinsicFunctionalConnectivity2021}
+included rs-fMRI data for 34 subjects with a mean age of 22 years (range: 18-32
+years), and 28 subjects with a mean age of 70 years (range: 61-80 years). For
+our analysis, we make the binary classification task the prediction of subject
+age-group membership, i.e. \code{younger v older}.
 
-\subsubsection{Bilinguality}
+% The final sample included 34 younger adults (1832 years old, Mage = 22.21, SD
+% = 3.65; 20 female) and 28 older adults (61-80 years old, Mage = 69.82, SD =
+% 5.64; 20 female). Table 1 shows that younger and older adults had comparable
+% education and working memory capacity, the latter measured by forward and
+% backward digit span (48). Older adults had higher vocabulary scores (49) and
+% slower processing speed (50) than younger adults. Resting-state data
+% comprised 300 measurements collected over one 10-minute run. This scan
+% duration produces reliable functional connectivity estimates within and
+% across participants (57). Participants were instructed to remain still and
+% awake with their eyes open. No stimuli appeared during this scan, and the
+% display was black. Functional scans were collected using an echo-planar image
+% sequence sensitive to blood-oxygen-level-dependent (BOLD) contrast (T2*; 32
+% slices with 4.0 mm thickness and no skip, TE = 30 ms, TR = 2000 ms, flip
+% angle = 70, FOV = 220 mm, matrix size = 74 × 74 × 32 voxels, A/P phase
+% encoding direction). Slices were collected in a descending order that covered
+% the entire cortex and partial cerebellum. At the beginning of the
+% resting-state scan, the scanner acquired and discarded two dummy scans.
 
-This dataset\cite{ds001747:1.0.0, goldcarrieelizabethExploringRestingState2018} examined English and
-Spanish-speaking monolinguals and multilinguals during a prolonged resting state. The original study
-grouped participants into three classes: early vs late bilinguals vs monolingual controls, and found
-connectivity differences only between the early and late
-bilinguals\cite{goldcarrieelizabethExploringRestingState2018}. We compared for our two classes just
-"bilingual" vs "monolingual". The data is most notable for the short TR (875ms) combined with long
-acquisition time (12 minutes).
+\subsubsection{Bilinguality Dataset}
+
+The \textit{Bilinguality} dataset\cite{ds001747:1.0.0,
+goldcarrieelizabethExploringRestingState2018} examined English and
+Spanish-speaking monolinguals and multilinguals during a prolonged resting
+state. The original study grouped participants into three subgroups: early vs
+late bilinguals vs monolingual
+controls\cite{goldcarrieelizabethExploringRestingState2018}. For our analysis,
+we make the binary classification task be to classify \code{monolingual v
+bilingual}.
 
@@
 
-\subsubsection{Learning}
+\subsubsection{Depression Dataset}
 
-This dataset\cite{schapiroHumanHippocampalReplay2018, schapiroHumanHippocampalReplay2020}. In this study, both rs-fMRI and task fMRI scans
-were available for all subjects, allowing for within-subject comparisons. The task is complex and difficult to summarize here adequately, but involved
-multiple phases where subjects were presented with various related abstract images, and then later tested for their memory of certain aspects of the learned images.
-For our study, we simply note that these task scans ought to be broadly more attentionally  and cognitively demanding than the resting state, and instead simply try to classify task vs.
-test scans using eigenfeatures.
+\subsubsection{Learning Dataset}
 
-\subsubsection{Dataset: PSYCH}
-The PSYCH Dataset  is detailed in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
-battery of psychological measurements, and various resting-state fMRI (rs-fMRI) whole-brain 7-Tesla
-scans were obtained. For the purpose of keeping with the ``attention'' theme of the analyses in this
-paper, three pairings of subgroups were constructed via z-score splits of the various sum-scores of
-scale items involving concepts closely related to attention.
+The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
+schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans were
+available for all subjects, and \code{task v rest} is chosen as the binary
+classification task for our analysis.  The task is complex and difficult to
+summarize here adequately, but involved multiple phases where subjects were
+presented with various related abstract images, and then later tested for their
+memory of certain aspects of the learned images. Thus the classification task
+might also be considered learning vs. rest.
 
-\emph{Trait Attention}.
-The PANAS-X \cite{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
-queries the extent to which subjects agree particular affective terms have applied to them over a
-recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) scale was constructed from the
-scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
-``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects were then placed in the high / low ``trait
-attention'' category based on a median split of the average z-score across items.
+\subsubsection{Attention Dataset}
 
-\emph{Task Attention}.
-The Conjunctive Continuous Performance Task \cite{shalevConjunctiveContinuousPerformance2011} (CCPT)
- is a behavioural measure of sustained attention
-where subjects must react only if they see a red square among other different coloured figures. For
-this paper, a sum score was constructed by subtracting the number of false positive and false
-negative responses from the number of success responses, and subjects with a z-score greater / less
-than zero on this scale were placed in high / low ``task attention'' category.
+The \textit{Attention} dataset\cite{gorgolewskiHighResolution7Tesla2015} is a
+high-resolution rs-fMRI dataset including a battery of psychological measures
+for each subject. Since previous studies employing RMT sometimes interpreted
+their findings with respect to attentional processes
+\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we
+divided subjects into various high vs. low attention binary classification
+tasks based on median splits of subsets of the metrics available in this study.
 
-\emph{Vigilance}.
-The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
-state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
-from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
-alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
-``vigilance'' conditions in the same manner as the previous two splits.
+The \code{viglant v nonvigilant} task was formed based on self-report
+questionnaire items involving
+"vigilance"\cite{gorgolewskiCorrespondenceIndividualDifferences2014}. The
+\code{trait\_attend v trait\_nonattend} task was formed using PANAS-X
+\cite{watsonPANASXManualPositive1994} items related to self-reported
+wakefulness and attention over the past weeks. Finally, the
+\code{task\_attend v tas\_nonattent} classification task was constructed
+using the scores of The Conjunctive Continuous Performance Task
+(CCPT\cite{shalevConjunctiveContinuousPerformance2011}). This is a behavioural
+performance task which requires the subject to quickly and selectively respond
+to only certain visual stimuli. Details allowing exact reproduction of these
+splits are available in this paper's
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/preprocess/unify\_attention\_data.py}{source
+code}.
 
-\subsubsection{Dataset 2}
-Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
-\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
-condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
-either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
-also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
-subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
-documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
-complementary to \cite{matharooSpontaneousBackpainAlters2020}.
+\subsubsection{Osteoarthritis Dataset}
+The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
+includes whiole-brain rs-fMRI scans of healthy, pain-free controls
+(\code{nopain} condition), and individuals with knee osteoarthritis.
+Osteoarthritic patients were treated for 2 weeks with either placebo
+(\code{pain} condition) or duloxetine (\code{duloxetine} condition).
+We use the three binary classification tasks \code{nopain v duloxetine},
+\code{nopain v pain}, and \code{pain v duloxetine} for our analysis.
 
-\subsubsection{Dataset 4}
-Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
-described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
-in Table \ref{table:1}. Subjects with non-demented Parkinson's Disease (PD) and healthy controls performed a
-number of repetitions of the Attention Network Task
-(ANT; \cite{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
-task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
-groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
-instead differ by virtue of the presence or absence of PD, any results here serve as a useful
-comparison point for the results from studies that should clearly involve attentional differences.
-This dataset also included extensively preprocessed scans, which were included in analyses to
-examine the effects of preprocessing on the extracted RMT features.
-
-\subsubsection{Dataset 5}
+\subsubsection{Parkinson's Dataset}
+The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
+subjects with non-demented Parkinson's Disease (PD) and
+healthy controls performed a number of repetitions of the Attention Network
+Task (ANT; \cite{fanActivationAttentionalNetworks2005}) during scans.
+The classification task for this dataset is Parkinson's versus controls, i.e.
+\code{park v ctrl}.
 
@@
 \end{tabular}
+
+\subsection{Classification Task Subgroups}
+
+For each dataset, we construct subgroups based either on existing labels in the
+data, or for the \textit{Attention} dataset, from psychological measures available
+from the data.
 

### 2022-11-04 paper/paper/SPIE/article.tex
[WIP]

@@
 releases when greater detail is needed, and highlight only the most basic
-aspects of each dataset here.
+aspects of each dataset here.  Dataset scan parameters and acquisition details
+are summarized in Table \ref{table:1}, and sample and subgroup sizes are
+summarized in Table \ref{table:2}.
 
-Dataset scan parameters and acquisition details are summarized in Table
-\ref{table:1}, and sample and subgroup sizes are summarized in Table
-\ref{table:2}.
+\begin{table}[h!]
+\caption{
+    ID = Identifier for paper. TR = Time of Repetition (seconds).
+    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
+    indicate P horizontal / axial slices each with dimensions M x N.
+}
+\label{table:1}
+\small
+\centering
+\begin{tabular}{ l c c c c c c }
+\hline
+\textbf{Name}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
+\hline
+Aging     & 74 × 74 × 32   & 3.0 × 3.0 × 4.0 & 2.0  & 300 & 62  \\
+Bilingual    & 100 × 100 × 72 & 1.8 × 1.8 × 1.8 & 0.88 & 823 & 90  \\
+Depress   & 112 × 112 × 25 & 2.0 × 2.0 × 5.0 & 2.5  & 100 & 72  \\
+Learn     & 64 × 64 × 36   & 3.0 × 3.0 × 3.0 & 2.0  & 195 & 432 \\
+Osteo     & 64 × 64 × 36   & 3.4 × 3.4 × 3.0 & 2.5  & 300 & 74  \\
+Park      & 80 × 80 × 43   & 3.0 × 3.0 × 3.0 & 2.4  & 149 & 552 \\
+Attention & 128 × 128 × 70 & 1.5 × 1.5 × 1.5 & 3.0  & 300 & 90  \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \subsubsection{Depression Dataset}
+
+The \textit{Depression} dataset\cite{bezmaternykhddRestingStateClosed2021,
+bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans frmo
+non-depressed controls and mildy- or moderately-depressed subjects. The our analysis,
+the binary classification task is \code{depress v control}.
+
+% The intergroup study involved 21 healthy volunteers received a compensation
+% for a participation and 51 participants featuring mild depression (F32.0),
+% moderate depression (F32.1), or dysthymia (F34.1, single patient) (see Table
+% 1). The fMRI study was carried out in the International Tomography Center,
+% Novosibirsk, using a 3 T Ingenia scanner (Philips). Functional T2∗-weighted
+% Ssh echo planar imaging scans were acquired using the following parameters:
+% voxel size 2×2×5mm, repetition time/echo time = 2500/35 ms, and fat
+% suppression mode. The reference anatomical image was obtained by the T1W 3D
+% turbo field echo method with a voxel size of 1×1×1mm. The instruction for
+% participants was to lie still with eyes closed for 6 minutes.
 
@@
 
-\begin{table}[h!]
-\caption{
-    ID = Identifier for paper. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P horizontal / axial slices each with dimensions M x N.
-}
-\label{table:1}
-\small
-\centering
-\begin{tabular}{ c c c c c c c }
-\hline
-\textbf{ID}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
-\hline
-AGING     & 74 x 74 x 32   & 3.0 x 3.0 x 4.0 & 2.0  & 300 & 62  \\
-BILING    & 100 x 100 x 72 & 1.8 x 1.8 x 1.8 & 0.88 & 823 & 90  \\
-DEPRESS   & 112 x 112 x 25 & 2.0 x 2.0 x 5.0 & 2.5  & 100 & 72  \\
-LEARN     & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0  & 195 & 432 \\
-OSTEO     & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5  & 300 & 74  \\
-PARK      & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4  & 149 & 552 \\
-PSYCH     & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0  & 300 & 90  \\
-\hline
-\end{tabular}
-
 \subsection{Classification Task Subgroups}
@@
 from the data.
-
-\end{table}
 
@@
         &  younger           & 34      & resting-state         & 1 \\
-\hline
+        &                    &         &                       &   \\
 BILING  &  bilingual          & 59      & resting-state        & 1 \\
         &  monolingual        & 33      & resting-state        & 1 \\
-\hline
+        &                    &         &                       &   \\
 DEPRESS &  depression         & 51      & resting-state        & 1 \\
         &  control            & 21      & resting-state        & 1 \\
-\hline
+        &                    &         &                       &   \\
 LEARN   &  task              & 24      & learn image sketches  & 16 \\
         &  rest              & 24      & resting-state         & 2 \\
-\hline
-OSTEO   &  duloxetine        & 19      & taking duloxetine     & 1 \\
-        &  pain              & 37      & placebo               & 1 \\
-        &  nopain            & 20      & control               & 1 \\
-\hline
+        &                    &         &                       &   \\
+OSTEO   &  duloxetine        & 19      & resting-state         & 1 \\
+        &  pain              & 37      & resting-state         & 1 \\
+        &  nopain            & 20      & resting-state         & 1 \\
+        &                    &         &                       &   \\
 PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
         &  control           & 21      & ANT                   & 12 \\
-\hline
+        &                    &         &                       &   \\
 PSYCH   &  vigilant           & 11     & resting-state        & 3 \\

### 2022-11-05 paper/paper/SPIE/article.tex
[WIP]

@@
 
-In order not to inundate readers with dataset details, and because this is an
-exploratory investigation of RMT which is not committed to any specific
-theory\footnote{In fact, datasets and, later, the classification tasks, were
-chosen without investigation into any findings or publications associated with
-the data, both in order to reduce bias in our decisions, and because the
-original authors analytic / statistical approaches and intentions have very
-limited relevance to our whole-brain, voxelwise, multiverse predictive
-approach.}, we refer the reader to the original publications and/or data
-releases when greater detail is needed, and highlight only the most basic
-aspects of each dataset here.  Dataset scan parameters and acquisition details
-are summarized in Table \ref{table:1}, and sample and subgroup sizes are
-summarized in Table \ref{table:2}.
-
 \begin{table}[h!]
@@
 Attention & 128 × 128 × 70 & 1.5 × 1.5 × 1.5 & 3.0  & 300 & 90  \\
+\hline
+\end{tabular}
+\end{table}
+
+In order not to inundate readers with dataset details, and because this is an
+exploratory investigation of RMT which is not committed to any specific
+theory\footnote{In fact, datasets and, later, the classification tasks, were
+chosen without investigation into any findings or publications associated with
+the data, both in order to reduce bias in our decisions, and because the
+original authors analytic / statistical approaches and intentions have very
+limited relevance to our whole-brain, voxelwise, multiverse predictive
+approach.}, we refer the reader to the original publications and/or data
+releases when greater detail is needed, and highlight only the most basic
+aspects of each dataset here.  Dataset scan parameters and acquisition details
+are summarized in Table \ref{table:1}, and sample and subgroup sizes are
+summarized in Table \ref{table:2}.
+
+\begin{table}[h!]
+\caption{
+    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
+    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
+}
+\label{table:2}
+\small
+\centering
+\begin{tabular}{ l r c c c }
+\hline
+\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
+\hline
+Aging   &  older             & 28      & resting-state        & 1 \\
+        &  younger           & 34      & resting-state        & 1 \\
+        &                    &         &                      &   \\
+Bilingual  &  bilingual          & 59      & resting-state    & 1 \\
+        &  monolingual        & 33      & resting-state       & 1 \\
+        &                    &         &                      &   \\
+Depress &  depression         & 51      & resting-state       & 1 \\
+        &  control            & 21      & resting-state       & 1 \\
+        &                    &         &                      &   \\
+Learn   &  task              & 24      & learn image sketches & 16\\
+        &  rest              & 24      & resting-state        & 2 \\
+        &                    &         &                      &   \\
+Osteo   &  duloxetine        & 19      & resting-state        & 1 \\
+        &  pain              & 37      & resting-state        & 1 \\
+        &  nopain            & 20      & resting-state        & 1 \\
+        &                    &         &                      &   \\
+Park    &  Parkinson's       & 25      & ANT                  & 12\\
+        &  control           & 21      & ANT                  & 12\\
+        &                    &         &                      &   \\
+Attention&  vigilant           & 11     & resting-state       & 2 \\
+        &  nonigilant         & 11     & resting-state        & 2 \\
+        &  trait-attentive     & 11    & resting-state        & 2 \\
+        &  trait-non-attentive & 11    & resting-state        & 2 \\
+        &  task-attentive     & 11     & resting-state        & 2 \\
+        &  task-nonattentive  & 11     & resting-state        & 2 \\
 \hline
@@
 
-\subsubsection{Bilinguality Dataset}
+% \subsubsection{Bilinguality Dataset}
 
@@
 
-\subsubsection{Depression Dataset}
+% \subsubsection{Depression Dataset}
 
@@
 
-\subsubsection{Learning Dataset}
+% \subsubsection{Learning Dataset}
 
@@
 
-\subsubsection{Attention Dataset}
+% \subsubsection{Osteoarthritis Dataset}
+The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
+includes whiole-brain rs-fMRI scans of healthy, pain-free controls
+(\code{nopain} condition), and individuals with knee osteoarthritis.
+Osteoarthritic patients were treated for 2 weeks with either placebo
+(\code{pain} condition) or duloxetine (\code{duloxetine} condition).
+We use the three binary classification tasks \code{nopain v duloxetine},
+\code{nopain v pain}, and \code{pain v duloxetine} for our analysis.
+
+% \subsubsection{Parkinson's Dataset}
+The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
+subjects with non-demented Parkinson's Disease (PD) and
+healthy controls performed a number of repetitions of the Attention Network
+Task (ANT; \cite{fanActivationAttentionalNetworks2005}) during scans.
+The classification task for this dataset is Parkinson's versus controls, i.e.
+\code{park v ctrl}.
+
+% \subsubsection{Attention Dataset}
 
@@
 
-\subsubsection{Osteoarthritis Dataset}
-The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
-includes whiole-brain rs-fMRI scans of healthy, pain-free controls
-(\code{nopain} condition), and individuals with knee osteoarthritis.
-Osteoarthritic patients were treated for 2 weeks with either placebo
-(\code{pain} condition) or duloxetine (\code{duloxetine} condition).
-We use the three binary classification tasks \code{nopain v duloxetine},
-\code{nopain v pain}, and \code{pain v duloxetine} for our analysis.
-
-\subsubsection{Parkinson's Dataset}
-The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
-subjects with non-demented Parkinson's Disease (PD) and
-healthy controls performed a number of repetitions of the Attention Network
-Task (ANT; \cite{fanActivationAttentionalNetworks2005}) during scans.
-The classification task for this dataset is Parkinson's versus controls, i.e.
-\code{park v ctrl}.
-
-\subsection{Classification Task Subgroups}
-
-For each dataset, we construct subgroups based either on existing labels in the
-data, or for the \textit{Attention} dataset, from psychological measures available
-from the data.
-
-\begin{table}[h!]
-\caption{
-    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
-    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
-}
-\label{table:2}
-\small
-\centering
-\begin{tabular}{ l l c c c }
-\hline
-\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
-\hline
-AGING   &  older             & 28      & resting-state         & 1 \\
-        &  younger           & 34      & resting-state         & 1 \\
-        &                    &         &                       &   \\
-BILING  &  bilingual          & 59      & resting-state        & 1 \\
-        &  monolingual        & 33      & resting-state        & 1 \\
-        &                    &         &                       &   \\
-DEPRESS &  depression         & 51      & resting-state        & 1 \\
-        &  control            & 21      & resting-state        & 1 \\
-        &                    &         &                       &   \\
-LEARN   &  task              & 24      & learn image sketches  & 16 \\
-        &  rest              & 24      & resting-state         & 2 \\
-        &                    &         &                       &   \\
-OSTEO   &  duloxetine        & 19      & resting-state         & 1 \\
-        &  pain              & 37      & resting-state         & 1 \\
-        &  nopain            & 20      & resting-state         & 1 \\
-        &                    &         &                       &   \\
-PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
-        &  control           & 21      & ANT                   & 12 \\
-        &                    &         &                       &   \\
-PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
-        &  nonigilant         & 11     & resting-state        & 3 \\
-        &  trait-attentive     & 11    & resting-state        & 3 \\
-        &  trait-non-attentive & 11    & resting-state        & 3 \\
-        &  task-attentive     & 11     & resting-state        & 3 \\
-        &  task-nonattentive  & 11     & resting-state        & 3 \\
-\hline
-\end{tabular}
-\end{table}
-
 \section{Methods}
@@
 
+\subsection{fMRI Preprocessing}
+
+We limited preprocessing to a simple pipeline of, in order, (1) brain
+extraction, (2) slice timing correction, (3), motion correction, and (4),
+registration to the MNI152
+\cite{fonovUnbiasedAverageAgeappropriate2011,fonovUnbiasedNonlinearAverage2009}
+2mm isotropic, asymmetric template version C made available through
+\href{https://www.templateflow.org/}{TemplateFlow}
+\cite{ciricTemplateFlowFAIRsharingMultiscale2021}.
+
+We limited preprocessing methods to these steps sometimes because data required
+for other typical preprocessing steps (e.g. field intensity maps, physiological
+measurements), but also because we saved all pipelined intermediates to later
+compare the effect of increasing degrees of pre-processing (see Section
+\ref{sec:multiverse}), and it was important to limit the number of
+preprocessing degrees to prevent excessive computational costs.
+
+Brain extraction was performed first with BET
+\cite{smithFastRobustAutomated2002}, and then with with
+ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
+by BET. BET results were visually inspected for each image in each dataset, and
+BET fractional anisotropy values were modified to ensure brain extraction was
+acceptable. Motion-correction was performed with FSL's MCFLIRT
+\cite{jenkinsonImprovedOptimizationRobust2002}, and slicetime correction also
+with FSL's \code{slicetimer}\cite{jenkinsonFSL2012} tool. Finally, functional
+images were registered directly to the MNI152 template via
+ANTs\cite{avantsReproducibleEvaluationANTs2011}.
+
+\subsection{Feature Extraction}
+
+For each preprocessing degree and fMRI image, features were extracted from all
+\(N\) non-constant brain voxels, where \(N\) may differ for each image. After
+this process, each scan yields an \(N \times T\) matrix \(\mathbf{M}\), where
+\(T\) is the number of volumes acquired. All features summarize this matrix
+of brain voxels \(\mathbf{M}\) in some manner.
+
+\subsubsection{Baseline Features}
+
+If RMT- or eigenvalue-based features fail to outperform simple reductions of
+the fMRI data in predictive utility, it is difficult to justify the greater
+computational and interpretational complexities of the former. These simple
+statistical reductions help assess the relative value of RMT features.
+
+Baseline reductions compute either a robust measure of location (mean, max,
+min) or non-robust measure of location (median, \(95^{\text{th}}\) percentile,
+\(5^{\text{th}}\) percentile), or a non-robust (standard deviation, range) or
+robust (interquartile range, difference between \(95^{\text{th}}\) percentile
+and \(5^{\text{th}}\) percentile) measure of scale summarizing across the voxel
+dimension, yielding a feature of \(T\) dimensions. Exact details can be found
+in the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
+source code}.
+
+\subsubsection{Baseline Features}
+MAYBE MOVE THIS SECTION AFTER RMT AND EIGEN FEATURES?
+
+fMRI global mean signal is
+
+\subsubsection{Non-RMT Eigenfeatures}
+
+Eigenfeatures summarize \(\mathbf{M}\) via some
+transformation of the eigenvalues of the correlation matrix of \(\mathbf{M}\),
+and thus yield a feature of  \(T - 1\) dimensions.
+
+\subsubsection{RMT Features}
 Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity, \(\Delta_3(L)\), is
@@
 
+\subsection{Multiverse Analysis}
+\label{sec:multiverse}
+
 \section{Results}
-\label{sect:results}
+\label{sec:results}
 

### 2022-11-06 paper/paper/SPIE/article.tex
[WIP]

@@
 
-\textbf{NOTE}: Forget academia for a moment, and write like its a blog. Then go
-back and add the citations and adjust the tone.
+Random Matrix Theory (RMT) has obvious potential.  In small-scale physical
+systems, RMT universalities have been observed in quantum chaotic systems,
+complex nuclei, atoms, molecules and disordered mesoscopic systems
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
+and at larger scales, RMT has been applied to atmospheric physics
+\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
+\cite{plerouRandomMatrixApproach2002} , social networks
+\cite{jalanUncoveringRandomnessSuccess2014}, random networks
+\cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
+liquids
+\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}
+and amorphous clusters
+\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
 
-Random Matrix Theory (RMT) has obvious potential. It has helped to understand
-the generalization behaviour of neural networks [add citation], can be used to
-guide any principal components analysis reduction [cite: Google search: "random
-matrix" "principal components"], has been used to help understand phenomena as
-diverse as [include the usual citations to buses, traffic, physics, etc], and
-can provide insights wherever there is a system of components interacting with
-some independence [cite Cambridge RMT handbook, others].
+Within biological systems, RMT has also been used to successfully model aspects
+of  amino acid functional relationships
+\cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in epileptic seizures
+\cite{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein
+interactions both in different species
+\cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
+\cite{raiRandomnessPreservedPatterns2015}.
 
@@
 matrices under certain assumptions or conditions. A random matrix is,
-generally, a matrix in which the entries can be considered to be independently
-and identically distributed (\textit{iid}). For example, one of the simplest
+generally, a matrix in which the entries are drawn from independently
+and identically distributed (\textit{iid}) distributions. For example, one of the simplest
 ensembles [cite handbook] is the Gaussian Unitary Ensemble (GUE), wherein [desc
@@
 
-For a system as complex as the human brain, RMT is thus tantalizing. It would
-be quite something if we could extract some matrix reduction of brain activity,
-and if the RMT description of this matrix either aligned nicely with RMT
-theory, or otherwise was broadly predictively useful.
+For a system as complex as the human brain, RMT is thus tantalizing. If some
+matrix representation of brain activity can be described, even partially, by
+aspects of RMT theory, this would mean  description of this matrix either aligned nicely with
+RMT theory, or otherwise was broadly predictively useful.
 
@@
 
+\subsubsection{RMT Features}
+\label{sec:rmt-features}
+
+To compare the observed eigenvalue distribution to some RMT theoretical
+predictions, the eigenvalues must be unfolded
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
+details and motivation behind the unfolding procedure are well documented in
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
+can be viewed as a smoothing and rescaling procedure where the originally
+observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
+an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
+if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
+function is not known, and so must be approximated
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
+a polynomial is chosen \cite{abul-magdUnfoldingSpectrumChaotic2014}.
+
+Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity,
+\(\Delta_3(L)\), is calculated for any positive real value \(L <
+\max(\Lambda)\) as:
+\[
+\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
+\]
+where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
+equal to \(\lambda\), \(\langle \cdot \rangle_c\) denotes the average with
+respect to all starting points \(c\), and where \(A\) and \(B\) denote the
+slope and intercept, respectively, of the least squares fit of a straight line
+to \(\eta(\lambda)\) on \([c, c+L]\)
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
+as a timeseries, the spectral rigidty for a value \(L\) can be is thus the
+\textit{average nonlinearity} of all intervals of length \(L\) over the series.
+
+The level number variance, \(\Sigma^2(L)\), or level variance, for short, is
+closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
+calculated as:
+\[
+\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
+\]
+where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
+and where \(c\), \(L\), and \(\langle \cdot \rangle_c\) are as above
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
+as an \textit{irregular} timeseries, the level number varaince is simply the
+variance of the number of samples of all intervals of length \(L\) over the
+series.
+
+Both the spectral rigidity and level number variance are also computed and
+investigated in this study. Because the computation of these statistics is
+non-trivial, we developed and release the separate, open-source Python library
+\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
+allows for efficient, parallel computation of the spectral rigidity and level
+variance via Monte Carlo methods, and automatically ensures convergence of the
+statistics to user-specified tolerances.
+
 \subsubsection{Baseline Features}
+\label{sec:baselines}
 
-If RMT- or eigenvalue-based features fail to outperform simple reductions of
-the fMRI data in predictive utility, it is difficult to justify the greater
-computational and interpretational complexities of the former. These simple
-statistical reductions help assess the relative value of RMT features.
+MAYBE MOVE THIS SECTION AFTER RMT AND EIGEN FEATURES?
 
-Baseline reductions compute either a robust measure of location (mean, max,
-min) or non-robust measure of location (median, \(95^{\text{th}}\) percentile,
-\(5^{\text{th}}\) percentile), or a non-robust (standard deviation, range) or
-robust (interquartile range, difference between \(95^{\text{th}}\) percentile
-and \(5^{\text{th}}\) percentile) measure of scale summarizing across the voxel
-dimension, yielding a feature of \(T\) dimensions. Exact details can be found
-in the
+If RMT- or eigenvalue-based features fail to outperform features that are simple
+summary statistics of the fMRI data in predictive utility, it is difficult to
+justify the greater computational and interpretational complexities of the
+former. We compute ten simple statistical reductions of the voxel time series
+("timeseries features") to help assess the relative value of RMT features.
+
+Baseline reductions use as statistical summaries: robust measures of location
+(mean, max, min); non-robust measures of location (median, \(95^{\text{th}}\)
+percentile, \(5^{\text{th}}\) percentile); non-robust measures of scale
+(standard deviation, range); and robust measures of scale (interquartile range,
+difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
+percentile). These statistics are used to summarizing / reduce the fMRI data
+along the voxel dimension, yielding a feature of \(T\) dimensions. E.g. the "mean"
+feature is just the global mean signal. Exact details can be found in the
 \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
@@
 
-\subsubsection{Baseline Features}
-MAYBE MOVE THIS SECTION AFTER RMT AND EIGEN FEATURES?
-
-fMRI global mean signal is
-
 \subsubsection{Non-RMT Eigenfeatures}
+\label{sec:eigs-only}
 
@@
 and thus yield a feature of  \(T - 1\) dimensions.
-
-\subsubsection{RMT Features}
-Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity, \(\Delta_3(L)\), is
-calculated for any positive real value \(L < \max(\Lambda)\) as:
-\[
-\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\]
-where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
-\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
-\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
-line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The spectral
-rigidty for a value \(L\) is thus the \textit{average nonlinearity}
-
-The level
-number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
-\cite{mehtaRandomMatrices2004}, and is calculated as:
-
-\[
-\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\]
-
-where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
-and \(\langle \cdot \rangle_c\) are as above \cite{guhrRandommatrixTheoriesQuantum1998a}. Both the
-spectral rigidity and level number variance are also computed and investigated in this study.
 

### 2022-11-07 paper/paper/SPIE/article.tex
wip

@@
 \max(\Lambda)\) as:
-\[
+\begin{equation}
+\label{eq:rigidity}
 \Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\]
+\end{equation}
 where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
@@
 calculated as:
-\[
+\begin{equation}
+\label{eq:levlvar}
 \Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\]
+\end{equation}
 where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
@@
 
+\textit{Trimming Variants}. As the unfolding procedure operates on the sorted
+eigenvalues, and involves fitting a smooth polynomial or exponential to these
+values, extreme values are often trimmed (or omitted from fitting)
+\cite{moralesImprovedUnfoldingDetrending2011,
+abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
+highly subjective, and not usually well-documented. In addition, with a large
+number of different source matrices, it is too destructive to use the same hard
+criterion (e.g. largest 10\%, largest 3) for all unfoldings.
+
+We develop and test four trimming variants: no trimming, precision-based,
+largest, and middle, with far less subjective criteria for determining
+trimming thresholds. However, we relegate the details of these trimming
+procedures to the appendix, or refer readers to the source code [link].
+
+\textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
+with a polynomial. The choice of degree has been known to dramatically impact
+certain analyses involving the spectral rigidity or level
+variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
+moralesImprovedUnfoldingDetrending2011, abueleninSpectralUnfoldingChaotic2018,
+fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}. We
+thus compute the rigidity and level variance with all possible combinations of
+trimming and unfolding polynomial degree.
+
+\subsubsection{Non-RMT Eigenfeatures}
+\label{sec:eigs-only}
+
+When computing the RMT eigenfeatures, the polynomial fitting during unfolding
+has a smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
+operations. However, it is possible other, simpler
+transformations of the eigenvalues (e.g. uniform or Savitsky-Golay smoothing)
+might have equal or greater predictive utility. We thus also test smoothed
+variants of the eigenvalues as predictive features: we smooth the sorted
+eigenvalues with a uniform (moving average) filter and Savitsky-Golay, using
+window sizes of 3, 5, 7, and 9 to yield 8 total "smoothed eigenvalue" features.
+
+\subsubsection{Slicing Variants}
+
+It is possible that only some eigenvalues
+contain predictively-useful information. Broadly, we might expect the largest,
+middle, or smallest eigenvalues to be most useful in characterizing some
+system. E.g. if the smallest eigenvalues correspond to random / noise aspects
+of a system, but it the differences in the nature of the noise that most
+differentiate between certain states, then here the smallest eigenvalues may
+have the most predictive utility.
+
+Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
+locality at which we summarize the spectrum. E.g. at small \(L\)-values, the
+spectral rigidity summarizes the non-linearity of eigenvalues thar are
+relatively close to each other. At large values of \(L\), the rigidity
+summarizes the long-range non-linearity of the spectrum. E.g. larger \(L\)
+values summarize the spectrum more globally, and smaller \(L\) values summarize
+it more locally.
+
+Since predictive utility may vary with summary locality, or with the region of
+the original spectrum, we investigate various contiguous front, middle, and end
+slices of each eigenfeature in all analyses, and where the size of each slice
+is either 5\%, 10\%, or 20\% of the full eigenfeatures for the front and end.
+silces, and 10\%, 20\%, and 40\% for the middle eigenvalues.
+
+\subsubsection{Feature Combinations}
+
+It could easily be the case that a combination of RMT and non-RMT eigenfeatures
+are more predictive than either alone. We thus test a variety of eigenfeature
+combinations, with the motivation for these particular combinations being to
+test whether RMT eigenfeatures add anything beyond smoothing.
+...
+
 \subsubsection{Baseline Features}
@@
 
-If RMT- or eigenvalue-based features fail to outperform features that are simple
-summary statistics of the fMRI data in predictive utility, it is difficult to
-justify the greater computational and interpretational complexities of the
-former. We compute ten simple statistical reductions of the voxel time series
-("timeseries features") to help assess the relative value of RMT features.
+If RMT- or eigenvalue-based features fail to predictively outperform features
+that are simple summary statistics of the fMRI data, it is difficult to justify
+the greater computational and interpretational complexities of the former. We
+compute ten simple statistical reductions of the voxel time series ("timeseries
+features") to help assess the relative value of RMT features.
 
@@
 \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
-source code}.
-
-\subsubsection{Non-RMT Eigenfeatures}
-\label{sec:eigs-only}
-
-Eigenfeatures summarize \(\mathbf{M}\) via some
-transformation of the eigenvalues of the correlation matrix of \(\mathbf{M}\),
-and thus yield a feature of  \(T - 1\) dimensions.
+source code}. Note we do not take slice variants of these baseline features.
+
+\subsubsection{Final Feature List}
+
+[INSERT TABLE OF ALL FEATURES]
 
@@
 \label{sec:multiverse}
+
+Our investigation can be structured hierarchically: we have a number of
+\textit{datasets}, and each dataset has one or more \textit{classification
+tasks}. Each classification task can be solved with a \textit{classifier}.
+Each classifier fit operates on a \textit{featyre}, and each feature is
+extracted via a certain \textit{preprocessing degree} and various
+\textit{analytic choices} (slicing, trimming, unfolding degree, normalization).
 

### 2022-11-07 paper/paper/SPIE/article.tex
[WIP]

@@
 \affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
-
-% \author[a]{First Author}
-% \author[a]{Second Author}
-% \author[b]{Third Author}
-% \author[a,b,*]{Fourth Author}
-% \affil[a]{University Name, Faculty Group, Department, Street Address, City, Country, Postal Code}
-% \affil[b]{Company Name, Street Address, City, Country, Postal Code}
 
@@
 
-At its most rudimentary, RMT simply describes the statistically-expected
-behaviour of the eigenvalues of some classes—or \textit{ensembles}—of random
-matrices under certain assumptions or conditions. A random matrix is,
-generally, a matrix in which the entries are drawn from independently
-and identically distributed (\textit{iid}) distributions. For example, one of the simplest
-ensembles [cite handbook] is the Gaussian Unitary Ensemble (GUE), wherein [desc
-of ensemble properties] and which appears to be useful in describing
-[real-world examples].
-
-This means that if we examine a system and find that the distribution of
-eigenvalues of that system resemble some RMT predicted spectrum, we potentially
-gain insight into the nature of that system. If, in addition, we find a large
-deviance from the spectrum predicted by RMT, we find contrapositive evidence
-that the system is \textit{not} like those described by RMT. Roughly, complex
-systems well-described by RMT are highly random and independent, whereas
-systems that are not so well-described are either highly dependent, non-random,
-or both.
-
-In this sense, RMT might have real \textit{explanatory} and
-\textit{statistical} value: the more we see that a system resembles an RMT
-system, the less we can distinguish it from one of random, non-interacting
-components, and the less the system resembles an RMT system, the more
-fundamentally complex and interactive the interactions.
-
-For a system as complex as the human brain, RMT is thus tantalizing. If some
-matrix representation of brain activity can be described, even partially, by
-aspects of RMT theory, this would mean  description of this matrix either aligned nicely with
-RMT theory, or otherwise was broadly predictively useful.
-
-[Transition to fMRI now...]
-
-In functional magnetic resonance imaging (fMRI), changes in the
-blood-oxygenation-level-dependent (BOLD) signals are related to neural
-activity. It is common to investigate statistical relationships between the
-BOLD signals through \textbf{functional connectivity} analyses, where
-correlations between collections of these signals are examined to infer
-connections between different voxels or regions of interest (ROIs) within the
-brain.
-
-Whether in the presence of experimental stimuli, or the relative absence, as in
-resting state fMRI (rs-fMRI), complex patterns in functional connectivity
-networks are ubiquitous
-\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
-This complexity suggests fMRI is a candidate to be studied using \textbf{Random
-Matrix Theory} (RMT), a set of mathematical tools originally developed some 50
-years ago to solve complex problems in nuclear physics
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}
-
-In various physical systems where nuclei can have different possible states,
-precisely modeling the entirety of the interactions between nuclei is often
-intractable. For such large, complex systems, the eigenvalues (or spectra)
-efficiently summarize the totality of the interactions between components of
-the system, and RMT describes the expected behaviour of such eigenvalues. By
-analysing the spectrum statistically, complex systems can be investigated by
-comparison to the universal properties predicted by RMT.
+At its most rudimentary, RMT describes the expected behaviour of the
+eigenvalues—also often called levels—of a number of classes of random
+matrices\cite{andersonIntroductionRandomMatrices2010,
+akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
+simplifying somewhat, a matrix in which the entries are drawn from
+independently and identically distributed (\textit{iid}) distributions.
+
+RMT likely has the most real-world explanatory potential when dealing with a
+complex system of many (hundreds or more\cite{mehtaRandomMatrices2004})
+interacting components. If such a system has a matrix representation, and the
+eigenvalues of this representation have a distribution similar to one predicted
+by RMT, it suggests the system is either highly random or chaotic. By contrast,
+if the observed spectra deviate significantly from RMT-predicted spectra, this
+suggests otherwise. A number of studies have used RMT to make such
+interpretations and comparisons between
+systems\cite{santhanamStatisticsAtmosphericCorrelations2001,
+jalanUncoveringRandomnessSuccess2014,
+bandyopadhyayUniversalityComplexNetworks2007,
+agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003, wangSpectralPropertiesTemporal2015,
+wangRandomMatrixTheory2016, matharooSpontaneousBackpainAlters2020}.
 
@@
 
-Recently, RMT has been applied to study brain functioning. The earliest study
+RMT may thus have real potential in understanding the human brain. For an
+imaging modality like functional magnetic resonance imaging (fMRI), where
+changes in the blood-oxygenation-level-dependent (BOLD) signals are related to
+neural activity, RMT may be an ideal starting point. Each voxel or collection
+of voxels—i.e. Region of Interest (ROI\footnote{We will use ROI to refer to
+\textit{any} collection of voxels of any size. E.g. a single voxel is still an
+ROI.}); any collection )—can be considered an interacting component of the system.
+
+Likewise, in \textbf{functional connectivity} analyses, statistical
+relationships of the BOLD ROI time-courses are investigated in the hope of
+gaining insights into brain functioniong (see
+\citenum{bastosTutorialReviewFunctional2016,
+vandenheuvelExploringBrainNetwork2010} for reviews, but also
+\citenum{nobleDecadeTestretestReliability2019} and
+\citenum{lurieQuestionsControversiesStudy2020} for challenges facing functional
+connectivity analyses). In this framework, each connection or correlation can
+be considered a system component, and the eigenvalues of such a correlation
+matrix can be examined from the perspective of RMT.
+
+The earliest study taking this approach
 demonstrated that spectra of the correlations between electroencephalographic
 (EEG) signals closely resemble those of the Gaussian Orthogonal Ensemble (GOE)
-\cite{sebaRandomMatrixAnalysis2003}. Somewhat more recently, RMT was used to
+\cite{sebaRandomMatrixAnalysis2003}. In fMRI, RMT has been used to
 evaluate the quality of whole brain features extracted from fMRI data
-\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has
-also been used in diffusion MRI to aid in the selection of the number of
+\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}, and in
+diffusion MRI to aid in the selection of the number of
 components to employ in principal-component reduction analysis and denoising
@@
 
+\subsubsection{Raw Eigenvalues}
+\label{sec:raw-eigs}
+
+The time and space complexities of computing the full \(N \times N\) correlation
+matrix of \(\mathbf{M}\), and the subsequent eigenvalues, are too large to be
+tractable. However, since transposition does not change the eigenvalues, we
+can use the transpose to efficiently compute these eigenvalues (see
+Appendix \ref{sec:transpose}).
+
 \subsubsection{RMT Features}
@@
 \cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as an \textit{irregular} timeseries, the level number varaince is simply the
-variance of the number of samples of all intervals of length \(L\) over the
+as an \textit{irregular} timeseries, the level number variance is the
+variation of the number of samples of all intervals of length \(L\) over the
 series.
@@
 
-\subsection{Tables}
-Tables are numbered in the order in which they are referenced.
+% \subsection{Tables}
+% Tables are numbered in the order in which they are referenced.
 They should appear in the document in numerical order and as close as possible to their first reference in the text.
@@
 
-\subsection{Multimedia}
-Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player.
+% \subsection{Multimedia}
+% Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player.
 The recommended maximum size for each multimedia file is 10-12 MB.
@@
 
-\begin{video}
-\begin{center}
-{\includegraphics[height=5cm]{satellite.eps}}
-\\
-\end{center}
-\caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
-\end{video}
+% \begin{video}
+% \begin{center}
+% {\includegraphics[height=5cm]{satellite.eps}}
+% \\
+% \end{center}
+% \caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
+% \end{video}
 
@@
 
-\section{Miscellaneous Formatting Details}
-\label{sect:misc}
-At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.
-
-\subsection{Formatting Equations}
-Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.
+\section{Correlation Eigenvalues via Transposition}
+\label{sec:transpose}
+
+Let \(\mathbf{X}\) is a real \(n \times p\) matrix with \(n > p\). Let
+\[
+\mathbf{Z} = \text{norm}(\mathbf{X}) =
+\left(
+\mathbf{X}_1 - \bar{\mathbf{X}}_1 | \dots | \mathbf{X}_p - \bar{\mathbf{X}}_p
+\right)
+\]
+where \(\mathbf{X}_i\) denotes column \(i\) of \(\mathbf{X}\). Denote the (unordered) set
+of eigenvalues of \(\mathbf{X}\) as \(\text{eigs}(\mathbf{X})\), and let \(r = (p - 1)^{-1}\).
+Denote the covariance matrix of \(\mathbf{X}\) as \(\text{cov}(\mathbf{X})\).  Then:
+
+\begin{align*}
+\text{eigs}\left( \text{cov}(\mathbf{X})  \right)
+&= \text{eigs}(r \cdot \mathbf{Z}\mathbf{Z}^{\top}) \\
+&= r \cdot \text{eigs}(\mathbf{Z}\mathbf{Z}^{\top}) \\
+&= r \cdot \text{eigs}((\mathbf{Z}\mathbf{Z}^{\top})^{\top}) \\
+&= r \cdot \text{eigs}(\mathbf{Z}^{\top}\mathbf{Z}) \\
+\end{align*}
+
+Denote the correlation matrix of \(\mathbf{X}\) as \(\text{corr}(\mathbf{X})\), and let
+\[
+\mathbf{Y} = \text{stand}(\mathbf{X}) =
+\left(
+(\mathbf{X}_1 - \bar{\mathbf{X}}_1) / \sigma_1 | \dots | (\mathbf{X}_p - \bar{\mathbf{X}}_p) / \sigma_p
+\right)
+\]
+
+Then
+
+\[
+\text{eigs}\left( \text{corr}(\mathbf{X})  \right)
+= \text{eigs}\left( \text{cov}(\mathbf{Y})  \right)
+= r \cdot \text{eigs}(\mathbf{Y}^{\top}\mathbf{Y})
+\]
+
+% \section{Miscellaneous Formatting Details}
+% \label{sect:misc}
+% At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.
+
+% \subsection{Formatting Equations}
+% Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.
 Important equations appear on their own line.
@@
 For example, ``The expression for the field of view is
-\begin{equation}
-\label{eq:fov}
-2 a = \frac{(b + 1)}{3c} \, ,
-\end{equation}
-where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.
-
-Equations are considered to be part of a sentence and should be punctuated accordingly.
+% \begin{equation}
+% \label{eq:fov}
+% 2 a = \frac{(b + 1)}{3c} \, ,
+% \end{equation}
+% where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.
+
+% Equations are considered to be part of a sentence and should be punctuated accordingly.
 In the above example, a comma appears after the equation because the next line is a subordinate clause.
@@
 
-References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise.
+% References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise.
 Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..
 
-\subsection{Formatting Theorems}
-
-To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.
+% \subsection{Formatting Theorems}
+
+% To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.
 The text of the theorem continues on the same line in normal, 10-pt.
@@
 
-\vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}
+% \vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}
 

### 2022-11-07 paper/paper/SPIE/article.tex
[WIP]

@@
 \usepackage{graphicx}
+\graphicspath{ {./figures/} }
 \usepackage{setspace}
@@
 dependent on analytic decisions, suggesting considerable caution when
-interprting past and future studies applying RMT.
+interpreting past and future studies applying RMT to fMRI.
 
@@
 
-Random Matrix Theory (RMT) has obvious potential.  In small-scale physical
-systems, RMT universalities have been observed in quantum chaotic systems,
-complex nuclei, atoms, molecules and disordered mesoscopic systems
+At its most rudimentary, RMT describes the expected behaviour of the
+eigenvalues—also often called levels—of a number of classes of random
+matrices\cite{andersonIntroductionRandomMatrices2010,
+akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
+simplifying somewhat, a matrix in which the entries are drawn from
+independently and identically distributed (\textit{iid}) distributions. Though
+first developed to describe the fluctuation of nuclei energy levels in quantum
+physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
+RMT has been shown have extremely broad potential.
+
+In small-scale physical systems, RMT universalities have been observed in
+quantum chaotic systems, complex nuclei, atoms, molecules and disordered
+mesoscopic systems
 \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
@@
 \cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
-
 Within biological systems, RMT has also been used to successfully model aspects
@@
 \cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
-\cite{raiRandomnessPreservedPatterns2015}.
-
-At its most rudimentary, RMT describes the expected behaviour of the
-eigenvalues—also often called levels—of a number of classes of random
-matrices\cite{andersonIntroductionRandomMatrices2010,
-akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
-simplifying somewhat, a matrix in which the entries are drawn from
-independently and identically distributed (\textit{iid}) distributions.
+\cite{raiRandomnessPreservedPatterns2015}. RMT has also been used to guide
+statistical decisions in principal components
+analyses\cite{franklinParallelAnalysisMethod1995,
+veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
+even more recently, has provided insights into the behaviours of deep
+neural networks\cite{martinImplicitSelfRegularizationDeep2021,
+martinPredictingTrendsQuality2021}.
 
@@
 
-RMT may thus have real potential in understanding the human brain. For an
-imaging modality like functional magnetic resonance imaging (fMRI), where
-changes in the blood-oxygenation-level-dependent (BOLD) signals are related to
-neural activity, RMT may be an ideal starting point. Each voxel or collection
-of voxels—i.e. Region of Interest (ROI\footnote{We will use ROI to refer to
-\textit{any} collection of voxels of any size. E.g. a single voxel is still an
-ROI.}); any collection )—can be considered an interacting component of the system.
-
-Likewise, in \textbf{functional connectivity} analyses, statistical
-relationships of the BOLD ROI time-courses are investigated in the hope of
-gaining insights into brain functioniong (see
+In the human brain, each neuron, collection of neurons, or region of interest
+(ROI) is a potentially-interacting component in a highly complex system. RMT
+may have potential in describing the totality of these interactions, provided
+that measurements of functioning can be obtained with sufficient spatial and
+temporal resolution to speak to some neurobiological or neuropsychological
+phenomenon of interest.
+
+For an imaging modality like fMRI, where changes in the
+blood-oxygenation-level-dependent (BOLD) signals are related to neural
+activity, RMT may be an ideal starting point. Each voxel time course (or
+collection of such time sources - i.e. ROIs) can be considered an interacting
+component of the system. Likewise, in \textit{functional connectivity}
+analyses, statistical relationships of the BOLD ROI time-courses are
+investigated in the hope of gaining insights into brain functioniong (see
 \citenum{bastosTutorialReviewFunctional2016,
@@
 
-Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to
+RMT has also been used in ROI-based fMRI functional-connectivity studies to
 investigate differences between rest and task states
@@
 \cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the
-spectra of resting or low-attention states exhibited properties closest to the
+spectra of resting or low-attention states exhibited properties close to the
 GOE. These findings suggest that certain aspects of psychological processes
@@
 
-\subsection{Motivating Eigenfeatures}
-
-The basic insight of RMT is thus that the \textit{eigenvalues alone} may
-provide interesting information about highly complex Neurobiological systems.
-However, in practice, the eigenvalues computed from empirically observed data
-are noisy. RMT is statistical in nature, describing only the \textit{expected}
-behaviour of the system spectra. To this end, a number of summary statistics or
-``spectral observables'' \cite{mehtaRandomMatrices2004} can be computed from
-empirically observed spectra, with these spectral observables sometimes being
-better suited for further analysis or comparison to theory. Two such summary
-statistics that have been somewhat popular are the \textit{spectral rigidity}
-and \textit{level number variance} (``rigidity'' and ``level variance'' for
-short; details in Section \ref{sect:methods}).
+\subsection{Eigenvalues and Prediction}
+
+The basic insight of RMT is thus that the eigenvalues alone may provide
+interesting information about highly complex systems. However, real systems are
+usually noisy, and involve a mixture of random and non-random components and
+interactions. RMT is statistical in nature, describing only the expected
+behaviour of the spectra of \textit{iid} random matrics. To this end, a number
+of summary statistics or ``spectral observables''
+\cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a} can be
+computed from empirically observed spectra, with these spectral observables
+sometimes being better suited for further analysis, or for the comparison of
+empirical observations to theory. Two such summary statistics that have been
+popular\cite{santhanamStatisticsAtmosphericCorrelations2001,
+jalanUncoveringRandomnessSuccess2014, matharooSpontaneousBackpainAlters2020,
+bandyopadhyayUniversalityComplexNetworks2007,
+agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
+sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,
+wangRandomMatrixTheory2016} are the \textit{spectral rigidity} and
+\textit{level number variance} (``rigidity'' and ``level variance'' for short;
+details in Section \ref{sect:methods}).
 
@@
 predictively-useful information. If the basic RMT insight is that the
-eigenvalues are useful features to understand a system, then those eigenvalues
-alone (or other simple transformations of them) ought also to be predictively
+eigenvalues alone can provide understanding of a system, then those eigenvalues
+(or other simple transformations of them) ought also to be predictively
 useful. We examine a number of such \textit{eigenfeatures} in this study.
@@
 Both the spectral rigidity and level number variance are also computed and
-investigated in this study. Because the computation of these statistics is
-non-trivial, we developed and release the separate, open-source Python library
+investigated in this study, for all \(L \in \{ 1, 2, \dots 20 \}\).
+Because the computation of these statistics is non-trivial, we developed and
+release the separate, open-source Python library
 \code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
@@
 trimming thresholds. However, we relegate the details of these trimming
-procedures to the appendix, or refer readers to the source code [link].
+procedures to the Appendix \ref{sec:trimming}, or refer readers to the \href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source code}.
 
@@
 
-It is possible that only some eigenvalues
-contain predictively-useful information. Broadly, we might expect the largest,
-middle, or smallest eigenvalues to be most useful in characterizing some
-system. E.g. if the smallest eigenvalues correspond to random / noise aspects
-of a system, but it the differences in the nature of the noise that most
-differentiate between certain states, then here the smallest eigenvalues may
-have the most predictive utility.
+We might expect the largest, middle, or smallest eigenvalues to be most useful
+in characterizing some system. E.g. if the smallest eigenvalues correspond to
+random / noise aspects of a system, but it is differences in the nature of the
+noise that most differentiate between certain systems, then here the smallest
+eigenvalues may have the most predictive utility.
 
 Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
-locality at which we summarize the spectrum. E.g. at small \(L\)-values, the
-spectral rigidity summarizes the non-linearity of eigenvalues thar are
-relatively close to each other. At large values of \(L\), the rigidity
-summarizes the long-range non-linearity of the spectrum. E.g. larger \(L\)
-values summarize the spectrum more globally, and smaller \(L\) values summarize
-it more locally.
+locality at which we summarize the spectrum: at small \(L\)-values, the
+spectral rigidity summarizes the non-linearity of eigenvalues that are
+relatively close to each other in magnitude. At large values of \(L\), the rigidity
+summarizes the long-range non-linearity of the spectrum.
 
 Since predictive utility may vary with summary locality, or with the region of
-the original spectrum, we investigate various contiguous front, middle, and end
-slices of each eigenfeature in all analyses, and where the size of each slice
-is either 5\%, 10\%, or 20\% of the full eigenfeatures for the front and end.
-silces, and 10\%, 20\%, and 40\% for the middle eigenvalues.
+the original spectrum, we investigate various front, middle, and end contiguous
+slices of each eigenfeature in all analyses, where the size of each slice is
+either the first or last 5\%, 10\%, or 20\% of the full eigenfeatures for
+non-middle slices, or the middle 10\%, 20\%, or 40\% of the full eigenfeature for
+the middle eigenvalues.
 
@@
 \label{sec:results}
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{gross_feature_overall_by_subgroup.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:main-results}
+AUROC distributions across gross feature groupings and comparison tasks.}
+\end{figure}
 
@@
 
+\section{Trimming Procedures}
+
+We implement three trimming procedures: precision-based, largest, and middle
+trimming. The
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
+code} is the definitive reference for the procedures, but we describe the
+motivations belows.
+
+In precision-based trimming, we trim away any eigenvalues that are close enough
+to zero to be considered a result of numerical error due to floating point
+representation. There are two thresholds we consider, including the one used by
+NumPy\cite{harrisArrayProgrammingNumPy2020} for the
+\href{https://numpy.org/doc/stable/reference/generated/numpy.linalg.matrix_rank.html}{determination
+of matrix rank}, and those recommended by LAPACK\cite{laug} in their user
+guide\cite{andersonLAPACKUsersGuide1999a} on the
+\href{https://netlib.org/lapack/lug/node89.html}{error bounds for symmetric
+eigenproblems}, and related
+\href{https://netlib.org/lapack/lug/node90.html}{additional details}. We trim
+each matrix eigenvalues to whichever threshold is largest for the matrix in
+question.
+
+For largest trimming, we must determine a threshold at which to separate
+"large" from "small" eigenvalues. The eigenvalues for our data tended to grow
+exponentially, so we instead looked at thresholding on the logarithms. We then
+used \(k\)-means with \(k=2\) on the precision-trimmed eigenvalues, and took
+the largest cluster (which also always had the smaller mean) cluster as the
+"largest" eigenvalues to trim away. "Middle" trimming simply reflects the
+threshold found by the largest trim method, e.g. if the largest trim method
+removes the last \(n\) precision-trimmed eigenvalues, then we also trim the
+first \(n\) smallest eigenvalues remaining after precision-trimming.
+
+We chose one-dimensional k-means partly due to efficiency and simplicity, and
+because of the general relation to classical threholding methods like the Otsu
+method\cite{liuOtsuMethodKmeans2009}.
+
 % \section{Miscellaneous Formatting Details}

### 2022-11-07 code/scripts/summarize_everything_updated.py

@@ -13,6 +13,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import seaborn as sbn
+from shutil import copyfile
 from joblib import Memory
 from matplotlib.axes import Axes
 from matplotlib.figure import Figure
@@ -32,8 +33,10 @@ from rmt.visualize import best_rect
 PROJECT = ROOT.parent
 MEMORY = Memory(PROJECT / "__JOBLIB_CACHE__")
 SPIE_OUTDIR = PROJECT / "results/plots/figures/SPIE"
-if not SPIE_OUTDIR.exists():
-    SPIE_OUTDIR.mkdir(exist_ok=True, parents=True)
+SPIE_PAPER_OUTDIR = PROJECT / "paper/paper/SPIE/figures"
+
+SPIE_OUTDIR.mkdir(exist_ok=True, parents=True)
+SPIE_PAPER_OUTDIR.mkdir(exist_ok=True, parents=True)
 
 SPIE_JMI_MAX_WIDTH_INCHES = 6.75
 SPIE_JMI_MAX_COL_WIDTH_INCHES = 3 + 5 / 16
@@ -1413,8 +1416,11 @@ def make_row_labels(grid: FacetGrid, col_order: list[str], row_order: list[str])
 def savefig(fig: Figure, filename: str) -> None:
     print("Saving...", end="", flush=True)
     outfile = SPIE_OUTDIR / filename
+    paper_outfile = SPIE_PAPER_OUTDIR / filename
     fig.savefig(outfile, dpi=600)
     print(f" saved figure to {outfile}")
+    copyfile(outfile, paper_outfile)
+    print(f"Copied saved figure to {paper_outfile}")
     plt.close()
 
 

### 2022-11-08 paper/paper/SPIE/article.tex
[WIP]

@@
 We develop and test four trimming variants: no trimming, precision-based,
-largest, and middle, with far less subjective criteria for determining
-trimming thresholds. However, we relegate the details of these trimming
-procedures to the Appendix \ref{sec:trimming}, or refer readers to the \href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source code}.
+largest, and middle, with far less subjective criteria for determining trimming
+thresholds. However, we relegate the details of these trimming procedures to
+the Appendix \ref{sec:trimming}, or refer readers to the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
+code}.
 
@@
 \section{Trimming Procedures}
+\label{sec:trimming}
 

### 2022-11-08 paper/paper/SPIE/article.tex
[WIP]

@@
 thus compute the rigidity and level variance with all possible combinations of
-trimming and unfolding polynomial degree.
+trimming and unfolding polynomial degrees of 3, 5, 7, and 9.
 
@@
 
-When computing the RMT eigenfeatures, the polynomial fitting during unfolding
+When computing RMT eigenfeatures, the polynomial fitting during unfolding
 has a smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
@@
 eigenvalues with a uniform (moving average) filter and Savitsky-Golay, using
-window sizes of 3, 5, 7, and 9 to yield 8 total "smoothed eigenvalue" features.
+window sizes of 3, 5, 7, and 9 to yield 8 total ``smoothed eigenvalue'' features.
 

### 2022-11-08 paper/paper/SPIE/article.tex
[WIP]

@@
 \usepackage{doi}
+\usepackage{csquotes}
+\MakeOuterQuote{"}
 
 \newcommand{\code}[1]{\small \texttt{#1} \normalsize}
+\newcommand{\tcode}[1]{\footnotesize \texttt{#1} \normalsize}
 
@@
 \section{Methods}
-\label{sect:methods}
+\label{sec:methods}
 
 \subsection{fMRI Preprocessing}
+\label{sec:preproc}
 
@@
 
-The time and space complexities of computing the full \(N \times N\) correlation
-matrix of \(\mathbf{M}\), and the subsequent eigenvalues, are too large to be
-tractable. However, since transposition does not change the eigenvalues, we
-can use the transpose to efficiently compute these eigenvalues (see
-Appendix \ref{sec:transpose}).
+The time and space complexities of computing the full \(N \times N\)
+correlation matrix of \(\mathbf{M}\), and the subsequent eigenvalues, are too
+large to be tractable. However, since transposition does not change the
+eigenvalues, we can use the transpose to efficiently compute these eigenvalues
+(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
+in Table \ref{tab:features}), most directly test the idea that eigenvalues
+alone are useful functional connectivity features.
+
+However, not all eigenvalues may have predictive utility. We also examine as
+features the central eigenvalues (defined as the middle 10\%, 20\%, or 40\% of
+the sorted eigenvalues, \code{eigsmiddle} in Table \ref{tab:features}), and
+tail eigenvalues (defined as the first and last 5\%,  10\%, and 20\% of each
+tail of the spectrum, \code{eigsminmax} in Table \ref{tab:features}).
 
@@
 
-\subsubsection{Non-RMT Eigenfeatures}
+\subsubsection{Smoothed Eigenfeatures}
 \label{sec:eigs-only}
@@
 
+\subsubsection{Feature Combinations}
+
+It could easily be the case that a combination of RMT and non-RMT eigenfeatures
+are more predictive than either alone. We thus test a variety of eigenfeature
+combinations, with the motivation for these particular combinations being to
+test whether RMT eigenfeatures add anything beyond smoothing. Combined features
+are formed by concatenation, so that if we have features \(f_1, \dots,  f_n\)
+with dimensions \(p_1, \dots, p_n\), then the combined feature is \([f_1;
+\dots; f_n]\) with \(\sum p_i\) dimensions.  The final combinations chosen can
+be found in Table \ref{tab:features}.
+
 \subsubsection{Slicing Variants}
@@
 
-\subsubsection{Feature Combinations}
-
-It could easily be the case that a combination of RMT and non-RMT eigenfeatures
-are more predictive than either alone. We thus test a variety of eigenfeature
-combinations, with the motivation for these particular combinations being to
-test whether RMT eigenfeatures add anything beyond smoothing.
-...
+For combined features, slicing variants are also computed, with slicing
+performed first on each feature to be combined. E.g. if we have features \(f_1,
+\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
+feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
+elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
+\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
 
@@
 \label{sec:baselines}
-
-MAYBE MOVE THIS SECTION AFTER RMT AND EIGEN FEATURES?
 
@@
 percentile). These statistics are used to summarizing / reduce the fMRI data
-along the voxel dimension, yielding a feature of \(T\) dimensions. E.g. the "mean"
+along the voxel dimension, yielding a feature of \(T\) dimensions. E.g. the ``mean''
 feature is just the global mean signal. Exact details can be found in the
 \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
-source code}. Note we do not take slice variants of these baseline features.
-
-\subsubsection{Final Feature List}
-
-[INSERT TABLE OF ALL FEATURES]
+source code}.
+
+We do not take slice variants of these baseline features, since the baseline
+features are still time series. However, we do also evaluate a number of
+smoothing degrees of these features, to account for noise and to be somewhat
+similar to the various smoothing variants for the eigenfeatures. Each baseline
+feature thus has a uniformly smoothed variant, where the size of the smoothing
+window is either 1 (no smoothing), 2, 4, 8, or 16.
+
+\subsection{Classifiers}
+
+We use a variety of stadard machine-learning classifiers available in the
+Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
+solve each classification task. We use a gradient-boosted decision tree (GBDT),
+random forest (RF) classifier, support vector classifier (SVC) with radial
+basis function, and k-nearest neighbours classifiers with k equal to 3, 5, and
+9 (KNN3, KNN5, and KNN9, respectively). We originally also attempted to test a
+simple logistic regression classifier, but found that this model to frequently
+failed to converge for a number of features, even after increasing iterations
+significantly. In all cases the default hyperparameter values were used.
+
+\subsection{Preprocessing Levels}
+
+Each preprocessing intermediate in the (1) brain extraction, (2) slice time correction,
+(3) motion correction, and (4) template registration pipeline was saved and used for
+entirely separate feature analyses. E.g. all features were extracted from a brain
+prepared with one of four levels or degrees of preprocessing, where level \(k\)
+includes preprocessings steps 1 to \(k\), inclusive, and starting at \(k = 1\).
+
+\begin{table}[h!]
+\caption{
+    Feature groupings for summarization. eigs = non-RMT eigenfeatures and their combinations with
+    other non-RMT eigenfeatures. rmt = RMT eigenfeatures and their combinations. tseries = baseline
+    timeseries reductions.\\
+}
+\label{tab:features}
+\small
+\centering
+\begin{tabular}{lllr}
+\hline
+\textbf{Gross Source} & \textbf{Feature Group} & \textbf{Feature id}   \\
+\hline
+eigs           & eigs          & \tcode{eigs}                           \\
+               & eigs max      & \tcode{eigsminmax10}                   \\
+               &               & \tcode{eigsminmax20}                   \\
+               &               & \tcode{eigsminmax5}                    \\
+               & eigs middle   & \tcode{eigsmiddle10}                   \\
+               &               & \tcode{eigsmiddle20}                   \\
+               &               & \tcode{eigsmiddle40}                   \\
+               & eigs smooth   & \tcode{eigs + eigs\_smooth}            \\
+               &               & \tcode{eigs + savgol}                  \\
+               &               & \tcode{eigs\_savgol}                   \\
+               &               & \tcode{eigs\_smooth}                   \\
+               &               &                                        \\
+rmt            & rmt + eigs    & \tcode{eigs + levelvar}                \\
+               &               & \tcode{eigs + rigidity}                \\
+               &               & \tcode{eigs + rigidity + levelvar}     \\
+               &               & \tcode{eigs + unfolded}                \\
+               &               & \tcode{eigs + unfolded + levelvar}     \\
+               &               & \tcode{eigs + unfolded + rigidity}     \\
+               & rmt only      & \tcode{levelvar}                       \\
+               &               & \tcode{rigidity}                       \\
+               &               & \tcode{rigidity + levelvar}            \\
+               &               & \tcode{unfolded}                       \\
+               &               & \tcode{unfolded + levelvar}            \\
+               &               & \tcode{unfolded + rigidity}            \\
+               &               & \tcode{unfolded + rigidity + levelvar} \\
+               &               &                                        \\
+tseries        & location      & \tcode{T-max}                          \\
+               &               & \tcode{T-mean}                         \\
+               &               & \tcode{T-med}                          \\
+               &               & \tcode{T-min}                          \\
+               &               & \tcode{T-p05}                          \\
+               &               & \tcode{T-p95}                          \\
+               & scale         & \tcode{T-iqr}                          \\
+               &               & \tcode{T-rng}                          \\
+               &               & \tcode{T-rrng}                         \\
+               &               & \tcode{T-std}                          \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-Our investigation can be structured hierarchically: we have a number of
-\textit{datasets}, and each dataset has one or more \textit{classification
-tasks}. Each classification task can be solved with a \textit{classifier}.
-Each classifier fit operates on a \textit{featyre}, and each feature is
-extracted via a certain \textit{preprocessing degree} and various
-\textit{analytic choices} (slicing, trimming, unfolding degree, normalization).
+We perform a \textit{multiverse
+analysis}\cite{steegenIncreasingTransparencyMultiverse2016} to assess the
+overall predictive potential of the various eigenfeatures across all
+previously-mentioned analytic choices. That is, for each combination of
+comparison task and classifier, we use 5-fold cross validation, and use the
+mean AUROC (mAUROC) across folds as our metric to evaluate feature predictive utility.
+
+Each non-baseline feature is evaluated, for each classification task and
+classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\)
+2 normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
+degrees  \(\times\) 10 slice choices. Each baseline feature is evaluated
+exactly 40 times for each comparison task and classifier: 4 preprocessing
+levels \(\times\) 2 normalization methods \(\times\) 5 smoothing window sizes.
+In total, with all the datasets examined in this study, this yields 2 053 920
+mAUROC values to summarize, with 5 main \textit{analytic factors}:
+preprocessing, normalization, trimming, degree, and slicing.
+
+\section{Results}
+
+Summarizing this amount of data requires some caution. Measures of location or
+scale, even when robust, are, for the most part, misleading and uninformative when distributions
+are skewed, and summarizing too broadly is likewise uninformative. E.g. the
+average mAUROC across all over \(2 \times 10^6\) model evaluations is too
+broad a summary to meaningfully interpret. Likewise, each analytic factor
+could produce an interaction: e.g. perhaps
+a meaningful number.
+
+Our multiverse analysis can be structured semi-hierarchically, where choices at each level
+in the hierarchy impact the final distribution of AUROCs observed. We have a
+number of \textit{datasets}, and each dataset has one or more
+\textit{classification tasks}. Each classification task can be solved with a
+\textit{classifier}. Each classifier fit operates on a \textit{feature}, and
+each feature is extracted via a certain \textit{preprocessing degree} and
+various \textit{analytic choices} (slicing, trimming, unfolding degree,
+normalization).
 

### 2022-11-08 code/scripts/summarize_everything_updated.py

@@ -2122,6 +2122,15 @@ def summary_stats_and_tables() -> None:
     )
 
 
+def make_feature_table() -> None:
+    df = load_combined()
+    feats = (
+        df.loc[:, ["gross_feature", "feature_group", "feature"]].value_counts().to_frame()
+    )
+    feats.sort_values(by=["gross_feature", "feature_group", "feature"], inplace=True)
+    print(feats.to_latex())
+
+
 if __name__ == "__main__":
     simplefilter(action="ignore", category=PerformanceWarning)
     pd.options.display.max_rows = 1000

### 2022-11-08 paper/paper/SPIE/article.tex
[WIP]

@@
 
-Each preprocessing intermediate in the (1) brain extraction, (2) slice time correction,
-(3) motion correction, and (4) template registration pipeline was saved and used for
-entirely separate feature analyses. E.g. all features were extracted from a brain
-prepared with one of four levels or degrees of preprocessing, where level \(k\)
-includes preprocessings steps 1 to \(k\), inclusive, and starting at \(k = 1\).
+Each preprocessing intermediate in the (1) brain extraction, (2) slice time
+correction, (3) motion correction, and (4) template registration pipeline was
+saved and used for entirely separate feature analyses. E.g. all features were
+extracted from a brain prepared with one of four levels or degrees of
+preprocessing, where level \(k\) includes preprocessings steps 1 to \(k\),
+inclusive, and starting at \(k = 1\).
+
+\subsection{Normalization}
+
+Because of the exponential distribution of the eigenfeatures, normalization is
+somewhat non-trivial, and a simple standardization or min-max normalization is
+ineffective. We instead first apply a logarithm to all eigenfeatures, and then
+test each classifier with un-normalized or min-max normalized versions of the
+log-features. Baseline features are also tested with un-normalized and min-max
+normalized versions, but without any log transform.
 

### 2022-11-09 paper/paper/SPIE/article.tex
[WIP]

@@
 Each non-baseline feature is evaluated, for each classification task and
-classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\)
-2 normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
+classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\) 2
+normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
 degrees  \(\times\) 10 slice choices. Each baseline feature is evaluated
@@
 In total, with all the datasets examined in this study, this yields 2 053 920
-mAUROC values to summarize, with 5 main \textit{analytic factors}:
-preprocessing, normalization, trimming, degree, and slicing.
+mAUROC values to summarize, with 5 main \textit{analytic factors}—
+preprocessing, normalization, trimming, unfolding/smoothing degree, and
+slicing—to consider.
 
 \section{Results}
+\label{sec:results}
 
 Summarizing this amount of data requires some caution. Measures of location or
-scale, even when robust, are, for the most part, misleading and uninformative when distributions
-are skewed, and summarizing too broadly is likewise uninformative. E.g. the
-average mAUROC across all over \(2 \times 10^6\) model evaluations is too
-broad a summary to meaningfully interpret. Likewise, each analytic factor
-could produce an interaction: e.g. perhaps
-a meaningful number.
-
-Our multiverse analysis can be structured semi-hierarchically, where choices at each level
-in the hierarchy impact the final distribution of AUROCs observed. We have a
-number of \textit{datasets}, and each dataset has one or more
-\textit{classification tasks}. Each classification task can be solved with a
-\textit{classifier}. Each classifier fit operates on a \textit{feature}, and
-each feature is extracted via a certain \textit{preprocessing degree} and
-various \textit{analytic choices} (slicing, trimming, unfolding degree,
-normalization).
-
-\section{Results}
+scale, even when robust, are, for the most part, misleading and uninformative
+when distributions are skewed. In most cases we find skewed mAUROC distributions,
+and so have chosen primarily to present our findings visually.
+
+In addition, each analytic factor could produce an interaction. For example, it
+could be that trimming impacts the overall mAUROC distribution only at a
+certain preprocessing level and for certain feature slicing. However, there are
+too many potential interactions to present, so we limit our presentation
+to the most significant ones.
+
+As there are a large number of figures, we summarize patterns of results across
+two abstract levels of feature groupings—"gross", and "feature group"—shown in
+Table \ref{tab:features}.
+
+Finally, it makes sense to exclude classification tasks or an analytic
+factor from summaries if there is no evidence that the classification task is
+solvable in general, or for that factor. We take as lack of evidence of solvability
+an mAUROC distribution that is either (1) roughly symmetric and has mean and median
+very close to 0.5, or (2) left-tailed, with median less than 0.5.
+
+\subsection{}
+
+initial investigation reveals no evidence
+that a certain compone
+
+  if a certain feature is simply not useful, or if a certain comparison
+task is not solvable with some or even all classifiers
+
 \label{sec:results}

### 2022-11-09 paper/paper/SPIE/article.tex
make supplementary material

@@
 \MakeOuterQuote{"}
+\usepackage{xr}
+\externaldocument[S-]{supplementary}
 
@@
 \hline
-\textbf{Gross Source} & \textbf{Feature Group} & \textbf{Feature id}   \\
+\textbf{Coarse Grouping} & \textbf{Fine Grouping} & \textbf{Feature id}   \\
 \hline
@@
 scale, even when robust, are, for the most part, misleading and uninformative
-when distributions are skewed. In most cases we find skewed mAUROC distributions,
-and so have chosen primarily to present our findings visually.
+when distributions are skewed. In most cases we find skewed mAUROC
+distributions, and so have chosen primarily to present our findings visually.
 
@@
 As there are a large number of figures, we summarize patterns of results across
-two abstract levels of feature groupings—"gross", and "feature group"—shown in
+two abstract levels of feature groupings—"coarse", and "fine"—shown in
 Table \ref{tab:features}.
 
-Finally, it makes sense to exclude classification tasks or an analytic
-factor from summaries if there is no evidence that the classification task is
-solvable in general, or for that factor. We take as lack of evidence of solvability
-an mAUROC distribution that is either (1) roughly symmetric and has mean and median
-very close to 0.5, or (2) left-tailed, with median less than 0.5.
-
-\subsection{}
-
-initial investigation reveals no evidence
-that a certain compone
-
-  if a certain feature is simply not useful, or if a certain comparison
-task is not solvable with some or even all classifiers
-
-\label{sec:results}
+Finally, it makes sense to exclude classification tasks or an analytic factor
+from summaries if there is no evidence that the classification task is solvable
+in general, or for that factor. We take as lack of evidence of solvability an
+mAUROC distribution that is either (1) roughly symmetric and has mean and
+median close to 0.5, or (2) with median and mode less than 0.5.
 
@@
 \begin{tabular}{c}
-\includegraphics[width=6.5in]{gross_feature_overall_by_subgroup.png}
+\includegraphics[width=6.5in]{coarse_feature_overall_by_subgroup.png}
 \end{tabular}
@@
 \end{figure}
+
+\subsection{Classification Tasks}
+
+\subsubsection{Non-Predictable Comparisons}
+Based on the mAUROC distributions for either coarsely-grouped (Figure
+\ref{fig:main-results}) or finely-grouped features (Figure
+S\ref{S-fig:feature-group-all}), neither bilinguality nor depression were
+predictable by either baseline or eigenfeatures. Additionally, the \code{pain v
+duloxetine} comparison in the \textit{Osteo} dataset, and the
+\code{trait\_attend v trait\_nonattend} conditions ("WeeklyAttend" in Figures
+\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the \textit{Attention}
+dataset were also not meaningfully predictable by any feature. As such, we exclude
+these classification tasks from further figures and discussion.
+
+\subsubsection{Predictable Comparisons}
+As visible in Figure \ref{fig:main-results}, when coarsely summarizing features,
+eigenfeatures were more likely to be predictively useful than not, and except
+for in the \code{task\_attend v task\_nonattend} comparison, were also more
+predictively useful than the baseline features. We also note that although
+orders of magnitude more models were fit for eigenfeatures (see Section \ref{sec:multiverse}),
+the mAUROC distributions for the baseline features are either equally as wide,
+or, for the Parkinson's data, wider.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{coarse_feature_by_preproc_predictive_subgroup.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:preproc} mAUROC distributions by preprocessing degree.}
+\end{figure}
+
+\subsection{Effect of Preprocessing}
+At a coarse level
+
+  if a certain feature is simply not useful, or if a certain comparison
+task is not solvable with some or even all classifiers
 

### 2022-11-09 paper/paper/SPIE/supplementary.tex
make supplementary material

--- 2022-11-08 paper/paper/SPIE/article.tex
@@
-\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
-%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
+% \documentclass[12pt]{supplementary}  % 12pt font required by SPIE;
+\documentclass{article}  % use this instead for A4 paper
 \usepackage{amsmath,amsfonts,amssymb}
+\usepackage{authblk}
+\usepackage{times}
+\usepackage{float}
+\usepackage{ifthen}
+\usepackage{setspace}
+\usepackage[super]{cite}[2003/11/04] % need vers. > 4.01
+\usepackage{color}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{graphicx}
@@
 \MakeOuterQuote{"}
+\usepackage{pdflscape}
+\usepackage{geometry}
+\geometry{
+  textwidth=11in,
+  textheight=8.5in,
+  headheight=0in,
+  headsep=0in,
+  marginparwidth=0in,
+  papersize={11in,8.5in},
+  left=0in,
+  lmargin=0in,
+  inner=0in,
+  right=0in,
+  rmargin=0in,
+  outer=0in,
+  top=0in,
+  tmarg=0in,
+  bottom=0in,
+  bmargin=0in,
+}
 
@@
 
-\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
+\pagestyle{empty}               % page numbers is default; use empty for no numbers
+
+\title{Supplementary Material}
 
@@
 \cftpagenumbersoff{table}
+
 \begin{document}
+
+\begin{center}
 \maketitle
+\end{center}
 
-% \begin{abstract} This document shows the required format and appearance of a
-% manuscript prepared for SPIE journals. It is prepared using LaTeX2e with the
-% class file \texttt{spieman.cls}. Please note that the following journals
-% require the use of structured abstracts in manuscript submissions:
-% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the
-% \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for
-% the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Guidelines
-% are available on the journal website. Whether structured or single-paragraph,
-% the abstract should be a summary of the paper and not an introduction.
-% Because the abstract may be used in abstracting and indexing databases, it
-% should be self-contained (i.e., no numerical references) and substantive in
-% nature, presenting concisely the objectives, methodology used, results
-% obtained, and their significance. A list of up to six keywords should
-% immediately follow. \end{abstract}
-
-% See
-% https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
-% for info on the structured abstract
-%
-% Purpose: This section presents the significance and aims, stating the broad
-% impact and the rationale or motivation of the work, including potentially
-% some background.
-%
-% Approach: This section briefly describes the materials and methods, including
-% the statistical analyses, used in the research.
-%
-% Results: This section provides a core summary of the findings from the work,
-% including study numbers, quantitative analyses, or discoveries.
-%
-% Conclusions: This section summarizes and interprets the approach and findings
-% from the work, stating also the importance or impact of the findings.
-
-% Purpose: Data-intensive modeling could provide insight on the broad
-% variability in outcomes in spine surgery. Previous studies were limited to
-% analysis of demographic and clinical characteristics. We report an analytic
-% framework called “SpineCloud” that incorporates quantitative features
-% extracted from perioperative images to predict spine surgery outcome.
-
-% Approach: A retrospective study was conducted in which patient demographics,
-% imaging, and outcome data were collected. Image features were automatically
-% computed from perioperative CT. Postoperative 3- and 12-month functional and
-% pain outcomes were analyzed in terms of improvement relative to the
-% preoperative state. A boosted decision tree classifier was trained to predict
-% outcome using demographic and image features as predictor variables.
-% Predictions were computed based on SpineCloud and conventional demographic
-% models, and features associated with poor outcome were identified from
-% weighting terms evident in the boosted tree.
-
-% Results Neither approach was predictive of 3- or 12-month outcomes based on
-% preoperative data alone in the current, preliminary study. However,
-% SpineCloud predictions incorporating image features obtained during and
-% immediately following surgery (i.e., intraoperative and immediate
-% postoperative images) exhibited significant improvement in area under the
-% receiver operating characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3
-% months and AUC=0.69 (CI95=0.55 to 0.82) at 12 months.
-
-% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved
-% by incorporation of image-based features compared to analysis based on
-% conventional demographic data. The SpineCloud framework could improve
-% understanding of factors underlying outcome variability and warrants further
-% investigation and validation in a larger patient cohort.
-\begin{abstract}
-
-\section*{Purpose}
-Random matrix theory (RMT) is an increasingly useful tool for understanding
-large, complex systems. Prior studies have examined functional Magnetic
-Resonance Imaging (fMRI) scans using tools from RMT, with some success.
-However, RMT computations are highly sensitive to a number of analytic choices,
-and the robustness of findings involving RMT remain in question. We
-systematically investigate the usefulness of RMT on a wide variety of fMRI
-datasets using a rigorous predictive framework.
-
-\section*{Approach}
-We develop open-source software to efficiently compute RMT features from fMRI
-images, and examine the cross-validated predictive potential of eigenvalue and
-RMT-based features (``eigenfeatures'') with classic machine-learning
-classifiers. We systematically vary pre-processing extent, normalization
-procedures, RMT unfolding procedures, and feature selection, and compare the
-impact of these analytic choices on the distributions of cross-validated
-prediction performance for each combination of dataset binary classification
-task, classifier, and feature. To deal with class imbalance, we use the
-area under the receiver operating characteristic (AUROC) as the main
-performance metric.
-
-\section*{Results}
-% df.groupby(["subgroup", "gross_feature"])["auroc"].median()
-Across all classification tasks and analytic choices, we find eiegenfeatures to
-more often have predictive utility than not (82.4\% of median AUROCs \(>\) 0.5;
-median AUROC range across classification tasks 0.47 - 0.64). Simple baseline
-reductions on source timeseries, by contrast, were less useful (58.8\% of
-median AUROCs \(>\) 0.5, median AUROC range across classification tasks 0.42 -
-0.62). Additionally, eigenfeature AUROC distributions were overall much more
-right-tailed than baseline features suggesting greater predictive potential.
-However, performance distributions were wide, and often significantly affected by
-analytic choices.
-
-\section*{Conclusions}
-Eigenfeatures clearly have potential for understanding and reducing fMRI
-functional connectivity. However, the utility of these features is strongly
-dependent on analytic decisions, suggesting considerable caution when
-interpreting past and future studies applying RMT to fMRI.
-
-% This document shows the required format and appearance of a manuscript
-% prepared for SPIE journals. It is prepared using LaTeX2e with the class file
-% \texttt{spieman.cls}. Please note that the following journals require the use
-% of structured abstracts in manuscript submissions: \textit{Neurophotonics},
-% the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical
-% Imaging}. Structured abstracts are encouraged for the \textit{Journal of
-% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the
-% journal website. Whether structured or single-paragraph, the abstract should
-% be a summary of the paper and not an introduction. Because the abstract may
-% be used in abstracting and indexing databases, it should be self-contained
-% (i.e., no numerical references) and substantive in nature, presenting
-% concisely the objectives, methodology used, results obtained, and their
-% significance. A list of up to six keywords should immediately follow.
-\end{abstract}
-
-% Include a list of up to six keywords after the abstract
-\keywords{random matrix, spectral rigidity, level number variance, fMRI, classification, machine-learning}
-
-% Include email contact information for corresponding author
-{\noindent \footnotesize\textbf{*}Jacob Levman,  \linkable{jlevman@stfx.ca} }
-
-\begin{spacing}{2}   % use double spacing for rest of manuscript
-
-\section{Introduction}
-\label{sect:intro}  % \label{} allows reference to this section
-% This document shows the format and appearance of a manuscript prepared for
-% submission to an SPIE journal. Note that this template is only intended to be
-% used as a guideline for author convenience. It is designed for optimum
-% clarity and ease of reading for editors and reviewers, but the template does
-% not reflect the final page layout of a published journal paper. Accepted
-% papers are professionally typeset in XML according to the layout and design
-% of the journal.
-
-At its most rudimentary, RMT describes the expected behaviour of the
-eigenvalues—also often called levels—of a number of classes of random
-matrices\cite{andersonIntroductionRandomMatrices2010,
-akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
-simplifying somewhat, a matrix in which the entries are drawn from
-independently and identically distributed (\textit{iid}) distributions. Though
-first developed to describe the fluctuation of nuclei energy levels in quantum
-physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
-RMT has been shown have extremely broad potential.
-
-In small-scale physical systems, RMT universalities have been observed in
-quantum chaotic systems, complex nuclei, atoms, molecules and disordered
-mesoscopic systems
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
-and at larger scales, RMT has been applied to atmospheric physics
-\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
-\cite{plerouRandomMatrixApproach2002} , social networks
-\cite{jalanUncoveringRandomnessSuccess2014}, random networks
-\cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
-liquids
-\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}
-and amorphous clusters
-\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
-Within biological systems, RMT has also been used to successfully model aspects
-of  amino acid functional relationships
-\cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in epileptic seizures
-\cite{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein
-interactions both in different species
-\cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
-\cite{raiRandomnessPreservedPatterns2015}. RMT has also been used to guide
-statistical decisions in principal components
-analyses\cite{franklinParallelAnalysisMethod1995,
-veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
-even more recently, has provided insights into the behaviours of deep
-neural networks\cite{martinImplicitSelfRegularizationDeep2021,
-martinPredictingTrendsQuality2021}.
-
-RMT likely has the most real-world explanatory potential when dealing with a
-complex system of many (hundreds or more\cite{mehtaRandomMatrices2004})
-interacting components. If such a system has a matrix representation, and the
-eigenvalues of this representation have a distribution similar to one predicted
-by RMT, it suggests the system is either highly random or chaotic. By contrast,
-if the observed spectra deviate significantly from RMT-predicted spectra, this
-suggests otherwise. A number of studies have used RMT to make such
-interpretations and comparisons between
-systems\cite{santhanamStatisticsAtmosphericCorrelations2001,
-jalanUncoveringRandomnessSuccess2014,
-bandyopadhyayUniversalityComplexNetworks2007,
-agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003, wangSpectralPropertiesTemporal2015,
-wangRandomMatrixTheory2016, matharooSpontaneousBackpainAlters2020}.
-
-\subsection{RMT and Neurobiological Signals}
-
-In the human brain, each neuron, collection of neurons, or region of interest
-(ROI) is a potentially-interacting component in a highly complex system. RMT
-may have potential in describing the totality of these interactions, provided
-that measurements of functioning can be obtained with sufficient spatial and
-temporal resolution to speak to some neurobiological or neuropsychological
-phenomenon of interest.
-
-For an imaging modality like fMRI, where changes in the
-blood-oxygenation-level-dependent (BOLD) signals are related to neural
-activity, RMT may be an ideal starting point. Each voxel time course (or
-collection of such time sources - i.e. ROIs) can be considered an interacting
-component of the system. Likewise, in \textit{functional connectivity}
-analyses, statistical relationships of the BOLD ROI time-courses are
-investigated in the hope of gaining insights into brain functioniong (see
-\citenum{bastosTutorialReviewFunctional2016,
-vandenheuvelExploringBrainNetwork2010} for reviews, but also
-\citenum{nobleDecadeTestretestReliability2019} and
-\citenum{lurieQuestionsControversiesStudy2020} for challenges facing functional
-connectivity analyses). In this framework, each connection or correlation can
-be considered a system component, and the eigenvalues of such a correlation
-matrix can be examined from the perspective of RMT.
-
-The earliest study taking this approach
-demonstrated that spectra of the correlations between electroencephalographic
-(EEG) signals closely resemble those of the Gaussian Orthogonal Ensemble (GOE)
-\cite{sebaRandomMatrixAnalysis2003}. In fMRI, RMT has been used to
-evaluate the quality of whole brain features extracted from fMRI data
-\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}, and in
-diffusion MRI to aid in the selection of the number of
-components to employ in principal-component reduction analysis and denoising
-\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
-
-RMT has also been used in ROI-based fMRI functional-connectivity studies to
-investigate differences between rest and task states
-\cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
-ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain states
-\cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the
-spectra of resting or low-attention states exhibited properties close to the
-GOE. These findings suggest that certain aspects of psychological processes
-might be characterized, in part, by features computed from the eigenvalues of
-fMRI correlation matrices, and that these features might vary in an
-interpretable manner across psychological processes. If this is the case, RMT
-could aid in understanding the functioning of the human brain.
-
-\subsection{Eigenvalues and Prediction}
-
-The basic insight of RMT is thus that the eigenvalues alone may provide
-interesting information about highly complex systems. However, real systems are
-usually noisy, and involve a mixture of random and non-random components and
-interactions. RMT is statistical in nature, describing only the expected
-behaviour of the spectra of \textit{iid} random matrics. To this end, a number
-of summary statistics or ``spectral observables''
-\cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a} can be
-computed from empirically observed spectra, with these spectral observables
-sometimes being better suited for further analysis, or for the comparison of
-empirical observations to theory. Two such summary statistics that have been
-popular\cite{santhanamStatisticsAtmosphericCorrelations2001,
-jalanUncoveringRandomnessSuccess2014, matharooSpontaneousBackpainAlters2020,
-bandyopadhyayUniversalityComplexNetworks2007,
-agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
-sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,
-wangRandomMatrixTheory2016} are the \textit{spectral rigidity} and
-\textit{level number variance} (``rigidity'' and ``level variance'' for short;
-details in Section \ref{sect:methods}).
-
-However, from a predictive standpoint, these summary statistics may eliminate
-predictively-useful information. If the basic RMT insight is that the
-eigenvalues alone can provide understanding of a system, then those eigenvalues
-(or other simple transformations of them) ought also to be predictively
-useful. We examine a number of such \textit{eigenfeatures} in this study.
-
-In addition, the functional connectivity - the matrix of correlations of
-various collections of voxels of an fMRI scan - is a priori a useful
-representation of the fMRI data. However, the full correlation matrix between
-all voxels is itself often computationally infeasible to work with, and thus is
-reduced in various ways prior to being used in analyses.
-
-Typical reductions include the use of a ``seed'' voxel or collection of voxels
-(i.e. ROI): the mean signal over that ROI is correlated with all other N ROIs
-of interest to generate a reduced functional connectivity matrix (or image)
-with one correlation value at each non-seed ROI. This sort of reduction reduces
-the functional connectivity to N values. This approach strongly biases the
-analysis to the seed ROI.
-
-Likewise, one can work with ROI mean signals, and work with the \(N \times N\)
-matrix of these ROI correlations. This analysis [maybe cite Poldrack book for
-this?] is, \textit{a priori}, sensitive to the choice of ROIs, and, in the case
-of anatomical atlases / parcellations, may also lack theoretical justification.
-
-A major disadvantage of both the above approaches, besides requiring
-parcellation into ROIs, is that most such parcellation methods are atlas-based,
-and thus require image registration, and, general, an entire preprocessing
-pipeline to ensure that registration is successful.
-
-RMT suggests a potentially useful reduction in the form of the eigenvalues of
-the voxelwise functional connectivity. An \(N \times t\) matrix \(\bf{A}\) of
-\(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
-correlation (or covariance) matrix with \(t - 1\) non-zero, positive
-real-valued eigenvalues \(\Lambda\). These eigenvalues can be computed highly
-efficiently via transposition of the voxelwise correlation or covariance
-matrix.
-
-\subsection{Limitations of Previous Work}
-
-However, these previous studies took largely descriptive / explanatory
-approaches, noting only differences in RMT across subgroups and/or conditions,
-in many cases, with no formal statistical model nor formal testing of the
-significance and robustness of these differences. No code was provided to allow
-the reproduction of the results, and the extraction of the spectral rigidity
-and level variance is non-trivial, and computationally demanding. Computation
-of these spectral observables requires an \textit{unfolding}
-procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004},
-which is, unfortunately, a poorly-documented fitting procedure that requires
-making a number of difficult-to-justify decisions, and these decisions can
-dramatically impact RMT conclusions
-\cite{abul-magdUnfoldingSpectrumChaotic2014,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012,moralesImprovedUnfoldingDetrending2011}.
-On top of the flexibility in the implementation and application of RMT to
-empirical data, there is also the flexibility introduced by the choices in the
-complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
-
-In addition these issues stemming from ``research degrees of freedom'' (cite
-Gelman), there is also a basic baseline comparison issue: RMT is mathematically
-non-trivial, the extraction of RMT spectral statistics is computationally
-demanding, and eigenvalues, while easy to understand mathematically, are
-difficult to interpret relative to simpler potential reductions of fMRI data.
-
-What is missing is a rigorous, systematic, reproducible investigation of the
-*predictice* value of RMT metrics across a wide variety of data, RMT analytic
-choices, choice of eigenfeature, and choice of preprocessing, and for these RMT
-eigenfeatures to be compared to simpler alternative baselines. The current
-study aims to do just this.
-
-\section{Datasets}
-\label{sec:datasets}
-
-\subsection{Overview}
-
-We selected fMRI data publicly available on the
-\href{https://openneuro.org/}{OpenNeuro} platform (36). Selection criteria was
-somewhat subjective, but we required (1), that all scans have the same spatial
-and temporal resolutions, (2), the the dataset comprise 10+ human subjects, and
-(3), that it was reasonable to classify at the level of an entire scan, i.e.
-neither task nor event timings were needed to determine class membership. This
-yielded seven datasets comprising XX total subjects, and over 5000 total fMRI
-volumes.
-
-\begin{table}[h!]
-\caption{
-    ID = Identifier for paper. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P horizontal / axial slices each with dimensions M x N.
-}
-\label{table:1}
-\small
-\centering
-\begin{tabular}{ l c c c c c c }
-\hline
-\textbf{Name}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
-\hline
-Aging     & 74 × 74 × 32   & 3.0 × 3.0 × 4.0 & 2.0  & 300 & 62  \\
-Bilingual    & 100 × 100 × 72 & 1.8 × 1.8 × 1.8 & 0.88 & 823 & 90  \\
-Depress   & 112 × 112 × 25 & 2.0 × 2.0 × 5.0 & 2.5  & 100 & 72  \\
-Learn     & 64 × 64 × 36   & 3.0 × 3.0 × 3.0 & 2.0  & 195 & 432 \\
-Osteo     & 64 × 64 × 36   & 3.4 × 3.4 × 3.0 & 2.5  & 300 & 74  \\
-Park      & 80 × 80 × 43   & 3.0 × 3.0 × 3.0 & 2.4  & 149 & 552 \\
-Attention & 128 × 128 × 70 & 1.5 × 1.5 × 1.5 & 3.0  & 300 & 90  \\
-\hline
-\end{tabular}
-\end{table}
-
-In order not to inundate readers with dataset details, and because this is an
-exploratory investigation of RMT which is not committed to any specific
-theory\footnote{In fact, datasets and, later, the classification tasks, were
-chosen without investigation into any findings or publications associated with
-the data, both in order to reduce bias in our decisions, and because the
-original authors analytic / statistical approaches and intentions have very
-limited relevance to our whole-brain, voxelwise, multiverse predictive
-approach.}, we refer the reader to the original publications and/or data
-releases when greater detail is needed, and highlight only the most basic
-aspects of each dataset here.  Dataset scan parameters and acquisition details
-are summarized in Table \ref{table:1}, and sample and subgroup sizes are
-summarized in Table \ref{table:2}.
-
-\begin{table}[h!]
-\caption{
-    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
-    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
-}
-\label{table:2}
-\small
-\centering
-\begin{tabular}{ l r c c c }
-\hline
-\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
-\hline
-Aging   &  older             & 28      & resting-state        & 1 \\
-        &  younger           & 34      & resting-state        & 1 \\
-        &                    &         &                      &   \\
-Bilingual  &  bilingual          & 59      & resting-state    & 1 \\
-        &  monolingual        & 33      & resting-state       & 1 \\
-        &                    &         &                      &   \\
-Depress &  depression         & 51      & resting-state       & 1 \\
-        &  control            & 21      & resting-state       & 1 \\
-        &                    &         &                      &   \\
-Learn   &  task              & 24      & learn image sketches & 16\\
-        &  rest              & 24      & resting-state        & 2 \\
-        &                    &         &                      &   \\
-Osteo   &  duloxetine        & 19      & resting-state        & 1 \\
-        &  pain              & 37      & resting-state        & 1 \\
-        &  nopain            & 20      & resting-state        & 1 \\
-        &                    &         &                      &   \\
-Park    &  Parkinson's       & 25      & ANT                  & 12\\
-        &  control           & 21      & ANT                  & 12\\
-        &                    &         &                      &   \\
-Attention&  vigilant           & 11     & resting-state       & 2 \\
-        &  nonigilant         & 11     & resting-state        & 2 \\
-        &  trait-attentive     & 11    & resting-state        & 2 \\
-        &  trait-non-attentive & 11    & resting-state        & 2 \\
-        &  task-attentive     & 11     & resting-state        & 2 \\
-        &  task-nonattentive  & 11     & resting-state        & 2 \\
-\hline
-\end{tabular}
-\end{table}
-
-\subsection{Datasets and Classification Tasks}
-
-The \textit{Aging} dataset\cite{wahlheimIntrinsicFunctionalConnectivity2021}
-included rs-fMRI data for 34 subjects with a mean age of 22 years (range: 18-32
-years), and 28 subjects with a mean age of 70 years (range: 61-80 years). For
-our analysis, we make the binary classification task the prediction of subject
-age-group membership, i.e. \code{younger v older}.
-
-% The final sample included 34 younger adults (1832 years old, Mage = 22.21, SD
-% = 3.65; 20 female) and 28 older adults (61-80 years old, Mage = 69.82, SD =
-% 5.64; 20 female). Table 1 shows that younger and older adults had comparable
-% education and working memory capacity, the latter measured by forward and
-% backward digit span (48). Older adults had higher vocabulary scores (49) and
-% slower processing speed (50) than younger adults. Resting-state data
-% comprised 300 measurements collected over one 10-minute run. This scan
-% duration produces reliable functional connectivity estimates within and
-% across participants (57). Participants were instructed to remain still and
-% awake with their eyes open. No stimuli appeared during this scan, and the
-% display was black. Functional scans were collected using an echo-planar image
-% sequence sensitive to blood-oxygen-level-dependent (BOLD) contrast (T2*; 32
-% slices with 4.0 mm thickness and no skip, TE = 30 ms, TR = 2000 ms, flip
-% angle = 70, FOV = 220 mm, matrix size = 74 × 74 × 32 voxels, A/P phase
-% encoding direction). Slices were collected in a descending order that covered
-% the entire cortex and partial cerebellum. At the beginning of the
-% resting-state scan, the scanner acquired and discarded two dummy scans.
-
-% \subsubsection{Bilinguality Dataset}
-
-The \textit{Bilinguality} dataset\cite{ds001747:1.0.0,
-goldcarrieelizabethExploringRestingState2018} examined English and
-Spanish-speaking monolinguals and multilinguals during a prolonged resting
-state. The original study grouped participants into three subgroups: early vs
-late bilinguals vs monolingual
-controls\cite{goldcarrieelizabethExploringRestingState2018}. For our analysis,
-we make the binary classification task be to classify \code{monolingual v
-bilingual}.
-
-% Resting state fMRI scans were performed using a T2*-weighted EPI sequence with the following
-% parameters: TR=875ms; TE=43.6ms; slice thickness=1.8mm; acquisition matrix=100 × 100; flip
-% angle=55°; number of slices=72; field of view=180 × 180mm; voxel size=1.8 × 1.8 × 1.8mm;
-% multi-band factor=8; volumes per run=823 (total scan time=12min). EPI scans were oriented
-% parallel to the long axis of the hippocampus.
-
-% \subsubsection{Depression Dataset}
-
-The \textit{Depression} dataset\cite{bezmaternykhddRestingStateClosed2021,
-bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans frmo
-non-depressed controls and mildy- or moderately-depressed subjects. The our analysis,
-the binary classification task is \code{depress v control}.
-
-% The intergroup study involved 21 healthy volunteers received a compensation
-% for a participation and 51 participants featuring mild depression (F32.0),
-% moderate depression (F32.1), or dysthymia (F34.1, single patient) (see Table
-% 1). The fMRI study was carried out in the International Tomography Center,
-% Novosibirsk, using a 3 T Ingenia scanner (Philips). Functional T2∗-weighted
-% Ssh echo planar imaging scans were acquired using the following parameters:
-% voxel size 2×2×5mm, repetition time/echo time = 2500/35 ms, and fat
-% suppression mode. The reference anatomical image was obtained by the T1W 3D
-% turbo field echo method with a voxel size of 1×1×1mm. The instruction for
-% participants was to lie still with eyes closed for 6 minutes.
-
-% \subsubsection{Learning Dataset}
-
-The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
-schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans were
-available for all subjects, and \code{task v rest} is chosen as the binary
-classification task for our analysis.  The task is complex and difficult to
-summarize here adequately, but involved multiple phases where subjects were
-presented with various related abstract images, and then later tested for their
-memory of certain aspects of the learned images. Thus the classification task
-might also be considered learning vs. rest.
-
-% \subsubsection{Osteoarthritis Dataset}
-The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
-includes whiole-brain rs-fMRI scans of healthy, pain-free controls
-(\code{nopain} condition), and individuals with knee osteoarthritis.
-Osteoarthritic patients were treated for 2 weeks with either placebo
-(\code{pain} condition) or duloxetine (\code{duloxetine} condition).
-We use the three binary classification tasks \code{nopain v duloxetine},
-\code{nopain v pain}, and \code{pain v duloxetine} for our analysis.
-
-% \subsubsection{Parkinson's Dataset}
-The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
-subjects with non-demented Parkinson's Disease (PD) and
-healthy controls performed a number of repetitions of the Attention Network
-Task (ANT; \cite{fanActivationAttentionalNetworks2005}) during scans.
-The classification task for this dataset is Parkinson's versus controls, i.e.
-\code{park v ctrl}.
-
-% \subsubsection{Attention Dataset}
-
-The \textit{Attention} dataset\cite{gorgolewskiHighResolution7Tesla2015} is a
-high-resolution rs-fMRI dataset including a battery of psychological measures
-for each subject. Since previous studies employing RMT sometimes interpreted
-their findings with respect to attentional processes
-\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we
-divided subjects into various high vs. low attention binary classification
-tasks based on median splits of subsets of the metrics available in this study.
-
-The \code{viglant v nonvigilant} task was formed based on self-report
-questionnaire items involving
-"vigilance"\cite{gorgolewskiCorrespondenceIndividualDifferences2014}. The
-\code{trait\_attend v trait\_nonattend} task was formed using PANAS-X
-\cite{watsonPANASXManualPositive1994} items related to self-reported
-wakefulness and attention over the past weeks. Finally, the
-\code{task\_attend v tas\_nonattent} classification task was constructed
-using the scores of The Conjunctive Continuous Performance Task
-(CCPT\cite{shalevConjunctiveContinuousPerformance2011}). This is a behavioural
-performance task which requires the subject to quickly and selectively respond
-to only certain visual stimuli. Details allowing exact reproduction of these
-splits are available in this paper's
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/preprocess/unify\_attention\_data.py}{source
-code}.
-
-\section{Methods}
-\label{sec:methods}
-
-\subsection{fMRI Preprocessing}
-\label{sec:preproc}
-
-We limited preprocessing to a simple pipeline of, in order, (1) brain
-extraction, (2) slice timing correction, (3), motion correction, and (4),
-registration to the MNI152
-\cite{fonovUnbiasedAverageAgeappropriate2011,fonovUnbiasedNonlinearAverage2009}
-2mm isotropic, asymmetric template version C made available through
-\href{https://www.templateflow.org/}{TemplateFlow}
-\cite{ciricTemplateFlowFAIRsharingMultiscale2021}.
-
-We limited preprocessing methods to these steps sometimes because data required
-for other typical preprocessing steps (e.g. field intensity maps, physiological
-measurements), but also because we saved all pipelined intermediates to later
-compare the effect of increasing degrees of pre-processing (see Section
-\ref{sec:multiverse}), and it was important to limit the number of
-preprocessing degrees to prevent excessive computational costs.
-
-Brain extraction was performed first with BET
-\cite{smithFastRobustAutomated2002}, and then with with
-ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
-by BET. BET results were visually inspected for each image in each dataset, and
-BET fractional anisotropy values were modified to ensure brain extraction was
-acceptable. Motion-correction was performed with FSL's MCFLIRT
-\cite{jenkinsonImprovedOptimizationRobust2002}, and slicetime correction also
-with FSL's \code{slicetimer}\cite{jenkinsonFSL2012} tool. Finally, functional
-images were registered directly to the MNI152 template via
-ANTs\cite{avantsReproducibleEvaluationANTs2011}.
-
-\subsection{Feature Extraction}
-
-For each preprocessing degree and fMRI image, features were extracted from all
-\(N\) non-constant brain voxels, where \(N\) may differ for each image. After
-this process, each scan yields an \(N \times T\) matrix \(\mathbf{M}\), where
-\(T\) is the number of volumes acquired. All features summarize this matrix
-of brain voxels \(\mathbf{M}\) in some manner.
-
-\subsubsection{Raw Eigenvalues}
-\label{sec:raw-eigs}
-
-The time and space complexities of computing the full \(N \times N\)
-correlation matrix of \(\mathbf{M}\), and the subsequent eigenvalues, are too
-large to be tractable. However, since transposition does not change the
-eigenvalues, we can use the transpose to efficiently compute these eigenvalues
-(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
-in Table \ref{tab:features}), most directly test the idea that eigenvalues
-alone are useful functional connectivity features.
-
-However, not all eigenvalues may have predictive utility. We also examine as
-features the central eigenvalues (defined as the middle 10\%, 20\%, or 40\% of
-the sorted eigenvalues, \code{eigsmiddle} in Table \ref{tab:features}), and
-tail eigenvalues (defined as the first and last 5\%,  10\%, and 20\% of each
-tail of the spectrum, \code{eigsminmax} in Table \ref{tab:features}).
-
-\subsubsection{RMT Features}
-\label{sec:rmt-features}
-
-To compare the observed eigenvalue distribution to some RMT theoretical
-predictions, the eigenvalues must be unfolded
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
-details and motivation behind the unfolding procedure are well documented in
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
-can be viewed as a smoothing and rescaling procedure where the originally
-observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
-an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
-if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
-function is not known, and so must be approximated
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
-a polynomial is chosen \cite{abul-magdUnfoldingSpectrumChaotic2014}.
-
-Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity,
-\(\Delta_3(L)\), is calculated for any positive real value \(L <
-\max(\Lambda)\) as:
-\begin{equation}
-\label{eq:rigidity}
-\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\end{equation}
-where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
-equal to \(\lambda\), \(\langle \cdot \rangle_c\) denotes the average with
-respect to all starting points \(c\), and where \(A\) and \(B\) denote the
-slope and intercept, respectively, of the least squares fit of a straight line
-to \(\eta(\lambda)\) on \([c, c+L]\)
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as a timeseries, the spectral rigidty for a value \(L\) can be is thus the
-\textit{average nonlinearity} of all intervals of length \(L\) over the series.
-
-The level number variance, \(\Sigma^2(L)\), or level variance, for short, is
-closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
-calculated as:
-\begin{equation}
-\label{eq:levlvar}
-\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\end{equation}
-where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
-and where \(c\), \(L\), and \(\langle \cdot \rangle_c\) are as above
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as an \textit{irregular} timeseries, the level number variance is the
-variation of the number of samples of all intervals of length \(L\) over the
-series.
-
-Both the spectral rigidity and level number variance are also computed and
-investigated in this study, for all \(L \in \{ 1, 2, \dots 20 \}\).
-Because the computation of these statistics is non-trivial, we developed and
-release the separate, open-source Python library
-\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
-allows for efficient, parallel computation of the spectral rigidity and level
-variance via Monte Carlo methods, and automatically ensures convergence of the
-statistics to user-specified tolerances.
-
-\textit{Trimming Variants}. As the unfolding procedure operates on the sorted
-eigenvalues, and involves fitting a smooth polynomial or exponential to these
-values, extreme values are often trimmed (or omitted from fitting)
-\cite{moralesImprovedUnfoldingDetrending2011,
-abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
-highly subjective, and not usually well-documented. In addition, with a large
-number of different source matrices, it is too destructive to use the same hard
-criterion (e.g. largest 10\%, largest 3) for all unfoldings.
-
-We develop and test four trimming variants: no trimming, precision-based,
-largest, and middle, with far less subjective criteria for determining trimming
-thresholds. However, we relegate the details of these trimming procedures to
-the Appendix \ref{sec:trimming}, or refer readers to the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
-code}.
-
-\textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
-with a polynomial. The choice of degree has been known to dramatically impact
-certain analyses involving the spectral rigidity or level
-variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
-moralesImprovedUnfoldingDetrending2011, abueleninSpectralUnfoldingChaotic2018,
-fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}. We
-thus compute the rigidity and level variance with all possible combinations of
-trimming and unfolding polynomial degrees of 3, 5, 7, and 9.
-
-\subsubsection{Smoothed Eigenfeatures}
-\label{sec:eigs-only}
-
-When computing RMT eigenfeatures, the polynomial fitting during unfolding
-has a smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
-operations. However, it is possible other, simpler
-transformations of the eigenvalues (e.g. uniform or Savitsky-Golay smoothing)
-might have equal or greater predictive utility. We thus also test smoothed
-variants of the eigenvalues as predictive features: we smooth the sorted
-eigenvalues with a uniform (moving average) filter and Savitsky-Golay, using
-window sizes of 3, 5, 7, and 9 to yield 8 total ``smoothed eigenvalue'' features.
-
-\subsubsection{Feature Combinations}
-
-It could easily be the case that a combination of RMT and non-RMT eigenfeatures
-are more predictive than either alone. We thus test a variety of eigenfeature
-combinations, with the motivation for these particular combinations being to
-test whether RMT eigenfeatures add anything beyond smoothing. Combined features
-are formed by concatenation, so that if we have features \(f_1, \dots,  f_n\)
-with dimensions \(p_1, \dots, p_n\), then the combined feature is \([f_1;
-\dots; f_n]\) with \(\sum p_i\) dimensions.  The final combinations chosen can
-be found in Table \ref{tab:features}.
-
-\subsubsection{Slicing Variants}
-
-We might expect the largest, middle, or smallest eigenvalues to be most useful
-in characterizing some system. E.g. if the smallest eigenvalues correspond to
-random / noise aspects of a system, but it is differences in the nature of the
-noise that most differentiate between certain systems, then here the smallest
-eigenvalues may have the most predictive utility.
-
-Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
-locality at which we summarize the spectrum: at small \(L\)-values, the
-spectral rigidity summarizes the non-linearity of eigenvalues that are
-relatively close to each other in magnitude. At large values of \(L\), the rigidity
-summarizes the long-range non-linearity of the spectrum.
-
-Since predictive utility may vary with summary locality, or with the region of
-the original spectrum, we investigate various front, middle, and end contiguous
-slices of each eigenfeature in all analyses, where the size of each slice is
-either the first or last 5\%, 10\%, or 20\% of the full eigenfeatures for
-non-middle slices, or the middle 10\%, 20\%, or 40\% of the full eigenfeature for
-the middle eigenvalues.
-
-For combined features, slicing variants are also computed, with slicing
-performed first on each feature to be combined. E.g. if we have features \(f_1,
-\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
-feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
-elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
-\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
-
-\subsubsection{Baseline Features}
-\label{sec:baselines}
-
-If RMT- or eigenvalue-based features fail to predictively outperform features
-that are simple summary statistics of the fMRI data, it is difficult to justify
-the greater computational and interpretational complexities of the former. We
-compute ten simple statistical reductions of the voxel time series ("timeseries
-features") to help assess the relative value of RMT features.
-
-Baseline reductions use as statistical summaries: robust measures of location
-(mean, max, min); non-robust measures of location (median, \(95^{\text{th}}\)
-percentile, \(5^{\text{th}}\) percentile); non-robust measures of scale
-(standard deviation, range); and robust measures of scale (interquartile range,
-difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
-percentile). These statistics are used to summarizing / reduce the fMRI data
-along the voxel dimension, yielding a feature of \(T\) dimensions. E.g. the ``mean''
-feature is just the global mean signal. Exact details can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
-source code}.
-
-We do not take slice variants of these baseline features, since the baseline
-features are still time series. However, we do also evaluate a number of
-smoothing degrees of these features, to account for noise and to be somewhat
-similar to the various smoothing variants for the eigenfeatures. Each baseline
-feature thus has a uniformly smoothed variant, where the size of the smoothing
-window is either 1 (no smoothing), 2, 4, 8, or 16.
-
-\subsection{Classifiers}
-
-We use a variety of stadard machine-learning classifiers available in the
-Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
-solve each classification task. We use a gradient-boosted decision tree (GBDT),
-random forest (RF) classifier, support vector classifier (SVC) with radial
-basis function, and k-nearest neighbours classifiers with k equal to 3, 5, and
-9 (KNN3, KNN5, and KNN9, respectively). We originally also attempted to test a
-simple logistic regression classifier, but found that this model to frequently
-failed to converge for a number of features, even after increasing iterations
-significantly. In all cases the default hyperparameter values were used.
-
-\subsection{Preprocessing Levels}
-
-Each preprocessing intermediate in the (1) brain extraction, (2) slice time
-correction, (3) motion correction, and (4) template registration pipeline was
-saved and used for entirely separate feature analyses. E.g. all features were
-extracted from a brain prepared with one of four levels or degrees of
-preprocessing, where level \(k\) includes preprocessings steps 1 to \(k\),
-inclusive, and starting at \(k = 1\).
-
-\subsection{Normalization}
-
-Because of the exponential distribution of the eigenfeatures, normalization is
-somewhat non-trivial, and a simple standardization or min-max normalization is
-ineffective. We instead first apply a logarithm to all eigenfeatures, and then
-test each classifier with un-normalized or min-max normalized versions of the
-log-features. Baseline features are also tested with un-normalized and min-max
-normalized versions, but without any log transform.
-
-\begin{table}[h!]
-\caption{
-    Feature groupings for summarization. eigs = non-RMT eigenfeatures and their combinations with
-    other non-RMT eigenfeatures. rmt = RMT eigenfeatures and their combinations. tseries = baseline
-    timeseries reductions.\\
-}
-\label{tab:features}
-\small
-\centering
-\begin{tabular}{lllr}
-\hline
-\textbf{Gross Source} & \textbf{Feature Group} & \textbf{Feature id}   \\
-\hline
-eigs           & eigs          & \tcode{eigs}                           \\
-               & eigs max      & \tcode{eigsminmax10}                   \\
-               &               & \tcode{eigsminmax20}                   \\
-               &               & \tcode{eigsminmax5}                    \\
-               & eigs middle   & \tcode{eigsmiddle10}                   \\
-               &               & \tcode{eigsmiddle20}                   \\
-               &               & \tcode{eigsmiddle40}                   \\
-               & eigs smooth   & \tcode{eigs + eigs\_smooth}            \\
-               &               & \tcode{eigs + savgol}                  \\
-               &               & \tcode{eigs\_savgol}                   \\
-               &               & \tcode{eigs\_smooth}                   \\
-               &               &                                        \\
-rmt            & rmt + eigs    & \tcode{eigs + levelvar}                \\
-               &               & \tcode{eigs + rigidity}                \\
-               &               & \tcode{eigs + rigidity + levelvar}     \\
-               &               & \tcode{eigs + unfolded}                \\
-               &               & \tcode{eigs + unfolded + levelvar}     \\
-               &               & \tcode{eigs + unfolded + rigidity}     \\
-               & rmt only      & \tcode{levelvar}                       \\
-               &               & \tcode{rigidity}                       \\
-               &               & \tcode{rigidity + levelvar}            \\
-               &               & \tcode{unfolded}                       \\
-               &               & \tcode{unfolded + levelvar}            \\
-               &               & \tcode{unfolded + rigidity}            \\
-               &               & \tcode{unfolded + rigidity + levelvar} \\
-               &               &                                        \\
-tseries        & location      & \tcode{T-max}                          \\
-               &               & \tcode{T-mean}                         \\
-               &               & \tcode{T-med}                          \\
-               &               & \tcode{T-min}                          \\
-               &               & \tcode{T-p05}                          \\
-               &               & \tcode{T-p95}                          \\
-               & scale         & \tcode{T-iqr}                          \\
-               &               & \tcode{T-rng}                          \\
-               &               & \tcode{T-rrng}                         \\
-               &               & \tcode{T-std}                          \\
-\hline
-\end{tabular}
-\end{table}
-
-\subsection{Multiverse Analysis}
-\label{sec:multiverse}
-
-We perform a \textit{multiverse
-analysis}\cite{steegenIncreasingTransparencyMultiverse2016} to assess the
-overall predictive potential of the various eigenfeatures across all
-previously-mentioned analytic choices. That is, for each combination of
-comparison task and classifier, we use 5-fold cross validation, and use the
-mean AUROC (mAUROC) across folds as our metric to evaluate feature predictive utility.
-
-Each non-baseline feature is evaluated, for each classification task and
-classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\)
-2 normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
-degrees  \(\times\) 10 slice choices. Each baseline feature is evaluated
-exactly 40 times for each comparison task and classifier: 4 preprocessing
-levels \(\times\) 2 normalization methods \(\times\) 5 smoothing window sizes.
-In total, with all the datasets examined in this study, this yields 2 053 920
-mAUROC values to summarize, with 5 main \textit{analytic factors}:
-preprocessing, normalization, trimming, degree, and slicing.
-
-\section{Results}
-
-Summarizing this amount of data requires some caution. Measures of location or
-scale, even when robust, are, for the most part, misleading and uninformative when distributions
-are skewed, and summarizing too broadly is likewise uninformative. E.g. the
-average mAUROC across all over \(2 \times 10^6\) model evaluations is too
-broad a summary to meaningfully interpret. Likewise, each analytic factor
-could produce an interaction: e.g. perhaps
-a meaningful number.
-
-Our multiverse analysis can be structured semi-hierarchically, where choices at each level
-in the hierarchy impact the final distribution of AUROCs observed. We have a
-number of \textit{datasets}, and each dataset has one or more
-\textit{classification tasks}. Each classification task can be solved with a
-\textit{classifier}. Each classifier fit operates on a \textit{feature}, and
-each feature is extracted via a certain \textit{preprocessing degree} and
-various \textit{analytic choices} (slicing, trimming, unfolding degree,
-normalization).
-
-\section{Results}
-\label{sec:results}
+\begin{spacing}{1}   % use single spacing for supplementary
 
@@
 \begin{center}
-\begin{tabular}{c}
-\includegraphics[width=6.5in]{gross_feature_overall_by_subgroup.png}
-\end{tabular}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{all_by_fine_feature_groups.png}
 \end{center}
 \caption
-{ \label{fig:main-results}
+{ \label{fig:feature-group-all}
 AUROC distributions across gross feature groupings and comparison tasks.}
@@
 
-First not non-predictable data, and then no longer discuss it in subsequent features.
-
-\subsection{Normalization}
-NO meaningful effect of normalization on AUROC distributions at level of gross feature, or feature
-group, so can safely ignore this.
-NO meaningful effect of normalization for any feature group by classifier either, so
-normalization doesn't seem to matter here.
-Maybe add table of correlations by groupings here
-
-\subsection{Classifier}
-When the data is predictable, timeseries features almost all have the mode / max <= 0.5 for all
-classifiers and subgroups, whereas eigenfeatures (RMT, eigenvalue only, eigenvalue + RMT) more often
-than not have mode greater than 0.5, and bulk of the distribution > 0.5.
-
-KNN3 consistently has least favourable AUROC distributions across subgroups and features
-
-\subsection{Page Setup and Fonts}
-
-All text and figures, including footnotes, must fit inside a text area 6.5 in.\ wide by 9 in.\ high (16.51 by 22.86 cm).
-Manuscripts must be formatted for US letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1 in.\ on the bottom, and 1 in.\ on the left and right.
-
-The Times New Roman font is used throughout the manuscript, in the sizes and styles shown in Table~\ref{tab:fonts}.
-If this font is not available, use a similar serif font.
-The manuscript should not contain headers or footers.
-Pages should be numbered.
-
-\begin{table}[ht]
-\caption{Fonts sizes and styles.}
-\label{tab:fonts}
-\begin{center}
-\begin{tabular}{|l|l|} %% this creates two columns
-%% |l|l| to left justify each column entry
-%% |c|c| to center each column entry
-%% use of \rule[]{}{} below opens up each row
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
-\hline\hline
-\rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\section{Parts of Manuscript}
-
-This section describes the normal structure of a manuscript and how each part should be handled.
-The appropriate vertical spacing between various parts of this document is achieved in LaTeX through the proper use of defined constructs, such as \verb|\section{}|.
-
-\subsection{Title and Author Information}
-\label{sect:title}
-The article title appears left justified at the top of the first page.
-The title font is 16 pt., bold.
-The rules for capitalizing the title are the same as for sentences; only the first word, proper nouns, and acronyms should be capitalized.
-Do not begin titles with articles (for example, a, an, the) or prepositions (for example, on, by, etc.).
-The word ``novel'' should not appear in the title, as publication will imply novelty.
-Avoid the use of acronyms in the title, unless they are widely understood.
-Appendix A contains more about acronyms.
-
-The list of authors immediately follows the title, 18 points below.
-The font is 12 pt., bold and the author names are left justified.
-The author affiliations and addresses follow the names, in 10-pt., normal font and left justified.
-For multiple affiliations, each affiliation should appear on a separate line.
-Superscript letters (a, b, c, etc.) should be used to associate multiple authors with their respective affiliations.
-The corresponding author should be identified with an asterisk, and that person's email address should be provided below the keywords.
-
-\subsection{Abstract}
-The abstract should be a summary of the paper and not an introduction.
-Because the abstract may be used in abstracting journals, it should be self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the objectives, methodology used, results obtained, and their significance.
-Please note that the following journals require the use of structured abstracts in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the\textit{ Journal of Medical Imaging}.
-Structured abstracts are encouraged for the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}.
-Helpful guidelines for structured abstracts are available on the website of the journal.
-
-\subsection{Subject terms/Keywords}
-Keywords are required. Please provide 3-6 keywords related to your paper.
-
-\subsection{Body of Paper}
-The body of the paper consists of numbered sections that present the main findings. These sections should be organized to best present the material.
-
-To provide transition elements in your paper, it is important to refer back (or forward) to specific sections.
-Such references are made by indicating the section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1 contained a description...'' If the word Section, Reference, Equation, or Figure starts a sentence, it is spelled out.
-When occurring in the middle of a sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
-
-At the first occurrence of an acronym, spell it out followed by the acronym in parentheses, for example, charge-coupled diode (CCD).
-
-\subsection{Footnotes}
-Textual footnotes should be used rarely to present important documentary or explanatory material whose inclusion in the text would be distracting.\footnote{Example of a footnote.} Due to problems with HTML display, use of footnotes should generally be avoided.
-If absolutely necessary, the footnote mark must come at the end of a sentence.
-To insert a footnote, use the {\verb|\footnote{}|} command.
-
-\subsection{Appendices}
-Brief appendices may be included when necessary, such as derivations of equations, proofs of theorems, and details of algorithms.
-Equations and figures appearing in appendices should continue sequential numbering from earlier in the paper.
-
-\subsection{Disclosures}
-Conflicts of interest should be declared under a separate header, above Acknowledgments.
-If the authors have no competing interests to declare, then a statement should be included declaring no conflicts of interest.
-For assistance generating a disclosure statement, see the form available from  the International Committee of Medical Journal Editors website: \linkable{http://www.icmje.org/conflicts-of-interest/}
-
-\subsection{Acknowledgments}
-Acknowledgments and funding information should be added after the conclusion, and before references.
-Include grant numbers and the full name of the funding body.
-The acknowledgments section does not have a section number.
-
-\subsection{Data, Materials, and Code Availability}
-Availability of data, materials, and/or code used in the research results reported in the manuscript may be declared under the heading ``Data, Materials, and Code Availability,'' following the Acknowledgments section.
-As relevant, provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).
-
-\subsection{References}
-The References section lists books, articles, and reports that are cited in the paper.
-This section does not have a section number.
-The references are numbered in the order in which they are cited.
-Examples of the format to be followed are given at the end of this document.
-
-The reference list at the end of this document is created using BibTeX, which looks through the file {\ttfamily report.bib} for the entries cited in the LaTeX source file.
-The format of the reference list is determined by the bibliography style file {\ttfamily spiejour.bst}, as specified in the \\ \verb|\bibliographystyle{spiejour}| command.
-Alternatively, the references may be directly formatted in the LaTeX source file.
-
-For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.
-Journal-article references \citenum{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.
-A reference to a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.
-References to an article in the SPIE Proceedings may include the conference name, as shown in Ref.~\citenum{Hanson93c}.
-
-The references are numbered in the order of their first citation.
-Citations to the references are made using superscripts, as demonstrated in the preceding paragraph.
-One may also directly refer to a reference within the text, for example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more references should be separated by a comma with no space between them.
-Multiple sequential references should be displayed with a dash between the first and last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
-
-\subsubsection{Reference linking and DOIs}
-A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a digital object, such as a journal article or a book chapter, that provides a persistent link to its location on the internet.
-The use of DOIs allows readers to easily access cited articles.
-Authors should include the DOI at the end of each reference in brackets if a DOI is available.
-See examples at the end of this manuscript.
-A free DOI lookup service is available from CrossRef at \\\linkable{http://www.crossref.org/freeTextQuery/}.
-The inclusion of DOIs will facilitate reference linking and is highly recommended.
-
-In the present LaTeX template, the author needs to add the DOI reference by including it in a ``note'' in the bibliography file, as shown in the file {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
-The DOI may be used by the reader to locate that document with the link: {\verb+http://dx.doi.org10.1117/12.154577+}.
-
-\subsection{Biographies}
-A brief professional biography of approximately 75 words may be provided for each author, if available.
-Biographies should be placed at the end of the paper, after the references.
-Personal information such as hobbies or birthplace/birthdate should not be included.
-Author photographs are not published.
-
-\section{Section Formatting}
-\label{sect:sections}
-In LaTeX, a new section is created with the \verb|\section{}| command, which automatically numbers the sections.
-Sections will be numbered sequentially, starting with the first section after the abstract, except for the acknowledgments and references.
-(Note that numbering of section headings is not required, but the numbering must be consistent if used.) All section headings should be left justified.
-
-Main section headings are in 12-pt. bold font, left-justified and in title case, where important words are capitalized.
-
-Paragraphs that immediately follow a section heading are leading paragraphs and should not be indented, according to standard publishing style.
-The same goes for leading paragraphs of subsections and sub-subsections.
-Subsequent paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation.
-There is no additional space between paragraphs.
-In LaTeX, paragraphs are separated by blank lines in the source file.
-Indentation of the first line of a paragraph may be avoided by starting it with \verb|\noindent|.
-
-\subsection{Subsection Headings}
-All important words in a subsection (level 1) header are capitalized.
-Subsection numbers consist of the section number, followed by a period, and the subsection number within that section, without a period at the end.
-The heading is left justified and its font is 12 pt.
-italic.
-
-\subsubsection{Sub-subsection headings}
-The first word of a sub-subsection is capitalized.
-The rest of the text is not capitalized, except for proper names and acronyms (the latter should only be used if well known).
-The heading is left justified and its font is 11 pt.
-italic.
-
-\section{Figures and Tables}
-
-\subsection{Figures}
-
-Figures are numbered in the order in which they are called out in the text.
-They should appear in the document in numerical order and as close as possible to their first reference in the text.
-It may be necessary to move figures or tables around to enhance readability.
-LaTeX will attempt to place figures at the top or bottom of a page in which they are first referenced.
-
-Figures, along with their captions, should be separated from the main text by  0.2 in.\ or 5 mm and centered.
-Figure captions are centered below the figure or graph.
-Figure captions start with the abbreviation ``Fig'' in front of the figure number, followed by a period, and the text in 10-pt.
-font.
-See Fig.~\ref{fig:example} for an example.
-
 \begin{figure}
 \begin{center}
-\begin{tabular}{c}
-\includegraphics[height=5.5cm]{mcr3b.eps}
-\end{tabular}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_by_preproc_predictive_subgroup.png}
 \end{center}
 \caption
-{ \label{fig:example}
-Example of a figure caption. }
+{ \label{fig:fine-preproc}
+AUROC distributions across fine feature groupings and predictable comparison tasks, with effect of preprocessing.}
 \end{figure}
-
-Authors may wish to create figures consisting of two or more images, in which case, they should be neatly arranged in a rectangular array.
-In no case, should the article's text be wrapped around a figure.
-Figure~\ref{fig:example2} shows two side-by-side images.
-When a figure contains more than one image, the author must submit them as a single image file.
-Further details about figure formatting can be found in the author guidelines for each specific SPIE journal: \\
-\linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
-\\
-(a) \hspace{5.1cm} (b)
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:example2}
-Example of a figure containing multiple images: (a) sun and (b) blob. Figures containing multiple images must be submitted to SPIE as a single image file.}
-\end{figure}
-
-% \subsection{Tables}
-% Tables are numbered in the order in which they are referenced.
-They should appear in the document in numerical order and as close as possible to their first reference in the text.
-It is preferable to have tables appear at the top or bottom of the page, if possible.
-Table captions are handled identically to those for figures, except that they appear above the table.
-See Table~\ref{tab:fonts} for an example.
-
-% \subsection{Multimedia}
-% Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player.
-The recommended maximum size for each multimedia file is 10-12 MB.
-Authors must insert a representative still image from the video file in the manuscript as a figure.
-The caption label will be linked by the publisher to the actual video file.
-The video may also be mentioned in an existing figure caption.
-Multimedia files are treated in the same manner as figures and they will be numbered sequentially with normal figures.
-The video number, file type, and file size should be included in parentheses at the end of the figure caption.
-See Figure \ref{vid:satellite} for an example.
-
-% \begin{video}
-% \begin{center}
-% {\includegraphics[height=5cm]{satellite.eps}}
-% \\
-% \end{center}
-% \caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
-% \end{video}
-
-\appendix    % this command starts appendixes
-
-\section{Correlation Eigenvalues via Transposition}
-\label{sec:transpose}
-
-Let \(\mathbf{X}\) is a real \(n \times p\) matrix with \(n > p\). Let
-\[
-\mathbf{Z} = \text{norm}(\mathbf{X}) =
-\left(
-\mathbf{X}_1 - \bar{\mathbf{X}}_1 | \dots | \mathbf{X}_p - \bar{\mathbf{X}}_p
-\right)
-\]
-where \(\mathbf{X}_i\) denotes column \(i\) of \(\mathbf{X}\). Denote the (unordered) set
-of eigenvalues of \(\mathbf{X}\) as \(\text{eigs}(\mathbf{X})\), and let \(r = (p - 1)^{-1}\).
-Denote the covariance matrix of \(\mathbf{X}\) as \(\text{cov}(\mathbf{X})\).  Then:
-
-\begin{align*}
-\text{eigs}\left( \text{cov}(\mathbf{X})  \right)
-&= \text{eigs}(r \cdot \mathbf{Z}\mathbf{Z}^{\top}) \\
-&= r \cdot \text{eigs}(\mathbf{Z}\mathbf{Z}^{\top}) \\
-&= r \cdot \text{eigs}((\mathbf{Z}\mathbf{Z}^{\top})^{\top}) \\
-&= r \cdot \text{eigs}(\mathbf{Z}^{\top}\mathbf{Z}) \\
-\end{align*}
-
-Denote the correlation matrix of \(\mathbf{X}\) as \(\text{corr}(\mathbf{X})\), and let
-\[
-\mathbf{Y} = \text{stand}(\mathbf{X}) =
-\left(
-(\mathbf{X}_1 - \bar{\mathbf{X}}_1) / \sigma_1 | \dots | (\mathbf{X}_p - \bar{\mathbf{X}}_p) / \sigma_p
-\right)
-\]
-
-Then
-
-\[
-\text{eigs}\left( \text{corr}(\mathbf{X})  \right)
-= \text{eigs}\left( \text{cov}(\mathbf{Y})  \right)
-= r \cdot \text{eigs}(\mathbf{Y}^{\top}\mathbf{Y})
-\]
-
-\section{Trimming Procedures}
-\label{sec:trimming}
-
-We implement three trimming procedures: precision-based, largest, and middle
-trimming. The
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
-code} is the definitive reference for the procedures, but we describe the
-motivations belows.
-
-In precision-based trimming, we trim away any eigenvalues that are close enough
-to zero to be considered a result of numerical error due to floating point
-representation. There are two thresholds we consider, including the one used by
-NumPy\cite{harrisArrayProgrammingNumPy2020} for the
-\href{https://numpy.org/doc/stable/reference/generated/numpy.linalg.matrix_rank.html}{determination
-of matrix rank}, and those recommended by LAPACK\cite{laug} in their user
-guide\cite{andersonLAPACKUsersGuide1999a} on the
-\href{https://netlib.org/lapack/lug/node89.html}{error bounds for symmetric
-eigenproblems}, and related
-\href{https://netlib.org/lapack/lug/node90.html}{additional details}. We trim
-each matrix eigenvalues to whichever threshold is largest for the matrix in
-question.
-
-For largest trimming, we must determine a threshold at which to separate
-"large" from "small" eigenvalues. The eigenvalues for our data tended to grow
-exponentially, so we instead looked at thresholding on the logarithms. We then
-used \(k\)-means with \(k=2\) on the precision-trimmed eigenvalues, and took
-the largest cluster (which also always had the smaller mean) cluster as the
-"largest" eigenvalues to trim away. "Middle" trimming simply reflects the
-threshold found by the largest trim method, e.g. if the largest trim method
-removes the last \(n\) precision-trimmed eigenvalues, then we also trim the
-first \(n\) smallest eigenvalues remaining after precision-trimming.
-
-We chose one-dimensional k-means partly due to efficiency and simplicity, and
-because of the general relation to classical threholding methods like the Otsu
-method\cite{liuOtsuMethodKmeans2009}.
-
-% \section{Miscellaneous Formatting Details}
-% \label{sect:misc}
-% At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.
-
-% \subsection{Formatting Equations}
-% Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.
-Important equations appear on their own line.
-Such equations are centered.
-For example, ``The expression for the field of view is
-% \begin{equation}
-% \label{eq:fov}
-% 2 a = \frac{(b + 1)}{3c} \, ,
-% \end{equation}
-% where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.
-
-% Equations are considered to be part of a sentence and should be punctuated accordingly.
-In the above example, a comma appears after the equation because the next line is a subordinate clause.
-If the equation ends the sentence, a period should follow the equation.
-The line following an equation should not be indented unless it is meant to start a new paragraph.
-Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.
-
-% References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise.
-Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..
-
-% \subsection{Formatting Theorems}
-
-% To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.
-The text of the theorem continues on the same line in normal, 10-pt.
-font, achieved in LaTeX using \verb|\footnotesize|.
-For example,
-
-% \vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}
-
-% \disclosures
-\subsection*{Disclosures}
-Conflicts of interest should be declared under a separate header.
-If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.
-
-\subsection* {Acknowledgments}
-This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.
-
-\subsection* {Data, Materials, and Code Availability}
-As relevant, the availability of data, materials, and/or software code used in the research results reported in the manuscript may be declared in this section.
-(Note: this section is required for the \textit{Journal of Biomedical Optics} and \textit{Neurophotonics}.) Provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).
-
-%%%%% References %%%%%
 
@@
 
-%%%%% Biographies of authors %%%%%
-
-\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering.
-He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.
-He is the author of more than 50 journal papers and has written three book chapters.
-His current research interests include optical interconnects, holography, and optoelectronic systems.
-He is a member of SPIE.
-
-\vspace{1ex}
-\noindent Biographies and photographs of the other authors are not available.
-
-\listoffigures
-\listoftables
-
 \end{spacing}

### 2022-11-09 code/scripts/summarize_everything_updated.py

@@ -149,7 +149,7 @@ OVERALL_PREDICTIVE_GROUP_ORDER = [
     "Aging - younger v older",
     "Osteo - nopain v duloxetine",
     "Osteo - nopain v pain",
-    "Osteo - pain v duloxetine",
+    # "Osteo - pain v duloxetine",
     "Parkinsons - ctrl v park",  # questionable, only tseries seem predictive
     "TaskAttention - task_attend v task_nonattend",
     "Vigilance - vigilant v nonvigilant",
@@ -241,9 +241,9 @@ def load_all_renamed() -> DataFrame:
     )
     df = df.loc[~dupes]
     df["subgroup"] = df["data"] + " - " + df["comparison"]
-    df["feature_group"] = df["feature"].apply(feature_grouping)
+    df["fine_feature"] = df["feature"].apply(fine_feature_grouping)
     df["slice_group"] = df["slice"].apply(slice_grouping)
-    df["gross_feature"] = df["feature"].apply(gross_feature_grouping)
+    df["coarse_feature"] = df["feature"].apply(coarse_feature_grouping)
     return df
 
 
@@ -279,9 +279,9 @@ def load_tseries() -> DataFrame:
     )
     df = df.loc[~dupes]
     df["subgroup"] = df["data"] + " - " + df["comparison"]
-    df["feature_group"] = df["feature"].apply(feature_grouping)
+    df["fine_feature"] = df["feature"].apply(fine_feature_grouping)
     df["slice_group"] = df["slice"].apply(slice_grouping)
-    df["gross_feature"] = df["feature"].apply(gross_feature_grouping)
+    df["coarse_feature"] = df["feature"].apply(coarse_feature_grouping)
     return df
 
 
@@ -366,7 +366,7 @@ def is_eigs_only(s: str) -> bool:
     return s == "eigs" or ("middle" in s)
 
 
-def feature_grouping(s: str) -> str:
+def fine_feature_grouping(s: str) -> str:
     if is_ts_loc(s):
         return "tseries loc"
     if is_ts_scale(s):
@@ -385,7 +385,7 @@ def feature_grouping(s: str) -> str:
         return "rmt + eigs"
 
 
-def gross_feature_grouping(s: str) -> str:
+def coarse_feature_grouping(s: str) -> str:
     if is_ts_loc(s) or is_ts_scale(s):
         return "tseries"
     if is_rmt_only(s) or is_rmt_plus(s):
[…]

### 2022-11-10 paper/paper/SPIE/article.tex
[WIP]

@@
 
-Summarizing this amount of data requires some caution. Measures of location or
+Summarizing this quantity of data requires some caution. Measures of location or
 scale, even when robust, are, for the most part, misleading and uninformative
 when distributions are skewed. In most cases we find skewed mAUROC
-distributions, and so have chosen primarily to present our findings visually.
+distributions, and so have chosen primarily to present our findings visually,
+with kernel density estimates.
 
@@
 certain preprocessing level and for certain feature slicing. However, there are
-too many potential interactions to present, so we limit our presentation
-to the most significant ones.
-
-As there are a large number of figures, we summarize patterns of results across
-two abstract levels of feature groupings—"coarse", and "fine"—shown in
-Table \ref{tab:features}.
+too many potential interactions to present, so we limit our presentation to the
+most significant ones.  As there are a large number of features, we also
+summarize patterns of results across two abstract levels of feature
+groupings—"coarse", and "fine"—shown in Table \ref{tab:features}.
 
@@
 
+Eigenfeatures most strongly and consistently demonstrated predictive utility
+in the \textit{Aging} dataset \code{older v younger} classification task, the
+\textit{Osteo} dataset \code{nopain v duloxetine} task, and in the \textit{Attention}
+\code{vigilant v nonvigilant} comparison.
+
 \begin{figure}
@@
 
+\subsection{Largest mAUROCs}
+Considering the various analytic choices as tunable parameters, it makes sense
+to examine the \textit{largest} portion of mAUROCs as an indication of the
+maximum predictive potential of the eigenfeatures. In this case, it is clear
+that eigenfeatures using RMT features almost always had the highest potential
+predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
+shows that this was primarily due to either the "rmt only" or "rmt + eigs"
+features (see Table \ref{tab:features}). However, these features were also most
+likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
+0.5; Figure S\ref{S-fig:fine-smallest}).
+
+Examining these RMT features more closely (Figure S\ref{S-fig:fig:fine-degree}),
+it is clear these features are more predictive mostly due to the unfolding
+procedure: the mAUROC distributions with largest medians and modes are due to
+features that include the unfolded eigenvalues. Combined features that use
+the unfolded eigenvalues plus some other RMT eigenfeature have identical
+mAUROC distributions to those using the unfolded eigenvalues alone.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{coarse_feature_largest_by_subgroup.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:largest} Distributions of largest 500 mAUROCs by coarse feature grouping.}
+\end{figure}
+
 \subsection{Effect of Preprocessing}
-At a coarse level
-
-  if a certain feature is simply not useful, or if a certain comparison
-task is not solvable with some or even all classifiers
+At a coarse level, slice time correction followed by motion correction tended
+to slightly increase the predictive utility of the eigenfeatures (Figure
+\ref{fig:preproc}) relative to brain extraction only. Subsequent registration
+following these steps did not generally impact mAUROC distributions further,
+except in the \code{task\_attend v task\_nonattend} comparison, where
+registration reduced the predictive utility of the eigenfeatures (Figure
+\ref{fig:preproc}, second-last column). When examining features more finely, it
+is clear that preprocessing most impacts the mAUROC distribution of the largest
+and central eigenvalues ("eigs middle" and "eigs max" in Figure
+S\ref{S-fig:fine-preproc})
+
+\subsection{Effect of Classifier}
+mAUROC distributions were generally similar across coarse or fine feature
+groupings (Figure S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}).
+In the \textit{Attention} \code{task\_attend v task\_nonattend} condition, eigenfeatures were modally predictive only with
+the random forest classifier, whereas in the \textit{Aging} data, a SVC was least
+likely to have mAUROC \(<\) 0.5. Overall, however, differences in the mAUROC distributions
+were small.
+
+\subsection{Effect of Trimming}
+Trimming based on numerical precision (see Section \ref{sec:rmt-features} and
+Appendix \ref{sec:trimming}), did not result in meaningfully different mAUROC
+distributions in any case (Figure S\ref{S-fig:fine-trim}).   While there was no
+clearly observable differences between the mAUROC distributions depending on
+whether
+
+\subsection{Effect of Unfolding Degree}
+The choice of polynomial unfolding degree significantly impacted the mAUROC
+distributions for most classification tasks and most RMT features, and most
+significantly for the level variance features (Figure
+S\ref{S-fig:fine-degree}). Overall, Figure S\ref{S-fig:fine-degree} weakly
+suggests that either smaller (degree = 3) or larger (degree = 9) unfolding
+degrees tend to yield the most predictively useful RMT features. However, when
+restricting to the most predictive RMT features (those including the unfolded
+eigenvalues), it seems clear from Figure S\ref{S-fig:fine-degree} the the
+largest unfolding degree of 9 produces the most favourable mAUROC
+distributions.
 
@@
 
-\subsection{Classifier}
-When the data is predictable, timeseries features almost all have the mode / max <= 0.5 for all
-classifiers and subgroups, whereas eigenfeatures (RMT, eigenvalue only, eigenvalue + RMT) more often
-than not have mode greater than 0.5, and bulk of the distribution > 0.5.
-
-KNN3 consistently has least favourable AUROC distributions across subgroups and features
-
-\subsection{Page Setup and Fonts}
-
-All text and figures, including footnotes, must fit inside a text area 6.5 in.\ wide by 9 in.\ high (16.51 by 22.86 cm).
-Manuscripts must be formatted for US letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1 in.\ on the bottom, and 1 in.\ on the left and right.
-
-The Times New Roman font is used throughout the manuscript, in the sizes and styles shown in Table~\ref{tab:fonts}.
-If this font is not available, use a similar serif font.
-The manuscript should not contain headers or footers.
-Pages should be numbered.
-
-\begin{table}[ht]
-\caption{Fonts sizes and styles.}
-\label{tab:fonts}
-\begin{center}
-\begin{tabular}{|l|l|} %% this creates two columns
-%% |l|l| to left justify each column entry
-%% |c|c| to center each column entry
-%% use of \rule[]{}{} below opens up each row
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
-\hline\hline
-\rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
-\hline
-\rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+% \subsection{Page Setup and Fonts}
+
+% All text and figures, including footnotes, must fit inside a text area 6.5 in.\
+% wide by 9 in.\ high (16.51 by 22.86 cm). Manuscripts must be formatted for US
+% letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1
+% in.\ on the bottom, and 1 in.\ on the left and right.
+
+% The Times New Roman font is used throughout the manuscript, in the sizes and
+% styles shown in Table~\ref{tab:fonts}. If this font is not available, use a
+% similar serif font. The manuscript should not contain headers or footers. Pages
+% should be numbered.
+
+% \begin{table}[ht]
+% \caption{Fonts sizes and styles.}
+% \label{tab:fonts}
+% \begin{center}
+% \begin{tabular}{|l|l|} %% this creates two columns
+% %% |l|l| to left justify each column entry
+% %% |c|c| to center each column entry
+% %% use of \rule[]{}{} below opens up each row
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
+% \hline\hline
+% \rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
+% \hline
+% \rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
 
@@
 
-This section describes the normal structure of a manuscript and how each part should be handled.
-The appropriate vertical spacing between various parts of this document is achieved in LaTeX through the proper use of defined constructs, such as \verb|\section{}|.
+This section describes the normal structure of a manuscript and how each part
+should be handled. The appropriate vertical spacing between various parts of
+this document is achieved in LaTeX through the proper use of defined
+constructs, such as \verb|\section{}|.
 
@@
 \label{sect:title}
-The article title appears left justified at the top of the first page.
-The title font is 16 pt., bold.
-The rules for capitalizing the title are the same as for sentences; only the first word, proper nouns, and acronyms should be capitalized.
-Do not begin titles with articles (for example, a, an, the) or prepositions (for example, on, by, etc.).
-The word ``novel'' should not appear in the title, as publication will imply novelty.
-Avoid the use of acronyms in the title, unless they are widely understood.
-Appendix A contains more about acronyms.
-
-The list of authors immediately follows the title, 18 points below.
-The font is 12 pt., bold and the author names are left justified.
-The author affiliations and addresses follow the names, in 10-pt., normal font and left justified.
-For multiple affiliations, each affiliation should appear on a separate line.
-Superscript letters (a, b, c, etc.) should be used to associate multiple authors with their respective affiliations.
-The corresponding author should be identified with an asterisk, and that person's email address should be provided below the keywords.
+The article title appears left justified at the top of the first page. The
+title font is 16 pt., bold. The rules for capitalizing the title are the same
+as for sentences; only the first word, proper nouns, and acronyms should be
+capitalized. Do not begin titles with articles (for example, a, an, the) or
+prepositions (for example, on, by, etc.). The word ``novel'' should not appear
+in the title, as publication will imply novelty. Avoid the use of acronyms in
+the title, unless they are widely understood. Appendix A contains more about
+acronyms.
+
+The list of authors immediately follows the title, 18 points below. The font is
+12 pt., bold and the author names are left justified. The author affiliations
+and addresses follow the names, in 10-pt., normal font and left justified. For
+multiple affiliations, each affiliation should appear on a separate line.
+Superscript letters (a, b, c, etc.) should be used to associate multiple
+authors with their respective affiliations. The corresponding author should be
+identified with an asterisk, and that person's email address should be provided
+below the keywords.
 
 \subsection{Abstract}
-The abstract should be a summary of the paper and not an introduction.
-Because the abstract may be used in abstracting journals, it should be self-contained (i.e., no numerical references) and substantive in nature, presenting concisely the objectives, methodology used, results obtained, and their significance.
-Please note that the following journals require the use of structured abstracts in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the\textit{ Journal of Medical Imaging}.
-Structured abstracts are encouraged for the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}.
-Helpful guidelines for structured abstracts are available on the website of the journal.
+The abstract should be a summary of the paper and not an introduction. Because
+the abstract may be used in abstracting journals, it should be self-contained
+(i.e., no numerical references) and substantive in nature, presenting concisely
+the objectives, methodology used, results obtained, and their significance.
+Please note that the following journals require the use of structured abstracts
+in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of
+Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured
+abstracts are encouraged for the \textit{Journal of Micro/Nanolithography,
+MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on
+the website of the journal.
 
@@
 \subsection{Body of Paper}
-The body of the paper consists of numbered sections that present the main findings. These sections should be organized to best present the material.
-
-To provide transition elements in your paper, it is important to refer back (or forward) to specific sections.
-Such references are made by indicating the section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1 contained a description...'' If the word Section, Reference, Equation, or Figure starts a sentence, it is spelled out.
-When occurring in the middle of a sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
-
-At the first occurrence of an acronym, spell it out followed by the acronym in parentheses, for example, charge-coupled diode (CCD).
+The body of the paper consists of numbered sections that present the main
+findings. These sections should be organized to best present the material.
+
+To provide transition elements in your paper, it is important to refer back (or
+forward) to specific sections. Such references are made by indicating the
+section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1
+contained a description...'' If the word Section, Reference, Equation, or
+Figure starts a sentence, it is spelled out. When occurring in the middle of a
+sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
+
+At the first occurrence of an acronym, spell it out followed by the acronym in
+parentheses, for example, charge-coupled diode (CCD).
 
 \subsection{Footnotes}
-Textual footnotes should be used rarely to present important documentary or explanatory material whose inclusion in the text would be distracting.\footnote{Example of a footnote.} Due to problems with HTML display, use of footnotes should generally be avoided.
-If absolutely necessary, the footnote mark must come at the end of a sentence.
-To insert a footnote, use the {\verb|\footnote{}|} command.
+Textual footnotes should be used rarely to present important documentary or
+explanatory material whose inclusion in the text would be
+distracting.\footnote{Example of a footnote.} Due to problems with HTML
+display, use of footnotes should generally be avoided. If absolutely necessary,
+the footnote mark must come at the end of a sentence. To insert a footnote, use
+the {\verb|\footnote{}|} command.
 
 \subsection{Appendices}
-Brief appendices may be included when necessary, such as derivations of equations, proofs of theorems, and details of algorithms.
-Equations and figures appearing in appendices should continue sequential numbering from earlier in the paper.
+Brief appendices may be included when necessary, such as derivations of
+equations, proofs of theorems, and details of algorithms. Equations and figures
+appearing in appendices should continue sequential numbering from earlier in
+the paper.
 
 \subsection{Disclosures}
-Conflicts of interest should be declared under a separate header, above Acknowledgments.
-If the authors have no competing interests to declare, then a statement should be included declaring no conflicts of interest.
-For assistance generating a disclosure statement, see the form available from  the International Committee of Medical Journal Editors website: \linkable{http://www.icmje.org/conflicts-of-interest/}
+Conflicts of interest should be declared under a separate header, above
+Acknowledgments. If the authors have no competing interests to declare, then a
+statement should be included declaring no conflicts of interest. For assistance
+generating a disclosure statement, see the form available from  the
+International Committee of Medical Journal Editors website:
+\linkable{http://www.icmje.org/conflicts-of-interest/}
 
 \subsection{Acknowledgments}
-Acknowledgments and funding information should be added after the conclusion, and before references.
-Include grant numbers and the full name of the funding body.
-The acknowledgments section does not have a section number.
+Acknowledgments and funding information should be added after the conclusion,
+and before references. Include grant numbers and the full name of the funding
+body. The acknowledgments section does not have a section number.
 
 \subsection{Data, Materials, and Code Availability}
-Availability of data, materials, and/or code used in the research results reported in the manuscript may be declared under the heading ``Data, Materials, and Code Availability,'' following the Acknowledgments section.
-As relevant, provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).
+Availability of data, materials, and/or code used in the research results
+reported in the manuscript may be declared under the heading ``Data, Materials,
+and Code Availability,'' following the Acknowledgments section. As relevant,
+provide specific access information or restrictions for data, materials, and
+computer code (i.e., links to repository access addresses with guidance on
+commercial or public access).
 
 \subsection{References}
-The References section lists books, articles, and reports that are cited in the paper.
-This section does not have a section number.
-The references are numbered in the order in which they are cited.
-Examples of the format to be followed are given at the end of this document.
-
-The reference list at the end of this document is created using BibTeX, which looks through the file {\ttfamily report.bib} for the entries cited in the LaTeX source file.
-The format of the reference list is determined by the bibliography style file {\ttfamily spiejour.bst}, as specified in the \\ \verb|\bibliographystyle{spiejour}| command.
-Alternatively, the references may be directly formatted in the LaTeX source file.
-
-For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list of authors (initials plus last name), book title (in italics), page or chapter numbers, publisher, city, and year of publication.
-Journal-article references \citenum{Metropolis53,Harris06} include the author list, title of the article (in quotes), journal name (in italics, properly abbreviated), volume number (in bold), inclusive page numbers or citation identifier, and year.
-A reference to a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes the author list, title of the article (in quotes), conference name (in italics), editors (if appropriate), volume title (in italics), volume number if applicable (in bold), inclusive page numbers, publisher, city, and year.
-References to an article in the SPIE Proceedings may include the conference name, as shown in Ref.~\citenum{Hanson93c}.
-
-The references are numbered in the order of their first citation.
-Citations to the references are made using superscripts, as demonstrated in the preceding paragraph.
-One may also directly refer to a reference within the text, for example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more references should be separated by a comma with no space between them.
-Multiple sequential references should be displayed with a dash between the first and last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
+The References section lists books, articles, and reports that are cited in the
+paper. This section does not have a section number. The references are numbered
+in the order in which they are cited. Examples of the format to be followed are
+given at the end of this document.
+
+The reference list at the end of this document is created using BibTeX, which
+looks through the file {\ttfamily report.bib} for the entries cited in the
+LaTeX source file.  The format of the reference list is determined by the
+bibliography style file {\ttfamily spiejour.bst}, as specified in the \\
+\verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may
+be directly formatted in the LaTeX source file.
+
+For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list
+of authors (initials plus last name), book title (in italics), page or chapter
+numbers, publisher, city, and year of publication.  Journal-article references
+\citenum{Metropolis53,Harris06} include the author list, title of the article
+(in quotes), journal name (in italics, properly abbreviated), volume number (in
+bold), inclusive page numbers or citation identifier, and year.  A reference to
+a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes
+the author list, title of the article (in quotes), conference name (in
+italics), editors (if appropriate), volume title (in italics), volume number if
+applicable (in bold), inclusive page numbers, publisher, city, and year.
+References to an article in the SPIE Proceedings may include the conference
+name, as shown in Ref.~\citenum{Hanson93c}.
+
+The references are numbered in the order of their first citation. Citations to
+the references are made using superscripts, as demonstrated in the preceding
+paragraph. One may also directly refer to a reference within the text, for
+example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more
+references should be separated by a comma with no space between them. Multiple
+sequential references should be displayed with a dash between the first and
+last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
 
 \subsubsection{Reference linking and DOIs}
-A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a digital object, such as a journal article or a book chapter, that provides a persistent link to its location on the internet.
-The use of DOIs allows readers to easily access cited articles.
-Authors should include the DOI at the end of each reference in brackets if a DOI is available.
-See examples at the end of this manuscript.
-A free DOI lookup service is available from CrossRef at \\\linkable{http://www.crossref.org/freeTextQuery/}.
-The inclusion of DOIs will facilitate reference linking and is highly recommended.
-
-In the present LaTeX template, the author needs to add the DOI reference by including it in a ``note'' in the bibliography file, as shown in the file {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
-The DOI may be used by the reader to locate that document with the link: {\verb+http://dx.doi.org10.1117/12.154577+}.
+A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a
+digital object, such as a journal article or a book chapter, that provides a
+persistent link to its location on the internet. The use of DOIs allows readers
+to easily access cited articles. Authors should include the DOI at the end of
+each reference in brackets if a DOI is available. See examples at the end of
+this manuscript. A free DOI lookup service is available from CrossRef at
+\\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will
+facilitate reference linking and is highly recommended.
+
+In the present LaTeX template, the author needs to add the DOI reference by
+including it in a ``note'' in the bibliography file, as shown in the file
+{\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
+The DOI may be used by the reader to locate that document with the link:
+{\verb+http://dx.doi.org10.1117/12.154577+}.
 
 \subsection{Biographies}
-A brief professional biography of approximately 75 words may be provided for each author, if available.
-Biographies should be placed at the end of the paper, after the references.
-Personal information such as hobbies or birthplace/birthdate should not be included.
-Author photographs are not published.
+A brief professional biography of approximately 75 words may be provided for
+each author, if available. Biographies should be placed at the end of the
+paper, after the references. Personal information such as hobbies or
+birthplace/birthdate should not be included. Author photographs are not
+published.
 
@@
 \label{sect:sections}
-In LaTeX, a new section is created with the \verb|\section{}| command, which automatically numbers the sections.
-Sections will be numbered sequentially, starting with the first section after the abstract, except for the acknowledgments and references.
-(Note that numbering of section headings is not required, but the numbering must be consistent if used.) All section headings should be left justified.
-
-Main section headings are in 12-pt. bold font, left-justified and in title case, where important words are capitalized.
-
-Paragraphs that immediately follow a section heading are leading paragraphs and should not be indented, according to standard publishing style.
-The same goes for leading paragraphs of subsections and sub-subsections.
-Subsequent paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation.
-There is no additional space between paragraphs.
-In LaTeX, paragraphs are separated by blank lines in the source file.
-Indentation of the first line of a paragraph may be avoided by starting it with \verb|\noindent|.
+In LaTeX, a new section is created with the \verb|\section{}| command, which
+automatically numbers the sections. Sections will be numbered sequentially,
+starting with the first section after the abstract, except for the
+acknowledgments and references. (Note that numbering of section headings is not
+required, but the numbering must be consistent if used.) All section headings
+should be left justified.
+
+Main section headings are in 12-pt. bold font, left-justified and in title
+case, where important words are capitalized.
+
+Paragraphs that immediately follow a section heading are leading paragraphs and
+should not be indented, according to standard publishing style. The same goes
+for leading paragraphs of subsections and sub-subsections. Subsequent
+paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no
+additional space between paragraphs. In LaTeX, paragraphs are separated by
+blank lines in the source file. Indentation of the first line of a paragraph
+may be avoided by starting it with \verb|\noindent|.
 
@@
 All important words in a subsection (level 1) header are capitalized.
-Subsection numbers consist of the section number, followed by a period, and the subsection number within that section, without a period at the end.
-The heading is left justified and its font is 12 pt.
+Subsection numbers consist of the section number, followed by a period, and the
+subsection number within that section, without a period at the end. The heading
+is left justified and its font is 12 pt. italic.
+
+\subsubsection{Sub-subsection headings}
+The first word of a sub-subsection is capitalized. The rest of the text is not
+capitalized, except for proper names and acronyms (the latter should only be
+used if well known). The heading is left justified and its font is 11 pt.
 italic.
 
-\subsubsection{Sub-subsection headings}
-The first word of a sub-subsection is capitalized.
-The rest of the text is not capitalized, except for proper names and acronyms (the latter should only be used if well known).
-The heading is left justified and its font is 11 pt.
-italic.
-
 \section{Figures and Tables}
@@
 Figures are numbered in the order in which they are called out in the text.
-They should appear in the document in numerical order and as close as possible to their first reference in the text.
-It may be necessary to move figures or tables around to enhance readability.
-LaTeX will attempt to place figures at the top or bottom of a page in which they are first referenced.
-
-Figures, along with their captions, should be separated from the main text by  0.2 in.\ or 5 mm and centered.
-Figure captions are centered below the figure or graph.
-Figure captions start with the abbreviation ``Fig'' in front of the figure number, followed by a period, and the text in 10-pt.
-font.
-See Fig.~\ref{fig:example} for an example.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[height=5.5cm]{mcr3b.eps}
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:example}
-Example of a figure caption. }
-\end{figure}
-
-Authors may wish to create figures consisting of two or more images, in which case, they should be neatly arranged in a rectangular array.
-In no case, should the article's text be wrapped around a figure.
-Figure~\ref{fig:example2} shows two side-by-side images.
-When a figure contains more than one image, the author must submit them as a single image file.
-Further details about figure formatting can be found in the author guidelines for each specific SPIE journal: \\
-\linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
-\\
-(a) \hspace{5.1cm} (b)
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:example2}
-Example of a figure containing multiple images: (a) sun and (b) blob. Figures containing multiple images must be submitted to SPIE as a single image file.}
-\end{figure}
+They should appear in the document in numerical order and as close as possible
+to their first reference in the text. It may be necessary to move figures or
+tables around to enhance readability. LaTeX will attempt to place figures at
+the top or bottom of a page in which they are first referenced.
+
+Figures, along with their captions, should be separated from the main text by
+0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or
+graph. Figure captions start with the abbreviation ``Fig'' in front of the
+figure number, followed by a period, and the text in 10-pt. font. See
+Fig.~\ref{fig:example} for an example.
+
+% \begin{figure}
+% \begin{center}
+% \begin{tabular}{c}
+% \includegraphics[height=5.5cm]{mcr3b.eps}
+% \end{tabular}
+% \end{center}
+% \caption
+% { \label{fig:example}
+% Example of a figure caption. }
+% \end{figure}
+
+% Authors may wish to create figures consisting of two or more images, in which
+% case, they should be neatly arranged in a rectangular array.  In no case,
+% should the article's text be wrapped around a figure. Figure~\ref{fig:example2}
+% shows two side-by-side images. When a figure contains more than one image, the
+% author must submit them as a single image file. Further details about figure
+% formatting can be found in the author guidelines for each specific SPIE
+% journal: \\
+
+% \linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.
+
+% \begin{figure}
+% \begin{center}
+% \begin{tabular}{c}
+% \includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
+% \\
+% (a) \hspace{5.1cm} (b)
+% \end{tabular}
+% \end{center}
+% \caption
+% { \label{fig:example2} Example of a figure containing multiple images: (a) sun
+% and (b) blob. Figures containing multiple images must be submitted to SPIE as a
+% single image file.}
+% \end{figure}
 

### 2022-11-10 paper/paper/SPIE/supplementary.tex
[WIP]

@@
-% \documentclass[12pt]{supplementary}  % 12pt font required by SPIE;
 \documentclass{article}  % use this instead for A4 paper
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{coarse_feature_group_by_predictive_subgroup_classifier.png}
+\end{center}
+\caption
+{ \label{fig:coarse-classifier}
+AUROC distributions across coarse feature groupings and predictable comparison tasks, by classifier.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_group_by_predictive_subgroup_classifier.png}
+\end{center}
+\caption
+{ \label{fig:fine-classifier}
+AUROC distributions across coarse feature groupings and predictable comparison tasks, by classifier.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-largest}
+Distributions of largest 500 mAUROCs across fine feature grouping, by comparison task.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-smallest}
+Distributions of smallest 500 mAUROCs across fine feature groupings, by comparison task.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_auroc_by_trim.png}
+\end{center}
+\caption
+{ \label{fig:fine-trim}
+Distributions of mAUROCs for unfolding-dependent RMT features, by degree.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_auroc_by_degree.png}
+\end{center}
+\caption
+{ \label{fig:fine-degree}
+Distributions of mAUROCs for unfolding-dependent RMT features, by degree.}
+\end{figure}
+
 \bibliography{report}   % bibliography data in report.bib

### 2022-11-10 code/scripts/summarize_everything_updated.py

@@ -82,6 +82,18 @@ SUBGROUPERS = [
     ["data", "comparison", "classifier", "preproc", "deg", "trim", "norm"],
 ]
 
+RMT_FEATURE_PALETTE = {
+    "unfolded": GREEN,
+    "rigidity": BLUE,
+    "levelvar": PINK,
+    "rigidity + levelvar": ORNG,
+    "unfolded + levelvar": BLCK,
+    "unfolded + rigidity": GREY,
+    "unfolded + rigidity + levelvar": PURP,
+}
+
+RMT_FEATURE_ORDER = list(RMT_FEATURE_PALETTE.keys())
+
 FEATURE_GROUP_PALETTE = {
     "tseries loc": RED,
     "tseries scale": PINK,
@@ -120,6 +132,8 @@ SLICE_ORDER = [
     "min-10",
     "min-05",
 ]
+
+DEGREE_ORDER = [3, 5, 7, 9]
 
 SUBGROUP_ORDER = [
     "Bilinguality - monolingual v bilingual",
@@ -1352,6 +1366,7 @@ def despine(grid: FacetGrid) -> None:
     for ax in grid.fig.axes:
         ax.set_yticks([])
 
+
 def thinify_lines(grid: FacetGrid) -> None:
     fig: Figure = grid.fig
     for ax in fig.axes:
@@ -1360,6 +1375,7 @@ def thinify_lines(grid: FacetGrid) -> None:
 
     for line in grid.legend.legendHandles:
         line.set_linewidth(1.0)
+
 
 def dashify_gross(grid: FacetGrid) -> None:
     fig: Figure = grid.fig
@@ -1381,6 +1397,7 @@ def dashify_gross(grid: FacetGrid) -> None:
         grid.legend.legendHandles[2].set_linestyle("solid")
     except IndexError:
         pass
+
 
 def dashify_trims(grid: FacetGrid) -> None:
     fig: Figure = grid.fig
@@ -1565,6 +1582,45 @@ def make_kde_plots() -> None:
         sbn.move_legend(grid, loc=(0.79, 0.09))
         savefig(fig, "coarse_feature_largest_by_subgroup.png")
 
+    def plot_largest_by_fine_feature_subgroup() -> None:
+        dfg = df.groupby(["subgroup", "fine_feature"]).apply(
+            lambda grp: grp.nlargest(500, "auroc")
+        )
+        print("Plotting...", end="", flush=True)
+        grid = sbn.displot(
+            data=dfg,
+            x="auroc",
[…]

### 2022-11-10 paper/paper/SPIE/article.tex
[WIP]

@@
 previously-mentioned analytic choices. That is, for each combination of
-comparison task and classifier, we use 5-fold cross validation, and use the
-mean AUROC (mAUROC) across folds as our metric to evaluate feature predictive utility.
+comparison task, classifier, and analytic choices, we use 5-fold cross
+validation, and use the mean AUROC (mAUROC) across folds as our metric to
+evaluate feature predictive utility.
 
@@
 most significant ones.  As there are a large number of features, we also
-summarize patterns of results across two abstract levels of feature
-groupings—"coarse", and "fine"—shown in Table \ref{tab:features}.
-
-Finally, it makes sense to exclude classification tasks or an analytic factor
+summarize patterns of results across two abstract groupings of features
+("coarse", and "fine") shown in Table \ref{tab:features}.
+
+It is also important to exclude a classification task or an analytic factor
 from summaries if there is no evidence that the classification task is solvable
@@
 and central eigenvalues ("eigs middle" and "eigs max" in Figure
-S\ref{S-fig:fine-preproc})
+S\ref{S-fig:fine-preproc}).
 
@@
 
+\subsection{Normalization}
+There was no visible effect of normalization on mAUROC distributions at any
+summary level.
+
 \subsection{Effect of Trimming}
@@
 Appendix \ref{sec:trimming}), did not result in meaningfully different mAUROC
-distributions in any case (Figure S\ref{S-fig:fine-trim}).   While there was no
-clearly observable differences between the mAUROC distributions depending on
-whether
+distributions in any case (Figure S\ref{S-fig:fine-trim}). However, trimming
+away the largest, or both largest and smallest eigenvalues, generally had a
+significant positive effect on the predictive quality of RMT features, most
+especially for RMT features involving the unfolded eigenvalues. When employing
+these trimming methods, these features were consistently more predictive than
+not.
 
@@
 
-First not non-predictable data, and then no longer discuss it in subsequent features.
-
-\subsection{Normalization}
-NO meaningful effect of normalization on AUROC distributions at level of gross feature, or feature
-group, so can safely ignore this.
-NO meaningful effect of normalization for any feature group by classifier either, so
-normalization doesn't seem to matter here.
-Maybe add table of correlations by groupings here
+\subsection{Effect of Slicing}
+Features were most predictive when using either the largest 5 or 10\% or middle
+25\% ("max-05", "max-10" and "mid-25" in Figure
+\ref{S-fig:best-params-slicing}), and least predictive when features primarily
+involved the smallest eigenvalues.
+
+\subsection{Best Case Scenario}
+Viewing all the impactful analytic choices as tunable hyperparameters, and
+selecting the overall "best" hyperparameters of \(\text{trimming} \in
+\{\text{middle}, \text{largest}\}\), \(\text{degree} \in \{3, 9\}\),
+\(\text{preprocessing} \in \{\text{MotionCorrect}, \text{MNIRegister}\}\),
+\(\text{slicing} \in \{\text{"max-10"}, \text{"mid-25"}\} \) and
+restricting to only the "unfolded" RMT feature, results in the mAUROC
+distributions in Figure S\ref{S-fig:best-params}. Using these analytic choices,
+the \code{depress v control} comparison in the \textit{Depression} dataset
+becomes weakly predictable, and the smoothed eigenvalue features tend to
+closely follow the unfolded eigenvalues in predictive utility.
 

### 2022-11-10 paper/paper/SPIE/supplementary.tex
[WIP]

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:best-params}
+Distributions of mAUROCs when choosing ``best'' analytic choices and features.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup_and_slicing.png}
+\end{center}
+\caption
+{ \label{fig:best-params-slicing}
+Distributions of mAUROCs when choosing ``best'' analytic choices and features, by slicing.}
+\end{figure}
+
 \bibliography{report}   % bibliography data in report.bib

### 2022-11-10 code/scripts/summarize_everything_updated.py

@@ -105,6 +105,15 @@ FEATURE_GROUP_PALETTE = {
     "rmt only": BLUE,
 }
 
+NON_BASELINE_PALETTE = {
+    "eigs": BLCK,
+    "eigs max": ORNG,
+    "eigs smooth": GREEN,
+    "eigs middle": GREY,
+    "rmt + eigs": PURP,
+    "rmt only": BLUE,
+}
+
 GROSS_FEATURE_PALETTE = {
     "tseries": ORNG,
     "eigs": BLCK,
@@ -1803,6 +1812,147 @@ def make_kde_plots() -> None:
         sbn.move_legend(grid, loc=(0.76, 0.11))
         savefig(fig, "all_by_fine_feature_groups.png")
 
+    def plot_all_by_fine_feature_groups_best_params() -> None:
+        rmt = df.loc[df.coarse_feature.isin(["rmt"])]
+        rmt = rmt.loc[rmt.trim.isin(["middle", "largest"])]
+        rmt = rmt.loc[rmt.deg.isin([3, 9])]
+        rmt = rmt.loc[rmt.preproc.isin(["MotionCorrect", "MNIRegister"])]
+        rmt = rmt.loc[rmt.slice.isin(["max-10", "mid-25"])]
+        idx = rmt.feature.isin(["unfolded"])
+        rmt = rmt[idx]
+
+        other = df.loc[df.coarse_feature.isin(["eigs", "tseries"])]
+        data = pd.concat([rmt, other], axis=0)
+
+        grid = sbn.displot(
+            data=data,
+            x="auroc",
+            kind="kde",
+            hue="fine_feature",
+            hue_order=list(FEATURE_GROUP_PALETTE.keys()),
+            palette=FEATURE_GROUP_PALETTE,
+            fill=False,
+            common_norm=False,
+            col="subgroup",
+            col_order=SUBGROUP_ORDER,
+            col_wrap=4,
+            bw_adjust=1.2,
+            alpha=0.8,
+            facet_kws=dict(sharey=False, xlim=(0.1, 1.0)),
+        )
+        thinify_lines(grid)
+        add_auroc_lines(grid, kind="vline")
+        despine(grid)
+        clean_titles(grid, "classifier = ")
+        clean_titles(grid, "subgroup = ", split_at="-")
+        fig = grid.fig
+        fig.suptitle("Overall Distribution of AUROCs for each Fine Feature Group")
+        fig.tight_layout()
+        fig.set_size_inches(w=10, h=8)
+        fig.subplots_adjust(
[…]

### 2022-11-10 paper/paper/SPIE/article.tex
[WIP]

@@
 \subsection{Normalization}
-There was no visible effect of normalization on mAUROC distributions at any
-summary level.
+There was no visible effect of feature normalization on mAUROC distributions.
 
@@
 \subsection{Effect of Slicing}
-Features were most predictive when using either the largest 5 or 10\% or middle
+Features were slightly more predictive when using either the largest 5 or 10\% or middle
 25\% ("max-05", "max-10" and "mid-25" in Figure
@@
 closely follow the unfolded eigenvalues in predictive utility.
+
+\section{Discussion}
+
+In this study, eigenfeatures inspired by RMT and derived from the eigenvalues
+of the full, whole-brain voxelwise fMRI correlation matrix, were found to have
+predictive utility across a wide variety of phenomena. Compared to simple
+baseline reductions of the fMRI data, these eigenfeatures had more consistent
+predictive utility, and a higher maximum predictive potential (Figures
+\ref{fig:main-results}-\ref{fig:largest}).  In addition to evidence from
+previous studies
+\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
+this suggests RMT may be a useful analytic and theoretical tool for
+understanding functional connectivity.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{unfolded_duloxetine_nopain.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvaleus (polynomial degree 9) for \textit{Osteo} dataset \code{duloxetine v nopain} classification task.}
+\end{figure}
+
+When RMT-based eigenfeatures were most predictive
 

### 2022-11-10 paper/paper/SPIE/supplementary.tex
[WIP]

@@
 { \label{fig:fine-classifier}
-AUROC distributions across coarse feature groupings and predictable comparison tasks, by classifier.}
+AUROC distributions across fine feature groupings and predictable comparison tasks, by classifier.}
 \end{figure}
@@
 { \label{fig:fine-trim}
-Distributions of mAUROCs for unfolding-dependent RMT features, by degree.}
+Distributions of mAUROCs for unfolding-dependent RMT features, by trimming.}
 \end{figure}

### 2022-11-10 code/scripts/summarize_everything_updated.py

@@ -26,7 +26,13 @@ from seaborn import FacetGrid
 from statsmodels.distributions.empirical_distribution import ECDF
 from tqdm import tqdm
 
-from rmt.updated_features import FEATURE_OUTFILES as PATHS
+from rmt.updated_dataset import UpdatedProcessedDataset
+from rmt.enumerables import UpdatedDataset, PreprocLevel, TrimMethod
+from rmt.updated_features import (
+    FEATURE_OUTFILES as PATHS,
+    Unfolded,
+    Eigenvalues,
+)
 from rmt.visualize import UPDATED_PLOT_OUTDIR as PLOT_OUTDIR
 from rmt.visualize import best_rect
 
@@ -1905,29 +1911,37 @@ def make_kde_plots() -> None:
         savefig(fig, "best_rmt_params_by_subgroup_and_slicing.png")
 
     def plot_all_by_fine_feature_groups_best_params_best_slice() -> None:
-        rmt = df.loc[df.coarse_feature.isin(["rmt"])]
+        rmt = df.loc[df.coarse_feature.isin(["rmt", "eigs"])]
         rmt = rmt.loc[rmt.trim.isin(["middle", "largest"])]
         rmt = rmt.loc[rmt.deg.isin([3, 9])]
         rmt = rmt.loc[rmt.preproc.isin(["MotionCorrect", "MNIRegister"])]
+        rmt = rmt.loc[rmt.slice.isin(["max-05", "max-10", "mid-25"])]
         idx = rmt.feature.isin(["unfolded"])
         rmt = rmt[idx]
 
-        other = df.loc[df.coarse_feature.isin(["eigs"])]
-        data = pd.concat([rmt, other], axis=0)
+        # other = df.loc[df.coarse_feature.isin(["eigs"])]
+        # data = pd.concat([rmt, other], axis=0)
+        data = rmt
+
+        palette = {**FEATURE_GROUP_PALETTE}
+        palette.pop("tseries loc")
+        palette.pop("tseries scale")
 
         grid = sbn.displot(
             data=data,
             x="auroc",
             kind="kde",
             hue="fine_feature",
-            hue_order=list(FEATURE_GROUP_PALETTE.keys()),
-            palette=FEATURE_GROUP_PALETTE,
+            hue_order=list(palette.keys()),
+            palette=palette,
             fill=False,
             common_norm=False,
             col="subgroup",
[…]

### 2022-11-11 paper/paper/SPIE/article.tex
[WIP]

@@
 saved and used for entirely separate feature analyses. E.g. all features were
-extracted from a brain prepared with one of four levels or degrees of
-preprocessing, where level \(k\) includes preprocessings steps 1 to \(k\),
+extracted from an fMRI image prepared with one of four levels or degrees of
+preprocessing, where preprocessing level \(k\) includes preprocessings steps 1 to \(k\),
 inclusive, and starting at \(k = 1\).
@@
 distributions, and so have chosen primarily to present our findings visually,
-with kernel density estimates.
+with kernel density estimates. However, we also provide overall numerical
+summaries of the feature performances in Table \ref{tab:numerical}.
+
+\begin{table}[h!]
+\caption{Numerical summaries of feature mAUROCs across predictable comparisons, and all combinations of
+analytic choices, sorted by 95\% percentile (robust max) value. Bold values indicate column "best" values, when reasonable. }
+\label{tab:numerical}
+\small
+\centering
+
+\begin{tabular}{lrrrrrrr}
+\hline
+Feature &      mean &       min &        5\% &       50\% &       95\% &       max &       std \\
+\hline
+eigs + eigs\_smooth            &  \textbf{0.583} &  0.188 &  0.408 &  0.567 &  \textbf{0.788} &  0.906 &  0.111 \\
+eigs + savgol                  &  0.580 &  0.188 &  0.408 &  \textbf{0.563} &  0.785 &  0.913 &  0.110 \\
+unfolded                       &  0.568 &  0.150 &  0.400 &  0.555 &  0.775 &  0.931 &  0.108 \\
+unfolded + levelvar            &  0.569 &  0.150 &  0.400 &  0.555 &  0.775 &  0.925 &  0.108 \\
+unfolded + rigidity            &  0.568 &  0.150 &  0.401 &  0.555 &  0.775 &  0.925 &  0.108 \\
+unfolded + rigidity + levelvar &  0.568 &  0.150 &  0.401 &  0.555 &  0.775 &  0.931 &  0.108 \\
+eigs + rigidity + levelvar     &  0.564 &  0.202 &  0.401 &  0.548 &  0.763 &  \textbf{0.937} &  0.105 \\
+eigs + unfolded                &  0.564 &  0.202 &  0.401 &  0.549 &  0.763 &  0.919 &  0.105 \\
+eigs + unfolded + levelvar     &  0.564 &  0.202 &  0.400 &  0.548 &  0.763 &  0.933 &  0.105 \\
+eigs + unfolded + rigidity     &  0.564 &  0.202 &  0.401 &  0.548 &  0.763 &  0.920 &  0.105 \\
+eigs + levelvar                &  0.563 &  0.202 &  0.400 &  0.548 &  0.758 &  0.914 &  0.104 \\
+eigs                           &  0.563 &  0.202 &  0.401 &  0.548 &  0.756 &  0.875 &  0.104 \\
+eigs + rigidity                &  0.563 &  0.202 &  0.400 &  0.548 &  0.756 &  0.882 &  0.104 \\
+eigs\_smooth                   &  0.569 &  0.188 &  0.417 &  0.556 &  0.755 &  0.907 &  0.102 \\
+T-p05                          &  0.526 &  0.249 &  0.340 &  0.500 &  0.754 &  0.856 &  0.110 \\
+eigs\_savgol                   &  0.566 &  0.188 &  0.410 &  0.553 &  0.748 &  0.909 &  0.102 \\
+eigsminmax20                   &  0.556 &  0.245 &  0.411 &  0.544 &  0.738 &  0.906 &  0.098 \\
+eigsminmax10                   &  0.550 &  0.231 &  0.409 &  0.538 &  0.737 &  0.906 &  0.095 \\
+eigsminmax5                    &  0.549 &  0.213 &  0.400 &  0.539 &  0.729 &  0.888 &  0.096 \\
+T-mean                         &  0.557 &  0.208 &  \textbf{0.427} &  0.558 &  0.717 &  0.814 &  0.085 \\
+levelvar                       &  0.534 &  0.075 &  0.367 &  0.527 &  0.712 &  0.879 &  0.101 \\
+eigsmiddle40                   &  0.547 &  \textbf{0.324} &  0.419 &  0.533 &  0.697 &  0.785 &  0.084 \\
+rigidity + levelvar            &  0.529 &  0.154 &  0.377 &  0.523 &  0.696 &  0.888 &  0.094 \\
+rigidity                       &  0.529 &  0.154 &  0.378 &  0.523 &  0.694 &  0.888 &  0.094 \\
+T-rrng                         &  0.545 &  0.281 &  0.384 &  0.542 &  0.688 &  0.759 &  0.090 \\
+eigsmiddle20                   &  0.545 &  0.321 &  0.419 &  0.533 &  0.686 &  0.798 &  0.080 \\
+T-iqr                          &  0.494 &  0.175 &  0.312 &  0.490 &  0.682 &  0.759 &  0.099 \\
+eigsmiddle10                   &  0.541 &  0.323 &  0.411 &  0.530 &  0.681 &  0.755 &  0.081 \\
+T-std                          &  0.513 &  0.252 &  0.392 &  0.500 &  0.679 &  0.729 &  0.086 \\
+T-rng                          &  0.538 &  0.279 &  0.392 &  0.539 &  0.678 &  0.779 &  0.092 \\
+T-max                          &  0.540 &  0.275 &  0.385 &  0.546 &  0.678 &  0.754 &  0.095 \\
+T-med                          &  0.539 &  0.221 &  0.399 &  0.541 &  0.665 &  0.702 &  0.077 \\
+T-p95                          &  0.524 &  0.294 &  0.402 &  0.514 &  0.653 &  0.693 &  0.077 \\
+T-min                          &  0.504 &  0.321 &  0.426 &  0.500 &  0.582 &  0.698 &  0.046 \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 certain preprocessing level and for certain feature slicing. However, there are
-too many potential interactions to present, so we limit our presentation to the
-most significant ones.  As there are a large number of features, we also
-summarize patterns of results across two abstract groupings of features
+too many potential interactions to present, so we limit our presentation to
+main effects.  As there are a large number of features, we also summarize
+patterns of results across two abstract groupings of similar features
 ("coarse", and "fine") shown in Table \ref{tab:features}.
@@
 \subsubsection{Predictable Comparisons}
-As visible in Figure \ref{fig:main-results}, when coarsely summarizing features,
-eigenfeatures were more likely to be predictively useful than not, and except
-for in the \code{task\_attend v task\_nonattend} comparison, were also more
-predictively useful than the baseline features. We also note that although
-orders of magnitude more models were fit for eigenfeatures (see Section \ref{sec:multiverse}),
-the mAUROC distributions for the baseline features are either equally as wide,
-or, for the Parkinson's data, wider.
-
-Eigenfeatures most strongly and consistently demonstrated predictive utility
-in the \textit{Aging} dataset \code{older v younger} classification task, the
-\textit{Osteo} dataset \code{nopain v duloxetine} task, and in the \textit{Attention}
-\code{vigilant v nonvigilant} comparison.
+As visible in Figure \ref{fig:main-results}, when coarsely summarizing
+features, eigenfeatures were more likely to be predictively useful than not,
+and except for in the \code{task\_attend v task\_nonattend} comparison, were
+also more predictively useful than the baseline features.  Eigenfeatures most
+strongly and consistently demonstrated predictive utility in the \textit{Aging}
+dataset \code{older v younger} classification task, the \textit{Osteo} dataset
+\code{nopain v duloxetine} task, and in the \textit{Attention} \code{vigilant v
+nonvigilant} comparison.
 
@@
 
-Examining these RMT features more closely (Figure S\ref{S-fig:fig:fine-degree}),
+Examining these RMT features more closely (Figure S\ref{S-fig:fine-degree}),
 it is clear these features are more predictive mostly due to the unfolding
@@
 
-\subsection{Best Case Scenario}
-Viewing all the impactful analytic choices as tunable hyperparameters, and
-selecting the overall "best" hyperparameters of \(\text{trimming} \in
-\{\text{middle}, \text{largest}\}\), \(\text{degree} \in \{3, 9\}\),
-\(\text{preprocessing} \in \{\text{MotionCorrect}, \text{MNIRegister}\}\),
-\(\text{slicing} \in \{\text{"max-10"}, \text{"mid-25"}\} \) and
-restricting to only the "unfolded" RMT feature, results in the mAUROC
-distributions in Figure S\ref{S-fig:best-params}. Using these analytic choices,
-the \code{depress v control} comparison in the \textit{Depression} dataset
-becomes weakly predictable, and the smoothed eigenvalue features tend to
-closely follow the unfolded eigenvalues in predictive utility.
-
 \section{Discussion}
@@
 of the full, whole-brain voxelwise fMRI correlation matrix, were found to have
-predictive utility across a wide variety of phenomena. Compared to simple
-baseline reductions of the fMRI data, these eigenfeatures had more consistent
-predictive utility, and a higher maximum predictive potential (Figures
-\ref{fig:main-results}-\ref{fig:largest}).  In addition to evidence from
-previous studies
+predictive utility across a wide variety of phenomena and analytic choices.
+Compared to simple baseline reductions of the fMRI data, these eigenfeatures
+had more consistent predictive utility, and a higher maximum predictive
+potential (Figures \ref{fig:main-results}-\ref{fig:largest}). In addition to
+evidence from previous studies
 \cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
@@
 \begin{tabular}{c}
-\includegraphics[width=6.5in]{unfolded_duloxetine_nopain.png}
+\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
 \end{tabular}
@@
 \caption
-{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvaleus (polynomial degree 9) for \textit{Osteo} dataset \code{duloxetine v nopain} classification task.}
+{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvalues (polynomial
+degree 9) for \textit{Osteo} dataset \footnotesize\texttt{duloxetine v nopain}
+classification task, brain extraction and slice time correction only. Median
+observed mAUROC = 0.643 (range = 0.346 - 0.925)}
 \end{figure}
 
-When RMT-based eigenfeatures were most predictive
+It is likely that the unfolded eigenvalues have predictive utility because of
+their smoothing and re-scaling effect. It is clear from Figure \ref{fig:unfolded},
+which depicts the raw eigenvalues and unfolded eigenvalues with different trimming
+procedures, that the raw eigenvalues have an extremely exponential distribution
+(Figure \ref{fig:unfolded} already uses a log-scale for plotting) due to the largest
+eigenvalues. The trimmed feature distributions are clearly far less pathologically distributed.
+Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier} that
+a simple combination of the raw eigenvalues plus smoothed versions of them are the most
+generally-predictive features examined here.
+
+It was also sometimes the case that adding RMT-specific features (e.g. the spectral
+rigidity, or level variance) to either the unfolded eigenvalues, or raw
+eigenvalues, would allow for a higher maximum mAUROC (Table
+\ref{tab:numerical}, Figure S\ref{S-fig:fine-largest}). One case where these
+features yielded the largest predictions was in the \textit{Aging} dataset.
+This is depicted in Figure \ref{fig:observables}, where it is clear that, after
+trimming, at almost any \(L\) value, subgroups are strongly and clearly
+visually-separable.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:observables} Rigidity and level variance features (unfolding
+polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
+older} classification task, brain extraction only. Median observed mAUROC =
+0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
+all subplots, as described in the \ref{sec:methods}.}
+\end{figure}
+
+However, the range of mAUROCs observed in all cases in this study suggests
+caution. All eigenfeatures and baselines features were sensitive to analytic
+choices such as the amount of preprocessing, feature slicing (including trimming) and
+feature smoothing / unfolding, and this sensitivity was such that all features
+resulted in mAUROCs on either side of 0.5. Most of these analytic choices are benign enough that
+it is difficult to rigorously justify one choice over another, meaning RMT-based
+analyses may be susceptible to post-hoc justification \cite{kerrHARKingHypothesizingResults1998}.
+
+It is also difficult to quantify the practical significance of the observed mAUROC values.
+With small datasets of rs- or task-fMRI data, binary classification using whole-brain
+features is challenging. For subject level classification, state of the art deep learning methods
+improve upon guessing by 17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
+\cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
+\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive compulsive disorder
+\cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual feature engineering with more separable
+conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
+\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated custom feature
+extraction methods can achieve near perfect accuracies at classifying task vs. rest
+\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional connectivity data
+and with classical machine learning algorithms (such as SVM) we in general only expect large
+prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
+strongly separated (e.g. Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}).
+
+In this study, the median accuracy improvement upon guessing was 0 for all features, although
+eigenfeatures could improve upon guessing by 16.2\% - 33.7\% (\ref{S-tab:acc-numerical}).
+
+It is somewhat surprising that eigenfeatures are, on net, predictively-useful
+at all. The reduction of the functional connectivity matrix to \(T - 1\)
+eigenvalues is aggressive, and the subsequent small-degree polynomial fit used
+in the unfolding procedure further reduces information in the raw data.
+Still, the subsequent features more often than not allow non-trivial 5-fold
+cross-validated mean prediction AUROCs or accuracies.
+
+We believe that future studies using RMT to inform their examinations be more
+careful. In the absence of a cross-validated predictive framework that takes
+into account the strong impact of researcher analytic choices, any apparent
+explanatory utility of RMT should be treated with skepticism. RMT deals with
+highly-ideal \textit{iid} matrices, but almost no real-world systems involve
+\textit{iid} elements. The basic RMT insight that eigenvalues can be used to
+describe matrix-representable systems is certainly vindicated here, but the use
+of mathematically non-trivial RMT statistics (like the spectral rigidity or
+level number variance) should not be used if simpler statistical descriptions
+of the eigenvalues are equally informative.
 
@@
 
-\section{Parts of Manuscript}
-
-This section describes the normal structure of a manuscript and how each part
-should be handled. The appropriate vertical spacing between various parts of
-this document is achieved in LaTeX through the proper use of defined
-constructs, such as \verb|\section{}|.
-
-\subsection{Title and Author Information}
-\label{sect:title}
-The article title appears left justified at the top of the first page. The
-title font is 16 pt., bold. The rules for capitalizing the title are the same
-as for sentences; only the first word, proper nouns, and acronyms should be
-capitalized. Do not begin titles with articles (for example, a, an, the) or
-prepositions (for example, on, by, etc.). The word ``novel'' should not appear
-in the title, as publication will imply novelty. Avoid the use of acronyms in
-the title, unless they are widely understood. Appendix A contains more about
-acronyms.
-
-The list of authors immediately follows the title, 18 points below. The font is
-12 pt., bold and the author names are left justified. The author affiliations
-and addresses follow the names, in 10-pt., normal font and left justified. For
-multiple affiliations, each affiliation should appear on a separate line.
-Superscript letters (a, b, c, etc.) should be used to associate multiple
-authors with their respective affiliations. The corresponding author should be
-identified with an asterisk, and that person's email address should be provided
-below the keywords.
-
-\subsection{Abstract}
-The abstract should be a summary of the paper and not an introduction. Because
-the abstract may be used in abstracting journals, it should be self-contained
-(i.e., no numerical references) and substantive in nature, presenting concisely
-the objectives, methodology used, results obtained, and their significance.
-Please note that the following journals require the use of structured abstracts
-in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of
-Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured
-abstracts are encouraged for the \textit{Journal of Micro/Nanolithography,
-MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on
-the website of the journal.
-
-\subsection{Subject terms/Keywords}
-Keywords are required. Please provide 3-6 keywords related to your paper.
-
-\subsection{Body of Paper}
-The body of the paper consists of numbered sections that present the main
-findings. These sections should be organized to best present the material.
-
-To provide transition elements in your paper, it is important to refer back (or
-forward) to specific sections. Such references are made by indicating the
-section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1
-contained a description...'' If the word Section, Reference, Equation, or
-Figure starts a sentence, it is spelled out. When occurring in the middle of a
-sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
-
-At the first occurrence of an acronym, spell it out followed by the acronym in
-parentheses, for example, charge-coupled diode (CCD).
-
-\subsection{Footnotes}
-Textual footnotes should be used rarely to present important documentary or
-explanatory material whose inclusion in the text would be
-distracting.\footnote{Example of a footnote.} Due to problems with HTML
-display, use of footnotes should generally be avoided. If absolutely necessary,
-the footnote mark must come at the end of a sentence. To insert a footnote, use
-the {\verb|\footnote{}|} command.
-
-\subsection{Appendices}
-Brief appendices may be included when necessary, such as derivations of
-equations, proofs of theorems, and details of algorithms. Equations and figures
-appearing in appendices should continue sequential numbering from earlier in
-the paper.
+% \section{Parts of Manuscript}
+
+% This section describes the normal structure of a manuscript and how each part
+% should be handled. The appropriate vertical spacing between various parts of
+% this document is achieved in LaTeX through the proper use of defined
+% constructs, such as \verb|\section{}|.
+
+% \subsection{Title and Author Information}
+% \label{sect:title}
+% The article title appears left justified at the top of the first page. The
+% title font is 16 pt., bold. The rules for capitalizing the title are the same
+% as for sentences; only the first word, proper nouns, and acronyms should be
+% capitalized. Do not begin titles with articles (for example, a, an, the) or
+% prepositions (for example, on, by, etc.). The word ``novel'' should not appear
+% in the title, as publication will imply novelty. Avoid the use of acronyms in
+% the title, unless they are widely understood. Appendix A contains more about
+% acronyms.
+
+% The list of authors immediately follows the title, 18 points below. The font is
+% 12 pt., bold and the author names are left justified. The author affiliations
+% and addresses follow the names, in 10-pt., normal font and left justified. For
+% multiple affiliations, each affiliation should appear on a separate line.
+% Superscript letters (a, b, c, etc.) should be used to associate multiple
+% authors with their respective affiliations. The corresponding author should be
+% identified with an asterisk, and that person's email address should be provided
+% below the keywords.
+
+% \subsection{Abstract}
+% The abstract should be a summary of the paper and not an introduction. Because
+% the abstract may be used in abstracting journals, it should be self-contained
+% (i.e., no numerical references) and substantive in nature, presenting concisely
+% the objectives, methodology used, results obtained, and their significance.
+% Please note that the following journals require the use of structured abstracts
+% in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of
+% Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured
+% abstracts are encouraged for the \textit{Journal of Micro/Nanolithography,
+% MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on
+% the website of the journal.
+
+% \subsection{Subject terms/Keywords}
+% Keywords are required. Please provide 3-6 keywords related to your paper.
+
+% \subsection{Body of Paper}
+% The body of the paper consists of numbered sections that present the main
+% findings. These sections should be organized to best present the material.
+
+% To provide transition elements in your paper, it is important to refer back (or
+% forward) to specific sections. Such references are made by indicating the
+% section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1
+% contained a description...'' If the word Section, Reference, Equation, or
+% Figure starts a sentence, it is spelled out. When occurring in the middle of a
+% sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
+
+% At the first occurrence of an acronym, spell it out followed by the acronym in
+% parentheses, for example, charge-coupled diode (CCD).
+
+% \subsection{Footnotes}
+% Textual footnotes should be used rarely to present important documentary or
+% explanatory material whose inclusion in the text would be
+% distracting.\footnote{Example of a footnote.} Due to problems with HTML
+% display, use of footnotes should generally be avoided. If absolutely necessary,
+% the footnote mark must come at the end of a sentence. To insert a footnote, use
+% the {\verb|\footnote{}|} command.
+
+% \subsection{Appendices}
+% Brief appendices may be included when necessary, such as derivations of
+% equations, proofs of theorems, and details of algorithms. Equations and figures
+% appearing in appendices should continue sequential numbering from earlier in
+% the paper.
 
@@
 
-\subsection{References}
-The References section lists books, articles, and reports that are cited in the
-paper. This section does not have a section number. The references are numbered
-in the order in which they are cited. Examples of the format to be followed are
-given at the end of this document.
-
-The reference list at the end of this document is created using BibTeX, which
-looks through the file {\ttfamily report.bib} for the entries cited in the
-LaTeX source file.  The format of the reference list is determined by the
-bibliography style file {\ttfamily spiejour.bst}, as specified in the \\
-\verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may
-be directly formatted in the LaTeX source file.
-
-For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list
-of authors (initials plus last name), book title (in italics), page or chapter
-numbers, publisher, city, and year of publication.  Journal-article references
-\citenum{Metropolis53,Harris06} include the author list, title of the article
-(in quotes), journal name (in italics, properly abbreviated), volume number (in
-bold), inclusive page numbers or citation identifier, and year.  A reference to
-a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes
-the author list, title of the article (in quotes), conference name (in
-italics), editors (if appropriate), volume title (in italics), volume number if
-applicable (in bold), inclusive page numbers, publisher, city, and year.
-References to an article in the SPIE Proceedings may include the conference
-name, as shown in Ref.~\citenum{Hanson93c}.
-
-The references are numbered in the order of their first citation. Citations to
-the references are made using superscripts, as demonstrated in the preceding
-paragraph. One may also directly refer to a reference within the text, for
-example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more
-references should be separated by a comma with no space between them. Multiple
-sequential references should be displayed with a dash between the first and
-last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
+% \subsection{References}
+% The References section lists books, articles, and reports that are cited in the
+% paper. This section does not have a section number. The references are numbered
+% in the order in which they are cited. Examples of the format to be followed are
+% given at the end of this document.
+
+% The reference list at the end of this document is created using BibTeX, which
+% looks through the file {\ttfamily report.bib} for the entries cited in the
+% LaTeX source file.  The format of the reference list is determined by the
+% bibliography style file {\ttfamily spiejour.bst}, as specified in the \\
+% \verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may
+% be directly formatted in the LaTeX source file.
+
+% For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list
+% of authors (initials plus last name), book title (in italics), page or chapter
+% numbers, publisher, city, and year of publication.  Journal-article references
+% \citenum{Metropolis53,Harris06} include the author list, title of the article
+% (in quotes), journal name (in italics, properly abbreviated), volume number (in
+% bold), inclusive page numbers or citation identifier, and year.  A reference to
+% a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes
+% the author list, title of the article (in quotes), conference name (in
+% italics), editors (if appropriate), volume title (in italics), volume number if
+% applicable (in bold), inclusive page numbers, publisher, city, and year.
+% References to an article in the SPIE Proceedings may include the conference
+% name, as shown in Ref.~\citenum{Hanson93c}.
+
+% The references are numbered in the order of their first citation. Citations to
+% the references are made using superscripts, as demonstrated in the preceding
+% paragraph. One may also directly refer to a reference within the text, for
+% example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more
+% references should be separated by a comma with no space between them. Multiple
+% sequential references should be displayed with a dash between the first and
+% last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
 
@@
 
-\section{Section Formatting}
-\label{sect:sections}
-In LaTeX, a new section is created with the \verb|\section{}| command, which
-automatically numbers the sections. Sections will be numbered sequentially,
-starting with the first section after the abstract, except for the
-acknowledgments and references. (Note that numbering of section headings is not
-required, but the numbering must be consistent if used.) All section headings
-should be left justified.
-
-Main section headings are in 12-pt. bold font, left-justified and in title
-case, where important words are capitalized.
-
-Paragraphs that immediately follow a section heading are leading paragraphs and
-should not be indented, according to standard publishing style. The same goes
-for leading paragraphs of subsections and sub-subsections. Subsequent
-paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no
-additional space between paragraphs. In LaTeX, paragraphs are separated by
-blank lines in the source file. Indentation of the first line of a paragraph
-may be avoided by starting it with \verb|\noindent|.
-
-\subsection{Subsection Headings}
-All important words in a subsection (level 1) header are capitalized.
-Subsection numbers consist of the section number, followed by a period, and the
-subsection number within that section, without a period at the end. The heading
-is left justified and its font is 12 pt. italic.
-
-\subsubsection{Sub-subsection headings}
-The first word of a sub-subsection is capitalized. The rest of the text is not
-capitalized, except for proper names and acronyms (the latter should only be
-used if well known). The heading is left justified and its font is 11 pt.
-italic.
-
-\section{Figures and Tables}
-
-\subsection{Figures}
-
-Figures are numbered in the order in which they are called out in the text.
-They should appear in the document in numerical order and as close as possible
-to their first reference in the text. It may be necessary to move figures or
-tables around to enhance readability. LaTeX will attempt to place figures at
-the top or bottom of a page in which they are first referenced.
-
-Figures, along with their captions, should be separated from the main text by
-0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or
-graph. Figure captions start with the abbreviation ``Fig'' in front of the
-figure number, followed by a period, and the text in 10-pt. font. See
-Fig.~\ref{fig:example} for an example.
+% \section{Section Formatting}
+% \label{sect:sections}
+% In LaTeX, a new section is created with the \verb|\section{}| command, which
+% automatically numbers the sections. Sections will be numbered sequentially,
+% starting with the first section after the abstract, except for the
+% acknowledgments and references. (Note that numbering of section headings is not
+% required, but the numbering must be consistent if used.) All section headings
+% should be left justified.
+
+% Main section headings are in 12-pt. bold font, left-justified and in title
+% case, where important words are capitalized.
+
+% Paragraphs that immediately follow a section heading are leading paragraphs and
+% should not be indented, according to standard publishing style. The same goes
+% for leading paragraphs of subsections and sub-subsections. Subsequent
+% paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no
+% additional space between paragraphs. In LaTeX, paragraphs are separated by
+% blank lines in the source file. Indentation of the first line of a paragraph
+% may be avoided by starting it with \verb|\noindent|.
+
+% \subsection{Subsection Headings}
+% All important words in a subsection (level 1) header are capitalized.
+% Subsection numbers consist of the section number, followed by a period, and the
+% subsection number within that section, without a period at the end. The heading
+% is left justified and its font is 12 pt. italic.
+
+% \subsubsection{Sub-subsection headings}
+% The first word of a sub-subsection is capitalized. The rest of the text is not
+% capitalized, except for proper names and acronyms (the latter should only be
+% used if well known). The heading is left justified and its font is 11 pt.
+% italic.
+
+% \section{Figures and Tables}
+
+% \subsection{Figures}
+
+% Figures are numbered in the order in which they are called out in the text.
+% They should appear in the document in numerical order and as close as possible
+% to their first reference in the text. It may be necessary to move figures or
+% tables around to enhance readability. LaTeX will attempt to place figures at
+% the top or bottom of a page in which they are first referenced.
+
+% Figures, along with their captions, should be separated from the main text by
+% 0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or
+% graph. Figure captions start with the abbreviation ``Fig'' in front of the
+% figure number, followed by a period, and the text in 10-pt. font. See
+% Fig.~\ref{fig:example} for an example.
 

### 2022-11-11 paper/paper/SPIE/supplementary.tex
[WIP]

@@
 \usepackage{float}
+\restylefloat{table}  % https://stackoverflow.com/questions/1673942/latex-table-positioning
 \usepackage{ifthen}
@@
   top=0in,
-  tmarg=0in,
+  tmargin=0in,
   bottom=0in,
@@
 
-\begin{spacing}{1}   % use single spacing for supplementary
+% \begin{spacing}{1}   % use single spacing for supplementary
 
-\begin{figure}
+\begin{figure}[H]
 \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[H]
 \begin{center}
@@
 
-\bibliography{report}   % bibliography data in report.bib
-\bibliographystyle{spiejour}   % makes bibtex use spiejour.bst
+\begin{table}[H]
+\caption{Numerical summaries of feature mean accuracy difference from guess across predictable comparisons, and all combinations of
+analytic choices, sorted by 95\% percentile (robust max) value.}
+\label{tab:acc-numerical}
+\small
+\centering
 
-\end{spacing}
+\begin{tabular}{lrrrrrrr}
+\hline
+Feature &      mean &       min &        5\% &       50\% &       95\% &       max &       std \\
+\hline
+unfolded                       &  0.021 & -0.437 & -0.123 &  0.000 &  0.196 &  0.312 &  0.091 \\
+unfolded + levelvar            &  0.021 & -0.437 & -0.122 &  0.000 &  0.195 &  0.312 &  0.091 \\
+unfolded + rigidity            &  0.021 & -0.437 & -0.122 &  0.000 &  0.194 &  0.312 &  0.091 \\
+unfolded + rigidity + levelvar &  0.021 & -0.437 & -0.123 &  0.000 &  0.193 &  0.312 &  0.090 \\
+eigs + eigs\_smooth             &  0.025 & -0.295 & -0.125 &  0.000 &  0.193 &  0.337 &  0.091 \\
+eigs + savgol                  &  0.022 & -0.295 & -0.125 &  0.000 &  0.190 &  0.337 &  0.091 \\
+eigs + unfolded                &  0.014 & -0.311 & -0.130 &  0.000 &  0.173 &  0.362 &  0.087 \\
+eigs + rigidity + levelvar     &  0.013 & -0.311 & -0.130 &  0.000 &  0.173 &  0.362 &  0.086 \\
+eigs + unfolded + levelvar     &  0.013 & -0.311 & -0.130 &  0.000 &  0.173 &  0.362 &  0.087 \\
+eigs + unfolded + rigidity     &  0.013 & -0.311 & -0.130 &  0.000 &  0.173 &  0.362 &  0.086 \\
+eigs\_savgol                    &  0.016 & -0.313 & -0.126 &  0.000 &  0.172 &  0.321 &  0.085 \\
+eigs\_smooth                    &  0.018 & -0.263 & -0.123 &  0.000 &  0.172 &  0.323 &  0.084 \\
+eigs + levelvar                &  0.013 & -0.311 & -0.131 &  0.000 &  0.171 &  0.321 &  0.086 \\
+eigs                           &  0.013 & -0.311 & -0.130 &  0.000 &  0.168 &  0.312 &  0.086 \\
+eigs + rigidity                &  0.013 & -0.311 & -0.131 &  0.000 &  0.162 &  0.312 &  0.085 \\
+eigsminmax20                   &  0.007 & -0.348 & -0.123 &  0.000 &  0.159 &  0.309 &  0.083 \\
+eigsminmax5                    &  0.002 & -0.367 & -0.120 &  0.000 &  0.159 &  0.284 &  0.084 \\
+T-p05                          & -0.004 & -0.298 & -0.193 &  0.000 &  0.155 &  0.205 &  0.091 \\
+eigsmiddle40                   &  0.008 & -0.225 & -0.114 &  0.000 &  0.151 &  0.234 &  0.074 \\
+eigsmiddle20                   &  0.006 & -0.210 & -0.120 &  0.000 &  0.147 &  0.215 &  0.074 \\
+eigsminmax10                   &  0.003 & -0.367 & -0.116 &  0.000 &  0.145 &  0.287 &  0.081 \\
+levelvar                       & -0.004 & -0.363 & -0.135 & -0.010 &  0.145 &  0.312 &  0.081 \\
+T-rrng                         &  0.002 & -0.249 & -0.140 &  0.000 &  0.144 &  0.184 &  0.079 \\
+eigsmiddle10                   &  0.004 & -0.224 & -0.127 &  0.000 &  0.136 &  0.206 &  0.074 \\
+rigidity + levelvar            & -0.006 & -0.335 & -0.127 & -0.009 &  0.130 &  0.312 &  0.076 \\
+rigidity                       & -0.006 & -0.332 & -0.127 & -0.009 &  0.130 &  0.309 &  0.075 \\
+T-mean                         &  0.011 & -0.202 & -0.080 &  0.009 &  0.119 &  0.191 &  0.058 \\
+T-med                          & -0.006 & -0.259 & -0.102 & -0.002 &  0.114 &  0.194 &  0.065 \\
+T-max                          & -0.003 & -0.192 & -0.106 & -0.002 &  0.112 &  0.180 &  0.066 \\
+T-rng                          & -0.004 & -0.173 & -0.104 & -0.002 &  0.110 &  0.169 &  0.063 \\
+T-iqr                          & -0.030 & -0.313 & -0.152 & -0.025 &  0.103 &  0.202 &  0.077 \\
+T-std                          & -0.025 & -0.202 & -0.126 & -0.026 &  0.094 &  0.147 &  0.065 \\
+T-p95                          & -0.011 & -0.228 & -0.115 & -0.002 &  0.079 &  0.180 &  0.061 \\
+T-min                          & -0.030 & -0.298 & -0.298 & -0.001 &  0.047 &  0.127 &  0.082 \\
+\hline
+\end{tabular}
+\end{table}
+
+% \bibliography{report}   % bibliography data in report.bib
+% \bibliographystyle{spiejour}   % makes bibtex use spiejour.bst
+
+% \end{spacing}
 \end{document}

### 2022-11-11 code/scripts/summarize_everything_updated.py

@@ -10,10 +10,13 @@ from typing import Literal
 from warnings import simplefilter
 
 import matplotlib.pyplot as plt
+from numpy import ndarray
 import numpy as np
 import pandas as pd
 import seaborn as sbn
+from numba import njit
 from shutil import copyfile
+from warnings import simplefilter
 from joblib import Memory
 from matplotlib.axes import Axes
 from matplotlib.figure import Figure
@@ -32,6 +35,8 @@ from rmt.updated_features import (
     FEATURE_OUTFILES as PATHS,
     Unfolded,
     Eigenvalues,
+    Rigidities,
+    Levelvars,
 )
 from rmt.visualize import UPDATED_PLOT_OUTDIR as PLOT_OUTDIR
 from rmt.visualize import best_rect
@@ -56,6 +61,7 @@ DROPS = [
     "classifier_GradientBoostingClassifier",
 ]
 BLUE = "#004cc7"
+LBLUE = "#8dacfb"
 ORNG = "#f68a0e"
 GREEN = "#01f91e"
 GREY = "#5c5c5c"
@@ -2660,7 +2666,7 @@ def plot_unfolded_duloxetine() -> None:
             vals[vals <= 0] = np.nan
             ax.plot(
                 vals,
-                color=BLUE,
+                color=BLCK,
                 lw=0.5,
                 label="duloxetine" if k == 0 else None,
                 alpha=0.5,
@@ -2669,7 +2675,7 @@ def plot_unfolded_duloxetine() -> None:
             vals = nopain.iloc[k, :]
             vals[vals <= 0] = np.nan
             ax.plot(
-                vals, color=ORNG, lw=0.5, label="nopain" if k == 0 else None, alpha=0.5
+                vals, color=BLUE, lw=0.5, label="nopain" if k == 0 else None, alpha=0.5
             )
 
     ax = axes[0]
@@ -2682,7 +2688,7 @@ def plot_unfolded_duloxetine() -> None:
         vals[vals <= 1e-3] = np.nan
         ax.plot(
             vals,
-            color=BLUE,
+            color=BLCK,
             lw=0.5,
             label="duloxetine" if k == 0 else None,
             alpha=0.5,
@@ -2690,7 +2696,7 @@ def plot_unfolded_duloxetine() -> None:
     for k in range(len(nopain)):
         vals = nopain.iloc[k, :]
         vals[vals <= 1e-3] = np.nan
[…]

### 2022-11-12 paper/paper/SPIE/article.tex
[WIP]

@@
 In the human brain, each neuron, collection of neurons, or region of interest
-(ROI) is a potentially-interacting component in a highly complex system. RMT
+(ROI) is a potentially-interacting component in a complex system. RMT
 may have potential in describing the totality of these interactions, provided
@@
 \textit{level number variance} (``rigidity'' and ``level variance'' for short;
-details in Section \ref{sect:methods}).
+details in Section \ref{sec:methods}).
 
@@
 Typical reductions include the use of a ``seed'' voxel or collection of voxels
-(i.e. ROI): the mean signal over that ROI is correlated with all other N ROIs
-of interest to generate a reduced functional connectivity matrix (or image)
-with one correlation value at each non-seed ROI. This sort of reduction reduces
-the functional connectivity to N values. This approach strongly biases the
-analysis to the seed ROI.
-
-Likewise, one can work with ROI mean signals, and work with the \(N \times N\)
-matrix of these ROI correlations. This analysis [maybe cite Poldrack book for
-this?] is, \textit{a priori}, sensitive to the choice of ROIs, and, in the case
-of anatomical atlases / parcellations, may also lack theoretical justification.
-
-A major disadvantage of both the above approaches, besides requiring
-parcellation into ROIs, is that most such parcellation methods are atlas-based,
-and thus require image registration, and, general, an entire preprocessing
-pipeline to ensure that registration is successful.
+(see e.g. \citenum{joelRelationshipSeedbasedICAbased2011,
+smithCharacterizingIndividualDifferences2014}): the mean signal over that ROI
+is correlated with all other N ROIs of interest to generate a reduced
+functional connectivity matrix (or image) with one correlation value at each
+non-seed ROI. This sort of reduction reduces the functional connectivity to N
+values, but necessarily biases the analysis to the seed ROI.
+
+Likewise, one can work with ROI mean signal reductions, or independent
+component analysis reductions\cite{joelRelationshipSeedbasedICAbased2011,
+smithCharacterizingIndividualDifferences2014}, and work with the \(N \times N\)
+matrix of the correlations of these reductions. This analysis is, \textit{a
+priori}, sensitive to the choice of ROIs, and, in the case of anatomical
+atlases / parcellations for ROIs, may also lack theoretical justification.
 
@@
 correlation (or covariance) matrix with \(t - 1\) non-zero, positive
-real-valued eigenvalues \(\Lambda\). These eigenvalues can be computed highly
+real-valued eigenvalues. These eigenvalues can be computed highly
 efficiently via transposition of the voxelwise correlation or covariance
@@
 
-However, these previous studies took largely descriptive / explanatory
-approaches, noting only differences in RMT across subgroups and/or conditions,
-in many cases, with no formal statistical model nor formal testing of the
-significance and robustness of these differences. No code was provided to allow
-the reproduction of the results, and the extraction of the spectral rigidity
-and level variance is non-trivial, and computationally demanding. Computation
-of these spectral observables requires an \textit{unfolding}
+Previous studies applying RMT to functional connectivity
+data\cite{wangRandomMatrixTheory2016,
+wangSpectralPropertiesTemporal2015a,matharooSpontaneousBackpainAlters2020,
+guRandomMatrixTheory2020} took largely descriptive / explanatory approaches,
+noting only differences in RMT across subgroups and/or conditions. For example
+\citenum{wangRandomMatrixTheory2016,guRandomMatrixTheory2020} use a
+Kolmogorov-Smirnov test, and \citenum{wangSpectralPropertiesTemporal2015a} uses
+a t-test to provide evidence of differences in various RMT metrics. However,
+statistically significant differences in a metric do not imply practical
+significance or predictive utility \cite{loWhySignificantVariables2015}.
+
+In addition, the previous papers do not provide publicly-accessible code to
+reproduce results. The extraction of the spectral rigidity and level variance
+is computationally-demanding and non-trivial, requiring an \textit{unfolding}
 procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004},
 which is, unfortunately, a poorly-documented fitting procedure that requires
-making a number of difficult-to-justify decisions, and these decisions can
+making a number of arguably subjective decisions. These fitting decisions can
 dramatically impact RMT conclusions
-\cite{abul-magdUnfoldingSpectrumChaotic2014,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012,moralesImprovedUnfoldingDetrending2011}.
+\cite{abul-magdUnfoldingSpectrumChaotic2014,
+abueleninSpectralUnfoldingChaotic2018, fossionRandommatrixSpectraTime2013,
+abueleninEffectUnfoldingSpectral2012, moralesImprovedUnfoldingDetrending2011}.
 On top of the flexibility in the implementation and application of RMT to
@@
 complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
-
-In addition these issues stemming from ``research degrees of freedom'' (cite
-Gelman), there is also a basic baseline comparison issue: RMT is mathematically
-non-trivial, the extraction of RMT spectral statistics is computationally
-demanding, and eigenvalues, while easy to understand mathematically, are
-difficult to interpret relative to simpler potential reductions of fMRI data.
+These "researcher degrees of
+freedom"\cite{simmonsFalsePositivePsychologyUndisclosed2011b} in combination
+with the absence of clear code, and exclusively descriptive/explanatory
+analytic approaches means that the practical utility of RMT for fMRI is largely
+uninvestigated.
 
 What is missing is a rigorous, systematic, reproducible investigation of the
-*predictice* value of RMT metrics across a wide variety of data, RMT analytic
-choices, choice of eigenfeature, and choice of preprocessing, and for these RMT
-eigenfeatures to be compared to simpler alternative baselines. The current
-study aims to do just this.
+predictive value of RMT metrics across a wide variety of data, analytic
+choices, RMT features, and preprocessing decisions, with RMT features being
+compared to simpler alternative baseline predictors. The current study aims to
+remedy this.
 
@@
 neither task nor event timings were needed to determine class membership. This
-yielded seven datasets comprising XX total subjects, and over 5000 total fMRI
-volumes.
+yielded seven datasets.
 
@@
 controls\cite{goldcarrieelizabethExploringRestingState2018}. For our analysis,
-we make the binary classification task be to classify \code{monolingual v
+we make the binary classification task \code{monolingual v
 bilingual}.
@@
 The \textit{Depression} dataset\cite{bezmaternykhddRestingStateClosed2021,
-bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans frmo
-non-depressed controls and mildy- or moderately-depressed subjects. The our analysis,
+bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans from
+non-depressed controls and mildy- or moderately-depressed subjects. In our analysis,
 the binary classification task is \code{depress v control}.
@@
 The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
-schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans were
+schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans
 available for all subjects, and \code{task v rest} is chosen as the binary
@@
 presented with various related abstract images, and then later tested for their
-memory of certain aspects of the learned images. Thus the classification task
+memory of certain aspects of the learned images. Thus, the classification task
 might also be considered learning vs. rest.
@@
 The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
-includes whiole-brain rs-fMRI scans of healthy, pain-free controls
+includes whole-brain rs-fMRI scans of healthy, pain-free controls
 (\code{nopain} condition), and individuals with knee osteoarthritis.
-Osteoarthritic patients were treated for 2 weeks with either placebo
+Osteoarthritic patients were treated for two weeks with either placebo
 (\code{pain} condition) or duloxetine (\code{duloxetine} condition).
@@
 healthy controls performed a number of repetitions of the Attention Network
-Task (ANT; \cite{fanActivationAttentionalNetworks2005}) during scans.
+Task (ANT\cite{fanActivationAttentionalNetworks2005}) during scans.
 The classification task for this dataset is Parkinson's versus controls, i.e.
@@
 wakefulness and attention over the past weeks. Finally, the
-\code{task\_attend v tas\_nonattent} classification task was constructed
+\code{task\_attend v tas\_nonattend} classification task was constructed
 using the scores of The Conjunctive Continuous Performance Task
@@
 for other typical preprocessing steps (e.g. field intensity maps, physiological
-measurements), but also because we saved all pipelined intermediates to later
+measurements) was missing, but also because all pipelined intermediates were saved to later
 compare the effect of increasing degrees of pre-processing (see Section
@@
 \cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as a timeseries, the spectral rigidty for a value \(L\) can be is thus the
+as a timeseries, the spectral rigidity for a value \(L\) is the
 \textit{average nonlinearity} of all intervals of length \(L\) over the series.
 
-The level number variance, \(\Sigma^2(L)\), or level variance, for short, is
+The level number variance, \(\Sigma^2(L)\), or \textit{level variance}, for short, is
 closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
@@
 \cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as an \textit{irregular} timeseries, the level number variance is the
-variation of the number of samples of all intervals of length \(L\) over the
+as an irregular timeseries, the level number variance is the
+variation of the number of samples in all intervals of length \(L\) over the
 series.
@@
 eigenvalues, and involves fitting a smooth polynomial or exponential to these
-values, extreme values are often trimmed (or omitted from fitting)
+values, extreme values are often trimmed (omitted from fitting)
 \cite{moralesImprovedUnfoldingDetrending2011,
 abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
-highly subjective, and not usually well-documented. In addition, with a large
-number of different source matrices, it is too destructive to use the same hard
-criterion (e.g. largest 10\%, largest 3) for all unfoldings.
+highly subjective, and not usually well-documented, and unfortunately, with a
+large number of different source matrices, it is too destructive to naively use
+the same hard criterion (e.g. largest 10\%, largest 3) for all unfoldings.
 
 We develop and test four trimming variants: no trimming, precision-based,
-largest, and middle, with far less subjective criteria for determining trimming
-thresholds. However, we relegate the details of these trimming procedures to
-the Appendix \ref{sec:trimming}, or refer readers to the
+largest, and middle, with less subjective criteria for determining trimming
+thresholds. The details of these trimming procedures are straightforward, and
+can be found in Appendix \ref{sec:trimming}, or the
 \href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
@@
 
-When computing RMT eigenfeatures, the polynomial fitting during unfolding
-has a smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
-operations. However, it is possible other, simpler
-transformations of the eigenvalues (e.g. uniform or Savitsky-Golay smoothing)
-might have equal or greater predictive utility. We thus also test smoothed
-variants of the eigenvalues as predictive features: we smooth the sorted
-eigenvalues with a uniform (moving average) filter and Savitsky-Golay, using
-window sizes of 3, 5, 7, and 9 to yield 8 total ``smoothed eigenvalue'' features.
+When computing RMT eigenfeatures, the polynomial fitting during unfolding has a
+smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
+operations. However, it is possible other, simpler transformations of the
+eigenvalues (e.g. uniform or Savitsky-Golay smoothing) might have equal or
+greater predictive utility. We thus also test smoothed variants of the
+eigenvalues as predictive features: the sorted eigenvalues are smoothed with
+either a uniform (moving average) filter or Savitsky-Golay, using window sizes
+of 3, 5, 7, and 9 to yield 8 total ``smoothed eigenvalue'' features.
 
@@
 
-It could easily be the case that a combination of RMT and non-RMT eigenfeatures
-are more predictive than either alone. We thus test a variety of eigenfeature
-combinations, with the motivation for these particular combinations being to
-test whether RMT eigenfeatures add anything beyond smoothing. Combined features
-are formed by concatenation, so that if we have features \(f_1, \dots,  f_n\)
-with dimensions \(p_1, \dots, p_n\), then the combined feature is \([f_1;
-\dots; f_n]\) with \(\sum p_i\) dimensions.  The final combinations chosen can
-be found in Table \ref{tab:features}.
+A combination of RMT and non-RMT eigenfeatures
+could be more predictive than either alone. We test this with  a variety of eigenfeature
+combinations. The particular combinations are especially chosen to
+test whether RMT eigenfeatures add anything beyond smoothing.
+
+Combined features are formed by concatenation, so that if we have features
+\(f_1, \dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the combined
+feature is \([f_1; \dots; f_n]\) with \(\sum p_i\) dimensions.  The final
+combinations chosen can be found in Table \ref{tab:features}.
 
@@
 
-We might expect the largest, middle, or smallest eigenvalues to be most useful
+The largest, middle, or smallest eigenvalues could be most useful
 in characterizing some system. E.g. if the smallest eigenvalues correspond to
-random / noise aspects of a system, but it is differences in the nature of the
-noise that most differentiate between certain systems, then here the smallest
+random / noise aspects of a system, but differences in the nature of the
+system noise most differentiate between systems, then the smallest
 eigenvalues may have the most predictive utility.
@@
 Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
-locality at which we summarize the spectrum: at small \(L\)-values, the
+locality at which we summarize the spectrum: with small \(L\)-values, the
 spectral rigidity summarizes the non-linearity of eigenvalues that are
@@
 difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
-percentile). These statistics are used to summarizing / reduce the fMRI data
-along the voxel dimension, yielding a feature of \(T\) dimensions. E.g. the ``mean''
-feature is just the global mean signal. Exact details can be found in the
+percentile). These statistics are used to reduce the fMRI data
+along the voxel dimension, yielding a feature of \(T\) dimensions (e.g. the ``mean''
+feature is just the global mean signal). Exact details can be found in the
 \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
@@
 smoothing degrees of these features, to account for noise and to be somewhat
-similar to the various smoothing variants for the eigenfeatures. Each baseline
-feature thus has a uniformly smoothed variant, where the size of the smoothing
+similar to the various smoothing variants for the eigenfeatures: each baseline
+feature has a uniformly smoothed variant, where the size of the smoothing
 window is either 1 (no smoothing), 2, 4, 8, or 16.
@@
 
-We use a variety of stadard machine-learning classifiers available in the
+We use a variety of standard machine-learning classifiers available in the
 Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
@@
 basis function, and k-nearest neighbours classifiers with k equal to 3, 5, and
-9 (KNN3, KNN5, and KNN9, respectively). We originally also attempted to test a
-simple logistic regression classifier, but found that this model to frequently
-failed to converge for a number of features, even after increasing iterations
-significantly. In all cases the default hyperparameter values were used.
+9 (KNN3, KNN5, and KNN9, respectively). In all cases the default hyperparameter
+values were used. We originally also attempted to test a simple logistic
+regression classifier, but found that this model frequently failed to converge
+for a number of features, even after increasing iterations significantly.
 
@@
 likely to have mAUROC \(<\) 0.5. Overall, however, differences in the mAUROC distributions
-were small.
+due to classifier were small.
 
@@
 these trimming methods, these features were consistently more predictive than
-not.
+not (Figure S\ref{S-fig:fine-trim}).
 
@@
 It is likely that the unfolded eigenvalues have predictive utility because of
-their smoothing and re-scaling effect. It is clear from Figure \ref{fig:unfolded},
-which depicts the raw eigenvalues and unfolded eigenvalues with different trimming
-procedures, that the raw eigenvalues have an extremely exponential distribution
-(Figure \ref{fig:unfolded} already uses a log-scale for plotting) due to the largest
-eigenvalues. The trimmed feature distributions are clearly far less pathologically distributed.
-Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier} that
+their smoothing and re-scaling effect. Figure \ref{fig:unfolded}, which depicts
+the raw eigenvalues and unfolded eigenvalues with different trimming
+procedures, shows how the raw eigenvalues have strongly exponential
+distributions, even with logarithmic axes. This is due to the
+magnitude of the largest eigenvalues, and the trimmed feature distributions in
+Figure \ref{fig:unfolded} are less pathologically distributed.
+
+Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier}, that
 a simple combination of the raw eigenvalues plus smoothed versions of them are the most
-generally-predictive features examined here.
+generally-predictive features examined.
 
@@
 strongly separated (e.g. Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}).
-
-In this study, the median accuracy improvement upon guessing was 0 for all features, although
-eigenfeatures could improve upon guessing by 16.2\% - 33.7\% (\ref{S-tab:acc-numerical}).
+In this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\% (\ref{S-tab:acc-numerical}).
 

### 2022-11-12 code/scripts/summarize_everything_updated.py

@@ -2432,9 +2432,9 @@ def make_kde_plots() -> None:
         fig.set_size_inches(w=10, h=8)
         fig.tight_layout()
         fig.subplots_adjust(
-            top=0.81, bottom=0.085, left=0.04, right=0.96, hspace=0.35, wspace=0.22
+            top=0.795, bottom=0.085, left=0.04, right=0.96, hspace=0.35, wspace=0.22
         )
-        sbn.move_legend(grid, loc=(0.01, 0.85))
+        sbn.move_legend(grid, loc=(0.01, 0.825))
         savefig(fig, "rmt_feature_auroc_by_trim.png")
 
     def plot_rmt_by_degree() -> None:
@@ -2561,11 +2561,11 @@ def make_kde_plots() -> None:
     # plot_by_coarse_predictive_feature_preproc_subgroup()
     # plot_by_fine_feature_group_predictive_norm_subgroup()
     # plot_by_fine_predictive_feature_preproc_subgroup()
-    plot_by_fine_feature_group_predictive_classifier_subgroup()
+    # plot_by_fine_feature_group_predictive_classifier_subgroup()
     # plot_largest_by_fine_feature_subgroup()
     # plot_smallest_by_fine_feature_subgroup()
     # plot_by_coarse_feature_group_predictive_classifier_subgroup()
-    # plot_rmt_by_trim()
+    plot_rmt_by_trim()
     # plot_rmt_by_degree()
     # plot_rmt_largest_by_degree()
     # plot_all_by_fine_feature_groups_best_params()
@@ -2947,6 +2947,8 @@ if __name__ == "__main__":
     # df_ses = load_combined(drop_ses=False)
     # df.to_json(PROJECT / "EVERYTHING.json")
     # print(f"Saved all combined data to {PROJECT / 'EVERYTHING.json'}")
+    make_kde_plots()
+
     feature_summary = (
         df[df.subgroup.isin(OVERALL_PREDICTIVE_GROUP_ORDER)]
         .groupby("feature")
@@ -2963,22 +2965,21 @@ if __name__ == "__main__":
     simplefilter("ignore", UserWarning)
     # plot_unfolded(UpdatedDataset.Vigilance, group1="vigilant", group2="nonvigilant")
     # plot_unfolded(UpdatedDataset.Older, group1="younger", group2="older")
-    plot_unfolded(
-        UpdatedDataset.Osteo,
-        preproc=PreprocLevel.SliceTimeAlign,
-        group1="duloxetine",
[…]

### 2022-11-12 paper/paper/SPIE/article.tex
[WIP]

@@
 
-\newcommand{\code}[1]{\small \texttt{#1} \normalsize}
-\newcommand{\tcode}[1]{\footnotesize \texttt{#1} \normalsize}
+\newcommand{\code}[1]{\small\texttt{#1}\normalsize}
+\newcommand{\tcode}[1]{\footnotesize\texttt{#1}\normalsize}
 
@@
 \section*{Results}
-% df.groupby(["subgroup", "gross_feature"])["auroc"].median()
-Across all classification tasks and analytic choices, we find eiegenfeatures to
-more often have predictive utility than not (82.4\% of median AUROCs \(>\) 0.5;
-median AUROC range across classification tasks 0.47 - 0.64). Simple baseline
-reductions on source timeseries, by contrast, were less useful (58.8\% of
-median AUROCs \(>\) 0.5, median AUROC range across classification tasks 0.42 -
-0.62). Additionally, eigenfeature AUROC distributions were overall much more
-right-tailed than baseline features suggesting greater predictive potential.
-However, performance distributions were wide, and often significantly affected by
-analytic choices.
+Across all classification tasks and analytic choices, we find RMT- and
+eigenvalue-based "eigenfeatures" to have predictive utility more often than
+not (82.4\% of median AUROCs \(>\) 0.5; median AUROC range across
+classification tasks 0.47 - 0.64). Simple baseline reductions on source
+timeseries, by contrast, were less useful (58.8\% of median AUROCs \(>\) 0.5,
+median AUROC range across classification tasks 0.42 - 0.62). Additionally,
+eigenfeature AUROC distributions were overall more right-tailed than
+baseline features, suggesting greater predictive potential. However, performance
+distributions were wide, and often significantly affected by analytic choices.
 
@@
 
-At its most rudimentary, RMT describes the expected behaviour of the
+At its most rudimentary, RMT describes the expected behavior of the
 eigenvalues—also often called levels—of a number of classes of random
@@
 veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
-even more recently, has provided insights into the behaviours of deep
+even more recently, has provided insights into the behaviors of deep
 neural networks\cite{martinImplicitSelfRegularizationDeep2021,
@@
 blood-oxygenation-level-dependent (BOLD) signals are related to neural
-activity, RMT may be an ideal starting point. Each voxel time course (or
+activity, RMT may be an ideal starting point, as each voxel time course (or
 collection of such time sources - i.e. ROIs) can be considered an interacting
-component of the system. Likewise, in \textit{functional connectivity}
+component of the system. Likewise, in functional connectivity
 analyses, statistical relationships of the BOLD ROI time-courses are
-investigated in the hope of gaining insights into brain functioniong (see
+investigated in the hope of gaining insights into brain functioning (see
 \citenum{bastosTutorialReviewFunctional2016,
@@
 
-RMT has also been used in ROI-based fMRI functional-connectivity studies to
+RMT has also been used in ROI-based fMRI functional connectivity studies to
 investigate differences between rest and task states
 \cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
-ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain states
-\cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the
-spectra of resting or low-attention states exhibited properties close to the
-GOE. These findings suggest that certain aspects of psychological processes
-might be characterized, in part, by features computed from the eigenvalues of
-fMRI correlation matrices, and that these features might vary in an
-interpretable manner across psychological processes. If this is the case, RMT
-could aid in understanding the functioning of the human brain.
-
-\subsection{Eigenvalues and Prediction}
-
-The basic insight of RMT is thus that the eigenvalues alone may provide
+attention-deficit hyperactive disorder (ADHD)
+\cite{wangRandomMatrixTheory2016}, between pain and non-pain states
+\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs.
+right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
+were found in the latter study, across the first three studies, the spectra of
+resting or low-attention states exhibited properties close to the GOE. These
+findings suggest that certain aspects of psychological processes might be
+characterized, in part, by features computed from the eigenvalues of fMRI
+correlation matrices, and that these features might vary in an interpretable
+manner across psychological processes. If this is the case, RMT could aid in
+understanding the functioning of the human brain.
+
+\subsection{Eigenvalue Features}
+
+The basic insight of RMT is thus that eigenvalues alone may provide
 interesting information about highly complex systems. However, real systems are
@@
 interactions. RMT is statistical in nature, describing only the expected
-behaviour of the spectra of \textit{iid} random matrics. To this end, a number
+behavior of the spectra of \textit{iid} random matrices. To this end, a number
 of summary statistics or ``spectral observables''
@@
 eigenvalues alone can provide understanding of a system, then those eigenvalues
-(or other simple transformations of them) ought also to be predictively
-useful. We examine a number of such \textit{eigenfeatures} in this study.
-
-In addition, the functional connectivity - the matrix of correlations of
-various collections of voxels of an fMRI scan - is a priori a useful
+(or other simple transformations of them) ought also to be predictively useful.
+We examine a number of such features in this study, and refer to both
+RMT-derived features like the rigidity and level variance, and non-RMT-derived
+features collectively as \textit{eigenfeatures} in this study.
+
+\subsection{Functional Connectivity Reduction}
+
+The functional connectivity—the matrix of correlations of
+various collections of voxels of an fMRI scan—is \textit{a priori} a useful
 representation of the fMRI data. However, the full correlation matrix between
@@
 RMT suggests a potentially useful reduction in the form of the eigenvalues of
-the voxelwise functional connectivity. An \(N \times t\) matrix \(\bf{A}\) of
+the voxelwise functional connectivity. An \(N \times t\) matrix of
 \(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
@@
 noting only differences in RMT across subgroups and/or conditions. For example
-\citenum{wangRandomMatrixTheory2016,guRandomMatrixTheory2020} use a
-Kolmogorov-Smirnov test, and \citenum{wangSpectralPropertiesTemporal2015a} uses
-a t-test to provide evidence of differences in various RMT metrics. However,
-statistically significant differences in a metric do not imply practical
-significance or predictive utility \cite{loWhySignificantVariables2015}.
+\citenum{wangRandomMatrixTheory2016} and \citenum{guRandomMatrixTheory2020} use
+a Kolmogorov-Smirnov test, and \citenum{wangSpectralPropertiesTemporal2015a}
+uses a t-test to provide evidence of differences in various RMT metrics.
+However, statistically significant differences in a metric do not imply
+practical significance or predictive utility
+\cite{loWhySignificantVariables2015}, nor do they imply generalization or
+replicability \cite{amrheinEarthFlat052017}. Cross-validation, by contrast,
+attempts to more directly assess these properties
+\cite{yarkoniChoosingPredictionExplanation2017}.
 
@@
 reproduce results. The extraction of the spectral rigidity and level variance
-is computationally-demanding and non-trivial, requiring an \textit{unfolding}
-procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004},
-which is, unfortunately, a poorly-documented fitting procedure that requires
-making a number of arguably subjective decisions. These fitting decisions can
-dramatically impact RMT conclusions
+is computationally-demanding and mathematically and algorithmically
+non-trivial, and additionally requires an \textit{unfolding}
+procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}.
+Unfolding is an exponential fitting procedure that involves a number of highly
+subjective decisions regarding outliers and the flexibility of the fitting
+function. These decisions are, unfortunately, often poorly or even entirely
+undocumented, but are known to often dramatically impact RMT conclusions
 \cite{abul-magdUnfoldingSpectrumChaotic2014,
@@
 abueleninEffectUnfoldingSpectral2012, moralesImprovedUnfoldingDetrending2011}.
+
 On top of the flexibility in the implementation and application of RMT to
@@
 freedom"\cite{simmonsFalsePositivePsychologyUndisclosed2011b} in combination
-with the absence of clear code, and exclusively descriptive/explanatory
-analytic approaches means that the practical utility of RMT for fMRI is largely
-uninvestigated.
+with the absence of reproducible code and data availability, and predominance
+of descriptive and explanatory approaches, raise doubts about the basic robustness
+and practical utility of RMT-based functional connectivity analyses.
 
@@
 \href{https://openneuro.org/}{OpenNeuro} platform (36). Selection criteria was
-somewhat subjective, but we required (1), that all scans have the same spatial
-and temporal resolutions, (2), the the dataset comprise 10+ human subjects, and
-(3), that it was reasonable to classify at the level of an entire scan, i.e.
-neither task nor event timings were needed to determine class membership. This
-yielded seven datasets.
+somewhat subjective, but we required (1), the dataset comprise 10+ human
+subjects, (2) that all fMRI scans have the same spatial and temporal
+resolutions, (3) that the fMRI images could be split into various classes for a
+number of binary classification tasks, and (4), that it was reasonable to
+classify each scan in its entirety, (i.e. neither task nor event timings
+were needed to determine class membership, nor did any scan volumes need to be removed
+to make the scan classifiable). This yielded seven datasets total.
 
@@
 Attention&  vigilant           & 11     & resting-state       & 2 \\
-        &  nonigilant         & 11     & resting-state        & 2 \\
+        &  nonvigilant         & 11     & resting-state        & 2 \\
         &  trait-attentive     & 11    & resting-state        & 2 \\
@@
 bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans from
-non-depressed controls and mildy- or moderately-depressed subjects. In our analysis,
+non-depressed controls and mildly- or moderately-depressed subjects. In our analysis,
 the binary classification task is \code{depress v control}.
@@
 wakefulness and attention over the past weeks. Finally, the
-\code{task\_attend v tas\_nonattend} classification task was constructed
-using the scores of The Conjunctive Continuous Performance Task
-(CCPT\cite{shalevConjunctiveContinuousPerformance2011}). This is a behavioural
+\code{task\_attend v task\_nonattend} classification task was constructed
+using the scores of the Conjunctive Continuous Performance Task
+(CCPT\cite{shalevConjunctiveContinuousPerformance2011}). This is a behavioral
 performance task which requires the subject to quickly and selectively respond
@@
 
-We limited preprocessing to a simple pipeline of, in order, (1) brain
+We limited preprocessing to a simple pipeline of, in order: (1) brain
 extraction, (2) slice timing correction, (3), motion correction, and (4),
@@
 \cite{ciricTemplateFlowFAIRsharingMultiscale2021}.
-
-We limited preprocessing methods to these steps sometimes because data required
-for other typical preprocessing steps (e.g. field intensity maps, physiological
-measurements) was missing, but also because all pipelined intermediates were saved to later
-compare the effect of increasing degrees of pre-processing (see Section
-\ref{sec:multiverse}), and it was important to limit the number of
-preprocessing degrees to prevent excessive computational costs.
 
@@
 
+We limited preprocessing methods to these steps sometimes because data required
+for other typical preprocessing steps (e.g. field intensity maps, physiological
+measurements) was missing, but also because all pipeline intermediates were
+saved to later compare the effect of increasing degrees of preprocessing on
+feature predictive utility (see Section \ref{sec:multiverse}), and it was
+important to limit the number of preprocessing steps to prevent excessive
+computational costs and comparisons.
+
 \subsection{Feature Extraction}
@@
 
-\subsubsection{Raw Eigenvalues}
-\label{sec:raw-eigs}
-
-The time and space complexities of computing the full \(N \times N\)
-correlation matrix of \(\mathbf{M}\), and the subsequent eigenvalues, are too
-large to be tractable. However, since transposition does not change the
-eigenvalues, we can use the transpose to efficiently compute these eigenvalues
-(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
-in Table \ref{tab:features}), most directly test the idea that eigenvalues
-alone are useful functional connectivity features.
-
-However, not all eigenvalues may have predictive utility. We also examine as
-features the central eigenvalues (defined as the middle 10\%, 20\%, or 40\% of
-the sorted eigenvalues, \code{eigsmiddle} in Table \ref{tab:features}), and
-tail eigenvalues (defined as the first and last 5\%,  10\%, and 20\% of each
-tail of the spectrum, \code{eigsminmax} in Table \ref{tab:features}).
-
-\subsubsection{RMT Features}
-\label{sec:rmt-features}
-
-To compare the observed eigenvalue distribution to some RMT theoretical
-predictions, the eigenvalues must be unfolded
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
-details and motivation behind the unfolding procedure are well documented in
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
-can be viewed as a smoothing and rescaling procedure where the originally
-observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
-an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
-if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
-function is not known, and so must be approximated
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
-a polynomial is chosen \cite{abul-magdUnfoldingSpectrumChaotic2014}.
-
-Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity,
-\(\Delta_3(L)\), is calculated for any positive real value \(L <
-\max(\Lambda)\) as:
-\begin{equation}
-\label{eq:rigidity}
-\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
-\end{equation}
-where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
-equal to \(\lambda\), \(\langle \cdot \rangle_c\) denotes the average with
-respect to all starting points \(c\), and where \(A\) and \(B\) denote the
-slope and intercept, respectively, of the least squares fit of a straight line
-to \(\eta(\lambda)\) on \([c, c+L]\)
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as a timeseries, the spectral rigidity for a value \(L\) is the
-\textit{average nonlinearity} of all intervals of length \(L\) over the series.
-
-The level number variance, \(\Sigma^2(L)\), or \textit{level variance}, for short, is
-closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
-calculated as:
-\begin{equation}
-\label{eq:levlvar}
-\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
-\end{equation}
-where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
-and where \(c\), \(L\), and \(\langle \cdot \rangle_c\) are as above
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
-as an irregular timeseries, the level number variance is the
-variation of the number of samples in all intervals of length \(L\) over the
-series.
-
-Both the spectral rigidity and level number variance are also computed and
-investigated in this study, for all \(L \in \{ 1, 2, \dots 20 \}\).
-Because the computation of these statistics is non-trivial, we developed and
-release the separate, open-source Python library
-\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
-allows for efficient, parallel computation of the spectral rigidity and level
-variance via Monte Carlo methods, and automatically ensures convergence of the
-statistics to user-specified tolerances.
-
-\textit{Trimming Variants}. As the unfolding procedure operates on the sorted
-eigenvalues, and involves fitting a smooth polynomial or exponential to these
-values, extreme values are often trimmed (omitted from fitting)
-\cite{moralesImprovedUnfoldingDetrending2011,
-abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
-highly subjective, and not usually well-documented, and unfortunately, with a
-large number of different source matrices, it is too destructive to naively use
-the same hard criterion (e.g. largest 10\%, largest 3) for all unfoldings.
-
-We develop and test four trimming variants: no trimming, precision-based,
-largest, and middle, with less subjective criteria for determining trimming
-thresholds. The details of these trimming procedures are straightforward, and
-can be found in Appendix \ref{sec:trimming}, or the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
-code}.
-
-\textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
-with a polynomial. The choice of degree has been known to dramatically impact
-certain analyses involving the spectral rigidity or level
-variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
-moralesImprovedUnfoldingDetrending2011, abueleninSpectralUnfoldingChaotic2018,
-fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}. We
-thus compute the rigidity and level variance with all possible combinations of
-trimming and unfolding polynomial degrees of 3, 5, 7, and 9.
-
-\subsubsection{Smoothed Eigenfeatures}
-\label{sec:eigs-only}
-
-When computing RMT eigenfeatures, the polynomial fitting during unfolding has a
-smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
-operations. However, it is possible other, simpler transformations of the
-eigenvalues (e.g. uniform or Savitsky-Golay smoothing) might have equal or
-greater predictive utility. We thus also test smoothed variants of the
-eigenvalues as predictive features: the sorted eigenvalues are smoothed with
-either a uniform (moving average) filter or Savitsky-Golay, using window sizes
-of 3, 5, 7, and 9 to yield 8 total ``smoothed eigenvalue'' features.
-
-\subsubsection{Feature Combinations}
-
-A combination of RMT and non-RMT eigenfeatures
-could be more predictive than either alone. We test this with  a variety of eigenfeature
-combinations. The particular combinations are especially chosen to
-test whether RMT eigenfeatures add anything beyond smoothing.
-
-Combined features are formed by concatenation, so that if we have features
-\(f_1, \dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the combined
-feature is \([f_1; \dots; f_n]\) with \(\sum p_i\) dimensions.  The final
-combinations chosen can be found in Table \ref{tab:features}.
-
-\subsubsection{Slicing Variants}
-
-The largest, middle, or smallest eigenvalues could be most useful
-in characterizing some system. E.g. if the smallest eigenvalues correspond to
-random / noise aspects of a system, but differences in the nature of the
-system noise most differentiate between systems, then the smallest
-eigenvalues may have the most predictive utility.
-
-Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
-locality at which we summarize the spectrum: with small \(L\)-values, the
-spectral rigidity summarizes the non-linearity of eigenvalues that are
-relatively close to each other in magnitude. At large values of \(L\), the rigidity
-summarizes the long-range non-linearity of the spectrum.
-
-Since predictive utility may vary with summary locality, or with the region of
-the original spectrum, we investigate various front, middle, and end contiguous
-slices of each eigenfeature in all analyses, where the size of each slice is
-either the first or last 5\%, 10\%, or 20\% of the full eigenfeatures for
-non-middle slices, or the middle 10\%, 20\%, or 40\% of the full eigenfeature for
-the middle eigenvalues.
-
-For combined features, slicing variants are also computed, with slicing
-performed first on each feature to be combined. E.g. if we have features \(f_1,
-\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
-feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
-elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
-\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
-
-\subsubsection{Baseline Features}
-\label{sec:baselines}
-
-If RMT- or eigenvalue-based features fail to predictively outperform features
-that are simple summary statistics of the fMRI data, it is difficult to justify
-the greater computational and interpretational complexities of the former. We
-compute ten simple statistical reductions of the voxel time series ("timeseries
-features") to help assess the relative value of RMT features.
-
-Baseline reductions use as statistical summaries: robust measures of location
-(mean, max, min); non-robust measures of location (median, \(95^{\text{th}}\)
-percentile, \(5^{\text{th}}\) percentile); non-robust measures of scale
-(standard deviation, range); and robust measures of scale (interquartile range,
-difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
-percentile). These statistics are used to reduce the fMRI data
-along the voxel dimension, yielding a feature of \(T\) dimensions (e.g. the ``mean''
-feature is just the global mean signal). Exact details can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
-source code}.
-
-We do not take slice variants of these baseline features, since the baseline
-features are still time series. However, we do also evaluate a number of
-smoothing degrees of these features, to account for noise and to be somewhat
-similar to the various smoothing variants for the eigenfeatures: each baseline
-feature has a uniformly smoothed variant, where the size of the smoothing
-window is either 1 (no smoothing), 2, 4, 8, or 16.
-
-\subsection{Classifiers}
-
-We use a variety of standard machine-learning classifiers available in the
-Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
-solve each classification task. We use a gradient-boosted decision tree (GBDT),
-random forest (RF) classifier, support vector classifier (SVC) with radial
-basis function, and k-nearest neighbours classifiers with k equal to 3, 5, and
-9 (KNN3, KNN5, and KNN9, respectively). In all cases the default hyperparameter
-values were used. We originally also attempted to test a simple logistic
-regression classifier, but found that this model frequently failed to converge
-for a number of features, even after increasing iterations significantly.
-
-\subsection{Preprocessing Levels}
-
-Each preprocessing intermediate in the (1) brain extraction, (2) slice time
-correction, (3) motion correction, and (4) template registration pipeline was
-saved and used for entirely separate feature analyses. E.g. all features were
-extracted from an fMRI image prepared with one of four levels or degrees of
-preprocessing, where preprocessing level \(k\) includes preprocessings steps 1 to \(k\),
-inclusive, and starting at \(k = 1\).
-
-\subsection{Normalization}
-
-Because of the exponential distribution of the eigenfeatures, normalization is
-somewhat non-trivial, and a simple standardization or min-max normalization is
-ineffective. We instead first apply a logarithm to all eigenfeatures, and then
-test each classifier with un-normalized or min-max normalized versions of the
-log-features. Baseline features are also tested with un-normalized and min-max
-normalized versions, but without any log transform.
-
 \begin{table}[h!]
 \caption{
+\label{tab:features}
 Feature groupings for summarization. eigs = non-RMT eigenfeatures and their combinations with
@@
 }
-\label{tab:features}
 \small
@@
 
+\subsubsection{Raw Eigenvalues}
+\label{sec:raw-eigs}
+
+The time and space complexities of computing the full \(N \times N\)
+correlation matrix of \(\mathbf{M}\) (and subsequent eigenvalues) are too
+large to be tractable. However, since transposition does not change the
+eigenvalues, we can use the transpose to efficiently compute these eigenvalues
+(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
+in Table \ref{tab:features}), most directly test the idea that eigenvalues
+alone are useful functional connectivity features.
+
+However, not all eigenvalues may have predictive utility. We also examine as
+features the central eigenvalues (defined as the middle 10\%, 20\%, or 40\% of
+the sorted eigenvalues, \code{eigsmiddle} in Table \ref{tab:features}), and
+tail eigenvalues (defined as the first and last 5\%,  10\%, and 20\% of each
+tail of the spectrum, \code{eigsminmax} in Table \ref{tab:features}).
+
+\subsubsection{RMT Features}
+\label{sec:rmt-features}
+
+To compare the observed eigenvalue distribution to some RMT theoretical
+predictions, the eigenvalues must be unfolded
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
+details and motivation behind the unfolding procedure are well documented in
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
+can be viewed as a smoothing and rescaling procedure where the originally
+observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
+an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
+if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
+function is not known, and so must be approximated
+\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
+a polynomial is chosen \cite{abul-magdUnfoldingSpectrumChaotic2014}.
+
+Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity,
+\(\Delta_3(L)\), is calculated for any positive real value \(L <
+\max(\Lambda)\) as:
+\begin{equation}
+\label{eq:rigidity}
+\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
+\end{equation}
+where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
+equal to \(\lambda\), \(\langle \cdot \rangle_c\) denotes the average with
+respect to all starting points \(c\), and where \(A\) and \(B\) denote the
+slope and intercept, respectively, of the least squares fit of a straight line
+to \(\eta(\lambda)\) on \([c, c+L]\)
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
+as a timeseries, the spectral rigidity for a value \(L\) is the
+\textit{average nonlinearity} of all intervals of length \(L\) over the series.
+
+The level number variance, \(\Sigma^2(L)\), or \textit{level variance}, for short, is
+closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
+calculated as:
+\begin{equation}
+\label{eq:levlvar}
+\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
+\end{equation}
+where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
+and where \(c\), \(L\), and \(\langle \cdot \rangle_c\) are as above
+\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
+as an irregular timeseries, the level number variance is the
+variation of the number of samples in all intervals of length \(L\) over the
+series.
+
+As eigenfeatures, we compute the unfolded eigenvalues, and the spectral
+rigidity and level number variance for all \(L \in \{ 1, 2, \dots 20 \}\).
+These are \code{unfolded}, \code{rigidity}, and \code{levelvar}, respectively,
+in Table \ref{tab:features}.
+
+\textit{Trimming Variants}. As the unfolding procedure operates on the sorted
+eigenvalues, and involves fitting a smooth polynomial or exponential to these
+values, extreme values are often omitted from fitting
+\cite{moralesImprovedUnfoldingDetrending2011,
+abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
+subjective, and unfortunately, with a large number of different source
+matrices, it is too destructive to naively use the same hard criterion (e.g.
+largest 10\%, largest three) for all unfoldings.
+
+We develop and test four trimming variants: no trimming, precision-based,
+largest, and middle, with less subjective criteria for determining trimming
+thresholds. The details of these trimming procedures are straightforward, and
+can be found in Appendix \ref{sec:trimming}, or the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
+code}.
+
+\textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
+with a polynomial. Because the choice of degree has been known to dramatically impact
+certain analyses involving the spectral rigidity or level
+variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
+moralesImprovedUnfoldingDetrending2011, abueleninSpectralUnfoldingChaotic2018,
+fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}, we
+compute the rigidity and level variance with all possible combinations of
+trimming and unfolding polynomial degrees of 3, 5, 7, and 9.
+
+\textit{Software Release}. Because the computation of these statistics is non-trivial, and to make the
+unfolding procedure more transparent and reproducible, we developed and release
+the separate, open-source Python library
+\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
+allows for efficient, parallel computation of the spectral rigidity and level
+variance via Monte Carlo methods, and automatically ensures convergence of the
+statistics to user-specified tolerances. The library also makes available
+various other functions useful for empirical RMT analyses, such as unfolding
+and trimming functions, and plotting facilities for classic RMT ensembles.
+
+\subsubsection{Smoothed Eigenfeatures}
+\label{sec:eigs-only}
+
+When computing RMT eigenfeatures, the polynomial fitting during unfolding has a
+smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
+operations. However, it is possible other, simpler transformations of the
+eigenvalues (e.g. uniform or Savitsky-Golay smoothing) might have equal or
+greater predictive utility. We thus also test smoothed variants of the
+eigenvalues as predictive features: the sorted eigenvalues are smoothed with
+either a uniform (moving average) filter or Savitsky-Golay, using window sizes
+of 3, 5, 7, and 9 to yield 8 total features (\code{eigs\_smooth} and
+\code{eigs\_savgol} in Table \ref{tab:features}).
+
+\subsubsection{Feature Combinations}
+
+A combination of RMT and non-RMT eigenfeatures could be more predictive than
+either alone. We test this with a variety of eigenfeature combinations, with a
+focus on combinations that involve the simplest features (e.g. raw eigenvalues,
+unfolded eigenvalues) and then features involving additional processing (e.g.
+smoothed eigenvalues, level variance, rigidity). Combined features are formed
+by concatenation, so that if we have features \(f_1, \dots,  f_n\) with
+dimensions \(p_1, \dots, p_n\), then the combined feature is \([f_1; \dots;
+f_n]\) with \(\sum p_i\) dimensions.  The final combinations chosen can be
+found in Table \ref{tab:features}, where the "\code{+}" symbol indicates concatenation.
+
+\subsubsection{Slicing Variants}
+
+The largest, middle, or smallest eigenvalues could be most useful in
+characterizing some system. E.g. if the smallest eigenvalues correspond to
+random / noise aspects of a system, but differences in the nature of the system
+noise most differentiate between systems, then the smallest eigenvalues may
+have the most predictive utility.  Likewise, the \(L\) value in each RMT
+eigenfeature defines the degree of locality at which we summarize the spectrum:
+with small \(L\)-values, the spectral rigidity summarizes the non-linearity of
+eigenvalues that are relatively close to each other in magnitude. At large
+values of \(L\), the rigidity summarizes the long-range non-linearity of the
+spectrum.
+
+Since predictive utility may vary with summary locality, or with the region of
+the original spectrum, we investigate various front, middle, and end contiguous
+slices of each eigenfeature in all analyses, where the size of each slice is
+either the first or last 5\%, 10\%, or 20\% of the full eigenfeatures for
+non-middle slices, or the middle 10\%, 20\%, or 40\% of the full eigenfeature for
+the middle eigenvalues.
+
+For combined features, slicing variants are also computed, with slicing
+performed first on each feature to be combined. E.g. if we have features \(f_1,
+\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
+feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
+elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
+\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
+
+\subsubsection{Baseline Features}
+\label{sec:baselines}
+
+If RMT- or eigenvalue-based features fail to predictively outperform features
+that are simple summary statistics of the fMRI data, it is difficult to justify
+the greater computational and interpretational complexities of the former. We
+compute ten simple statistical reductions of the voxel time series ( "tseries"
+features in Table \ref{tab:features}) to help assess the relative value of RMT
+features.
+
+The baseline reductions used as statistical summaries were: robust measures of
+location (mean, max, min); non-robust measures of location (median,
+\(95^{\text{th}}\) percentile, \(5^{\text{th}}\) percentile); non-robust
+measures of scale (standard deviation and the range—\code{T-rng} in Table
+\ref{tab:features}); and robust measures of scale (interquartile range and
+difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
+percentile—\code{T-rrng} in Table \ref{tab:features}). Each statistic reduces
+the fMRI data along the voxel dimension, yielding a feature of \(T\) dimensions
+(e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
+Exact details can be found in the
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
+source code}.
+
+We do not take slice variants of these baseline features, since the baseline
+features are still time series. However, we do also evaluate a number of
+smoothing degrees of these features, to account for noise and to be somewhat
+similar to the various smoothing variants for the eigenfeatures. Each baseline
+feature is tested with a degree of uniform smoothing, where the size of the
+smoothing window is either 1 (no smoothing), 2, 4, 8, or 16.
+
+\subsection{Classifiers}
+
+We use a variety of standard machine-learning classifiers available in the
+Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
+solve each classification task. We use a gradient-boosted decision tree (GBDT),
+random forest (RF) classifier, support vector classifier (SVC) with radial
+basis function, and k-nearest neighbors classifiers with k equal to 3, 5, and 9
+(KNN3, KNN5, and KNN9, respectively), in all cases with the default
+hyperparameter values. We originally also attempted to test a simple logistic
+regression classifier, but found that this model frequently failed to converge
+for a number of features, even after increasing iterations significantly.
+
+\subsection{Preprocessing Levels}
+
+Each preprocessing intermediate in the (1) brain extraction, (2) slice time
+correction, (3) motion correction, and (4) template registration pipeline was
+saved and used for entirely separate feature analyses. E.g. all features were
+extracted from an fMRI image prepared with one of four levels or degrees of
+preprocessing, where preprocessing level \(k\) includes preprocessing steps 1 to \(k\),
+inclusive, and starting at \(k = 1\).
+
+\subsection{Normalization}
+
+Because of the exponential distribution of the eigenfeatures, normalization is
+somewhat non-trivial, and a simple standardization or min-max normalization is
+ineffective. We instead first apply a logarithm to all eigenfeatures, and then
+test each classifier with un-normalized or min-max normalized versions of the
+log-features. Baseline features are also tested with un-normalized and min-max
+normalized versions, but without any log transform.
+
 \subsection{Multiverse Analysis}
@@
 evaluate feature predictive utility.
+
+The AUROC metric was chosen because it handles class imbalances present in the
+various datasets, and is naturally normalized and interpretable such that
+mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
+mAUROC \(>\) 0.5 indicates positive predictive
+utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
+collected as performance metrics the accuracy, adjusted accuracy (accuracy
+minus the proportion of the largest class), and the f1 score, and these metrics
+are available with all the raw data for this study \ref{sec:data-and-code}.
+
+the different
 
@@
 \begin{table}[h!]
-\caption{Numerical summaries of feature mAUROCs across predictable comparisons, and all combinations of
-analytic choices, sorted by 95\% percentile (robust max) value. Bold values indicate column "best" values, when reasonable. }
+\caption{Numerical summaries of feature mAUROCs across predictable comparisons,
+and all combinations of analytic choices, sorted by 95\% percentile (robust
+max) value. Bold values indicate column "best" values, when reasonable. We
+note, however, that aggregating across datasets in this manner can be deeply
+misleading\cite{brabecModelEvaluationNonconstant2020}, and thus the values here
+are intended solely as supplements to distribution plots. }
 \label{tab:numerical}
@@
 
-It is also difficult to quantify the practical significance of the observed mAUROC values.
-With small datasets of rs- or task-fMRI data, binary classification using whole-brain
-features is challenging. For subject level classification, state of the art deep learning methods
-improve upon guessing by 17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
+It is also difficult to quantify the practical significance of the observed
+mAUROC values. With small datasets of rs- or task-fMRI data, binary
+classification using whole-brain features is challenging. For subject level
+classification, state of the art deep learning methods improve upon guessing by
+17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
 \cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
-\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive compulsive disorder
-\cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual feature engineering with more separable
-conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
-\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated custom feature
-extraction methods can achieve near perfect accuracies at classifying task vs. rest
-\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional connectivity data
-and with classical machine learning algorithms (such as SVM) we in general only expect large
-prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
-strongly separated (e.g. Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}).
-In this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\% (\ref{S-tab:acc-numerical}).
+\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive
+compulsive disorder \cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual
+feature engineering with more separable conditions (e.g. schizophrenia) can
+result in classification accuracies well above 90\%
+\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated
+custom feature extraction methods can achieve near perfect accuracies at
+classifying task vs. rest
+\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional
+connectivity data and with classical machine learning algorithms (such as SVM)
+we in general only expect large prediction accuracies (e.g. greater than 80\%)
+when the group functional connectivities are already strongly separated (e.g.
+Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}). In
+this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\%
+(\ref{S-tab:acc-numerical}).
 
@@
 of the eigenvalues are equally informative.
+
+\section{Conclusion}
+
+\section{Disclosures}
+
+\section{Acknowledgments and Funding}
+
+\section{Data, Materials, and Code Availability}
+\label{sec:data-and-code}
+
+All raw fMRI data used in this study is publicly available on the
+\href{https://openneuro.org/}{OpenNeuro}
+platform\cite{poldrackOpenfMRIOpenSharing2017,
+markiewiczOpenNeuroResourceSharing2021}.
+
+All code used specifically to perform preprocessing and analyses for this paper
+is publicly available \href{https://github.com/DM-Berger/random-matrix-fmri/}{online} in
+the paper GitHub repository\cite{dm-bergerDMBergerRandommatrixfmriPaper2022}.
+This repository also includes all produced prediction data in a single table
+available in the file \code{all\_produced\_prediction\_data.json.xz} available
+\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/all\_produced\_prediction\_data.json.xz}{in
+the paper repository}.
+
+Code used to compute the spectral observables and
+perform unfolding is publicly available in the
+\href{https://github.com/stfxecutables/empyricalRMT}\code{empyricalRMT}
+library\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}.
 
@@
 We chose one-dimensional k-means partly due to efficiency and simplicity, and
-because of the general relation to classical threholding methods like the Otsu
+because of the general relation to classical thresholding methods like the Otsu
 method\cite{liuOtsuMethodKmeans2009}.

### 2022-11-12 code/scripts/summarize_everything_updated.py

@@ -2947,18 +2947,18 @@ if __name__ == "__main__":
     # df_ses = load_combined(drop_ses=False)
     # df.to_json(PROJECT / "EVERYTHING.json")
     # print(f"Saved all combined data to {PROJECT / 'EVERYTHING.json'}")
-    make_kde_plots()
+    # make_kde_plots()
 
     feature_summary = (
         df[df.subgroup.isin(OVERALL_PREDICTIVE_GROUP_ORDER)]
         .groupby("feature")
-        .describe(percentiles=[0.05, 0.5, 0.95])["auroc"]
+        .describe(percentiles=[0.05, 0.5, 0.95])["f1"]
         .sort_values(by=["95%", "feature"], ascending=[False, True])
         .drop(columns=["count"])
         .loc[:, ["mean", "min", "5%", "50%", "95%", "max", "std"]]
         .round(3)
     )
-    print(feature_summary.to_latex())
+    print(feature_summary.to_markdown(tablefmt="simple"))
 
     # plot_unfolded_duloxetine()
     # plot_unfolded_aging()

### 2022-11-12 paper/paper/SPIE/article.tex
[WIP]

@@
 collected as performance metrics the accuracy, adjusted accuracy (accuracy
-minus the proportion of the largest class), and the f1 score, and these metrics
-are available with all the raw data for this study \ref{sec:data-and-code}.
-
-the different
+minus the proportion of the largest class), and the F1 score, and a complete
+table with these additional metrics is available with the data for this study
+(Section \ref{sec:data-and-code}).
 
@@
 
+\subsection{Overview}
+
 Summarizing this quantity of data requires some caution. Measures of location or
@@
 distributions, and so have chosen primarily to present our findings visually,
-with kernel density estimates. However, we also provide overall numerical
-summaries of the feature performances in Table \ref{tab:numerical}.
+with kernel density estimates.
+
+Because aggregating across datasets with different class imbalances can
+be extremely misleading, even with a metric like the mAUROC\cite{brabecModelEvaluationNonconstant2020},
+we restrict such summaries to Table \ref{tab:numerical} only, and note that
+the table is intended as a \textit{supplement only} to the distribution plots: no conclusions
+should be drawn about feature performance from the table values alone.
 
@@
 and all combinations of analytic choices, sorted by 95\% percentile (robust
-max) value. Bold values indicate column "best" values, when reasonable. We
-note, however, that aggregating across datasets in this manner can be deeply
-misleading\cite{brabecModelEvaluationNonconstant2020}, and thus the values here
-are intended solely as supplements to distribution plots. }
+max) value. Bold values indicate column "best" values, when reasonable.}
 \label{tab:numerical}
@@
 
-In addition, each analytic factor could produce an interaction. For example, it
-could be that trimming impacts the overall mAUROC distribution only at a
-certain preprocessing level and for certain feature slicing. However, there are
-too many potential interactions to present, so we limit our presentation to
-main effects.  As there are a large number of features, we also summarize
-patterns of results across two abstract groupings of similar features
-("coarse", and "fine") shown in Table \ref{tab:features}.
-
-It is also important to exclude a classification task or an analytic factor
-from summaries if there is no evidence that the classification task is solvable
-in general, or for that factor. We take as lack of evidence of solvability an
-mAUROC distribution that is either (1) roughly symmetric and has mean and
-median close to 0.5, or (2) with median and mode less than 0.5.
+In addition, a large number of interactions are possible between each analytic
+factor in this study. For example, it could be that trimming impacts the
+overall mAUROC distribution only at a certain preprocessing level and for
+certain feature slicing. However, as there are too many potential interactions
+to present, we limit our presentation to main effects.
+
+To aid in summarizing the performances of the large number of different
+features, we also summarize patterns of results across two abstract groupings
+of similar features ("coarse", and "fine") shown in Table \ref{tab:features}.
+Thus, for example, figures depicting the coarse feature grouping "eigs" in fact
+depict all observed mAUROC values for the the fine feature groupings of raw,
+max, middle, and smoothed eigenvalues, and the fine feature grouping "eigs
+smooth", for example, depicts all observed mAUROC values for both the
+\code{eigs\_savgol} and \code{eigs\_smooth} features.
 
@@
 \subsubsection{Non-Predictable Comparisons}
+
+It is important, to reduce figure clutter and complexity, to exclude a
+classification task or an analytic factor from summaries if there is no
+evidence that the classification task is solvable in general, or for that
+factor. We take as lack of evidence of solvability an mAUROC distribution that
+is either (1) roughly symmetric and has mean and median close to 0.5, i.e.
+performance appears random, or (2) with median and mode less than 0.5.
+
 Based on the mAUROC distributions for either coarsely-grouped (Figure
@@
 
+\subsection{Largest mAUROCs}
+Considering the various analytic choices as tunable parameters, it makes sense
+to examine the \textit{largest} portion of mAUROCs as an indication of the
+maximum predictive potential of the eigenfeatures. In this case, it is clear
+that eigenfeatures using RMT features almost always had the highest potential
+predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
+shows that this was primarily due to either the "rmt only" or "rmt + eigs"
+features (see Table \ref{tab:features}). However, RMT eigenfeatures were also most
+likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
+0.5; Figure S\ref{S-fig:fine-smallest}).
+
+Examining these RMT features more closely, it is clear these features'
+performance distributions differ mostly due to the unfolding procedure. That
+is, combined features that used the unfolded eigenvalues plus some other RMT
+eigenfeature tended to have visually-indistinguishable mAUROC distributions to
+those using the unfolded eigenvalues alone (Figures
+S\ref{S-fig:fine-trim}-\ref{S-fig:fine-degree}). The mAUROC distributions of
+these features differed most strongly in the tails (Figures
+S\ref{S-fig:fine-largest}-\ref{S-fig:fine-smallest}).
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{coarse_feature_largest_by_subgroup.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:largest} Distributions of largest 500 mAUROCs by coarse feature grouping.}
+\end{figure}
+
+\subsection{Effect of Preprocessing}
+At a coarse level of feature grouping, slice time correction followed by motion
+correction tended to slightly increase the predictive utility of the
+eigenfeatures (Figure \ref{fig:preproc}) relative to brain extraction only.
+Subsequent registration following these steps did not generally impact mAUROC
+distributions further, except in the \code{task\_attend v task\_nonattend}
+comparison, where registration reduced the predictive utility of the
+eigenfeatures (Figure \ref{fig:preproc}, second-last column). When examining
+features more finely, it is clear that preprocessing most impacts the mAUROC
+distribution of the largest and central eigenvalues ("eigs middle" and "eigs
+max" in Figure S\ref{S-fig:fine-preproc}).
+
 \begin{figure}
@@
 
-\subsection{Largest mAUROCs}
-Considering the various analytic choices as tunable parameters, it makes sense
-to examine the \textit{largest} portion of mAUROCs as an indication of the
-maximum predictive potential of the eigenfeatures. In this case, it is clear
-that eigenfeatures using RMT features almost always had the highest potential
-predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
-shows that this was primarily due to either the "rmt only" or "rmt + eigs"
-features (see Table \ref{tab:features}). However, these features were also most
-likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
-0.5; Figure S\ref{S-fig:fine-smallest}).
-
-Examining these RMT features more closely (Figure S\ref{S-fig:fine-degree}),
-it is clear these features are more predictive mostly due to the unfolding
-procedure: the mAUROC distributions with largest medians and modes are due to
-features that include the unfolded eigenvalues. Combined features that use
-the unfolded eigenvalues plus some other RMT eigenfeature have identical
-mAUROC distributions to those using the unfolded eigenvalues alone.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[width=6.5in]{coarse_feature_largest_by_subgroup.png}
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:largest} Distributions of largest 500 mAUROCs by coarse feature grouping.}
-\end{figure}
-
-\subsection{Effect of Preprocessing}
-At a coarse level, slice time correction followed by motion correction tended
-to slightly increase the predictive utility of the eigenfeatures (Figure
-\ref{fig:preproc}) relative to brain extraction only. Subsequent registration
-following these steps did not generally impact mAUROC distributions further,
-except in the \code{task\_attend v task\_nonattend} comparison, where
-registration reduced the predictive utility of the eigenfeatures (Figure
-\ref{fig:preproc}, second-last column). When examining features more finely, it
-is clear that preprocessing most impacts the mAUROC distribution of the largest
-and central eigenvalues ("eigs middle" and "eigs max" in Figure
-S\ref{S-fig:fine-preproc}).
-
 \subsection{Effect of Classifier}
-mAUROC distributions were generally similar across coarse or fine feature
-groupings (Figure S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}).
-In the \textit{Attention} \code{task\_attend v task\_nonattend} condition, eigenfeatures were modally predictive only with
-the random forest classifier, whereas in the \textit{Aging} data, a SVC was least
-likely to have mAUROC \(<\) 0.5. Overall, however, differences in the mAUROC distributions
-due to classifier were small.
+Within any feature grouping (coarse or fine), and within a classification task,
+mAUROC distributions were generally similar across classifiers (Figure
+S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}). Additionally,
+these figures also show that within each classification task, choice of
+classifier does not result in dramatic changes to the rough rank ordering of
+features. E.g. if features are ranked on predictive utility, using the median,
+modal, or bulk of the  mAUROC values, this rank ordering appears to remain
+similar across classifiers.
+
+In the \textit{Attention} \code{task\_attend v task\_nonattend} condition,
+eigenfeatures were modally predictive only with the random forest classifier,
+whereas in the \textit{Aging} data, a SVC was least likely to have mAUROC \(<\)
+0.5. Overall, however, differences in the mAUROC distributions due to
+classifier were small and inconsistent.
 
@@
 \subsection{Effect of Slicing}
-Features were slightly more predictive when using either the largest 5 or 10\% or middle
-25\% ("max-05", "max-10" and "mid-25" in Figure
-\ref{S-fig:best-params-slicing}), and least predictive when features primarily
-involved the smallest eigenvalues.
+The best slice size and location depended complexly on the classification task
+and feature, and few general summaries can be made of these interactions.
+However, features including the full spectrum (e.g. raw eigenvalues, smoothed
+eigenvalues, and their combinations with RMT eigenfeatures) were slightly more
+predictive when using the largest portions ("max-XX", rows in Figure
+S\ref{S-fig:best-params-slicing}), and usually least predictive when features
+primarily involved the smallest or central portions.
 
@@
 In this study, eigenfeatures inspired by RMT and derived from the eigenvalues
-of the full, whole-brain voxelwise fMRI correlation matrix, were found to have
+of the full, whole-brain voxelwise fMRI correlation matrix were found to have
 predictive utility across a wide variety of phenomena and analytic choices.
@@
 understanding functional connectivity.
+
+However, mAUROC values observed in this study were highly sensitive to , and mean and
+median mAUROC v
 

### 2022-11-12 paper/paper/SPIE/supplementary.tex
[WIP]

@@
 \begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-largest} Distributions of largest 500 mAUROCs across fine
+feature grouping, by comparison task. Note "rmt only" and "rmt + eigs" features
+tend to have the best possible performances across predictable tasks. }
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-smallest} Distributions of smallest 500 mAUROCs across fine
+feature groupings, by comparison task. Note "rmt only" and "rmt + eigs"
+features tend to have the worse possible performances across predictable
+tasks.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
 \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_by_preproc_predictive_subgroup.png}
@@
 \caption
-{ \label{fig:coarse-classifier}
-AUROC distributions across coarse feature groupings and predictable comparison tasks, by classifier.}
+{ \label{fig:coarse-classifier} AUROC distributions across coarse feature
+groupings and predictable comparison tasks, by classifier. Note the general
+similarity of each distribution within a particular classification task
+(column) and within each feature grouping.}
 \end{figure}
@@
 \caption
-{ \label{fig:fine-classifier}
-AUROC distributions across fine feature groupings and predictable comparison tasks, by classifier.}
-\end{figure}
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_by_subgroup.png}
-\end{center}
-\caption
-{ \label{fig:fine-largest}
-Distributions of largest 500 mAUROCs across fine feature grouping, by comparison task.}
-\end{figure}
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_by_subgroup.png}
-\end{center}
-\caption
-{ \label{fig:fine-smallest}
-Distributions of smallest 500 mAUROCs across fine feature groupings, by comparison task.}
+{ \label{fig:fine-classifier} AUROC distributions across fine feature groupings
+and predictable comparison tasks, by classifier. Note the general similarity of
+each distribution within a particular classification task (column) and within
+each feature grouping. Note also that, within a classification task (column),
+that the rank ordering of features, based on wither the median, mode, or mean,
+does not change dramatically or consistently from classifier to classifier. }
 \end{figure}
@@
 \caption
-{ \label{fig:fine-trim}
-Distributions of mAUROCs for unfolding-dependent RMT features, by trimming.}
+{ \label{fig:fine-trim} Distributions of mAUROCs for unfolding-dependent RMT
+features, by trimming. Note the tendency for a rightward shift in the mAUROC
+distributions of the features involving the unfolded eigenvalues when using
+largest or middle trimming (most dramatic in the Osteo nopain v duloxetine
+condition). The impact of these trimming methods on the rigidity and level
+variance features, however, was mixed (compare Vigilance data to Osteo nopain v
+duloxetine condition.)}
 \end{figure}
@@
 \begin{center}
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup.png}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_eigs_aurocs_by_subgroup_and_slicing.png}
 \end{center}
 \caption
-{ \label{fig:best-params}
-Distributions of mAUROCs when choosing ``best'' analytic choices and features.}
+{ \label{fig:best-params-slicing} Distributions of mAUROCs by slicing. Features
+involving the full spectrum (raw eigenvalues, smoothed eigenvalues, and rmt +
+eigs) sometimes have most positive mAUROC distributions when using the larger
+eigenfeature values (first three columns) or middle values (Osteo nopain v pain
+condition, Vigilance classification task).}
 \end{figure}
 
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup_and_slicing.png}
-\end{center}
-\caption
-{ \label{fig:best-params-slicing}
-Distributions of mAUROCs when choosing ``best'' analytic choices and features, by slicing.}
-\end{figure}
+% \begin{figure}[H]
+% \begin{center}
+% \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup.png}
+% \end{center}
+% \caption
+% { \label{fig:best-params}
+% Distributions of mAUROCs when choosing ``best'' analytic choices and features.}
+% \end{figure}
 

### 2022-11-12 code/scripts/summarize_everything_updated.py

@@ -1867,17 +1867,8 @@ def make_kde_plots() -> None:
         sbn.move_legend(grid, loc=(0.76, 0.09))
         savefig(fig, "best_rmt_params_by_subgroup.png")
 
-    def plot_all_by_fine_feature_groups_best_params_slice() -> None:
-        rmt = df.loc[df.coarse_feature.isin(["rmt"])]
-        # rmt = rmt.loc[rmt.trim.isin(["middle", "largest"])]
-        # rmt = rmt.loc[rmt.deg.isin([3, 9])]
-        # rmt = rmt.loc[rmt.preproc.isin(["MotionCorrect", "MNIRegister"])]
-        idx = rmt.feature.isin(["unfolded"])
-        rmt = rmt[idx]
-
-        other = df.loc[df.coarse_feature.isin(["eigs"])]
-        data = pd.concat([rmt, other], axis=0)
-
+    def plot_all_by_fine_feature_groups_slice() -> None:
+        data = df.loc[df.coarse_feature.isin(["rmt", "eigs"])]
         grid = sbn.displot(
             data=data,
             x="auroc",
@@ -1908,13 +1899,13 @@ def make_kde_plots() -> None:
         fig.suptitle(
             "Overall Distribution of AUROCs for each Fine Feature Group by Slicing"
         )
-        fig.set_size_inches(w=18, h=8)
+        fig.set_size_inches(w=10, h=8)
         fig.tight_layout()
         fig.subplots_adjust(
-            top=0.908, bottom=0.07, left=0.021, right=0.92, hspace=1.0, wspace=0.128
+            top=0.82, bottom=0.07, left=0.036, right=0.99, hspace=1.0, wspace=0.128
         )
-        sbn.move_legend(grid, loc=(0.93, 0.50))
-        savefig(fig, "best_rmt_params_by_subgroup_and_slicing.png")
+        sbn.move_legend(grid, loc=(0.01, 0.85))
+        savefig(fig, "rmt_eigs_aurocs_by_subgroup_and_slicing.png")
 
     def plot_all_by_fine_feature_groups_best_params_best_slice() -> None:
         rmt = df.loc[df.coarse_feature.isin(["rmt", "eigs"])]
@@ -2565,7 +2556,8 @@ def make_kde_plots() -> None:
     # plot_largest_by_fine_feature_subgroup()
     # plot_smallest_by_fine_feature_subgroup()
     # plot_by_coarse_feature_group_predictive_classifier_subgroup()
[…]

### 2022-11-13 paper/paper/SPIE/article.tex
[WIP]

@@
 utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
-collected as performance metrics the accuracy, adjusted accuracy (accuracy
-minus the proportion of the largest class), and the F1 score, and a complete
+collected as performance metrics the accuracy, adjusted accuracy (proportion
+of samples in the entire dataset largest class minus accuracy), and the F1 score, and a complete
 table with these additional metrics is available with the data for this study
@@
 \subsection{Largest mAUROCs}
+\label{sec:unfolding-importance}
 Considering the various analytic choices as tunable parameters, it makes sense
@@
 
+\subsection{Choice of Summary Metric}
+We note briefly that most of the above findings regarding the impacts of
+analytic factors, and rank ordering of feature predictive utilities, are
+similar when using the adjusted accuracy (see Section \ref{sec:multiverse}),
+instead of the mAUROC (see Figures
+S\ref{S-fig:feature-group-all-acc}-\ref{S-fig:best-params-slicing-acc}).
+However, if comparison task predictability is defined as requiring adjusted
+accuracies to be more positive than negative, then the \textit{Learning} and
+\textit{Parkinson's} datasets do not appear to be predictable with any feature
+(Figure S\ref{S-fig:feature-group-all-acc}).
+
 \section{Discussion}
+\label{sec:discussion}
 
@@
 
-However, mAUROC values observed in this study were highly sensitive to , and mean and
-median mAUROC v
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
+However, eigenfeature mAUROC values observed in this study were highly
+sensitive to the overall analytic procedure, and there was no single analytic
+choice (e.g. choice of trimming procedure, unfolding polynomial degree, number
+of preprocessing steps, or feature slicing) that ensured, for any combination
+of feature and classification task, that all mAUROC values were greater than
+0.5. In addition, the mean, median and modal mAUROCs were generally close to
+0.5, and adjusted mean and median accuracies also tended to be close to zero.
+Thus we find limited evidence that functional-connectivity-based eigenfeatures
+have general, "out of the box" predictive utility, with general
+utility likely requiring either careful tuning, or different preprocessing
+decisions and analytic choices than those examined here.
+
+\begin{table}[h!]
+\caption{
+\label{tab:top3-fine}
+Top three robust maximum (\(95^{\text{th}}\) percentile) mAUROC and adjusted
+accuracy (acc+) values for each predictable classification task and fine
+feature grouping, sorted by mAUROC. A dash indicates that the fine feature
+grouping for that row was not in the top three, i.e., that the top three
+performing features differed depending the performance metric. }
+\small
+\centering
+\begin{tabular}{llcc}
+\hline
+                                   &               & mAUROC &   acc+ \\
+Classification Task                & Source Feature&        &        \\
+\hline
+Aging - younger v older            & eigs smooth   &  0.834 &  0.226 \\
+                                   & rmt + eigs    &  0.828 &  0.212 \\
+                                   & eigs          &  0.823 &  0.211 \\
+Learning - rest v task             & rmt + eigs    &  0.638 &  0.007 \\
+                                   & eigs          &  0.636 &  0.007 \\
+                                   & eigs max      &  0.631 &    –   \\
+                                   & eigs middle   &    –   &  0.012 \\
+Osteo - nopain v duloxetine        & rmt only      &  0.819 &  0.259 \\
+                                   & eigs max      &  0.812 &  0.226 \\
+                                   & rmt + eigs    &  0.806 &    –   \\
+                                   & eigs smooth   &    –   &  0.212 \\
+Osteo - nopain v pain              & tseries loc   &  0.762 &  0.104 \\
+                                   & tseries scale &  0.697 &    –   \\
+                                   & eigs smooth   &  0.684 &    –   \\
+                                   & eigs middle   &    –   &  0.069 \\
+                                   & eigs          &    –   &  0.034 \\
+Parkinsons - ctrl v park           & tseries scale &  0.687 &  0.107 \\
+                                   & tseries loc   &  0.657 &  0.063 \\
+                                   & rmt only      &  0.590 &  0.027 \\
+TaskAttention - task\_attend v task\_nonattend     & tseries loc &  0.666 &
+0.135 \\
+                                   & rmt only      &  0.660 &  0.121 \\
+                                   & tseries scale &  0.656 &  0.124 \\
+Vigilance - vigilant v nonvigilant & eigs middle   &  0.733 &  0.184 \\
+                                   & eigs smooth   &  0.726 &    –   \\
+                                   & rmt + eigs    &  0.719 &  0.162 \\
+                                   & eigs          &    –   &  0.162 \\
+\hline
 \end{tabular}
-\end{center}
-\caption
-{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvalues (polynomial
-degree 9) for \textit{Osteo} dataset \footnotesize\texttt{duloxetine v nopain}
-classification task, brain extraction and slice time correction only. Median
-observed mAUROC = 0.643 (range = 0.346 - 0.925)}
-\end{figure}
-
-It is likely that the unfolded eigenvalues have predictive utility because of
-their smoothing and re-scaling effect. Figure \ref{fig:unfolded}, which depicts
-the raw eigenvalues and unfolded eigenvalues with different trimming
-procedures, shows how the raw eigenvalues have strongly exponential
-distributions, even with logarithmic axes. This is due to the
-magnitude of the largest eigenvalues, and the trimmed feature distributions in
-Figure \ref{fig:unfolded} are less pathologically distributed.
-
-Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier}, that
-a simple combination of the raw eigenvalues plus smoothed versions of them are the most
-generally-predictive features examined.
-
-It was also sometimes the case that adding RMT-specific features (e.g. the spectral
-rigidity, or level variance) to either the unfolded eigenvalues, or raw
-eigenvalues, would allow for a higher maximum mAUROC (Table
-\ref{tab:numerical}, Figure S\ref{S-fig:fine-largest}). One case where these
-features yielded the largest predictions was in the \textit{Aging} dataset.
-This is depicted in Figure \ref{fig:observables}, where it is clear that, after
-trimming, at almost any \(L\) value, subgroups are strongly and clearly
-visually-separable.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:observables} Rigidity and level variance features (unfolding
-polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
-older} classification task, brain extraction only. Median observed mAUROC =
-0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
-all subplots, as described in the \ref{sec:methods}.}
-\end{figure}
-
-However, the range of mAUROCs observed in all cases in this study suggests
-caution. All eigenfeatures and baselines features were sensitive to analytic
-choices such as the amount of preprocessing, feature slicing (including trimming) and
-feature smoothing / unfolding, and this sensitivity was such that all features
-resulted in mAUROCs on either side of 0.5. Most of these analytic choices are benign enough that
-it is difficult to rigorously justify one choice over another, meaning RMT-based
-analyses may be susceptible to post-hoc justification \cite{kerrHARKingHypothesizingResults1998}.
-
-It is also difficult to quantify the practical significance of the observed
-mAUROC values. With small datasets of rs- or task-fMRI data, binary
-classification using whole-brain features is challenging. For subject level
-classification, state of the art deep learning methods improve upon guessing by
-17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
+\end{table}
+
+Nevertheless, in all datasets, there were combinations of analytic choices that
+resulted in cross-validated mean prediction performances well beyond mere guessing
+(Figures S\ref{S-fig:fine-largest}, S\ref{S-fig:fine-largest-acc}, Table
+\ref{tab:top3-fine}). Whether or not these should be considered to have
+practical relevance depends on one's goals, however, we note that with small
+datasets of rs- or task-fMRI data, binary, subject-level classification using
+whole-brain features is generally challenging.
+
+For example, deep learning methods improve upon guessing by 17\% for autism
+\cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
 \cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
@@
 when the group functional connectivities are already strongly separated (e.g.
-Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}). In
-this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\%
-(\ref{S-tab:acc-numerical}).
-
-It is somewhat surprising that eigenfeatures are, on net, predictively-useful
-at all. The reduction of the functional connectivity matrix to \(T - 1\)
-eigenvalues is aggressive, and the subsequent small-degree polynomial fit used
-in the unfolding procedure further reduces information in the raw data.
-Still, the subsequent features more often than not allow non-trivial 5-fold
-cross-validated mean prediction AUROCs or accuracies.
+Cohen’s \(d > 1.0\))\cite{dansereauStatisticalPowerPrediction2017}. In this
+study, the (robust) maximum improvements upon guessing are shown in Table
+\ref{tab:top3-fine}), and vary from 3-26\%.
+
+It is somewhat surprising that the eigenfeatures examined here ever have net
+cross-validated predictive utility. The reduction of the functional
+connectivity matrix to the sorted \(T - 1\) eigenvalues destroys a large amount
+of information (e.g. radically different matrices can have identical sorted
+eigenvalues), and the subsequent small-degree polynomial fit used in the
+unfolding procedure further reduces variance in the raw data.  All brain voxels
+(including grey matter voxels) are also used with equal importance in the
+computation of the correlation matrix, and the number of subjects involved in
+each dataset is generally quite small.
+
+We speculate that the unfolded eigenvalues may have predictive utility in part
+because of their smoothing and re-scaling effect (see also Section
+\ref{sec:unfolding-importance}). Figure \ref{fig:unfolded}, which depicts the
+raw eigenvalues and unfolded eigenvalues with different trimming procedures,
+shows how the raw eigenvalues have strongly exponential distributions, even
+with logarithmic axes. This is due to the magnitude of the largest eigenvalues,
+and the unfolded and trimmed feature distribution in Figure \ref{fig:unfolded}
+are far less pathologically distributed.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:unfolded} Raw eigenvalues (first subplot) and unfolded eigenvalues
+(polynomial degree 9; last 3 subplots) for \textit{Osteo} dataset
+\footnotesize\texttt{duloxetine v nopain} classification task, brain extraction
+and slice time correction only. Median observed mAUROC = 0.643 (range = 0.346 -
+0.925)}
+\end{figure}
+
+It was also sometimes the case that adding RMT-specific features (e.g. the spectral
+rigidity, or level variance) to either the unfolded eigenvalues, or raw
+eigenvalues, would allow for a higher maximum mAUROC (Tables
+\ref{tab:numerical}-\ref{tab:top3-fine}, and Figure S\ref{S-fig:fine-largest}). One case where these
+features yielded the largest predictions was in the \textit{Aging} dataset.
+This is depicted in Figure \ref{fig:observables}, where it is clear that, after
+trimming, at almost any \(L\) value, subgroups either cluster in bands, or tend to
+generally have lower or higher rigidity of level variance curves.
+
+\begin{figure}
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
+\end{tabular}
+\end{center}
+\caption
+{ \label{fig:observables} Rigidity and level variance features (unfolding
+polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
+older} classification task, brain extraction only. Median observed mAUROC =
+0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
+all subplots, as described in Section \ref{sec:methods}. Note banding / clusters
+related to classes.}
+\end{figure}
+
+However, the range of mAUROCs observed in all cases in this study suggests
+caution. All eigenfeatures and baselines features were sensitive to analytic
+choices such as the amount of preprocessing, feature slicing (including trimming) and
+feature smoothing / unfolding, and this sensitivity was such that all features
+resulted in mAUROCs on either side of 0.5. Most of these analytic choices are benign enough that
+it is difficult to rigorously justify one choice over another, meaning RMT-based
+analyses may be susceptible to post-hoc justification \cite{kerrHARKingHypothesizingResults1998}.
 

### 2022-11-13 paper/paper/SPIE/supplementary.tex
[WIP]

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% BEGIN ACC+ FIGURES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{all_accs_by_fine_feature_groups.png}
+\end{center}
+\caption
+{ \label{fig:feature-group-all-acc} Adjusted accuracy distributions across
+fine feature groupings and comparison tasks.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_accs_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-largest-acc} Distributions of largest 500 mean adjusted
+accuracies across fine feature grouping, by comparison task. Note "rmt only"
+and "rmt + eigs" features tend to have the best possible performances across
+predictable tasks. }
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_accs_by_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-smallest-acc} Distributions of smallest 500 mean adjusted
+accuracies across fine feature groupings, by comparison task. Note "rmt only"
+and "rmt + eigs" features tend to have the worse possible performances across
+predictable tasks.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_accs_by_preproc_predictive_subgroup.png}
+\end{center}
+\caption
+{ \label{fig:fine-preproc-acc} Adjusted accuracy distributions across fine
+feature groupings and predictable comparison tasks, with effect of
+preprocessing.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{coarse_feature_group_accs_by_predictive_subgroup_classifier.png}
+\end{center}
+\caption
+{ \label{fig:coarse-classifier-acc} Adjusted accuracy distributions across
+coarse feature groupings and predictable comparison tasks, by classifier. Note
+the general similarity of each distribution within a particular classification
+task (column) and within each feature grouping.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_group_accs_by_predictive_subgroup_classifier.png}
+\end{center}
+\caption
+{ \label{fig:fine-classifier-acc} Adjusted accuracy distributions across fine
+feature groupings and predictable comparison tasks, by classifier. Note the
+general similarity of each distribution within a particular classification task
+(column) and within each feature grouping. Note also that, within a
+classification task (column), that the rank ordering of features, based on
+wither the median, mode, or mean, does not change dramatically or consistently
+from classifier to classifier. }
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_accs_by_trim.png}
+\end{center}
+\caption
+{ \label{fig:fine-trim-acc} Distributions of mean adjusted accuracies for
+unfolding-dependent RMT features, by trimming. Note the tendency for a
+rightward shift in the distributions of the features involving the
+unfolded eigenvalues when using largest or middle trimming (most dramatic in
+the Osteo nopain v duloxetine condition). The impact of these trimming methods
+on the rigidity and level variance features, however, was mixed (compare
+Vigilance data to Osteo nopain v duloxetine condition.)}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_accs_by_degree.png}
+\end{center}
+\caption
+{ \label{fig:fine-degree-acc}
+Distributions of mean adjusted accuracies for unfolding-dependent RMT features, by degree.}
+\end{figure}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_eigs_accs_by_subgroup_and_slicing.png}
+\end{center}
+\caption
+{ \label{fig:best-params-slicing-acc} Distributions of mean adjusted accuracies
+by slicing. Features involving the full spectrum (raw eigenvalues, smoothed
+eigenvalues, and rmt + eigs) sometimes have most positive adjusted accuracy
+distributions when using the larger eigenfeature values (first two columns)
+or middle values (Osteo nopain v pain condition, Vigilance classification
+task).}
+\end{figure}
+
 \begin{table}[H]
-\caption{Numerical summaries of feature mean accuracy difference from guess across predictable comparisons, and all combinations of
-analytic choices, sorted by 95\% percentile (robust max) value.}
+\caption{Numerical summaries of feature mean accuracy difference from guess
+across predictable comparisons, and all combinations of analytic choices,
+sorted by 95\% percentile (robust max) value.}
 \label{tab:acc-numerical}

### 2022-11-13 code/scripts/summarize_everything_updated.py

@@ -3011,8 +3011,9 @@ if __name__ == "__main__":
     # df_ses = load_combined(drop_ses=False)
     # df.to_json(PROJECT / "EVERYTHING.json")
     # print(f"Saved all combined data to {PROJECT / 'EVERYTHING.json'}")
-    make_kde_plots()
-    sys.exit()
+
+    # make_kde_plots()
+    # sys.exit()
 
     df = load_combined()
     feature_summary = (
@@ -3025,6 +3026,27 @@ if __name__ == "__main__":
         .round(3)
     )
     print(feature_summary.to_markdown(tablefmt="simple"))
+
+    preds = df[df.subgroup.isin(OVERALL_PREDICTIVE_GROUP_ORDER)].copy()
+    largest_aurocs = (
+        preds.groupby(["subgroup", "fine_feature"])[["auroc"]]
+        .quantile(0.95)
+        .groupby("subgroup")
+        .apply(lambda grp: grp.nlargest(3, columns=["auroc"]))
+        .droplevel(1)
+    )
+    largest_accs = (
+        preds.groupby(["subgroup", "fine_feature"])[["acc+"]]
+        .quantile(0.95)
+        .groupby("subgroup")
+        .apply(lambda grp: grp.nlargest(3, columns=["acc+"]))
+        .droplevel(1)
+    )
+    largests = pd.concat([largest_aurocs, largest_accs], axis=1)
+    largests = largests.sort_values(
+        by=["subgroup", "auroc", "acc+"], ascending=[True, False, False]
+    )
+    print(largests.round(3).to_latex())
 
     # plot_unfolded_duloxetine()
     # plot_unfolded_aging()

### 2022-11-13 paper/paper/SPIE/article.tex
[WIP]

@@
 \usepackage{setspace}
+\usepackage{hyperref}
 \usepackage{tocloft}
@@
 \section*{Conclusions}
-Eigenfeatures clearly have potential for understanding and reducing fMRI
-functional connectivity. However, the utility of these features is strongly
-dependent on analytic decisions, suggesting considerable caution when
+Eigenfeatures clearly have potential for understanding fMRI functional
+connectivity in a wide variety of scenarios. However, the utility of these
+features is strongly dependent on analytic decisions, suggesting caution when
 interpreting past and future studies applying RMT to fMRI.
@@
 veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
-even more recently, has provided insights into the behaviors of deep
+more recently, has provided insights into the behaviors of deep
 neural networks\cite{martinImplicitSelfRegularizationDeep2021,
@@
 
-The functional connectivity—the matrix of correlations of
+The functional connectivity—often, the matrix of correlations of
 various collections of voxels of an fMRI scan—is \textit{a priori} a useful
@@
 }
-\label{table:1}
+\label{tab:data-dimensions}
 \small
@@
 aspects of each dataset here.  Dataset scan parameters and acquisition details
-are summarized in Table \ref{table:1}, and sample and subgroup sizes are
-summarized in Table \ref{table:2}.
+are summarized in Table \ref{tab:data-dimensions}, and sample and subgroup sizes are
+summarized in Table \ref{tab:data-subjects}.
 
@@
 }
-\label{table:2}
+\label{tab:data-subjects}
 \small
@@
 The time and space complexities of computing the full \(N \times N\)
-correlation matrix of \(\mathbf{M}\) (and subsequent eigenvalues) are too
+\(\mathbf{M}\) of Pearson correlation coefficients (and subsequent eigenvalues) are too
 large to be tractable. However, since transposition does not change the
@@
 elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
-\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
+\dots;\; \hat{f}_n]\).
+% \footnote{Exact implementation can be found in the
+%\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
 
@@
 (e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
-Exact details can be found in the
-\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
-source code}.
+% Exact details can be found in the
+% \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
+% source code}.
 
@@
 utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
-collected as performance metrics the accuracy, adjusted accuracy (proportion
-of samples in the entire dataset largest class minus accuracy), and the F1 score, and a complete
-table with these additional metrics is available with the data for this study
-(Section \ref{sec:data-and-code}).
-
-Each non-baseline feature is evaluated, for each classification task and
+collected as performance metrics the accuracy, adjusted accuracy (proportion of
+samples in the entire dataset largest class minus accuracy), and the F1 score.
+A complete table with these additional metrics is available with the
+\hyperref[sec:data-and-code]{data for this study}.
+
+Each non-baseline feature is thus evaluated, for each classification task and
 classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\) 2
@@
 
-Summarizing this quantity of data requires some caution. Measures of location or
-scale, even when robust, are, for the most part, misleading and uninformative
-when distributions are skewed. In most cases we find skewed mAUROC
-distributions, and so have chosen primarily to present our findings visually,
-with kernel density estimates.
-
-Because aggregating across datasets with different class imbalances can
-be extremely misleading, even with a metric like the mAUROC\cite{brabecModelEvaluationNonconstant2020},
-we restrict such summaries to Table \ref{tab:numerical} only, and note that
-the table is intended as a \textit{supplement only} to the distribution plots: no conclusions
-should be drawn about feature performance from the table values alone.
+Summarizing this quantity of data requires some caution. Measures of location
+or scale, even when robust, are, for the most part, misleading and
+uninformative when distributions are skewed. In most cases we find skewed
+mAUROC distributions, and so have chosen primarily to present our findings
+visually, with kernel density estimates.  Because aggregating across datasets
+with different class imbalances can be extremely misleading, even with a metric
+like the mAUROC\cite{brabecModelEvaluationNonconstant2020}, we restrict such
+summaries to Table \ref{tab:numerical} only, and note that the table is
+intended as a \textit{supplement only} to the distribution plots: no
+conclusions should be drawn about feature performance from the table values
+alone.
 
@@
 those using the unfolded eigenvalues alone (Figures
-S\ref{S-fig:fine-trim}-\ref{S-fig:fine-degree}). The mAUROC distributions of
-these features differed most strongly in the tails (Figures
+S\ref{S-fig:fine-trim}-\ref{S-fig:fine-degree}). Instead, the mAUROC
+distributions of these features differed mostly in the tails (Figures
 S\ref{S-fig:fine-largest}-\ref{S-fig:fine-smallest}).
@@
 of preprocessing steps, or feature slicing) that ensured, for any combination
-of feature and classification task, that all mAUROC values were greater than
-0.5. In addition, the mean, median and modal mAUROCs were generally close to
-0.5, and adjusted mean and median accuracies also tended to be close to zero.
-Thus we find limited evidence that functional-connectivity-based eigenfeatures
-have general, "out of the box" predictive utility, with general
-utility likely requiring either careful tuning, or different preprocessing
-decisions and analytic choices than those examined here.
+of feature and classification task, that all all other analytic choices
+resulted in mAUROC values greater than 0.5. In addition, the mean, median and
+modal mAUROCs were generally close to 0.5, and adjusted mean and median
+accuracies also tended to be close to zero. Thus we find limited evidence that
+functional-connectivity-based eigenfeatures have general, "out of the box"
+predictive utility, with general utility likely requiring either careful
+tuning, or different preprocessing decisions and analytic choices than those
+examined here.
 
@@
 cross-validated predictive utility. The reduction of the functional
-connectivity matrix to the sorted \(T - 1\) eigenvalues destroys a large amount
-of information (e.g. radically different matrices can have identical sorted
-eigenvalues), and the subsequent small-degree polynomial fit used in the
-unfolding procedure further reduces variance in the raw data.  All brain voxels
-(including grey matter voxels) are also used with equal importance in the
-computation of the correlation matrix, and the number of subjects involved in
-each dataset is generally quite small.
+connectivity matrix to the sorted \(T - 1\) eigenvalues uses all brain voxels
+(including grey matter voxels), and destroys a large amount of information
+(radically different matrices can have identical sorted eigenvalues). The
+subsequent small-degree polynomial fit used in the unfolding procedure further
+reduces variance in the raw data, and all eigenfeatures, due to the eigenvalue
+sorting, are monotonically increasing curves (or approximately monotonic). All
+such curves could likely be fit near-perfectly with 3-5 parameters, i.e. the
+inherent dimensionality of these features is quite small. Interpreting the eigenvalues
+as the magnitudes of the principal components of the standardized data, this
+suggests that a rough summary of the magnitudes of the principal components can
+often be surprisingly predictive in fMRI.
 
@@
 
-It was also sometimes the case that adding RMT-specific features (e.g. the spectral
-rigidity, or level variance) to either the unfolded eigenvalues, or raw
-eigenvalues, would allow for a higher maximum mAUROC (Tables
-\ref{tab:numerical}-\ref{tab:top3-fine}, and Figure S\ref{S-fig:fine-largest}). One case where these
-features yielded the largest predictions was in the \textit{Aging} dataset.
-This is depicted in Figure \ref{fig:observables}, where it is clear that, after
-trimming, at almost any \(L\) value, subgroups either cluster in bands, or tend to
-generally have lower or higher rigidity of level variance curves.
-
-\begin{figure}
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
-\end{tabular}
-\end{center}
-\caption
-{ \label{fig:observables} Rigidity and level variance features (unfolding
-polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
-older} classification task, brain extraction only. Median observed mAUROC =
-0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
-all subplots, as described in Section \ref{sec:methods}. Note banding / clusters
-related to classes.}
-\end{figure}
-
-However, the range of mAUROCs observed in all cases in this study suggests
-caution. All eigenfeatures and baselines features were sensitive to analytic
-choices such as the amount of preprocessing, feature slicing (including trimming) and
-feature smoothing / unfolding, and this sensitivity was such that all features
-resulted in mAUROCs on either side of 0.5. Most of these analytic choices are benign enough that
-it is difficult to rigorously justify one choice over another, meaning RMT-based
-analyses may be susceptible to post-hoc justification \cite{kerrHARKingHypothesizingResults1998}.
-
-We believe that future studies using RMT to inform their examinations be more
-careful. In the absence of a cross-validated predictive framework that takes
-into account the strong impact of researcher analytic choices, any apparent
-explanatory utility of RMT should be treated with skepticism. RMT deals with
-highly-ideal \textit{iid} matrices, but almost no real-world systems involve
-\textit{iid} elements. The basic RMT insight that eigenvalues can be used to
-describe matrix-representable systems is certainly vindicated here, but the use
-of mathematically non-trivial RMT statistics (like the spectral rigidity or
-level number variance) should not be used if simpler statistical descriptions
-of the eigenvalues are equally informative.
+% It was also sometimes the case that adding RMT-specific features (e.g. the spectral
+% rigidity, or level variance) to either the unfolded or raw
+% eigenvalues would allow for a higher maximum mAUROC (Tables
+% \ref{tab:numerical}-\ref{tab:top3-fine}, and Figure S\ref{S-fig:fine-largest}). One case where these
+% features yielded the largest predictions was in the \textit{Aging} dataset.
+% This is depicted in Figure \ref{fig:observables}, where it is clear that, after
+% trimming, at almost any \(L\) value, subgroups either cluster in bands, or tend to
+% generally have lower or higher rigidity of level variance curves.
+
+% \begin{figure}
+% \begin{center}
+% \begin{tabular}{c}
+% \includegraphics[width=6.5in]{observables_older_younger_v_older.png}
+% \end{tabular}
+% \end{center}
+% \caption
+% { \label{fig:observables} Rigidity and level variance features (unfolding
+% polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
+% older} classification task, brain extraction only. Median observed mAUROC =
+% 0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
+% all subplots, as described in Section \ref{sec:methods}. Note banding / clusters
+% related to classes.}
+% \end{figure}
+
+\subsection{Limitations}
+As is unfortunately typical of fMRI research\cite{turnerSmallSampleSizes2018},
+the number of subjects in each dataset was quite small (Table
+\ref{tab:data-subjects}). With such small numbers of subjects, randomization
+cannot be expected to effectively control group differences, and it is possible
+the predictive utility of the eigenfeatures was due to capitalization of such
+differences. For example, eigenfeatures were generally predictive in the
+\textit{Aging} dataset, and it is quite possible for randomization failures to
+introduce age confounding across classes.
+
+Likewise, one of the other more predictable classification tasks was the
+\code{duloxetine v nopain} task in the osteoarthritis dataset. An active drug
+could introduce any number of physiological confounds relevant to
+fMRI\cite{murphyRestingstateFMRIConfounds2013}, but we could not control for
+such effects do to the absence of physiological recording in most datasets.
+
+In general, much larger fMRI datasets are needed to adequately control for and test
+\textit{what} exactly the connectivity eigenvalues actually predict, and to test
+if these predictions generalize to larger, different populations.
 
@@
 
+Eigenvalue-based features inspired by RMT, and extracted from fMRI functional
+connectivity matrices were found to have predictive utility across a wide
+variety of datasets and classification tasks. However, the predictive utilities
+were modest, and highly dependent on preprocessing steps and other fitting and
+feature selection procedures. Future studies employing RMT should carefully
+investigate the sensitivity of any findings to such analytic decisions before
+drawing any conclusions.
+
 \section{Disclosures}
-
-\section{Acknowledgments and Funding}
-
-\section{Data, Materials, and Code Availability}
+The authors have no relevant financial interests in the manuscript and no other
+potential conflicts of interest to disclose.
+
+% Conflicts of interest should be declared under a separate header, above
+% Acknowledgments. If the authors have no competing interests to declare, then a
+% statement should be included declaring no conflicts of interest. For assistance
+% generating a disclosure statement, see the form available from  the
+% International Committee of Medical Journal Editors website:
+% \linkable{http://www.icmje.org/conflicts-of-interest/}
+
+\section*{Acknowledgments and Funding}
+This work was supported by the Natural Science and Engineering Research
+Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
+Natural Science and Engineering Research Council of Canada Discovery Grant to
+JL, a Canada Foundation for Innovation and Nova Scotia Research and Innovation
+Trust infrastructure grant (R0176004) to JL, a St. Francis Xavier University
+research startup grant to JL (grant number R0168020), and a St. Francis Xavier
+University UCR grant to JL.
+
+\section*{Data, Materials, and Code Availability}
 \label{sec:data-and-code}
@@
 
-\subsection{Disclosures}
-Conflicts of interest should be declared under a separate header, above
-Acknowledgments. If the authors have no competing interests to declare, then a
-statement should be included declaring no conflicts of interest. For assistance
-generating a disclosure statement, see the form available from  the
-International Committee of Medical Journal Editors website:
-\linkable{http://www.icmje.org/conflicts-of-interest/}
-
-\subsection{Acknowledgments}
-Acknowledgments and funding information should be added after the conclusion,
-and before references. Include grant numbers and the full name of the funding
-body. The acknowledgments section does not have a section number.
-
-\subsection{Data, Materials, and Code Availability}
-Availability of data, materials, and/or code used in the research results
-reported in the manuscript may be declared under the heading ``Data, Materials,
-and Code Availability,'' following the Acknowledgments section. As relevant,
-provide specific access information or restrictions for data, materials, and
-computer code (i.e., links to repository access addresses with guidance on
-commercial or public access).
+% \subsection{Disclosures}
+% Conflicts of interest should be declared under a separate header, above
+% Acknowledgments. If the authors have no competing interests to declare, then a
+% statement should be included declaring no conflicts of interest. For assistance
+% generating a disclosure statement, see the form available from  the
+% International Committee of Medical Journal Editors website:
+% \linkable{http://www.icmje.org/conflicts-of-interest/}
+
+% \subsection{Acknowledgments}
+% Acknowledgments and funding information should be added after the conclusion,
+% and before references. Include grant numbers and the full name of the funding
+% body. The acknowledgments section does not have a section number.
+
+% \subsection{Data, Materials, and Code Availability}
+% Availability of data, materials, and/or code used in the research results
+% reported in the manuscript may be declared under the heading ``Data, Materials,
+% and Code Availability,'' following the Acknowledgments section. As relevant,
+% provide specific access information or restrictions for data, materials, and
+% computer code (i.e., links to repository access addresses with guidance on
+% commercial or public access).
 
@@
 
-\subsubsection{Reference linking and DOIs}
-A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a
-digital object, such as a journal article or a book chapter, that provides a
-persistent link to its location on the internet. The use of DOIs allows readers
-to easily access cited articles. Authors should include the DOI at the end of
-each reference in brackets if a DOI is available. See examples at the end of
-this manuscript. A free DOI lookup service is available from CrossRef at
-\\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will
-facilitate reference linking and is highly recommended.
-
-In the present LaTeX template, the author needs to add the DOI reference by
-including it in a ``note'' in the bibliography file, as shown in the file
-{\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
-The DOI may be used by the reader to locate that document with the link:
-{\verb+http://dx.doi.org10.1117/12.154577+}.
-
-\subsection{Biographies}
-A brief professional biography of approximately 75 words may be provided for
-each author, if available. Biographies should be placed at the end of the
-paper, after the references. Personal information such as hobbies or
-birthplace/birthdate should not be included. Author photographs are not
-published.
+% \subsubsection{Reference linking and DOIs}
+% A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a
+% digital object, such as a journal article or a book chapter, that provides a
+% persistent link to its location on the internet. The use of DOIs allows readers
+% to easily access cited articles. Authors should include the DOI at the end of
+% each reference in brackets if a DOI is available. See examples at the end of
+% this manuscript. A free DOI lookup service is available from CrossRef at
+% \\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will
+% facilitate reference linking and is highly recommended.
+
+% In the present LaTeX template, the author needs to add the DOI reference by
+% including it in a ``note'' in the bibliography file, as shown in the file
+% {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
+% The DOI may be used by the reader to locate that document with the link:
+% {\verb+http://dx.doi.org10.1117/12.154577+}.
+
+% \subsection{Biographies}
+% A brief professional biography of approximately 75 words may be provided for
+% each author, if available. Biographies should be placed at the end of the
+% paper, after the references. Personal information such as hobbies or
+% birthplace/birthdate should not be included. Author photographs are not
+% published.
 
@@
 
-Let \(\mathbf{X}\) is a real \(n \times p\) matrix with \(n > p\). Let
+Let \(\mathbf{X}\) be a real \(n \times p\) matrix with \(n > p\). Let
 \[
@@
 \[
-\mathbf{Y} = \text{stand}(\mathbf{X}) =
+\mathbf{Y} = \text{standardize}(\mathbf{X}) =
 \left(
@@
 % \disclosures
-\subsection*{Disclosures}
-Conflicts of interest should be declared under a separate header.
+% \subsection*{Disclosures}
+% Conflicts of interest should be declared under a separate header.
 If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.
 
-\subsection* {Acknowledgments}
-This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.
+% \subsection* {Acknowledgments}
+% This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.
 
@@
 
-\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering.
+% \vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering.
 He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.

### 2022-11-14 paper/paper/SPIE/cover_letter.tex
finalize

--- 2021-11-17 paper/latex/PMB/cover_letter.tex
@@
 \usepackage{graphicx}
-\graphicspath{ {figs/} }
+\graphicspath{ {figures/} }
 \usepackage{geometry}
@@
 ]{biblatex}
-\addbibresource{cover.bib}
+\addbibresource{report.bib}
 \usepackage{fancyhdr}
@@
 \today, \\
-To the Editors of Network Neuroscience, \\
+Dear Dr. Giger, \\
 
-We hope you will consider our manuscript \emph{Random Matrix Theory Tools for the Analysis of
-Functional Magnetic Resonance Imaging Examinations} for publication in Network Neuroscience as a
-\emph{Research} article. In it, we apply mathematical tools from random matrix theory (RMT) to
-analyze a wide variety of functional magnetic resonance imaging (fMRI) data, and release
-open-source software \citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} to facilitate such analyses with a wide variety of data.
+We hope you will consider our manuscript \emph{Random Matrix Theory Tools for
+the Predictive Analysis of Functional Magnetic Resonance Imaging Examinations}
+for publication in SPIE Journal of Medical Imaging.
 
-Numerous fMRI studies examine correlations amongst regions of interest (ROI) in the brain. These
-``functional connectivity'' analyses typically use ROIs or seed regions in order to simplify
-connectivity computations \citep{huettelFunctionalMagneticResonance2004}. Instead, we use
-statistics from RMT to summarize \emph{all} voxelwise correlations with no reductions. In
-addition, rather than simply reporting findings which are statistically significant, we report all
-analyses from a cross-validated predictive framework, where multiple machine learning models are
-compared against both each other and proper baselines (e.g. naive / dummy classifiers).
+We confirm that this work is original and has not been published elsewhere nor
+is it currently under consideration for publication elsewhere.
 
-We show that, across a variety of data and with fairly small data, and without any model tuning,
-that these RMT reductions have real predictive utility and potential. This suggests RMT could
-be a powerful and efficient method for summarizing functional connectivity information. As much of the
-research in network neuroscience involves fMRI, functional connectivity, and statistical physics, as
-per the \href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope} of \emph{Network
-Neuroscience}, we believe our manuscript to be an excellent fit for submission to your journal.
+In this paper, we extract a number of eigenvalue-based features inspired by
+Random Matrix Theory (RMT) from a wide variety of functional magnetic resonance
+imaging (fMRI) data. We subsequently examine the cross-validated predictive
+utility of those extracted features on a number of binary classification tasks
+that can be constructed for each dataset. The predictive utility is examined
+systematically, in a  multiverse-style analysis
+\citep{steegenIncreasingTransparencyMultiverse2016}, which reveals the
+sensitivity of results to various typical preprocessing decisions involved in
+fMRI and RMT.
 
-Should you consider to review our manuscript, we have suggested a number of potential reviewers on
-your submission website that we believe would be suitable due to expertise in RMT and similar
-interdisciplinary research. This paper is not under consideration at any other journa, although a
-poster of some preliminary results of our study was presented at the International Symposium for
-Magnetic Resonance in Medicine 2021 annual conference \citep{bergerOpenSourceRandom2021}. Our lab
-has REB ethics approval for conducting retrospective analyses (although we are unaware of the need,
-this being an analysis of anonymized public data). All authors approve the manuscript in its
-submitted form, and have no conflicts of interest, financial or otherwise.
+This paper is most unique in that it:
+
+\begin{enumerate}
+  \item employs RMT to generate reductions of the full voxelwise fMRI
+        correlation matrix
+  \item evaluates RMT-related metrics in a cross-validated predictive
+        framework, instead of the more typical descriptive or explanatory
+        frameworks
+  \item systematically investigates the impact of the various methodological
+        decisions involved in the findings
+  \item makes all code used in the paper publicly available, including the release of a
+        separate, general, open-source RMT
+        library \citep{dm-bergerStfxecutablesEmpyricalRMTV12022} that will allow
+        RMT metrics to be easily and reproducibly computed on a wide variety of
+        data
+\end{enumerate}
+
+Please address all correspondence concerning this manuscript to me at \href{jlevman@stfx.ca}{jlevman@stfx.ca}.
+
+Thank you for your consideration of this manuscript.
+
+% apply mathematical tools from random matrix theory (RMT) to analyze
+
+% , and release
+% open-source software \citep{bergerStfxecutablesEmpyricalRMTPreliminary2020} to
+% facilitate such analyses with a wide variety of data.
+
+% Numerous fMRI studies examine correlations amongst regions of interest (ROI) in the brain. These
+% ``functional connectivity'' analyses typically use ROIs or seed regions in order to simplify
+% connectivity computations \citep{huettelFunctionalMagneticResonance2004}. Instead, we use
+% statistics from RMT to summarize \emph{all} voxelwise correlations with no reductions. In
+% addition, rather than simply reporting findings which are statistically significant, we report all
+% analyses from a cross-validated predictive framework, where multiple machine learning models are
+% compared against both each other and proper baselines (e.g. naive / dummy classifiers).
+
+% We show that, across a variety of data and with fairly small data, and without any model tuning,
+% that these RMT reductions have real predictive utility and potential. This suggests RMT could
+% be a powerful and efficient method for summarizing functional connectivity information. As much of the
+% research in network neuroscience involves fMRI, functional connectivity, and statistical physics, as
+% per the \href{https://direct.mit.edu/netn/pages/submission-guidelines#aims}{Aims and Scope} of \emph{Network
+% Neuroscience}, we believe our manuscript to be an excellent fit for submission to your journal.
+
+% Should you consider to review our manuscript, we have suggested a number of potential reviewers on
+% your submission website that we believe would be suitable due to expertise in RMT and similar
+% interdisciplinary research. This paper is not under consideration at any other journa, although a
+% poster of some preliminary results of our study was presented at the International Symposium for
+% Magnetic Resonance in Medicine 2021 annual conference \citep{bergerOpenSourceRandom2021}. Our lab
+% has REB ethics approval for conducting retrospective analyses (although we are unaware of the need,
+% this being an analysis of anonymized public data). All authors approve the manuscript in its
+% submitted form, and have no conflicts of interest, financial or otherwise.
 
@@
 Jacob Levman, PhD (jlevman@stfx.ca) \\
-Canada Research Chair in Bioinformatics \\
+% Canada Research Chair in Bioinformatics \\
 Department of Computer Science \\

### 2022-11-14 code/scripts/summarize_everything_updated.py
autoformat

@@ -6,38 +6,28 @@ sys.path.append(str(ROOT))
 # fmt: on
 
 import re
+from shutil import copyfile
 from typing import Literal
 from warnings import simplefilter
 
 import matplotlib.pyplot as plt
-from numpy import ndarray
 import numpy as np
 import pandas as pd
 import seaborn as sbn
-from numba import njit
-from shutil import copyfile
-from warnings import simplefilter
 from joblib import Memory
 from matplotlib.axes import Axes
 from matplotlib.figure import Figure
-from matplotlib.lines import Line2D
 from matplotlib.patches import Patch
+from numba import njit
+from numpy import ndarray
 from pandas import DataFrame
 from pandas.errors import PerformanceWarning
-from scipy.optimize import NonlinearConstraint, minimize, minimize_scalar
 from seaborn import FacetGrid
-from statsmodels.distributions.empirical_distribution import ECDF
 from tqdm import tqdm
 
-from rmt.updated_dataset import UpdatedProcessedDataset
-from rmt.enumerables import UpdatedDataset, PreprocLevel, TrimMethod
-from rmt.updated_features import (
-    FEATURE_OUTFILES as PATHS,
-    Unfolded,
-    Eigenvalues,
-    Rigidities,
-    Levelvars,
-)
+from rmt.enumerables import PreprocLevel, TrimMethod, UpdatedDataset
+from rmt.updated_features import FEATURE_OUTFILES as PATHS
+from rmt.updated_features import Eigenvalues, Levelvars, Rigidities, Unfolded
 from rmt.visualize import UPDATED_PLOT_OUTDIR as PLOT_OUTDIR
 from rmt.visualize import best_rect
 
@@ -2685,7 +2675,7 @@ def summary_stats_and_tables() -> None:
     print("Proportion of AUROCs greater than 0.5 (Including Ses)")
     print(g05_ses.describe().T)
 
-    p90 = df.groupby(["subgroup", "mega_feature"])["auroc"].quantile(0.90).unstack()
+    df.groupby(["subgroup", "mega_feature"])["auroc"].quantile(0.90).unstack()
     p90_ses = (
         df_ses.groupby(["subgroup", "mega_feature"])["auroc"].quantile(0.90).unstack()
     )
@@ -2790,7 +2780,6 @@ def plot_unfolded(
         preproc=preproc,
         norm=True,
     )
-    deg = 9
[…]

### 2022-11-14 code/scripts/summarize_everything_updated.py
delint

@@ -59,6 +59,8 @@ def s_xlim(
             return (-0.1, 0.3)
         else:
             return (-0.4, 0.1)
+    else:
+        raise NotImplementedError()
 
 
 PROJECT = ROOT.parent
@@ -247,7 +249,7 @@ AGGREGATES = get_aggregates(SUBGROUPERS)
 def topk_outdir(k: int) -> Path:
     outdir = PLOT_OUTDIR / f"top-{k}_plots"
     outdir.mkdir(exist_ok=True, parents=True)
-    return outdir
+    return Path(outdir)
 
 
 def corr_renamer(s: str) -> str:
@@ -438,6 +440,7 @@ def fine_feature_grouping(s: str) -> str:
         return "rmt only"
     if is_rmt_plus(s):
         return "rmt + eigs"
+    raise ValueError(f"Invalid feature: {s}")
 
 
 def coarse_feature_grouping(s: str) -> str:
@@ -752,38 +755,6 @@ def summarize_performance_by_aggregation(
     resize_fig()
     fig.subplots_adjust(top=0.93, right=0.945, bottom=0.105)
     plt.show()
-    # """
-
-    return
-    """
-    # The more levels we include, the less we "generalize" our claims
-    for grouper, aggregates in zip(SUBGROUPERS, AGGREGATES):
-        min_summarized_per_feature = np.unique(df.groupby(grouper + ["feature"]).count())[
-            0
-        ]
-        grouped = df.groupby(grouper)
-        # if you do:
-        #   grouped.hist()
-        # here, you get a plot for each sub-frame induced by "grouper"
-        # I think if you reset_index, or just leave it as is, seaborn is also
-        # going to be able to do a decent job here
-        if summarizer == "median":
-            summary = df.groupby(grouper + ["feature"]).median(numeric_only=True)
-
-        else:
-            summary = df.groupby(grouper + ["feature"]).max(numeric_only=True)
-
-        summary.hist(color="black")
-        plt.gcf().suptitle(f"Grouping by {grouper}\n(aggregating across: {aggregates})")
-        plt.show()
-
-        print()
-        # bests = (
-        #     summary.reset_index()
-        #     .groupby(grouper)
-        #     .apply(lambda g: g.nlargest(k, columns="auroc"))
-        # )
-    """
 
 
 def plot_topk_features_by_aggregation(sorter: str, k: int = 5) -> None:
[…]

### 2022-11-15 paper/paper/SPIE/article.tex
finalize
tidy up and make corrections to paper

@@
 
-\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}
+\title{Random Matrix Theory Tools for the Predictive Analysis of Functional Magnetic Resonance Imaging Examinations}
 
 \author[a]{Derek Berger}
-\author[a,*]{Jacob Levman}
-\author[b]{Gurpreet M. Matharoo}
-\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
-\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
+\author[b,c]{Gurpreet S. Matharoo}
+\author[a*,d,e,f]{Jacob Levman}
+
+\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+\affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+\affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
+\affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
+\affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}
 
@@
 
-% \begin{abstract} This document shows the required format and appearance of a
-% manuscript prepared for SPIE journals. It is prepared using LaTeX2e with the
-% class file \texttt{spieman.cls}. Please note that the following journals
-% require the use of structured abstracts in manuscript submissions:
-% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the
-% \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for
-% the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Guidelines
-% are available on the journal website. Whether structured or single-paragraph,
-% the abstract should be a summary of the paper and not an introduction.
-% Because the abstract may be used in abstracting and indexing databases, it
-% should be self-contained (i.e., no numerical references) and substantive in
-% nature, presenting concisely the objectives, methodology used, results
-% obtained, and their significance. A list of up to six keywords should
-% immediately follow. \end{abstract}
-
-% See
-% https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
-% for info on the structured abstract
-%
-% Purpose: This section presents the significance and aims, stating the broad
-% impact and the rationale or motivation of the work, including potentially
-% some background.
-%
-% Approach: This section briefly describes the materials and methods, including
-% the statistical analyses, used in the research.
-%
-% Results: This section provides a core summary of the findings from the work,
-% including study numbers, quantitative analyses, or discoveries.
-%
-% Conclusions: This section summarizes and interprets the approach and findings
-% from the work, stating also the importance or impact of the findings.
-
-% Purpose: Data-intensive modeling could provide insight on the broad
-% variability in outcomes in spine surgery. Previous studies were limited to
-% analysis of demographic and clinical characteristics. We report an analytic
-% framework called “SpineCloud” that incorporates quantitative features
-% extracted from perioperative images to predict spine surgery outcome.
-
-% Approach: A retrospective study was conducted in which patient demographics,
-% imaging, and outcome data were collected. Image features were automatically
-% computed from perioperative CT. Postoperative 3- and 12-month functional and
-% pain outcomes were analyzed in terms of improvement relative to the
-% preoperative state. A boosted decision tree classifier was trained to predict
-% outcome using demographic and image features as predictor variables.
-% Predictions were computed based on SpineCloud and conventional demographic
-% models, and features associated with poor outcome were identified from
-% weighting terms evident in the boosted tree.
-
-% Results Neither approach was predictive of 3- or 12-month outcomes based on
-% preoperative data alone in the current, preliminary study. However,
-% SpineCloud predictions incorporating image features obtained during and
-% immediately following surgery (i.e., intraoperative and immediate
-% postoperative images) exhibited significant improvement in area under the
-% receiver operating characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3
-% months and AUC=0.69 (CI95=0.55 to 0.82) at 12 months.
-
-% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved
-% by incorporation of image-based features compared to analysis based on
-% conventional demographic data. The SpineCloud framework could improve
-% understanding of factors underlying outcome variability and warrants further
-% investigation and validation in a larger patient cohort.
 \begin{abstract}
@@
 task, classifier, and feature. To deal with class imbalance, we use the
-area under the receiver operating characteristic (AUROC) as the main
+area under the receiver operating characteristic curve (AUROC) as the main
 performance metric.
@@
 Eigenfeatures clearly have potential for understanding fMRI functional
-connectivity in a wide variety of scenarios. However, the utility of these
-features is strongly dependent on analytic decisions, suggesting caution when
-interpreting past and future studies applying RMT to fMRI.
-
-% This document shows the required format and appearance of a manuscript
-% prepared for SPIE journals. It is prepared using LaTeX2e with the class file
-% \texttt{spieman.cls}. Please note that the following journals require the use
-% of structured abstracts in manuscript submissions: \textit{Neurophotonics},
-% the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical
-% Imaging}. Structured abstracts are encouraged for the \textit{Journal of
-% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the
-% journal website. Whether structured or single-paragraph, the abstract should
-% be a summary of the paper and not an introduction. Because the abstract may
-% be used in abstracting and indexing databases, it should be self-contained
-% (i.e., no numerical references) and substantive in nature, presenting
-% concisely the objectives, methodology used, results obtained, and their
-% significance. A list of up to six keywords should immediately follow.
+connectivity in a wide variety of scenarios. The utility of these features is
+strongly dependent on analytic decisions, suggesting caution when interpreting
+past and future studies applying RMT to fMRI. However, this study demonstrates
+that the inclusion of RMT statistics in fMRI investigations could improve
+prediction performances across a wide variety of phenomena.
+
 \end{abstract}
 
-% Include a list of up to six keywords after the abstract
 \keywords{random matrix, spectral rigidity, level number variance, fMRI, classification, machine-learning}
 
-% Include email contact information for corresponding author
-{\noindent \footnotesize\textbf{*}Jacob Levman,  \linkable{jlevman@stfx.ca} }
-
-\begin{spacing}{2}   % use double spacing for rest of manuscript
+{
+\noindent
+\footnotesize
+\textbf{*}Jacob Levman,  \linkable{jlevman@stfx.ca}
+(primary), \linkable{jlevman@mgh.harvard.edu}
+}
+
+\begin{spacing}{2}
 
 \section{Introduction}
-\label{sect:intro}  % \label{} allows reference to this section
-% This document shows the format and appearance of a manuscript prepared for
-% submission to an SPIE journal. Note that this template is only intended to be
-% used as a guideline for author convenience. It is designed for optimum
-% clarity and ease of reading for editors and reviewers, but the template does
-% not reflect the final page layout of a published journal paper. Accepted
-% papers are professionally typeset in XML according to the layout and design
-% of the journal.
+\label{sect:intro}
 
@@
 physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
-RMT has been shown have extremely broad potential.
-
-In small-scale physical systems, RMT universalities have been observed in
-quantum chaotic systems, complex nuclei, atoms, molecules and disordered
-mesoscopic systems
-\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
-and at larger scales, RMT has been applied to atmospheric physics
+RMT has been shown to have extremely broad potential. In small-scale physical
+systems, RMT universalities have been observed in quantum chaotic systems,
+complex nuclei, atoms, molecules and disordered mesoscopic systems
+\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004,
+brodyRandommatrixPhysicsSpectrum1981, beenakkerRandommatrixTheoryQuantum1997,
+bohigasHigherOrderCorrelationsSpectra1985, wintgenLevelStatisticsQuantized1988,
+pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has
+been applied to atmospheric physics
 \cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
-\cite{plerouRandomMatrixApproach2002} , social networks
+\cite{plerouRandomMatrixApproach2002}, social networks
 \cite{jalanUncoveringRandomnessSuccess2014}, random networks
 \cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
-liquids
-\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}
-and amorphous clusters
-\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
-Within biological systems, RMT has also been used to successfully model aspects
-of  amino acid functional relationships
-\cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in epileptic seizures
-\cite{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein
-interactions both in different species
+liquids\cite{sastrySpectralStatisticsInstantaneous2001,
+matharooSpectralStatisticsQuenched2009} and amorphous clusters
+\cite{sarkarUniversalityVibrationalSpectra2004,
+matharooVibrationalSpectraAmorphous2005,
+matharooUniversalityVibrationalSpectra2008}. Within biological systems, RMT has
+also been used to successfully model aspects of  amino acid functional
+relationships \cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in
+epileptic seizures \cite{osorioPhasesynchronizationRandommatrixBased2011}, and
+in protein-protein interactions both in different species
 \cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
@@
 veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
-more recently, has provided insights into the behaviors of deep
-neural networks\cite{martinImplicitSelfRegularizationDeep2021,
+more recently, has provided insights into the behaviors of deep neural
+networks\cite{martinImplicitSelfRegularizationDeep2021,
 martinPredictingTrendsQuality2021}.
@@
 activity, RMT may be an ideal starting point, as each voxel time course (or
-collection of such time sources - i.e. ROIs) can be considered an interacting
+collection of such time sources, i.e. ROIs) can be considered an interacting
 component of the system. Likewise, in functional connectivity
 analyses, statistical relationships of the BOLD ROI time-courses are
-investigated in the hope of gaining insights into brain functioning (see
+investigated in the hope of gaining insights into brain function (see
 \citenum{bastosTutorialReviewFunctional2016,
@@
 
-The earliest study taking this approach
-demonstrated that spectra of the correlations between electroencephalographic
-(EEG) signals closely resemble those of the Gaussian Orthogonal Ensemble (GOE)
-\cite{sebaRandomMatrixAnalysis2003}. In fMRI, RMT has been used to
-evaluate the quality of whole brain features extracted from fMRI data
-\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}, and in
-diffusion MRI to aid in the selection of the number of
-components to employ in principal-component reduction analysis and denoising
-\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.
+The earliest study taking this approach demonstrated that spectra of the
+correlations between electroencephalographic (EEG) signals closely resemble
+those of the Gaussian Orthogonal Ensemble (GOE)
+\cite{sebaRandomMatrixAnalysis2003}. In fMRI, RMT has been used to evaluate the
+quality of whole brain features extracted from fMRI data
+\cite{voultsidouFeatureEvaluationFMRI2007, verganiRestingStateFMRI2019}, and in
+diffusion MRI to aid in the selection of the number of components to employ in
+principal-component reduction analysis and denoising
+\cite{veraartDenoisingDiffusionMRI2016, verganiRestingStateFMRI2019,
+ulfarssonDimensionEstimationNoisy2008}.
 
@@
 smithCharacterizingIndividualDifferences2014}): the mean signal over that ROI
-is correlated with all other N ROIs of interest to generate a reduced
+is correlated with all other \(N\) ROIs of interest to generate a reduced
 functional connectivity matrix (or image) with one correlation value at each
-non-seed ROI. This sort of reduction reduces the functional connectivity to N
-values, but necessarily biases the analysis to the seed ROI.
+non-seed ROI. This sort of reduction reduces the functional connectivity to
+\(N\) values, but necessarily biases the analysis to the seed ROI.
 
@@
 Previous studies applying RMT to functional connectivity
-data\cite{wangRandomMatrixTheory2016,
-wangSpectralPropertiesTemporal2015a,matharooSpontaneousBackpainAlters2020,
-guRandomMatrixTheory2020} took largely descriptive / explanatory approaches,
-noting only differences in RMT across subgroups and/or conditions. For example
-\citenum{wangRandomMatrixTheory2016} and \citenum{guRandomMatrixTheory2020} use
-a Kolmogorov-Smirnov test, and \citenum{wangSpectralPropertiesTemporal2015a}
-uses a t-test to provide evidence of differences in various RMT metrics.
-However, statistically significant differences in a metric do not imply
-practical significance or predictive utility
-\cite{loWhySignificantVariables2015}, nor do they imply generalization or
-replicability \cite{amrheinEarthFlat052017}. Cross-validation, by contrast,
-attempts to more directly assess these properties
-\cite{yarkoniChoosingPredictionExplanation2017}.
+data\cite{wangRandomMatrixTheory2016, wangSpectralPropertiesTemporal2015a,
+matharooSpontaneousBackpainAlters2020, guRandomMatrixTheory2020} took largely
+descriptive / explanatory approaches, noting only differences in RMT across
+subgroups and/or conditions. For example Wang et al.
+(2016)\cite{wangRandomMatrixTheory2016} and Gu et al.
+(2020)\cite{guRandomMatrixTheory2020} use a Kolmogorov-Smirnov test, and Wang
+et al. (2015)\cite{wangSpectralPropertiesTemporal2015a} uses a t-test to
+provide evidence of differences in various RMT metrics. However, statistically
+significant differences in a metric do not necessarily imply practical
+significance or predictive utility \cite{loWhySignificantVariables2015}, nor do
+they imply generalization or replicability \cite{amrheinEarthFlat052017}.
+Cross-validation, by contrast, attempts to more directly assess these
+properties \cite{yarkoniChoosingPredictionExplanation2017}.
 
@@
 non-trivial, and additionally requires an \textit{unfolding}
-procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}.
+procedure\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004}.
 Unfolding is an exponential fitting procedure that involves a number of highly
@@
 
-On top of the flexibility in the implementation and application of RMT to
-empirical data, there is also the flexibility introduced by the choices in the
-complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
-These "researcher degrees of
-freedom"\cite{simmonsFalsePositivePsychologyUndisclosed2011b} in combination
+In addition to the flexibility in the implementation and application of RMT to
+empirical data, there is also flexibility introduced by analytic choices made
+in the complex preprocessing pipelines of
+fMRI\cite{parkerBenefitSliceTiming2019}. These "researcher degrees of
+freedom"\cite{simmonsFalsePositivePsychologyUndisclosed2011b}, in combination
 with the absence of reproducible code and data availability, and predominance
-of descriptive and explanatory approaches, raise doubts about the basic robustness
-and practical utility of RMT-based functional connectivity analyses.
+of descriptive and explanatory approaches, raise doubts about the basic
+robustness and practical utility of RMT-based functional connectivity analyses.
 
@@
 We selected fMRI data publicly available on the
-\href{https://openneuro.org/}{OpenNeuro} platform (36). Selection criteria was
-somewhat subjective, but we required (1), the dataset comprise 10+ human
-subjects, (2) that all fMRI scans have the same spatial and temporal
-resolutions, (3) that the fMRI images could be split into various classes for a
-number of binary classification tasks, and (4), that it was reasonable to
-classify each scan in its entirety, (i.e. neither task nor event timings
-were needed to determine class membership, nor did any scan volumes need to be removed
-to make the scan classifiable). This yielded seven datasets total.
+\href{https://openneuro.org/}{OpenNeuro}
+platform\cite{markiewiczOpenNeuroResourceSharing2021}. Selection criteria was
+somewhat subjective, but we required each dataset to (1) comprise 10 or more
+human subjects, (2) have all fMRI scans have the same spatial and temporal
+resolutions, (3) have fMRI images that can be split into various classes for a
+number of binary classification tasks, and (4), allow reasonably classifying
+each scan in its entirety (i.e. neither task nor event timings are needed to
+determine class membership, nor do any scan volumes need to be removed to make
+the scan classifiable). This yielded seven datasets total.
 
@@
 \caption{
-    ID = Identifier for paper. TR = Time of Repetition (seconds).
-    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
-    indicate P horizontal / axial slices each with dimensions M x N.
+    \label{tab:data-dimensions}
+    Included fMRI dataset details.
+    Name = Identifier for paper.
+    Dimensions listed as M x N x P, indicate P axial slices each with dimensions M x N.
+    TR = Time of Repetition (seconds).
+    n\_scans = Total number of images in dataset.
 }
-\label{tab:data-dimensions}
 \small
@@
 aspects of each dataset here.  Dataset scan parameters and acquisition details
-are summarized in Table \ref{tab:data-dimensions}, and sample and subgroup sizes are
-summarized in Table \ref{tab:data-subjects}.
+are summarized in Table \ref{tab:data-dimensions}, and sample and subgroup
+sizes are summarized in Table \ref{tab:data-subjects}.
 
@@
 \caption{
-    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
+    \label{tab:data-subjects}
+    Sizes and other details for classification task subgroups.
+    ID = Identifier for paper. Subgroup = name of subgroup used in classification task. Subjects = Number of subjects.
     Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
 }
-\label{tab:data-subjects}
 \small
@@
 
-% Resting state fMRI scans were performed using a T2*-weighted EPI sequence with the following
-% parameters: TR=875ms; TE=43.6ms; slice thickness=1.8mm; acquisition matrix=100 × 100; flip
-% angle=55°; number of slices=72; field of view=180 × 180mm; voxel size=1.8 × 1.8 × 1.8mm;
-% multi-band factor=8; volumes per run=823 (total scan time=12min). EPI scans were oriented
-% parallel to the long axis of the hippocampus.
+% Resting state fMRI scans were performed using a T2*-weighted EPI sequence
+% with the following parameters: TR=875ms; TE=43.6ms; slice thickness=1.8mm;
+% acquisition matrix=100 × 100; flip angle=55°; number of slices=72; field of
+% view=180 × 180mm; voxel size=1.8 × 1.8 × 1.8mm; multi-band factor=8; volumes
+% per run=823 (total scan time=12min). EPI scans were oriented parallel to the
+% long axis of the hippocampus.
 
@@
 bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans from
-non-depressed controls and mildly- or moderately-depressed subjects. In our analysis,
-the binary classification task is \code{depress v control}.
+non-depressed controls and mildly- or moderately-depressed subjects. In our
+analysis, the binary classification task is \code{depress v control}.
 
@@
 
-% \subsubsection{Learning Dataset}
-
 The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
@@
 
-% \subsubsection{Osteoarthritis Dataset}
 The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
@@
 
-% \subsubsection{Parkinson's Dataset}
 The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
@@
 \code{park v ctrl}.
-
-% \subsubsection{Attention Dataset}
 
@@
 
-The \code{viglant v nonvigilant} task was formed based on self-report
+The \code{vigilant v nonvigilant} task was formed based on self-report
 questionnaire items involving
@@
 registration to the MNI152
-\cite{fonovUnbiasedAverageAgeappropriate2011,fonovUnbiasedNonlinearAverage2009}
+\cite{fonovUnbiasedAverageAgeappropriate2011, fonovUnbiasedNonlinearAverage2009}
 2mm isotropic, asymmetric template version C made available through
@@
 Brain extraction was performed first with BET
-\cite{smithFastRobustAutomated2002}, and then with with
+\cite{smithFastRobustAutomated2002}, and then with
 ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
@@
 acceptable. Motion-correction was performed with FSL's MCFLIRT
-\cite{jenkinsonImprovedOptimizationRobust2002}, and slicetime correction also
-with FSL's \code{slicetimer}\cite{jenkinsonFSL2012} tool. Finally, functional
+\cite{jenkinsonImprovedOptimizationRobust2002}, and slice time correction
+with FSL's slicetimer\cite{jenkinsonFSL2012} tool. Finally, functional
 images were registered directly to the MNI152 template via
@@
 
-We limited preprocessing methods to these steps sometimes because data required
-for other typical preprocessing steps (e.g. field intensity maps, physiological
+We limited preprocessing methods to these steps because data required for other
+typical preprocessing steps (e.g. field intensity maps, physiological
 measurements) was missing, but also because all pipeline intermediates were
@@
 \(N\) non-constant brain voxels, where \(N\) may differ for each image. After
-this process, each scan yields an \(N \times T\) matrix \(\mathbf{M}\), where
-\(T\) is the number of volumes acquired. All features summarize this matrix
-of brain voxels \(\mathbf{M}\) in some manner.
+this process, each scan yields an \(N \times t\) matrix \(\mathbf{M}\), where
+\(t\) is the number of volumes acquired. All features extracted summarize
+this matrix \(\mathbf{M}\).
 
@@
 \label{tab:features}
-Feature groupings for summarization. eigs = non-RMT eigenfeatures and their combinations with
-other non-RMT eigenfeatures. rmt = RMT eigenfeatures and their combinations. tseries = baseline
-timeseries reductions.\\
+Feature groupings for summarization. eigs = non-RMT eigenfeatures and their
+combinations with other non-RMT eigenfeatures. rmt = RMT eigenfeatures and
+their combinations. tseries = baseline timeseries reductions.\\
 }
@@
 The time and space complexities of computing the full \(N \times N\)
-\(\mathbf{M}\) of Pearson correlation coefficients (and subsequent eigenvalues) are too
-large to be tractable. However, since transposition does not change the
-eigenvalues, we can use the transpose to efficiently compute these eigenvalues
-(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
-in Table \ref{tab:features}), most directly test the idea that eigenvalues
-alone are useful functional connectivity features.
+\(\mathbf{M}\) matrix of Pearson correlation coefficients (and subsequent
+eigenvalues) are too large to be tractable. However, since transposition does
+not change the eigenvalues, we can use the transpose to efficiently compute
+these eigenvalues (see Appendix \ref{sec:transpose}). These raw eigenvalues
+(feature \code{eigs} in Table \ref{tab:features}), most directly test the idea
+that eigenvalues alone are useful functional connectivity features.
 
@@
 \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
-details and motivation behind the unfolding procedure are well documented in
-\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
-can be viewed as a smoothing and rescaling procedure where the originally
-observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
-an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
-if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
-function is not known, and so must be approximated
+details and motivation behind the unfolding procedure are well documented
+elsewhere\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the
+unfolding process can be viewed as a smoothing and rescaling procedure where
+the originally observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is
+smoothly mapped to an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum
+d_i/n \approx 1\), if \(d_i = e_{i+1} - e_i \). For most empirical data, the
+true smoothing function is not known, and so must be approximated
 \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
@@
 \textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
-with a polynomial. Because the choice of degree has been known to dramatically impact
-certain analyses involving the spectral rigidity or level
+with a polynomial. Because the choice of degree has been known to dramatically
+impact certain analyses involving the spectral rigidity or level
 variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
@@
 fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}, we
-compute the rigidity and level variance with all possible combinations of
-trimming and unfolding polynomial degrees of 3, 5, 7, and 9.
-
-\textit{Software Release}. Because the computation of these statistics is non-trivial, and to make the
-unfolding procedure more transparent and reproducible, we developed and release
-the separate, open-source Python library
+compute the rigidity and level variance with all possible combinations of our
+trimming procedures, and unfolding polynomial degrees of 3, 5, 7, and 9.
+
+\textit{Software Release}. Because the computation of these statistics is
+non-trivial, and to make the unfolding procedure more transparent and
+reproducible, we developed and release a separate, open-source Python library
 \code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
@@
 The largest, middle, or smallest eigenvalues could be most useful in
-characterizing some system. E.g. if the smallest eigenvalues correspond to
+characterizing any given system. E.g. if the smallest eigenvalues correspond to
 random / noise aspects of a system, but differences in the nature of the system
@@
 percentile—\code{T-rrng} in Table \ref{tab:features}). Each statistic reduces
-the fMRI data along the voxel dimension, yielding a feature of \(T\) dimensions
+the fMRI data along the voxel dimension, yielding a feature of \(t\) dimensions
 (e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
@@
 extracted from an fMRI image prepared with one of four levels or degrees of
-preprocessing, where preprocessing level \(k\) includes preprocessing steps 1 to \(k\),
-inclusive, and starting at \(k = 1\).
+preprocessing, where preprocessing level \(k\) includes preprocessing steps 1
+to \(k\), inclusive, and starting at \(k = 1\).
 
@@
 
-Each non-baseline feature is thus evaluated, for each classification task and
-classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\) 2
-normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
-degrees  \(\times\) 10 slice choices. Each baseline feature is evaluated
-exactly 40 times for each comparison task and classifier: 4 preprocessing
-levels \(\times\) 2 normalization methods \(\times\) 5 smoothing window sizes.
-In total, with all the datasets examined in this study, this yields 2 053 920
-mAUROC values to summarize, with 5 main \textit{analytic factors}—
-preprocessing, normalization, trimming, unfolding/smoothing degree, and
-slicing—to consider.
+Each non-baseline feature is thus evaluated, for each dataset classification
+task and each classifier, exactly 1280 times: there are 4 preprocessing levels
+\(\times\) 2 normalization methods \(\times\) 4 trimming choices \(\times\) 4
+unfolding degrees  \(\times\) 10 slice choices. Each baseline feature is
+evaluated exactly 40 times for each comparison task and classifier: 4
+preprocessing levels \(\times\) 2 normalization methods \(\times\) 5 smoothing
+window sizes. In total, with all the datasets and features examined in this
+study, this yields 2,053,920 mAUROC values to summarize, with 5 main
+\textit{analytic factors}— preprocessing, normalization, trimming,
+unfolding/smoothing degree, and slicing—to consider.
 
@@
 Thus, for example, figures depicting the coarse feature grouping "eigs" in fact
-depict all observed mAUROC values for the the fine feature groupings of raw,
+depict all observed mAUROC values for the fine feature groupings of raw,
 max, middle, and smoothed eigenvalues, and the fine feature grouping "eigs
@@
 classification task or an analytic factor from summaries if there is no
-evidence that the classification task is solvable in general, or for that
-factor. We take as lack of evidence of solvability an mAUROC distribution that
-is either (1) roughly symmetric and has mean and median close to 0.5, i.e.
-performance appears random, or (2) with median and mode less than 0.5.
-
-Based on the mAUROC distributions for either coarsely-grouped (Figure
+evidence that the classification task is solvable in general, or if, when
+restricting the analytic factor to a particular instance, there is no evidence
+of solvability. E.g. if a particular trimming procedure were to render all
+classification tasks unsolvable, it would be better to note this and exclude
+the associated mAUROCs from further visualizations, rather than to have the
+mAUROC distributions diluted by the bad procedure.
+
+We take as lack of evidence of solvability an mAUROC distribution that is
+either (1) roughly symmetric and has mean and median close to 0.5, i.e.
+performance appears random, or (2) with median and mode less than 0.5. Based on
+the mAUROC distributions for either coarsely-grouped (Figure
 \ref{fig:main-results}) or finely-grouped features (Figure
@@
 \code{trait\_attend v trait\_nonattend} conditions ("WeeklyAttend" in Figures
-\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the \textit{Attention}
-dataset were also not meaningfully predictable by any feature. As such, we exclude
-these classification tasks from further figures and discussion.
+\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the
+\textit{Attention} dataset were also not meaningfully predictable by any
+feature. As such, we exclude these classification tasks from further figures
+and discussion. We did not, however, find that any analytic factor resulted in
+unsolvability.
 
@@
 shows that this was primarily due to either the "rmt only" or "rmt + eigs"
-features (see Table \ref{tab:features}). However, RMT eigenfeatures were also most
-likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
-0.5; Figure S\ref{S-fig:fine-smallest}).
+features (see Table \ref{tab:features}). However, RMT eigenfeatures were also
+most likely to cause poor performance and overfitting (indicated by an mAUROC
+\(<\) 0.5; Figure S\ref{S-fig:fine-smallest}).
 
@@
 restricting to the most predictive RMT features (those including the unfolded
-eigenvalues), it seems clear from Figure S\ref{S-fig:fine-degree} the the
+eigenvalues), it seems clear from Figure S\ref{S-fig:fine-degree} that the
 largest unfolding degree of 9 produces the most favourable mAUROC
@@
 potential (Figures \ref{fig:main-results}-\ref{fig:largest}). In addition to
-evidence from previous studies
-\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
-this suggests RMT may be a useful analytic and theoretical tool for
-understanding functional connectivity.
+evidence from previous studies \cite{sebaRandomMatrixAnalysis2003,
+wangRandomMatrixTheory2016, matharooSpontaneousBackpainAlters2020}, this
+suggests RMT may be a useful analytic and theoretical tool for understanding
+functional connectivity.
 
@@
 of preprocessing steps, or feature slicing) that ensured, for any combination
-of feature and classification task, that all all other analytic choices
-resulted in mAUROC values greater than 0.5. In addition, the mean, median and
-modal mAUROCs were generally close to 0.5, and adjusted mean and median
-accuracies also tended to be close to zero. Thus we find limited evidence that
+of feature and classification task, that all other analytic choices resulted in
+mAUROC values greater than 0.5. In addition, the mean, median and modal mAUROCs
+were generally close to 0.5, and adjusted mean and median accuracies also
+tended to be close to zero. Thus we find limited evidence that
 functional-connectivity-based eigenfeatures have general, "out of the box"
@@
 grouping for that row was not in the top three, i.e., that the top three
-performing features differed depending the performance metric. }
+performing features differed depending on the performance metric. }
 \small
@@
 Nevertheless, in all datasets, there were combinations of analytic choices that
-resulted in cross-validated mean prediction performances well beyond mere guessing
-(Figures S\ref{S-fig:fine-largest}, S\ref{S-fig:fine-largest-acc}, Table
-\ref{tab:top3-fine}). Whether or not these should be considered to have
+resulted in cross-validated mean prediction performances well beyond mere
+guessing (Figures S\ref{S-fig:fine-largest}, S\ref{S-fig:fine-largest-acc},
+Table \ref{tab:top3-fine}). Whether or not these should be considered to have
 practical relevance depends on one's goals, however, we note that with small
@@
 \cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional
-connectivity data and with classical machine learning algorithms (such as SVM)
+connectivity data and with classical machine learning algorithms (such as SVC)
 we in general only expect large prediction accuracies (e.g. greater than 80\%)
@@
 study, the (robust) maximum improvements upon guessing are shown in Table
-\ref{tab:top3-fine}), and vary from 3-26\%.
+\ref{tab:top3-fine}, and vary from 3-26\%.
 
@@
 cross-validated predictive utility. The reduction of the functional
-connectivity matrix to the sorted \(T - 1\) eigenvalues uses all brain voxels
+connectivity matrix to the sorted \(t - 1\) eigenvalues uses all brain voxels
 (including grey matter voxels), and destroys a large amount of information
@@
 such curves could likely be fit near-perfectly with 3-5 parameters, i.e. the
-inherent dimensionality of these features is quite small. Interpreting the eigenvalues
-as the magnitudes of the principal components of the standardized data, this
-suggests that a rough summary of the magnitudes of the principal components can
-often be surprisingly predictive in fMRI.
+inherent dimensionality of these features is quite small. Interpreting the
+eigenvalues as the magnitudes of the principal components of the standardized
+data, this suggests that a rough summary of the magnitudes of the principal
+components can often be surprisingly predictive in fMRI.
 
@@
 
-% It was also sometimes the case that adding RMT-specific features (e.g. the spectral
-% rigidity, or level variance) to either the unfolded or raw
+% It was also sometimes the case that adding RMT-specific features (e.g. the
+% spectral rigidity, or level variance) to either the unfolded or raw
 % eigenvalues would allow for a higher maximum mAUROC (Tables
-% \ref{tab:numerical}-\ref{tab:top3-fine}, and Figure S\ref{S-fig:fine-largest}). One case where these
-% features yielded the largest predictions was in the \textit{Aging} dataset.
-% This is depicted in Figure \ref{fig:observables}, where it is clear that, after
-% trimming, at almost any \(L\) value, subgroups either cluster in bands, or tend to
-% generally have lower or higher rigidity of level variance curves.
+% \ref{tab:numerical}-\ref{tab:top3-fine}, and Figure
+% S\ref{S-fig:fine-largest}). One case where these features yielded the largest
+% predictions was in the \textit{Aging} dataset. This is depicted in Figure
+% \ref{fig:observables}, where it is clear that, after trimming, at almost any
+% \(L\) value, subgroups either cluster in bands, or tend to generally have
+% lower or higher rigidity of level variance curves.
 
@@
 \textit{Aging} dataset, and it is quite possible for randomization failures to
-introduce age confounding across classes.
+introduce age imbalances across classes.
 
@@
 
-In general, much larger fMRI datasets are needed to adequately control for and test
-\textit{what} exactly the connectivity eigenvalues actually predict, and to test
-if these predictions generalize to larger, different populations.
+In general, much larger fMRI datasets are needed to adequately control for and
+test \textit{what} exactly the connectivity eigenvalues actually predict, and
+to test if these predictions generalize to larger, different populations.
 
@@
 investigate the sensitivity of any findings to such analytic decisions before
-drawing any conclusions.
-
-\section{Disclosures}
-The authors have no relevant financial interests in the manuscript and no other
-potential conflicts of interest to disclose.
-
-% Conflicts of interest should be declared under a separate header, above
-% Acknowledgments. If the authors have no competing interests to declare, then a
-% statement should be included declaring no conflicts of interest. For assistance
-% generating a disclosure statement, see the form available from  the
-% International Committee of Medical Journal Editors website:
-% \linkable{http://www.icmje.org/conflicts-of-interest/}
+drawing any conclusions. However, this study demonstrates that the inclusion of
+RMT statistics and other eigenfeatures extracted from fMRI images could aid in
+the predictive modeling of a wide variety of phenomena.
+
+\newpage
+
+\section*{Disclosures}
+JL is founder of Time Will Tell Technologies, Inc.. The authors have no
+relevant financial interests in the manuscript and no other potential conflicts
+of interest to disclose.
 
@@
 All code used specifically to perform preprocessing and analyses for this paper
-is publicly available \href{https://github.com/DM-Berger/random-matrix-fmri/}{online} in
-the paper GitHub repository\cite{dm-bergerDMBergerRandommatrixfmriPaper2022}.
-This repository also includes all produced prediction data in a single table
+is publicly available
+\href{https://github.com/DM-Berger/random-matrix-fmri/}{online} in the paper
+GitHub repository\cite{dm-bergerDMBergerRandommatrixfmriPaper2022}. This
+repository also includes all produced prediction data in a single table
 available in the file \code{all\_produced\_prediction\_data.json.xz} available
@@
 
-Code used to compute the spectral observables and
-perform unfolding is publicly available in the
+Code used to compute the spectral observables and perform unfolding is publicly
+available in the
 \href{https://github.com/stfxecutables/empyricalRMT}\code{empyricalRMT}
@@
 
-% \subsection{Page Setup and Fonts}
-
-% All text and figures, including footnotes, must fit inside a text area 6.5 in.\
-% wide by 9 in.\ high (16.51 by 22.86 cm). Manuscripts must be formatted for US
-% letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1
-% in.\ on the bottom, and 1 in.\ on the left and right.
-
-% The Times New Roman font is used throughout the manuscript, in the sizes and
-% styles shown in Table~\ref{tab:fonts}. If this font is not available, use a
-% similar serif font. The manuscript should not contain headers or footers. Pages
-% should be numbered.
-
-% \begin{table}[ht]
-% \caption{Fonts sizes and styles.}
-% \label{tab:fonts}
-% \begin{center}
-% \begin{tabular}{|l|l|} %% this creates two columns
-% %% |l|l| to left justify each column entry
-% %% |c|c| to center each column entry
-% %% use of \rule[]{}{} below opens up each row
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
-% \hline\hline
-% \rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
-% \hline
-% \rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
-
-% \section{Parts of Manuscript}
-
-% This section describes the normal structure of a manuscript and how each part
-% should be handled. The appropriate vertical spacing between various parts of
-% this document is achieved in LaTeX through the proper use of defined
-% constructs, such as \verb|\section{}|.
-
-% \subsection{Title and Author Information}
-% \label{sect:title}
-% The article title appears left justified at the top of the first page. The
-% title font is 16 pt., bold. The rules for capitalizing the title are the same
-% as for sentences; only the first word, proper nouns, and acronyms should be
-% capitalized. Do not begin titles with articles (for example, a, an, the) or
-% prepositions (for example, on, by, etc.). The word ``novel'' should not appear
-% in the title, as publication will imply novelty. Avoid the use of acronyms in
-% the title, unless they are widely understood. Appendix A contains more about
-% acronyms.
-
-% The list of authors immediately follows the title, 18 points below. The font is
-% 12 pt., bold and the author names are left justified. The author affiliations
-% and addresses follow the names, in 10-pt., normal font and left justified. For
-% multiple affiliations, each affiliation should appear on a separate line.
-% Superscript letters (a, b, c, etc.) should be used to associate multiple
-% authors with their respective affiliations. The corresponding author should be
-% identified with an asterisk, and that person's email address should be provided
-% below the keywords.
-
-% \subsection{Abstract}
-% The abstract should be a summary of the paper and not an introduction. Because
-% the abstract may be used in abstracting journals, it should be self-contained
-% (i.e., no numerical references) and substantive in nature, presenting concisely
-% the objectives, methodology used, results obtained, and their significance.
-% Please note that the following journals require the use of structured abstracts
-% in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of
-% Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured
-% abstracts are encouraged for the \textit{Journal of Micro/Nanolithography,
-% MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on
-% the website of the journal.
-
-% \subsection{Subject terms/Keywords}
-% Keywords are required. Please provide 3-6 keywords related to your paper.
-
-% \subsection{Body of Paper}
-% The body of the paper consists of numbered sections that present the main
-% findings. These sections should be organized to best present the material.
-
-% To provide transition elements in your paper, it is important to refer back (or
-% forward) to specific sections. Such references are made by indicating the
-% section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1
-% contained a description...'' If the word Section, Reference, Equation, or
-% Figure starts a sentence, it is spelled out. When occurring in the middle of a
-% sentence, these words are abbreviated Sec., Ref., Eq., and Fig.
-
-% At the first occurrence of an acronym, spell it out followed by the acronym in
-% parentheses, for example, charge-coupled diode (CCD).
-
-% \subsection{Footnotes}
-% Textual footnotes should be used rarely to present important documentary or
-% explanatory material whose inclusion in the text would be
-% distracting.\footnote{Example of a footnote.} Due to problems with HTML
-% display, use of footnotes should generally be avoided. If absolutely necessary,
-% the footnote mark must come at the end of a sentence. To insert a footnote, use
-% the {\verb|\footnote{}|} command.
-
-% \subsection{Appendices}
-% Brief appendices may be included when necessary, such as derivations of
-% equations, proofs of theorems, and details of algorithms. Equations and figures
-% appearing in appendices should continue sequential numbering from earlier in
-% the paper.
-
-% \subsection{Disclosures}
-% Conflicts of interest should be declared under a separate header, above
-% Acknowledgments. If the authors have no competing interests to declare, then a
-% statement should be included declaring no conflicts of interest. For assistance
-% generating a disclosure statement, see the form available from  the
-% International Committee of Medical Journal Editors website:
-% \linkable{http://www.icmje.org/conflicts-of-interest/}
-
-% \subsection{Acknowledgments}
-% Acknowledgments and funding information should be added after the conclusion,
-% and before references. Include grant numbers and the full name of the funding
-% body. The acknowledgments section does not have a section number.
-
-% \subsection{Data, Materials, and Code Availability}
-% Availability of data, materials, and/or code used in the research results
-% reported in the manuscript may be declared under the heading ``Data, Materials,
-% and Code Availability,'' following the Acknowledgments section. As relevant,
-% provide specific access information or restrictions for data, materials, and
-% computer code (i.e., links to repository access addresses with guidance on
-% commercial or public access).
-
-% \subsection{References}
-% The References section lists books, articles, and reports that are cited in the
-% paper. This section does not have a section number. The references are numbered
-% in the order in which they are cited. Examples of the format to be followed are
-% given at the end of this document.
-
-% The reference list at the end of this document is created using BibTeX, which
-% looks through the file {\ttfamily report.bib} for the entries cited in the
-% LaTeX source file.  The format of the reference list is determined by the
-% bibliography style file {\ttfamily spiejour.bst}, as specified in the \\
-% \verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may
-% be directly formatted in the LaTeX source file.
-
-% For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list
-% of authors (initials plus last name), book title (in italics), page or chapter
-% numbers, publisher, city, and year of publication.  Journal-article references
-% \citenum{Metropolis53,Harris06} include the author list, title of the article
-% (in quotes), journal name (in italics, properly abbreviated), volume number (in
-% bold), inclusive page numbers or citation identifier, and year.  A reference to
-% a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes
-% the author list, title of the article (in quotes), conference name (in
-% italics), editors (if appropriate), volume title (in italics), volume number if
-% applicable (in bold), inclusive page numbers, publisher, city, and year.
-% References to an article in the SPIE Proceedings may include the conference
-% name, as shown in Ref.~\citenum{Hanson93c}.
-
-% The references are numbered in the order of their first citation. Citations to
-% the references are made using superscripts, as demonstrated in the preceding
-% paragraph. One may also directly refer to a reference within the text, for
-% example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more
-% references should be separated by a comma with no space between them. Multiple
-% sequential references should be displayed with a dash between the first and
-% last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.
-
-% \subsubsection{Reference linking and DOIs}
-% A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a
-% digital object, such as a journal article or a book chapter, that provides a
-% persistent link to its location on the internet. The use of DOIs allows readers
-% to easily access cited articles. Authors should include the DOI at the end of
-% each reference in brackets if a DOI is available. See examples at the end of
-% this manuscript. A free DOI lookup service is available from CrossRef at
-% \\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will
-% facilitate reference linking and is highly recommended.
-
-% In the present LaTeX template, the author needs to add the DOI reference by
-% including it in a ``note'' in the bibliography file, as shown in the file
-% {\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
-% The DOI may be used by the reader to locate that document with the link:
-% {\verb+http://dx.doi.org10.1117/12.154577+}.
-
-% \subsection{Biographies}
-% A brief professional biography of approximately 75 words may be provided for
-% each author, if available. Biographies should be placed at the end of the
-% paper, after the references. Personal information such as hobbies or
-% birthplace/birthdate should not be included. Author photographs are not
-% published.
-
-% \section{Section Formatting}
-% \label{sect:sections}
-% In LaTeX, a new section is created with the \verb|\section{}| command, which
-% automatically numbers the sections. Sections will be numbered sequentially,
-% starting with the first section after the abstract, except for the
-% acknowledgments and references. (Note that numbering of section headings is not
-% required, but the numbering must be consistent if used.) All section headings
-% should be left justified.
-
-% Main section headings are in 12-pt. bold font, left-justified and in title
-% case, where important words are capitalized.
-
-% Paragraphs that immediately follow a section heading are leading paragraphs and
-% should not be indented, according to standard publishing style. The same goes
-% for leading paragraphs of subsections and sub-subsections. Subsequent
-% paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no
-% additional space between paragraphs. In LaTeX, paragraphs are separated by
-% blank lines in the source file. Indentation of the first line of a paragraph
-% may be avoided by starting it with \verb|\noindent|.
-
-% \subsection{Subsection Headings}
-% All important words in a subsection (level 1) header are capitalized.
-% Subsection numbers consist of the section number, followed by a period, and the
-% subsection number within that section, without a period at the end. The heading
-% is left justified and its font is 12 pt. italic.
-
-% \subsubsection{Sub-subsection headings}
-% The first word of a sub-subsection is capitalized. The rest of the text is not
-% capitalized, except for proper names and acronyms (the latter should only be
-% used if well known). The heading is left justified and its font is 11 pt.
-% italic.
-
-% \section{Figures and Tables}
-
-% \subsection{Figures}
-
-% Figures are numbered in the order in which they are called out in the text.
-% They should appear in the document in numerical order and as close as possible
-% to their first reference in the text. It may be necessary to move figures or
-% tables around to enhance readability. LaTeX will attempt to place figures at
-% the top or bottom of a page in which they are first referenced.
-
-% Figures, along with their captions, should be separated from the main text by
-% 0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or
-% graph. Figure captions start with the abbreviation ``Fig'' in front of the
-% figure number, followed by a period, and the text in 10-pt. font. See
-% Fig.~\ref{fig:example} for an example.
-
-% \begin{figure}
-% \begin{center}
-% \begin{tabular}{c}
-% \includegraphics[height=5.5cm]{mcr3b.eps}
-% \end{tabular}
-% \end{center}
-% \caption
-% { \label{fig:example}
-% Example of a figure caption. }
-% \end{figure}
-
-% Authors may wish to create figures consisting of two or more images, in which
-% case, they should be neatly arranged in a rectangular array.  In no case,
-% should the article's text be wrapped around a figure. Figure~\ref{fig:example2}
-% shows two side-by-side images. When a figure contains more than one image, the
-% author must submit them as a single image file. Further details about figure
-% formatting can be found in the author guidelines for each specific SPIE
-% journal: \\
-
-% \linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.
-
-% \begin{figure}
-% \begin{center}
-% \begin{tabular}{c}
-% \includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
-% \\
-% (a) \hspace{5.1cm} (b)
-% \end{tabular}
-% \end{center}
-% \caption
-% { \label{fig:example2} Example of a figure containing multiple images: (a) sun
-% and (b) blob. Figures containing multiple images must be submitted to SPIE as a
-% single image file.}
-% \end{figure}
-
-% \subsection{Tables}
-% Tables are numbered in the order in which they are referenced.
-They should appear in the document in numerical order and as close as possible to their first reference in the text.
-It is preferable to have tables appear at the top or bottom of the page, if possible.
-Table captions are handled identically to those for figures, except that they appear above the table.
-See Table~\ref{tab:fonts} for an example.
-
-% \subsection{Multimedia}
-% Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player.
-The recommended maximum size for each multimedia file is 10-12 MB.
-Authors must insert a representative still image from the video file in the manuscript as a figure.
-The caption label will be linked by the publisher to the actual video file.
-The video may also be mentioned in an existing figure caption.
-Multimedia files are treated in the same manner as figures and they will be numbered sequentially with normal figures.
-The video number, file type, and file size should be included in parentheses at the end of the figure caption.
-See Figure \ref{vid:satellite} for an example.
-
-% \begin{video}
-% \begin{center}
-% {\includegraphics[height=5cm]{satellite.eps}}
-% \\
-% \end{center}
-% \caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
-% \end{video}
-
-\appendix    % this command starts appendixes
+\newpage
+
+\appendix
 
@@
 \]
-where \(\mathbf{X}_i\) denotes column \(i\) of \(\mathbf{X}\). Denote the (unordered) set
-of eigenvalues of \(\mathbf{X}\) as \(\text{eigs}(\mathbf{X})\), and let \(r = (p - 1)^{-1}\).
-Denote the covariance matrix of \(\mathbf{X}\) as \(\text{cov}(\mathbf{X})\).  Then:
-
+where \(\mathbf{X}_i\) denotes column \(i\) of \(\mathbf{X}\). Denote the
+(unordered) set of eigenvalues of \(\mathbf{X}\) as
+\(\text{eigs}(\mathbf{X})\), and let \(r = (p - 1)^{-1}\). Denote the
+covariance matrix of \(\mathbf{X}\) as \(\text{cov}(\mathbf{X})\).  Then:
 \begin{align*}
@@
 
-Denote the correlation matrix of \(\mathbf{X}\) as \(\text{corr}(\mathbf{X})\), and let
+Denote the correlation matrix of \(\mathbf{X}\) as \(\text{corr}(\mathbf{X})\),
+and let
 \[
@@
 \]
-
 Then
-
 \[
@@
 
-% \section{Miscellaneous Formatting Details}
-% \label{sect:misc}
-% At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.
-
-% \subsection{Formatting Equations}
-% Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.
-Important equations appear on their own line.
-Such equations are centered.
-For example, ``The expression for the field of view is
-% \begin{equation}
-% \label{eq:fov}
-% 2 a = \frac{(b + 1)}{3c} \, ,
-% \end{equation}
-% where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.
-
-% Equations are considered to be part of a sentence and should be punctuated accordingly.
-In the above example, a comma appears after the equation because the next line is a subordinate clause.
-If the equation ends the sentence, a period should follow the equation.
-The line following an equation should not be indented unless it is meant to start a new paragraph.
-Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.
-
-% References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise.
-Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..
-
-% \subsection{Formatting Theorems}
-
-% To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.
-The text of the theorem continues on the same line in normal, 10-pt.
-font, achieved in LaTeX using \verb|\footnotesize|.
-For example,
-
-% \vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}
-
-% \disclosures
-% \subsection*{Disclosures}
-% Conflicts of interest should be declared under a separate header.
-If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.
-
-% \subsection* {Acknowledgments}
-% This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.
-
-\subsection* {Data, Materials, and Code Availability}
-As relevant, the availability of data, materials, and/or software code used in the research results reported in the manuscript may be declared in this section.
-(Note: this section is required for the \textit{Journal of Biomedical Optics} and \textit{Neurophotonics}.) Provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).
-
 %%%%% References %%%%%

### 2022-11-15 paper/paper/SPIE/supplementary.tex
[WIP]
finalize
tidy up and make corrections to paper

@@
 }
+\usepackage[margin=1in]{caption}
 
@@
 
-\title{Supplementary Material}
+\title{Random Matrix Theory Tools for the Predictive Analysis of Functional Magnetic Resonance Imaging Examinations - Supplementary Material}
 
 \author[a]{Derek Berger}
-\author[a,*]{Jacob Levman}
-\author[b]{Gurpreet M. Matharoo}
-\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
-\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
+\author[b,c]{Gurpreet S. Matharoo}
+\author[a*,d,e,f]{Jacob Levman}
+
+\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+\affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+\affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
+\affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
+\affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}
 

### 2022-11-15 paper/paper/SPIE/article.tex
finalize

@@
 \usepackage{doi}
+\usepackage{lineno}
+\linenumbers
 \usepackage{csquotes}
@@
 \affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
-
 \affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
-
-\affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
-\affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
+\affil[d]{Athinoula A. Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
+\affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts, United States, 02115}
 \affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}

### 2022-11-15 paper/paper/SPIE/cover_letter.tex
finalize

@@
-\documentclass[10pt,a4paper]{article}
+\documentclass[10pt,letter]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{geometry}
+\geometry{
+  lmargin=2cm,
+  rmargin=2cm,
+  bmargin=2cm,
+}
 \usepackage{amsmath}
@@
 We confirm that this work is original and has not been published elsewhere nor
-is it currently under consideration for publication elsewhere.
+is it currently under consideration for publication elsewhere. However, a
+poster of limited preliminary results related to this study was presented at
+the International Symposium for Magnetic Resonance in Medicine 2021 annual
+conference, and is available in abstract form \citep{bergerOpenSourceRandom2021}.
 
@@
   \item systematically investigates the impact of the various methodological
-        decisions involved in the findings
+        decisions that affect the findings
   \item makes all code used in the paper publicly available, including the release of a
@@
 Jacob Levman, PhD (jlevman@stfx.ca) \\
-% Canada Research Chair in Bioinformatics \\
-Department of Computer Science \\
-St. Francis Xavier University \\
+\footnotesize
+\hspace*{0.25cm}St. Francis Xavier University, Department of Computer Science  \\
+\hspace*{0.25cm}Martinos Center for Biomedical Imaging, Massachusetts General Hospital \\
+\hspace*{0.25cm}Harvard Medical School, Department of Radiology (jlevman@mgh.harvard.edu) \\
+\hspace*{0.25cm}Nova Scotia Health Authority, Research Affiliate (jacob.levman@nshealth.ca) \\
+\normalsize
 
+\small
 \noindent
 \textbf{Additional Contacts} [should the corresponding author be unavailable]\\
-Derek Berger (dberger@stfx.ca), Department of Computer Science, St. Francis Xavier University\\
-Gurpreet S. Matharoo (gmatharo@stfx.ca), Department of Physics, St. Francis Xavier University
+Derek Berger (dberger@stfx.ca)\\
+\footnotesize
+\hspace*{0.25cm} St. Francis Xavier University, Department of Computer Science \\
+\small
+Gurpreet S. Matharoo (gmatharo@stfx.ca) \\
+\footnotesize
+\hspace*{0.25cm} St. Francis Xavier University, Department of Physics \\
+\hspace*{0.25cm} St. Francis Xavier University, ACENET
+\normalsize
+
+% \affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+% \affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}
+
+% \affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
+% \affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
+% \affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}
 

### 2023-04-03 paper/paper/SPIE/response_letter.tex
[WIP] - draft response cover letter

\documentclass[10pt,letter]{article}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[british]{babel}
\usepackage{url}
\usepackage{graphicx}
\graphicspath{ {figures/} }
\usepackage{geometry}
\geometry{
  lmargin=2cm,
  rmargin=2cm,
  bmargin=2cm,
}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[
    backend=biber,
    style=authoryear,
    natbib=true,
    sortlocale=en_US,
    url=false,
    doi=true,
    eprint=false
]{biblatex}
\addbibresource{report.bib}
\usepackage{fancyhdr}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.,0.,0.4}
\definecolor{darkred}{rgb}{0.5,0.,0.}
\usepackage[pdftex,colorlinks=true,linkcolor=darkblue,citecolor=darkred,urlcolor=blue]{hyperref}
\usepackage{xcolor}

\newcommand{\longcomment}[1]{}

\begin{document}
% \bibliographystyle{apa}
\fancypagestyle{nofooter}{\fancyfoot{}}
\thispagestyle{nofooter}  %  https://tex.stackexchange.com/questions/136384/remove-footer-without-removing-the-header
\setlength{\headheight}{30.0pt}
\fancyhead[C]{\includegraphics[height=0.5in]{stfx_text.png}}
\fancyhead[L]{\includegraphics[height=0.5in]{crest.png}}
\fancyhead[R]{
    \sffamily
    \scriptsize
    4130 University Avenue \\
    B2G 2W5 \\
    Antigonish, Nova Scotia \\
    Canada
}
\normalsize

\title{Cover Letter}
% \begin{center}
% \includegraphics[width=4in]{stfx_logo.png}
% \end{center}

\longcomment{
Specific comments:

% 1. Lines 42-45: I think it's unlikely that most readers of SPIE will be
% familiar with RMT and so it would be a good idea to have a bit more of an
% introduction - what's a random matrix?, what does RMT try to do (e.g.,
% characterize systems that can be described using matrices, statistically,
% rather than deterministically.) Later they talk about what RMT has been applied
% to, but it would be good to orient the reader to the general idea.

% 2. Line 44: "independently" should be "independent"

% 3. Line 56: This is a pretty broad statement, because it would seem that it's
% an open question whether all complex systems would lend themselves to RMT-based
% analysis.

% 4. Line 100: Do the authors really mean "eliminate" here? Do they instead mean
% "mask", given that the information presumably doesn't disappear.

% 5. Line 121: "functional connectivity". Do the authors mean "functional
% connectivity matrix"?

% 6. Line 138: This sentence is a bit garbled. Do the authors mean "often poorly
% documented or even missing..."?

% 7. Lines 143-145: I don't think these absences raise doubts about these two
% things in general, but rather that one has to carefully document the workflows
% and make the data and code available to the public.

% 8. Line 156: Do the authors mean "allow a reasonable classification of each
% scan without resorting to other information". Unclear what they mean by
% "entirety".

% 9. Table 1: Volumes and n_scans are not defined sufficiently. Are "Volumes" the
% number of volumes per run? Is n_scans to total number of runs? Is there one run
% per participant? The participant info is in Table 2, but Table 1 should be
% independently clear.

% 10. Line 180: "learning vs. rest" should be in a different font to be
% consistent with the other dataset descriptions.

% 11. Line 208: Are the "fractional anisotropy" values mentioned here the input
% variables for BET which govern how aggressive BET is and the gradient
% parameter?

12. Section 3.2: While I understand that the authors are trying to make a
general statement that will apply to all features, this section is unclear
because the reader has no idea at this point what features are. Perhaps this
could be expanded.

% 13. Line 223: How about "computing time" and "storage space", assuming that is
% what is meant.

% 14. Line 267: It's not clear that this section should be here, maybe at the end
% of the methods section?

% 15. Line 292: This is a minor point, but sentences shouldn't start with an
% abbreviation (E.g.) better "For example,". This also occurs in line 387.

% 16. Line 311: I see what they mean by "voxel time series" because of lines
% 317-318, so it might be clearer to move those lines to directly after line 311,
% so it is clear without having to read further.

17. The authors do a nice job of presenting the relevant results, but the
conclusions were rather general, being along the lines of "investigate the
sensitivity of choices made in the processing and analysis". That's an
important result, given previous work and shows that RMT might not "work" with
different data sets and processing workflows. The question is: what's the
takeaway: that the method is not one in which you can have a default set of
parameters that will generally give good results (like say, the recon-all
pipeline for FreeSurfer)? If RMT so sensitive to preprocessing and choice of
parameters, is this a useful method to be applied widely on fMRI data? It would
be good for this to be discussed and perhaps some recommendations given.
}

\noindent
\today, \\
Dear Dr. Gimi, \\

Thank you for your careful consideration and comments on our manuscript
\emph{Random Matrix Theory Tools for the Predictive Analysis of Functional
Magnetic Resonance Imaging Examinations} submitted for publication in SPIE
Journal of Medical Imaging.

In response to reviewer comments, we have made a number of changes to the
manuscript. Minor changes relating to wording or for which an edit is
straightforward (changing or adding a word or two) are identified in
\textcolor{orange}{orange} in the updated \.pdf manuscript (or with
\verb|\textcolor{orange}{text}| in the \.tex file). Other more significant
changes requiring the addition (or movement) of 1-3 sentences are identified
in \textcolor{cyan}{cyan} (\verb|\textcolor{cyan}{text}|).

\begin{description}
\item[Comment 1]: An introduction to the basics of Random Matrix Theory (RMT) has been
added to the introduction, to improve the basic accessibility of RMT to readers
\end{description}

Reviewer comments 2, 4, 5, 6, 10, 11, 13, 15, 16 noted easily-correctable
issues relating to phrasing and clarity, and we have made changes in line with
those comments. Reviewer comments 3 and 7 noted some overly-broad or
insufficiently specific claims on our part. We agree these need qualification,
and have altered these statements to be avoid implying anything overly-broad.
Reviewer comments 8, 9 noted some issues with vagueness involving descriptions
of the fMRI scans in the text and tables. Both the text and table captions have
been improved to make these details clear. Comment 14 also suggested moving some
material to methods, and this was also done.

Comment 1 suggested expanding the introduction to Random Matrix Theory (RMT)
somewhat to better cover the basic aspects of RMT, and help with orienting
readers unfamiliar with the area. We have done this expansion, and hope that it
covers the basics both in a way that presumes neither too little nor too much
from readers (always a very difficult balance to strike with something like
RMT, unfortunately).

Our reviewer's final comment suggested perhaps strengthening our conclusions
re: the utility of RMT, given sensitivity to preprocessing.

Please address all correspondence concerning this manuscript to me at \href{jlevman@stfx.ca}{jlevman@stfx.ca}.

\bigskip

\noindent
Sincerely, \\

\small
\noindent
\textbf{Corresponding Author}\\
Jacob Levman, PhD (jlevman@stfx.ca) \\
\footnotesize
\hspace*{0.25cm}St. Francis Xavier University, Department of Computer Science  \\
\hspace*{0.25cm}Martinos Center for Biomedical Imaging, Massachusetts General Hospital \\
\hspace*{0.25cm}Harvard Medical School, Department of Radiology (jlevman@mgh.harvard.edu) \\
\hspace*{0.25cm}Nova Scotia Health Authority, Research Affiliate (jacob.levman@nshealth.ca) \\
\normalsize

\small
\noindent
\textbf{Additional Contacts} [should the corresponding author be unavailable]\\
Derek Berger (dberger@stfx.ca)\\
\footnotesize
\hspace*{0.25cm} St. Francis Xavier University, Department of Computer Science \\
\small
Gurpreet S. Matharoo (gmatharo@stfx.ca) \\
\footnotesize
\hspace*{0.25cm} St. Francis Xavier University, Department of Physics \\
\hspace*{0.25cm} St. Francis Xavier University, ACENET
\normalsize

% \affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}

% \affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}

% \affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
% \affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
% \affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}

% \printbibliography % biblatex
% \renewcommand{\bibsection}{}  % https://tex.stackexchange.com/questions/132646/how-to-remove-the-references-title
% \smallskip
% \bibliographystyle{mystyle}
% \bibliography{cover}  % natbib

\footnotesize
\begin{center}
\noindent\rule{6cm}{0.4pt}
\end{center}

\AtNextBibliography{\footnotesize}
\printbibliography[heading=none]

\end{document}

### 2023-04-03 paper/paper/SPIE/article.tex
add xcolor dependency
[WIP] make paper corrections

@@
 \externaldocument[S-]{supplementary}
+\usepackage{xcolor}
 
@@
 simplifying somewhat, a matrix in which the entries are drawn from
-independently and identically distributed (\textit{iid}) distributions. Though
-first developed to describe the fluctuation of nuclei energy levels in quantum
-physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
-RMT has been shown to have extremely broad potential. In small-scale physical
-systems, RMT universalities have been observed in quantum chaotic systems,
-complex nuclei, atoms, molecules and disordered mesoscopic systems
-\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004,
-brodyRandommatrixPhysicsSpectrum1981, beenakkerRandommatrixTheoryQuantum1997,
+\textcolor{orange}{independent} and identically distributed (\textit{iid})
+distributions. Though first developed to describe the fluctuation of nuclei
+energy levels in quantum physics \cite{mehtaRandomMatrices2004,
+guhrRandommatrixTheoriesQuantum1998a}, RMT has been shown to have extremely
+broad potential. In small-scale physical systems, RMT universalities have been
+observed in quantum chaotic systems, complex nuclei, atoms, molecules and
+disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,
+mehtaRandomMatrices2004, brodyRandommatrixPhysicsSpectrum1981,
+beenakkerRandommatrixTheoryQuantum1997,
 bohigasHigherOrderCorrelationsSpectra1985, wintgenLevelStatisticsQuantized1988,
@@
 \cite{wangRandomMatrixTheory2016}, between pain and non-pain states
-\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs.
+\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs. \
 right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
@@
 
-However, from a predictive standpoint, these summary statistics may eliminate
-predictively-useful information. If the basic RMT insight is that the
-eigenvalues alone can provide understanding of a system, then those eigenvalues
-(or other simple transformations of them) ought also to be predictively useful.
+However, from a predictive standpoint, these summary statistics may
+\textcolor{orange}{mask} predictively-useful information. If the basic RMT
+insight is that the eigenvalues alone can provide understanding of a system,
+then those eigenvalues (or other simple transformations of them) ought also to
+be predictively useful.
 We examine a number of such features in this study, and refer to both
@@
 RMT suggests a potentially useful reduction in the form of the eigenvalues of
-the voxelwise functional connectivity. An \(N \times t\) matrix of
-\(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
-correlation (or covariance) matrix with \(t - 1\) non-zero, positive
-real-valued eigenvalues. These eigenvalues can be computed highly
+the voxelwise functional connectivity \textcolor{orange}{matrix}. An \(N \times
+t\) matrix of \(N\) time series of length \(t\), and where \(t \ll N\) will
+have a symmetric correlation (or covariance) matrix with \(t - 1\) non-zero,
+positive real-valued eigenvalues. These eigenvalues can be computed highly
 efficiently via transposition of the voxelwise correlation or covariance
@@
 subjective decisions regarding outliers and the flexibility of the fitting
-function. These decisions are, unfortunately, often poorly or even entirely
-undocumented, but are known to often dramatically impact RMT conclusions
-\cite{abul-magdUnfoldingSpectrumChaotic2014,
+function. These decisions are, unfortunately, often poorly
+\textcolor{orange}{documented} or even entirely \textcolor{orange}{missing from
+method descriptions, despite being} known to often dramatically impact RMT
+conclusions \cite{abul-magdUnfoldingSpectrumChaotic2014,
 abueleninSpectralUnfoldingChaotic2018, fossionRandommatrixSpectraTime2013,
@@
 memory of certain aspects of the learned images. Thus, the classification task
-might also be considered learning vs. rest.
+might also be considered \textcolor{orange}{\texttt{learning v rest}}.
 
@@
 by BET. BET results were visually inspected for each image in each dataset, and
-BET fractional anisotropy values were modified to ensure brain extraction was
-acceptable. Motion-correction was performed with FSL's MCFLIRT
+BET fractional \textcolor{orange}{intensity threshold (\texttt{-f}
+argument)} values were modified to ensure brain extraction was acceptable.
+Motion-correction was performed with FSL's MCFLIRT
 \cite{jenkinsonImprovedOptimizationRobust2002}, and slice time correction
@@
 
-The time and space complexities of computing the full \(N \times N\)
-\(\mathbf{M}\) matrix of Pearson correlation coefficients (and subsequent
-eigenvalues) are too large to be tractable. However, since transposition does
-not change the eigenvalues, we can use the transpose to efficiently compute
-these eigenvalues (see Appendix \ref{sec:transpose}). These raw eigenvalues
-(feature \code{eigs} in Table \ref{tab:features}), most directly test the idea
-that eigenvalues alone are useful functional connectivity features.
+The \textcolor{orange}{computing} time and \textcolor{orange}{memory
+requirements} for calculating the full \(N \times N\) \(\mathbf{M}\) matrix of
+Pearson correlation coefficients (and subsequent eigenvalues) are too large to
+be tractable. However, since transposition does not change the eigenvalues, we
+can use the transpose to efficiently compute these eigenvalues (see Appendix
+\ref{sec:transpose}). These raw eigenvalues (feature \code{eigs} in Table
+\ref{tab:features}), most directly test the idea that eigenvalues alone are
+useful functional connectivity features.
 
@@
 The largest, middle, or smallest eigenvalues could be most useful in
-characterizing any given system. E.g. if the smallest eigenvalues correspond to
-random / noise aspects of a system, but differences in the nature of the system
-noise most differentiate between systems, then the smallest eigenvalues may
-have the most predictive utility.  Likewise, the \(L\) value in each RMT
-eigenfeature defines the degree of locality at which we summarize the spectrum:
-with small \(L\)-values, the spectral rigidity summarizes the non-linearity of
-eigenvalues that are relatively close to each other in magnitude. At large
-values of \(L\), the rigidity summarizes the long-range non-linearity of the
-spectrum.
+characterizing any given system. \textcolor{orange}{For example,} if the
+smallest eigenvalues correspond to random / noise aspects of a system, but
+differences in the nature of the system noise most differentiate between
+systems, then the smallest eigenvalues may have the most predictive utility.
+Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
+locality at which we summarize the spectrum: with small \(L\)-values, the
+spectral rigidity summarizes the non-linearity of eigenvalues that are
+relatively close to each other in magnitude. At large values of \(L\), the
+rigidity summarizes the long-range non-linearity of the spectrum.
 
@@
 features in Table \ref{tab:features}) to help assess the relative value of RMT
-features.
+features.  \textcolor{orange}{That is, each statistic reduces the fMRI data along
+the voxel dimension, yielding a feature of \(t\) dimensions (e.g. the ``mean''
+feature, \code{T-mean}, is the usual global mean signal).}
 
@@
 difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
-percentile—\code{T-rrng} in Table \ref{tab:features}). Each statistic reduces
-the fMRI data along the voxel dimension, yielding a feature of \(t\) dimensions
-(e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
+percentile—\code{T-rrng} in Table \ref{tab:features}).
+
 % Exact details can be found in the
@@
 restricting the analytic factor to a particular instance, there is no evidence
-of solvability. E.g. if a particular trimming procedure were to render all
-classification tasks unsolvable, it would be better to note this and exclude
-the associated mAUROCs from further visualizations, rather than to have the
-mAUROC distributions diluted by the bad procedure.
+of solvability. \textcolor{orange}{For example,} if a particular trimming
+procedure were to render all classification tasks unsolvable, it would be
+better to note this and exclude the associated mAUROCs from further
+visualizations, rather than to have the mAUROC distributions diluted by the bad
+procedure.
 

### 2023-04-07 paper/paper/SPIE/response_letter.tex
[WIP] make paper corrections
wip

@@
 
-% 2. Line 44: "independently" should be "independent"
+%%%%%% 2. Line 44: "independently" should be "independent"
 
-% 3. Line 56: This is a pretty broad statement, because it would seem that it's
-% an open question whether all complex systems would lend themselves to RMT-based
-% analysis.
+%%%%%%% 3. Line 56: This is a pretty broad statement, because it would seem that it's
+%%%%%%% an open question whether all complex systems would lend themselves to RMT-based
+%%%%%%% analysis.
 
-% 4. Line 100: Do the authors really mean "eliminate" here? Do they instead mean
-% "mask", given that the information presumably doesn't disappear.
+%%%%%% 4. Line 100: Do the authors really mean "eliminate" here? Do they instead mean
+%%%%%% "mask", given that the information presumably doesn't disappear.
 
-% 5. Line 121: "functional connectivity". Do the authors mean "functional
-% connectivity matrix"?
+%%%%%% 5. Line 121: "functional connectivity". Do the authors mean "functional
+%%%%%% connectivity matrix"?
 
-% 6. Line 138: This sentence is a bit garbled. Do the authors mean "often poorly
-% documented or even missing..."?
+%%%%%% 6. Line 138: This sentence is a bit garbled. Do the authors mean "often poorly
+%%%%%% documented or even missing..."?
 
-% 7. Lines 143-145: I don't think these absences raise doubts about these two
-% things in general, but rather that one has to carefully document the workflows
-% and make the data and code available to the public.
+%%%%%% 7. Lines 143-145: I don't think these absences raise doubts about these two
+%%%%%% things in general, but rather that one has to carefully document the workflows
+%%%%%% and make the data and code available to the public.
 
-% 8. Line 156: Do the authors mean "allow a reasonable classification of each
-% scan without resorting to other information". Unclear what they mean by
-% "entirety".
+%%%%%% 8. Line 156: Do the authors mean "allow a reasonable classification of each
+%%%%%% scan without resorting to other information". Unclear what they mean by
+%%%%%% "entirety".
 
-% 9. Table 1: Volumes and n_scans are not defined sufficiently. Are "Volumes" the
-% number of volumes per run? Is n_scans to total number of runs? Is there one run
-% per participant? The participant info is in Table 2, but Table 1 should be
-% independently clear.
+%%%%%% 9. Table 1: Volumes and n_scans are not defined sufficiently. Are "Volumes" the
+%%%%%% number of volumes per run? Is n_scans to total number of runs? Is there one run
+%%%%%% per participant? The participant info is in Table 2, but Table 1 should be
+%%%%%% independently clear.
 
-% 10. Line 180: "learning vs. rest" should be in a different font to be
-% consistent with the other dataset descriptions.
+%%%%%% 10. Line 180: "learning vs. rest" should be in a different font to be
+%%%%%% consistent with the other dataset descriptions.
 
-% 11. Line 208: Are the "fractional anisotropy" values mentioned here the input
-% variables for BET which govern how aggressive BET is and the gradient
-% parameter?
+%%%%%% 11. Line 208: Are the "fractional anisotropy" values mentioned here the input
+%%%%%% variables for BET which govern how aggressive BET is and the gradient
+%%%%%% parameter?
 
-12. Section 3.2: While I understand that the authors are trying to make a
-general statement that will apply to all features, this section is unclear
-because the reader has no idea at this point what features are. Perhaps this
-could be expanded.
+%%%%%% 12. Section 3.2: While I understand that the authors are trying to make a
+%%%%%% general statement that will apply to all features, this section is unclear
+%%%%%% because the reader has no idea at this point what features are. Perhaps this
+%%%%%% could be expanded.
 
-% 13. Line 223: How about "computing time" and "storage space", assuming that is
-% what is meant.
+%%%%%% 13. Line 223: How about "computing time" and "storage space", assuming that is
+%%%%%% what is meant.
 
-% 14. Line 267: It's not clear that this section should be here, maybe at the end
-% of the methods section?
+%%%%%% 14. Line 267: It's not clear that this section should be here, maybe at the end
+%%%%%% of the methods section?
 
-% 15. Line 292: This is a minor point, but sentences shouldn't start with an
-% abbreviation (E.g.) better "For example,". This also occurs in line 387.
+%%%%%% 15. Line 292: This is a minor point, but sentences shouldn't start with an
+%%%%%% abbreviation (E.g.) better "For example,". This also occurs in line 387.
 
-% 16. Line 311: I see what they mean by "voxel time series" because of lines
-% 317-318, so it might be clearer to move those lines to directly after line 311,
-% so it is clear without having to read further.
+%%%%%% 16. Line 311: I see what they mean by "voxel time series" because of lines
+%%%%%% 317-318, so it might be clearer to move those lines to directly after line 311,
+%%%%%% so it is clear without having to read further.
 
@@
 
-Reviewer comments 2, 4, 5, 6, 10, 11, 13, 15, 16 noted easily-correctable
-issues relating to phrasing and clarity, and we have made changes in line with
-those comments. Reviewer comments 3 and 7 noted some overly-broad or
-insufficiently specific claims on our part. We agree these need qualification,
-and have altered these statements to be avoid implying anything overly-broad.
+Reviewer comments 2, 4, 5, 6, 10, 11, 13, 14, 15, 16 noted easily-correctable
+and uncontroversial issues relating to phrasing, clarity or ordering, and we
+have made changes in line with those comments.
+
+Comment 3 interprets our statement as making a broad claim, but the statement was
+in fact intended to qualify or limit the scope of applicability of RMT. This has been
+made clearer. Likewise, we don't think there is any disagreement between our
+intended meaning and the sentiment in comment 7, but have made it clearer that
+our doubts are confined to \textit{past} studies (e.g. with poorly documentation,
+as mentioned in that paragraph). Of course we don't think poor past studies mean no
+future for RMT in general.
+
 Reviewer comments 8, 9 noted some issues with vagueness involving descriptions
-of the fMRI scans in the text and tables. Both the text and table captions have
-been improved to make these details clear. Comment 14 also suggested moving some
-material to methods, and this was also done.
+of the fMRI scans and runs in the text and tables. We hope these descriptions are now
+unambiguous.
+
+In response to comment 12, while we did, and still do take as given that our
+reviewer and a reader of SPIE:JMI knows what features and feature extraction
+are, we can see how at this point in the paper, it may not be clear
+\textit{why} we are extracting features. We have added a short paragraph here
+that we hope resolves this, and makes it clear we are comparing the predictive
+utility of different features / feature extraction methods (and that "feature"
+has the usual meaning in this context).
 

### 2023-04-07 paper/paper/SPIE/article.tex
wip

@@
 
-At its most rudimentary, RMT describes the expected behavior of the
-eigenvalues—also often called levels—of a number of classes of random
-matrices\cite{andersonIntroductionRandomMatrices2010,
-akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
-simplifying somewhat, a matrix in which the entries are drawn from
-\textcolor{orange}{independent} and identically distributed (\textit{iid})
-distributions. Though first developed to describe the fluctuation of nuclei
+\color{cyan}
+
+At its most rudimentary, the basic insight of Random Matrix Theory (RMT) is
+that matrices can be described by their eigenvalues, and that
+large matrices (or large collections of matrices) can be
+well-described by the statistical behaviour of those eigenvalues. Such a
+description is obtained by assuming that the large matrix (or collection
+of matrices) is sampled from a \textit{class} or \textit{ensemble} of random
+matrices.
+
+Simplifying somewhat, a random matrix is a matrix that has entries that can be
+well-described as being drawn from a particular independent and identically
+distributed (\textit{iid}) distribution, or, equivalently, is a matrix where
+each entry is itself a random variable. A \textit{class} or \textit{ensemble}
+of random matrices is generally defined by the shape of the \textit{iid}
+distribution involved. For example, the Gaussian Orthogonal Ensemble (GOE) is a
+\textit{class} or \textit{ensemble} of random matrices that comprises
+orthogonal matrices with each entry being sampled from a standard Gaussian, and
+the Marchenko-Pastur distribution describes the limiting behaviour of the
+singular values of \textit{iid} rectangular random matrices where the
+\textit{iid} distribution is arbitrary\cite{mehtaRandomMatrices2004}. Of
+course, given a single particular matrix, one can only estimate a
+\textiit{probability} that it is a member of a certain class, since certain RMT
+classes can yield arbitrary matrices (albeit with extremely low probability).
+
+and eigenvalue distributions of GOE random matrices have a
+number of well-understood properties\cite{mehtaRandomMatrices2004,
+guhrRandommatrixTheoriesQuantum1998a}. Even more generally,
+
+Specifically, RMT finds that often, in the infinite limit, that the expected
+distribution of the eigenvalues of a number of classes of random matrices can
+be described quite precisely.\cite{andersonIntroductionRandomMatrices2010,
+akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. Although these
+assumptions are quite specific and restrictive, nevertheless, RMT-predicted
+eigenvalue distributions have appeared in diverse natural phenomena.
+
+\color{black}
+
+\textcolor{cyan}{Thus, while} first developed to describe the fluctuation of nuclei
 energy levels in quantum physics \cite{mehtaRandomMatrices2004,
@@
 
-RMT likely has the most real-world explanatory potential when dealing with a
-complex system of many (hundreds or more\cite{mehtaRandomMatrices2004})
-interacting components. If such a system has a matrix representation, and the
-eigenvalues of this representation have a distribution similar to one predicted
-by RMT, it suggests the system is either highly random or chaotic. By contrast,
-if the observed spectra deviate significantly from RMT-predicted spectra, this
+\textcolor{orange}{When or if RMT has} real-world explanatory potential,
+\textcolor{orange}{this will most likely be} when dealing with a complex system
+of many (hundreds or more\cite{mehtaRandomMatrices2004}) interacting
+components, \textcolor{orange}{rather than a system that is too small for
+statistical regularities ro reliably surface}. If such a system has a matrix
+representation, and the eigenvalues of this representation have a distribution
+similar to one predicted by RMT, it suggests the system is either highly random
+or chaotic. By contrast, if the observed spectra deviate significantly from
+RMT-predicted spectra, this
 suggests otherwise. A number of studies have used RMT to make such
@@
 with the absence of reproducible code and data availability, and predominance
-of descriptive and explanatory approaches, raise doubts about the basic
-robustness and practical utility of RMT-based functional connectivity analyses.
+of descriptive and explanatory approaches, \textcolor{orange}{may} raise doubts about the basic
+robustness and practical utility of \textcolor{orange}{past} RMT-based functional connectivity analyses.
 
@@
 somewhat subjective, but we required each dataset to (1) comprise 10 or more
-human subjects, (2) have all fMRI scans have the same spatial and temporal
-resolutions, (3) have fMRI images that can be split into various classes for a
-number of binary classification tasks, and (4), allow reasonably classifying
-each scan in its entirety (i.e. neither task nor event timings are needed to
-determine class membership, nor do any scan volumes need to be removed to make
-the scan classifiable). This yielded seven datasets total.
+human subjects, (2) have all fMRI \textcolor{orange}{runs} have the same
+spatial and temporal resolutions, (3) have fMRI images that can be split into
+various classes for a number of binary classification tasks, and (4), allow
+reasonably classifying each run \textcolor{orange}{using all of the run 4D
+voxel data only.} \textcolor{cyan}{That is, for this last point, classification
+of a run should be reasonable in the absence of run-specific task or event
+timings or details, and should not require using only some subset of the total
+set of 3D volumes (e.g. those associated only with some task or event
+timings)}. This yielded seven datasets total.
 
@@
     TR = Time of Repetition (seconds).
-    n\_scans = Total number of images in dataset.
+    \textcolor{orange}{Volumes = Number of 3D volumes per run.}
+    n\_scans = Total number of \textcolor{orange}{4D images in dataset (number of subjects times number of runs per subject)}.
 }
@@
 
-For each preprocessing degree and fMRI image, features were extracted from all
-\(N\) non-constant brain voxels, where \(N\) may differ for each image. After
-this process, each scan yields an \(N \times t\) matrix \(\mathbf{M}\), where
-\(t\) is the number of volumes acquired. All features extracted summarize
-this matrix \(\mathbf{M}\).
+\textcolor{cyan}{When dealing with large inputs, any predictive analysis must
+reduce that input into predictively-useful features. Any method that reduces a
+larger set of data is potentially valid as a feature extraction method, and RMT,
+which summarizes matrix data by summarizing the eigenvalue distributions, may
+may be useful for feature extraction. However, different feature
+extraction methods may have different advantages and disadvantages in terms of
+the computational complexity, amount of summary or selection involved, and in
+interpretability. RMT is both mathematically and computationally complex, and
+thus RMT-based features should be compared to alternate features that may be
+easier to understand or compute.}
+
+Thus, for each preprocessing degree and fMRI image, features were extracted
+from all \(N\) non-constant brain voxels, where \(N\) may differ for each
+image. After this process, each scan yields an \(N \times t\) matrix
+\(\mathbf{M}\), where \(t\) is the number of volumes acquired. All features
+extracted summarize this matrix \(\mathbf{M}\).
 
@@
 
-\textit{Software Release}. Because the computation of these statistics is
-non-trivial, and to make the unfolding procedure more transparent and
-reproducible, we developed and release a separate, open-source Python library
-\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
-allows for efficient, parallel computation of the spectral rigidity and level
-variance via Monte Carlo methods, and automatically ensures convergence of the
-statistics to user-specified tolerances. The library also makes available
-various other functions useful for empirical RMT analyses, such as unfolding
-and trimming functions, and plotting facilities for classic RMT ensembles.
-
 \subsubsection{Smoothed Eigenfeatures}
@@
 unfolding/smoothing degree, and slicing—to consider.
+
+\subsection{Software Release}
+
+\textcolor{cyan}{Because the computation of these statistics is
+non-trivial, and to make the unfolding procedure more transparent and
+reproducible, we developed and release a separate, open-source Python library
+\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
+allows for efficient, parallel computation of the spectral rigidity and level
+variance via Monte Carlo methods, and automatically ensures convergence of the
+statistics to user-specified tolerances. The library also makes available
+various other functions useful for empirical RMT analyses, such as unfolding
+and trimming functions, and plotting facilities for classic RMT ensembles.}
 

### 2023-04-07 paper/paper/SPIE/article.tex
WIP

@@
 course, given a single particular matrix, one can only estimate a
-\textiit{probability} that it is a member of a certain class, since certain RMT
+\textit{probability} that it is a member of a certain class, since certain RMT
 classes can yield arbitrary matrices (albeit with extremely low probability).
@@
 were modest, and highly dependent on preprocessing steps and other fitting and
-feature selection procedures. Future studies employing RMT should carefully
+feature selection procedures.
+
+\color{cyan}
+Previous studies have not explored the sensitivity of RMT-based analyses to
+analytic decisions. It is far too early to say whether RMT may prove fruitful
+for understanding fMRI, however, we hope our analyses show that, \textit{if}
+RMT has such potential, that potential must be contextualized by the analytic
+decisions involved. Absent this context, we advise caution with regard to
+claims that RMT offers insights into fMRI analyses.
+
+\color{black}
+
+Future studies employing RMT should carefully
 investigate the sensitivity of any findings to such analytic decisions before

### 2023-04-07 paper/paper/SPIE/response_letter.tex
WIP

@@
 
-Thank you for your careful consideration and comments on our manuscript
+Thank you for your careful consideration of our manuscript
 \emph{Random Matrix Theory Tools for the Predictive Analysis of Functional
 Magnetic Resonance Imaging Examinations} submitted for publication in SPIE
-Journal of Medical Imaging.
+Journal of Medical Imaging. We would especially like to thank the reviewer,
+who and offered especially constructive and specific comments that we
 
@@
 RMT, unfortunately).
+
+Finally, comment 17 suggested (by our reading) to opt for less caution and make
+a stronger recommendation in our conclusions.
 

### 2023-04-07 paper/paper/SPIE/article.tex
WIP

@@
 analytic decisions. It is far too early to say whether RMT may prove fruitful
-for understanding fMRI, however, we hope our analyses show that, \textit{if}
-RMT has such potential, that potential must be contextualized by the analytic
-decisions involved. Absent this context, we advise caution with regard to
-claims that RMT offers insights into fMRI analyses.
+for understanding fMRI, however, we hope our analyses show that, when RMT has
+such potential, that potential must be contextualized by the analytic decisions
+involved. Absent this analytical context, we advise considerable caution with
+regard to claims that RMT offers insights into fMRI analyses.
 

### 2023-04-15 paper/paper/SPIE/response_letter.tex
mostly finalize review response

--- 2023-04-07 paper/paper/SPIE/response_letter.tex
@@
 
-Thank you for your careful consideration and comments on our manuscript
+Thank you for your careful consideration of our manuscript
 \emph{Random Matrix Theory Tools for the Predictive Analysis of Functional
 Magnetic Resonance Imaging Examinations} submitted for publication in SPIE
-Journal of Medical Imaging.
+Journal of Medical Imaging. We especially thank our anonymous reviewer,
+who and offered especially constructive and specific comments on this manuscript.
 
-In response to reviewer comments, we have made a number of changes to the
-manuscript. Minor changes relating to wording or for which an edit is
-straightforward (changing or adding a word or two) are identified in
-\textcolor{orange}{orange} in the updated \.pdf manuscript (or with
-\verb|\textcolor{orange}{text}| in the \.tex file). Other more significant
-changes requiring the addition (or movement) of 1-3 sentences are identified
-in \textcolor{cyan}{cyan} (\verb|\textcolor{cyan}{text}|).
+The updated manuscript indicates changes with coloration, as suggested by
+SPIE:JMI. Minor wording or other uncontroversial changes are identified in
+\textcolor{orange}{orange} in the updated .pdf manuscript (or with
+\verb|\textcolor{orange}{text}| in the \texttt{.tex} file). More significant
+changes are identified in \textcolor{cyan}{cyan} (\verb|\textcolor{cyan}{text}|).
 
-\begin{description}
-\item[Comment 1]: An introduction to the basics of Random Matrix Theory (RMT) has been
-added to the introduction, to improve the basic accessibility of RMT to readers
-\end{description}
+\begin{itemize}
+\item Reviewer comments 2, 4, 5, 6, 10, 11, 13, 14, 15, 16 noted easily-correctable
+and uncontroversial issues relating to phrasing, clarity or ordering, and we
+have made changes in line with those comments on lines 55, 115, 136, 153,
+196, 224, 247, 381-387, 413 and 309, and 329-331, respectively.
 
-Reviewer comments 2, 4, 5, 6, 10, 11, 13, 14, 15, 16 noted easily-correctable
-and uncontroversial issues relating to phrasing, clarity or ordering, and we
-have made changes in line with those comments.
+\item Comment 3 interprets our statement as making a broad claim, but the statement
+was in fact intended to qualify or limit the scope of applicability of RMT.
+This has been made clearer (lines 69-71). Likewise, we don't think there is
+any disagreement between our intended meaning and the sentiment in comment 7,
+and have revised line 159 to make it clear that our doubts are confined to
+\textit{past} studies.
 
-Comment 3 interprets our statement as making a broad claim, but the statement was
-in fact intended to qualify or limit the scope of applicability of RMT. This has been
-made clearer. Likewise, we don't think there is any disagreement between our
-intended meaning and the sentiment in comment 7, but have made it clearer that
-our doubts are confined to \textit{past} studies (e.g. with poorly documentation,
-as mentioned in that paragraph). Of course we don't think poor past studies mean no
-future for RMT in general.
+\item Reviewer comments 8 and 9 noted some issues with vagueness involving
+descriptions of the fMRI scans and runs in the text and tables. We hope these
+descriptions are now unambiguous (lines 169-174 and Table 1 caption).
 
-Reviewer comments 8, 9 noted some issues with vagueness involving descriptions
-of the fMRI scans and runs in the text and tables. We hope these descriptions are now
-unambiguous.
-
-In response to comment 12, while we did, and still do take as given that our
+\item In response to comment 12, while we did, and still do take as given that our
 reviewer and a reader of SPIE:JMI knows what features and feature extraction
 are, we can see how at this point in the paper, it may not be clear
-\textit{why} we are extracting features. We have added a short paragraph here
-that we hope resolves this, and makes it clear we are comparing the predictive
-utility of different features / feature extraction methods (and that "feature"
-has the usual meaning in this context).
+\textit{why} we are extracting features. We have added a short paragraph (lines
+234-241) that we hope resolves this, and makes it clear we are comparing the
+predictive utility of different features / feature extraction methods (and that
+``feature'' has the usual meaning in this context).
+\end{itemize}
 
-Comment 1 suggested expanding the introduction to Random Matrix Theory (RMT)
-somewhat to better cover the basic aspects of RMT, and help with orienting
-readers unfamiliar with the area. We have done this expansion, and hope that it
-covers the basics both in a way that presumes neither too little nor too much
-from readers (always a very difficult balance to strike with something like
-RMT, unfortunately).
+In response to our reviewer's first comment, we have expanded the introduction
+to include both the definition of a random matrix, and some concrete examples of
+random matrix ensembles or classes, and also a brief explanation of how RMT
+theory is used in empirical contexts. We hope that this covers the basics both
+in a way that presumes neither too little nor too much from readers (a very
+difficult balance to strike with something like RMT, unfortunately).
 
-Our reviewer's final comment suggested perhaps strengthening our conclusions
-re: the utility of RMT, given sensitivity to preprocessing.
+Our reviewer's final comment suggested (by our reading) to give our conclusion
+a bit more punch. The point is well-taken, and we have reworked our conclusion
+to more strongly imply a need for caution and/or carefulness in the
+presentation and interpretation of RMT-fMRI-related findings.
 
-Please address all correspondence concerning this manuscript to me at \href{jlevman@stfx.ca}{jlevman@stfx.ca}.
+Please address all correspondence concerning this manuscript to \href{jlevman@stfx.ca}{jlevman@stfx.ca}.
 

### 2023-04-15 paper/paper/SPIE/article.tex
mostly finalize review response

--- 2023-04-07 paper/paper/SPIE/article.tex
@@
 
-\color{cyan}
-
-At its most rudimentary, the basic insight of Random Matrix Theory (RMT) is
-that matrices can be described by their eigenvalues, and that
-large matrices (or large collections of matrices) can be
-well-described by the statistical behaviour of those eigenvalues. Such a
-description is obtained by assuming that the large matrix (or collection
-of matrices) is sampled from a \textit{class} or \textit{ensemble} of random
-matrices.
-
-Simplifying somewhat, a random matrix is a matrix that has entries that can be
-well-described as being drawn from a particular independent and identically
-distributed (\textit{iid}) distribution, or, equivalently, is a matrix where
-each entry is itself a random variable. A \textit{class} or \textit{ensemble}
-of random matrices is generally defined by the shape of the \textit{iid}
-distribution involved. For example, the Gaussian Orthogonal Ensemble (GOE) is a
-\textit{class} or \textit{ensemble} of random matrices that comprises
-orthogonal matrices with each entry being sampled from a standard Gaussian, and
-the Marchenko-Pastur distribution describes the limiting behaviour of the
-singular values of \textit{iid} rectangular random matrices where the
-\textit{iid} distribution is arbitrary\cite{mehtaRandomMatrices2004}. Of
-course, given a single particular matrix, one can only estimate a
-\textiit{probability} that it is a member of a certain class, since certain RMT
-classes can yield arbitrary matrices (albeit with extremely low probability).
-
-and eigenvalue distributions of GOE random matrices have a
-number of well-understood properties\cite{mehtaRandomMatrices2004,
-guhrRandommatrixTheoriesQuantum1998a}. Even more generally,
-
-Specifically, RMT finds that often, in the infinite limit, that the expected
-distribution of the eigenvalues of a number of classes of random matrices can
-be described quite precisely.\cite{andersonIntroductionRandomMatrices2010,
-akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. Although these
-assumptions are quite specific and restrictive, nevertheless, RMT-predicted
-eigenvalue distributions have appeared in diverse natural phenomena.
-
-\color{black}
-
-\textcolor{cyan}{Thus, while} first developed to describe the fluctuation of nuclei
-energy levels in quantum physics \cite{mehtaRandomMatrices2004,
-guhrRandommatrixTheoriesQuantum1998a}, RMT has been shown to have extremely
-broad potential. In small-scale physical systems, RMT universalities have been
-observed in quantum chaotic systems, complex nuclei, atoms, molecules and
-disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,
-mehtaRandomMatrices2004, brodyRandommatrixPhysicsSpectrum1981,
-beenakkerRandommatrixTheoryQuantum1997,
+Though first developed to describe the fluctuation of nuclei energy levels in
+quantum physics \cite{mehtaRandomMatrices2004,
+guhrRandommatrixTheoriesQuantum1998a}, Random Matrix Theory (RMT) has been
+shown to have extremely broad potential. In small-scale physical systems, RMT
+universalities have been observed in quantum chaotic systems, complex nuclei,
+atoms, molecules and disordered mesoscopic systems
+\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004,
+brodyRandommatrixPhysicsSpectrum1981, beenakkerRandommatrixTheoryQuantum1997,
 bohigasHigherOrderCorrelationsSpectra1985, wintgenLevelStatisticsQuantized1988,
@@
 
+\color{cyan}
+At its most rudimentary, RMT describes the expected behaviour of the
+eigenvalues—also often called \textcolor{orange}{the \textit{spectra} or}
+\textit{levels}—of a number of classes of random
+matrices\cite{andersonIntroductionRandomMatrices2010,
+akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}.  A random matrix is
+a matrix where each entry is a random variable. Typically, each random variable
+is \textcolor{orange}{independent} and identically distributed (\textit{iid}), however, modern
+extensions allow for some
+dependence\cite{bouchaudFinancialApplicationsRandom2009}. A \textit{class} or
+\textit{ensemble} of random matrices can be defined by the type of
+\textit{iid} distribution involved. For example, the Gaussian Orthogonal
+Ensemble (GOE\cite{mehtaRandomMatrices2004,
+guhrRandommatrixTheoriesQuantum1998a}) is a \textit{class} or \textit{ensemble}
+of random matrices that comprises orthogonal matrices with each entry being
+sampled from a standard Gaussian, and the Marchenko-Pastur distribution
+describes the limiting behaviour of the singular values of \textit{iid}
+rectangular random matrices where the \textit{iid} distribution is
+arbitrary\cite{mehtaRandomMatrices2004}. RMT often finds that in the infinite
+limit, the expected eigenvalue distribution of a number of classes of random
+matrices can be described quite
+precisely\cite{andersonIntroductionRandomMatrices2010,
+akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}.
+
+Of course, given a particular non-random matrix or instantiation of a random
+matrix, one can, in general, only estimate the likelihood that it is a member
+of a certain class or ensemble, since certain RMT ensembles (such as the GOE)
+can yield arbitrary matrices with extremely low probability. Thus, RMT also
+includes statistical measures ("spectral observables") and tools to help
+compare empirical observations to theory\cite{mehtaRandomMatrices2004,
+guhrRandommatrixTheoriesQuantum1998a}.
+
+% At its most rudimentary, the basic insight of Random Matrix Theory (RMT) is
+% that matrices can be described by their eigenvalues, and that
+% large matrices (or large collections of matrices) can be
+% well-described by the statistical behaviour of those eigenvalues.
+
+\color{black}
+
 \textcolor{orange}{When or if RMT has} real-world explanatory potential,
@@
 statistical regularities ro reliably surface}. If such a system has a matrix
-representation, and the eigenvalues of this representation have a distribution
-similar to one predicted by RMT, it suggests the system is either highly random
-or chaotic. By contrast, if the observed spectra deviate significantly from
-RMT-predicted spectra, this
-suggests otherwise. A number of studies have used RMT to make such
-interpretations and comparisons between
+representation, and the eigenvalues \textcolor{orange}{and spectral
+observables} of this representation have distributions similar to
+\textcolor{orange}{those} predicted by RMT, it suggests the system is either
+highly random or chaotic. By contrast, if the observed spectra deviate
+significantly from RMT-predicted spectra, this suggests otherwise. A number of
+studies have used RMT to make such interpretations and comparisons between
 systems\cite{santhanamStatisticsAtmosphericCorrelations2001,
@@
 The \textcolor{orange}{computing} time and \textcolor{orange}{memory
-requirements} for calculating the full \(N \times N\) \(\mathbf{M}\) matrix of
+requirements} for calculating the full \(N \times N\) matrix \(\mathbf{M}\) of
 Pearson correlation coefficients (and subsequent eigenvalues) are too large to
@@
 For combined features, slicing variants are also computed, with slicing
-performed first on each feature to be combined. E.g. if we have features \(f_1,
-\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
-feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
-elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
-\dots;\; \hat{f}_n]\).
+performed first on each feature to be combined. \textcolor{orange}{For example}
+if we have features \(f_1, \dots,  f_n\) with dimensions \(p_1, \dots, p_n\),
+then the sliced max 25\% feature for each component feature, \(\hat{f}_i\) is
+the last \(p_i / 4\) elements of \(f_i\), and the combined sliced feature is
+\([\hat{f}_1;\; \dots;\; \hat{f}_n]\).
 % \footnote{Exact implementation can be found in the
-%\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.
+%\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source
+%code}}.
 
@@
 
-\textcolor{cyan}{Because the computation of these statistics is
+\textcolor{orange}{Because the computation of these statistics is
 non-trivial, and to make the unfolding procedure more transparent and
@@
 were modest, and highly dependent on preprocessing steps and other fitting and
-feature selection procedures. Future studies employing RMT should carefully
-investigate the sensitivity of any findings to such analytic decisions before
-drawing any conclusions. However, this study demonstrates that the inclusion of
-RMT statistics and other eigenfeatures extracted from fMRI images could aid in
-the predictive modeling of a wide variety of phenomena.
+feature selection procedures.
+
+\textcolor{cyan}{Given the sensitivity to these decisions observed in this
+study, and considering the lack of consensus regarding the preprocessing of
+fMRI or other complex biological signal data, RMT should probably not be
+currently considered a tool with "out-of-the-box" utility in these domains.
+While RMT likely still has potential to yield insights in a variety of
+contexts, our results suggest that, absent strong evidence otherwise, it is not
+safe to assume that these insights will generalize broadly beyond the specific
+preprocessing and analytic pipelines involved.}
+
+\color{black}
+
+\textcolor{cyan}{Further research might establish specific sets of analytic
+choices that allow RMT to consistently extract useful information in a wide
+variety of contexts. However, it is also possible that each unique context
+might require a specific combination of choices. In this latter case, there
+should be \textit{strong} theoretical justification for the use of
+RMT, and for each of the various analytic choices involved in its use: the
+variability of results seen in this study suggests that there may often be a
+set of analytic choices that provide favourable results, and which can be
+weakly justified by plausible, but ultimately \textit{post hoc} justifications.}
+\textcolor{cyan}{Provided that these precautions are followed, and that} future
+studies employing RMT \textcolor{orange}{also} carefully investigate the
+sensitivity of any findings to such analytic decisions, \textcolor{cyan}{then
+there is likely considerable untapped potential for RMT in the analysis of fMRI.}
 

### 2023-05-03 paper/paper/SPIE/response_letter.tex
update reponse letter and article to Jacob suggestions

@@
 \usepackage{csquotes}
+\MakeOuterQuote{"}
 \usepackage[british]{babel}
@@
 Magnetic Resonance Imaging Examinations} submitted for publication in SPIE
-Journal of Medical Imaging. We especially thank our anonymous reviewer,
-who and offered especially constructive and specific comments on this manuscript.
+Journal of Medical Imaging. We also thank our anonymous reviewer,
+who offered highly constructive and helpful comments on the manuscript.
 
@@
 \begin{itemize}
-\item Reviewer comments 2, 4, 5, 6, 10, 11, 13, 14, 15, 16 noted easily-correctable
+\item Reviewer comments 2, 4, 5, 6, 10, 11, 13, 14, 15, 16 (see \hyperref[comments]{below}) noted easily-correctable
 and uncontroversial issues relating to phrasing, clarity or ordering, and we
@@
 
-Please address all correspondence concerning this manuscript to \href{jlevman@stfx.ca}{jlevman@stfx.ca}.
-
 \bigskip
@@
 \hspace*{0.25cm}St. Francis Xavier University, Department of Computer Science  \\
-\hspace*{0.25cm}Martinos Center for Biomedical Imaging, Massachusetts General Hospital \\
-\hspace*{0.25cm}Harvard Medical School, Department of Radiology (jlevman@mgh.harvard.edu) \\
+\hspace*{0.25cm}Martinos Center for Biomedical Imaging, Massachusetts General Hospital, Harvard Medical School (jlevman@mgh.harvard.edu) \\
 \hspace*{0.25cm}Nova Scotia Health Authority, Research Affiliate (jacob.levman@nshealth.ca) \\
@@
 
+\section*{Reviewer Comments} \label{comments}
+\begin{enumerate}
+\item Lines 42-45: I think it's unlikely that most readers of SPIE will be
+familiar with RMT and so it would be a good idea to have a bit more of an
+introduction - what's a random matrix?, what does RMT try to do (e.g.,
+characterize systems that can be described using matrices, statistically,
+rather than deterministically.) Later they talk about what RMT has been applied
+to, but it would be good to orient the reader to the general idea.
+
+\item Line 44: "independently" should be "independent"
+
+\item Line 56: This is a pretty broad statement, because it would seem that it's
+an open question whether all complex systems would lend themselves to RMT-based
+analysis.
+
+\item Line 100: Do the authors really mean "eliminate" here? Do they instead mean
+"mask", given that the information presumably doesn't disappear.
+
+\item Line 121: "functional connectivity". Do the authors mean "functional
+connectivity matrix"?
+
+\item Line 138: This sentence is a bit garbled. Do the authors mean "often poorly
+documented or even missing..."?
+
+\item Lines 143-145: I don't think these absences raise doubts about these two
+things in general, but rather that one has to carefully document the workflows
+and make the data and code available to the public.
+
+\item Line 156: Do the authors mean "allow a reasonable classification of each
+scan without resorting to other information". Unclear what they mean by
+"entirety".
+
+\item Table 1: Volumes and n\_scans are not defined sufficiently. Are "Volumes"
+the number of volumes per run? Is n\_scans to total number of runs? Is there
+one run per participant? The participant info is in Table 2, but Table 1 should
+be independently clear.
+
+\item Line 180: "learning vs.\ rest" should be in a different font to be
+consistent with the other dataset descriptions.
+
+\item Line 208: Are the "fractional anisotropy" values mentioned here the input
+variables for BET which govern how aggressive BET is and the gradient
+parameter?
+
+\item Section 3.2: While I understand that the authors are trying to make a
+general statement that will apply to all features, this section is unclear
+because the reader has no idea at this point what features are. Perhaps this
+could be expanded.
+
+\item Line 223: How about "computing time" and "storage space", assuming that is
+what is meant.
+
+\item Line 267: It's not clear that this section should be here, maybe at the end
+of the methods section?
+
+\item Line 292: This is a minor point, but sentences shouldn't start with an
+abbreviation (E.g.) better "For example,". This also occurs in line 387.
+
+\item Line 311: I see what they mean by "voxel time series" because of lines
+317-318, so it might be clearer to move those lines to directly after line 311,
+so it is clear without having to read further.
+
+\item The authors do a nice job of presenting the relevant results, but the
+conclusions were rather general, being along the lines of "investigate the
+sensitivity of choices made in the processing and analysis". That's an
+important result, given previous work and shows that RMT might not "work" with
+different data sets and processing workflows. The question is: what's the
+takeaway: that the method is not one in which you can have a default set of
+parameters that will generally give good results (like say, the recon-all
+pipeline for FreeSurfer)? If RMT so sensitive to preprocessing and choice of
+parameters, is this a useful method to be applied widely on fMRI data? It would
+be good for this to be discussed and perhaps some recommendations given.
+\end{enumerate}
+
 \end{document}

### 2023-05-03 paper/paper/SPIE/article.tex
update reponse letter and article to Jacob suggestions

@@
 components, \textcolor{orange}{rather than a system that is too small for
-statistical regularities ro reliably surface}. If such a system has a matrix
+statistical regularities to reliably surface}. If such a system has a matrix
 representation, and the eigenvalues \textcolor{orange}{and spectral
@@
 \cite{wangRandomMatrixTheory2016}, between pain and non-pain states
-\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs. \
-right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
+\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided
+vs.\ right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
 were found in the latter study, across the first three studies, the spectra of
@@
 which summarizes matrix data by summarizing the eigenvalue distributions, may
-may be useful for feature extraction. However, different feature
+be useful for feature extraction. However, different feature
 extraction methods may have different advantages and disadvantages in terms of
@@
 contexts, our results suggest that, absent strong evidence otherwise, it is not
-safe to assume that these insights will generalize broadly beyond the specific
-preprocessing and analytic pipelines involved.}
+necessarily safe to assume that these insights will generalize broadly beyond
+the specific preprocessing and analytic pipelines involved.}
 
